---
title: "04 — Item Specificity: The Coefficient Map"
subtitle: "How the econ → DV relationship varies by item type and country"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

**Purpose**: The paper's central diagnostic — compare the economic evaluation
coefficient across *all* democratic attitude items and both countries. The
coefficient map (Figure 4) visualises how econ → DV relationships differ
by item type (satisfaction, abstract normative, authoritarian rejection, evaluative)
and by country. The 4.5× ratio calculation quantifies the key cross-item contrast.

- **Input**: `analysis/results/analysis_data.rds`
- **Output**: `analysis/results/item_specificity_results.rds`,
  figures → `analysis/figures/fig04_coefficient_map.{pdf,png}`

```{r setup}
library(tidyverse)
library(broom)
library(sandwich)
library(lmtest)
library(knitr)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/03b_sk_satisfaction_paradox")
analysis_dir <- file.path(paper_dir, "analysis")
results_dir  <- file.path(analysis_dir, "results")
fig_dir      <- file.path(analysis_dir, "figures")

dir.create(results_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(fig_dir,     showWarnings = FALSE, recursive = TRUE)

source(file.path(paper_dir, "R", "helpers.R"))

dat <- readRDS(file.path(results_dir, "analysis_data.rds"))

# Standard controls string
controls_str <- "age_n + gender + edu_n + urban_rural + polint_n"

# is_korea indicator for cross-country interaction models
dat <- dat |>
  mutate(is_korea = as.integer(country_label == "Korea"))

cat("Loaded analysis_data.rds:", nrow(dat), "obs\n")
cat("Countries:", paste(levels(dat$country_label), collapse = ", "), "\n")
cat("Waves:", paste(sort(unique(dat$wave)), collapse = ", "), "\n")
```

## The big regression table — all DVs, both countries {#sec-big-table}

Every democratic attitude DV in the dataset is regressed on `econ_index` with
wave fixed effects and the standard controls. This produces the full matrix of
econ → DV coefficients that structures the paper's item-specificity argument.

```{r dv-list}
# DV list: name, display label, category
all_dvs <- list(
  c("sat_democracy_n",     "Satisfaction: democracy",     "Satisfaction"),
  c("sat_govt_n",          "Satisfaction: government",    "Satisfaction"),
  c("qual_pref_dem_n",     "Dem always preferable",       "Abstract normative"),
  c("auth_reject_index",   "Auth rejection (index)",      "Auth rejection"),
  c("strongman_reject_n",  "Reject: strongman",           "Auth rejection"),
  c("military_reject_n",   "Reject: military rule",       "Auth rejection"),
  c("expert_reject_n",     "Reject: expert rule",         "Auth rejection"),
  c("singleparty_reject_n","Reject: single-party",        "Auth rejection"),
  c("qual_extent_n",       "Dem extent (current)",        "Evaluative"),
  c("qual_sys_support_n",  "System deserves support",     "Evaluative"),
  c("qual_sys_change_n",   "No major change needed",      "Evaluative")
)

# Preserve display order
dv_order <- map_chr(all_dvs, 1)
dv_labels <- setNames(map_chr(all_dvs, 2), map_chr(all_dvs, 1))
dv_cats   <- setNames(map_chr(all_dvs, 3), map_chr(all_dvs, 1))

countries <- c("Korea", "Taiwan")
```

```{r estimate-all}
item_spec_rows <- map_dfr(countries, function(cntry) {
  df_c <- dat |> filter(country_label == cntry)

  map_dfr(all_dvs, function(dv_entry) {
    dv       <- dv_entry[1]
    dv_label <- dv_entry[2]
    category <- dv_entry[3]

    # Skip if DV is entirely missing for this country
    if (all(is.na(df_c[[dv]]))) {
      return(tibble(
        country   = cntry,
        dv        = dv,
        dv_label  = dv_label,
        category  = category,
        estimate  = NA_real_, std.error = NA_real_,
        p.value   = NA_real_,
        conf.low  = NA_real_, conf.high = NA_real_,
        n         = 0L,
        stars     = ""
      ))
    }

    formula_str <- paste(dv, "~ econ_index + factor(wave) +", controls_str)
    m <- lm(as.formula(formula_str), data = df_c)
    n_obs <- sum(!is.na(df_c[[dv]]) & !is.na(df_c$econ_index))

    extract_coef(m, "econ_index") |>
      mutate(
        country  = cntry,
        dv       = dv,
        dv_label = dv_label,
        category = category,
        n        = n_obs,
        stars    = sig_stars(p.value)
      )
  })
})

item_spec_df <- item_spec_rows |>
  mutate(
    dv       = factor(dv,       levels = dv_order),
    dv_label = factor(dv_label, levels = dv_labels[dv_order]),
    category = factor(category, levels = c("Satisfaction", "Abstract normative",
                                           "Auth rejection", "Evaluative"))
  )
```

```{r big-table-display}
item_spec_df |>
  arrange(category, dv, country) |>
  select(country, category, dv_label, estimate, std.error,
         conf.low, conf.high, p.value, stars, n) |>
  mutate(across(c(estimate, std.error, conf.low, conf.high, p.value),
                \(x) round(x, 4))) |>
  kbl(
    caption = paste(
      "Economic evaluation → democratic attitude DVs, pooled OLS with wave FE.",
      "Both countries, all items. β on econ_index (0–1 scale)."
    ),
    col.names = c("Country", "Category", "DV", "β", "SE",
                  "CI low", "CI high", "p", "Sig", "N")
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  collapse_rows(columns = c(1, 2), valign = "top")
```

## The 4.5× ratio calculation {#sec-ratio}

The core claim: the econ → abstract normative gap between Korea and Taiwan is
approximately 4.5× larger than the mean econ → authoritarian rejection gap.
This ratio is the quantitative anchor for the item-specificity argument.

```{r ratio}
# Abstract item gap (Taiwan β - Korea β on qual_pref_dem_n)
# Expected direction: Taiwan = strong negative (econ good → prefer dem less?),
# Korea ≈ 0 (no relationship). Gap = Taiwan - Korea.
tw_pref <- item_spec_df |>
  filter(country == "Taiwan", dv == "qual_pref_dem_n") |>
  pull(estimate)

kr_pref <- item_spec_df |>
  filter(country == "Korea", dv == "qual_pref_dem_n") |>
  pull(estimate)

pref_gap <- tw_pref - kr_pref

# Auth rejection gaps (Korea - Taiwan for each item)
# Korea is expected to be more positive (econ good → more anti-auth) relative to Taiwan
auth_items <- c("auth_reject_index", "strongman_reject_n", "military_reject_n",
                "expert_reject_n", "singleparty_reject_n")

auth_betas <- item_spec_df |>
  filter(dv %in% auth_items) |>
  select(country, dv, estimate) |>
  pivot_wider(names_from = country, values_from = estimate)

# Gap = Korea - Taiwan for each item
auth_gaps <- auth_betas$Korea - auth_betas$Taiwan
mean_auth_gap <- mean(auth_gaps, na.rm = TRUE)

ratio <- abs(pref_gap) / abs(mean_auth_gap)

cat("── Item-specificity ratio ──────────────────────────────────\n")
cat(sprintf("Abstract item (qual_pref_dem_n):\n"))
cat(sprintf("  Taiwan β: %.4f\n", tw_pref))
cat(sprintf("  Korea  β: %.4f\n", kr_pref))
cat(sprintf("  Gap (Taiwan - Korea): %.4f\n\n", pref_gap))

cat("Auth rejection gaps (Korea - Taiwan) by item:\n")
walk2(auth_betas$dv, auth_gaps, \(dv, gap)
  cat(sprintf("  %-25s: %.4f\n", dv, gap))
)
cat(sprintf("  Mean auth rejection gap: %.4f\n\n", mean_auth_gap))

cat(sprintf("Ratio |pref_gap| / |mean_auth_gap| = %.1fx\n", ratio))
```

```{r ratio-table}
# Formatted table of the ratio inputs
ratio_tbl <- bind_rows(
  tibble(
    item      = "qual_pref_dem_n (abstract normative)",
    korea_b   = kr_pref,
    taiwan_b  = tw_pref,
    gap       = pref_gap,
    note      = "Korea – Taiwan gap used in ratio numerator"
  ),
  auth_betas |>
    mutate(
      gap  = Korea - Taiwan,
      note = "Korea – Taiwan gap, auth rejection items"
    ) |>
    rename(item = dv, korea_b = Korea, taiwan_b = Taiwan)
) |>
  add_row(
    item     = "Mean auth rejection gap",
    korea_b  = NA, taiwan_b = NA,
    gap      = mean_auth_gap,
    note     = sprintf("Ratio = %.1fx", ratio)
  )

ratio_tbl |>
  mutate(across(c(korea_b, taiwan_b, gap), \(x) round(x, 4))) |>
  kbl(
    caption = sprintf(
      "Ratio calculation: |pref_gap| / |mean_auth_gap| = %.1fx", ratio
    ),
    col.names = c("Item", "Korea β", "Taiwan β", "Gap", "Note")
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  row_spec(nrow(ratio_tbl), bold = TRUE, background = "#f0f4ff")
```

## Cross-country interaction by item type {#sec-cross-country}

For each DV, the model `dv ~ econ_index * is_korea + factor(wave) + controls`
is estimated on the pooled both-country dataset. The interaction term
`econ_index:is_korea` captures the *differential* Korea vs. Taiwan effect.
The item-specificity argument predicts a large, positive interaction for
`qual_pref_dem_n` and a small or sign-reversed interaction for auth rejection items.

```{r cross-country-by-item}
xc_item_rows <- map_dfr(all_dvs, function(dv_entry) {
  dv       <- dv_entry[1]
  dv_label <- dv_entry[2]
  category <- dv_entry[3]

  if (all(is.na(dat[[dv]]))) {
    return(tibble(
      dv = dv, dv_label = dv_label, category = category,
      term = "econ_index:is_korea",
      estimate = NA_real_, std.error = NA_real_,
      statistic = NA_real_, p.value = NA_real_,
      conf.low = NA_real_, conf.high = NA_real_,
      stars = ""
    ))
  }

  formula_str <- paste(
    dv, "~ econ_index * is_korea + factor(wave) +", controls_str
  )
  m <- lm(as.formula(formula_str), data = dat)
  tidy_hc2(m) |>
    filter(term == "econ_index:is_korea") |>
    mutate(
      dv       = dv,
      dv_label = dv_label,
      category = category,
      stars    = sig_stars(p.value)
    ) |>
    select(dv, dv_label, category, term,
           estimate, std.error, statistic, p.value, conf.low, conf.high, stars)
})

xc_item_df <- xc_item_rows |>
  mutate(
    dv       = factor(dv,       levels = dv_order),
    dv_label = factor(dv_label, levels = dv_labels[dv_order]),
    category = factor(category, levels = levels(item_spec_df$category))
  )

xc_item_df |>
  arrange(category, dv) |>
  select(category, dv_label, estimate, std.error, conf.low, conf.high, p.value, stars) |>
  mutate(across(c(estimate, std.error, conf.low, conf.high, p.value),
                \(x) round(x, 4))) |>
  kbl(
    caption = paste(
      "Cross-country interaction: econ_index:is_korea coefficient by DV.",
      "Positive = Korea effect stronger than Taiwan."
    ),
    col.names = c("Category", "DV", "β (interaction)", "SE",
                  "CI low", "CI high", "p", "Sig")
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  collapse_rows(columns = 1, valign = "top")
```

## Coefficient map figure (Figure 4) {#sec-figure}

Figure 4 is the paper's visual centrepiece. It shows the `econ_index` pooled-OLS
coefficient (with 95% CI) for every DV, side by side for Korea and Taiwan, grouped
by item category.

```{r coef-map-figure}
# Build display-ready data frame with factor ordering preserved
fig_data <- item_spec_df |>
  filter(!is.na(estimate)) |>
  mutate(
    # Reverse label order so the first item in each category appears at the top
    dv_label = fct_rev(dv_label)
  )

fig_coef_map <- fig_data |>
  ggplot(aes(x = estimate, y = dv_label, colour = country)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_pointrange(
    aes(xmin = conf.low, xmax = conf.high),
    position = position_dodge(width = 0.6),
    size     = 0.4,
    linewidth = 0.7
  ) +
  facet_grid(category ~ ., scales = "free_y", space = "free_y") +
  scale_colour_manual(values = c("Korea" = "#2166AC", "Taiwan" = "#D55E00")) +
  labs(
    title    = "Economic evaluations and democratic attitudes: Korea vs. Taiwan",
    subtitle = "Pooled OLS with wave FE and controls. 95% CIs.",
    x        = "\u03b2 (economic evaluation index \u2192 DV, 0\u20131 scales)",
    y        = NULL,
    colour   = NULL,
    caption  = paste(
      "Controls: age, gender, education, urban/rural, political interest.",
      "\nSource: Asian Barometer Survey, waves 1\u20136."
    )
  ) +
  theme_pub +
  theme(
    strip.text.y  = element_text(angle = 0, hjust = 0),
    panel.spacing = unit(0.8, "lines")
  )

ggsave(file.path(fig_dir, "fig04_coefficient_map.pdf"), fig_coef_map,
       width = 8, height = 7)
ggsave(file.path(fig_dir, "fig04_coefficient_map.png"), fig_coef_map,
       width = 8, height = 7, dpi = 300)

fig_coef_map
```

```{r coef-map-notes}
cat("Figure 4 saved:\n")
cat("  ", file.path(fig_dir, "fig04_coefficient_map.pdf"), "\n")
cat("  ", file.path(fig_dir, "fig04_coefficient_map.png"), "\n")
```

## Coefficient map — interaction version {#sec-figure-interaction}

Supplementary version of Figure 4 showing the `econ_index:is_korea` interaction
coefficient (Korea differential effect) for each DV. Items where Korea differs
most strongly from Taiwan are immediately visible.

```{r coef-map-interaction}
fig_xc <- xc_item_df |>
  filter(!is.na(estimate)) |>
  mutate(dv_label = fct_rev(dv_label)) |>
  ggplot(aes(x = estimate, y = dv_label)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  geom_pointrange(
    aes(xmin = conf.low, xmax = conf.high),
    colour    = "#2166AC",
    size      = 0.4,
    linewidth = 0.7
  ) +
  facet_grid(category ~ ., scales = "free_y", space = "free_y") +
  labs(
    title    = "Korea – Taiwan differential: econ_index:is_korea interaction",
    subtitle = "Positive = Korea effect on DV stronger than Taiwan. Pooled OLS, wave FE.",
    x        = "\u03b2 (econ_index:is_korea interaction coefficient)",
    y        = NULL,
    caption  = paste(
      "Controls: age, gender, education, urban/rural, political interest.",
      "\nSource: Asian Barometer Survey, waves 1\u20136."
    )
  ) +
  theme_pub +
  theme(
    strip.text.y  = element_text(angle = 0, hjust = 0),
    panel.spacing = unit(0.8, "lines")
  )

ggsave(file.path(fig_dir, "fig04b_coef_map_interaction.pdf"), fig_xc,
       width = 7, height = 7)
ggsave(file.path(fig_dir, "fig04b_coef_map_interaction.png"), fig_xc,
       width = 7, height = 7, dpi = 300)

fig_xc
```

## Save {#sec-save}

```{r save}
item_specificity_results <- list(
  all_dvs       = item_spec_df,
  ratio         = ratio,
  pref_gap      = pref_gap,
  mean_auth_gap = mean_auth_gap,
  cross_country = xc_item_df
)

saveRDS(item_specificity_results,
        file.path(results_dir, "item_specificity_results.rds"))

cat("✓ Saved item_specificity_results.rds\n")
cat(sprintf("  all_dvs:       %d rows (%d DVs × %d countries)\n",
            nrow(item_spec_df), length(all_dvs), length(countries)))
cat(sprintf("  cross_country: %d rows\n", nrow(xc_item_df)))
cat(sprintf("  Ratio:         %.1fx\n", ratio))
cat(sprintf("  pref_gap:      %.4f\n",  pref_gap))
cat(sprintf("  mean_auth_gap: %.4f\n",  mean_auth_gap))
```
