---
title: "Appendix I & Appendix A Supplement: Reliability and Polychoric Correlations"
subtitle: "03b_sk_satisfaction_paradox"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
execute:
  echo: false
  warning: false
  message: false
  cache: false
---

```{r setup}
library(tidyverse)
library(psych)
library(kableExtra)

# Paths — hardcoded for standalone rendering outside the Quarto project
project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/03b_sk_satisfaction_paradox")
results_dir  <- file.path(paper_dir, "analysis/results")

source(file.path(paper_dir, "R/helpers.R"))

# Load analysis data
dat <- readRDS(file.path(results_dir, "analysis_data.rds"))
```

```{r item-definitions}
# Item sets
econ_items <- c(
  "econ_national_n",
  "econ_family_n",
  "econ_outlook_n",
  "econ_fam_outlook_n",
  "econ_change_n",
  "econ_fam_change_n"
)

auth_items <- c(
  "strongman_reject_n",
  "military_reject_n",
  "expert_reject_n",
  "singleparty_reject_n"
)

dem_var <- "dem_always_preferable"

# Confirm all variables exist
stopifnot(all(econ_items %in% names(dat)))
stopifnot(all(auth_items %in% names(dat)))
stopifnot(dem_var %in% names(dat))

# Attach survey year
dat <- dat |>
  left_join(survey_years_lookup, by = c("wave", "country_label" = "country"))
```

## 1  Polychoric / Pearson Correlations: Democratic Preference and Auth Rejection Items

The normalized (`_n`) variables are continuous [0, 1], so we use Pearson correlations
throughout. The raw ordinal versions could support polychoric estimation, but the
normalized variables are used in all main analyses.

```{r polychoric-by-country}
# ── Pooled by country ──────────────────────────────────────────────────────────
poly_by_country <- dat |>
  filter(!is.na(country_label)) |>
  group_by(country_label) |>
  group_map(function(df, grp) {
    vars_to_use <- c(dem_var, auth_items)
    mat <- df |> select(all_of(vars_to_use)) |> drop_na()
    r   <- cor(mat, use = "pairwise.complete.obs")
    n   <- nrow(mat)
    list(
      country = grp$country_label,
      n       = n,
      cor_mat = r
    )
  })

names(poly_by_country) <- sapply(poly_by_country, `[[`, "country")
```

```{r polychoric-by-country-wave}
# ── By country × wave ──────────────────────────────────────────────────────────
poly_by_cw <- dat |>
  filter(!is.na(country_label), !is.na(wave)) |>
  group_by(country_label, wave) |>
  group_map(function(df, grp) {
    vars_to_use <- c(dem_var, auth_items)
    mat <- df |> select(all_of(vars_to_use)) |> drop_na()
    n   <- nrow(mat)
    yr  <- unique(df$year)
    yr  <- if (length(yr) == 1) yr else NA_integer_
    # Return NULL-safe placeholder if too few complete cases
    if (n < 3) {
      return(list(
        country = grp$country_label,
        wave    = grp$wave,
        year    = yr,
        n       = n,
        cor_mat = NULL
      ))
    }
    r <- cor(mat, use = "pairwise.complete.obs")
    list(
      country = grp$country_label,
      wave    = grp$wave,
      year    = yr,
      n       = n,
      cor_mat = r
    )
  })

names(poly_by_cw) <- paste0(
  sapply(poly_by_cw, `[[`, "country"), "_W",
  sapply(poly_by_cw, `[[`, "wave")
)
```

```{r display-country-cors}
# Build a tidy summary table: one row per country × auth_item
cor_summary <- map_dfr(poly_by_country, function(lst) {
  tibble(
    country     = lst$country,
    n           = lst$n,
    item        = auth_items,
    r_with_dem  = lst$cor_mat[dem_var, auth_items]
  )
})

cor_summary |>
  mutate(
    item = recode(item,
      strongman_reject_n = "Strongman rejection",
      military_reject_n  = "Military rejection",
      expert_reject_n    = "Expert-rule rejection",
      singleparty_reject_n = "Single-party rejection"
    )
  ) |>
  kable(
    caption = "Pearson correlations: dem_always_preferable × auth rejection items (pooled across waves)",
    col.names = c("Country", "N", "Item", "r"),
    digits = 3
  ) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

```{r display-cw-cors}
# Build tidy summary: one row per country × wave × auth_item
# Skip entries with NULL cor_mat (e.g. wave 2 where auth items were not asked)
cor_cw_summary <- map_dfr(poly_by_cw, function(lst) {
  if (is.null(lst$cor_mat)) {
    return(tibble(
      country    = lst$country,
      wave       = lst$wave,
      year       = lst$year,
      n          = lst$n,
      item       = auth_items,
      r_with_dem = NA_real_
    ))
  }
  tibble(
    country    = lst$country,
    wave       = lst$wave,
    year       = lst$year,
    n          = lst$n,
    item       = auth_items,
    r_with_dem = lst$cor_mat[dem_var, auth_items]
  )
})

cor_cw_summary |>
  mutate(
    item = recode(item,
      strongman_reject_n   = "Strongman",
      military_reject_n    = "Military",
      expert_reject_n      = "Expert",
      singleparty_reject_n = "Single-party"
    )
  ) |>
  pivot_wider(
    id_cols     = c(country, wave, year, n),
    names_from  = item,
    values_from = r_with_dem
  ) |>
  arrange(country, wave) |>
  kable(
    caption = "Pearson correlations: dem_always_preferable × auth rejection items, by country × wave",
    digits = 3
  ) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) |>
  pack_rows("Korea", 1, 6) |>
  pack_rows("Taiwan", 7, 12)
```

## 2  Cronbach's Alpha — Economic Index (R3)

6-item economic evaluation index: `econ_national_n`, `econ_family_n`,
`econ_outlook_n`, `econ_fam_outlook_n`, `econ_change_n`, `econ_fam_change_n`.

```{r alpha-econ}
alpha_econ <- dat |>
  filter(!is.na(country_label), !is.na(wave)) |>
  group_by(country_label, wave) |>
  group_map(function(df, grp) {
    mat <- df |> select(all_of(econ_items)) |> drop_na()
    n   <- nrow(mat)
    yr  <- unique(df$year)
    yr  <- if (length(yr) == 1) yr else NA_integer_
    if (n < 10) {
      return(tibble(
        country = grp$country_label, wave = grp$wave,
        year = yr, n = n, alpha = NA_real_
      ))
    }
    a <- psych::alpha(mat, check.keys = TRUE, warnings = FALSE)
    tibble(
      country = grp$country_label,
      wave    = grp$wave,
      year    = yr,
      n       = n,
      alpha   = a$total$raw_alpha
    )
  }) |>
  bind_rows()

alpha_econ |>
  arrange(country, wave) |>
  kable(
    caption = "Cronbach's alpha: Economic index (6 items), by country × wave",
    col.names = c("Country", "Wave", "Year", "N", "Alpha"),
    digits = 3
  ) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## 3  Cronbach's Alpha — Authoritarian Rejection Index (R3)

4-item auth rejection index: `strongman_reject_n`, `military_reject_n`,
`expert_reject_n`, `singleparty_reject_n`.

```{r alpha-auth}
alpha_auth <- dat |>
  filter(!is.na(country_label), !is.na(wave)) |>
  group_by(country_label, wave) |>
  group_map(function(df, grp) {
    mat <- df |> select(all_of(auth_items)) |> drop_na()
    n   <- nrow(mat)
    yr  <- unique(df$year)
    yr  <- if (length(yr) == 1) yr else NA_integer_
    if (n < 10) {
      return(tibble(
        country = grp$country_label, wave = grp$wave,
        year = yr, n = n, alpha = NA_real_
      ))
    }
    a <- psych::alpha(mat, check.keys = TRUE, warnings = FALSE)
    tibble(
      country = grp$country_label,
      wave    = grp$wave,
      year    = yr,
      n       = n,
      alpha   = a$total$raw_alpha
    )
  }) |>
  bind_rows()

alpha_auth |>
  arrange(country, wave) |>
  kable(
    caption = "Cronbach's alpha: Auth rejection index (4 items), by country × wave",
    col.names = c("Country", "Wave", "Year", "N", "Alpha"),
    digits = 3
  ) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## 4  McDonald's Omega — Econ and Auth Indices

Omega total is a better reliability estimate than alpha when item loadings are
unequal. Computed here via `psych::omega()` (minres factor model, 1 factor).

```{r omega-econ}
omega_econ <- dat |>
  filter(!is.na(country_label), !is.na(wave)) |>
  group_by(country_label, wave) |>
  group_map(function(df, grp) {
    mat <- df |> select(all_of(econ_items)) |> drop_na()
    n   <- nrow(mat)
    yr  <- unique(df$year)
    yr  <- if (length(yr) == 1) yr else NA_integer_
    if (n < 20) {
      return(tibble(
        country = grp$country_label, wave = grp$wave,
        year = yr, n = n, omega_t = NA_real_
      ))
    }
    om <- tryCatch(
      psych::omega(mat, nfactors = 1, plot = FALSE, warnings = FALSE),
      error = function(e) NULL
    )
    tibble(
      country = grp$country_label,
      wave    = grp$wave,
      year    = yr,
      n       = n,
      omega_t = if (!is.null(om)) om$omega.tot else NA_real_
    )
  }) |>
  bind_rows()

omega_econ |>
  arrange(country, wave) |>
  kable(
    caption = "McDonald's omega (total): Economic index (6 items), by country × wave",
    col.names = c("Country", "Wave", "Year", "N", "Omega"),
    digits = 3
  ) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

```{r omega-auth}
omega_auth <- dat |>
  filter(!is.na(country_label), !is.na(wave)) |>
  group_by(country_label, wave) |>
  group_map(function(df, grp) {
    mat <- df |> select(all_of(auth_items)) |> drop_na()
    n   <- nrow(mat)
    yr  <- unique(df$year)
    yr  <- if (length(yr) == 1) yr else NA_integer_
    if (n < 20) {
      return(tibble(
        country = grp$country_label, wave = grp$wave,
        year = yr, n = n, omega_t = NA_real_
      ))
    }
    om <- tryCatch(
      psych::omega(mat, nfactors = 1, plot = FALSE, warnings = FALSE),
      error = function(e) NULL
    )
    tibble(
      country = grp$country_label,
      wave    = grp$wave,
      year    = yr,
      n       = n,
      omega_t = if (!is.null(om)) om$omega.tot else NA_real_
    )
  }) |>
  bind_rows()

omega_auth |>
  arrange(country, wave) |>
  kable(
    caption = "McDonald's omega (total): Auth rejection index (4 items), by country × wave",
    col.names = c("Country", "Wave", "Year", "N", "Omega"),
    digits = 3
  ) |>
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```

## 5  Save Results

```{r save-results}
reliability_results <- list(
  polychoric_by_country      = poly_by_country,
  polychoric_by_country_wave = poly_by_cw,
  alpha_econ                 = alpha_econ,
  omega_econ                 = omega_econ,
  alpha_auth                 = alpha_auth,
  omega_auth                 = omega_auth
)

saveRDS(reliability_results, file.path(results_dir, "reliability_results.rds"))

cat("Saved reliability_results.rds to:", file.path(results_dir, "reliability_results.rds"), "\n")
cat("Components:", paste(names(reliability_results), collapse = ", "), "\n")
```
