---
title: "The Meaning of Losing: How Democratic Erosion Reshapes Citizens' Conceptions of Democracy"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: today
abstract: |
  Does democratic erosion reshape not just citizens' satisfaction with democracy but their understanding of what democracy means? Using multinomial logit models on four waves of Asian Barometer Survey data (2005–2022) across eleven countries, this article shows that electoral losers systematically prioritize liberal-procedural conceptions of democracy—free expression, media freedom, judicial accountability—while winners emphasize substantive outcomes like economic equality and welfare. This divergence responds dynamically to political context. In Thailand, the gap between winners' and losers' conceptions widened from near zero to over twenty percentage points as democratic institutions eroded after the 2014 coup; in South Korea, where a comparable crisis was resolved constitutionally, the gap remained negligible. A fairness interaction test confirms that the effect intensifies among losers perceiving elections as compromised, consistent with a positional updating mechanism rather than stable cultural orientations. The finding carries a troubling implication for democratic resilience: where erosion is most severe, procedural commitments become concentrated among those whom the political system has rendered least capable of defending them—the citizens who most value democratic safeguards are precisely those who have lost the power to demand them.
keywords:
  - democracy
  - electoral losers
  - procedural democracy
  - Asian Barometer Survey
  - democratic consolidation
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: false
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \raggedbottom
          \floatplacement{figure}{H}
          \newcommand{\keywords}[1]{\vspace{0.5em}\begin{singlespace}\noindent\textbf{Keywords:} #1\end{singlespace}}
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
  docx:
    reference-doc: custom-reference.docx
    number-sections: false
csl: chicago-author-date.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.pos = "H", out.width = "100%"
)

here::i_am("papers/meaning-of-democracy-revision/manuscript/manuscript.qmd")

# Load results from multinomial analysis pipeline
results_path <- here("papers", "meaning-of-democracy-revision", "analysis", "revised", "results")

res <- readRDS(file.path(results_path, "mlogit_results.rds"))
cw_gap <- read_csv(file.path(results_path, "country_wave_proc_sub_gap.csv"), show_col_types = FALSE)
cw_items <- read_csv(file.path(results_path, "country_wave_item_ames.csv"), show_col_types = FALSE)

# Helpers
fmt_n <- function(x) format(x, big.mark = ",")
fmt_pct <- function(x) sprintf("%.1f", x * 100)
fmt_pp <- function(x) sprintf("%+.1f", x * 100)

# Key quantities
pooled <- res$all_results
n_countries <- n_distinct(cw_gap$country_name)

# Thailand trajectory
thai <- cw_gap %>% filter(country_name == "Thailand") %>% arrange(wave)
thai_w3_gap <- thai %>% filter(wave == 3) %>% pull(proc_sub_gap)
thai_w4_gap <- thai %>% filter(wave == 4) %>% pull(proc_sub_gap)
thai_w6_gap <- thai %>% filter(wave == 6) %>% pull(proc_sub_gap)
thai_change <- (thai_w6_gap - thai_w3_gap) * 100

# Total N (approximate from item-level data)
total_n <- cw_items %>%
  filter(set == "Set1") %>%
  group_by(country_name, wave_label) %>%
  slice(1) %>%
  ungroup() %>%
  pull(n) %>%
  sum()

# COVID robustness summary
covid_sum <- readRDS(file.path(results_path, "robustness_covid_summary.rds"))
covid_proc_base <- covid_sum %>% filter(item_type == "procedural", controls == "base") %>% pull(mean_ame)
covid_proc_covid <- covid_sum %>% filter(item_type == "procedural", controls == "covid") %>% pull(mean_ame)
covid_sub_base <- covid_sum %>% filter(item_type == "substantive", controls == "base") %>% pull(mean_ame)
covid_sub_covid <- covid_sum %>% filter(item_type == "substantive", controls == "covid") %>% pull(mean_ame)
```

\keywords{democracy, electoral losers, procedural democracy, Asian Barometer Survey, democratic consolidation}


# Introduction

Between 2010 and 2020, Thailand experienced a constitutional court's annulment of an election, a military coup, six years of junta governance, and the return to civilian rule under a military-drafted constitution that ensured the army's continued dominance. During this same decade, the gap between how electoral winners and losers conceptualized democracy widened from near zero to over `r round(thai_w6_gap * 100)` percentage points. Supporters of the excluded Thaksin-aligned parties increasingly defined democracy in terms of free elections, civil liberties, and government accountability; supporters of the military-backed order increasingly emphasized economic welfare and public services. This is not a story about who was more satisfied with democracy or more trusting of government---the attitudes that the winner-loser literature has studied extensively [@Anderson2005-uk; @Singh2012-hj]. It is a story about what citizens came to believe democracy *means*.

South Korea offers the contrast. Over the same period, the country weathered its own crisis: the impeachment and criminal prosecution of President Park Geun-hye, mass candlelight protests, and the election of a progressive successor. Yet the gap between winners' and losers' conceptions of democracy never exceeded five percentage points. The critical difference was institutional. South Korea's crisis was resolved through constitutional procedures---impeachment, judicial review, a free election---that demonstrated the system's capacity for self-correction. Thailand's crisis was resolved by tanks. Where institutions held, winners and losers converged on what democracy meant; where institutions collapsed, democratic meaning itself fractured along the winner-loser divide.

These contrasting trajectories point to a question that two established research literatures have, surprisingly, not addressed together. One body of scholarship documents how citizens conceptualize democracy---whether in procedural terms emphasizing elections and rights, or substantive terms emphasizing welfare and equality---and finds considerable variation across and within countries [@Russell-J-Dalton-Doh-C-Shin-and-Willy-Jou2007-gm; @Chu2008-fk; @Ferrin2016-se]. A separate literature demonstrates that electoral winners and losers differ systematically in their satisfaction with democracy, their institutional trust, and their willingness to accept political outcomes [@Anderson2005-uk; @Anderson2002-cv; @Esaiasson2011-po]. But while the first literature has treated democratic conceptions as relatively stable cultural orientations, and the second has assumed that winners and losers are evaluating the same thing when they express views about "democracy," neither has examined whether electoral status shapes not just how citizens *evaluate* democracy but how they *understand* it.

This article argues that it does. Using multinomial logit models on four waves of Asian Barometer Survey data (2005--2022) encompassing more than `r fmt_n(total_n)` voters across `r n_countries` countries, I show that electoral losers systematically prioritize procedural elements of democracy---competitive elections, civil liberties, government accountability---over substantive outcomes such as economic equality or welfare provision. This "loser effect" is individually modest in pooled cross-national estimates, typically two to four percentage points per item, but it is directionally consistent across twenty items from five separate survey batteries. More importantly, the effect responds dynamically to political context: it tracks democratic erosion in Thailand, remains stable amid institutional resilience in South Korea, and persists at high levels in Cambodia's competitive authoritarian regime where the stakes of losing were always extreme.

The finding carries a troubling implication for the study of democratic resilience. A growing body of work examines whether mass democratic attitudes can serve as a bulwark against backsliding [@Claassen2020-hk; @Svolik2019-zi; @Graham2020-jr]. The present results suggest that where erosion is most severe, procedural commitments become concentrated among those whom the political system has rendered least capable of defending them. The citizens who most value free elections and civil liberties are precisely those who have lost the institutional power to demand them. If this dynamic generalizes, aggregate measures of public support for democracy may overstate democratic resilience by averaging across winners whose procedural commitments are shallow and losers whose commitments are deep but politically ineffectual.

The analysis contributes to the literature in two ways. First, it bridges the study of democratic conceptions and the winner-loser gap by demonstrating that electoral status shapes democratic *meaning*, not merely democratic *evaluation*---and that this effect is dynamic rather than fixed, responding to political conditions within countries over time. Second, by preserving the full structure of respondents' choices across twenty survey items through multinomial logit models rather than collapsing responses into binary procedural-substantive indicators, it reveals heterogeneity that prior approaches would obscure---including the finding that losers' procedural orientation centers on the liberal components of democracy (media freedom, free expression, the right to organize) rather than on elections per se, a distinction with implications for how we understand the relationship between electoral experience and democratic commitment.

# Positional Updating and Democratic Conceptions

The argument proceeds in three steps. First, I establish that citizens hold meaningfully different conceptions of democracy, varying in the weight they assign to procedural rules versus substantive outcomes. Second, I draw on the winner-loser gap literature to argue that electoral status creates systematically different orientations toward these dimensions. Third, I specify the mechanism---positional updating---and derive predictions that distinguish it from competing accounts.

## The procedural-substantive distinction

A foundational line of comparative research distinguishes procedural from substantive conceptions of democracy. Procedural definitions emphasize the rules and institutions through which political decisions are made: competitive elections, civil liberties, rule of law, and mechanisms of accountability [@Dahl1971-lu; @Schumpeter2010-jp; @Schedler1998-dl]. Substantive definitions focus on outcomes: economic equality, social welfare, and responsiveness to citizen demands [@Diamond1999-wd; @Linz1996-qs]. Empirical research confirms that ordinary citizens hold both types of conceptions, with the balance varying across contexts. Citizens in newer democracies tend to emphasize substantive outcomes, while those in established democracies more frequently cite procedural elements [@Russell-J-Dalton-Doh-C-Shin-and-Willy-Jou2007-gm; @Bratton2001-ug]. In Asia specifically, publics hold complex, multidimensional understandings that blend procedural and substantive elements, with substantial variation both within and between countries [@Chu2008-fk; @Chu2010-nb; @Cho2014-tj; @Shin2018-mq].

This variation matters because how citizens define democracy shapes how they evaluate it. If democracy means free elections, competitive voting may suffice for legitimacy; if it means economic equality, procedural compliance without material progress breeds disillusionment. Yet existing scholarship has largely treated conceptions as stable attributes of individuals or societies---cultural inheritances shaped by historical experience, regime type, or modernization trajectories [@Shin2011-kc; @Inglehart1997-gw; @Canache2012-fb; @Ferrin2016-se]. Relatively little work has asked what drives *within-country, within-period* variation in how individual citizens conceptualize democracy.

## Electoral status as a source of conceptual divergence

The winner-loser gap literature provides a candidate answer. A large body of research demonstrates that citizens whose preferred party won the most recent election differ systematically from those whose party lost: winners express greater institutional trust, higher satisfaction with democracy, and greater willingness to accept political outcomes [@Anderson2005-uk; @Anderson2002-cv; @Nadeau1993-vo]. The magnitude of these gaps varies with institutional context---losers' discontent is more pronounced in majoritarian systems [@Anderson1997-dh; @Singh2012-hj]---and procedural fairness perceptions mediate the relationship [@Esaiasson2011-po; @Esaiasson2020-av]. Recent extensions have confirmed these dynamics in non-Western settings and across regime types [@Blais2025-qq; @Mauk2020-bu].

This literature, however, has focused on *evaluations*---how much citizens support, trust, or feel satisfied with democracy---while treating the underlying object of evaluation as shared. Winners and losers are assumed to mean the same thing when they express views about "democracy"; they simply differ in how positively they assess it. If electoral status shapes not only evaluations but *conceptions*---if winners and losers are in fact evaluating different things---then the winner-loser gap may be deeper than the existing literature has recognized, and the two literatures' isolation from one another represents a missed theoretical opportunity.

## Positional updating: mechanism and predictions

I propose that electoral status shapes democratic conceptions through a process of positional updating. The core logic rests on the differential salience of procedures and outcomes for those in and out of power. When one's preferred party controls government, substantive outcomes command attention: what policies will be enacted, what material benefits will flow. For those excluded from power, procedures become paramount: will there be a fair chance to compete again? Will civil liberties protect against retribution? Will accountability mechanisms constrain those who govern?

This selective attention is driven by motivated cognition [@Kunda1990-uw; @Taber2006-kx]. Citizens do not arrive at their understanding of democracy through disinterested reflection; rather, they attend to and weight those dimensions of democracy that are instrumentally relevant to their political position. Losers orient toward procedural features because procedures are the tools of the excluded---the mechanisms through which power can be contested, checked, and ultimately recaptured. Winners orient toward substantive outcomes because they occupy the position from which outcomes are delivered. In this sense, the mechanism parallels the political conditioning of economic perceptions documented in other domains: just as partisan position shapes whether citizens view the economy favorably [@Evans2006-np; @Fiorina1978-wf], electoral status shapes which features of democracy citizens attend to and prioritize. Emphasizing procedural democracy also serves losers' discursive interests: a government can be criticized for violating democratic procedures in ways that a government cannot easily be criticized for failing to realize one's preferred substantive vision.

The positional account generates three testable predictions, each of which distinguishes it from the most plausible alternative---that democratic conceptions are stable cultural orientations or fixed psychological dispositions.

First, if conceptions reflect positional interests, losers should prioritize procedural over substantive elements of democracy relative to winners. This prediction is shared with the cultural account, which might attribute any observed differences to the broader political orientations of parties in opposition.

*H1:* Electoral losers will be more likely than winners to prioritize procedural over substantive conceptions of democracy.

Second, the positional logic implies that the magnitude of the gap should vary with the stakes of losing. Where power alternates regularly and institutional checks limit what winners can do, the positional distinction carries less weight---both sides retain access to democratic procedures regardless. Where losing is consequential and potentially permanent, the need for procedural protections becomes acute. Loss aversion [@Kahneman1979-af] reinforces this dynamic: the threat of losing established rights weighs more heavily than the prospect of gaining new benefits, amplifying losers' orientation toward procedural safeguards.

*H2:* The loser effect will be larger in political contexts where losing is more consequential---where democratic institutions are weaker or power alternation less frequent.

Third, and most critically for distinguishing positional updating from cultural accounts: if conceptions are cultural inheritances or stable dispositions, they should not shift appreciably within countries over periods shorter than a generation. The positional account, by contrast, predicts that conceptions will respond dynamically to changes in the political stakes of losing---growing when institutions weaken and shrinking when institutional protections are restored through alternation or constitutional resolution of crises.

*H3:* Within countries, the loser effect will respond to changes in political context, growing when democratic institutions weaken and shrinking when power alternates.

H3 is the hypothesis that most sharply distinguishes the positional account from its alternatives, and the one whose confirmation would represent the strongest evidence for the theoretical framework advanced here. The within-country trajectories examined below provide the most direct test.

# Data and Methods

## Data

The analysis draws on the Asian Barometer Survey (ABS), a regional comparative project tracking democratic values across Asia since 2001 [@ABS-W2; @ABS-W3; @ABS-W4; @ABS-W6].[^abs] The ABS covers countries spanning the full range of regime types [@Chu2016-js] and employs forced-choice items requiring respondents to prioritize among competing democratic values, providing cleaner identification of relative orientations than agree-disagree batteries.

[^abs]: Data analyzed in this article were collected by the Asian Barometer Project (2005--2008, 2010--2012, 2013--2016, 2020--2022), co-directed by Professors Fu Hu, Yun-han Chu, and Min-hua Huang, with major funding support from Taiwan's Ministry of Education, the National Science and Technology Council, Academia Sinica, and National Taiwan University. The Asian Barometer Project Office (www.asianbarometer.org) is solely responsible for data distribution. The views expressed herein are the author's own.

```{r data-overview}
#| include: false
# Count waves and country-waves
n_cw <- n_distinct(paste(cw_gap$country_name, cw_gap$wave_label))
waves_used <- c("W2 (2005-2008)", "W3 (2010-2012)", "W4 (2014-2016)", "W6 (2019-2022)")
```

Four waves are analyzed: Wave 2 (2005--2008), Wave 3 (2010--2012), Wave 4 (2014--2016), and Wave 6 (2019--2022). Wave 5 is excluded because it did not include the relevant items. China and Vietnam are excluded because their single-party systems do not produce meaningful electoral winners and losers. The analysis is restricted to respondents who reported voting in the most recent national election, as winner-loser status cannot be meaningfully assigned to non-voters. This restriction is not merely a methodological necessity but a theoretically motivated choice: voters are the citizens whose democratic conceptions most directly translate into political demands through electoral participation, party mobilization, and protest. A study of how electoral outcomes reshape democratic understanding is, by definition, a study of the electorally engaged. After these restrictions, the sample comprises `r fmt_n(total_n)` observations across `r n_countries` countries.

## Measurement

In Waves 3, 4, and 6, the ABS presented respondents with four item sets, each containing four statements describing different aspects of democracy. Respondents indicated which statement they considered most important. Each set included procedural items (e.g., "People can choose their leaders in free and fair elections," "Government protects people's freedom of speech"), substantive items (e.g., "Government reduces the gap between the rich and the poor," "Government provides everyone with the basic necessities of life"), and governance items (e.g., "Government does not waste public money," "People's safety from crime is guaranteed"). Wave 2 employed a single forced-choice item with different response options and is analyzed separately.

The key independent variable is electoral status, derived from the ABS vote-choice item (q34a in Waves 4 and 6, q33a in Wave 3, q39a in Wave 2). For each country-wave, the ABS research team codes each respondent's self-reported vote choice against the outcome of the most recent national election prior to fieldwork, classifying respondents as winners (voted for the governing party, president, or coalition partner) or losers (voted for any non-governing party). The analysis adopts this binary coding throughout. Respondents who did not vote, refused to answer, or reported an invalid vote choice are excluded, restricting the sample to citizens whose electoral participation permits meaningful assignment of winner-loser status. One country-wave is dropped entirely: Thailand Wave 4, where the February 2014 election was annulled by the Constitutional Court and the military seized power before ABS fieldwork began, leaving no valid prior election to reference. Appendix Section B documents the reference election, winning party, and runner-up for each country-wave; reports an independent verification of the ABS coding against official election commission data for Taiwan, South Korea, and Thailand; and discusses additional edge cases including Thailand's government-formation coding and Cambodia's opposition dissolution. Control variables include age, gender, education level, and urban-rural residence.

## Analytical Strategy

Rather than collapsing responses into a binary procedural–substantive indicator, the analysis models each item set as a separate multinomial logistic regression, preserving all four response categories within each set. For Waves 3, 4, and 6, which share a common instrument of four item sets (Sets 1–4), data are pooled across waves and modeled jointly. Each set takes the form:

$$
\Pr(Y_{ij} = k) = \frac{\exp(\beta_{0k} + \beta_{1k}\text{Loser}_i + \mathbf{X}_i\boldsymbol{\gamma}_k + \alpha_{ck} + \delta_{tk})}{\sum_{m=1}^{4}\exp(\beta_{0m} + \beta_{1m}\text{Loser}_i + \mathbf{X}_i\boldsymbol{\gamma}_m + \alpha_{cm} + \delta_{tm})}
$$

where $Y_{ij}$ denotes the item selected by respondent $i$ in set $j$, $k$ indexes the four response categories within each set, $\mathbf{X}_i$ is a vector of demographic controls (age, gender, education, and urban–rural residence), and $\alpha_{ck}$ and $\delta_{tk}$ represent category-specific country and wave fixed effects. The reference category is set internally by the estimation routine; because results are reported as average marginal effects (AMEs), the choice of baseline category does not affect interpretation. AMEs express the effect of loser status as percentage-point changes in the predicted probability of selecting each item, offering a direct and intuitive quantity of interest across all twenty items.

Wave 2, which employed a single forced-choice item with a distinct set of response options, is modeled separately using the same multinomial logit specification (without wave fixed effects, as it draws from a single survey round). Including Wave 2 provides an independent check on the robustness of the loser effect across different measurement instruments.

Standard errors are computed via wild cluster bootstrap at the country level, following @Cameron2008-aw. With only eleven country clusters in the pooled sample, conventional cluster-robust variance estimators are known to produce downward-biased standard errors and over-reject the null hypothesis. The wild bootstrap imposes the null hypothesis on the bootstrap data-generating process, providing more reliable inference with few clusters. This approach yields conservative tests throughout the analysis.

To examine country-level variation and within-country trajectories (H2 and H3), the multinomial logit models are re-estimated separately for each country-wave combination. For each estimate, I compute the procedural–substantive gap: the difference between the mean AME of loser status across procedural items and the mean AME across substantive items within a given country-wave. A positive gap indicates that losers disproportionately favor procedural conceptions relative to substantive ones. Confidence intervals for these gaps are obtained via the same wild bootstrap procedure. Tracking these gaps across waves within countries allows direct assessment of whether conceptions of democracy respond to changing political conditions—the central prediction distinguishing the positional account from cultural or dispositional alternatives.


# Results

## The Baseline Loser Effect

```{r pooled-summary}
#| include: false
# Summary stats for inline text
proc_items <- pooled %>% filter(item_type == "procedural")
sub_items <- pooled %>% filter(item_type == "substantive")
gov_items <- pooled %>% filter(item_type == "governance")

n_proc_pos <- sum(proc_items$ame > 0)
n_proc_total <- nrow(proc_items)
n_proc_sig <- sum(proc_items$p < 0.05)

n_sub_neg <- sum(sub_items$ame < 0)
n_sub_total <- nrow(sub_items)
n_sub_sig <- sum(sub_items$p < 0.05)

proc_range_lo <- fmt_pct(min(proc_items$ame[proc_items$ame > 0]))
proc_range_hi <- fmt_pct(max(proc_items$ame))

sub_range_lo <- fmt_pct(min(sub_items$ame))
sub_range_hi <- fmt_pct(max(sub_items$ame[sub_items$ame < 0]))

# Largest effects
largest_proc <- proc_items %>% filter(ame == max(ame))
largest_sub <- sub_items %>% filter(ame == min(ame))
largest_gov <- gov_items %>% filter(ame == min(ame))

# Free elections item
free_elections <- proc_items %>% filter(item_label == "Free elections")
```

Table 1 reports the average marginal effect of loser status on the probability of selecting each of the twenty items across the pooled sample. The pattern is consistent with H1: `r n_proc_pos` of `r n_proc_total` procedural items show positive effects (losers more likely to select), while all `r n_sub_total` substantive items show negative effects (winners more likely to select). Procedural effects range from `r proc_range_lo` to `r proc_range_hi` percentage points, with `r largest_proc$item_label` showing the largest effect (`r fmt_pp(largest_proc$ame)` pp, $p$ `r ifelse(largest_proc$p < 0.001, "< 0.001", sprintf("= %.3f", largest_proc$p))`). Substantive effects range from `r sub_range_lo` to `r sub_range_hi` pp, with `r largest_sub$item_label` as the largest (`r fmt_pp(largest_sub$ame)` pp, $p$ `r ifelse(largest_sub$p < 0.001, "< 0.001", sprintf("= %.3f", largest_sub$p))`).

```{r pooled-table}
#| output: asis

# Build the main results table
tbl <- pooled %>%
  arrange(desc(ame)) %>%
  mutate(
    Type = str_to_title(item_type),
    Item = item_label,
    `AME (pp)` = sprintf("%+.1f", ame * 100),
    SE = sprintf("%.1f", se * 100),
    Sig = sig
  ) %>%
  select(Item, Type, `AME (pp)`, SE, Sig)

if (knitr::is_latex_output()) {
  tbl_out <- tbl %>%
    kbl(
      booktabs = TRUE,
      longtable = FALSE,
      linesep = "",
      caption = "Average Marginal Effect of Loser Status on Item Selection"
    ) %>%
    kable_styling(
      latex_options = c("striped", "hold_position"),
      font_size = 10,
      full_width = TRUE
    ) %>%
    footnote(
      general = "Average marginal effects from multinomial logit with country and wave fixed effects and demographic controls (age, gender, education, urban residence). Positive values indicate losers are more likely to select the item. Bootstrap SEs clustered at the country level.",
      symbol = c("$\\\\dagger$ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001"),
      general_title = "Note: ",
      footnote_as_chunk = TRUE,
      escape = FALSE
    )
  # Make footnote rows wrap and single-space
  tbl_str <- as.character(tbl_out)
  tbl_str <- gsub("\\multicolumn{5}{l}{", "\\multicolumn{5}{p{\\linewidth}}{\\singlespacing ", tbl_str, fixed = TRUE)
  cat(tbl_str)
} else {
  tbl %>%
    kable(format = "pipe", align = c("l", "l", "l", "r", "r", "c"))
}
```

The individual effects are modest---typically two to four percentage points---but the consistency of the pattern across twenty items from five separate batteries is notable. This is not a result driven by one or two items. Winners and losers differ systematically in how they conceptualize democracy, with losers oriented toward the rules governing political competition and winners oriented toward the material benefits government provides.

One item warrants specific attention. Free elections---the most canonical procedural item---shows a *negative* AME (`r fmt_pp(free_elections$ame)` pp), meaning losers are slightly *less* likely than winners to identify free elections as the most important feature of democracy. This runs counter to H1's general prediction, but the pattern becomes interpretable in light of the remaining procedural results. The items that losers *do* favor are the liberal components of democratic governance: free expression (`r fmt_pp(proc_items %>% filter(item_label == "Free expression") %>% pull(ame))` pp), media freedom (`r fmt_pp(proc_items %>% filter(item_label == "Media freedom") %>% pull(ame))` pp), the right to organize groups (`r fmt_pp(proc_items %>% filter(item_label == "Organize groups") %>% pull(ame))` pp), and party competition (`r fmt_pp(proc_items %>% filter(item_label == "Party competition") %>% pull(ame))` pp). Losers who have experienced electoral manipulation or whose preferred parties have been dissolved by courts may have reason to distrust the electoral mechanism itself while valuing the broader liberal ecosystem---media, expression, association, judicial accountability---that enables political contestation outside formal elections. The loser effect, in other words, is not an elections effect but a *liberal-democratic* effect: losers gravitate toward the institutional protections that sustain opposition rather than the specific electoral procedure that has failed to deliver victory.

The governance items reveal an additional pattern. Winners are consistently more likely to select governance items emphasizing order and state capacity, with `r largest_gov$item_label` showing the largest effect (`r fmt_pp(largest_gov$ame)` pp, $p$ `r ifelse(largest_gov$p < 0.001, "< 0.001", sprintf("= %.3f", largest_gov$p))`). Winners conceive of democracy as a framework for stability and effective administration; losers conceive of it as a framework for contestation and the protection of rights. This order-versus-contestation tension is explored further in the discussion.

## Democratic Erosion and Institutional Resilience

```{r trajectory-stats}
#| include: false
# Key trajectory stats
# Thailand
thai_w3_pp <- round(thai_w3_gap * 100, 1)
thai_w4_pp <- round(thai_w4_gap * 100, 1)
thai_w6_pp <- round(thai_w6_gap * 100, 1)
thai_w6_ci <- thai %>% filter(wave == 6)

# South Korea
kor <- cw_gap %>% filter(country_name == "South Korea") %>% arrange(wave)
kor_w3_pp <- round(kor %>% filter(wave == 3) %>% pull(proc_sub_gap) * 100, 1)
kor_w4_pp <- round(kor %>% filter(wave == 4) %>% pull(proc_sub_gap) * 100, 1)
kor_w6_pp <- round(kor %>% filter(wave == 6) %>% pull(proc_sub_gap) * 100, 1)

# Japan
jpn <- cw_gap %>% filter(country_name == "Japan") %>% arrange(wave)

# Cambodia
khm <- cw_gap %>% filter(country_name == "Cambodia") %>% arrange(wave)
```

Figure 1 plots the procedural--substantive gap for each country across survey waves. A positive value indicates that losers disproportionately favor procedural conceptions. The within-country trajectories provide the analysis's most compelling evidence for positional updating.

```{r fig-trajectories, fig.width=10, fig.height=6}
#| fig-cap: "Loser Effect Trajectories: Procedural--Substantive Gap by Country and Wave"
#| fig-alt: "Line plot showing procedural-substantive gap trajectories across three survey waves for eleven Asian countries. Thailand shows dramatic increase from near zero to over 20 percentage points. South Korea remains near zero throughout. Cambodia maintains a high gap around 13 points. Other countries shown in gray."

# Plot countries with 2+ waves
plot_data <- cw_gap %>%
  group_by(country_name) %>%
  filter(n() >= 2) %>%
  ungroup()

highlight <- c("Thailand", "South Korea", "Japan", "Cambodia")

plot_data %>%
  mutate(
    highlighted = country_name %in% highlight,
    country_label = if_else(highlighted, country_name, "Other")
  ) %>%
  ggplot(aes(x = wave, y = proc_sub_gap * 100, group = country_name)) +
  geom_line(
    data = . %>% filter(!highlighted),
    color = "gray75", linewidth = 0.5, alpha = 0.7
  ) +
  geom_point(
    data = . %>% filter(!highlighted),
    color = "gray75", size = 1.5, alpha = 0.7
  ) +
  geom_ribbon(
    data = . %>% filter(highlighted),
    aes(ymin = proc_sub_gap_ci_low * 100, ymax = proc_sub_gap_ci_high * 100,
        fill = country_name),
    alpha = 0.15, color = NA
  ) +
  geom_line(
    data = . %>% filter(highlighted),
    aes(color = country_name), linewidth = 1.2
  ) +
  geom_point(
    data = . %>% filter(highlighted),
    aes(color = country_name), size = 3
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous(
    breaks = c(3, 4, 6),
    labels = c("W3\n(2010-12)", "W4\n(2014-16)", "W6\n(2019-22)")
  ) +
  scale_color_manual(values = c(
    "Thailand" = "#D62728", "Cambodia" = "#FF7F0E",
    "South Korea" = "#2CA02C", "Japan" = "#1F77B4"
  )) +
  scale_fill_manual(values = c(
    "Thailand" = "#D62728", "Cambodia" = "#FF7F0E",
    "South Korea" = "#2CA02C", "Japan" = "#1F77B4"
  )) +
  labs(
    x = "Survey Wave",
    y = "Procedural - Substantive Gap\n(percentage points)",
    color = NULL, fill = NULL,
    caption = "Note: Gray lines show remaining countries. Shaded bands show 95% CIs for highlighted countries.\nPositive values = losers favor procedural items relative to substantive items."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

Thailand provides the most dramatic support for H3. In Wave 3 (2010--2012), during the Democrat Party government of Abhisit Vejjajiva, the procedural--substantive gap was negligible (`r thai_w3_pp` pp, not significant). By Wave 4 (2014--2016), conducted around the time of the May 2014 coup that removed Prime Minister Yingluck Shinawatra, the gap had widened to `r thai_w4_pp` pp ($p < 0.05$). By Wave 6 (2019--2022), after six years of military or military-backed governance, the gap reached `r thai_w6_pp` pp ($p < 0.001$, 95% CI: \[`r round(thai_w6_ci$proc_sub_gap_ci_low * 100, 1)`, `r round(thai_w6_ci$proc_sub_gap_ci_high * 100, 1)`\]). This trajectory---from zero to over twenty percentage points in a decade---tracks Thailand's democratic erosion with remarkable precision [@Kuhonta2014-qk; @Kongkirati2020-fl]. As coups and judicial interventions repeatedly overturned electoral outcomes, supporters of the excluded Thaksin-aligned faction increasingly defined democracy in procedural terms, while supporters of the military-backed order emphasized substantive outcomes.

South Korea presents the mirror image. Despite experiencing its own political upheaval during this period---the impeachment and removal of President Park Geun-hye in 2016--2017 and the subsequent election of progressive Moon Jae-in---the procedural--substantive gap remained modest throughout, fluctuating between `r kor_w3_pp` pp in Wave 3 and `r kor_w4_pp` pp in Wave 4, never exceeding five percentage points. The critical difference is institutional: South Korea's crisis was resolved through constitutional procedures rather than military intervention. The impeachment demonstrated that accountability mechanisms functioned, and the 2017 election delivered genuine alternation. Under these conditions, the positional logic predicts convergence, and convergence is what the data show.

Cambodia shows a persistently large gap (approximately 13 pp across all three waves) that, notably, did not grow following the dissolution of the main opposition party (CNRP) in 2017. In a competitive authoritarian context, the stakes of losing were already effectively maximized---losers' procedural orientation had little room to intensify further. Among the remaining countries (@tbl-country-gaps), Malaysia shows a large gap that narrows across waves, potentially reflecting the political opening that culminated in the 2018 alternation. Myanmar's single observation captures a strikingly large gap consistent with the high stakes of its fragile democratic transition. Taiwan's trajectory is distinctive, reversing from positive to negative by Wave 6---a pattern that may reflect the unusual dynamics of cross-strait identity politics. Japan provides the expected baseline for a consolidated democracy: modest and stable across waves.

```{r tbl-country-gaps}
#| tbl-cap: "Procedural--Substantive Gap by Country and Wave"

gap_tbl <- cw_gap %>%
  mutate(
    gap_label = sprintf("%+.1f [%+.1f, %+.1f]",
                        proc_sub_gap * 100,
                        proc_sub_gap_ci_low * 100,
                        proc_sub_gap_ci_high * 100)
  ) %>%
  select(country_name, wave_label, gap_label) %>%
  pivot_wider(names_from = wave_label, values_from = gap_label, values_fill = "---") %>%
  rename(Country = country_name) %>%
  select(Country, W3, W4, W6)

if (knitr::is_latex_output()) {
  gap_tbl %>%
    kbl(
      booktabs = TRUE,
      longtable = FALSE,
      linesep = ""
    ) %>%
    kable_styling(
      latex_options = c("striped", "hold_position"),
      font_size = 10
    ) %>%
    column_spec(1, width = "3cm") %>%
    column_spec(2, width = "3.5cm") %>%
    column_spec(3, width = "3.5cm") %>%
    column_spec(4, width = "3.5cm") %>%
    footnote(
      general = "Gap = (Mean Procedural AME) - (Mean Substantive AME), in percentage points, with 95% bootstrap CIs. Positive values indicate losers favor procedural items relative to substantive items. Dashes indicate country not surveyed in that wave.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE,
      threeparttable = TRUE
    )
} else {
  gap_tbl %>%
    kable(format = "pipe", align = c("l", "c", "c", "c"))
}
```

## Mechanism Test: Perceived Electoral Fairness

```{r fairness-data}
#| include: false
# Load fairness interaction results
fairness <- read_csv(file.path(results_path, "robustness_fairness_interaction.csv"), show_col_types = FALSE)

# Key interaction effects
court_int <- fairness %>% filter(item == "Court protection") %>% pull(interaction)
media_int <- fairness %>% filter(item == "Media freedom") %>% pull(interaction)
expression_int <- fairness %>% filter(item == "Free expression") %>% pull(interaction)
elections_int <- fairness %>% filter(item == "Free elections") %>% pull(interaction)

# Count positive interactions among procedural items
proc_fairness <- fairness %>% filter(item_type == "procedural")
n_proc_pos_int <- sum(proc_fairness$interaction > 0)
n_proc_int_total <- nrow(proc_fairness)
```

The positional updating account predicts that the loser effect should intensify among losers who perceive elections as unfair, since these citizens face the most acute procedural threat. If instead the loser effect reflects stable normative commitments to procedural democracy, it should appear regardless of whether losers perceive the process as fair---committed proceduralists would prioritize procedures as a matter of principle, not instrumental concern. To test this, I interact loser status with a binary indicator of whether respondents perceived the most recent election as unfair, following @Mauk2022-xx.

```{r fig-fairness}
#| fig-cap: "Fairness Interaction: Loser Effect by Perceived Electoral Fairness"
#| fig-alt: "Dot plot showing loser effects separately for respondents perceiving fair vs. unfair elections, across sixteen items. Liberal procedural items show substantially larger loser effects among those perceiving unfair elections."
#| fig-width: 8
#| fig-height: 6

# Plot interaction effects
fairness %>%
  mutate(
    item = fct_reorder(item, interaction),
    item_type = str_to_title(item_type)
  ) %>%
  pivot_longer(
    cols = c(loser_effect_fair, loser_effect_unfair),
    names_to = "condition",
    values_to = "loser_effect"
  ) %>%
  mutate(
    condition = if_else(condition == "loser_effect_fair", "Fair election", "Unfair election")
  ) %>%
  ggplot(aes(x = loser_effect * 100, y = item, color = condition, shape = condition)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  facet_wrap(~ item_type, scales = "free_y", ncol = 1) +
  scale_color_manual(values = c("Fair election" = "#1F77B4", "Unfair election" = "#D62728")) +
  labs(
    x = "Loser Effect (percentage points)",
    y = NULL,
    color = NULL, shape = NULL,
    caption = "Note: Loser AMEs from multinomial logit interacting loser status with perceived electoral fairness.\nCountry and wave fixed effects with demographic controls."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

The results support the positional account. Among `r n_proc_int_total` procedural items, `r n_proc_pos_int` show a positive interaction---meaning the loser effect is larger among those perceiving elections as unfair. The amplification is especially pronounced for the liberal procedural items: court protection shows an interaction of `r sprintf("%+.1f", court_int * 100)` percentage points, media freedom `r sprintf("%+.1f", media_int * 100)` pp, and free expression `r sprintf("%+.1f", expression_int * 100)` pp. Losers who perceive the electoral process as compromised do not simply withdraw from procedural commitments; they double down on precisely the institutional safeguards---judicial protection, press freedom, expressive liberty---that would constrain those who benefit from unfair elections. The free elections item, by contrast, shows the opposite pattern (interaction: `r sprintf("%+.1f", elections_int * 100)` pp), reinforcing the interpretation from the pooled results: losers who perceive elections as unfair are *less* inclined to identify elections themselves as democracy's essential feature, even as they intensify their commitment to the liberal infrastructure surrounding elections.

These interaction patterns are difficult to reconcile with a stable-commitments account. If the loser effect reflected fixed normative orientations, it should not vary systematically with perceptions of electoral fairness. That it does---and that the amplification concentrates on liberal protections rather than elections per se---suggests that democratic conceptions respond to the perceived severity of procedural threat, consistent with the positional updating mechanism.

# Robustness

Several additional tests assess the sensitivity of the main findings. First, a weighted least squares (WLS) estimation that aggregates the data to the country-wave level sidesteps the clustering problem entirely: nine of ten procedural items show positive loser effects (six significant), all six substantive items show negative effects (five significant), with mean effect sizes comparable to the individual-level estimates (+2.2 pp procedural, -2.4 pp substantive; Appendix G).

Second, Wave 2 of the ABS, which employed a different forced-choice instrument, provides an independent replication. Both procedural items show positive AMEs and both substantive items show negative AMEs, with directional consistency that increases confidence the loser effect is not an artifact of the Wave 3--6 battery design. Open-ended responses from Wave 2 confirm convergent validity: respondents selecting procedural items were significantly more likely to offer procedural definitions spontaneously ($r = 0.28$, $p < .001$; Appendix F).[^1]

[^1]: Following established conceptual frameworks, open-ended responses were coded as procedural, substantive, or excluded. A respondent-level procedural proportion score (0--1) based on up to three codeable responses correlates at $r = 0.28$ with the closed-ended item. Given the noise inherent in open-ended coding, this represents meaningful convergent validity.

Third, Wave 6 fieldwork coincided with the COVID-19 pandemic, which may have compressed the procedural-substantive gap by increasing the salience of state capacity for all citizens. Re-estimating Wave 6 models with controls for pandemic attitudes yields essentially unchanged results (mean procedural AME: `r fmt_pp(covid_proc_base)` → `r fmt_pp(covid_proc_covid)` pp; mean substantive AME: `r fmt_pp(covid_sub_base)` → `r fmt_pp(covid_sub_covid)` pp). Thailand's gap persists through the pandemic, consistent with democratic erosion overwhelming the homogenizing pull of COVID-19.

Fourth, non-voters resemble winners rather than losers in their procedural-substantive gap (14.3 pp vs. 13.9 pp for winners and 17.3 pp for losers), confirming that restricting the sample to voters does not inflate the estimated loser effect (Appendix H). Fifth, a three-way decomposition distinguishing key opposition supporters from other losers yields no clear differentiation across the three countries where it was feasible, likely reflecting limited coverage and additional missingness (Appendix J). The binary winner-loser distinction provides broader coverage and cleaner identification.

Finally, the use of repeated cross-sections rather than panel data means within-country trajectories cannot definitively rule out compositional change. The Thai case offers the strongest counterargument: party loyalties in Thailand's color-coded politics are notably stable, and the gap's tracking of documented political events---rather than gradual demographic shifts---favors the positional interpretation.

# Discussion

The central finding of this article is not that electoral losers prefer procedural democracy — though they do, consistently, across twenty items and eleven countries. The finding that matters is that this preference *moves*. Thailand's trajectory from near-zero to over twenty percentage points across a single decade of democratic erosion cannot be reconciled with accounts treating democratic conceptions as stable cultural inheritances or fixed psychological dispositions [@Norris2011-zt; @Inglehart1997-gw]. Nor is it plausibly compositional: Thailand's color-coded partisan alignments remained remarkably stable even as the parties themselves were repeatedly dissolved and reconstituted [@Sinpeng2021-hd; @McCargo2019-rx]. Citizens updated their understanding of what democracy fundamentally means in response to changes in the political stakes of losing — the pattern predicted by the positional updating account and confirmed by the fairness interaction results, which show that the loser effect intensifies precisely among those who perceive elections as compromised.

This dynamic carries a troubling implication for the study of democratic resilience. A growing literature asks whether mass public support for democracy can serve as a bulwark against backsliding [@Claassen2020-hk; @Svolik2019-zi; @Graham2020-jr]. The present results suggest a structural problem with this expectation: where erosion is most severe, procedural commitments become concentrated among those whom the political system has rendered least capable of defending them. Aggregate measures of democratic support may therefore overstate resilience by averaging across winners whose procedural commitments are shallow and losers whose commitments are deep but politically ineffectual. The mechanisms through which democracies erode — executive aggrandizement, judicial capture, media suppression [@Levitsky2018-dd; @Waldner2018-yw; @Luhrmann2019-ow] — are precisely the mechanisms that disarm the constituency most committed to procedural norms. Whether losers' procedural orientation can translate into effective resistance, through protest, civil disobedience, or transnational advocacy, remains an open and urgent question.

The results also complicate the winner-loser gap literature in a way that has implications for survey research. Previous work has assumed that winners and losers share a common understanding of democracy when they report their satisfaction with it or their trust in its institutions [@Anderson2005-uk; @Esaiasson2011-po; @Norris2012-wj]. If winners and losers define democracy differently — winners emphasizing outcomes, losers emphasizing procedures — then the standard "satisfaction with democracy" question is not measuring the same thing for both groups. Losers may report dissatisfaction because they perceive procedural violations; winners may report satisfaction because they perceive policy delivery. This disconnect suggests that the much-studied winner-loser gap in democratic satisfaction may partly reflect a *measurement* artifact: not different evaluations of the same object, but evaluations of different objects.

The item-level results add a further dimension. The loser effect is not a generic procedural orientation but a specifically *liberal-democratic* one. Losers gravitate toward media freedom, free expression, the right to organize, and judicial accountability — the institutional infrastructure that sustains political opposition — rather than toward elections per se. This pattern, reinforced by the negative interaction between loser status and perceived electoral fairness on the free elections item, suggests that electoral experience teaches losers a nuanced lesson: the value of democracy lies not in the electoral mechanism that has failed them but in the broader liberal ecosystem that preserves the possibility of contestation. The distinction between electoral and liberal conceptions of procedural democracy, latent in theoretical frameworks from @Dahl1971-lu to @Schedler1998-dl, appears to be psychologically real among citizens navigating the consequences of electoral loss.

Finally, the order-versus-contestation tension in the governance items deserves further investigation. Winners' strong preference for law and order as a defining feature of democracy, combined with losers' emphasis on civil liberties and accountability, captures a fault line that @Slater2010-ba has identified as central to the politics of order in Southeast Asia. Where democratic institutions are robust, the demands for order and for contestation coexist in productive tension. Where they are not, the two conceptions may become mutually exclusive — and in the current period of global democratic recession [@Diamond2015-yb; @Foa2016-lq], understanding which conception prevails, and among whom, has consequences well beyond the region studied here.

Several limitations temper these conclusions. The analysis relies on repeated cross-sections rather than panel data, precluding direct observation of individual-level attitude change; the within-country trajectories are suggestive but not definitive. The sample comprises only eleven countries, limiting cross-national inference. The Wave 6 compression across most countries outside Thailand remains incompletely explained, though the COVID-19 pandemic's homogenizing effect on the salience of state capacity is a plausible contributor — a prediction that post-pandemic survey waves will be positioned to test. And the exclusion of non-voters, while theoretically motivated, means the findings characterize the politically engaged electorate rather than the population at large.

# Conclusion

This article has shown that electoral status shapes not merely citizens' satisfaction with democracy but their understanding of what democracy fundamentally means. Losers prioritize the liberal-procedural infrastructure — free expression, media freedom, judicial accountability, the right to organize — that sustains the possibility of political contestation. Winners prioritize substantive outcomes and the governance of order. This divergence is not a reflection of stable values but a positional response to the political stakes of exclusion, as the dramatic within-country trajectories in Thailand and South Korea demonstrate.

The most consequential implication is structural. Democratic erosion does not merely reduce the quality of governance or diminish civil liberties; it reshapes the very meaning of democracy along the winner-loser divide, concentrating procedural commitments among those least able to act on them. If the positional logic holds beyond Asia — a question that comparative research in Latin America and Africa is well positioned to address — then the relationship between public attitudes and democratic resilience may be more fragile than the existing literature suggests. Future research should examine whether losers' procedural orientation translates into distinct patterns of political behavior, particularly unconventional participation, testing whether democratic conceptions serve as blueprints for resistance or remain, ultimately, survey artifacts.

# References {.unnumbered}

\singlespacing

::: {#refs}
:::
