---
title: "The Meaning of Losing: How Democratic Erosion Reshapes Citizens' Conceptions of Democracy"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: today
abstract: |
  Does democratic erosion reshape not just citizens' satisfaction with democracy but their understanding of what democracy means? Using multinomial logit models on four waves of Asian Barometer Survey data (2005–2022) across eleven countries, this article finds that electoral losers systematically prioritize liberal-procedural conceptions of democracy—free expression, media freedom, judicial accountability—while winners emphasize substantive outcomes like economic equality and welfare. This association varies dynamically with political context. In Thailand, the gap between winners' and losers' conceptions widened from near zero to over twenty percentage points as democratic institutions eroded after the 2014 coup; in South Korea, where a comparable crisis was resolved constitutionally, the gap remained negligible. Multiple robustness checks support the positional updating interpretation: the loser effect (1) amplifies among those perceiving elections as unfair, (2) does not extend to non-procedural domains (placebo test), and (3) persists after demographic reweighting (composition test). If this pattern reflects positional updating rather than stable sorting, the implications are troubling: erosion would concentrate procedural commitments among those whom the political system has rendered least capable of defending them—the citizens who most value democratic safeguards are precisely those who have lost the power to demand them.
keywords:
  - democracy
  - electoral losers
  - procedural democracy
  - Asian Barometer Survey
  - democratic consolidation
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: false
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \raggedbottom
          \floatplacement{figure}{H}
          \newcommand{\keywords}[1]{\vspace{0.5em}\begin{singlespace}\noindent\textbf{Keywords:} #1\end{singlespace}}
  html:
    toc: true
    toc-depth: 3
    number-sections: false
    theme: cosmo
  docx:
    reference-doc: custom-reference.docx
    number-sections: false
csl: chicago-author-date.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.pos = "H", out.width = "100%"
)

here::i_am("papers/meaning-of-democracy-revision/manuscript/manuscript.qmd")

# Load results from multinomial analysis pipeline
results_path <- here("papers", "meaning-of-democracy-revision", "analysis", "revised", "results")

res <- readRDS(file.path(results_path, "mlogit_results.rds"))
cw_gap <- read_csv(file.path(results_path, "country_wave_proc_sub_gap.csv"), show_col_types = FALSE)
cw_items <- read_csv(file.path(results_path, "country_wave_item_ames.csv"), show_col_types = FALSE)

# Helpers
fmt_n <- function(x) format(x, big.mark = ",")
fmt_pct <- function(x) sprintf("%.1f", x * 100)
fmt_pp <- function(x) sprintf("%+.1f", x * 100)

# Key quantities
pooled <- res$all_results
n_countries <- n_distinct(cw_gap$country_name)

# Thailand trajectory
thai <- cw_gap %>% filter(country_name == "Thailand") %>% arrange(wave)
thai_w3_gap <- thai %>% filter(wave == 3) %>% pull(proc_sub_gap)
thai_w4_gap <- thai %>% filter(wave == 4) %>% pull(proc_sub_gap)
thai_w6_gap <- thai %>% filter(wave == 6) %>% pull(proc_sub_gap)
thai_change <- (thai_w6_gap - thai_w3_gap) * 100

# Total N (approximate from item-level data)
total_n <- cw_items %>%
  filter(set == "Set1") %>%
  group_by(country_name, wave_label) %>%
  slice(1) %>%
  ungroup() %>%
  pull(n) %>%
  sum()

# COVID robustness summary
covid_sum <- readRDS(file.path(results_path, "robustness_covid_summary.rds"))
covid_proc_base <- covid_sum %>% filter(item_type == "procedural", controls == "base") %>% pull(mean_ame)
covid_proc_covid <- covid_sum %>% filter(item_type == "procedural", controls == "covid") %>% pull(mean_ame)
covid_sub_base <- covid_sum %>% filter(item_type == "substantive", controls == "base") %>% pull(mean_ame)
covid_sub_covid <- covid_sum %>% filter(item_type == "substantive", controls == "covid") %>% pull(mean_ame)

# Placebo and composition robustness stats
placebo <- readRDS(file.path(results_path, "placebo_stats.rds"))
composition <- readRDS(file.path(results_path, "composition_stats.rds"))
endog <- readRDS(file.path(results_path, "endogeneity_stats.rds"))

# Helper functions for formatting
fmt_coef <- function(x) sprintf("%.3f", x)
fmt_p <- function(p) {
  if (p < 0.001) return("< 0.001")
  sprintf("%.3f", p)
}
```

\keywords{democracy, electoral losers, procedural democracy, Asian Barometer Survey, democratic consolidation}

# Introduction

Between 2010 and 2020, Thailand experienced a constitutional court's annulment of an election, a military coup, six years of junta governance, and the return to civilian rule under a military-drafted constitution that ensured the army's continued dominance. During this same decade, the gap between how electoral winners and losers conceptualized democracy widened from near zero to over `r round(thai_w6_gap * 100)` percentage points. Supporters of the excluded Thaksin-aligned parties increasingly defined democracy in terms of free elections, civil liberties, and government accountability; supporters of the military-backed order increasingly emphasized economic welfare and public services. This is not a story about who was more satisfied with democracy or more trusting of government—the attitudes that the winner-loser literature has studied extensively [@Anderson2005-uk; @Singh2012-hj]. It is a story about what citizens came to believe democracy *means*.

South Korea offers the contrast. Over the same period, the country weathered its own crisis: the impeachment and criminal prosecution of President Park Geun-hye, mass candlelight protests, and the election of a progressive successor. Yet the gap between winners' and losers' conceptions of democracy never exceeded five percentage points. The critical difference was institutional. South Korea's crisis was resolved through constitutional procedures—impeachment, judicial review, a free election—that demonstrated the system's capacity for self-correction. Thailand's crisis was resolved by tanks. Where institutions held, winners and losers converged on what democracy meant; where institutions collapsed, democratic meaning itself fractured along the winner-loser divide.

These contrasting trajectories point to a question that two established research literatures have, surprisingly, not addressed together. One body of scholarship documents how citizens conceptualize democracy—whether in procedural terms emphasizing elections and rights, or substantive terms emphasizing welfare and equality—and finds considerable variation across and within countries [@Russell-J-Dalton-Doh-C-Shin-and-Willy-Jou2007-gm; @Chu2008-fk; @Ferrin2016-se]. A separate literature demonstrates that electoral winners and losers differ systematically in their satisfaction with democracy, their institutional trust, and their willingness to accept political outcomes [@Anderson2005-uk; @Anderson2002-cv; @Esaiasson2011-po]. But while the first literature has treated democratic conceptions as relatively stable cultural orientations, and the second has assumed that winners and losers are evaluating the same thing when they express views about "democracy," neither has examined whether electoral status shapes not just how citizens *evaluate* democracy but how they *understand* it.

This article demonstrates that it does. Using multinomial logit models on four waves of Asian Barometer Survey data (2005--2022) encompassing more than `r fmt_n(total_n)` voters across `r n_countries` countries, I find that electoral losers systematically prioritize procedural elements of democracy—competitive elections, civil liberties, government accountability—over substantive outcomes such as economic equality or welfare provision. This "loser effect" is individually modest in pooled cross-national estimates, typically two to four percentage points per item, but it is directionally consistent across twenty items from five separate survey batteries. More importantly, the effect varies with political context: it tracks democratic erosion in Thailand, remains stable amid institutional resilience in South Korea, and persists at high levels in Cambodia's competitive authoritarian regime where the stakes of losing were always extreme.

The pattern is consistent with positional updating, supported by three converging pieces of evidence. First, the loser effect amplifies among those perceiving elections as unfair—precisely where procedural threat is most salient. Second, a placebo test shows the amplification does not extend to material welfare perceptions, ruling out generalized discontent. Third, demographic reweighting confirms the Thailand trajectory is not driven by compositional change. While these tests do not definitively establish causation, they are difficult to reconcile with accounts treating democratic conceptions as stable cultural orientations, and they tilt the evidence toward a positional interpretation.

The finding carries a troubling implication for the study of democratic resilience. A growing body of work examines whether mass democratic attitudes can serve as a bulwark against backsliding [@Claassen2020-hk; @Svolik2019-zi; @Graham2020-jr]. The present results suggest that where erosion is most severe, procedural commitments become concentrated among those whom the political system has rendered least capable of defending them. The citizens who most value free elections and civil liberties are precisely those who have lost the institutional power to demand them. If this dynamic generalizes, aggregate measures of public support for democracy may overstate democratic resilience by averaging across winners whose procedural commitments are shallow and losers whose commitments are deep but politically ineffectual.

The analysis contributes to the literature in two ways. First, it bridges the study of democratic conceptions and the winner-loser gap by demonstrating that electoral status shapes democratic *meaning*, not merely democratic *evaluation*—and that this effect is dynamic rather than fixed, responding to political conditions within countries over time. Second, by preserving the full structure of respondents' choices across twenty survey items through multinomial logit models rather than collapsing responses into binary procedural-substantive indicators, it reveals heterogeneity that prior approaches would obscure—including the finding that losers' procedural orientation centers on the liberal components of democracy (media freedom, free expression, the right to organize) rather than on elections per se, a distinction with implications for how we understand the relationship between electoral experience and democratic commitment.

# Positional Updating and Democratic Conceptions

The argument proceeds in three steps. First, I establish that citizens hold meaningfully different conceptions of democracy, varying in the weight they assign to procedural rules versus substantive outcomes. Second, I draw on the winner-loser gap literature to argue that electoral status creates systematically different orientations toward these dimensions. Third, I specify the mechanism—positional updating—and derive predictions that distinguish it from competing accounts.

## The procedural-substantive distinction

A foundational line of comparative research distinguishes procedural from substantive conceptions of democracy. Procedural definitions emphasize the rules and institutions through which political decisions are made: competitive elections, civil liberties, rule of law, and mechanisms of accountability [@Dahl1971-lu; @Schumpeter2010-jp; @Schedler1998-dl]. Substantive definitions focus on outcomes: economic equality, social welfare, and responsiveness to citizen demands [@Diamond1999-wd; @Linz1996-qs]. Empirical research confirms that ordinary citizens hold both types of conceptions, with the balance varying across contexts. Citizens in newer democracies tend to emphasize substantive outcomes, while those in established democracies more frequently cite procedural elements [@Russell-J-Dalton-Doh-C-Shin-and-Willy-Jou2007-gm; @Bratton2001-ug]. In Asia specifically, publics hold complex, multidimensional understandings that blend procedural and substantive elements, with substantial variation both within and between countries [@Chu2008-fk; @Chu2010-nb; @Cho2014-tj; @Shin2018-mq]. Partisan dynamics shape these orientations: @Hsiao2020-fz show that affective polarization in Taiwan erodes support for democracy, suggesting that political position—not just cultural background—conditions how citizens relate to democratic principles.

This variation matters because how citizens define democracy shapes how they evaluate it. If democracy means free elections, competitive voting may suffice for legitimacy; if it means economic equality, procedural compliance without material progress breeds disillusionment. Yet existing scholarship has largely treated conceptions as stable attributes of individuals or societies—cultural inheritances shaped by historical experience, regime type, or modernization trajectories [@Shin2011-kc; @Inglehart1997-gw; @Canache2012-fb; @Ferrin2016-se]. Relatively little work has asked what drives *within-country, within-period* variation in how individual citizens conceptualize democracy.

## Electoral status as a source of conceptual divergence

The winner-loser gap literature provides a candidate answer. A large body of research demonstrates that citizens whose preferred party won the most recent election differ systematically from those whose party lost: winners express greater institutional trust, higher satisfaction with democracy, and greater willingness to accept political outcomes [@Anderson2005-uk; @Anderson2002-cv; @Nadeau1993-vo]. The magnitude of these gaps varies with institutional context—losers' discontent is more pronounced in majoritarian systems [@Anderson1997-dh; @Singh2012-hj]—and procedural fairness perceptions mediate the relationship [@Esaiasson2011-po; @Esaiasson2020-av]. Recent extensions have confirmed these dynamics in non-Western settings and across regime types [@Blais2025-qq; @Mauk2020-bu]. @Ricks2025-il demonstrate that even in Thailand's hybrid regime, the 2023 election produced winner-loser divergences in attitudes toward state institutions, suggesting the mechanism operates beyond fully democratic contexts.

This literature, however, has focused on *evaluations*—how much citizens support, trust, or feel satisfied with democracy—while treating the underlying object of evaluation as shared. Winners and losers are assumed to mean the same thing when they express views about "democracy"; they simply differ in how positively they assess it. If electoral status shapes not only evaluations but *conceptions*—if winners and losers are in fact evaluating different things—then the winner-loser gap may be deeper than the existing literature has recognized, and the two literatures' isolation from one another represents a missed theoretical opportunity.

## Positional updating: mechanism and predictions

I propose that electoral status shapes democratic conceptions through a process of positional updating. The core logic rests on the differential salience of procedures and outcomes for those in and out of power. When one's preferred party controls government, substantive outcomes command attention: what policies will be enacted, what material benefits will flow. For those excluded from power, procedures become paramount: will there be a fair chance to compete again? Will civil liberties protect against retribution? Will accountability mechanisms constrain those who govern?

This selective attention is driven by motivated cognition [@Kunda1990-uw; @Taber2006-kx]. Citizens do not arrive at their understanding of democracy through disinterested reflection; rather, they attend to and weight those dimensions of democracy that are instrumentally relevant to their political position. Losers orient toward procedural features because procedures are the tools of the excluded—the mechanisms through which power can be contested, checked, and ultimately recaptured. Winners orient toward substantive outcomes because they occupy the position from which outcomes are delivered. In this sense, the mechanism parallels the political conditioning of economic perceptions documented in other domains: just as partisan position shapes whether citizens view the economy favorably [@Evans2006-np; @Fiorina1978-wf], electoral status shapes which features of democracy citizens attend to and prioritize. Emphasizing procedural democracy also serves losers' discursive interests: a government can be criticized for violating democratic procedures in ways that a government cannot easily be criticized for failing to realize one's preferred substantive vision.

Critically, the positional updating framework advanced here does not require that losers strategically calculate their prospects of returning to power. Rather, the mechanism operates through a more fundamental channel: exclusion from the exercise of state authority. Citizens who supported losing parties or candidates—regardless of their partisan organizational strength—find themselves outside the coalition that controls government institutions, shapes policy, and distributes resources. This structural position of exclusion generates heightened attention to the procedural safeguards that constrain incumbent power, not because losers adopt a calculated rhetorical posture, but because procedural protections become practically salient for those who cannot rely on access to executive authority to advance their interests.

This exclusion-based account carries a distinct empirical implication. If positional updating were driven primarily by partisan strategy—with opposition elites signaling procedural commitments to mobilize supporters—one would expect the effect to concentrate among supporters of the principal opposition party, whose leaders have both the organizational capacity and the electoral incentive to champion procedural norms. By contrast, if positional updating reflects the broader condition of exclusion from power, the procedural orientation should extend across all categories of losers, including supporters of minor parties and political independents who lack any realistic path back to government. As the supplementary analysis in Appendix J demonstrates, this is precisely what the data reveal.

The positional account generates three testable predictions, each of which distinguishes it from the most plausible alternative—that democratic conceptions are stable cultural orientations or fixed psychological dispositions.

First, if conceptions reflect positional interests, losers should prioritize procedural over substantive elements of democracy relative to winners. This prediction is shared with the cultural account, which might attribute any observed differences to the broader political orientations of parties in opposition.

*H1:* Electoral losers will be more likely than winners to prioritize procedural over substantive conceptions of democracy.

Second, the positional logic implies that the magnitude of the gap should vary with the stakes of losing. Where power alternates regularly and institutional checks limit what winners can do, the positional distinction carries less weight—both sides retain access to democratic procedures regardless. Where losing is consequential and potentially permanent, the need for procedural protections becomes acute. Loss aversion [@Kahneman1979-af] reinforces this dynamic: the threat of losing established rights weighs more heavily than the prospect of gaining new benefits, amplifying losers' orientation toward procedural safeguards.

*H2:* The loser effect will be larger in political contexts where losing is more consequential—where democratic institutions are weaker or power alternation less frequent.

Third, and most critically for distinguishing positional updating from cultural accounts: if conceptions are cultural inheritances or stable dispositions, they should not shift appreciably within countries over periods shorter than a generation. The positional account, by contrast, predicts that conceptions will respond dynamically to changes in the political stakes of losing—growing when institutions weaken and shrinking when institutional protections are restored through alternation or constitutional resolution of crises.

*H3:* Within countries, the loser effect will respond to changes in political context, growing when democratic institutions weaken and shrinking when power alternates.

H3 is the hypothesis that most sharply distinguishes the positional account from its alternatives, and the one whose confirmation would represent the strongest evidence for the theoretical framework advanced here. The within-country trajectories examined below provide the most direct test.

# Data and Methods

## Data

The analysis draws on the Asian Barometer Survey (ABS), a regional comparative project tracking democratic values across Asia since 2001 [@ABS-W2; @ABS-W3; @ABS-W4; @ABS-W6].[^1] The ABS covers countries spanning the full range of regime types [@Chu2016-js] and employs forced-choice items requiring respondents to prioritize among competing democratic values, providing cleaner identification of relative orientations than agree-disagree batteries.

[^1]: Data analyzed in this article were collected by the Asian Barometer Project (2005--2008, 2010--2012, 2013--2016, 2020--2022), co-directed by Professors Fu Hu, Yun-han Chu, and Min-hua Huang, with major funding support from Taiwan's Ministry of Education, the National Science and Technology Council, Academia Sinica, and National Taiwan University. The Asian Barometer Project Office (www.asianbarometer.org) is solely responsible for data distribution. The views expressed herein are the author's own.

```{r data-overview}
#| include: false
# Count waves and country-waves
n_cw <- n_distinct(paste(cw_gap$country_name, cw_gap$wave_label))
waves_used <- c("W2 (2005-2008)", "W3 (2010-2012)", "W4 (2014-2016)", "W6 (2019-2022)")
```

Four waves are analyzed: Wave 2 (2005--2008), Wave 3 (2010--2012), Wave 4 (2014--2016), and Wave 6 (2019--2022). Wave 5 is excluded because it did not include the relevant items. China and Vietnam are excluded because their single-party systems do not produce meaningful electoral winners and losers. The analysis is restricted to respondents who reported voting in the most recent national election, as winner-loser status cannot be meaningfully assigned to non-voters. This restriction is not merely a methodological necessity but a theoretically motivated choice: voters are the citizens whose democratic conceptions most directly translate into political demands through electoral participation, party mobilization, and protest. A study of how electoral outcomes reshape democratic understanding is, by definition, a study of the electorally engaged. After these restrictions, the sample comprises `r fmt_n(total_n)` observations across `r n_countries` countries.

## Measurement

In Waves 3, 4, and 6, the ABS presented respondents with four item sets, each containing four statements describing different aspects of democracy. Respondents indicated which statement they considered most important. Each set included procedural items (e.g., "People can choose their leaders in free and fair elections," "Government protects people's freedom of speech"), substantive items (e.g., "Government reduces the gap between the rich and the poor," "Government provides everyone with the basic necessities of life"), and governance items (e.g., "Government does not waste public money," "People's safety from crime is guaranteed"). Wave 2 employed a single forced-choice item with different response options and is analyzed separately.

The key independent variable is electoral status, derived from the ABS vote-choice item (q34a in Waves 4 and 6, q33a in Wave 3, q39a in Wave 2). For each country-wave, the ABS research team codes each respondent's self-reported vote choice against the outcome of the most recent national election prior to fieldwork, classifying respondents as winners (voted for the governing party, president, or coalition partner) or losers (voted for any non-governing party). The analysis adopts this binary coding throughout. Respondents who did not vote, refused to answer, or reported an invalid vote choice are excluded, restricting the sample to citizens whose electoral participation permits meaningful assignment of winner-loser status. Thailand Wave 4 (2014--2016) presents a distinctive case: fieldwork occurred after the May 2014 military coup, which followed the annulment of the February 2014 election. The ABS nevertheless coded winner/loser status for 382 respondents based on prior electoral behavior; the analysis retains this country-wave, and a robustness check verifies (Appendix Table K1) that all core findings hold when it is excluded.[^2]

[^2]: The relatively high rate of missing data on q34a in Thailand Wave 4 (51.1%) likely reflects respondent reluctance to disclose vote choice under military rule rather than a failure of the survey instrument. The 382 respondents with valid coding yield winner and loser subsamples of 186 and 227, respectively—sufficient for stable estimation. Appendix Section B documents the reference election, winning party, and runner-up for each country-wave; reports an independent verification of the ABS coding against official election commission data for Taiwan, South Korea, and Thailand; and discusses additional edge cases including Thailand's government-formation coding and Cambodia's opposition dissolution. Control variables include age, gender, education level, and urban-rural residence.

## Analytical Strategy

Rather than collapsing responses into a binary procedural–substantive indicator, the analysis models each item set as a separate multinomial logistic regression, preserving all four response categories within each set. For Waves 3, 4, and 6, which share a common instrument of four item sets (Sets 1–4), data are pooled across waves and modeled jointly. Each set takes the form:

$$
\Pr(Y_{ij} = k) = \frac{\exp(\beta_{0k} + \beta_{1k}\text{Loser}_i + \mathbf{X}_i\boldsymbol{\gamma}_k + \alpha_{ck} + \delta_{tk})}{\sum_{m=1}^{4}\exp(\beta_{0m} + \beta_{1m}\text{Loser}_i + \mathbf{X}_i\boldsymbol{\gamma}_m + \alpha_{cm} + \delta_{tm})}
$$

where $Y_{ij}$ denotes the item selected by respondent $i$ in set $j$, $k$ indexes the four response categories within each set, $\mathbf{X}_i$ is a vector of demographic controls (age, gender, education, and urban–rural residence), and $\alpha_{ck}$ and $\delta_{tk}$ represent category-specific country and wave fixed effects. The reference category is set internally by the estimation routine; because results are reported as average marginal effects (AMEs), the choice of baseline category does not affect interpretation. AMEs express the effect of loser status as percentage-point changes in the predicted probability of selecting each item, offering a direct and intuitive quantity of interest across all twenty items.

Wave 2, which employed a single forced-choice item with a distinct set of response options, is modeled separately using the same multinomial logit specification (without wave fixed effects, as it draws from a single survey round). Including Wave 2 provides an independent check on the robustness of the loser effect across different measurement instruments.

Standard errors are computed via wild cluster bootstrap at the country level, following @Cameron2008-aw. With only eleven country clusters in the pooled sample, conventional cluster-robust variance estimators are known to produce downward-biased standard errors and over-reject the null hypothesis. The wild bootstrap imposes the null hypothesis on the bootstrap data-generating process, providing more reliable inference with few clusters. This approach yields conservative tests throughout the analysis.

To examine country-level variation and within-country trajectories (H2 and H3), the multinomial logit models are re-estimated separately for each country-wave combination. For each estimate, I compute the procedural–substantive gap: the difference between the mean AME of loser status across procedural items and the mean AME across substantive items within a given country-wave. A positive gap indicates that losers disproportionately favor procedural conceptions relative to substantive ones. Confidence intervals for these gaps are obtained via the same wild bootstrap procedure. Tracking these gaps across waves within countries allows direct assessment of whether conceptions of democracy respond to changing political conditions—the central prediction distinguishing the positional account from cultural or dispositional alternatives.

# Results

## The Baseline Loser Effect

```{r pooled-summary}
#| include: false
# Summary stats for inline text
proc_items <- pooled %>% filter(item_type == "procedural")
sub_items <- pooled %>% filter(item_type == "substantive")
gov_items <- pooled %>% filter(item_type == "governance")

n_proc_pos <- sum(proc_items$ame > 0)
n_proc_total <- nrow(proc_items)
n_proc_sig <- sum(proc_items$p < 0.05)

n_sub_neg <- sum(sub_items$ame < 0)
n_sub_total <- nrow(sub_items)
n_sub_sig <- sum(sub_items$p < 0.05)

proc_range_lo <- fmt_pct(min(proc_items$ame[proc_items$ame > 0]))
proc_range_hi <- fmt_pct(max(proc_items$ame))

sub_range_lo <- fmt_pct(min(sub_items$ame))
sub_range_hi <- fmt_pct(max(sub_items$ame[sub_items$ame < 0]))

# Largest effects
largest_proc <- proc_items %>% filter(ame == max(ame))
largest_sub <- sub_items %>% filter(ame == min(ame))
largest_gov <- gov_items %>% filter(ame == min(ame))

# Free elections item
free_elections <- proc_items %>% filter(item_label == "Free elections")
```

Table 1 reports the average marginal effect of loser status on the probability of selecting each of the twenty items across the pooled sample. The pattern is consistent with H1: `r n_proc_pos` of `r n_proc_total` procedural items show positive effects (losers more likely to select), while all `r n_sub_total` substantive items show negative effects (winners more likely to select). Procedural effects range from `r proc_range_lo` to `r proc_range_hi` percentage points, with `r largest_proc$item_label` showing the largest effect (`r fmt_pp(largest_proc$ame)` pp, $p$ `r ifelse(largest_proc$p < 0.001, "< 0.001", sprintf("= %.3f", largest_proc$p))`). Substantive effects range from `r sub_range_lo` to `r sub_range_hi` pp, with `r largest_sub$item_label` as the largest (`r fmt_pp(largest_sub$ame)` pp, $p$ `r ifelse(largest_sub$p < 0.001, "< 0.001", sprintf("= %.3f", largest_sub$p))`).

```{r pooled-table}
#| output: asis

# Build the main results table
tbl <- pooled %>%
  arrange(desc(ame)) %>%
  mutate(
    Type = str_to_title(item_type),
    Item = item_label,
    `AME (pp)` = sprintf("%+.1f", ame * 100),
    SE = sprintf("%.1f", se * 100),
    Sig = sig
  ) %>%
  select(Item, Type, `AME (pp)`, SE, Sig)

if (knitr::is_latex_output()) {
  tbl_out <- tbl %>%
    kbl(
      booktabs = TRUE,
      longtable = FALSE,
      linesep = "",
      caption = "Average Marginal Effect of Loser Status on Item Selection"
    ) %>%
    kable_styling(
      latex_options = c("striped", "hold_position"),
      font_size = 10,
      full_width = TRUE
    ) %>%
    footnote(
      general = "Average marginal effects from multinomial logit with country and wave fixed effects and demographic controls (age, gender, education, urban residence). Positive values indicate losers are more likely to select the item. Bootstrap SEs clustered at the country level.",
      symbol = c("$\\\\dagger$ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001"),
      general_title = "Note: ",
      footnote_as_chunk = TRUE,
      escape = FALSE
    )
  # Make footnote rows wrap and single-space
  tbl_str <- as.character(tbl_out)
  tbl_str <- gsub("\\multicolumn{5}{l}{", "\\multicolumn{5}{p{\\linewidth}}{\\singlespacing ", tbl_str, fixed = TRUE)
  cat(tbl_str)
} else {
  tbl %>%
    kable(format = "pipe", align = c("l", "l", "l", "r", "r", "c"))
}
```

The individual effects are modest—typically two to four percentage points—but the consistency of the pattern across twenty items from five separate batteries is notable. This is not a result driven by one or two items. Winners and losers differ systematically in how they conceptualize democracy, with losers oriented toward the rules governing political competition and winners oriented toward the material benefits government provides.

One item warrants specific attention. Free elections—the most canonical procedural item—shows a *negative* AME (`r fmt_pp(free_elections$ame)` pp), meaning losers are slightly *less* likely than winners to identify free elections as the most important feature of democracy. This runs counter to H1's general prediction, but the pattern becomes interpretable in light of the remaining procedural results. The items that losers *do* favor are the liberal components of democratic governance: free expression (`r fmt_pp(proc_items %>% filter(item_label == "Free expression") %>% pull(ame))` pp), media freedom (`r fmt_pp(proc_items %>% filter(item_label == "Media freedom") %>% pull(ame))` pp), the right to organize groups (`r fmt_pp(proc_items %>% filter(item_label == "Organize groups") %>% pull(ame))` pp), and party competition (`r fmt_pp(proc_items %>% filter(item_label == "Party competition") %>% pull(ame))` pp). Losers who have experienced electoral manipulation or whose preferred parties have been dissolved by courts may have reason to distrust the electoral mechanism itself while valuing the broader liberal ecosystem—media, expression, association, judicial accountability—that enables political contestation outside formal elections. The loser effect, in other words, is not an elections effect but a *liberal-democratic* effect: losers gravitate toward the institutional protections that sustain opposition rather than the specific electoral procedure that has failed to deliver victory.

This pattern is consistent with what might be termed *cynical proceduralism*: losers do not naively embrace all democratic procedures but discriminate between constraint-based safeguards (courts, media, expression) whose protective value is independent of electoral outcomes and selection-based mechanisms (elections) whose credibility is contingent on perceived integrity.

The governance items reveal an additional pattern. Winners are consistently more likely to select governance items emphasizing order and state capacity, with `r largest_gov$item_label` showing the largest effect (`r fmt_pp(largest_gov$ame)` pp, $p$ `r ifelse(largest_gov$p < 0.001, "< 0.001", sprintf("= %.3f", largest_gov$p))`). Winners conceive of democracy as a framework for stability and effective administration; losers conceive of it as a framework for contestation and the protection of rights. This order-versus-contestation tension is explored further in the discussion.

## Democratic Erosion and Institutional Resilience

```{r trajectory-stats}
#| include: false
# Key trajectory stats
# Thailand
thai_w3_pp <- round(thai_w3_gap * 100, 1)
thai_w4_pp <- round(thai_w4_gap * 100, 1)
thai_w6_pp <- round(thai_w6_gap * 100, 1)
thai_w6_ci <- thai %>% filter(wave == 6)

# South Korea
kor <- cw_gap %>% filter(country_name == "South Korea") %>% arrange(wave)
kor_w3_pp <- round(kor %>% filter(wave == 3) %>% pull(proc_sub_gap) * 100, 1)
kor_w4_pp <- round(kor %>% filter(wave == 4) %>% pull(proc_sub_gap) * 100, 1)
kor_w6_pp <- round(kor %>% filter(wave == 6) %>% pull(proc_sub_gap) * 100, 1)

# Japan
jpn <- cw_gap %>% filter(country_name == "Japan") %>% arrange(wave)

# Cambodia
khm <- cw_gap %>% filter(country_name == "Cambodia") %>% arrange(wave)
```

The pooled estimates establish that losers systematically favor procedural conceptions, but the theory predicts more than a cross-sectional association: it predicts that the gap should respond to changing political conditions within countries over time (H2 and H3). Figure 1 plots the procedural-substantive gap for each country across survey waves, where a positive value indicates that losers disproportionately favor procedural conceptions.

```{r fig-trajectories, fig.width=10, fig.height=6}
#| fig-cap: "Loser Effect Trajectories: Procedural--Substantive Gap by Country and Wave"
#| fig-alt: "Line plot showing procedural-substantive gap trajectories across three survey waves for eleven Asian countries. Thailand shows dramatic increase from near zero to over 20 percentage points. South Korea remains near zero throughout. Cambodia maintains a high gap around 13 points. Other countries shown in gray."

# Plot countries with 2+ waves
plot_data <- cw_gap %>%
  group_by(country_name) %>%
  filter(n() >= 2) %>%
  ungroup()

highlight <- c("Thailand", "South Korea", "Japan", "Cambodia")

plot_data %>%
  mutate(
    highlighted = country_name %in% highlight,
    country_label = if_else(highlighted, country_name, "Other")
  ) %>%
  ggplot(aes(x = wave, y = proc_sub_gap * 100, group = country_name)) +
  geom_line(
    data = . %>% filter(!highlighted),
    color = "gray75", linewidth = 0.5, alpha = 0.7
  ) +
  geom_point(
    data = . %>% filter(!highlighted),
    color = "gray75", size = 1.5, alpha = 0.7
  ) +
  geom_ribbon(
    data = . %>% filter(highlighted),
    aes(ymin = proc_sub_gap_ci_low * 100, ymax = proc_sub_gap_ci_high * 100,
        fill = country_name),
    alpha = 0.15, color = NA
  ) +
  geom_line(
    data = . %>% filter(highlighted),
    aes(color = country_name), linewidth = 1.2
  ) +
  geom_point(
    data = . %>% filter(highlighted),
    aes(color = country_name), size = 3
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous(
    breaks = c(3, 4, 6),
    labels = c("W3\n(2010-12)", "W4\n(2014-16)", "W6\n(2019-22)")
  ) +
  scale_color_manual(values = c(
    "Thailand" = "#D62728", "Cambodia" = "#FF7F0E",
    "South Korea" = "#2CA02C", "Japan" = "#1F77B4"
  )) +
  scale_fill_manual(values = c(
    "Thailand" = "#D62728", "Cambodia" = "#FF7F0E",
    "South Korea" = "#2CA02C", "Japan" = "#1F77B4"
  )) +
  labs(
    x = "Survey Wave",
    y = "Procedural - Substantive Gap\n(percentage points)",
    color = NULL, fill = NULL,
    caption = "Note: Gray lines show remaining countries. Shaded bands show 95% CIs for highlighted countries.\nPositive values = losers favor procedural items relative to substantive items."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

Thailand provides the most dramatic support for H3. In Wave 3 (2010--2012), during the Democrat Party government of Abhisit Vejjajiva, the procedural--substantive gap was negligible (`r thai_w3_pp` pp, not significant). By Wave 4 (2014--2016), conducted after the May 2014 coup that removed Prime Minister Yingluck Shinawatra ($N = 382$ respondents with valid winner/loser coding; see Methods for discussion of post-coup coding), the gap had widened to `r thai_w4_pp` pp ($p < 0.05$). The positive gap is driven by Democrat and minor-party voters---the electoral losers under ABS coding---who increasingly emphasized procedural conceptions, while Pheu Thai supporters---electoral winners despite the coup that removed their government---shifted toward substantive definitions. By Wave 6 (2019--2022), after six years of military or military-backed governance, the gap reached `r thai_w6_pp` pp ($p < 0.001$, 95% CI: \[`r round(thai_w6_ci$proc_sub_gap_ci_low * 100, 1)`, `r round(thai_w6_ci$proc_sub_gap_ci_high * 100, 1)`\]). This trajectory—from zero to over twenty percentage points in a decade—tracks Thailand's democratic erosion with remarkable precision [@Kuhonta2014-qk; @Kongkirati2020-fl]. As coups and judicial interventions repeatedly overturned electoral outcomes, supporters of the excluded Thaksin-aligned faction increasingly defined democracy in procedural terms, while supporters of the military-backed order emphasized substantive outcomes.

South Korea presents the mirror image. Despite experiencing its own political upheaval during this period—the impeachment and removal of President Park Geun-hye in 2016--2017 and the subsequent election of progressive Moon Jae-in—the procedural--substantive gap remained modest throughout, fluctuating between `r kor_w3_pp` pp in Wave 3 and `r kor_w4_pp` pp in Wave 4, never exceeding five percentage points. The critical difference is institutional: South Korea's crisis was resolved through constitutional procedures rather than military intervention. The impeachment demonstrated that accountability mechanisms functioned, and the 2017 election delivered genuine alternation. Under these conditions, the positional logic predicts convergence, and convergence is what the data show.

Cambodia shows a persistently large gap (approximately 13 pp across all three waves) that, notably, did not grow following the dissolution of the main opposition party (CNRP) in 2017. In a competitive authoritarian context, the stakes of losing were already effectively maximized—losers' procedural orientation had little room to intensify further. Among the remaining countries (@tbl-country-gaps), Malaysia shows a large gap that narrows across waves, potentially reflecting the political opening that culminated in the 2018 alternation. Myanmar's single observation captures a strikingly large gap consistent with the high stakes of its fragile democratic transition. Taiwan's trajectory is distinctive, reversing from positive to negative by Wave 6—a pattern that may reflect the unusual dynamics of cross-strait identity politics. Japan provides the expected baseline for a consolidated democracy: modest and stable across waves.

```{r tbl-country-gaps}
#| tbl-cap: "Procedural--Substantive Gap by Country and Wave"

gap_tbl <- cw_gap %>%
  mutate(
    gap_label = sprintf("%+.1f [%+.1f, %+.1f]",
                        proc_sub_gap * 100,
                        proc_sub_gap_ci_low * 100,
                        proc_sub_gap_ci_high * 100)
  ) %>%
  select(country_name, wave_label, gap_label) %>%
  pivot_wider(names_from = wave_label, values_from = gap_label, values_fill = "—") %>%
  rename(Country = country_name) %>%
  select(Country, W3, W4, W6)

if (knitr::is_latex_output()) {
  gap_tbl %>%
    kbl(
      booktabs = TRUE,
      longtable = FALSE,
      linesep = ""
    ) %>%
    kable_styling(
      latex_options = c("striped", "hold_position"),
      font_size = 10
    ) %>%
    column_spec(1, width = "3cm") %>%
    column_spec(2, width = "3.5cm") %>%
    column_spec(3, width = "3.5cm") %>%
    column_spec(4, width = "3.5cm") %>%
    footnote(
      general = "Gap = (Mean Procedural AME) - (Mean Substantive AME), in percentage points, with 95% bootstrap CIs. Positive values indicate losers favor procedural items relative to substantive items. Dashes indicate country not surveyed in that wave.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE,
      threeparttable = TRUE
    )
} else {
  gap_tbl %>%
    kable(format = "pipe", align = c("l", "c", "c", "c"))
}
```

## Testing the Positional Mechanism

The consistent pattern of loser-winner differences across twenty items, combined with the divergent country trajectories, raises the question of mechanism. Do these associations reflect positional updating, in which citizens revise their democratic priorities in response to the political stakes of exclusion, or stable ideological sorting, whereby different types of people sort into winner versus loser categories?

Three complementary tests address this question, each designed to rule out a competing explanation.

### Test 1: Fairness Amplification

If conceptions reflect positional updating, the loser effect should amplify when procedural threat is salient—i.e., among losers who perceive elections as unfair. If instead it reflects stable normative commitments, it should appear regardless of fairness perceptions.

However, not all procedural conceptions carry the same relationship to electoral status, and specifying this structure *ex ante* disciplines the interpretation of the interaction results. A distinction can be drawn between two classes of procedural items. The first encompasses *protective* procedural elements: institutional safeguards that constrain incumbent power and protect political minorities regardless of who governs, including an independent judiciary, media freedom, and free expression. For electoral losers, these protections are unambiguously valuable—they represent the institutional infrastructure that prevents winners from converting temporary electoral mandates into permanent domination. The positional updating framework therefore predicts that the loser effect should be consistently positive for protective procedural items, and this expectation should hold across varying levels of perceived electoral fairness.

The second class consists of *participatory* procedural elements: the mechanisms through which political competition itself is organized, most notably free and fair elections and multiparty contestation. The relationship between electoral status and the valuation of participatory procedures is theoretically more complex. Under conditions of well-functioning democracy, losers may value electoral mechanisms highly, recognizing them as the vehicle through which power can eventually be transferred. However, under conditions of democratic erosion—particularly when erosion targets electoral integrity itself—losers confront a distinctive dilemma. The very mechanism that is supposed to enable alternation in power has produced an outcome they regard as illegitimate, compromised by manipulation, unfair media coverage, or institutional bias. In such contexts, losers may devalue elections specifically while intensifying their commitment to non-electoral safeguards.

This distinction generates a structured set of conditional expectations: (a) a uniformly positive loser effect for protective procedural items across fairness conditions; (b) a positive loser effect for participatory procedural items under conditions of perceived electoral integrity; and (c) an attenuated or negative loser effect for participatory procedural items when losers perceive the electoral process itself as compromised.

|   | Perceived Electoral Integrity | Perceived Electoral Manipulation |
|------------------------|------------------------|------------------------|
| **Protective procedures** (courts, media, expression) | Loser effect positive | Loser effect positive (strengthened) |
| **Participatory procedures** (free elections, multiparty competition) | Loser effect positive | Loser effect attenuated or negative |

```{r fairness-data}
#| include: false
# Load fairness interaction results
fairness <- read_csv(file.path(results_path, "robustness_fairness_interaction.csv"), show_col_types = FALSE)

# Key interaction effects
court_int <- fairness %>% filter(item == "Court protection") %>% pull(interaction)
media_int <- fairness %>% filter(item == "Media freedom") %>% pull(interaction)
expression_int <- fairness %>% filter(item == "Free expression") %>% pull(interaction)
elections_int <- fairness %>% filter(item == "Free elections") %>% pull(interaction)

# Count positive interactions among procedural items
proc_fairness <- fairness %>% filter(item_type == "procedural")
n_proc_pos_int <- sum(proc_fairness$interaction > 0)
n_proc_int_total <- nrow(proc_fairness)
```

The fairness interaction results align closely with the conditional expectations derived from the protective-participatory distinction. Among `r n_proc_int_total` procedural items, `r n_proc_pos_int` show a positive interaction—meaning the loser effect is larger among those perceiving elections as unfair. The amplification is especially pronounced for protective procedural items: court protection shows an interaction of `r sprintf("%+.1f", court_int * 100)` percentage points, media freedom `r sprintf("%+.1f", media_int * 100)` pp, and free expression `r sprintf("%+.1f", expression_int * 100)` pp. Losers who perceive the electoral process as compromised do not simply withdraw from procedural commitments; they double down on precisely the non-electoral safeguards that would constrain those who benefit from unfair elections.

The pattern diverges for participatory procedural items as predicted. The interaction between loser status and perceived fairness for the free elections item runs negative (`r sprintf("%+.1f", elections_int * 100)` pp), indicating that losers who perceive the electoral process as compromised are *less* likely to prioritize elections as a defining feature of democracy. Rather than constituting a theoretical anomaly, this finding reflects a coherent response: citizens who believe the primary mechanism of democratic alternation has been corrupted shift their definitional emphasis toward institutional safeguards that operate independently of the electoral process itself. The negative coefficient does not undermine the positional updating framework; it reveals its internal structure. Losers do not uniformly inflate the value of all procedural elements—they attend most to those procedures that offer genuine protection given the specific institutional threats they perceive.

This structured pattern also helps rule out an alternative interpretation—that the loser effect reflects nothing more than generalized negativity or sour-grapes reasoning. If losers were simply devaluing everything associated with the political system, the effect should be uniformly negative across both protective and participatory items when perceived fairness is low. The observed divergence—positive for protective items, negative for participatory items under conditions of perceived manipulation—is more consistent with a discriminating assessment of which institutional features remain credible safeguards and which have been compromised.

```{r fig-fairness}
#| fig-cap: "Fairness Interaction: Loser Effect by Perceived Electoral Fairness"
#| fig-alt: "Dot plot showing loser effects separately for respondents perceiving fair vs. unfair elections, across sixteen items. Liberal procedural items show substantially larger loser effects among those perceiving unfair elections."
#| fig-width: 8
#| fig-height: 6

# Plot interaction effects
fairness %>%
  mutate(
    item = fct_reorder(item, interaction),
    item_type = str_to_title(item_type)
  ) %>%
  pivot_longer(
    cols = c(loser_effect_fair, loser_effect_unfair),
    names_to = "condition",
    values_to = "loser_effect"
  ) %>%
  mutate(
    condition = if_else(condition == "loser_effect_fair", "Fair election", "Unfair election")
  ) %>%
  ggplot(aes(x = loser_effect * 100, y = item, color = condition, shape = condition)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  facet_wrap(~ item_type, scales = "free_y", ncol = 1) +
  scale_color_manual(values = c("Fair election" = "#1F77B4", "Unfair election" = "#D62728")) +
  labs(
    x = "Loser Effect (percentage points)",
    y = NULL,
    color = NULL, shape = NULL,
    caption = "Note: Loser AMEs from multinomial logit interacting loser status with perceived electoral fairness.\nCountry and wave fixed effects with demographic controls."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

An important caveat qualifies these findings. Perceived unfairness is endogenous to loser status: losers are approximately `r sprintf("%.1f", endog$ratio)` times more likely to perceive elections as unfair (`r sprintf("%.1f", endog$pct_unfair_losers)`% vs. `r sprintf("%.1f", endog$pct_unfair_winners)`% of winners), though the point-biserial correlation is modest ($r$ = `r sprintf("%.3f", endog$r_pb)`, $N$ = `r format(endog$n_complete, big.mark = ",")`), indicating substantial overlap rather than collinearity. It remains unclear whether the interaction reflects genuine updating or pre-existing differences in outlook. The protective-participatory divergence mitigates but does not eliminate this concern, since motivated reasoning could produce domain-specific patterns. However, a pure "sore loser" interpretation is difficult to reconcile with the item-level structure of the results: if losers who perceive unfairness were simply expressing generalized grievance, the interaction should run uniformly in the same direction across all procedural items rather than diverging sharply between protective items (positive) and participatory items (negative). Because this test alone cannot definitively distinguish the mechanisms, two additional tests are needed.

### Test 2: Domain Specificity (Placebo)

If the loser-unfairness interaction simply captures generalized discontent ("losers who think elections are unfair are negative about everything"), it should extend to non-procedural domains like material welfare. If it reflects procedural-specific updating, the amplification should be confined to procedural safeguards.

I test whether the interaction extends to perceptions of basic necessities (q109: "People have food, clothes, and shelter"). While losers are slightly more likely than winners to agree people have basic necessities ($\beta$ = `r fmt_coef(placebo$main_loser_coef)`, SE = `r fmt_coef(placebo$main_loser_se)`, p \< 0.001)—a level effect potentially reflecting socioeconomic composition—the Loser × Unfairness interaction is not significant ($\beta$ = `r fmt_coef(placebo$interaction_coef)`, SE = `r fmt_coef(placebo$interaction_se)`, p = `r fmt_p(placebo$interaction_p)`).

Among those perceiving fair elections, the loser effect is `r sprintf("%.1f", placebo$ame_fair)` percentage points; among those perceiving unfair elections, it is `r sprintf("%.1f", placebo$ame_unfair)` points—a difference of only `r sprintf("%.1f", placebo$ame_diff)`pp (95% CI: \[`r sprintf("%.1f", placebo$ci_low)`, `r sprintf("%.1f", placebo$ci_high)`\]pp). This near-zero interaction contrasts sharply with procedural items, where the amplification averages 5–8 percentage points for liberal safeguards.

The mechanism is domain-specific, not a general pattern of negativity. Material welfare taps output legitimacy (whether government delivers); procedural items tap input legitimacy (whether processes are fair). The fact that unfairness perceptions amplify the loser effect for the latter but not the former is consistent with positional updating focused on procedural threat rather than motivated reasoning or diffuse discontent.

### Test 3: Not Compositional (Thailand Reweighting)

The Thailand trajectory—widening from near-zero to over 20 percentage points—could reflect compositional change: if opposition coalitions realigned to include more urban, educated voters who intrinsically favor procedural democracy, the gap would widen without individual-level updating.

I reweight each Thailand wave to match the Wave 3 demographic distribution (age × gender × education × urban residence), creating a counterfactual trajectory under constant demographics. Reweighting changes loser effects by an average of `r sprintf("%.2f", composition$mean_abs_change)` percentage points (maximum: `r sprintf("%.2f", composition$max_abs_change)`pp for `r composition$max_item`), with only `r composition$n_sign_changes` of `r composition$n_total_items` items changing sign—and that item (`r composition$sign_change_detail`) flips between values both near zero. All substantively meaningful effects retain their direction and magnitude.

The trajectory survives demographic reweighting, indicating it reflects changing attitudes conditional on demographics rather than changing demographics per se. The analysis cannot rule out realignment on unobserved dimensions, including ideology, political interest, or unmeasured socioeconomic characteristics. However, age, education, and urbanization are the primary demographic cleavages in Thai politics [@Ockey2004-ek; @McCargo2019-rx]. The fact that controlling for these produces negligible change suggests compositional shifts are not the driver of Thailand's widening gap.

### Cumulative Interpretation

No single test is definitive. Taken together, however, the three checks converge on the positional interpretation:

1.  Amplification (Test 1): The loser effect is larger among those perceiving unfairness (5–8pp for liberal items vs. 2–3pp for those perceiving fairness), showing sensitivity to perceived threat rather than stable commitments.

2.  Specificity (Test 2): The amplification does not extend to material welfare. The loser-unfairness interaction for basic necessities is `r sprintf("%.1f", placebo$ame_diff)`pp (p = `r fmt_p(placebo$interaction_p)`), contrasting with 5–8pp for procedural items. The mechanism is procedurally bounded, not generalized discontent.

3.  Stability (Test 3): The Thailand trajectory survives demographic reweighting. Holding age, education, gender, and urban residence constant changes loser effects by an average of `r sprintf("%.1f", composition$mean_abs_change)`pp, with one near-zero sign flip. Compositional change does not explain the widening gap.

While these patterns do not constitute proof of individual-level attitude change—repeated cross-sections cannot observe within-person dynamics—they are difficult to reconcile with accounts treating democratic conceptions as stable cultural inheritances, fixed psychological traits, or artifacts of coalition realignment. The alternative explanation would require that ideological or dispositional differences between winner and loser coalitions widened dramatically in Thailand while remaining stable in South Korea, Cambodia, and most other countries—a pattern for which no independent evidence exists. The associations documented here point toward positional updating: citizens revising their understanding of democracy in response to their political position and the institutional context in which they find themselves.

# Robustness

Several additional tests assess the sensitivity of the main findings. First, a weighted least squares (WLS) estimation that aggregates the data to the country-wave level sidesteps the clustering problem entirely: nine of ten procedural items show positive loser effects (six significant), all six substantive items show negative effects (five significant), with mean effect sizes comparable to the individual-level estimates (+2.2 pp procedural, -2.4 pp substantive; Appendix G).

Second, Wave 2 of the ABS, which employed a different forced-choice instrument, provides an independent replication. Both procedural items show positive AMEs and both substantive items show negative AMEs, reinforcing confidence that the loser effect is not an artifact of the Wave 3--6 battery design. Open-ended responses from Wave 2 confirm convergent validity: respondents selecting procedural items were significantly more likely to offer procedural definitions spontaneously ($r = 0.28$, $p < .001$; Appendix F).[^3]

[^3]: Following established conceptual frameworks, open-ended responses were coded as procedural, substantive, or excluded. A respondent-level procedural proportion score (0--1) based on up to three codeable responses correlates at $r = 0.28$ with the closed-ended item. Given the noise inherent in open-ended coding, this represents meaningful convergent validity.

Third, Wave 6 fieldwork coincided with the COVID-19 pandemic, which may have compressed the procedural-substantive gap by increasing the salience of state capacity for all citizens. Re-estimating Wave 6 models with controls for pandemic attitudes yields essentially unchanged results (mean procedural AME: `r fmt_pp(covid_proc_base)` → `r fmt_pp(covid_proc_covid)` pp; mean substantive AME: `r fmt_pp(covid_sub_base)` → `r fmt_pp(covid_sub_covid)` pp). Thailand's gap persists through the pandemic, consistent with democratic erosion overwhelming the homogenizing pull of COVID-19.

Fourth, non-voters' procedural-substantive gap (14.3 pp) resembles that of winners (13.9 pp) rather than losers (17.3 pp), confirming that restricting the sample to voters does not inflate the estimated loser effect (Appendix H). Fifth, a three-way decomposition distinguishing key opposition supporters from other losers finds that both subgroups display comparable procedural orientations across the three countries where decomposition was feasible (Appendix J). This pattern is more consistent with an exclusion-based mechanism—where distance from state authority, rather than partisan strategic signaling, drives procedural attention—than with an elite-cue account that would predict concentration among principal opposition supporters.

Sixth, because Thailand Wave 4 was fielded under military rule with elevated missing data on vote choice (51.1%), re-estimating all pooled models with Thailand W4 excluded confirms that no coefficient changes by more than 0.28 percentage points (Appendix Table K1).

Finally, the use of repeated cross-sections rather than panel data means within-country trajectories cannot definitively rule out compositional change. The Thai case offers the strongest counterargument: party loyalties in Thailand's color-coded politics are notably stable, and the gap's tracking of documented political events—rather than gradual demographic shifts—favors the positional interpretation.

# Discussion

The central finding of this article is not simply that electoral losers prefer procedural democracy, though they do, consistently, across twenty items and eleven countries. The more consequential finding is that this association varies systematically with political context. Thailand's trajectory from near-zero to over twenty percentage points across a single decade of democratic erosion is difficult to reconcile with accounts treating democratic conceptions as stable cultural inheritances or fixed psychological dispositions [@Norris2011-zt; @Inglehart1997-gw]. Nor is it plausibly compositional: Thailand's color-coded partisan alignments remained remarkably stable even as the parties themselves were repeatedly dissolved and reconstituted [@Sinpeng2021-hd; @McCargo2019-rx]. Three pieces of evidence tilt toward the positional interpretation: (1) the fairness interaction shows amplification where procedural threat is most salient, (2) the placebo test shows the mechanism is domain-specific, not generalized discontent, and (3) demographic reweighting confirms the Thailand trajectory is not compositionally driven.

While repeated cross-sections cannot definitively establish individual-level attitude change, the within-country dynamics documented here, particularly the Thailand-South Korea contrast, point toward positional updating rather than stable sorting on unobserved characteristics. The alternative explanation would require that ideological or dispositional differences between winner and loser coalitions widened dramatically in Thailand while remaining stable in South Korea, Cambodia, and most other countries—a pattern for which no independent evidence exists.

This dynamic carries implications for democratic resilience, though the direction depends on whether the associations reflect updating or sorting. A growing literature asks whether mass public support for democracy can serve as a bulwark against backsliding [@Claassen2020-hk; @Svolik2019-zi; @Graham2020-jr]. If the patterns documented here reflect positional updating, the results suggest a structural problem with this expectation: where erosion is most severe, procedural commitments become concentrated among those whom the political system has rendered least capable of defending them. Aggregate measures of democratic support may therefore overstate resilience by averaging across winners whose procedural commitments are shallow and losers whose commitments are deep but politically ineffectual. The very mechanisms through which democracies erode, including executive aggrandizement, judicial capture, and media suppression [@Levitsky2018-dd; @Waldner2018-yw; @Luhrmann2019-ow], are precisely the mechanisms that disarm the constituency most committed to procedural norms.

If, alternatively, the patterns reflect stable sorting, the implications are less dire but still consequential: it would mean democratic conceptions are demographically or ideologically clustered in ways that make procedural support vulnerable to electoral outcomes. Whether losers' procedural orientation can translate into effective resistance, through protest, civil disobedience, or transnational advocacy, remains an open and urgent question.

The results also complicate the winner-loser gap literature in a way that has implications for survey research. Previous work has assumed that winners and losers share a common understanding of democracy when they report their satisfaction with it or their trust in its institutions [@Anderson2005-uk; @Esaiasson2011-po; @Norris2012-wj]. If winners and losers define democracy differently—winners emphasizing outcomes, losers emphasizing procedures—then the standard "satisfaction with democracy" question is not measuring the same thing for both groups. Losers may report dissatisfaction because they perceive procedural violations; winners may report satisfaction because they perceive policy delivery. @Rich2025-gd offers suggestive evidence for this interpretation: using Asian Barometer data from Taiwan, he finds that the winner-loser gap appears in trust toward electoral institutions but not in overall satisfaction with democracy, consistent with winners and losers differentiating between dimensions of democratic performance rather than evaluating a single shared object. This disconnect suggests that the much-studied winner-loser gap in democratic satisfaction may partly reflect a *measurement* artifact: not different evaluations of the same object, but evaluations of different objects.

The item-level results add a further dimension. The loser effect is not a generic procedural orientation but a specifically *liberal-democratic* one. Losers gravitate toward media freedom, free expression, the right to organize, and judicial accountability—the institutional infrastructure that sustains political opposition—rather than toward elections per se. The protective-participatory distinction introduced in the fairness analysis clarifies this pattern: losers consistently favor protective procedural elements (non-electoral safeguards that constrain incumbent power) while their orientation toward participatory elements (elections and multiparty competition) is conditional on whether they perceive the electoral process as compromised. The negative fairness interaction on the free elections item reflects not an anomaly but a coherent response—citizens who believe the mechanism of democratic alternation has been corrupted shift their definitional emphasis toward safeguards that operate independently of elections. The distinction between protective and participatory conceptions of procedural democracy, latent in theoretical frameworks from @Dahl1971-lu to @Schedler1998-dl, appears to be psychologically real among citizens navigating the consequences of electoral loss.

Finally, the order-versus-contestation tension in the governance items deserves further investigation. Winners' strong preference for law and order as a defining feature of democracy, combined with losers' emphasis on civil liberties and accountability, captures a fault line that @Slater2010-ba has identified as central to the politics of order in Southeast Asia. Where democratic institutions are robust, the demands for order and for contestation coexist in productive tension. Where they are not, the two conceptions may become mutually exclusive—and in the current period of global democratic recession [@Diamond2015-yb; @Foa2016-lq], understanding which conception prevails, and among whom, has consequences well beyond the region studied here.

Several limitations temper these conclusions. The analysis relies on repeated cross-sections rather than panel data, precluding direct observation of individual-level attitude change; the within-country trajectories are suggestive but not definitive. The sample comprises only eleven countries, limiting cross-national inference. The Wave 6 compression across most countries outside Thailand remains incompletely explained, though the COVID-19 pandemic's homogenizing effect on the salience of state capacity is a plausible contributor, a prediction that post-pandemic survey waves will be positioned to test. And the exclusion of non-voters, while theoretically motivated, means the findings characterize the politically engaged electorate rather than the population at large.

# Conclusion

This article has demonstrated that electoral status is systematically associated with how citizens conceptualize democracy. Losers prioritize the liberal-procedural infrastructure—free expression, media freedom, judicial accountability, the right to organize—that sustains the possibility of political contestation. Winners prioritize substantive outcomes and the governance of order. Three pieces of evidence support a positional updating interpretation rather than stable sorting: the loser effect (1) amplifies among those perceiving electoral unfairness, (2) does not extend to non-procedural domains, and (3) persists after demographic reweighting. The dramatic within-country trajectories in Thailand and South Korea—diverging precisely as their institutional paths diverged—are difficult to reconcile with accounts treating democratic conceptions as fixed cultural inheritances.

If this positional logic holds beyond Asia—a question that comparative research in Latin America and Africa is well positioned to address—then the relationship between public attitudes and democratic resilience may be more fragile than the existing literature suggests. Democratic erosion would not merely reduce the quality of governance or diminish civil liberties; it would reshape the very meaning of democracy along the winner-loser divide, concentrating procedural commitments among those least able to act on them. Future research should examine whether losers' procedural orientation translates into distinct patterns of political behavior, particularly unconventional participation, testing whether democratic conceptions serve as blueprints for resistance or remain, ultimately, survey artifacts.

The limits of repeated cross-sectional data caution against definitive causal claims. The analysis cannot directly observe individual-level attitude change, nor can it fully rule out that the observed patterns reflect ideological realignment on unobserved dimensions. What can be said is that the associations documented here—their consistency across items, their sensitivity to political context, their concentration among those perceiving procedural threat—point toward a dynamic relationship between electoral experience and democratic meaning that the literature has not previously recognized. Whether that relationship is best understood as updating, sorting, or some combination remains an open question, but the evidence presented here tilts toward the positional account and away from cultural determinism.

# References {.unnumbered}

\singlespacing

::: {#refs}
:::