---
title: "The Meaning of Losing: How Democratic Erosion Reshapes Citizens' Conceptions of Democracy"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: today
abstract: |
  Does democratic erosion reshape not just how citizens evaluate democracy but what they understand it to mean? This article estimates multinomial logit models on four waves of Asian Barometer Survey data (2005–2022), drawing on more than 75,000 voters across eleven countries to examine whether electoral winners and losers hold systematically different conceptions of democracy—and whether those differences respond to political context. The central finding is that they do: losers consistently gravitate toward liberal-procedural definitions of democracy, emphasizing free expression, media freedom, and judicial accountability, while winners orient toward substantive outcomes such as economic equality and welfare provision. The gap is individually small. What makes it consequential is context. In Thailand, the procedural-substantive gap widened from near zero to over twenty percentage points across a single decade of democratic erosion culminating in the 2014 military coup and its aftermath; in South Korea, where a political crisis of comparable severity was resolved through constitutional channels, the gap barely moved. Three empirical checks support a positional updating interpretation over stable cultural sorting: the loser effect amplifies among citizens who perceive elections as unfair, does not extend to non-procedural domains such as material welfare, and persists after demographic reweighting that holds coalition composition constant. If these patterns reflect updating rather than sorting, the implications for democratic resilience are troubling—erosion would concentrate procedural commitments among citizens whom the political system has rendered least capable of defending them.
keywords:
  - democratic conceptions
  - electoral losers
  - procedural democracy
  - Thailand
  - South Korea
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: false
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \raggedbottom
          \raggedright
          \floatplacement{figure}{H}
          \newcommand{\keywords}[1]{\vspace{0.5em}\begin{singlespace}\noindent\textbf{Keywords:} #1\end{singlespace}}
  html:
    toc: true
    toc-depth: 3
    number-sections: false
    theme: cosmo
  docx:
    reference-doc: custom-reference.docx
    number-sections: false
bibliography: references.bib
csl: chicago-author-date.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE, warning = FALSE,
  fig.align = "center", fig.pos = "H", out.width = "100%"
)

here::i_am("papers/meaning-of-democracy-revision/manuscript/md-manuscript.qmd")

# Load results from multinomial analysis pipeline
results_path <- here("papers", "meaning-of-democracy-revision", "analysis", "revised", "results")

res <- readRDS(file.path(results_path, "mlogit_results.rds"))
cw_gap <- read_csv(file.path(results_path, "country_wave_proc_sub_gap.csv"), show_col_types = FALSE)
cw_items <- read_csv(file.path(results_path, "country_wave_item_ames.csv"), show_col_types = FALSE)

# Helpers
fmt_n <- function(x) format(x, big.mark = ",")
fmt_pct <- function(x) sprintf("%.1f", x * 100)
fmt_pp <- function(x) sprintf("%+.1f", x * 100)

# Key quantities
pooled <- res$all_results
n_countries <- n_distinct(cw_gap$country_name)

# Thailand trajectory
thai <- cw_gap %>% filter(country_name == "Thailand") %>% arrange(wave)
thai_w3_gap <- thai %>% filter(wave == 3) %>% pull(proc_sub_gap)
thai_w4_gap <- thai %>% filter(wave == 4) %>% pull(proc_sub_gap)
thai_w6_gap <- thai %>% filter(wave == 6) %>% pull(proc_sub_gap)
thai_change <- (thai_w6_gap - thai_w3_gap) * 100

# Total N (approximate from item-level data)
total_n <- cw_items %>%
  filter(set == "Set1") %>%
  group_by(country_name, wave_label) %>%
  slice(1) %>%
  ungroup() %>%
  pull(n) %>%
  sum()

# COVID robustness summary
covid_sum <- readRDS(file.path(results_path, "robustness_covid_summary.rds"))
covid_proc_base <- covid_sum %>% filter(item_type == "procedural", controls == "base") %>% pull(mean_ame)
covid_proc_covid <- covid_sum %>% filter(item_type == "procedural", controls == "covid") %>% pull(mean_ame)
covid_sub_base <- covid_sum %>% filter(item_type == "substantive", controls == "base") %>% pull(mean_ame)
covid_sub_covid <- covid_sum %>% filter(item_type == "substantive", controls == "covid") %>% pull(mean_ame)

# Placebo and composition robustness stats
placebo <- readRDS(file.path(results_path, "placebo_stats.rds"))
composition <- readRDS(file.path(results_path, "composition_stats.rds"))
endog <- readRDS(file.path(results_path, "endogeneity_stats.rds"))

# Helper functions for formatting
fmt_coef <- function(x) sprintf("%.3f", x)
fmt_p <- function(p) {
  if (p < 0.001) return("< 0.001")
  sprintf("%.3f", p)
}
```

\keywords{democracy, electoral losers, procedural democracy, Asian Barometer Survey, democratic consolidation}

# Introduction

Between 2010 and 2020, Thailand experienced a constitutional court's annulment of an election, a military coup, six years of junta governance, and the return to civilian rule under a military-drafted constitution that ensured the army's continued dominance. During this same decade, the gap between how electoral winners and losers conceptualized democracy widened from near zero to over `r round(thai_w6_gap * 100)` percentage points. Supporters of the excluded Thaksin-aligned parties increasingly defined democracy in terms of free elections, civil liberties, and government accountability; supporters of the military-backed order increasingly emphasized economic welfare and public services. This is not a story about who was more satisfied with democracy or more trusting of government—the attitudes that the winner-loser literature has studied extensively [@Anderson2005-uk; @Singh2012-hj]. It is a story about what citizens came to believe democracy *means*.

South Korea offers the contrast. Over the same period, the country weathered its own crisis: the impeachment and criminal prosecution of President Park Geun-hye, mass candlelight protests, and the election of a progressive successor. Yet the gap between winners' and losers' conceptions of democracy never exceeded five percentage points. The critical difference was institutional. South Korea's crisis was resolved through constitutional procedures—impeachment, judicial review, a free election—that demonstrated the system's capacity for self-correction. Thailand's crisis was resolved by tanks. Where institutions held, winners and losers converged on what democracy meant; where institutions collapsed, democratic meaning itself fractured along the winner-loser divide.

These contrasting trajectories point to a question that two established research literatures have not addressed together. One body of scholarship documents how citizens conceptualize democracy—whether in procedural terms emphasizing elections and rights, or substantive terms emphasizing welfare and equality—and finds considerable variation across and within countries [@Dalton2007-ei; @Chu2008-fk; @Ferrin2016-se]. A separate literature demonstrates that electoral winners and losers differ systematically in their satisfaction with democracy, their institutional trust, and their willingness to accept political outcomes [@Anderson2005-uk; @Anderson2002-cv; @Esaiasson2011-po]. But while the first literature has treated democratic conceptions as relatively stable cultural orientations, and the second has assumed that winners and losers are evaluating the same thing when they express views about "democracy," neither has examined whether electoral status shapes not just how citizens *evaluate* democracy but how they *understand* it.

This article presents evidence that it does. I estimate multinomial logit models on four waves of Asian Barometer Survey data (2005--2022), drawing on more than `r fmt_n(total_n)` voters in `r n_countries` countries, and find that electoral losers consistently gravitate toward procedural elements of democracy—competitive elections, civil liberties, government accountability—at the expense of substantive outcomes like economic equality or welfare provision. In pooled cross-national estimates, these effects are individually small, on the order of two to four percentage points per item. But the pattern holds across twenty items drawn from five separate survey batteries, and that directional consistency is what makes the finding substantively important. What matters most, though, is that the effect is not constant. It tracks democratic erosion in Thailand, holds steady amid institutional resilience in South Korea, and sits persistently high in Cambodia, where the competitive authoritarian regime ensured the stakes of losing were extreme from the start.

I interpret this pattern as evidence of positional updating, and three empirical checks support the interpretation. The loser effect grows stronger among respondents who perceive elections as unfair—exactly the circumstance in which procedural threat is most palpable. A placebo test demonstrates that this amplification does not spill over into material welfare perceptions, which rules out a generalized discontent story. And a demographic reweighting exercise confirms that Thailand's widening gap is not an artifact of changing coalition composition. None of these tests can nail down causation with certainty, but taken together they sit uneasily with accounts that treat democratic conceptions as stable cultural orientations. The weight of the evidence favors a positional reading.

These findings have an uncomfortable implication for how we think about democratic resilience. Scholars have recently asked whether broad public commitment to democracy might function as a check on backsliding [@Claassen2020-hk; @Svolik2019-zi; @Graham2020-jr]. But the results here suggest something close to the opposite: in the places where erosion cuts deepest, it is the people with the least political leverage who end up caring most about procedural safeguards. The citizens who prize free elections and civil liberties above all else are the same citizens whose side has been shut out of power. If this dynamic is widespread, then aggregate measures of democratic support may paint an overly reassuring picture—averaging together winners who do not care much about procedures and losers who care deeply but cannot do anything about it.

The analysis bridges the study of democratic conceptions and the winner-loser gap by demonstrating that electoral status shapes democratic *meaning*, not merely democratic *evaluation*—and that this effect is dynamic, tracking political conditions within countries over time. The multinomial logit approach, which preserves the full structure of respondents' choices across twenty items rather than collapsing responses into binary procedural-substantive indicators, reveals an additional finding: losers' procedural orientation centers on the liberal components of democracy (media freedom, free expression, the right to organize), not elections per se, a distinction with implications for how we understand the relationship between electoral experience and democratic commitment.

# Positional Updating and Democratic Conceptions

The argument begins with the observation that citizens hold meaningfully different conceptions of democracy, varying in the weight they assign to procedural rules versus substantive outcomes. From there it draws on the winner-loser gap literature to argue that electoral status creates systematically different orientations toward these dimensions, before specifying the mechanism—positional updating—and deriving predictions that distinguish it from competing accounts.

## The procedural-substantive distinction

A foundational line of comparative research draws a distinction between procedural and substantive visions of democracy. On one side sit definitions anchored in rules and institutions—competitive elections, civil liberties, the rule of law, and the accountability mechanisms that keep officeholders in check [@Dahl1977-wj; @Schumpeter2013-em; @Schedler1998-dl]. On the other sit definitions organized around what government actually delivers: reduced inequality, social protection, and responsiveness to popular demands [@Diamond1999-wd; @Linz1996-qs]. Survey research has repeatedly shown that ordinary people hold both kinds of conceptions, though the balance tips depending on context. Publics in younger democracies lean toward substantive criteria, while citizens in consolidated democracies invoke procedural ones more readily [@Dalton2007-ei; @Bratton2001-ug]. Across Asia, the picture is especially layered: people blend procedural and substantive elements in ways that vary enormously both across and within countries [@Chu2008-fk; @Chu2010-nb; @Cho2014-tj; @Shin2018-mq]. Partisan dynamics matter too. @Hsiao2020-fz demonstrate that affective polarization in Taiwan attenuates support for democracy, indicating that political position—not just cultural background—mediates how citizens relate to democratic principles.

The stakes of this variation are concrete: how citizens define democracy shapes how they evaluate it. If democracy means free elections, competitive voting may suffice for legitimacy; if it means economic equality, procedural compliance without material progress breeds disillusionment. Yet existing scholarship has largely treated conceptions as stable attributes of individuals or societies: cultural inheritances shaped by historical experience, regime type, or modernization trajectories [@Shin2011-kc; @Inglehart1997-gw; @Canache2012-fb; @Ferrin2016-se]. What remains largely unexplored is what drives *within-country, within-period* variation in how individual citizens conceptualize democracy.

## Electoral status as a source of conceptual divergence

The winner-loser gap literature provides a candidate answer. Decades of research have established that citizens whose preferred party prevailed in the most recent election look different attitudinally from those whose party lost: winners report higher institutional trust, greater satisfaction with democracy, and more willingness to accept political outcomes [@Anderson2005-uk; @Anderson2002-cv; @Nadeau1993-vo]. How large these gaps are depends on institutional design—losers in majoritarian systems tend to be more discontented than losers in consensus democracies [@Anderson1997-dh; @Singh2012-hj]—and perceptions of procedural fairness play a mediating role [@Esaiasson2011-po; @Esaiasson2020-av]. More recent work has extended these findings to non-Western contexts and hybrid regimes [@Blais2025-qq; @Mauk2020-bu]. In a particularly relevant study, @Ricks2025-il show that Thailand's 2023 election—held under a hybrid regime with military-drafted rules—nonetheless generated the classic winner-loser divergence in attitudes toward state institutions, evidence that the mechanism is not confined to full democracies.

The limitation is one of scope: this literature has focused on *evaluations*—how much citizens support, trust, or feel satisfied with democracy—while treating the underlying object of evaluation as shared. Winners and losers are assumed to mean the same thing when they express views about "democracy"; they simply differ in how positively they assess it. If electoral status shapes not only evaluations but *conceptions*—if winners and losers are in fact evaluating different things—then the winner-loser gap may be deeper than the existing literature has recognized, and the two literatures' isolation from one another represents a missed theoretical opportunity.

The question of whether electoral status shapes conceptions, not just evaluations, has begun to attract attention, though the existing work stops short of the dynamic, erosion-conditional account developed here. @Cohen2023-sm invert the classic "losers' consent" framework by tracking Brazilian voters through Bolsonaro's 2018 victory; they find that winners' abstract support for democracy rose even as their tolerance for executive self-coups—closing Congress, sidelining the Supreme Court—increased in tandem. The result is a portrait of superficial democratic commitment: winners embrace the system's label while hollowing out its procedural substance. Cohen et al. do not, however, examine whether winners and losers come to define democracy differently; their dependent variables remain evaluations of democratic institutions rather than conceptions of what those institutions should look like. @Wu2022-hz take up the conceptual question directly, using the same Asian Barometer data employed here to show that the winner-loser gap in liberal democratic values is substantially larger in competitive authoritarian regimes than in democracies—a finding consistent with H2 above. Yet their framework treats regime type as a static cross-sectional moderator; the analysis cannot capture the emergence of conceptual divergence as a country slides from democracy into authoritarianism, which is precisely what the Thai trajectory documents. @Bryan2023-eu pushes furthest, demonstrating across seventy-four countries that partisans in power hold more illiberal conceptions of democracy while maintaining explicit support for the democratic label, framing the pattern as partisan-motivated reasoning operating universally. The global scope is impressive, but the universality of the account is also its limitation: if conceptual divergence is a general feature of being in power, then it should appear everywhere and always, which sits uneasily with cases like South Korea where the gap remains negligible despite dramatic political upheaval. The positional updating framework advanced here differs from all three in specifying democratic erosion as the conditioning variable—not a single authoritarian victory (Cohen et al.), not a static regime category (Wu and Chang), and not an unconditional cognitive bias (Bryan), but a dynamic process that raises the stakes of exclusion and thereby amplifies losers' procedural orientation.

## Positional updating: mechanism and predictions

I propose that electoral status shapes democratic conceptions through what I call positional updating. The core logic turns on asymmetric salience: procedures and outcomes carry different weight depending on whether one is in or out of power. If your party runs the government, your attention naturally drifts toward policy outputs—what is being built, distributed, reformed. If your party lost, you start to care a great deal about the rules of the game: Is there a fair shot at winning next time? Can the press still criticize? Can the courts still push back?

This selective attention runs on motivated cognition [@Kunda1990-uw; @Taber2006-kx]. Citizens do not arrive at their understanding of democracy through disinterested reflection; rather, they attend to and weight those dimensions of democracy that are instrumentally relevant to their political position. Losers orient toward procedural features because procedures are the tools of the excluded—the mechanisms through which power can be contested, checked, and ultimately recaptured. Winners orient toward substantive outcomes because they occupy the position from which outcomes are delivered. In this sense, the mechanism parallels the political conditioning of economic perceptions documented in other domains: just as partisan position shapes whether citizens view the economy favorably [@Evans2006-np; @Fiorina1978-wf], electoral status shapes which features of democracy citizens attend to and prioritize. Emphasizing procedural democracy also serves losers' discursive interests: a government can be criticized for violating democratic procedures in ways that a government cannot easily be criticized for failing to realize one's preferred substantive vision.

A clarification is in order: this framework does not require losers to sit down and calculate their odds of returning to power. The mechanism is more basic than that. People who backed the losing side—whatever their party's organizational heft—find themselves on the outside of the coalition that runs the state, makes policy, and allocates resources. That structural fact sharpens their awareness of the procedural guardrails that limit what incumbents can do. It is not that losers cynically adopt a procedural vocabulary; it is that procedural protections become more relevant to people who lack access to executive power.

This exclusion-based reading yields a testable distinction. If positional updating were really about partisan strategy—opposition leaders rallying their base around procedural rhetoric—we would expect the effect to cluster among supporters of the main opposition party, whose leadership has both the organizational muscle and the electoral motive to wave the procedural banner. But if what drives the effect is the more diffuse experience of being locked out of government, then the procedural orientation should show up across the board: minor-party supporters, independents, anyone who did not back the winner. The supplementary analysis in Appendix J confirms the latter pattern.

From this positional account I derive three predictions. Each one helps separate the positional story from the most obvious rival—that democratic conceptions are essentially stable cultural orientations or fixed personality traits.

The first prediction is shared with the cultural account: losers should lean procedural and winners should lean substantive. A cultural explanation could generate the same cross-sectional pattern by pointing to the ideological profiles of parties that happen to be in opposition. So H1 alone cannot adjudicate between the two.

*H1:* Electoral losers will be more likely than winners to prioritize procedural over substantive conceptions of democracy.

The second prediction is where the positional logic starts to bite. If what matters is the *stakes* of losing, then the gap should widen where losing really hurts—where institutions are weak, alternation is rare, and exclusion from power threatens to become permanent. Loss aversion [@Kahneman1979-af] reinforces the point: the prospect of losing established rights weighs more heavily than the prospect of gaining new benefits, which should amplify losers' procedural orientation in high-stakes environments.

*H2:* The loser effect will be larger in political contexts where losing is more consequential—where democratic institutions are weaker or power alternation less frequent.

Third, and most critically for distinguishing positional updating from cultural accounts: if conceptions are cultural inheritances or stable dispositions, they should not shift appreciably within countries over periods shorter than a generation. The positional account, by contrast, predicts that conceptions will respond dynamically to changes in the political stakes of losing—growing when institutions weaken and shrinking when institutional protections are restored through alternation or constitutional resolution of crises.

*H3:* Within countries, the loser effect will respond to changes in political context, growing when democratic institutions weaken and shrinking when power alternates.

H3 is the hypothesis that most sharply distinguishes the positional account from its alternatives, and the one whose confirmation would represent the strongest evidence for the theoretical framework advanced here. The within-country trajectories examined below provide the most direct test.

# Data and Methods

## Data

The analysis draws on the Asian Barometer Survey (ABS), a regional comparative project tracking democratic values across Asia since 2001 [@ABS-W2; @ABS-W3; @ABS-W4; @ABS-W6].[^1] The ABS covers countries spanning the full range of regime types [@Chu2016-va] and employs forced-choice items requiring respondents to prioritize among competing democratic values, providing cleaner identification of relative orientations than agree-disagree batteries.

[^1]: Data analyzed in this article were collected by the Asian Barometer Project (2005--2008, 2010--2012, 2013--2016, 2020--2022), co-directed by Professors Fu Hu, Yun-han Chu, and Min-hua Huang, with major funding support from Taiwan's Ministry of Education, the National Science and Technology Council, Academia Sinica, and National Taiwan University. The Asian Barometer Project Office (www.asianbarometer.org) is solely responsible for data distribution. The views expressed herein are the author's own.

```{r data-overview}
#| include: false
# Count waves and country-waves
n_cw <- n_distinct(paste(cw_gap$country_name, cw_gap$wave_label))
waves_used <- c("W2 (2005-2008)", "W3 (2010-2012)", "W4 (2014-2016)", "W6 (2019-2022)")
```

Four waves are analyzed: Wave 2 (2005--2008), Wave 3 (2010--2012), Wave 4 (2014--2016), and Wave 6 (2019--2022). Wave 5 is excluded because it did not include the relevant items. China and Vietnam are excluded because their single-party systems do not produce meaningful electoral winners and losers. The analysis is restricted to respondents who reported voting in the most recent national election, as winner-loser status cannot be meaningfully assigned to non-voters. This restriction is not merely a methodological necessity but a theoretically motivated choice: voters are the citizens whose democratic conceptions most directly translate into political demands through electoral participation, party mobilization, and protest. A study of how electoral outcomes reshape democratic understanding is, by definition, a study of the electorally engaged. After these restrictions, the sample comprises `r fmt_n(total_n)` observations across `r n_countries` countries.

## Measurement

In Waves 3, 4, and 6, the ABS presented respondents with four item sets, each containing four statements describing different aspects of democracy. Respondents indicated which statement they considered most important. Each set included procedural items (e.g., "People can choose their leaders in free and fair elections," "Government protects people's freedom of speech"), substantive items (e.g., "Government reduces the gap between the rich and the poor," "Government provides everyone with the basic necessities of life"), and governance items (e.g., "Government does not waste public money," "People's safety from crime is guaranteed"). Wave 2 employed a single forced-choice item with different response options and is analyzed separately.

The key independent variable is electoral status, derived from the ABS vote-choice item (q34a in Waves 4 and 6, q33a in Wave 3, q39a in Wave 2). For each country-wave, the ABS research team codes each respondent's self-reported vote choice against the outcome of the most recent national election prior to fieldwork, classifying respondents as winners (voted for the governing party, president, or coalition partner) or losers (voted for any non-governing party). The analysis adopts this binary coding throughout—a simplification, certainly, since it treats a voter for a marginal coalition partner identically to a voter for the dominant governing party, but one that is standard in the literature and that preserves comparability across very different party systems. Respondents who did not vote, refused to answer, or reported an invalid vote choice are excluded, restricting the sample to citizens whose electoral participation permits meaningful assignment of winner-loser status. Thailand Wave 4 (2014--2016) presents a distinctive case: fieldwork occurred after the May 2014 military coup, which followed the annulment of the February 2014 election. The ABS nevertheless coded winner/loser status for 382 respondents based on prior electoral behavior; the analysis retains this country-wave, and a robustness check verifies (Appendix Table K1) that all core findings hold when it is excluded.[^2]

[^2]: The relatively high rate of missing data on q34a in Thailand Wave 4 (51.1%) likely reflects respondent reluctance to disclose vote choice under military rule rather than a failure of the survey instrument. The 382 respondents with valid coding yield winner and loser subsamples of 186 and 227, respectively—sufficient for stable estimation. Appendix Section B documents the reference election, winning party, and runner-up for each country-wave; reports an independent verification of the ABS coding against official election commission data for Taiwan, South Korea, and Thailand; and discusses additional edge cases including Thailand's government-formation coding and Cambodia's opposition dissolution. Control variables include age, gender, education level, and urban-rural residence.

## Analytical Strategy

Collapsing responses into a binary procedural–substantive indicator would sacrifice the item-level heterogeneity that is central to the analysis. Instead, each item set is modeled as a separate multinomial logistic regression, preserving all four response categories within each set. For Waves 3, 4, and 6, which share a common instrument of four item sets (Sets 1–4), data are pooled across waves and modeled jointly. Each set takes the form:

$$
\Pr(Y_{ij} = k) = \frac{\exp(\beta_{0k} + \beta_{1k}\text{Loser}_i + \mathbf{X}_i\boldsymbol{\gamma}_k + \alpha_{ck} + \delta_{tk})}{\sum_{m=1}^{4}\exp(\beta_{0m} + \beta_{1m}\text{Loser}_i + \mathbf{X}_i\boldsymbol{\gamma}_m + \alpha_{cm} + \delta_{tm})}
$$

where $Y_{ij}$ denotes the item selected by respondent $i$ in set $j$, $k$ indexes the four response categories within each set, $\mathbf{X}_i$ is a vector of demographic controls (age, gender, education, and urban–rural residence), and $\alpha_{ck}$ and $\delta_{tk}$ represent category-specific country and wave fixed effects. The reference category is set internally by the estimation routine; because results are reported as average marginal effects (AMEs), the choice of baseline category does not affect interpretation. AMEs express the effect of loser status as percentage-point changes in the predicted probability of selecting each item—a quantity that is directly interpretable and comparable across all twenty items without requiring the reader to mentally translate log-odds ratios.

Wave 2, which employed a single forced-choice item with a distinct set of response options, is modeled separately using the same multinomial logit specification (without wave fixed effects, as it draws from a single survey round). The inclusion is not merely a methodological nicety: because Wave 2's instrument differs in structure from the later waves, replication across both formats provides a check on whether the loser effect is a feature of the underlying political psychology or an artifact of a particular battery design.

Standard errors are computed via wild cluster bootstrap at the country level [@Cameron2008-aw]. The rationale is straightforward—with only eleven country clusters, conventional cluster-robust variance estimators produce downward-biased standard errors and over-reject the null—but the practical consequences are worth stating. By imposing the null hypothesis on the bootstrap data-generating process, this procedure trades statistical power for honest inference. The tests throughout the analysis are accordingly conservative, which means that the effects reported below had to clear a higher bar than in most comparative survey research with similar sample sizes.

To examine country-level variation and within-country trajectories (H2 and H3), the multinomial logit models are re-estimated separately for each country-wave combination. For each estimate, I compute the procedural–substantive gap: the difference between the mean AME of loser status across procedural items and the mean AME across substantive items within a given country-wave. A positive gap indicates that losers disproportionately favor procedural conceptions relative to substantive ones. Confidence intervals for these gaps are obtained via the same wild bootstrap procedure. Tracking these gaps across waves within countries is where the analysis moves beyond cross-sectional description, allowing direct assessment of whether conceptions of democracy respond to changing political conditions—the central prediction that distinguishes the positional account from cultural or dispositional alternatives.

# Results

## The Baseline Loser Effect

```{r pooled-summary}
#| include: false
# Summary stats for inline text
proc_items <- pooled %>% filter(item_type == "procedural")
sub_items <- pooled %>% filter(item_type == "substantive")
gov_items <- pooled %>% filter(item_type == "governance")

n_proc_pos <- sum(proc_items$ame > 0)
n_proc_total <- nrow(proc_items)
n_proc_sig <- sum(proc_items$p < 0.05)

n_sub_neg <- sum(sub_items$ame < 0)
n_sub_total <- nrow(sub_items)
n_sub_sig <- sum(sub_items$p < 0.05)

proc_range_lo <- fmt_pct(min(proc_items$ame[proc_items$ame > 0]))
proc_range_hi <- fmt_pct(max(proc_items$ame))

sub_range_lo <- fmt_pct(min(sub_items$ame))
sub_range_hi <- fmt_pct(max(sub_items$ame[sub_items$ame < 0]))

# Largest effects
largest_proc <- proc_items %>% filter(ame == max(ame))
largest_sub <- sub_items %>% filter(ame == min(ame))
largest_gov <- gov_items %>% filter(ame == min(ame))

# Free elections item
free_elections <- proc_items %>% filter(item_label == "Free elections")
```

Table 1 reports the average marginal effect of loser status on the probability of selecting each of the twenty items across the pooled sample. The pattern is consistent with H1: `r n_proc_pos` of `r n_proc_total` procedural items show positive effects (losers more likely to select), while all `r n_sub_total` substantive items show negative effects (winners more likely to select). Procedural effects range from `r proc_range_lo` to `r proc_range_hi` percentage points, with `r largest_proc$item_label` showing the largest effect (`r fmt_pp(largest_proc$ame)` pp, $p$ `r ifelse(largest_proc$p < 0.001, "< 0.001", sprintf("= %.3f", largest_proc$p))`). Substantive effects range from `r sub_range_lo` to `r sub_range_hi` pp, with `r largest_sub$item_label` as the largest (`r fmt_pp(largest_sub$ame)` pp, $p$ `r ifelse(largest_sub$p < 0.001, "< 0.001", sprintf("= %.3f", largest_sub$p))`).

```{r pooled-table}
#| output: asis

# Build the main results table
tbl <- pooled %>%
  arrange(desc(ame)) %>%
  mutate(
    Type = str_to_title(item_type),
    Item = item_label,
    `AME (pp)` = sprintf("%+.1f", ame * 100),
    SE = sprintf("%.1f", se * 100),
    Sig = sig
  ) %>%
  select(Item, Type, `AME (pp)`, SE, Sig)

if (knitr::is_latex_output()) {
  tbl_out <- tbl %>%
    kbl(
      booktabs = TRUE,
      longtable = FALSE,
      linesep = "",
      caption = "Average Marginal Effect of Loser Status on Item Selection"
    ) %>%
    kable_styling(
      latex_options = c("striped", "hold_position"),
      font_size = 10,
      full_width = TRUE
    ) %>%
    footnote(
      general = "Average marginal effects from multinomial logit with country and wave fixed effects and demographic controls (age, gender, education, urban residence). Positive values indicate losers are more likely to select the item. Bootstrap SEs clustered at the country level.",
      symbol = c("$\\\\dagger$ p < 0.10, * p < 0.05, ** p < 0.01, *** p < 0.001"),
      general_title = "Note: ",
      footnote_as_chunk = TRUE,
      escape = FALSE
    )
  # Make footnote rows wrap and single-space
  tbl_str <- as.character(tbl_out)
  tbl_str <- gsub("\\multicolumn{5}{l}{", "\\multicolumn{5}{p{\\linewidth}}{\\singlespacing ", tbl_str, fixed = TRUE)
  cat(tbl_str)
} else {
  tbl %>%
    kable(format = "pipe", align = c("l", "l", "l", "r", "r", "c"))
}
```

Individually, these effects are not large—on the order of two to four percentage points per item. But individual effect sizes are not the point. What stands out is the sheer consistency of the pattern across twenty items spread over five separate batteries. No single item drives the result; no single battery drives it. Winners and losers occupy measurably different positions in the conceptual space of democratic meaning, and they do so in the same direction virtually everywhere you look: losers zero in on the rules of political competition, winners on the material payoffs of governance.

One item warrants specific attention. Free elections, the most canonical procedural item, shows a *negative* AME (`r fmt_pp(free_elections$ame)` pp), meaning losers are slightly *less* likely than winners to identify free elections as the most important feature of democracy. This runs counter to H1's general prediction, but the pattern becomes interpretable in light of the remaining procedural results. The items that losers *do* favor are the liberal components of democratic governance: free expression (`r fmt_pp(proc_items %>% filter(item_label == "Free expression") %>% pull(ame))` pp), media freedom (`r fmt_pp(proc_items %>% filter(item_label == "Media freedom") %>% pull(ame))` pp), the right to organize groups (`r fmt_pp(proc_items %>% filter(item_label == "Organize groups") %>% pull(ame))` pp), and party competition (`r fmt_pp(proc_items %>% filter(item_label == "Party competition") %>% pull(ame))` pp). Losers who have experienced electoral manipulation or whose preferred parties have been dissolved by courts may have reason to distrust the electoral mechanism itself while valuing the broader liberal ecosystem (media, expression, association, judicial accountability) that enables political contestation outside formal elections. The loser effect, in other words, is not an elections effect but a *liberal-democratic* effect: losers gravitate toward the institutional protections that sustain opposition rather than the specific electoral procedure that has failed to deliver victory.

This pattern is consistent with what might be termed *cynical proceduralism*: losers do not naively embrace all democratic procedures but discriminate between constraint-based safeguards (courts, media, expression) whose protective value is independent of electoral outcomes and selection-based mechanisms (elections) whose credibility is contingent on perceived integrity.

The governance items reveal an additional pattern. Winners are consistently more likely to select governance items emphasizing order and state capacity, with `r largest_gov$item_label` showing the largest effect (`r fmt_pp(largest_gov$ame)` pp, $p$ `r ifelse(largest_gov$p < 0.001, "< 0.001", sprintf("= %.3f", largest_gov$p))`). Winners conceive of democracy as a framework for stability and effective administration; losers conceive of it as a framework for contestation and the protection of rights. This order-versus-contestation tension is explored further in the discussion.

## Democratic Erosion and Institutional Resilience

```{r trajectory-stats}
#| include: false
# Key trajectory stats
# Thailand
thai_w3_pp <- round(thai_w3_gap * 100, 1)
thai_w4_pp <- round(thai_w4_gap * 100, 1)
thai_w6_pp <- round(thai_w6_gap * 100, 1)
thai_w6_ci <- thai %>% filter(wave == 6)

# South Korea
kor <- cw_gap %>% filter(country_name == "South Korea") %>% arrange(wave)
kor_w3_pp <- round(kor %>% filter(wave == 3) %>% pull(proc_sub_gap) * 100, 1)
kor_w4_pp <- round(kor %>% filter(wave == 4) %>% pull(proc_sub_gap) * 100, 1)
kor_w6_pp <- round(kor %>% filter(wave == 6) %>% pull(proc_sub_gap) * 100, 1)

# Japan
jpn <- cw_gap %>% filter(country_name == "Japan") %>% arrange(wave)

# Cambodia
khm <- cw_gap %>% filter(country_name == "Cambodia") %>% arrange(wave)
```

The pooled estimates establish that losers systematically favor procedural conceptions, but the theory predicts more than a cross-sectional association: it predicts that the gap should respond to changing political conditions within countries over time (H2 and H3). Figure 1 plots the procedural-substantive gap for each country across survey waves, where a positive value indicates that losers disproportionately favor procedural conceptions.

```{r fig-trajectories, fig.width=10, fig.height=6}
#| fig-cap: "Loser Effect Trajectories: Procedural--Substantive Gap by Country and Wave"
#| fig-alt: "Line plot showing procedural-substantive gap trajectories across three survey waves for eleven Asian countries. Thailand shows dramatic increase from near zero to over 20 percentage points. South Korea remains near zero throughout. Cambodia maintains a high gap around 13 points. Other countries shown in gray."

# Plot countries with 2+ waves
plot_data <- cw_gap %>%
  group_by(country_name) %>%
  filter(n() >= 2) %>%
  ungroup()

highlight <- c("Thailand", "South Korea", "Japan", "Cambodia")

plot_data %>%
  mutate(
    highlighted = country_name %in% highlight,
    country_label = if_else(highlighted, country_name, "Other")
  ) %>%
  ggplot(aes(x = wave, y = proc_sub_gap * 100, group = country_name)) +
  geom_line(
    data = . %>% filter(!highlighted),
    color = "gray75", linewidth = 0.5, alpha = 0.7
  ) +
  geom_point(
    data = . %>% filter(!highlighted),
    color = "gray75", size = 1.5, alpha = 0.7
  ) +
  geom_ribbon(
    data = . %>% filter(highlighted),
    aes(ymin = proc_sub_gap_ci_low * 100, ymax = proc_sub_gap_ci_high * 100,
        fill = country_name),
    alpha = 0.15, color = NA
  ) +
  geom_line(
    data = . %>% filter(highlighted),
    aes(color = country_name), linewidth = 1.2
  ) +
  geom_point(
    data = . %>% filter(highlighted),
    aes(color = country_name), size = 3
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous(
    breaks = c(3, 4, 6),
    labels = c("W3\n(2010-12)", "W4\n(2014-16)", "W6\n(2019-22)")
  ) +
  scale_color_manual(values = c(
    "Thailand" = "#D62728", "Cambodia" = "#FF7F0E",
    "South Korea" = "#2CA02C", "Japan" = "#1F77B4"
  )) +
  scale_fill_manual(values = c(
    "Thailand" = "#D62728", "Cambodia" = "#FF7F0E",
    "South Korea" = "#2CA02C", "Japan" = "#1F77B4"
  )) +
  labs(
    x = "Survey Wave",
    y = "Procedural - Substantive Gap\n(percentage points)",
    color = NULL, fill = NULL,
    caption = "Note: Gray lines show remaining countries. Shaded bands show 95% CIs for highlighted countries.\nPositive values = losers favor procedural items relative to substantive items."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

Thailand provides the most dramatic support for H3. In Wave 3 (2010--2012), during the Democrat Party government of Abhisit Vejjajiva, the procedural--substantive gap was negligible (`r thai_w3_pp` pp, not significant). By Wave 4 (2014--2016), conducted after the May 2014 coup that removed Prime Minister Yingluck Shinawatra ($N = 382$ respondents with valid winner/loser coding; see Methods for discussion of post-coup coding), the gap had widened to `r thai_w4_pp` pp ($p < 0.05$). The positive gap is driven by Democrat and minor-party voters---the electoral losers under ABS coding---who increasingly emphasized procedural conceptions, while Pheu Thai supporters---electoral winners despite the coup that removed their government---shifted toward substantive definitions. By Wave 6 (2019--2022), after six years of military or military-backed governance, the gap reached `r thai_w6_pp` pp ($p < 0.001$, 95% CI: \[`r round(thai_w6_ci$proc_sub_gap_ci_low * 100, 1)`, `r round(thai_w6_ci$proc_sub_gap_ci_high * 100, 1)`\]). This trajectory, from zero to over twenty percentage points in a decade, tracks the contours of Thailand's democratic erosion [@Kuhonta2014-qk; @Ricks2019-hs]. This period also saw the determinants of political engagement shift fundamentally, with conventional socioeconomic predictors giving way to group mobilization and party attachment as the primary drivers of participation [@Pankaew2022-pr]. As the coup and a string of judicial interventions kept overturning electoral results, voters aligned with the excluded Thaksin faction came to define democracy increasingly in procedural terms—free speech, media independence, accountability. Backers of the military-aligned establishment moved in the opposite direction, toward substantive outcomes.

South Korea presents a strikingly divergent trajectory. The country went through its own political earthquake during this same period—the impeachment and removal of President Park Geun-hye in 2016--2017, followed by the election of the progressive Moon Jae-in—yet the procedural--substantive gap barely moved. It hovered between `r kor_w3_pp` pp in Wave 3 and `r kor_w4_pp` pp in Wave 4, never cracking five percentage points. What separates the two cases is institutional: South Korea resolved its crisis through constitutional channels—impeachment, judicial review, a free election—rather than military intervention. The impeachment showed that accountability mechanisms actually worked, and the 2017 election produced genuine turnover. Under these conditions, the positional logic predicts convergence, and convergence is what the data show.

Cambodia illuminates a different boundary condition: a persistently large gap (approximately 13 pp across all three waves) that, notably, did not grow following the dissolution of the main opposition party (CNRP) in 2017. In a competitive authoritarian context, the stakes of losing were already effectively maximized—losers' procedural orientation had little room to intensify further. Among the remaining countries (Table 2), Malaysia's gap narrows across waves, potentially reflecting the political opening that would culminate in the extraordinary 2018 alternation—the first change in government since independence. Taiwan reverses from positive to negative by Wave 6, a pattern that likely reflects cross-strait identity politics more than the winner-loser dynamic examined here. (Myanmar's single data point captures a strikingly large gap consistent with the stakes of its fragile transition, though little can be inferred from one observation.) Japan, as expected for a consolidated democracy, shows a modest and stable gap throughout.

```{r country-gaps}
#| output: asis

gap_tbl <- cw_gap %>%
  mutate(
    gap_label = sprintf("%+.1f [%+.1f, %+.1f]",
                        proc_sub_gap * 100,
                        proc_sub_gap_ci_low * 100,
                        proc_sub_gap_ci_high * 100)
  ) %>%
  select(country_name, wave_label, gap_label) %>%
  pivot_wider(names_from = wave_label, values_from = gap_label, values_fill = "—") %>%
  rename(Country = country_name) %>%
  select(Country, W3, W4, W6)

if (knitr::is_latex_output()) {
  tbl_out <- gap_tbl %>%
    kbl(
      booktabs = TRUE,
      longtable = FALSE,
      linesep = "",
      label = NULL,
      caption = "Procedural--Substantive Gap by Country and Wave"
    ) %>%
    kable_styling(
      latex_options = c("striped", "hold_position"),
      font_size = 10
    ) %>%
    column_spec(1, width = "3cm") %>%
    column_spec(2, width = "3.5cm") %>%
    column_spec(3, width = "3.5cm") %>%
    column_spec(4, width = "3.5cm") %>%
    footnote(
      general = "Gap = (Mean Procedural AME) - (Mean Substantive AME), in percentage points, with 95% bootstrap CIs. Positive values indicate losers favor procedural items relative to substantive items. Dashes indicate country not surveyed in that wave.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE,
      threeparttable = TRUE
    )
  tbl_str <- as.character(tbl_out)
  tbl_str <- gsub("\\label{tab:country-gaps}", "", tbl_str, fixed = TRUE)
  cat(tbl_str)
} else {
  gap_tbl %>%
    kable(format = "pipe", align = c("l", "c", "c", "c"))
}
```

## Testing the Positional Mechanism

The twenty-item pattern and the divergent country trajectories raise the question of mechanism. Are losers actually revising what democracy means to them in response to the political stakes they face? Or are we simply observing the fact that different sorts of people end up on the winning and losing sides? Three tests, each targeting a different alternative explanation, help adjudicate.

### Test 1: Fairness Amplification

If conceptions reflect positional updating, the loser effect should amplify when procedural threat is salient, that is, among losers who perceive elections as unfair. If instead it reflects stable normative commitments, it should appear regardless of fairness perceptions.

However, not all procedural conceptions carry the same relationship to electoral status, and specifying this structure *ex ante* disciplines the interpretation of the interaction results. A distinction can be drawn between two classes of procedural items. The first encompasses *protective* procedural elements: institutional safeguards that constrain incumbent power and protect political minorities regardless of who governs, including an independent judiciary, media freedom, and free expression. For electoral losers, these protections are unambiguously valuable—they represent the institutional infrastructure that prevents winners from converting temporary electoral mandates into permanent domination. If losers are updating their conceptions in response to positional threat, we should expect the loser effect on protective items to remain positive regardless of whether respondents think elections are fair—courts and a free press are useful to the excluded either way. But what the positional framework predicts for *participatory* items is less obvious and more interesting.

In a well-functioning democracy, losers have good reason to prize elections: elections are how you get back into power. But when democratic erosion has corrupted the electoral process itself, losers face a bind. The institution that is supposed to enable turnover has, in their view, been rigged. Under those circumstances, losers may start to discount elections *specifically* while doubling down on non-electoral safeguards. The prediction is not a generic amplification of all procedural items among disaffected losers; it is a structured pattern where protective items strengthen and participatory items weaken or reverse.

This yields a set of conditional expectations: (a) protective procedural items should show a positive loser effect under both fairness conditions, strengthening where elections are seen as manipulated; (b) participatory procedural items should show a positive loser effect where elections are seen as fair; and (c) participatory items should attenuate or flip negative where losers believe the electoral process itself has been compromised.

|   | Perceived Electoral Integrity | Perceived Electoral Manipulation |
|------------------------|------------------------|------------------------|
| **Protective procedures** (courts, media, expression) | Loser effect positive | Loser effect positive (strengthened) |
| **Participatory procedures** (free elections, multiparty competition) | Loser effect positive | Loser effect attenuated or negative |

```{r fairness-data}
#| include: false
# Load fairness interaction results
fairness <- read_csv(file.path(results_path, "robustness_fairness_interaction.csv"), show_col_types = FALSE)

# Key interaction effects
court_int <- fairness %>% filter(item == "Court protection") %>% pull(interaction)
media_int <- fairness %>% filter(item == "Media freedom") %>% pull(interaction)
expression_int <- fairness %>% filter(item == "Free expression") %>% pull(interaction)
elections_int <- fairness %>% filter(item == "Free elections") %>% pull(interaction)

# Count positive interactions among procedural items
proc_fairness <- fairness %>% filter(item_type == "procedural")
n_proc_pos_int <- sum(proc_fairness$interaction > 0)
n_proc_int_total <- nrow(proc_fairness)
```

The fairness interaction results align closely with the conditional expectations derived from the protective-participatory distinction. Among `r n_proc_int_total` procedural items, `r n_proc_pos_int` show a positive interaction—meaning the loser effect is larger among those perceiving elections as unfair. The amplification is especially pronounced for protective procedural items: court protection shows an interaction of `r sprintf("%+.1f", court_int * 100)` percentage points, media freedom `r sprintf("%+.1f", media_int * 100)` pp, and free expression `r sprintf("%+.1f", expression_int * 100)` pp. Losers who perceive the electoral process as compromised do not simply withdraw from procedural commitments; they double down on precisely the non-electoral safeguards that would constrain those who benefit from unfair elections.

The participatory items tell a different story, exactly as the theory anticipated. For free elections, the interaction between loser status and perceived fairness runs *negative* (`r sprintf("%+.1f", elections_int * 100)` pp): losers who think the electoral process is rigged are actually *less* inclined to name free elections as democracy's most important feature. This is not a puzzle—it is a coherent response. If you believe that the primary mechanism for changing governments has been corrupted, you do not abandon democracy; you redefine it around the safeguards that still work independently of elections. The negative coefficient does not undercut the positional updating framework. It reveals the framework's internal architecture. Losers are not mechanically inflating everything procedural; they are paying closest attention to whichever procedures offer real protection given the specific threats they perceive.

This structured divergence also helps dispose of a simpler alternative reading—that the loser effect is just generalized electoral discontent expressed through procedural vocabulary. If losers who feel cheated were merely being negative about the whole system, the interaction should push in the same direction for every procedural item. Instead we see the opposite: protective items go up, participatory items go down. That is harder to square with generalized grievance and easier to square with a discriminating assessment of which institutions still function as credible constraints.

```{r fig-fairness}
#| fig-cap: "Fairness Interaction: Loser Effect by Perceived Electoral Fairness"
#| fig-alt: "Dot plot showing loser effects separately for respondents perceiving fair vs. unfair elections, across sixteen items. Liberal procedural items show substantially larger loser effects among those perceiving unfair elections."
#| fig-width: 8
#| fig-height: 6

# Plot interaction effects
fairness %>%
  mutate(
    item = fct_reorder(item, interaction),
    item_type = str_to_title(item_type)
  ) %>%
  pivot_longer(
    cols = c(loser_effect_fair, loser_effect_unfair),
    names_to = "condition",
    values_to = "loser_effect"
  ) %>%
  mutate(
    condition = if_else(condition == "loser_effect_fair", "Fair election", "Unfair election")
  ) %>%
  ggplot(aes(x = loser_effect * 100, y = item, color = condition, shape = condition)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  geom_point(size = 3, position = position_dodge(width = 0.5)) +
  facet_wrap(~ item_type, scales = "free_y", ncol = 1) +
  scale_color_manual(values = c("Fair election" = "#1F77B4", "Unfair election" = "#D62728")) +
  labs(
    x = "Loser Effect (percentage points)",
    y = NULL,
    color = NULL, shape = NULL,
    caption = "Note: Loser AMEs from multinomial logit interacting loser status with perceived electoral fairness.\nCountry and wave fixed effects with demographic controls."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold")
  )
```

A qualification is necessary. Perceived unfairness is endogenous to loser status: losers are approximately `r sprintf("%.1f", endog$ratio)` times more likely to perceive elections as unfair (`r sprintf("%.1f", endog$pct_unfair_losers)`% vs. `r sprintf("%.1f", endog$pct_unfair_winners)`% of winners), though the point-biserial correlation is modest ($r$ = `r sprintf("%.3f", endog$r_pb)`, $N$ = `r format(endog$n_complete, big.mark = ",")`), indicating substantial overlap rather than collinearity. It remains unclear whether the interaction reflects genuine updating or pre-existing differences in outlook. The protective-participatory divergence mitigates but does not eliminate this concern, since motivated reasoning could produce domain-specific patterns. However, a pure "sore loser" interpretation is difficult to reconcile with the item-level structure of the results: if losers who perceive unfairness were simply expressing generalized grievance, the interaction should run uniformly in the same direction across all procedural items rather than diverging sharply between protective items (positive) and participatory items (negative). Because this test alone cannot definitively distinguish the mechanisms, two additional tests are needed.

### Test 2: Domain Specificity (Placebo)

If the loser-unfairness interaction simply captures generalized discontent ("losers who think elections are unfair are negative about everything"), it should extend to non-procedural domains like material welfare. If it reflects procedural-specific updating, the amplification should be confined to procedural safeguards.

I test whether the interaction extends to perceptions of basic necessities (q109: "People have food, clothes, and shelter"). While losers are slightly more likely than winners to agree people have basic necessities ($\beta$ = `r fmt_coef(placebo$main_loser_coef)`, SE = `r fmt_coef(placebo$main_loser_se)`, p \< 0.001)—a level effect potentially reflecting socioeconomic composition—the Loser × Unfairness interaction is not significant ($\beta$ = `r fmt_coef(placebo$interaction_coef)`, SE = `r fmt_coef(placebo$interaction_se)`, p = `r fmt_p(placebo$interaction_p)`).

Among those perceiving fair elections, the loser effect is `r sprintf("%.1f", placebo$ame_fair)` percentage points; among those perceiving unfair elections, it is `r sprintf("%.1f", placebo$ame_unfair)` points—a difference of only `r sprintf("%.1f", placebo$ame_diff)`pp (95% CI: \[`r sprintf("%.1f", placebo$ci_low)`, `r sprintf("%.1f", placebo$ci_high)`\]pp). This near-zero interaction contrasts sharply with procedural items, where the amplification averages 5–8 percentage points for liberal safeguards.

The mechanism is domain-specific, not a general pattern of negativity. Material welfare taps output legitimacy (whether government delivers); procedural items tap input legitimacy (whether processes are fair). The fact that unfairness perceptions amplify the loser effect for the latter but not the former is consistent with positional updating focused on procedural threat rather than motivated reasoning or diffuse discontent.

### Test 3: Not Compositional (Thailand Reweighting)

Thailand's dramatic trajectory—from near-zero to over 20 percentage points—invites a compositional objection. Perhaps opposition coalitions reshuffled over the decade, pulling in more urban and educated voters who were already predisposed toward procedural definitions. If so, the gap could widen without anyone actually changing their mind.

To address this, I reweight each Thailand wave to match the Wave 3 demographic profile along age, gender, education, and urban residence, creating a counterfactual trajectory that holds demographics constant. The exercise barely moves the needle. Loser effects shift by an average of `r sprintf("%.2f", composition$mean_abs_change)` percentage points (the largest shift is `r sprintf("%.2f", composition$max_abs_change)`pp, for `r composition$max_item`). Only `r composition$n_sign_changes` of `r composition$n_total_items` item-level effects changes sign, and the item in question (`r composition$sign_change_detail`) flips between values that are both essentially zero. Every substantively meaningful effect keeps its direction and its size.

So the trajectory is not about who ended up in the winner and loser camps—it is about how people in those camps came to think differently about democracy over time. The exercise does not rule out realignment along unmeasured dimensions like ideology or political interest. But age, education, and urbanization are the fault lines that structure Thai politics most powerfully [@Ockey2004-ek; @McCargo2019-rx], and controlling for them changes almost nothing.

### Cumulative Interpretation

No single test is definitive. But the three converge on the positional interpretation in a way that would be hard to orchestrate from any of the rival accounts. The fairness amplification shows that the loser effect responds to perceived threat—5–8 percentage points for liberal items among those perceiving unfair elections, compared with 2–3 among those perceiving fairness. That the placebo test finds no such amplification for basic necessities (`r sprintf("%.1f", placebo$ame_diff)`pp, p = `r fmt_p(placebo$interaction_p)`) rules out generalized discontent as the driver. And Thailand's demographic story holds up under reweighting: pinning age, education, gender, and urban residence to their Wave 3 distributions shifts the loser effects by `r sprintf("%.1f", composition$mean_abs_change)`pp on average, with a single near-zero sign flip.

These three pieces of evidence do not amount to proof that individuals changed their minds—establishing individual-level change would require panel data, and repeated cross-sections cannot deliver it. But the patterns are hard to square with the idea that democratic conceptions are stable cultural inheritances, hardwired personality traits, or byproducts of coalition shuffling. For the rival explanation to work, you would need ideological or dispositional differences between the winning and losing coalitions to have blown apart in Thailand while holding perfectly steady in South Korea, Cambodia, and nearly everywhere else—a scenario for which there is no independent evidence. The evidence at hand, taken as a whole, points toward positional updating: people rethinking what democracy means in light of where they stand politically and what the institutional landscape looks like around them.

# Robustness

The main findings survive a battery of sensitivity checks, each targeting a different potential vulnerability.

Wave 2 of the ABS provides the most stringent independent replication, since it employed a different forced-choice instrument with different response options than the Sets 1–4 used in later waves. Both procedural items show positive AMEs and both substantive items show negative AMEs, reinforcing confidence that the loser effect is not an artifact of the Wave 3–6 battery design. Open-ended responses from Wave 2 offer an additional form of validation: respondents selecting procedural items were significantly more likely to offer procedural definitions spontaneously ($r = 0.28$, $p < .001$; Appendix F), confirming that the forced-choice items capture something real about how people think rather than an artifact of how the questions were framed.[^3]

[^3]: Following established conceptual frameworks, open-ended responses were coded as procedural, substantive, or excluded. A respondent-level procedural proportion score (0--1) based on up to three codeable responses correlates at $r = 0.28$ with the closed-ended item. Given the noise inherent in open-ended coding, this represents meaningful convergent validity.

A range of more routine checks confirm the same pattern. WLS estimation at the country-wave level, COVID-attitude controls on Wave 6 models, and re-estimation excluding Thailand Wave 4 entirely all replicate the core findings (Appendix G, K1). That the pattern holds at two levels of analysis—individual respondents and country-wave cells—and survives a pandemic-period confound and an unusual post-coup country-wave makes it harder to attribute to any single modeling decision.

Non-voters provide another angle on the voter-restriction choice. Their procedural-substantive gap (14.3 pp) resembles that of winners (13.9 pp) rather than losers (17.3 pp), confirming that restricting the sample to voters does not inflate the estimated loser effect (Appendix H). A three-way decomposition distinguishing key opposition supporters from other losers pushes on the mechanism from a different direction: both subgroups display comparable procedural orientations across the three countries where decomposition was feasible (Appendix J). If the effect were driven by partisan strategic signaling—opposition leaders rallying their base around procedural rhetoric—it should concentrate among principal opposition supporters. That it does not is more consistent with the exclusion-based mechanism, where distance from state authority rather than organizational affiliation drives procedural attention.

Finally, the use of repeated cross-sections rather than panel data means within-country trajectories cannot definitively rule out compositional change. The Thai case offers the strongest counterargument: party loyalties in Thailand's color-coded politics are notably stable, and the gap's tracking of documented political events, rather than gradual demographic shifts, favors the positional interpretation.

# Discussion

The central finding of this article is not simply that electoral losers prefer procedural democracy, though they do, consistently, across twenty items and eleven countries. The more consequential finding is that this association varies systematically with political context. Thailand's trajectory from near-zero to over twenty percentage points across a single decade of democratic erosion is difficult to reconcile with accounts treating democratic conceptions as stable cultural inheritances or fixed psychological dispositions [@Norris2011-qz; @Inglehart1997-gw]. Nor is it plausibly compositional: Thailand's color-coded partisan alignments remained remarkably stable even as the parties themselves were repeatedly dissolved and reconstituted [@Sinpeng2021-hd; @McCargo2019-rx]. The cumulative evidence—fairness amplification, domain specificity, and compositional stability—tilts toward the positional interpretation.

While repeated cross-sections cannot pin down individual-level attitude change with certainty, the within-country dynamics—especially the Thailand-South Korea contrast—fit the positional updating story far better than they fit stable sorting on unobservable characteristics. The alternative would require that ideological or dispositional differences between winners and losers widened dramatically in Thailand during this period while staying flat in South Korea, Cambodia, and nearly everywhere else. No one has produced evidence for that kind of asymmetric realignment. The conditional character of the effect also helps distinguish positional updating from the motivated reasoning account advanced by @Bryan2023-eu, who treats conceptual divergence as an automatic consequence of holding power. If the mechanism were purely cognitive—partisans reflexively molding democratic meaning to fit their position—then South Korea's crisis should have produced the same divergence as Thailand's, since both cases involved sharp partisan conflict over institutional legitimacy. That it did not suggests the mechanism is better understood as a response to the institutional stakes of exclusion rather than to partisan status per se.

What does this mean for democratic resilience? The answer depends on whether we are looking at updating or sorting, and the two scenarios lead to quite different places. A growing body of work has asked whether broadly shared democratic attitudes can act as a brake on backsliding [@Claassen2020-hk; @Svolik2019-zi; @Graham2020-jr]. If what I have documented here really is positional updating, then those hopes face a structural problem: in the countries where erosion has gone furthest, the people who care most about procedural safeguards are precisely the people who have been pushed to the political margins. Aggregate measures of democratic support would then paint a misleadingly rosy picture—blending together winners who are indifferent to procedures with losers who are passionately committed but politically sidelined. And here is the bitter irony: the very tools through which democracies erode—executive overreach, the capture of courts, the muzzling of media [@Levitsky2018-dd; @Waldner2018-yw; @Luhrmann2019-ow]—are also the tools that neutralize the constituency most invested in procedural norms. @Cohen2023-sm document a parallel dynamic in Brazil: after Bolsonaro's victory, winners who expressed broad support for the democratic system simultaneously became more willing to countenance executive overreach, suggesting that system-level endorsement and procedural commitment can move in opposite directions. The Asian patterns reported here indicate that this decoupling is not confined to dramatic authoritarian victories but unfolds gradually across extended erosion trajectories.

If the patterns instead reflect stable sorting, the implications are less alarming but not trivial. It would mean that procedural commitment is clustered demographically or ideologically in ways that make it vulnerable to electoral fortunes. Either way, whether losers' procedural orientation ever translates into effective resistance—through protest, civil disobedience, transnational advocacy—is a question that remains wide open.

The results also complicate the winner-loser gap literature in a way that has implications for survey research. Previous work has assumed that winners and losers share a common understanding of democracy when they report their satisfaction with it or their trust in its institutions [@Anderson2005-uk; @Esaiasson2011-po; @Norris2012-er]. If winners and losers define democracy differently—winners emphasizing outcomes, losers emphasizing procedures—then the standard "satisfaction with democracy" question is not measuring the same thing for both groups. This is a measurement problem masquerading as a substantive finding. Losers may report dissatisfaction because they perceive procedural violations; winners may report satisfaction because they perceive policy delivery. @Rich2025-gd offers suggestive evidence for this interpretation: using Asian Barometer data from Taiwan, he documents that the winner-loser gap materializes in trust toward electoral institutions but not in overall satisfaction with democracy, consistent with winners and losers differentiating between dimensions of democratic performance rather than evaluating a single shared object. This disconnect suggests that the much-studied winner-loser gap in democratic satisfaction may partly reflect a *measurement* artifact: not different evaluations of the same object, but evaluations of different objects. @Wu2022-hz identify this problem in the East Asian context specifically, noting that in competitive authoritarian regimes, ruling party supporters who report high "support for democracy" may be endorsing a substantive vision centered on economic delivery rather than the procedural safeguards that external observers associate with the term. The present findings extend their insight temporally: the measurement disconnect is not simply a feature of regime type but one that deepens as democratic institutions erode within a given country.

The item-level results add a further dimension. The loser effect is not a generic procedural orientation but a specifically *liberal-democratic* one. Losers gravitate toward media freedom, free expression, the right to organize, and judicial accountability—the institutional infrastructure that sustains political opposition—rather than toward elections per se. The protective-participatory distinction introduced in the fairness analysis clarifies this pattern: losers consistently favor protective procedural elements (non-electoral safeguards that constrain incumbent power) while their orientation toward participatory elements (elections and multiparty competition) is conditional on whether they perceive the electoral process as compromised. The negative fairness interaction on the free elections item reflects not an anomaly but a coherent response—citizens who believe the mechanism of democratic alternation has been corrupted shift their definitional emphasis toward safeguards that operate independently of elections. The protective-participatory distinction, which has been implicit in theoretical treatments from @Dahl1977-wj through @Schedler1998-dl, appears to be psychologically real for citizens living through the consequences of electoral defeat. People do not relate to "procedural democracy" as a monolith; they parse it into the pieces that matter given the particular threats they face.

The order-versus-contestation tension visible in the governance items opens up another line of inquiry. Winners' marked preference for law and order as a defining feature of democracy, set against losers' emphasis on civil liberties and accountability, maps onto what @Slater2010-ba has called the politics of order in Southeast Asia. Where democratic institutions are sturdy enough, these two impulses—order and contestation—coexist in a productive tension. Where institutions are weak, the two visions of democracy can become mutually exclusive. In an era of global democratic recession [@Diamond2015-yb; @Foa2016-lq], knowing which vision predominates, and among whom, matters well beyond the region examined here.

The analysis faces limitations worth stating directly. The repeated cross-sectional design means I cannot directly watch individuals change their minds; the within-country trajectories are suggestive rather than dispositive. Eleven countries is not a large sample for cross-national inference—though it is worth noting that the Asian Barometer is among the very few surveys that has tracked democratic conceptions with forced-choice items over nearly two decades, and the consistency of the loser effect across such diverse political systems is itself informative. The compression of the gap in Wave 6 across most countries outside Thailand remains a puzzle. The pandemic's tendency to make state capacity salient for everyone is a plausible explanation, though not a fully satisfying one, and post-pandemic survey rounds should be able to test it. And the restriction to voters, while theoretically justified, means the findings describe the electorally active public, not the population at large.

# Conclusion

This article has demonstrated that electoral status is systematically associated with how citizens conceptualize democracy. Losers prioritize the liberal-procedural infrastructure—free expression, media freedom, judicial accountability, the right to organize—that sustains the possibility of political contestation. Winners prioritize substantive outcomes and the governance of order. The positional updating interpretation draws support from three converging empirical checks: the loser effect amplifies among those perceiving electoral unfairness, does not extend to non-procedural domains, and persists after demographic reweighting. The within-country trajectories do the heaviest lifting. Thailand and South Korea, pulling apart precisely as their institutional paths diverged, sit poorly with any account that treats democratic conceptions as fixed cultural products.

If this positional logic extends beyond Asia—and comparative work in Latin America and Africa is well placed to find out—then the link between public attitudes and democratic resilience may be more fragile than the literature has assumed. Democratic erosion would not merely degrade governance or shrink civil liberties; it would fracture the very meaning of democracy along the winner-loser divide, concentrating procedural commitments among the people least positioned to act on them. A priority for future research is whether losers' procedural orientation actually shapes what they do politically—whether it feeds into protest, civil disobedience, or transnational advocacy, or whether it remains, in the end, a pattern that shows up in surveys and nowhere else.

The limits of repeated cross-sections warrant caution about causal claims. I cannot directly observe anyone changing their mind, and I cannot entirely rule out that the patterns reflect ideological sorting along dimensions my data do not measure. But the associations documented here—their consistency across items, their responsiveness to political context, their concentration among those who perceive procedural threats—point toward a dynamic relationship between electoral experience and democratic meaning that previous work has overlooked. Whether that relationship is best described as updating, sorting, or some blend of the two remains to be settled. What the evidence does suggest, fairly clearly, is that the positional account has more going for it than cultural determinism does.

# References {.unnumbered}

\singlespacing

::: {#refs}
:::
