---
title: "The Education of a Citizenry: Hope, Demobilization, and Authoritarian Normalization in Cambodia, 2008–2021"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: now
date-format: "MMMM D, YYYY [at] h:mm A"
abstract: |
  Between 2008 and 2021, Cambodian citizens underwent one of the most dramatic transformations in political orientation recorded in contemporary survey research. Drawing on four waves of Asian Barometer Survey data (Waves 2, 3, 4, and 6; N ≈ 1,000–1,242 per wave), this article traces the arc of that transformation as a sequential process: from post-conflict acquiescence, through a remarkable peak of democratic aspiration and civic mobilization, to a final condition of political withdrawal and regime normalization following the elimination of organized opposition. The Cambodian case offers an unusually clean empirical window into how dominant-party authoritarian regimes produce political quiescence—not through the cultivation of enthusiastic support, but through a process the article terms *demobilization by subtraction*, in which the removal of credible democratic alternatives restructures citizen orientations across multiple attitudinal and behavioral dimensions simultaneously. The findings challenge accounts that treat authoritarian legitimacy as a product of performance or ideology, suggesting instead that the most effective mechanism of regime stabilization may be the systematic closure of the political imagination.
keywords:
  - Cambodia
  - Asian Barometer Survey
  - authoritarian normalization
  - demobilization by subtraction
  - competitive authoritarianism
  - democratic aspirations
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{caption}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \usepackage{colortbl}
          \usepackage{xcolor}
          \captionsetup{width=\textwidth}
          \setlength\LTcapwidth{\textwidth}
          \floatplacement{figure}{H}
          \floatplacement{table}{H}
          \raggedright
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
  docx:
    toc: true
    number-sections: true
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r setup}
#| include: false
library(tidyverse)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/cambodia-fairy-tale")
results_dir  <- file.path(paper_dir, "analysis/results")
fig_dir      <- file.path(paper_dir, "analysis/figures")
tbl_dir      <- file.path(paper_dir, "analysis/tables")

source(file.path(paper_dir, "R", "helpers.R"))

# dat <- readRDS(file.path(results_dir, "analysis_data.rds"))
stats <- readRDS(file.path(results_dir, "inline_stats.rds"))
```

# Introduction

In 2012, Cambodians rated the democratic future of their country at `r round(stats$dem_future_w3, 1)` out of 10.

It is worth pausing on that number. On a scale designed to capture the full range of democratic expectation, Cambodian respondents placed themselves at the extreme ceiling—more optimistic about their democratic trajectory than citizens of virtually any other country surveyed by the Asian Barometer [@Hu2023-ib]. This was not a country with free and fair elections, an independent judiciary, or a vibrant free press. It was a dominant-party authoritarian state under the continuous rule of Hun Sen and the Cambodian People's Party (CPP), a regime that had held power since 1985 and showed no signs of voluntary departure. And yet Cambodians believed, with remarkable consistency, that democracy was coming.

Nine years later, that same measure had fallen to `r round(stats$dem_future_w6, 1)` out of 10. In the same period, independent civic engagement—contacting influential figures outside the state, discussing politics, attending to political news—had collapsed to historic lows, even as participation through officially sanctioned channels quietly rose. Acceptance of authoritarian governance alternatives, including strongman rule, single-party rule, expert rule, and military rule, had risen sharply. Reported encounters with corruption had plummeted from `r round(stats$corrupt_witnessed_w3, 0)` percent to `r round(stats$corrupt_witnessed_w6, 0)` percent. And voter turnout in national elections, now conducted without any viable opposition, had paradoxically climbed to `r round(stats$voted_w6, 0)` percent.

What happened in between is the subject of this article. The answer, stated plainly, is that the Cambodian People's Party dissolved the only credible opposition party, the Cambodia National Rescue Party (CNRP), in November 2017, and in doing so restructured the political field so thoroughly that Cambodian citizens adjusted their attitudes, behaviors, and expectations in a remarkably coherent pattern. This article argues that this adjustment is best understood not as a story of authoritarian consolidation in the conventional sense—not as the triumph of ideology, coercion, or performance legitimacy—but as something quieter and, in its way, more profound: the education of a citizenry in the limits of the politically possible.

The article proceeds as a structured analytic narrative, organized around the four survey waves as acts in a political drama. This is a deliberate methodological choice. Cambodia's trajectory between 2008 and 2021 has the unusual property of being both analytically tractable—the survey data capture discrete moments in a well-documented political sequence—and narratively coherent, with each wave mapping onto a distinct phase of political life. The narrative structure allows the article to preserve the temporal dynamics that conventional cross-sectional analysis tends to flatten, while the survey data anchor the story in systematic evidence rather than impressionistic accounts.

The question of how dominant-party regimes endure has generated a rich architectural vocabulary in contemporary scholarship. We are accustomed to analyzing the structural mechanics of survival: the skewed playing fields of competitive authoritarianism, the meticulous menu of manipulation required to manage electoral autocracies, and the stabilizing pillars of legitimation, repression, and co-optation.[^1] Yet, this institutional focus, while vital for understanding the resilience of the state, often obscures the quiet, downstream effects on the citizenry. Theories of regime survival tend to treat the public as a variable to be managed—either repressed into submission or bought off through patronage. This article shifts the analytical lens. By tracing the Cambodian electorate's trajectory from 2008 to 2021, it focuses not on how the regime maintained power, but on what happens to citizen orientations when the democratic alternative is systematically removed. The resulting quiescence, this article argues, is not a product of active authoritarian legitimation, but of demobilization by subtraction.

[^1]: For the foundational typologies of these institutional survival mechanics, see @Levitsky2010-bc; @Schedler2013-qn; and @Gerschewski2013-en. While these frameworks masterfully explain elite-level regime consolidation, they are less equipped to explain the simultaneous, multi-dimensional shifts in mass public opinion observed in the Asian Barometer Survey data during periods of opposition collapse.

The concept advanced here—demobilization by subtraction—is distinct from the two mechanisms most commonly invoked to explain political quiescence under authoritarianism. The first, repression-driven demobilization, holds that citizens withdraw from political life because the costs of participation have been raised through coercion, surveillance, or the threat of punishment.[^3] The second, co-optation, holds that citizens acquiesce because they have been incorporated into the regime's distributional networks and calculate that loyalty pays better than resistance.[^4] Both mechanisms presuppose an active regime strategy directed at the population. Demobilization by subtraction operates differently. Its proximate cause is not an action taken against citizens but an action taken against the political field itself: the removal of the credible alternative around which civic mobilization had organized. The result is not fear-based withdrawal or interest-based compliance but something closer to what Gaventa termed "quiescence"—the internalization of powerlessness that follows from the sustained absence of viable channels for political expression.[^5] Where Lukes' third dimension of power operates by shaping preferences so that grievances never form, demobilization by subtraction operates one step later: grievances may persist, but the perceived vehicle for their expression has been eliminated, producing a cascade of behavioral and attitudinal adjustment that is visible across multiple survey dimensions simultaneously.[^6]

[^3]: On repression as a mechanism of demobilization, the foundational treatment is @Davenport2007-ww. For a comprehensive typology of repressive strategies, see also the essays collected in @Davenport2005-mk. On the specific mechanisms through which states demobilize protest, see @Earl2011-et.

[^4]: The co-optation framework is developed most fully in @Gandhi2008-dk. On the equilibrium logic of authoritarian power-sharing, see @Przeworski2000-ed.

[^5]: Gaventa's central insight—that sustained powerlessness produces not resistance but the internalization of quiescence—provides the closest existing theoretical analogue to the pattern observed in the Cambodian data, though his analysis is situated in a very different political context [@Gaventa1980-ob].

[^6]: The distinction matters: Lukes' third dimension operates *before* grievances form; demobilization by subtraction operates *after* grievances have formed and been politically activated, removing the vehicle for their expression rather than the grievances themselves [@Lukes2005-bc].

The remainder of this article is organized as four acts corresponding to the four available survey waves. Each section opens with a brief account of the political conditions prevailing at the time of the survey, then turns to the Asian Barometer data to trace how those conditions registered in citizen attitudes and behaviors. The analysis tracks five thematic domains across waves: political participation, authoritarian governance preferences, democratic expectations, corruption experience and perception, and media engagement and political interest. A concluding section draws out the theoretical implications of the Cambodian sequence for the study of authoritarian durability and democratic aspiration.

### The Quiet Kingdom: Cambodia in 2008

The first act of this story is a silence, and it is important to understand what kind of silence it was.^[Unless otherwise noted, the political chronology and electoral data in this section draw on @Hughes2010-ns; @Strangio2014-vi; and @Un2013-ge.]

By 2008, Cambodia had been under continuous CPP rule for over two decades. The Paris Peace Accords of 1991 and the UNTAC-supervised elections of 1993 had produced a brief, chaotic experiment in multiparty democracy, but the CPP had never genuinely relinquished power, and after a violent coup in 1997 and disputed elections in 1998 and 2003, the party had consolidated its dominance through a combination of patronage, institutional control, and selective coercion. The 2008 National Assembly elections, which the CPP won with 58 of 123 seats and subsequently expanded through coalition politics, were competitive in form but not seriously contested in outcome.

The CPP's dominance by 2008 rested on institutional foundations that went far beyond electoral mechanics. The party had constructed a dense patronage network linking the central state to village-level administration, with commune chiefs serving as the capillary system through which resources, information, and political loyalty flowed.^[On the CPP's patronage architecture, see Caroline Hughes and Kheang Un, *Cambodia's Economic Transformation* (Copenhagen: NIAS Press, 2011); @Un2012-im.] The military and security services were thoroughly integrated into the party apparatus, and the judiciary operated as an instrument of executive power rather than an independent check upon it.^[@Blake2019-ji; @Morgenbesser2016-vf.] Economic growth, driven by the garment export sector, construction, and an expanding tourism industry, provided a material foundation for the regime's implicit social contract: stability and modest prosperity in exchange for political acquiescence.^[@Hughes2011-zf]

The opposition landscape was fragmented and demoralized. The Sam Rainsy Party (SRP), the most prominent non-CPP formation, commanded a loyal but limited urban constituency and suffered from periodic leadership crises as Sam Rainsy himself cycled between exile and return. The Human Rights Party, led by Kem Sokha, drew support from a distinct base but lacked the organizational reach to challenge the CPP's rural dominance. Neither party, operating independently, presented a credible threat to CPP hegemony—a structural condition that would change dramatically with their merger four years later.^[@Noren-Nilsson2015-yi provides the most detailed account of the opposition landscape in this period and the narratives that would eventually fuel the CNRP's formation.]

The Wave 2 Asian Barometer data from this period paint a portrait of a population that was neither enthusiastic nor resistant, but simply quiet.^[Unless otherwise noted, the following scale conventions apply throughout: gate participation items are reported as the percentage of respondents who ever engaged in each activity (binary: ever vs. never); community leader contact is reported as a mean on a 1–5 scale (1 = never, 5 = often); authoritarian governance preferences on a 1–4 scale (1 = very bad, 4 = very good); democratic assessments on a 1–10 scale; corruption perceptions on a 1–4 scale (1 = hardly anyone, 4 = almost everyone); and media and political interest items on a 1–4 scale (1 = none, 4 = a great deal). Internet news frequency is harmonized to a 1–6 scale (1 = never, 6 = daily or more) across waves with differing original response categories; see the dedicated footnote in the Wave 3 discussion for details. Witnessed corruption is reported as a binary proportion. Voter turnout is self-reported and expressed as a percentage.] Political participation was present but low in intensity: `r round(stats$gate_influential_w2, 1)` percent of respondents reported ever having contacted a person of influence, yet community leader contact averaged just `r round(stats$community_leader_w2, 2)` on a five-point frequency scale—figures that capture a population with prior exposure to civic channels but with limited ongoing engagement.^[Direct contact with elected officials and civil servants uses a broadly-worded 'at any level' framing in Wave 2 that differs from the narrower, problem-motivated framing in Wave 3 and from the Wave 4 and 6 formulations; these items are not used for cross-wave comparisons involving Wave 2 but are reported in the four-wave tables with appropriate documentation.]

```{r table1}
t1 <- readRDS(file.path(tbl_dir, "table1_w2_baseline.rds"))

t1 |>
  dplyr::select(label, w2_fmt) |>
  kbl(
    col.names = c("Variable", "W2 (2008)"),
    align     = "lc",
    caption   = "Wave 2 (2008) Baseline: Political Orientations in Cambodia",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 10,
    full_width = FALSE
  ) |>
  pack_rows("Political Participation", 1, 8) |>
  pack_rows("Authoritarian Preferences (1–4: very bad to very good)", 9, 12) |>
  pack_rows("Democratic Expectations (0–10 scale)", 13, 15) |>
  pack_rows("Corruption", 16, 18) |>
  pack_rows("Media & Political Interest", 19, 22) |>
  footnote(
    general = "Wave 2 N = 1,000. Participation rows 1–6 report the percentage of respondents who ever engaged in each activity (gate question); valid N shown in parentheses. Rows 7–8 report means (community leader contact: 1–5 scale) or proportions (voted). '—' indicates item not fielded in this wave. Authoritarian preferences: 1 = very bad, 4 = very good. Democratic expectations: 0–10. Corruption witnessed: binary proportion. National/local corruption: 1 = not at all, 4 = extremely.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

The authoritarian attitude measures are perhaps the most telling. Single-party rule was rated at `r round(stats$single_party_w2, 2)` on a four-point scale (where 4 represents "very good"), strongman rule at `r round(stats$strongman_w2, 2)`, and military rule at `r round(stats$military_w2, 2)`. These are not the numbers of a population that embraces authoritarianism as an ideal; they are the numbers of a population that has learned to live within it. The slightly elevated tolerance for military rule is consistent with the CPP's own origins in Vietnamese-backed military intervention and the residual security anxieties of a post-genocide society, but even this measure sits below the scale midpoint.

What is most notable about the 2008 data, viewed in retrospect, is how stable everything appears. There are no dramatic spikes, no anomalous values, no signs of the upheaval to come. Cambodia in 2008 was a country in political equilibrium—not a democratic equilibrium, but not one characterized by active authoritarian mobilization either. It was, to extend the metaphor, a kingdom in which the subjects had learned that the king was permanent and had arranged their lives accordingly. The question that the next four years would pose, with startling force, was what would happen when that arrangement was briefly and dramatically disrupted.

### The Awakening: Cambodia in 2012

Something extraordinary happened in Cambodia between 2008 and 2012, and the Wave 3 Asian Barometer data capture its imprint with unusual clarity.^[Unless otherwise noted, the political chronology and electoral data in this section draw on @Un2013-zu; @Um2014-ft; @Strangio2014-vi; and @Noren-Nilsson2015-yi. On the 2013 election results specifically, see @COMFREL-The-Committee-for-Free-and-Fair-Elections-in-Cambodia-2013-kt.]

Across nearly every measure of civic engagement and political aspiration, Cambodian respondents in 2012 registered at or near the highest values in the dataset's history. Contact with influential persons—the gate item with consistent framing across waves—rose from `r round(stats$gate_influential_w2, 1)` percent in 2008 to `r round(stats$gate_influential_w3, 1)` percent. Community leader contact averaged `r round(stats$community_leader_w3, 2)` on a five-point frequency scale, up from `r round(stats$community_leader_w2, 2)`. And the political activities newly tracked beginning in Wave 3—attending demonstrations (`r round(stats$gate_demo_w3, 1)` percent ever) and signing petitions (`r round(stats$gate_petition_w3, 1)` percent ever)—debuted at notable levels consistent with a population in active political motion.

These are not the numbers of a quiescent population. They are the numbers of a population that had been politically activated.

The catalyst was the formation of the Cambodia National Rescue Party in 2012, a merger of the Sam Rainsy Party and the Human Rights Party under the joint leadership of Sam Rainsy and Kem Sokha. The CNRP did something that no previous Cambodian opposition had managed: it unified the fragmented non-CPP political space into a single, credible alternative with a coherent message of democratic reform, anti-corruption, and economic justice. The party drew its energy from an emerging urban youth demographic—the first generation of Cambodians born after the Khmer Rouge, educated in the expanding university system, connected to the wider world through social media, and impatient with the political settlement their parents had accepted.

The CNRP's formation in July 2012 was itself a product of political learning. Previous opposition efforts had foundered on precisely the fragmentation that the CPP exploited: multiple parties splitting the non-CPP vote, personal rivalries preventing coordination, and the absence of a unified organizational infrastructure capable of matching the CPP's commune-level reach. The merger of the SRP and the Human Rights Party, brokered through extended negotiations between Sam Rainsy and Kem Sokha, resolved the coordination problem at a stroke.^[@Noren-Nilsson2015-yi; @Un2013-ss.] The resulting party offered voters something genuinely new in Cambodian politics: a single, unified alternative to the CPP with recognizable leaders, a coherent platform, and—critically—the organizational capacity to mobilize supporters in both urban and peri-urban constituencies.

The mobilization dynamics visible in the 2012 survey data intensified dramatically in the months that followed. The July 2013 national election became the most competitive in Cambodia's post-UNTAC history, driven by a potent combination of youth demographics, social media penetration, and economic grievance. Facebook, which had reached an estimated 1.5 million Cambodian users by 2013, served as both an organizational tool and an alternative information ecosystem that partially circumvented CPP control of broadcast media.^[@Noren-Nilsson2021-is on Fresh News and Cambodia's digital media landscape. On the broader role of social media in authoritarian contexts, see @Morgenbesser2020-kz.] The CNRP's campaign messaging—centered on anti-corruption, land rights, wages for garment workers, and a promise to bring genuine democratic accountability—resonated with a generation of voters who had no memory of the Khmer Rouge and little patience for the stability-first bargain their parents had accepted.^[@Noren-Nilsson2021-sq on youth mobilization and power reproduction in Cambodia's authoritarian system.]

The election results shocked the political establishment. The CNRP won 55 of 123 National Assembly seats, reducing the CPP's margin to its narrowest since 1993. Allegations of systematic electoral fraud triggered months of mass protest in Phnom Penh, with demonstrations regularly drawing crowds estimated in the tens of thousands. For observers and participants alike, the political settlement that had governed Cambodia since the late 1990s appeared to be unraveling.^[@Strangio2014-vi; @Morgenbesser2017-cu.]

The democratic expectation measures tell the same story from a different angle. When asked to rate the democratic future of their country on a ten-point scale, Cambodian respondents in 2012 produced a mean of `r round(stats$dem_future_w3, 2)`. This is, by any standard, a remarkably high figure, indicating near-universal optimism. It suggests a population that had collectively decided that democratic change was not merely desirable but imminent—that the political order they had lived under for decades was on the verge of fundamental transformation.

The paradox, of course, is that this expectation was not based on any structural change in the regime. The CPP still controlled the state apparatus, the military, the courts, the media, and the vast patronage networks that sustained its rural base. Hun Sen showed no inclination to reform or retire. What had changed was not the regime but the perception of the regime's permanence—and that perceptual shift, as the subsequent waves would demonstrate, proved far more fragile than the political moment suggested.

The specific configuration of the democratic assessment triad in 2012 is theoretically revealing. The past-present-future structure—low past (`r round(stats$dem_past_w3, 2)`), middling present (`r round(stats$dem_present_w3, 2)`), exceptionally high future (`r round(stats$dem_future_w3, 2)`)—is not the signature of a population evaluating its democratic institutions on the basis of lived experience. A population grounded in experiential assessment would be expected to rate the future as a modest extrapolation from the present, not as a near-perfect score that exceeds the present by nearly four points. What the 2012 configuration captures instead is a hope-driven democratic orientation: the conviction that the current political order is transitional and that the democratic destination is both certain and imminent. This orientation, as subsequent waves would reveal, proved extraordinarily sensitive to political events—precisely because it was grounded in expectation rather than experience.

Even the corruption data reflect the mobilization moment. Witnessed corruption stood at `r round(stats$corrupt_witnessed_w3, 1)` percent in Wave 3, up from `r round(stats$corrupt_witnessed_w2, 1)` percent in Wave 2. This increase is almost certainly not an increase in actual corruption—the CPP's patronage system was well-established long before 2012—but rather reflects a population that had become more willing to identify and report corruption as a political problem, emboldened by the opposition's anti-corruption messaging and the expanded civic space that accompanied the CNRP's rise.

There are also subtler patterns in the Wave 3 data that resist easy interpretation. Traditional values measures—obedience to parents, deference to teacher authority—also peaked in this wave, an unexpected finding given that the broader mobilization was associated with youth activism and challenges to established authority. One possibility is that the CNRP's own messaging, which blended democratic aspiration with Khmer cultural conservatism and Buddhist moral frameworks, activated traditional values alongside rather than in opposition to democratic ones. Another is that the elevated values reflect a general intensification of political and social engagement in which all expressed attitudes, whether progressive or traditional, registered at higher levels. A third possibility, which cannot be ruled out, is survey context effects: the charged political environment of 2012 may have produced a general social desirability bias toward expressing strong opinions of any kind.

The media and political engagement measures complete the portrait of a mobilized citizenry. Political interest stood at `r round(stats$pol_interest_w3, 2)` on a four-point scale, and following political news registered at `r round(stats$pol_news_w3, 2)`—both the highest values in the Cambodian ABS series. Political discussion, at `r round(stats$pol_discuss_w3, 2)`, remained modest, suggesting that the activation captured in 2012 was more a matter of individual attention and institutional engagement than of deliberative civic culture. Internet news consumption, harmonized to a common six-point frequency scale across waves, stood at `r round(stats$news_internet_w3, 2)`—in the "never to hardly ever" range, consistent with the still very limited internet penetration of Cambodia in 2012.^[The ABS internet news item underwent substantial changes in response scale across waves (6-point in W2/W3, 8-point in W4, 9-point in W5/W6). The values reported here reflect harmonization to a common 1–6 frequency scale (1 = never, 6 = daily or more). The coarser W2/W3 scale cannot distinguish within-daily usage patterns available in later waves, but the harmonized values are comparable at the level of weekly, monthly, and annual frequency.]

The 2012 data, taken together, present a portrait of a political community at its moment of maximum activation. Every dimension the Asian Barometer measures—participation, aspiration, attention, even the willingness to name corruption—points in the same direction. It is tempting, with the knowledge of what came after, to read this as the high-water mark of Cambodian democracy, the moment before the tide turned. But that reading, while emotionally satisfying, may be too simple. What the Wave 3 data actually capture is something rarer and more theoretically interesting: the approximate moment at which a population collectively updated its beliefs about what was politically possible, right before the regime demonstrated, with devastating effectiveness, that those beliefs were wrong.

### The Reckoning: Cambodia in 2015

That demonstration began at the ballot box and ended in the streets.^[Unless otherwise noted, the political chronology in this section draws on @Chheang2015-ud; @Morgenbesser2017-cu; and @Noren-Nilsson2019-ck.] The 2013 national election, held just months after the Wave 3 survey captured such exceptional optimism, confirmed the CNRP’s mobilization power: the opposition won 55 seats to the CPP’s 68, the narrowest margin in the country's modern history. Yet, the structural ceiling of the dominant-party state held firm. Amid widespread allegations of electoral fraud, the CNRP boycotted parliament, triggering months of mass demonstrations that soon merged with massive garment worker strikes. For a brief window, genuine political transformation appeared imminent. It did not arrive. Following a lethal state crackdown on protesters in January 2014, and a subsequent political compromise that coaxed the opposition into the National Assembly under a fragile and temporary "Culture of Dialogue," the momentum of the streets fractured.

The "Culture of Dialogue" that emerged from the July 2014 agreement between Sam Rainsy and Hun Sen was, in retrospect, less a genuine political opening than a controlled decompression. The CNRP entered parliament and gained access to committee chairmanships, lending the arrangement a veneer of power-sharing. But the underlying dynamics had not changed: the CPP retained control of the security forces, the judiciary, and the National Election Committee, while the CNRP's capacity for mass mobilization—its primary source of leverage—was progressively constrained through targeted legal harassment of its leaders and selective restrictions on public assembly.^[@Morgenbesser2017-cu; @Noren-Nilsson2019-ck. On the broader pattern of managed political openings in dominant-party systems, see @Schedler2013-qn.] By 2015, the dialogue had largely stalled, and the political atmosphere had shifted from the hopeful if chaotic energy of 2013 to something more wary and uncertain. It was in this atmosphere that the Wave 4 Asian Barometer survey entered the field.

The Wave 4 data from 2015 capture a population in the midst of this reckoning—no longer at the peak of mobilization but not yet in the trough of withdrawal. The behavioral participation measures tell a more ambiguous story than might be expected. Contact with influential persons held essentially steady at `r round(stats$gate_influential_w4, 1)` percent, and demonstration attendance, petition-signing, and civil servant contact all edged slightly upward from their Wave 3 levels. Community leader contact averaged `r round(stats$community_leader_w4, 2)` on a five-point frequency scale, essentially unchanged from 2012. The "selective retreat" of 2015 is not, in other words, a story written primarily in the behavioral data. It is written instead in the attitudinal record.

```{r table2}
t2 <- readRDS(file.path(tbl_dir, "table2_w3w4_comparison.rds"))

t2 |>
  dplyr::select(label, w3_fmt, w4_fmt, delta_fmt) |>
  kbl(
    col.names = c("Variable", "W3 (2012)", "W4 (2015)", "\\Delta (W4\u2212W3)"),
    align     = "lccc",
    caption   = "Selective Retreat, 2012--2015: Wave 3 to Wave 4 Comparison",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 10,
    full_width = FALSE
  ) |>
  pack_rows("Political Participation", 1, 8) |>
  pack_rows("Authoritarian Preferences (1–4)", 9, 12) |>
  pack_rows("Democratic Expectations (0–10)", 13, 15) |>
  pack_rows("Corruption", 16, 18) |>
  pack_rows("Media & Political Interest", 19, 22) |>
  footnote(
    general = "W3 N = 1,200; W4 N = 1,200. Participation rows 1–6 report percentages (gate question); valid N shown in parentheses. $\\\\Delta$ for participation rows is in percentage points (pp); for other rows, in scale units. '—' indicates item not fielded. Internet news harmonized to 1–6 frequency scale (1 = never, 6 = daily or more).",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE,
    escape = FALSE
  )
```

This attitudinal disengagement was not yet accompanied by a corresponding rise in authoritarian acceptance. Single-party rule actually dipped slightly from `r round(stats$single_party_w3, 2)` to `r round(stats$single_party_w4, 2)`, and military rule fell from `r round(stats$military_w3, 2)` to `r round(stats$military_w4, 2)`. Expert rule and strongman rule edged up marginally but remained below 2.0. This pattern is analytically important: it suggests that the demobilization observed between 2012 and 2015 was driven more by disillusionment and tactical withdrawal than by a genuine reorientation toward authoritarian preferences. Cambodians were pulling back from political action, but they had not yet accepted the authoritarian order as normatively appropriate.

The democratic expectation measures, however, had already begun their descent. The democratic future rating fell from `r round(stats$dem_future_w3, 2)` to `r round(stats$dem_future_w4, 2)`—a drop of nearly two full points on the ten-point scale. The democratic present, as assessed through government performance, declined from `r round(stats$dem_present_w3, 2)` to `r round(stats$dem_present_w4, 2)`. Only the democratic past held relatively steady at `r round(stats$dem_past_w4, 2)`. What this configuration suggests is that the hope that had animated the 2012 moment was already eroding, not because Cambodians had embraced an alternative vision, but because the events of 2013-2014—the election that changed nothing, the protests that were suppressed, the deal that left the CPP in power—had begun to teach a lesson about the limits of popular mobilization within a dominant-party system.

The corruption data from Wave 4 introduce a striking counterpoint. Witnessed corruption surged to its highest level in the dataset: `r round(stats$corrupt_witnessed_w4, 1)` percent of respondents reported having personally encountered corruption. This is unlikely to reflect merely increased willingness to report; the political environment of 2015 was, if anything, less permissive of anti-government speech than 2012. The more likely explanation is that the post-2013 period, with its political bargaining, institutional jockeying, and expanded CNRP presence in parliament, created more visible sites of corruption—more interactions between citizens and competing political actors, more opportunities to observe the transactional nature of Cambodian governance up close.

The informational dimensions of political life registered the earliest signs of the withdrawal that would accelerate dramatically by 2021. Political interest declined from `r round(stats$pol_interest_w3, 2)` to `r round(stats$pol_interest_w4, 2)`, and following political news fell from `r round(stats$pol_news_w3, 2)` to `r round(stats$pol_news_w4, 2)`—modest drops in absolute terms, but significant as the leading edge of a trend that would deepen sharply. Political discussion held essentially flat at `r round(stats$pol_discuss_w4, 2)`, suggesting that Cambodians were beginning to disengage from formal political information channels while maintaining informal conversational patterns. Internet news consumption, meanwhile, continued its modest climb from `r round(stats$news_internet_w3, 2)` to `r round(stats$news_internet_w4, 2)`, reflecting the early stages of broadening internet access even as political engagement with that access was beginning to wane.

The 2015 wave, viewed within the larger narrative, represents the hinge point—the moment when the trajectory could still have gone either way. The CNRP was weakened but present, its leaders alternately accommodated and harassed, its supporters disillusioned but not yet defeated. The data register this ambiguity: democratic hope was falling but had not crashed. Behavioral participation remained largely intact while attitudinal engagement had already begun its descent. Authoritarian acceptance had not yet risen. Corruption was more visible than ever. It was a moment of unstable equilibrium, poised between the democratic aspiration of 2012 and the authoritarian normalization that was about to descend.

### The Silence: Cambodia in 2021

On November 16, 2017, the Supreme Court of Cambodia dissolved the Cambodia National Rescue Party.^[Unless otherwise noted, the political chronology in this section draws on @Un2020-dq; @Chheang2021-td; and @Morgenbesser2019-mb.] The ruling, which cited a vaguely defined conspiracy to overthrow the government, banned 118 CNRP officials from political activity for five years and transferred the party's parliamentary seats to smaller, CPP-aligned parties. Kem Sokha had already been arrested in September; Sam Rainsy remained in exile. In the July 2018 national election, the CPP won all 125 seats in the National Assembly. Cambodia became, for all practical purposes, a one-party state.

The CNRP's dissolution did not occur in isolation. It was the centerpiece of a broader campaign of political closure that systematically dismantled the infrastructure of independent civic life. The *Cambodia Daily*, the country's most prominent English-language newspaper and a persistent thorn in the CPP's side, was forced to close in September 2017 after receiving a disputed tax bill of $6.3 million.^[The Cambodia Daily's closure was widely interpreted as politically motivated. See @Noren-Nilsson2021-is on the restructuring of Cambodia's media landscape during this period.] Radio Free Asia's Cambodian bureau was shuttered, and several of its journalists faced criminal charges. Independent radio stations that had broadcast opposition content were pressured to switch to pro-government programming or cease operations entirely. Civil society organizations faced intensified regulatory scrutiny under the 2015 Law on Associations and Non-Governmental Organizations (LANGO), which gave the government broad discretionary power to dissolve organizations deemed to threaten public order.^[On LANGO and its effects on civil society space, see @Noren-Nilsson2020-ev.]

The combined effect was the elimination not merely of the political opposition but of the broader ecosystem—media, civil society, informal networks—through which citizens had accessed alternative political information and organized collective action. The transition, in Levitsky and Way's typology, was from competitive authoritarianism, in which opposition parties exist and can occasionally win, to hegemonic authoritarianism, in which the electoral arena is formally maintained but substantively emptied.^[@Levitsky2010-bc; @Schedler2002-eg. @Morgenbesser2019-mb applies this framework specifically to Cambodia's post-2017 trajectory.] The 2018 national election, conducted without the CNRP on the ballot, produced a CPP sweep of all 125 National Assembly seats—a result that formalized what the dissolution had already accomplished.

The Wave 6 data from 2021 record the consequences.

The participation data from 2021 present a more differentiated picture than a simple collapse narrative would suggest. Contact with influential persons—the item most directly measuring independent civic engagement—fell sharply, from `r round(stats$gate_influential_w4, 1)` percent in 2015 to `r round(stats$gate_influential_w6, 1)` percent, representing the steepest single-item decline in the dataset. Community leader contact fell to a mean of `r round(stats$community_leader_w6, 2)` on the frequency scale, and political discussion dropped to `r round(stats$pol_discuss_w6, 2)` on a four-point scale—a population that had largely stopped talking about politics with one another. Demonstration attendance held near its 2015 level at `r round(stats$gate_demo_w6, 1)` percent.

What complicates any uniform "participation collapse" characterization is that several formal channel indicators actually rose substantially by 2021: petition-signing reached `r round(stats$gate_petition_w6, 1)` percent, civil servant contact climbed to `r round(stats$gate_civil_w6, 1)` percent, and media contact rose to `r round(stats$gate_media_w6, 1)` percent. These increases likely reflect two overlapping dynamics: the rapid expansion of internet access enabling digital petitions and online contact, and the transformation of formal institutional channels into vehicles for routine patronage interaction under a now-uncontested regime. Participation through officially sanctioned channels rises precisely as independent civic action—contacting influential figures outside the state apparatus, organizing demonstrations, discussing politics—contracts. The divergence between these two trajectories is itself evidence of demobilization by subtraction: what citizens can still do expands; what citizens choose to do in pursuit of genuine political alternatives collapses.

```{r table3}
t3 <- readRDS(file.path(tbl_dir, "table3_four_wave_trajectory.rds"))

t3 |>
  dplyr::select(label, w2_fmt, w3_fmt, w4_fmt, w6_fmt, delta_fmt) |>
  kbl(
    col.names = c("Variable", "W2 (2008)", "W3 (2012)", "W4 (2015)",
                  "\\textbf{W6 (2021)}", "\\Delta W3{\\textrightarrow}W6"),
    align     = "lccccc",
    caption   = "Four-Wave Trajectory of Political Orientations in Cambodia, 2008--2021",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 10,
    full_width = FALSE
  ) |>
  column_spec(5, bold = TRUE) |>
  pack_rows("Political Participation", 1, 8) |>
  pack_rows("Authoritarian Preferences (1–4: very bad to very good)", 9, 12) |>
  pack_rows("Democratic Expectations (0–10 scale)", 13, 15) |>
  pack_rows("Corruption", 16, 18) |>
  pack_rows("Media & Political Interest", 19, 22) |>
  footnote(
    general = "N per wave: W2 = 1,000; W3 = 1,200; W4 = 1,200; W6 = 1,242. Rows 1–6 (Political Participation) report the percentage of respondents who ever engaged in each activity (gate proportion), with per-cell valid N in parentheses; $\\\\Delta$ for these rows is in percentage points. Rows 7–8 and all remaining rows report unweighted wave means with per-cell valid N in parentheses; $\\\\Delta$ is the arithmetic difference. '---' indicates variable not fielded in that wave. $\\\\Delta$ W3$\\\\rightarrow$W6 = W6 value minus W3 value (peak-to-trough change). W6 column in bold marks the endpoint of the trajectory. Voted last election and corruption witnessed are binary proportions.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE,
    escape = FALSE
  )
```

```{r figure1, fig.cap="Political Orientations in Cambodia by Domain, 2008--2021. Wave means for selected indicators across five thematic domains. Dashed vertical line marks 2017 CNRP dissolution. Y-axes are free-scaled; each panel uses its own metric range. Source: Asian Barometer Survey, Waves 2, 3, 4, 6.", out.width="100%", fig.pos="!htbp"}
knitr::include_graphics(file.path(fig_dir, "fig1_trend_panels.pdf"))
```

The authoritarian acceptance measures moved in mirror image. All four alternative governance indicators rose to their highest levels: expert rule to `r round(stats$expert_rule_w6, 2)`, single-party rule to `r round(stats$single_party_w6, 2)`, strongman rule to `r round(stats$strongman_w6, 2)`, and military rule to `r round(stats$military_w6, 2)`. The jump was concentrated entirely in the W4-to-W6 period—the six years that encompassed the CNRP's destruction and the subsequent one-party election. What had been stable or declining through 2015 now surged.

There is an important interpretive question here, and the article's argument depends on how it is answered. Does the rise in authoritarian acceptance represent genuine preference change—Cambodians coming to believe that authoritarian governance is normatively desirable—or does it represent something else? Three considerations suggest the latter. First, the magnitude of the shift, while statistically meaningful, is modest in absolute terms: all four measures remain around 2.2 on a four-point scale, below the midpoint. Cambodians are not rating authoritarianism as "good"; they are rating it as slightly less bad than before. Second, the timing—concentrated after the removal of the democratic alternative—suggests a mechanism closer to adaptive preference formation than genuine conversion. Third, and most importantly, the shift co-occurs with the collapse of democratic future expectations (from `r round(stats$dem_future_w4, 2)` to `r round(stats$dem_future_w6, 2)`), rising perceptions of government information withholding, and declining political interest. This is not the attitudinal profile of a population that has found a new political faith; it is the profile of a population that has stopped believing the old one was achievable.

The article proposes the term authoritarian normalization to describe this process, distinguishing it from authoritarian legitimation. Legitimation implies that citizens have come to view the regime as rightfully entitled to rule, whether on the basis of performance, ideology, or tradition. Normalization implies something weaker but potentially more durable: citizens have come to view the regime as the only available reality and have adjusted their expressed preferences to align with that reality. The distinction matters theoretically because it suggests different mechanisms of regime stability—and different vulnerabilities. A regime sustained by legitimation is threatened when performance falters or ideology loses appeal. A regime sustained by normalization is threatened when alternatives re-enter the political imagination.

The normalization framework finds a useful analogue in Wedeen's analysis of political compliance in Hafez al-Assad's Syria, where citizens participated in elaborate public rituals of regime praise not because they believed the regime's claims but because acting as if they believed was the condition of ordinary life.^[@Wedeen1998-qz; @Wedeen2000-bg. Wedeen's concept of "acting as if" captures the performative dimension of political compliance under conditions where genuine belief is neither required nor expected by the regime.] The Cambodian case differs in important respects—the ABS data capture private survey responses rather than public performances, and the CPP's demands on citizens are less theatrically extravagant than the Assad cult of personality. But the underlying logic is similar: when the political field is restructured so that only one arrangement appears viable, citizens adjust their expressed orientations to accommodate that arrangement, not out of conviction but out of a practical reckoning with the available reality. The survey data cannot definitively distinguish between genuine preference change and adaptive accommodation, but the pattern of evidence—modest rather than dramatic shifts in authoritarian acceptance, coinciding with collapsing democratic expectations and civic withdrawal—is more consistent with normalization than with conversion.

The corruption data offer perhaps the most unsettling evidence of normalization. Witnessed corruption collapsed from `r round(stats$corrupt_witnessed_w4, 1)` percent in 2015 to `r round(stats$corrupt_witnessed_w6, 1)` percent in 2021. Perceived national government corruption also declined, from `r round(stats$corrupt_national_w4, 2)` to `r round(stats$corrupt_national_w6, 2)` on the four-point scale. On the surface, these numbers might suggest genuine anti-corruption progress. The CPP did conduct selective anti-corruption campaigns during this period, targeting rivals and politically inconvenient officials. But the magnitude of the witnessed-corruption decline—from nearly two-thirds to one-seventh of the population—is too large to explain through policy alone.

The more parsimonious explanation is that the corruption decline reflects the same demobilization process visible in the participation data. When citizens withdraw from political life—when they stop contacting officials, stop attending meetings, stop engaging with governance institutions—they also reduce their exposure to the sites where corruption is experienced and observed. A population that does not petition, does not demonstrate, and does not contact civil servants is also a population that has fewer occasions to witness a bribe demanded or a favor exchanged. Demobilization, in this reading, does not merely reduce political action; it reduces the perceptual evidence that political action might address. The regime becomes less corrupt not because it has reformed but because citizens have stopped looking.

An alternative, and not mutually exclusive, explanation for the corruption decline invokes the survey environment itself. In a political climate where the opposition has been criminalized and independent media shuttered, respondents may reasonably calculate that acknowledging witnessed corruption—even in a nominally confidential survey—carries risks of self-incrimination or unwanted official attention. The social desirability bias literature has long recognized that survey responses on sensitive topics shift in predictable directions under conditions of political constraint, with respondents offering answers that align with perceived official expectations.^[On the reliability of survey responses under electoral authoritarianism, see @Schedler2007-pv. @Frye2016-zm demonstrate that even in semi-competitive authoritarian contexts, survey-measured approval can closely track genuine opinion—but their findings also imply that as political space closes further, the gap between expressed and private attitudes is likely to widen.] The post-2017 Cambodian environment, in which public criticism of the government carried tangible legal consequences, represents precisely the conditions under which such bias would be expected to intensify. The two explanations—reduced exposure through demobilization and increased reticence through social desirability pressure—likely operate in tandem, each reinforcing the other to produce the dramatic decline observed in the data.

The voter turnout paradox completes the portrait. Despite the collapse of all other forms of political engagement, self-reported voting in the last election rose from `r round(stats$voted_w3, 1)` percent in 2012 to `r round(stats$voted_w4, 1)` percent in 2015 to `r round(stats$voted_w6, 1)` percent in 2021. This pattern—rising electoral participation amid collapsing civic engagement—is a well-documented feature of hegemonic authoritarian systems, where elections serve not as mechanisms of political choice but as rituals of regime affirmation. The 2018 election made the ritual character of Cambodian voting unusually visible. The CNRP-in-exile’s “clean finger” campaign—urging voters to boycott and display uninked fingers as proof of refusal—transformed the indelible ink applied at polling stations from an anti-fraud measure into a tool of social surveillance. Factory workers were warned that returning to work without inked fingers would result in wage deductions; village heads distributed polling slips with the expectation that residents would comply; government employees were required to participate.^[@McCargo2018-ms provides a detailed firsthand account of the turnout-maximization dynamics surrounding the 2018 election.] Voting in a one-party election is not a political act in the same sense as voting in a competitive one; it is closer to a demonstration of compliance, a public performance of belonging to the political community as the regime defines it. The fact that turnout rises precisely as meaningful engagement collapses suggests that Cambodians understand the distinction even as they perform the ritual.

The international orientation data add a final dimension to the normalization portrait. Cambodian assessments of Chinese influence in the region grew more positive between Waves 4 and 6, rising from `r round(stats$china_asia_goodharm_w4, 2)` to `r round(stats$china_asia_goodharm_w6, 2)` on a four-point scale measuring perceived benefit versus harm. More striking still, expectations of future Asian—implicitly Chinese—regional influence climbed from `r round(stats$future_influence_asia_w4, 2)` in Wave 4 to `r round(stats$future_influence_asia_w6, 2)` in Wave 6, a trajectory that places Cambodia among the most China-positive publics in the ABS dataset. This is consistent with Cambodia's deepening alignment with Beijing during this period, visible in Belt and Road infrastructure investment, the Ream Naval Base controversy, and Cambodia's consistent support for Chinese positions within ASEAN. But it also fits the normalization logic: as democratic futures recede and Western-aligned political alternatives are eliminated, alignment with the ascendant authoritarian power in the region becomes the natural orientation of a population adjusting to the available political reality. A paired comparison with Vietnam or the Philippines—where territorial disputes drive China perceptions in sharply negative directions despite similar economic interdependence—would sharpen this point considerably.^[@Strangio2020-gw provides the most comprehensive account of Cambodia's positioning within China's Southeast Asian strategy. On the divergent China perception trajectories across ASEAN, the ABS cross-national data offer direct comparative evidence.]

The informational withdrawal that began tentatively in 2015 reached its full expression by 2021. Following political news fell to `r round(stats$pol_news_w6, 2)`, political interest declined to `r round(stats$pol_interest_w6, 2)`, and political discussion dropped to `r round(stats$pol_discuss_w6, 2)`—the lowest values recorded across all four waves. These figures describe a population that has not merely withdrawn from political action but has largely ceased to attend to political information at all. The internet news variable throws this withdrawal into particularly sharp relief: harmonized internet news consumption rose to `r round(stats$news_internet_w6, 2)` by Wave 6—approaching "at least once a week"—reflecting the rapid expansion of internet infrastructure and smartphone penetration across Cambodia. Cambodians in 2021 had far greater access to online information than at any previous survey wave, yet their political interest, news following, and willingness to discuss politics all fell to historic lows. The infrastructure for engagement had expanded; the will to use it for political purposes had collapsed. The self-reinforcing logic is straightforward: citizens who do not participate have less reason to follow political developments; citizens who do not follow political developments have less basis on which to participate. Demobilization, once initiated through the removal of the political alternative, generates its own sustaining momentum through this informational feedback loop—a dynamic that suggests the effects of the CNRP's dissolution may prove more durable than the act itself.

The Cambodia of 2021 is, in a sense, the inverse of the Cambodia of 2012. Where the earlier moment was characterized by activation across all dimensions—participation, aspiration, attention, even the willingness to name corruption—the later moment is characterized by withdrawal across all the same dimensions. The symmetry is striking and, for the theoretical argument, essential. It suggests that what changed between the two moments was not any single attitude or behavior but the underlying orientation from which attitudes and behaviors derive—what might be called the perceived horizon of political possibility. When that horizon expanded, everything expanded with it. When it contracted, everything contracted.

### The Education

The story told in the preceding pages is, in one sense, particular to Cambodia—a product of specific historical circumstances, a specific regime, a specific opposition movement, and a specific act of political closure. In another sense, it is a story with much wider application, because the mechanism it illustrates—the restructuring of citizen orientations through the removal of political alternatives—operates wherever dominant-party regimes consolidate power. This concluding section draws out four theoretical implications of the Cambodian sequence and considers their relevance beyond the single case.

The first concerns the mechanism itself. Demobilization by subtraction, as the Cambodian data illustrate it, is distinct from the two processes most commonly invoked to explain political quiescence. It is not repression-driven demobilization: while the CPP certainly employed coercion, the attitudinal shifts observed in the ABS data—rising authoritarian acceptance, collapsing democratic expectations, declining political interest—are not the signature of a frightened population. A population demobilized primarily through fear would be expected to maintain private democratic commitments even as public participation declined; the Cambodian data show both dimensions moving in the same direction.^[This pattern is consistent with the distinction between "preference falsification," in which private attitudes diverge from public behavior, and genuine preference adjustment. The Cambodian case appears closer to the latter, though survey data alone cannot fully resolve the question. On preference falsification under authoritarianism, see @Kuran1995-up.] Nor is it co-optation: the CPP's patronage networks were already well-established before 2012, and there is no evidence of a dramatic expansion of distributive benefits between 2015 and 2021 that could account for the attitudinal shifts. What changed was not the cost of participation or the rewards of compliance but the perceived availability of an alternative—the object around which civic mobilization had organized. When that object was removed, the entire architecture of engagement collapsed with it.

The second implication concerns the relationship between hope and democratic attitudes. The Wave 3 data suggest that democratic support in authoritarian contexts may be driven less by abstract normative commitments than by perceived feasibility. When democracy appeared possible in 2012, every democratic indicator in the ABS surged—not only future expectations but participation, political interest, and even the willingness to name corruption. When democracy appeared impossible by 2021, those same indicators collapsed. This pattern challenges a foundational assumption in the democratic values literature: that survey-measured democratic attitudes reflect stable, deeply held orientations that exist independently of political circumstances.^[The canonical formulation of democratic values as a stable cultural attribute is found in the political culture tradition from @Inglehart1997-gw onward. The Cambodian data suggest a more situationally contingent model, closer to what Bratton and Mattes have termed "demand for democracy" as a function of perceived supply [@Bratton2004-ke; @Mattes2007-ej].] The Cambodian evidence suggests instead that measured democratic attitudes may be substantially a function of perceived democratic possibility—that hope is not merely the emotional accompaniment of democratic aspiration but its primary cognitive driver. If this is correct, cross-national surveys that measure democratic support without accounting for the structure of political alternatives available to respondents may be capturing something quite different from what they intend.

The third implication concerns corruption measurement. The finding that witnessed corruption declined precipitously alongside civic participation suggests a troubling methodological artifact: survey-based corruption indicators may systematically understate corruption in demobilized populations. When citizens withdraw from engagement with state institutions, they also withdraw from the sites where corruption is experienced and observed. A regime that successfully demobilizes its population will, by this logic, appear less corrupt on standard survey measures even if actual corruption is unchanged or increasing. This has practical consequences for the evaluation of anti-corruption campaigns in authoritarian contexts. The CPP’s selective anti-corruption efforts during this period may well have had some genuine effect, but the dramatic scale of the decline—from `r round(stats$corrupt_witnessed_w3, 0)` percent to `r round(stats$corrupt_witnessed_w6, 0)` percent—is far more parsimoniously explained by the withdrawal of citizens from the institutional encounters where corruption becomes visible. Transparency International’s Corruption Perception Index and similar instruments may be particularly vulnerable to this dynamic in consolidating authoritarian regimes.

The fourth implication is diagnostic. The divergence between electoral participation and civic participation—turnout rising to `r round(stats$voted_w6, 0)` percent even as independent civic engagement collapsed—may serve as a generalizable indicator of authoritarian normalization. Where elections are maintained as regime rituals rather than competitive mechanisms, rising turnout signals not democratic health but its opposite: the transformation of voting from an act of political choice into an act of political compliance.^[On the ritual function of elections in hegemonic authoritarian systems, see @Schedler2002-eg; @Magaloni2010-wl.] The Cambodian case suggests that this turnout-participation divergence emerges specifically in the wake of opposition elimination, and that its appearance in other contexts—wherever rising turnout coincides with collapsing civic engagement—should be read as evidence that the electoral arena has been substantively emptied even if it remains formally intact.

These findings carry inherent limitations. Four survey waves across thirteen years provide a suggestive but not definitive empirical foundation; the absence of Waves 1 and 5 from the Cambodian ABS series limits the precision with which structural breaks can be identified. The Wave 3 peak, while consistent with the CNRP mobilization narrative, may partly reflect survey timing effects rather than durable attitudinal change. The Wave 6 data, collected during the COVID-19 pandemic, are subject to the concern that mobility restrictions and heightened state authority may have independently depressed participation measures and elevated compliance orientations. The analysis has addressed this concern by noting that the key attitudinal shifts—particularly the decline in democratic future expectations—began in Wave 4, well before the pandemic. But the possibility of pandemic amplification effects in the 2021 data cannot be entirely excluded. Finally, any single-country study faces the inherent limitation of generalizability: the mechanisms identified here may be specific to Cambodia’s particular configuration of post-conflict politics, dominant-party rule, and opposition dynamics.

The most productive response to this limitation is not defensive qualification but comparative extension. The demobilization sequence observed in Cambodia—civic participation collapse, followed by rising authoritarian acceptance, followed by declining democratic aspiration—generates a testable temporal ordering that can be examined in other cases of opposition elimination. Turkey after the suppression of the Peoples’ Democratic Party (HDP), Russia following the banning of Navalny’s organizational network, and Thailand in the years following the 2014 coup all represent contexts where credible political alternatives were systematically removed from the political field. The Asian Barometer dataset itself offers the most direct comparative leverage: Thailand’s post-coup trajectory and the Philippines under Duterte provide cases where similar dynamics may be observable using identical survey instruments. Future research could test whether the Cambodian sequence represents a general pattern of authoritarian normalization or a configuration specific to the conditions of this particular case.

This article began with a number: `r round(stats$dem_future_w3, 1)` out of 10. It is worth ending with the question of what that number meant. In 2012, Cambodian citizens looked at their country—a dominant-party authoritarian state with controlled media, a compromised judiciary, and a history of political violence—and saw a democratic future so bright they could barely imagine anything brighter. Nine years later, they had learned otherwise. The democratic future had not arrived. The party that promised it had been dissolved. The leaders who championed it were in prison or exile. And the population, having briefly imagined a different political life, had returned to the quiet arrangement their parents had known—participating in elections that offered no choice, accepting governance structures they had once rejected, and gradually, measurably, ceasing to pay attention.

Any reliance on longitudinal survey data carries inherent vulnerabilities, and this analysis of Cambodia’s trajectory is no exception. The dramatic peaks of 2012 and the stark valleys of 2021 are undeniably colored by the unique atmospheric conditions of those exact moments.[^2] Yet, the theoretical weight of this study does not rest on the absolute value of any single data point, but on the architecture of the sequence. By anchoring this narrative in within-case temporal variation—observing the same population across four distinct acts of a remarkably well-documented political closure—the noise of individual survey waves gives way to a clear, structural signal. It provides an unusually clean empirical window into how a population recalibrates its political reality when the horizon of possibility collapses.

[^2]: Two contextual anomalies warrant specific acknowledgment. First, the Wave 3 (2012) survey was fielded during an unprecedented window of opposition mobilization just prior to the 2013 elections, meaning the data likely capture a temporary zenith of democratic optimism rather than a stable baseline. Second, the Wave 6 (2021) data were collected during the COVID-19 pandemic, a period where public health restrictions inherently depressed physical civic participation (e.g., attending demonstrations) and may have temporarily inflated reliance on state authority. However, the fact that ideological demobilization (such as the drop in democratic future expectations) began in Wave 4, well before the pandemic, suggests the 2021 results reflect deeper authoritarian normalization rather than mere epidemiological artifacts. On the methodological challenges of surveying under authoritarian conditions more generally, see @Schedler2007-pv. Furthermore, the single-country focus of this study naturally bounds its immediate generalizability, though the mechanisms observed here offer a framework for testing similar dynamics in other consolidating hegemonic regimes.

The fairy tale has no happy ending. But it has a lesson, and the lesson is not about Cambodia alone. It is about the fragility of democratic aspiration in the absence of democratic infrastructure, and about the remarkable efficiency with which authoritarian regimes can reshape citizen orientations simply by removing the object around which those orientations had formed. The education of Cambodia's citizenry between 2008 and 2021 was not accomplished through indoctrination or ideology. It was accomplished through subtraction—the removal of an alternative, and the long silence that followed.

\newpage

# References {.unnumbered}

::: {#refs}
:::