---
title: "Online Appendix"
subtitle: "EconDev-AuthPref"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
date: today
bibliography: refs.bib
csl: ../../apa.csl
number-sections: true
---

# Additional Descriptive Statistics {#sec-app-descriptives}

Additional descriptive statistics and data exploration.

```{r}
#| label: app-descriptives
#| echo: false

# Load data
source("../../src/r/utils/load_data.R")
data <- read_rds("../../data/processed/survey_data.rds")

# Additional descriptive tables
```

# Robustness Checks {#sec-app-robustness}

## Alternative Model Specifications

```{r}
#| label: app-alt-models
#| echo: false

# Alternative specifications
```

## Sensitivity Analyses

```{r}
#| label: app-sensitivity
#| echo: false

# Sensitivity analyses
```

# Survey Instrument {#sec-app-instrument}

## Questionnaire Items

Include key questionnaire items used in the analysis.

**Dependent Variable:**

- Item text here
- Response scale: [scale description]

**Independent Variables:**

- Item 1 text
- Item 2 text

## Variable Construction

Describe how variables were constructed from raw survey items.

# Additional Figures and Tables {#sec-app-additional}

```{r}
#| label: app-fig-distribution
#| fig-cap: "Distribution of Key Variables"
#| echo: false

# Distribution plots
```

```{r}
#| label: app-tbl-correlations
#| tbl-cap: "Correlation Matrix"
#| echo: false

# Correlation matrix
```

# References {.unnumbered}

::: {#refs}
:::

