---
title: "The Education of a Citizenry: Hope, Demobilization, and Authoritarian Normalization in Cambodia, 2008–2021"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: now
date-format: "MMMM D, YYYY [at] h:mm A"
abstract: |
  Between 2008 and 2021, Cambodian citizens underwent one of the most dramatic transformations in political orientation recorded in contemporary survey research. Drawing on four waves of Asian Barometer Survey data (Waves 2, 3, 4, and 6; N ≈ 1,000–1,242 per wave), supplemented by the 2017 Cambodia Socio-Economic Survey (N = 3,363), this article traces the arc of that transformation as a sequential process: from post-conflict acquiescence, through a remarkable peak of democratic aspiration and civic mobilization, to a final condition of political withdrawal and regime normalization following the elimination of organized opposition. The Cambodian case offers an unusually clean empirical window into how dominant-party authoritarian regimes produce political quiescence---not through the cultivation of enthusiastic support, but through a process the article terms *demobilization by subtraction*. The forced dissolution of the Cambodia National Rescue Party in 2017 was the keystone of a broader campaign of political closure---encompassing independent media shutdowns, civil society restrictions, and an uncontested national election---that collectively restructured citizen orientations across multiple attitudinal and behavioral dimensions simultaneously. The article documents a theoretically consequential divergence: abstract democratic commitment---the belief that democracy is the best form of government---remained largely intact across the study period, while democratic prioritization, feasibility expectations, and independent civic engagement collapsed in parallel. Citizens did not abandon democratic ideals; they ceased to believe those ideals were achievable and ceased to prioritize them against competing goods. The findings challenge accounts that treat authoritarian legitimacy as a product of performance or ideology, suggesting instead that the most effective mechanism of regime stabilization may be the systematic closure of the political imagination.
keywords:
  - Cambodia
  - Asian Barometer Survey
  - authoritarian normalization
  - demobilization by subtraction
  - competitive authoritarianism
  - democratic aspirations
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{caption}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \usepackage{colortbl}
          \usepackage{xcolor}
          \captionsetup{width=\textwidth}
          \setlength\LTcapwidth{\textwidth}
          \floatplacement{figure}{H}
          \floatplacement{table}{H}
          \raggedright
          \AtBeginEnvironment{CSLReferences}{\setstretch{1}}
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
bibliography: references.bib
csl: chicago-author-date.csl
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r setup}
#| include: false
library(tidyverse)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/04_cambodia_fairy_tale")
results_dir  <- file.path(paper_dir, "analysis/results")
fig_dir      <- file.path(paper_dir, "analysis/figures")
tbl_dir      <- file.path(paper_dir, "analysis/tables")
rr_dir       <- file.path(paper_dir, "analysis/reviewer_response")

source(file.path(paper_dir, "R", "helpers.R"))

# dat <- readRDS(file.path(results_dir, "analysis_data.rds"))
stats <- readRDS(file.path(results_dir, "inline_stats.rds"))
```

# Introduction

In 2012, Cambodians rated the democratic future of their country at `r round(stats$dem_future_w3, 1)` out of 10.

It is worth pausing on that number. On a scale designed to capture the full range of democratic expectation, Cambodian respondents placed themselves at the extreme ceiling---more optimistic about their democratic trajectory than citizens of virtually any other country surveyed by the Asian Barometer [@Hu2023-ib]. This was not a country with free and fair elections, an independent judiciary, or a vibrant free press. It was a dominant-party authoritarian state under the continuous rule of Hun Sen and the Cambodian People's Party (CPP), a regime that had held power since 1985 and showed no signs of voluntary departure. And yet Cambodians believed, with remarkable consistency, that democracy was coming.

Nine years later, that same measure had fallen to `r round(stats$dem_future_w6, 1)` out of 10. In the same period, independent civic engagement---contacting influential figures outside the state, discussing politics, attending to political news---had collapsed to historic lows, even as participation through officially sanctioned channels quietly rose. Acceptance of authoritarian governance alternatives, including strongman rule, single-party rule, expert rule, and military rule, had risen sharply. Reported encounters with corruption had plummeted from `r round(stats$corrupt_witnessed_w3, 0)` percent to `r round(stats$corrupt_witnessed_w6, 0)` percent. And voter turnout in national elections, now conducted without any viable opposition, had paradoxically climbed to `r round(stats$voted_w6, 0)` percent.

What happened in between is the subject of this article. The answer, stated plainly, is that the Cambodian People's Party dissolved the only credible opposition party, the Cambodia National Rescue Party (CNRP), in November 2017---and simultaneously dismantled the independent media, civil society organizations, and informal networks through which citizens had accessed alternative political information and organized collective action. The combined effect restructured the political field so thoroughly that Cambodian citizens adjusted their attitudes, behaviors, and expectations in a remarkably coherent pattern. This article argues that this adjustment is best understood not as a story of authoritarian consolidation in the conventional sense---not as the triumph of ideology, coercion, or performance legitimacy---but as something quieter and, in its way, more profound: the education of a citizenry in the limits of the politically possible.

The phenomenon under investigation---the systematic elimination of organized opposition within an otherwise functioning electoral system---has become increasingly common across consolidating authoritarian regimes, from Turkey's suppression of the Peoples' Democratic Party to Russia's dismantling of Navalny's organizational network. Yet existing scholarship has focused primarily on the institutional mechanics of opposition elimination rather than its downstream effects on mass political orientations. This article proposes that opposition elimination---understood not as a single act but as the systematic dismantling of the political alternative and the institutional ecosystem that sustained it---produces a distinctive observable signature in public opinion: synchronized, multi-domain shifts in which democratic expectations, independent civic engagement, political attention, and willingness to report corruption all decline in parallel, while acceptance of authoritarian governance alternatives and participation through regime-managed channels rise. The critical diagnostic indicator is a divergence between electoral and civic participation---rising voter turnout amid collapsing independent engagement---that signals the transformation of elections from mechanisms of political choice into rituals of regime compliance. If this signature proves generalizable beyond the Cambodian case, it offers both a theoretical framework for understanding how authoritarian regimes stabilize through subtraction rather than through active legitimation, and a practical diagnostic tool for identifying democratic closure in real time using existing cross-national survey infrastructure.

The article proceeds as a structured analytic narrative, organized around the four survey waves as acts in a political drama. This is a deliberate methodological choice. Cambodia's trajectory between 2008 and 2021 has the unusual property of being both analytically tractable---the survey data capture discrete moments in a well-documented political sequence---and narratively coherent, with each wave mapping onto a distinct phase of political life. The narrative structure allows the article to preserve the temporal dynamics that conventional cross-sectional analysis tends to flatten, while the survey data anchor the story in systematic evidence rather than impressionistic accounts.

The question of how dominant-party regimes endure has generated a rich architectural vocabulary in contemporary scholarship. We are accustomed to analyzing the structural mechanics of survival: the skewed playing fields of competitive authoritarianism, the meticulous menu of manipulation required to manage electoral autocracies, and the stabilizing pillars of legitimation, repression, and co-optation.[^1] Yet, this institutional focus, while vital for understanding the resilience of the state, often obscures the quiet, downstream effects on the citizenry. Theories of regime survival tend to treat the public as a variable to be managed---either repressed into submission or bought off through patronage. This article shifts the analytical lens. By tracing the Cambodian electorate's trajectory from 2008 to 2021, it focuses not on how the regime maintained power, but on what happens to citizen orientations when the democratic alternative is systematically removed. The resulting quiescence, this article argues, is not a product of active authoritarian legitimation, but of demobilization by subtraction.

[^1]: For the foundational typologies of these institutional survival mechanics, see @Levitsky2010-bc; @Schedler2013-qk; and @Gerschewski2013-en. While these frameworks masterfully explain elite-level regime consolidation, they are less equipped to explain the simultaneous, multi-dimensional shifts in mass public opinion observed in the Asian Barometer Survey data during periods of opposition collapse.

The concept advanced here---demobilization by subtraction---is distinct from the two mechanisms most commonly invoked to explain political quiescence under authoritarianism. The first, repression-driven demobilization, holds that citizens withdraw from political life because the costs of participation have been raised through coercion, surveillance, or the threat of punishment.[^2] The second, co-optation, holds that citizens acquiesce because they have been incorporated into the regime's distributional networks and calculate that loyalty pays better than resistance.[^3] Both mechanisms presuppose an active regime strategy directed at the population. Demobilization by subtraction operates differently. Its proximate cause is not an action taken against citizens but an action taken against the political field itself: the removal of the credible alternative around which civic mobilization had organized. In practice, subtraction rarely occurs as a single surgical act; the Cambodian case involved the coordinated elimination of the opposition party, independent media outlets, and civil society infrastructure within a compressed period. The theoretical claim is that the decisive mechanism is the closure of the political field---the elimination of the vehicle for collective political expression---rather than the fear, material deprivation, or ideological conversion that characterize repression and co-optation respectively. The result is not fear-based withdrawal or interest-based compliance but something closer to what Gaventa termed "quiescence"---the internalization of powerlessness that follows from the sustained absence of viable channels for political expression.[^4] Where Lukes' third dimension of power operates by shaping preferences so that grievances never form, demobilization by subtraction operates one step later: grievances may persist, but the perceived vehicle for their expression has been eliminated, producing a cascade of behavioral and attitudinal adjustment that is visible across multiple survey dimensions simultaneously.[^5]

[^2]: On repression as a mechanism of demobilization, the foundational treatment is @Davenport2007-ww. For a comprehensive typology of repressive strategies, see also the essays collected in @Davenport2005-mk. On the specific mechanisms through which states demobilize protest, see @Earl2011-et.

[^3]: The co-optation framework is developed most fully in @Gandhi2008-he. On the equilibrium logic of authoritarian power-sharing, see @Przeworski2000-ed.

[^4]: Gaventa's central insight---that sustained powerlessness produces not resistance but the internalization of quiescence---provides the closest existing theoretical analogue to the pattern observed in the Cambodian data, though his analysis is situated in a very different political context [@Gaventa1980-ob].

[^5]: The distinction matters: Lukes' third dimension operates *before* grievances form; demobilization by subtraction operates *after* grievances have formed and been politically activated, removing the vehicle for their expression rather than the grievances themselves [@Lukes2005-bc].

The remainder of this article proceeds as follows. A data and measurement section describes the Asian Barometer Survey waves, sampling procedures, variable operationalization, and cross-wave comparability. The analytic narrative is then organized as four acts corresponding to the four available survey waves. Each section opens with a brief account of the political conditions prevailing at the time of the survey, then turns to the ABS data to trace how those conditions registered in citizen attitudes and behaviors. The analysis tracks five thematic domains across waves: political participation, authoritarian governance preferences, democratic expectations, corruption experience and perception, and media engagement and political interest. Following the narrative, a section on alternative mechanisms adjudicates the demobilization-by-subtraction interpretation against competing explanations, drawing on subgroup decompositions, response-style diagnostics, and placebo tests. A concluding section draws out the theoretical implications of the Cambodian sequence for the study of authoritarian durability and democratic aspiration.

# Data, Measures, and Comparability

This article draws on the Asian Barometer Survey (ABS), a repeated cross-sectional survey program administered across East and Southeast Asia since 2001. Cambodia has been included in four of the six completed waves: Wave 2 (fielded 2008, N = 1,000), Wave 3 (2012, N = 1,200), Wave 4 (2015, N = 1,200), and Wave 6 (2021, N = 1,242). Cambodia was not included in Wave 1 (2001–2003) or Wave 5 (2018–2019); the latter omission is particularly consequential, as it means no ABS data exist from the immediate aftermath of the CNRP dissolution and the 2018 one-party election. The survey's Wave 5 fieldwork period coincided with the immediate aftermath of the forced dissolution of the CNRP in November 2017 and the July 2018 national election in which the CPP ran virtually unopposed---conditions under which credible public opinion research could not be conducted. The absence of Wave 5 is thus not an artifact of research design but a reflection of the political closure this study documents: the very phenomenon under investigation rendered standard survey fieldwork untenable. This gap means the analysis cannot observe the immediate post-dissolution period directly, a limitation addressed below through adjudication tests designed to distinguish between an acute fear response and the longer-term normalization pattern proposed here. The four available waves nevertheless bracket the critical political sequence: pre-mobilization (2008), peak mobilization (2012), post-crackdown selective retreat (2015), and post-dissolution normalization (2021).

The ABS employs face-to-face interviews with nationally representative probability samples drawn through multistage stratified random sampling. The Cambodian samples are stratified by province and urban/rural classification, with primary sampling units selected proportional to population size and households selected through random walk procedures within enumeration areas. The survey instrument is administered in Khmer. Response rates across the four waves are consistently high, ranging from approximately 80 to 90 percent, which is typical for face-to-face surveys in Southeast Asia. No post-stratification weights are applied in the present analysis; the samples are treated as self-weighting, following the ABS's standard practice for single-country analyses.[^6]

[^6]: On ABS methodology and sampling design, see @Chu2008-fk The ABS Technical Report for each wave provides country-level details on sampling frames, fieldwork dates, and response rates.

## Supplementary Data---The 2017 Cambodia Socio-Economic Survey

The ABS data are supplemented by a single cross-section from the Cambodia Socio-Economic Survey (SES), a nationally representative household survey administered by the National Institute of Statistics. The SES wave employed here (deposited with the World Bank Microdata Library as KHM_2018_SAIE_v01; N = 3,363) was collected through face-to-face interviews with an inferred fieldwork period of early-to-mid 2017, based on internal temporal references: the questionnaire references the 2016–17 school year as current, the 2016 commune budget as the most recently completed fiscal cycle, and "additional money paid to officials in 2016" as a retrospective question about the immediately preceding year. This timing places the SES fieldwork in the final months before the intensification of political closure---after the collapse of the Culture of Dialogue but before Kem Sokha's arrest in September 2017 and the CNRP dissolution in November 2017.

Six governance items from the SES are relevant to the present analysis: perceived responsiveness of commune councillors and village chiefs (binary agree/disagree), perceived freedom of speech and freedom of association at the commune level (binary), perceived freedom to vote for any party without fear (binary), and reported payment of "tea money"---the Cambodian vernacular for petty bribery---to government officials in 2016 (binary). These items are not directly comparable to the ABS instruments in scale or sampling design, and are not treated as equivalent measures. Their value is triangulative: they establish the state of perceived civic freedom and corruption exposure in the immediate pre-dissolution period, a window the ABS does not cover. The SES data are reported in Appendix D.

Detailed SES governance modules were not publicly available for the 2018–2020 fieldwork period; a phone-based COVID-focused survey conducted in 2020 carried no political orientation or corruption items. The absence of detailed governance survey data during the years immediately following the CNRP dissolution is consistent with the broader pattern of political closure documented in this article.

The analysis tracks five thematic domains across waves. For each domain, the specific indicators, scale constructions, and cross-wave comparability considerations are as follows.

*Political participation.* The ABS participation battery employs a two-part structure: a gate question ("Have you ever done X?") asked of all respondents, followed by a frequency question asked only of those who answer affirmatively. This article reports the gate proportions---the share of the full sample reporting ever having engaged in each activity---rather than the frequency means, which are conditional on participation and based on much smaller subsamples. Six gate items are tracked: contacting elected officials, contacting civil servants, contacting influential persons, signing petitions, attending demonstrations, and contacting the media. Of these, contacting influential persons and attending demonstrations employ consistent wording across all four waves and are treated as fully comparable. The elected official and civil servant contact items use a narrower, problem-motivated framing in Wave 3 ("to complain about or seek help with a problem") that differs from the broader framing in Waves 2, 4, and 6; cross-wave comparisons involving Wave 3 for these items are flagged accordingly. Petitions, demonstrations, and media contact were not fielded in Wave 2. Community leader contact frequency (1–5 scale) and self-reported voter turnout (binary) supplement the gate items. A full comparability matrix documenting exact wording by wave is provided in the appendix (Table A1).

*Authoritarian governance preferences.* Four items assess respondent evaluations of non-democratic governance alternatives: single-party rule, strongman rule, military rule, and expert rule. Each is measured on a consistent 1–4 scale (1 = "very bad," 4 = "very good") across all waves in which the item appears. Expert rule was not fielded in Wave 2. These items are among the most stable in the ABS instrument, with identical wording and response options across waves.

## Democratic Orientations

The current analysis expands the treatment of democratic attitudes beyond the 0–10 expectation items employed in earlier drafts, incorporating measures of democratic commitment, democratic prioritization, and democratic satisfaction alongside democratic expectations. The expanded battery addresses a measurement concern that has become increasingly salient in the comparative democratic attitudes literature: that survey-measured "democratic support" is not a unitary construct but a composite of analytically separable dimensions that may respond differently to political conditions (Bratton et al. 2004; Mattes and Bratton 2007).

*Democratic commitment.* Two items measure abstract endorsement of democracy as a regime form. The first, fielded in all four waves, presents respondents with a three-option forced choice: "Democracy is always preferable to any other kind of government," "Under some circumstances, an authoritarian government can be preferable," or "For people like me, it doesn't matter whether we have a democratic or non-democratic government." The harmonized coding runs 1 (democracy always preferable) to 3 (doesn't matter), such that lower means indicate stronger democratic commitment. The second, fielded in Waves 3 through 6, asks respondents to evaluate the statement "Democracy may have its problems, but it is still the best form of government" on a standard 1–4 agree/disagree scale (higher = more pro-democratic). Both items employ identical wording across all waves in which they appear. Nonresponse rates on these items are low: under 2 percent in Waves 3 and 4, rising to 10–13 percent in Wave 6---elevated, but far below the levels observed on the evaluative-projective items discussed below.

*Democratic prioritization.* One item measures the trade-off between political freedom and economic equality, asking respondents to choose between "reducing economic inequality" and "protecting political freedom" on a five-point scale (1 = inequality reduction definitely more important, 5 = political freedom definitely more important), with a midpoint option of "both equally important." The item was fielded in Waves 3, 4, and 6 with identical wording. Nonresponse is minimal (1–8 percent across waves). A second trade-off item, pitting democracy against economic development on a five-point scale, is available across all four waves but proves empirically less dynamic in the Cambodian case; it is reported in the appendix for completeness.

*Democratic satisfaction.* Respondents in all four waves were asked: "On the whole, how satisfied or dissatisfied are you with the way democracy works in \[Cambodia\]?" on a four-point scale (1 = not at all satisfied, 4 = very satisfied). Wording is identical across waves. Nonresponse is low (0.3–6.4 percent). This item occupies an intermediate position between normative democratic support and regime performance evaluation; its trajectory proves substantively central to the normalization argument developed below.

*Democratic expectations.* Three items fielded in Waves 3, 4, and 6 measure perceived democratic quality on 0–10 scales: democratic past (how democratic was the country ten years ago?), democratic present (how democratic is the country today?), and democratic future (how democratic will the country be in ten years?). These items carry the highest nonresponse rates in the democratic battery---reaching 63 percent on the democratic future item and 43 percent on the democratic present item in Wave 6. The elevated missingness, concentrated in Wave 6, is itself analytically informative and is addressed at length in the alternative mechanisms section. Because the expectations items provide the most vivid illustration of the temporal arc documented in this article but rest on the weakest data foundations, the analytic strategy treats them as supporting evidence for patterns that are established more robustly on the commitment, prioritization, and satisfaction measures.

*Corruption experience and perception.* Witnessed corruption is measured as a binary indicator (ever personally witnessed or experienced government corruption). The Wave 2 item uses narrower wording focused specifically on bribe solicitation, while Waves 3–6 use a broader "witnessed or experienced" framing; Wave 2 estimates for this item should therefore be interpreted with caution. Perceived corruption in national and local government is measured on consistent 1–4 scales across all waves.

*Media engagement and political interest.* Political interest, political news following, and political discussion are measured on consistent scales (1–4 or 1–3) across all waves. Internet news frequency underwent substantial changes in response scale across waves (6-point in Waves 2 and 3, 8-point in Wave 4, 9-point in Wave 6) and has been harmonized to a common 1–6 frequency scale (1 = never, 6 = daily or more); details of the harmonization procedure are documented in a footnote at first use.

*Analytic strategy.* The analysis is primarily descriptive, reporting wave-level means, proportions, and cross-wave differences for each indicator. All estimates are accompanied by 95 percent confidence intervals computed from standard errors (binomial/Wilson intervals for proportions; normal-approximation intervals for means). Cross-wave differences for headline statistics are evaluated using two-sample t-tests (for means) or two-sample proportion tests (for proportions). The article does not employ regression modeling; the analytic leverage derives from within-case temporal variation across a well-documented political sequence rather than from cross-sectional covariate adjustment. This design cannot establish causality in the formal sense, but it can identify whether the timing, direction, and domain-specificity of observed shifts are consistent with the demobilization-by-subtraction mechanism and inconsistent with plausible alternatives. The alternative mechanisms section addresses this adjudication directly.

# The Quiet Kingdom: Cambodia in 2008

The first act of this story is a silence, and it is important to understand what kind of silence it was.[^7]

[^7]: Unless otherwise noted, the political chronology and electoral data in this section draw on @Hughes2010-ns; @Strangio2014-vi; @Un2013-ss; and @Un2013-zu

By 2008, Cambodia had been under continuous CPP rule for over two decades. The Paris Peace Accords of 1991 and the UNTAC-supervised elections of 1993 had produced a brief, chaotic experiment in multiparty democracy, but the CPP had never genuinely relinquished power, and after a violent coup in 1997 and disputed elections in 1998 and 2003, the party had consolidated its dominance through a combination of patronage, institutional control, and selective coercion. The 2008 National Assembly elections, which the CPP won with 58 of 123 seats and subsequently expanded through coalition politics, were competitive in form but not seriously contested in outcome.

The CPP's dominance by 2008 rested on institutional foundations that went far beyond electoral mechanics. The party had constructed a dense patronage network linking the central state to village-level administration, with commune chiefs serving as the capillary system through which resources, information, and political loyalty flowed.[^8] The military and security services were thoroughly integrated into the party apparatus, and the judiciary operated as an instrument of executive power rather than an independent check upon it [@Blake2019-ji; @Morgenbesser2016-vf]. Economic growth, driven by the garment export sector, construction, and an expanding tourism industry, provided a material foundation for the regime's implicit social contract: stability and modest prosperity in exchange for political acquiescence [@Hughes2011-zf].

[^8]: On the CPP's patronage architecture, see Caroline Hughes and Kheang Un, *Cambodia's Economic Transformation* (Copenhagen: NIAS Press, 2011); @Un2012-im

The opposition landscape was fragmented and demoralized. The Sam Rainsy Party (SRP), the most prominent non-CPP formation, commanded a loyal but limited urban constituency and suffered from periodic leadership crises as Sam Rainsy himself cycled between exile and return. The Human Rights Party, led by Kem Sokha, drew support from a distinct base but lacked the organizational reach to challenge the CPP's rural dominance. Neither party, operating independently, presented a credible threat to CPP hegemony---a structural condition that would change dramatically with their merger four years later [@Noren-Nilsson2015-xa].

The Wave 2 Asian Barometer data from this period paint a portrait of a population that was neither enthusiastic nor resistant, but simply quiet.[^9] Political participation was present but low in intensity: `r round(stats$gate_influential_w2, 1)` percent of respondents reported ever having contacted a person of influence, yet community leader contact averaged just `r round(stats$community_leader_w2, 2)` on a five-point frequency scale---figures that capture a population with prior exposure to civic channels but with limited ongoing engagement.[^10]

[^9]: Unless otherwise noted, the following scale conventions apply throughout: gate participation items are reported as the percentage of respondents who ever engaged in each activity (binary: ever vs. never); community leader contact is reported as a mean on a 1–5 scale (1 = never, 5 = often); authoritarian governance preferences on a 1–4 scale (1 = very bad, 4 = very good); democratic assessments on a 1–10 scale; corruption perceptions on a 1–4 scale (1 = hardly anyone, 4 = almost everyone); and media and political interest items on a 1–4 scale (1 = none, 4 = a great deal). Internet news frequency is harmonized to a 1–6 scale (1 = never, 6 = daily or more) across waves with differing original response categories; see the dedicated footnote in the Wave 3 discussion for details. Witnessed corruption is reported as a binary proportion. Voter turnout is self-reported and expressed as a percentage.

[^10]: Direct contact with elected officials and civil servants uses a broadly-worded 'at any level' framing in Wave 2 that differs from the narrower, problem-motivated framing in Wave 3 and from the Wave 4 and 6 formulations; these items are not used for cross-wave comparisons involving Wave 2 but are reported in the four-wave tables with appropriate documentation.

```{r tbl-table1}
#| tbl-cap: "Wave 2 (2008) Baseline: Political Orientations in Cambodia"
t1 <- readRDS(file.path(tbl_dir, "table1_w2_baseline.rds"))

t1 |>
  dplyr::select(label, w2_fmt) |>
  kbl(
    col.names = c("Variable", "W2 (2008)"),
    align     = "lc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 10,
    full_width = FALSE
  ) |>
  pack_rows("Political Participation", 1, 8) |>
  pack_rows("Authoritarian Preferences (1–4: very bad to very good)", 9, 12) |>
  pack_rows("Democratic Expectations (0–10 scale)", 13, 15) |>
  pack_rows("Corruption", 16, 18) |>
  pack_rows("Media & Political Interest", 19, 22) |>
  footnote(
    general = "Wave 2 N = 1,000. Participation rows 1–6 report the percentage of respondents who ever engaged in each activity (gate question); 95\\\\% CI (Wilson interval) in parentheses. Rows 7–8 report means (community leader contact: 1–5 scale) or proportions (voted); 95\\\\% CI in parentheses. '---' indicates item not fielded in this wave. Authoritarian preferences: 1 = very bad, 4 = very good. Democratic expectations: 0–10. Corruption witnessed: binary proportion. National/local corruption: 1 = not at all, 4 = extremely.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

The authoritarian attitude measures are perhaps the most telling. Single-party rule was rated at `r round(stats$single_party_w2, 2)` on a four-point scale (where 4 represents "very good"), strongman rule at `r round(stats$strongman_w2, 2)`, and military rule at `r round(stats$military_w2, 2)`. These are not the numbers of a population that embraces authoritarianism as an ideal; they are the numbers of a population that has learned to live within it. The slightly elevated tolerance for military rule is consistent with the CPP's own origins in Vietnamese-backed military intervention and the residual security anxieties of a post-genocide society, but even this measure sits below the scale midpoint.

The democratic orientation measures from 2008 establish a baseline of modest, largely passive democratic sentiment. On the forced-choice democracy preference item, 61.2 percent of respondents selected "democracy is always preferable," while 30.5 percent---nearly a third---chose "for people like me, it doesn't matter." Only 8.3 percent endorsed the view that authoritarian government could sometimes be preferable. The mean (1.69 on the 1–3 scale, where 1 represents the strongest democratic commitment) reflects a population that leaned democratic but without intensity: a plurality endorsed democracy as an abstract principle, but the large "doesn't matter" share signals the political indifference of a citizenry that had experienced little meaningful democratic choice. Democracy satisfaction stood at 2.97 on a four-point scale---essentially at the midpoint between "not very satisfied" and "fairly satisfied." The democracy-versus-economic-development trade-off item produced a mean of 2.43 on a five-point scale, leaning slightly toward economic development. These are not the numbers of a population invested in democratic politics. They are the numbers of a population for whom democracy was neither an aspiration nor a grievance---simply not a salient dimension of political life.

What is most notable about the 2008 data, viewed in retrospect, is how stable everything appears. There are no dramatic spikes, no anomalous values, no signs of the upheaval to come. Cambodia in 2008 was a country in political equilibrium---not a democratic equilibrium, but not one characterized by active authoritarian mobilization either. It was, to extend the metaphor, a kingdom in which the subjects had learned that the king was permanent and had arranged their lives accordingly. The question that the next four years would pose, with startling force, was what would happen when that arrangement was briefly and dramatically disrupted.

# The Awakening: Cambodia in 2012

Something extraordinary happened in Cambodia between 2008 and 2012, and the Wave 3 Asian Barometer data capture its imprint with unusual clarity.[^11]

[^11]: Unless otherwise noted, the political chronology and electoral data in this section draw on @Un2013-zu; @Um2014-ft; @Strangio2014-vi; and @Noren-Nilsson2015-xa On the 2013 election results specifically, see @COMFREL-The-Committee-for-Free-and-Fair-Elections-in-Cambodia-2013-kt

Across nearly every measure of civic engagement and political aspiration, Cambodian respondents in 2012 registered at or near the highest values in the dataset's history. Contact with influential persons---the gate item with consistent framing across waves---rose from `r round(stats$gate_influential_w2, 1)` percent in 2008 to `r round(stats$gate_influential_w3, 1)` percent. Community leader contact averaged `r round(stats$community_leader_w3, 2)` on a five-point frequency scale, up from `r round(stats$community_leader_w2, 2)`. And the political activities newly tracked beginning in Wave 3---attending demonstrations (`r round(stats$gate_demo_w3, 1)` percent ever) and signing petitions (`r round(stats$gate_petition_w3, 1)` percent ever)---debuted at notable levels consistent with a population in active political motion.

These are not the numbers of a quiescent population. They are the numbers of a population that had been politically activated.

The catalyst was the formation of the Cambodia National Rescue Party in 2012, a merger of the Sam Rainsy Party and the Human Rights Party under the joint leadership of Sam Rainsy and Kem Sokha. The CNRP did something that no previous Cambodian opposition had managed: it unified the fragmented non-CPP political space into a single, credible alternative with a coherent message of democratic reform, anti-corruption, and economic justice. The party drew its energy from an emerging urban youth demographic---the first generation of Cambodians born after the Khmer Rouge, educated in the expanding university system, connected to the wider world through social media, and impatient with the political settlement their parents had accepted.

The CNRP's formation in July 2012 was itself a product of political learning. Previous opposition efforts had foundered on precisely the fragmentation that the CPP exploited: multiple parties splitting the non-CPP vote, personal rivalries preventing coordination, and the absence of a unified organizational infrastructure capable of matching the CPP's commune-level reach. The merger of the SRP and the Human Rights Party, brokered through extended negotiations between Sam Rainsy and Kem Sokha, resolved the coordination problem at a stroke [@Noren-Nilsson2015-xa; @Un2013-ss]. The resulting party offered voters something genuinely new in Cambodian politics: a single, unified alternative to the CPP with recognizable leaders, a coherent platform, and---critically---the organizational capacity to mobilize supporters in both urban and peri-urban constituencies.

The mobilization dynamics visible in the 2012 survey data intensified dramatically in the months that followed. The July 2013 national election became the most competitive in Cambodia's post-UNTAC history, driven by a potent combination of youth demographics, social media penetration, and economic grievance. Facebook, which had reached an estimated 1.5 million Cambodian users by 2013, served as both an organizational tool and an alternative information ecosystem that partially circumvented CPP control of broadcast media [@Noren-Nilsson2021-is; @Morgenbesser2020-kz]. The CNRP's campaign messaging---centered on anti-corruption, land rights, wages for garment workers, and a promise to bring genuine democratic accountability---resonated with a generation of voters who had no memory of the Khmer Rouge and little patience for the stability-first bargain their parents had accepted [@Noren-Nilsson2021-sq].

The election results shocked the political establishment. The CNRP won 55 of 123 National Assembly seats, reducing the CPP's margin to its narrowest since 1993. Allegations of systematic electoral fraud triggered months of mass protest in Phnom Penh, with demonstrations regularly drawing crowds estimated in the tens of thousands. For observers and participants alike, the political settlement that had governed Cambodia since the late 1990s appeared to be unraveling [@Strangio2014-vi; @Morgenbesser2017-cu].

The democratic expectation measures tell the same story from a different angle. When asked to rate the democratic future of their country on a ten-point scale, Cambodian respondents in 2012 produced a mean of `r round(stats$dem_future_w3, 2)`. This is, by any standard, a remarkably high figure, indicating near-universal optimism. It suggests a population that had collectively decided that democratic change was not merely desirable but imminent---that the political order they had lived under for decades was on the verge of fundamental transformation.

The broader democratic orientation battery from Wave 3 reveals a more textured picture than the expectations items alone suggest. On the forced-choice democracy preference, 56.9 percent chose "democracy is always preferable"---actually *lower* than in 2008 (61.2 percent), with 30.2 percent selecting "doesn't matter" and 12.9 percent endorsing conditional authoritarianism. The 2012 moment was not, in other words, a peak of abstract democratic commitment; it was a peak of democratic *expectation* and *prioritization*. The distinction matters. Respondents who rated their country's democratic future at 9.58 on a ten-point scale were not necessarily more committed to democracy as an ideal than their 2008 counterparts---they were more convinced that democracy was arriving. The commitment-expectation gap in 2012---modest abstract endorsement alongside extraordinary feasibility optimism---is the signature of a population whose democratic orientation was driven by perceived possibility rather than by normative conviction.

Where prioritization is concerned, the 2012 data register the high-water mark. On the democracy-versus-equality trade-off, the pro-political-freedom side accounted for 47.1 percent of respondents (25.1 percent "somewhat," 22.0 percent "definitely"), while the pro-equality side stood at 50.4 percent, with a negligible 2.5 percent selecting "both equally important." The population was almost evenly divided, but nearly half were willing to prioritize political freedom over material equality---a level of democratic prioritization that would not be observed again. The "democracy is the best form of government" item debuted in this wave at 3.26 on a four-point scale (between "agree" and "strongly agree"), the highest value it would record across the three waves in which it was fielded. Democracy satisfaction held steady at 2.98, essentially unchanged from 2008.

The paradox, of course, is that this expectation was not based on any structural change in the regime. The CPP still controlled the state apparatus, the military, the courts, the media, and the vast patronage networks that sustained its rural base. Hun Sen showed no inclination to reform or retire. What had changed was not the regime but the perception of the regime's permanence---and that perceptual shift, as the subsequent waves would demonstrate, proved far more fragile than the political moment suggested.

The specific configuration of the democratic assessment triad in 2012 is theoretically revealing. The past-present-future structure---low past (`r round(stats$dem_past_w3, 2)`), middling present (`r round(stats$dem_present_w3, 2)`), exceptionally high future (`r round(stats$dem_future_w3, 2)`)---is not the signature of a population evaluating its democratic institutions on the basis of lived experience. A population grounded in experiential assessment would be expected to rate the future as a modest extrapolation from the present, not as a near-perfect score that exceeds the present by nearly four points. What the 2012 configuration captures instead is a hope-driven democratic orientation: the conviction that the current political order is transitional and that the democratic destination is both certain and imminent. This orientation, as subsequent waves would reveal, proved extraordinarily sensitive to political events---precisely because it was grounded in expectation rather than experience.

Even the corruption data reflect the mobilization moment. Witnessed corruption stood at `r round(stats$corrupt_witnessed_w3, 1)` percent in Wave 3, up from `r round(stats$corrupt_witnessed_w2, 1)` percent in Wave 2. This increase is almost certainly not an increase in actual corruption---the CPP's patronage system was well-established long before 2012---but rather reflects a population that had become more willing to identify and report corruption as a political problem, emboldened by the opposition's anti-corruption messaging and the expanded civic space that accompanied the CNRP's rise.

There are also subtler patterns in the Wave 3 data that resist easy interpretation. Traditional values measures---obedience to parents, deference to teacher authority---also peaked in this wave, an unexpected finding given that the broader mobilization was associated with youth activism and challenges to established authority. One possibility is that the CNRP's own messaging, which blended democratic aspiration with Khmer cultural conservatism and Buddhist moral frameworks, activated traditional values alongside rather than in opposition to democratic ones. Another is that the elevated values reflect a general intensification of political and social engagement in which all expressed attitudes, whether progressive or traditional, registered at higher levels. A third possibility, which cannot be ruled out, is survey context effects: the charged political environment of 2012 may have produced a general social desirability bias toward expressing strong opinions of any kind.

The media and political engagement measures complete the portrait of a mobilized citizenry. Political interest stood at `r round(stats$pol_interest_w3, 2)` on a four-point scale, and following political news registered at `r round(stats$pol_news_w3, 2)`---both the highest values in the Cambodian ABS series. Political discussion, at `r round(stats$pol_discuss_w3, 2)`, remained modest, suggesting that the activation captured in 2012 was more a matter of individual attention and institutional engagement than of deliberative civic culture. Internet news consumption, harmonized to a common six-point frequency scale across waves, stood at `r round(stats$news_internet_w3, 2)`---in the "never to hardly ever" range, consistent with the still very limited internet penetration of Cambodia in 2012.[^12]

[^12]: The ABS internet news item underwent substantial changes in response scale across waves (6-point in W2/W3, 8-point in W4, 9-point in W5/W6). The values reported here reflect harmonization to a common 1–6 frequency scale (1 = never, 6 = daily or more). The coarser W2/W3 scale cannot distinguish within-daily usage patterns available in later waves, but the harmonized values are comparable at the level of weekly, monthly, and annual frequency.

The 2012 data, taken together, present a portrait of a political community at its moment of maximum activation. Every dimension the Asian Barometer measures---participation, aspiration, attention, even the willingness to name corruption---points in the same direction. It is tempting, with the knowledge of what came after, to read this as the high-water mark of Cambodian democracy, the moment before the tide turned. But that reading, while emotionally satisfying, may be too simple. What the Wave 3 data actually capture is something rarer and more theoretically interesting: the approximate moment at which a population collectively updated its beliefs about what was politically possible, right before the regime demonstrated, with devastating effectiveness, that those beliefs were wrong.

# The Reckoning: Cambodia in 2015

That demonstration began at the ballot box and ended in the streets.[^13] The 2013 national election, held just months after the Wave 3 survey captured such exceptional optimism, confirmed the CNRP’s mobilization power: the opposition won 55 seats to the CPP’s 68, the narrowest margin in the country's modern history. Yet, the structural ceiling of the dominant-party state held firm. Amid widespread allegations of electoral fraud, the CNRP boycotted parliament, triggering months of mass demonstrations that soon merged with massive garment worker strikes. For a brief window, genuine political transformation appeared imminent. It did not arrive. Following a lethal state crackdown on protesters in January 2014, and a subsequent political compromise that coaxed the opposition into the National Assembly under a fragile and temporary "Culture of Dialogue," the momentum of the streets fractured.

[^13]: Unless otherwise noted, the political chronology in this section draws on @Chheang2015-ud; @Morgenbesser2017-cu; and @Noren-Nilsson2019-ck

The "Culture of Dialogue" that emerged from the July 2014 agreement between Sam Rainsy and Hun Sen was, in retrospect, less a genuine political opening than a controlled decompression. The CNRP entered parliament and gained access to committee chairmanships, lending the arrangement a veneer of power-sharing. But the underlying dynamics had not changed: the CPP retained control of the security forces, the judiciary, and the National Election Committee, while the CNRP's capacity for mass mobilization---its primary source of leverage---was progressively constrained through targeted legal harassment of its leaders and selective restrictions on public assembly [@Morgenbesser2017-cu; @Noren-Nilsson2019-ck; @Schedler2013-qk]. By 2015, the dialogue had largely stalled, and the political atmosphere had shifted from the hopeful if chaotic energy of 2013 to something more wary and uncertain. It was in this atmosphere that the Wave 4 Asian Barometer survey entered the field.

The Wave 4 data from 2015 capture a population in the midst of this reckoning---no longer at the peak of mobilization but not yet in the trough of withdrawal. The behavioral participation measures tell a more ambiguous story than might be expected. Contact with influential persons held essentially steady at `r round(stats$gate_influential_w4, 1)` percent (the 3.3 percentage-point increase from Wave 3 is not statistically significant; *p* = 0.11), and demonstration attendance, petition-signing, and civil servant contact all edged slightly upward from their Wave 3 levels. Community leader contact averaged `r round(stats$community_leader_w4, 2)` on a five-point frequency scale, essentially unchanged from 2012. The "selective retreat" of 2015 is not, in other words, a story written primarily in the behavioral data. It is written instead in the attitudinal record.

```{r tbl-table2}
#| tbl-cap: "Selective Retreat, 2012--2015: Wave 3 to Wave 4 Comparison"
t2 <- readRDS(file.path(tbl_dir, "table2_w3w4_comparison.rds"))

t2 |>
  dplyr::select(label, w3_fmt, w4_fmt, delta_fmt) |>
  kbl(
    col.names = c("Variable", "W3 (2012)", "W4 (2015)", "\\Delta (W4\u2212W3)"),
    align     = "lccc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size = 10,
    full_width = FALSE
  ) |>
  pack_rows("Political Participation", 1, 8) |>
  pack_rows("Authoritarian Preferences (1–4)", 9, 12) |>
  pack_rows("Democratic Expectations (0–10)", 13, 15) |>
  pack_rows("Corruption", 16, 18) |>
  pack_rows("Media & Political Interest", 19, 22) |>
  footnote(
    general = "W3 N = 1,200; W4 N = 1,200. Participation rows 1–6 report percentages (gate question); 95\\\\% CI (Wilson interval) in parentheses. All other rows report means or binary proportions with 95\\\\% CI (t-based) in parentheses. $\\\\Delta$ for participation rows is in percentage points (pp); for other rows, in scale units. '---' indicates item not fielded. Internet news harmonized to 1–6 frequency scale (1 = never, 6 = daily or more).",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE,
    escape = FALSE
  )
```

This attitudinal disengagement was not yet accompanied by a corresponding rise in authoritarian acceptance. Single-party rule actually dipped slightly from `r round(stats$single_party_w3, 2)` to `r round(stats$single_party_w4, 2)`, and military rule fell from `r round(stats$military_w3, 2)` to `r round(stats$military_w4, 2)`. Expert rule and strongman rule edged up marginally but remained below 2.0. This pattern is analytically important: it suggests that the demobilization observed between 2012 and 2015 was driven more by disillusionment and tactical withdrawal than by a genuine reorientation toward authoritarian preferences. Cambodians were pulling back from political action, but they had not yet accepted the authoritarian order as normatively appropriate.

The democratic orientation measures from Wave 4 reveal the first signs of the decoupling that would define the Cambodian trajectory: the simultaneous intensification of abstract democratic commitment and erosion of democratic expectation. On the forced-choice democracy preference, 74.9 percent of respondents selected "democracy is always preferable"---the highest proportion across all four waves, up sharply from 56.9 percent in 2012. The "doesn't matter" category collapsed from 30.2 to 18.9 percent. Citizens who had been indifferent to the democracy question in earlier waves were now choosing sides, and choosing democracy. The shift is consistent with a political environment in which democracy had become contested and salient---the 2013 election, the mass protests, the crackdown---rather than the background condition it had been in 2008. Commitment to democracy as an ideal intensified precisely because the ideal was under visible threat. "Democracy may have its problems but it is still the best form of government" dipped modestly from 3.26 to 3.15 (*p* \< 0.001), a statistically significant but substantively small decline that left the mean firmly in "agree" territory.

At the same moment, democratic prioritization began its descent. The democracy-versus-equality trade-off shifted significantly toward the equality pole: the pro-political-freedom side contracted from 47.1 to 34.6 percent, while the "both equally" midpoint expanded from 2.5 to 8.2 percent (mean: 2.91 → 2.58, *p* \< 0.001). Citizens were declaring their commitment to democracy in the abstract while becoming less willing to prioritize it when placed in competition with economic equality. The pattern is not contradictory; it reflects a population that had learned from the events of 2013–2014 that democratic aspiration carried real costs---political instability, economic disruption, state violence---and was beginning to weigh those costs against the ideal even while affirming the ideal itself.

The democratic expectation measures had already begun their descent. The democratic future rating fell from `r round(stats$dem_future_w3, 2)` to `r round(stats$dem_future_w4, 2)`---a drop of nearly two full points on the ten-point scale (*p* \< 0.001). The democratic present declined from `r round(stats$dem_present_w3, 2)` to `r round(stats$dem_present_w4, 2)`. Only the democratic past held relatively steady at `r round(stats$dem_past_w4, 2)`. Democracy satisfaction fell to 2.71---the lowest value in the four-wave series---consistent with a widening gap between democratic expectations (still elevated at `r round(stats$dem_future_w4, 2)`) and perceived democratic reality (`r round(stats$dem_present_w4, 2)`). The gap between what citizens hoped for and what they observed produced dissatisfaction; the narrowing of that gap by Wave 6, accomplished not by improving reality but by lowering expectations, would produce the opposite.

The configuration of the 2015 democratic orientation data is, in retrospect, the clearest signal of the normalization process in its early stages. Ideals intensified. Prioritization weakened. Expectations fell. Satisfaction declined. The four dimensions, which might be expected to move in concert if democratic attitudes were a unitary construct, instead moved in characteristically different directions---a pattern that the elimination of the CNRP two years later would amplify to its full expression.

The corruption data from Wave 4 introduce a striking counterpoint. Witnessed corruption surged to its highest level in the dataset: `r round(stats$corrupt_witnessed_w4, 1)` percent of respondents reported having personally encountered corruption. This is unlikely to reflect merely increased willingness to report; the political environment of 2015 was, if anything, less permissive of anti-government speech than 2012. The more likely explanation is that the post-2013 period, with its political bargaining, institutional jockeying, and expanded CNRP presence in parliament, created more visible sites of corruption---more interactions between citizens and competing political actors, more opportunities to observe the transactional nature of Cambodian governance up close.

The informational dimensions of political life registered the earliest signs of the withdrawal that would accelerate dramatically by 2021. Political interest declined from `r round(stats$pol_interest_w3, 2)` to `r round(stats$pol_interest_w4, 2)`, and following political news fell from `r round(stats$pol_news_w3, 2)` to `r round(stats$pol_news_w4, 2)`---modest drops in absolute terms, but significant as the leading edge of a trend that would deepen sharply. Political discussion held essentially flat at `r round(stats$pol_discuss_w4, 2)`, suggesting that Cambodians were beginning to disengage from formal political information channels while maintaining informal conversational patterns. Internet news consumption, meanwhile, continued its modest climb from `r round(stats$news_internet_w3, 2)` to `r round(stats$news_internet_w4, 2)`, reflecting the early stages of broadening internet access even as political engagement with that access was beginning to wane.

The 2015 wave, viewed within the larger narrative, represents the hinge point---the moment when the trajectory could still have gone either way. The CNRP was weakened but present, its leaders alternately accommodated and harassed, its supporters disillusioned but not yet defeated. The data register this ambiguity: democratic hope was falling but had not crashed. Behavioral participation remained largely intact while attitudinal engagement had already begun its descent. Authoritarian acceptance had not yet risen. Corruption was more visible than ever. It was a moment of unstable equilibrium, poised between the democratic aspiration of 2012 and the authoritarian normalization that was about to descend.

# The Silence: Cambodia in 2021

On November 16, 2017, the Supreme Court of Cambodia dissolved the Cambodia National Rescue Party.[^14] The ruling, which cited a vaguely defined conspiracy to overthrow the government, banned 118 CNRP officials from political activity for five years and transferred the party's parliamentary seats to smaller, CPP-aligned parties. Kem Sokha had already been arrested in September; Sam Rainsy remained in exile. In the July 2018 national election, the CPP won all 125 seats in the National Assembly. Cambodia became, for all practical purposes, a one-party state.

[^14]: Unless otherwise noted, the political chronology in this section draws on @Un2020-dq; @Chheang2021-td; and @Morgenbesser2019-mb

The CNRP's dissolution did not occur in isolation. It was the centerpiece of a broader campaign of political closure that systematically dismantled the infrastructure of independent civic life. The *Cambodia Daily*, the country's most prominent English-language newspaper and a persistent thorn in the CPP's side, was forced to close in September 2017 after receiving a disputed tax bill of \$6.3 million [@Noren-Nilsson2021-is]. Radio Free Asia's Cambodian bureau was shuttered, and several of its journalists faced criminal charges. Independent radio stations that had broadcast opposition content were pressured to switch to pro-government programming or cease operations entirely. Civil society organizations faced intensified regulatory scrutiny under the 2015 Law on Associations and Non-Governmental Organizations (LANGO), which gave the government broad discretionary power to dissolve organizations deemed to threaten public order [@Noren-Nilsson2020-ev].

The combined effect was the elimination not merely of the political opposition but of the broader ecosystem---media, civil society, informal networks---through which citizens had accessed alternative political information and organized collective action. The transition, in Levitsky and Way's typology, was from competitive authoritarianism, in which opposition parties exist and can occasionally win, to hegemonic authoritarianism, in which the electoral arena is formally maintained but substantively emptied [@Levitsky2010-bc; @Schedler2002-eg; @Morgenbesser2019-mb]. The 2018 national election, conducted without the CNRP on the ballot, produced a CPP sweep of all 125 National Assembly seats---a result that formalized what the dissolution had already accomplished.

The Wave 6 data from 2021 record the consequences.

The participation data from 2021 present a more differentiated picture than a simple collapse narrative would suggest. Contact with influential persons---the item most directly measuring independent civic engagement---fell sharply, from `r round(stats$gate_influential_w4, 1)` percent in 2015 to `r round(stats$gate_influential_w6, 1)` percent (*p* \< 0.001), a 33 percentage-point decline that represents one of the steepest single-item shifts in the dataset. Community leader contact fell to a mean of `r round(stats$community_leader_w6, 2)` on the frequency scale, and political discussion dropped to `r round(stats$pol_discuss_w6, 2)` on a four-point scale---a population that had largely stopped talking about politics with one another. Demonstration attendance held near its 2015 level at `r round(stats$gate_demo_w6, 1)` percent.

What complicates any uniform "participation collapse" characterization is that several formal channel indicators actually rose substantially by 2021: petition-signing reached `r round(stats$gate_petition_w6, 1)` percent, civil servant contact climbed to `r round(stats$gate_civil_w6, 1)` percent, and media contact rose to `r round(stats$gate_media_w6, 1)` percent. These increases likely reflect two overlapping dynamics: the rapid expansion of internet access enabling digital petitions and online contact, and the transformation of formal institutional channels into vehicles for routine patronage interaction under a now-uncontested regime. Participation through officially sanctioned channels rises precisely as independent civic action---contacting influential figures outside the state apparatus, organizing demonstrations, discussing politics---contracts. The divergence between these two trajectories is itself evidence of demobilization by subtraction: what citizens can still do expands; what citizens choose to do in pursuit of genuine political alternatives collapses.

The rise in civil servant contact to `r round(stats$gate_civil_w6, 1)` percent in Wave 6, against the backdrop of collapsing independent engagement, warrants careful interpretation. This pattern need not reflect passive acquiescence alone. In a political environment where opposition parties, independent media, and civil society organizations have been systematically dismantled, contacting a civil servant may represent the only remaining channel through which citizens can address grievances or secure material needs. The education of a citizenry, in this sense, involves not only the lowering of democratic expectations but also a pragmatic recalibration of political strategy: citizens learn which levers still function and redirect their efforts accordingly. This reading preserves a degree of agency within the normalization framework---demobilized citizens are not necessarily inert, but their participation has been redirected into channels that reinforce rather than challenge the regime's institutional architecture. The result is a form of engagement that is simultaneously rational at the individual level and depoliticizing at the collective level.

```{r tbl-table3}
#| tbl-cap: "Four-Wave Trajectory of Political Orientations in Cambodia, 2008--2021"
t3 <- readRDS(file.path(tbl_dir, "table3_four_wave_trajectory.rds"))

t3 |>
  dplyr::select(label, w2_fmt, w3_fmt, w4_fmt, w6_fmt, delta_w3w6_fmt, delta_w4w6_fmt) |>
  kbl(
    col.names = c("Variable", "W2 (2008)", "W3 (2012)", "W4 (2015)",
                  "\\textbf{W6 (2021)}", "\\Delta W3{\\textrightarrow}W6",
                  "\\Delta W4{\\textrightarrow}W6"),
    align     = "lcccccc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position", "scale_down"),
    font_size = 10,
    full_width = FALSE
  ) |>
  column_spec(5, bold = TRUE) |>
  pack_rows("Political Participation", 1, 8) |>
  pack_rows("Authoritarian Preferences (1–4: very bad to very good)", 9, 12) |>
  pack_rows("Democratic Expectations (0–10 scale)", 13, 15) |>
  pack_rows("Corruption", 16, 18) |>
  pack_rows("Media & Political Interest", 19, 22) |>
  footnote(
    general = "N per wave: W2 = 1,000; W3 = 1,200; W4 = 1,200; W6 = 1,242. Rows 1–6 (Political Participation) report the percentage of respondents who ever engaged in each activity (gate proportion), with 95\\\\% CI (Wilson interval) in parentheses; $\\\\Delta$ for these rows is in percentage points. Rows 7–8 and all remaining rows report unweighted wave means with 95\\\\% CI (t-based) in parentheses; $\\\\Delta$ is the arithmetic difference. '---' indicates variable not fielded in that wave. $\\\\Delta$ W3$\\\\rightarrow$W6 = peak-to-trough change (W6 minus W3). $\\\\Delta$ W4$\\\\rightarrow$W6 = post-dissolution change (W6 minus W4). W6 column in bold marks the endpoint of the trajectory. Voted last election and corruption witnessed are binary proportions.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE,
    escape = FALSE
  )
```

```{r figure1, fig.cap="Political Orientations in Cambodia by Domain, 2008--2021. Point estimates with 95\\% confidence intervals (Wilson intervals for proportions; t-based for means) across six thematic domains. Dashed vertical line marks 2017 CNRP dissolution. Y-axes are free-scaled; each panel uses its own metric range. Participation panel reports gate proportions (\\% ever engaged). Democratic Expectations panel: 0--10 scale. Democratic Commitment panel: dem\\_best\\_form (1--4 mean) and dem\\_vs\\_equality (1--5 mean), showing stable abstract democratic commitment diverging from declining expectations. Source: Asian Barometer Survey, Waves 2, 3, 4, 6.", out.width="100%", fig.pos="!htbp"}
knitr::include_graphics(file.path(rr_dir, "fig1_with_uncertainty.pdf"))
```

# Democratic Orientations in 2021

The democratic orientation data from Wave 6 complete the divergence pattern that first appeared in 2015, now with the full force of the post-dissolution political closure behind it.

Abstract democratic commitment, remarkably, survived the dissolution largely intact. "Democracy may have its problems but it is still the best form of government" registered a mean of 3.20 in Wave 6---between "agree" and "strongly agree," statistically indistinguishable from its Wave 4 value (3.15; *p* = 0.080) and only marginally below the Wave 3 peak of 3.26 (*p* = 0.014). The stability of this item across a period in which every other democratic indicator declined sharply is among the most important findings in the dataset. Cambodians in 2021 continued to endorse democracy as an ideal at rates essentially unchanged from the peak mobilization of 2012. Whatever the dissolution accomplished, it did not produce ideological conversion.

The forced-choice democracy preference item tells a subtler story. The proportion selecting "democracy is always preferable" declined from its Wave 4 peak of 74.9 percent to 66.5 percent---still a clear majority, but a significant contraction. The critical diagnostic is *where the movement went*. The "doesn't matter" category, which had contracted sharply between 2012 and 2015 (30.2 to 18.9 percent), continued to decline to 17.5 percent---its lowest level across all four waves. The growth was concentrated entirely in "under some circumstances, an authoritarian government can be preferable," which more than doubled from 6.1 to 16.0 percent. This distribution is inconsistent with generalized political apathy, which would inflate the "doesn't matter" category, and inconsistent with fear-driven acquiescence, which would push respondents toward whichever option appeared safest. It is consistent with a specific form of pragmatic recalculation: a subset of citizens who had committed to democracy as an ideal in 2015, when the opposition was still present, reassessed that commitment in light of a political reality in which democratic alternatives no longer existed. They did not become indifferent; they became conditionally authoritarian. The distinction between apathy and conditional acceptance is central to the normalization framework proposed here.

Democratic prioritization---measured on an item with 8.1 percent nonresponse, among the lowest rates in the political battery---reached its lowest recorded level. The democracy-versus-equality trade-off declined to a mean of 2.31 (*p* \< 0.001 for both the W3→W6 and W4→W6 comparisons). The pro-political-freedom side of the distribution, which had accounted for 47.1 percent of respondents in 2012, contracted to 21.7 percent---a collapse of more than half. The movement was not uniform across the distribution. The "somewhat pro-freedom" category suffered the steepest proportional decline (25.1 to 8.4 percent), while "definitely pro-freedom" eroded more modestly (22.0 to 13.3 percent). The citizens most susceptible to deprioritization were those whose commitment to political freedom had been moderate and conditional---precisely the pragmatic middle that the normalization framework predicts would be most responsive to the closure of democratic possibility. Meanwhile, the "both equally important" midpoint expanded from 2.5 percent in 2012 to 15.1 percent---a sixfold increase in the population that had retreated to noncommittal ambivalence. These respondents are the attitudinal analogue of the behavioral withdrawal documented in the participation data: citizens who previously held a directional political preference and have ceased to hold one.

The democratic expectation items continued their descent, though these estimates rest on substantially weaker data foundations than the commitment, prioritization, and satisfaction measures. With 63 percent nonresponse on the democratic future item and 43 percent on the democratic present item in Wave 6, the expectation measures describe the orientation of the minority of respondents who retained enough political engagement to formulate a numerical judgment about Cambodia's democratic trajectory. The patterns are reported here as supporting evidence for the normalization dynamics established above on more robust measures, not as independent findings. The democratic future rating fell from `r round(stats$dem_future_w4, 2)` in 2015 to `r round(stats$dem_future_w6, 2)` in 2021 (*p* \< 0.001), extending the decline from the extraordinary `r round(stats$dem_future_w3, 2)` recorded in 2012 to a level that, while still modestly optimistic, represents a drop of nearly three points---roughly 30 percent of the available scale range---across nine years. Democratic present stood at `r round(stats$dem_present_w6, 2)`, essentially unchanged from 2012 (`r round(stats$dem_present_w3, 2)`), while the democratic past rose to `r round(stats$dem_past_w6, 2)`---the only democratic measure to increase between 2012 and 2021. Among respondents who answered, the upward revision of the past alongside the downward revision of the future is suggestive of a flattened temporal perspective---though this interpretation necessarily rests on a selected subsample and should be treated as illustrative rather than definitive.

The most robust single finding from the Wave 6 democratic orientation battery---measured on an item with only 6.4 percent nonresponse, on identical wording across all four waves---is the satisfaction paradox [@Linde2003-an; @Valgarsson2022-cb]. Democracy satisfaction rose to 3.10, the highest value across all four waves and a statistically significant increase from both the Wave 4 trough of 2.71 (*p* \< 0.001) and the Wave 3 value of 2.98 (*p* \< 0.001). On the surface, this is paradoxical: satisfaction with democracy reached its peak in the least democratic moment in the country's modern history, three years after a one-party election in which the CPP won every seat. The paradox dissolves under the normalization framework. Satisfaction is a function of the gap between expectations and perceived reality. In 2015, expectations remained elevated while reality was visibly deteriorating, producing a large gap and low satisfaction. By 2021, expectations had collapsed while the regime was delivering stability and modest service provision through its patronage infrastructure. The gap closed from below, and satisfaction rose. Citizens became more satisfied with "the way democracy works" not because democratic quality improved---by any external assessment it deteriorated dramatically---but because they had recalibrated their understanding of what "democracy" meant in the Cambodian context. The satisfaction recovery is the clearest single indicator of normalization in the dataset: it shows adaptive preference formation operating in real time, as citizens adjusted their evaluative standards to accommodate the available political reality.

The authoritarian acceptance measures moved in mirror image. All four alternative governance indicators rose to their highest levels: expert rule to `r round(stats$expert_rule_w6, 2)`, single-party rule to `r round(stats$single_party_w6, 2)`, strongman rule to `r round(stats$strongman_w6, 2)`, and military rule to `r round(stats$military_w6, 2)`. The jump was concentrated entirely in the W4-to-W6 period---the six years that encompassed the CNRP's destruction and the subsequent one-party election. What had been stable or declining through 2015 now surged.

There is an important interpretive question here, and the article's argument depends on how it is answered. Does the rise in authoritarian acceptance represent genuine preference change---Cambodians coming to believe that authoritarian governance is normatively desirable---or does it represent something else? Three considerations suggest the latter. First, the magnitude of the shift, while statistically meaningful, is modest in absolute terms: all four measures remain around 2.2 on a four-point scale, below the midpoint. Cambodians are not rating authoritarianism as "good"; they are rating it as slightly less bad than before. Second, the timing---concentrated after the removal of the democratic alternative---suggests a mechanism closer to adaptive preference formation than genuine conversion. Third, and most importantly, the shift co-occurs with the collapse of democratic future expectations (from `r round(stats$dem_future_w4, 2)` to `r round(stats$dem_future_w6, 2)`), rising perceptions of government information withholding, and declining political interest. This is not the attitudinal profile of a population that has found a new political faith; it is the profile of a population that has stopped believing the old one was achievable.

The article proposes the term authoritarian normalization to describe this process, distinguishing it from authoritarian legitimation. Legitimation implies that citizens have come to view the regime as rightfully entitled to rule, whether on the basis of performance, ideology, or tradition. Normalization implies something weaker but potentially more durable: citizens have come to view the regime as the only available reality and have adjusted their expressed preferences to align with that reality. The distinction matters theoretically because it suggests different mechanisms of regime stability---and different vulnerabilities. A regime sustained by legitimation is threatened when performance falters or ideology loses appeal. A regime sustained by normalization is threatened when alternatives re-enter the political imagination.

The normalization framework finds a productive analogue in Wedeen's analysis of political compliance in Hafez al-Assad's Syria, where citizens participated in elaborate public rituals of regime praise not because they believed the regime's claims but because acting as if they believed was the condition of ordinary life [@Wedeen1998-qz; @Wedeen2000-bg]. The Cambodian case differs from Wedeen's Syria in a respect that is not incidental but analytically fundamental. Wedeen's compliance is performative: citizens maintain private skepticism while publicly enacting belief, producing a gap between inner conviction and outer behavior that defines the regime's mode of domination. The Cambodian pattern visible in the ABS data is not performative in this sense. The survey responses are private---administered in face-to-face interviews with confidentiality protocols, not in public settings where social surveillance operates. The attitudinal shifts documented here appear to reflect genuine reorientation, not strategic concealment: citizens moved into conditional authoritarianism rather than into noncommittal safe responses, and the shifts cut across demographic subgroups and partisan categories in ways that fear-based models do not predict. What the Cambodian case shares with Wedeen's Syria is not the mechanism but the outcome: a population that has arranged its cognitive and evaluative life around the available political reality, not because it has been persuaded of the regime's virtue but because it has incorporated the regime's permanence as a background condition of practical reasoning. The difference is that Wedeen's Syrians *performed* accommodation while potentially preserving private dissent; the Cambodian data suggest something closer to the internalization of accommodation---or at minimum, the erosion of the psychological infrastructure required to sustain dissent in the absence of institutional support.

The democracy-versus-equality trajectory---from near-parity in 2012 to a two-to-one margin favoring equality over political freedom by 2021---bears the hallmarks of what Elster termed adaptive preference formation: the adjustment of desires to match perceived possibilities, the "sour grapes" mechanism through which agents come to devalue what they cannot obtain [@Elster2016-la, orig. 1983; @Przeworski2022-om]. The Cambodian case, however, reveals a more complex process than Elster's formulation anticipates. The adaptation is partial. Citizens deprioritized political freedom when placed in explicit competition with equality, and they lowered their expectations of democratic progress, but they did not revise their abstract endorsement of democracy as the best form of government. The ideal persisted even as the practical commitment to it eroded. This partial adaptation is theoretically consequential: it suggests that democratic values operate at multiple psychological levels, some of which are more responsive to political context than others [@Jones2025-tf]. The surface layer---abstract endorsement---proves resilient to even dramatic changes in the structure of political alternatives. The operational layer---the willingness to bear costs for democratic outcomes---proves highly elastic. A regime that successfully eliminates alternatives need not worry about what citizens believe in principle; it need only wait for the practical implications of that belief to atrophy.

The corruption data offer perhaps the most unsettling evidence of normalization. Witnessed corruption collapsed from `r round(stats$corrupt_witnessed_w4, 1)` percent in 2015 to `r round(stats$corrupt_witnessed_w6, 1)` percent in 2021. Perceived national government corruption also declined, from `r round(stats$corrupt_national_w4, 2)` to `r round(stats$corrupt_national_w6, 2)` on the four-point scale. On the surface, these numbers might suggest genuine anti-corruption progress. The CPP did conduct selective anti-corruption campaigns during this period, targeting rivals and politically inconvenient officials. But the magnitude of the witnessed-corruption decline---from nearly two-thirds to one-seventh of the population---is too large to explain through policy alone.

The more parsimonious explanation is that the corruption decline reflects the same demobilization process visible in the participation data. When citizens withdraw from political life---when they stop contacting officials, stop attending meetings, stop engaging with governance institutions---they also reduce their exposure to the sites where corruption is experienced and observed. A population that does not petition, does not demonstrate, and does not contact civil servants is also a population that has fewer occasions to witness a bribe demanded or a favor exchanged. Demobilization, in this reading, does not merely reduce political action; it reduces the perceptual evidence that political action might address. The regime becomes less corrupt not because it has reformed but because citizens have stopped looking.

External assessments corroborate the interpretation that declining witnessed corruption reflects perceptual withdrawal rather than actual improvement. Cambodia's score on Transparency International's Corruption Perceptions Index remained essentially flat throughout the study period, moving from 21 in 2015 to 23 in 2021 on a 100-point scale---consistently ranking the country among the most corrupt in Southeast Asia and globally [@Transparency_International2023-im]. Reports from domestic and international monitoring organizations continued to document endemic corruption in land concessions, public procurement, and the judiciary throughout the period in which survey respondents' reported witnessing of corruption collapsed from `r round(stats$corrupt_witnessed_w3, 0)` to `r round(stats$corrupt_witnessed_w6, 0)` percent. The gap between external expert assessments, which registered no meaningful improvement, and citizens' reported experience, which registered a fourfold decline, is difficult to reconcile with genuine anti-corruption progress. It is, however, consistent with a population that has progressively withdrawn from the administrative and political sites where corruption is encountered---not because corruption diminished, but because the reasons and opportunities for engaging with state institutions outside officially sanctioned channels have.

An alternative, and not mutually exclusive, explanation for the corruption decline invokes the survey environment itself. In a political climate where the opposition has been criminalized and independent media shuttered, respondents may reasonably calculate that acknowledging witnessed corruption---even in a nominally confidential survey---carries risks of self-incrimination or unwanted official attention. The social desirability bias literature has long recognized that survey responses on sensitive topics shift in predictable directions under conditions of political constraint, with respondents offering answers that align with perceived official expectations.[^15] The post-2017 Cambodian environment, in which public criticism of the government carried tangible legal consequences, represents precisely the conditions under which such bias would be expected to intensify. The two explanations---reduced exposure through demobilization and increased reticence through social desirability pressure---likely operate in tandem, each reinforcing the other to produce the dramatic decline observed in the data.

[^15]: On the reliability of survey responses under electoral authoritarianism, see @Schedler2007-pv @Frye2017-qo demonstrate that even in semi-competitive authoritarian contexts, survey-measured approval can closely track genuine opinion---but their findings also imply that as political space closes further, the gap between expressed and private attitudes is likely to widen.

The voter turnout paradox completes the portrait. Despite the collapse of all other forms of political engagement, self-reported voting in the last election rose from `r round(stats$voted_w3, 1)` percent in 2012 to `r round(stats$voted_w4, 1)` percent in 2015 to `r round(stats$voted_w6, 1)` percent in 2021. This pattern---rising electoral participation amid collapsing civic engagement---is a well-documented feature of hegemonic authoritarian systems, where elections serve not as mechanisms of political choice but as rituals of regime affirmation. The 2018 election made the ritual character of Cambodian voting unusually visible. The CNRP-in-exile’s “clean finger” campaign---urging voters to boycott and display uninked fingers as proof of refusal---transformed the indelible ink applied at polling stations from an anti-fraud measure into a tool of social surveillance. Factory workers were warned that returning to work without inked fingers would result in wage deductions; village heads distributed polling slips with the expectation that residents would comply; government employees were required to participate [@McCargo2018-ms]. Voting in a one-party election is not a political act in the same sense as voting in a competitive one; it is closer to a demonstration of compliance, a public performance of belonging to the political community as the regime defines it. The fact that turnout rises precisely as meaningful engagement collapses suggests that Cambodians understand the distinction even as they perform the ritual.

The international orientation data add a final dimension to the normalization portrait. Cambodian assessments of Chinese influence in the region grew more positive between Waves 4 and 6, rising from `r round(stats$china_asia_goodharm_w4, 2)` to `r round(stats$china_asia_goodharm_w6, 2)` on a four-point scale measuring perceived benefit versus harm. More striking still, expectations of future Asian---implicitly Chinese---regional influence climbed from `r round(stats$future_influence_asia_w4, 2)` in Wave 4 to `r round(stats$future_influence_asia_w6, 2)` in Wave 6, a trajectory that places Cambodia among the most China-positive publics in the ABS dataset. This is consistent with Cambodia's deepening alignment with Beijing during this period, visible in Belt and Road infrastructure investment, the Ream Naval Base controversy, and Cambodia's consistent support for Chinese positions within ASEAN. But it also fits the normalization logic: as democratic futures recede and Western-aligned political alternatives are eliminated, alignment with the ascendant authoritarian power in the region becomes the natural orientation of a population adjusting to the available political reality. A paired comparison with Vietnam or the Philippines---where territorial disputes drive China perceptions in sharply negative directions despite similar economic interdependence---would sharpen this point considerably [@Strangio2020-gw].

The informational withdrawal that began tentatively in 2015 reached its full expression by 2021. Following political news fell to `r round(stats$pol_news_w6, 2)`, political interest declined to `r round(stats$pol_interest_w6, 2)`, and political discussion dropped to `r round(stats$pol_discuss_w6, 2)`---the lowest values recorded across all four waves. These figures describe a population that has not merely withdrawn from political action but has largely ceased to attend to political information at all. The internet news variable throws this withdrawal into particularly sharp relief: harmonized internet news consumption rose to `r round(stats$news_internet_w6, 2)` by Wave 6---approaching "at least once a week"---reflecting the rapid expansion of internet infrastructure and smartphone penetration across Cambodia. Cambodians in 2021 had far greater access to online information than at any previous survey wave, yet their political interest, news following, and willingness to discuss politics all fell to historic lows. The infrastructure for engagement had expanded; the will to use it for political purposes had collapsed. The self-reinforcing logic is straightforward: citizens who do not participate have less reason to follow political developments; citizens who do not follow political developments have less basis on which to participate. Demobilization, once initiated through the removal of the political alternative, generates its own sustaining momentum through this informational feedback loop---a dynamic that suggests the effects of the CNRP's dissolution may prove more durable than the act itself.

The Cambodia of 2021 is, in a sense, the inverse of the Cambodia of 2012. Where the earlier moment was characterized by activation across all dimensions---participation, aspiration, attention, even the willingness to name corruption---the later moment is characterized by withdrawal across all the same dimensions. The symmetry is striking and, for the theoretical argument, essential. It suggests that what changed between the two moments was not any single attitude or behavior but the underlying orientation from which attitudes and behaviors derive---what might be called the perceived horizon of political possibility. When that horizon expanded, everything expanded with it. When it contracted, everything contracted.

# Alternative Mechanisms and Adjudication

The preceding narrative has interpreted the synchronized attitudinal and behavioral shifts across ABS waves as evidence of demobilization by subtraction. Before drawing theoretical implications, it is necessary to adjudicate this interpretation against four competing explanations: preference falsification under intensified repression, pandemic-era behavioral disruption, technological channel substitution, and generational replacement. Each generates distinct observable predictions that can be evaluated against the available data.

*Preference falsification.* The most direct challenge to the normalization interpretation holds that the observed shifts do not reflect genuine attitude change but rather increased survey reticence under harsher post-2017 authoritarian conditions. If respondents in 2021 were concealing pro-democratic views and understating civic participation out of fear, the measured declines would overstate actual demobilization. The response-style diagnostics offer partial support for this concern: item nonresponse on democratic expectations items rose substantially between Wave 3 and Wave 6, from `r round(stats$na_dem_future_w3, 0)` percent to `r round(stats$na_dem_future_w6, 0)` percent on the democratic future item, and from `r round(stats$na_dem_past_w3, 0)` to `r round(stats$na_dem_past_w6, 0)` percent on the democratic past item.[^16] This elevated nonresponse is consistent with a political environment in which some respondents feel uncomfortable offering evaluations of their country's democratic trajectory.

[^16]: Mean nonresponse across all political items also rose, from approximately `r round(stats$mean_na_political_w3, 0)` percent in Wave 3 to `r round(stats$mean_na_political_w6, 0)` percent in Wave 6. This general pattern is consistent with either increased reluctance to venture political opinions or decreased political engagement reducing the salience of these questions.

Three considerations, however, limit the scope of this alternative. First, if survey reticence were the primary driver, one would expect it to manifest broadly---through increased straightlining, elevated "don't know" responses across all political batteries, or systematic acquiescence. The data show a more selective pattern: straightlining rates on the authoritarian preferences battery actually declined from Wave 4 (`r round(stats$straightline_w4, 0)` percent) to Wave 6 (`r round(stats$straightline_w6, 0)` percent), and acquiescence rates (proportion selecting the most positive response) remained low and stable across waves. Respondents who did answer appear to have engaged substantively with the questions rather than defaulting to safe or formulaic responses. Second, the demobilization pattern began before the most intensive period of political closure: democratic future expectations fell by nearly two points between Wave 3 and Wave 4, and political interest declined between the same waves, well before the CNRP dissolution and its associated repressive apparatus. If fear were the primary mechanism, the timing should concentrate in the post-2017 period rather than emerging incrementally from 2012 onward. Third, the placebo battery provides the strongest evidence of domain specificity. Non-political items that should not be affected by opposition elimination---national pride (`r round(stats$placebo_natpride_w3, 2)` in Wave 3, `r round(stats$placebo_natpride_w4, 2)` in Wave 4, `r round(stats$placebo_natpride_w6, 2)` in Wave 6), family economic situation (`r round(stats$placebo_econFamily_w3, 2)`, `r round(stats$placebo_econFamily_w4, 2)`, `r round(stats$placebo_econFamily_w6, 2)`), and perceived economic change (`r round(stats$placebo_econChange_w3, 2)`, `r round(stats$placebo_econChange_w4, 2)`, `r round(stats$placebo_econChange_w6, 2)`)---show no systematic decline across waves. If generalized survey reticence were suppressing responses, these items would be expected to drift downward as well; they do not.

The most direct empirical test of the preference-falsification hypothesis draws on two ABS items asking respondents to rate, on a four-point scale from strongly disagree to strongly agree, whether "people are free to speak what they think without fear" (*dem_free_speech*) and "people can join any organization they like without fear" (*gov_free_to_organize*). Perceived freedom to organize declined steadily across waves---from a mean of `r round(stats$free_org_w2, 2)` in Wave 2 to `r round(stats$free_org_w3, 2)` in Wave 3, `r round(stats$free_org_w4, 2)` in Wave 4, and `r round(stats$free_org_w6, 2)` in Wave 6---confirming that the perceived civic environment tightened substantially over the study period, consistent with the documentary record. Perceived freedom of speech was comparatively flat (W2: `r round(stats$free_speech_w2, 2)`; W3: `r round(stats$free_speech_w3, 2)`; W4: `r round(stats$free_speech_w4, 2)`; W6: `r round(stats$free_speech_w6, 2)`). If preference falsification were the primary driver of the observed attitudinal shifts, controlling for these freedom perceptions should attenuate the wave-level coefficients. Adding both variables as covariates in OLS models predicting democratic future expectations, democracy satisfaction, and democratic preference produces virtually no attenuation: the Wave 6 coefficient on democratic future expectations changes by `r round(stats$attenuation_fut_w6, 1)` percent; the Wave 4 coefficient by `r round(stats$attenuation_fut_w4, 1)` percent. Attenuation on democracy satisfaction (`r round(stats$attenuation_sat_w4, 1)` percent at Wave 4) and democratic preference (`r round(stats$attenuation_pref_w4, 1)` percent at Wave 4) is similarly negligible. Freedom perceptions absorb essentially none of the temporal variance in democratic orientations. Furthermore, the moderation pattern runs opposite to the preference-falsification prediction: adding an interaction between wave and perceived freedom of speech reveals that the declines in democratic future expectations are concentrated among respondents who perceive *more* freedom of speech---not less (Wave 4 interaction: β = `r round(stats$interact_speech_fut_w4, 2)`, *p* \< .001; Wave 6: β = `r round(stats$interact_speech_fut_w6, 2)`, *p* \< .001). If survey reticence were suppressing pro-democratic responses, the effect should concentrate among those who feel least free to speak; instead, the deepest disillusionment runs through precisely those who feel most free to do so---a pattern consistent with genuine preference revision rather than strategic concealment.

The expanded democratic orientation battery provides additional leverage against the preference-falsification hypothesis. The response distributions on two forced-choice items are particularly informative. On the democracy-versus-equality trade-off, the Wave 6 shift moved respondents primarily into the noncommittal midpoint ("both equally important," expanding from 2.5 to 15.1 percent) rather than toward the pro-regime pole (which grew modestly). A frightened population seeking the socially safe answer in an authoritarian context would be expected to choose the answer that aligns with perceived regime preferences---in this case, prioritizing equality over political freedom, consistent with the CPP's developmental rhetoric. Instead, the largest category of movement was into ambivalence. Similarly, on the forced-choice democracy preference, the Wave 6 growth concentrated in "authoritarian sometimes preferable" (6.1 to 16.0 percent) rather than in "doesn't matter" (which actually continued to decline, from 18.9 to 17.5 percent). If survey reticence were the primary driver, the noncommittal option---the safest refuge for a respondent unwilling to stake a position---would be expected to expand. It contracted. The growth of conditional authoritarianism alongside the contraction of apathy is more consistent with citizens who have genuinely reassessed the trade-offs between regime types than with citizens concealing stable preferences behind safe responses.

The most parsimonious reading is that the elevated nonresponse on democratic expectation items reflects a combination of genuine disengagement (respondents who no longer attend to politics and therefore have no opinion to offer) and some degree of reluctance, while the responses that are provided reflect real attitudinal shifts. This interpretation is consistent with Frye et al.'s finding that survey-measured approval in semi-competitive authoritarian contexts can closely track genuine opinion, though the gap between expressed and private attitudes likely widens as political space narrows further [@Frye2017-qo].

The nonresponse gradient across the democratic orientation battery in Wave 6 is itself analytically informative. The items with the lowest nonresponse---"democracy is the best form of government" (9.9 percent), the democracy-versus-equality trade-off (8.1 percent), the forced-choice democracy preference (13.4 percent)---are agree/disagree or forced-choice items that require respondents to select among pre-defined positions. The items with the highest nonresponse---democratic future (63 percent), democratic present (42.5 percent)---are evaluative-projective items that require respondents to generate a numerical judgment about Cambodia's democratic trajectory. The cognitive and political demands of these two classes of items differ substantially. Placing Cambodia's democratic future on a ten-point scale requires both the political engagement to hold an opinion about the country's trajectory and the willingness to commit that opinion to a number. Selecting "agree" or "disagree" on whether democracy is the best form of government requires neither.

The nonresponse gradient thus maps onto the normalization framework. Citizens can still answer whether democracy is desirable in the abstract (low cost, low sensitivity), but increasingly decline to evaluate where their country sits on a democratic continuum (high cost, high sensitivity, or---per the theory---genuine absence of a formed opinion because the question has lost practical relevance). The 63 percent who did not answer the democratic future item and the 37 percent who answered with diminished expectations are not competing narratives but complementary expressions of the same process. The former have disengaged from the question entirely; the latter have revised their answer downward.

Bounds analysis confirms that the headline decline is robust to extreme assumptions about the missing respondents. If all nonrespondents on the democratic future item in Wave 6 are imputed at the maximum scale value of 10---the most conservative possible assumption, granting every missing respondent the most optimistic democratic expectation---the resulting population mean is 8.77. Even under this extreme scenario, the decline from the Wave 3 value of 9.58 remains 0.81 points---substantial and directionally unambiguous even under this most favorable possible assumption. The core finding that democratic expectations fell substantially between 2012 and 2021 is not an artifact of differential nonresponse.

# Partisan Heterogeneity and the Scope of Demobilization

If demobilization by subtraction operates specifically through the removal of the opposition---rather than through the generalized closure of political space---the resulting attitudinal shifts should concentrate among citizens who were politically attached to the dissolved party. Former CNRP supporters, who lost their vehicle for political expression, should exhibit steeper declines in democratic expectation and civic engagement than CPP supporters, whose political home remained intact. The ABS vote choice items, which identify respondents' reported partisan behavior in the most recent national election, permit a direct test of this prediction.

The partisan distribution across waves is itself the first piece of evidence. In Wave 4 (2015), CNRP voters constituted 21.3 percent of the sample and CPP voters 50.8 percent, with 17.2 percent nonvoters and 7.2 percent declining to answer. By Wave 6 (2021), the CNRP share had collapsed to 2.7 percent---34 respondents, a residual so small it is analytically uninformative except as a marker of the dissolution's completeness. The CPP share returned to 60.1 percent, absorbing an unknown proportion of former CNRP supporters who reported voting CPP in the uncontested 2018 election. The DK/Refuse category nearly tripled, from 7.2 to 18.8 percent. The redistribution of the CNRP's former constituency---into CPP voting, nonvoting, or refusal to answer---*is* the subtraction documented at the level of individual survey responses.

The W4-to-W6 decomposition of key outcomes by partisan group yields a finding that initially appears to cut against the subtraction-specific mechanism: the shifts are broadly uniform. Contact with influential persons---the clearest measure of independent civic engagement---dropped by approximately 30 percentage points across all groups: 35.1 points among CPP voters, 30.9 among CNRP voters, 29.9 among nonvoters, and 25.6 among those declining to answer. Democratic future expectations declined by 1.27 to 1.39 points across groups. Single-party rule acceptance rose by 0.29 to 0.47 points. The absence of partisan heterogeneity is robust: no group was spared the demobilization, and no group experienced it with distinctive severity.

Two considerations complicate the interpretation of this uniformity as evidence against the subtraction mechanism. The first is methodological. The W6 partisan categories are themselves *products* of the dissolution. The 746 respondents classified as "CPP voters" in Wave 6 include an unknown number of former CNRP sympathizers who voted CPP in a one-party election---citizens for whom "CPP voter" is a post-dissolution category imposed by the absence of alternatives rather than a stable partisan identity. The DK/Refuse category, which tripled to 18.8 percent, likely contains the most resistant former CNRP supporters---those who could neither bring themselves to report voting CPP nor risk identifying with a criminalized party. Their democratic orientation profile in Wave 6 is consistent with this reading: they registered the lowest proportion selecting "democracy always preferable" (61.1 percent) and the highest rate of conditional authoritarianism (18.1 percent) of any group---a pattern more consistent with suppressed democratic preferences under duress than with genuine authoritarian conversion. The dissolution did not merely change attitudes; it destroyed the partisan categories through which differential effects could be measured.

The second consideration is theoretical, and it reframes the uniformity finding as consistent with---rather than contradictory to---the subtraction mechanism, once the mechanism is specified at the correct level of analysis. The prediction that CNRP supporters should be differentially affected assumes that subtraction operates like losing one's team: supporters of the dissolved party lose their political vehicle while everyone else is unaffected. The framework advanced in this article operates at a different level. The removal of the credible alternative restructures the political field itself---the horizon of possibility for all participants, not only those who identified with the opposition. CPP voters in 2015 inhabited a competitive political system in which their party could face genuine electoral challenge, in which political information served a decision-making function, and in which contacting influential persons outside the party apparatus had strategic rationale. CPP voters in 2021 inhabited a one-party state in which none of these conditions obtained. The elimination of the CNRP depoliticized the system as a whole, reducing the informational and participatory incentives for all citizens regardless of partisan attachment. Under this reading, the uniformity of the observed shifts is precisely what a field-level mechanism would predict---and what a party-level mechanism would not.

# Supplementary Evidence from the 2017 Socio-Economic Survey

The 2017 Cambodia Socio-Economic Survey provides a second-instrument snapshot of civic orientations in the final months before the CNRP dissolution, partially addressing the temporal gap between ABS Waves 4 (2015) and 6 (2021). The SES data are not directly comparable to the ABS in scale or design, but the directional evidence is informative.

In early-to-mid 2017, 86.1 percent of SES respondents agreed that "people in this commune are free to speak what they think without fear," and 87.9 percent agreed that "people in this commune can join any organisation they like without fear." The perceived freedom to vote for any party without fear stood at 94.0 percent---near-universal agreement, recorded in the final months before the elimination of the only meaningful alternative party. These figures establish that on the eve of the crackdown, Cambodian citizens overwhelmingly perceived themselves as living in a civic environment that permitted free expression, association, and partisan choice. The political closure that followed was not acting on a population that already felt constrained. The subjective distance between the pre-dissolution civic environment (as perceived by citizens) and the post-dissolution environment documented in the ABS Wave 6 data was, by these measures, substantial.

The tea money item corroborates the corruption trajectory from a different instrument and a more behaviorally concrete measure. In the SES, 34.9 percent of respondents reported that they or a household member had paid additional money to government officials in 2016 to "get things done or get things done faster." This rate---measuring direct personal bribery experience rather than the broader "witnessed or experienced" formulation used in the ABS---confirms that petty corruption remained widespread in the immediate pre-dissolution period. The ABS witnessed-corruption item registered `r round(stats$corrupt_witnessed_w4, 0)` percent in 2015 and `r round(stats$corrupt_witnessed_w6, 0)` percent in 2021; the SES tea money rate of 34.9 percent for 2016, falling between these values on a narrower and more concrete measure, is consistent with a corruption environment that had not materially improved before the dissolution. The subsequent ABS collapse is thus more difficult to attribute to genuine anti-corruption progress and more consistent with the exposure-reduction and social-desirability mechanisms discussed above.

Satisfaction with local governance was high: 88.2 percent agreed that commune councillors are "helpful and responsive at solving problems," and 89.4 percent agreed that the village chief "generally works for the benefit of citizens." These figures are consistent with the CPP's patronage infrastructure functioning effectively at the commune level---the same infrastructure into which, the article argues, citizens redirected their engagement after the dissolution foreclosed independent channels. The post-dissolution rise in civil servant contact (from `r round(stats$gate_civil_w4, 0)` percent in Wave 4 to `r round(stats$gate_civil_w6, 0)` percent in Wave 6) landed on institutional ground that citizens already found functional.

The broader pattern of SES data availability reinforces the temporal argument. Detailed governance modules comparable to the items analyzed here were not publicly released for the 2018–2020 period; the only post-dissolution SES data available contain a single item on satisfaction with the government's handling of COVID-19. The systematic unavailability of civic orientation data from the years immediately following the CNRP dissolution---across both the ABS and the SES---reflects the same political closure the article documents and limits the temporal precision with which the attitudinal shifts can be dated.

*Pandemic-era behavioral disruption.* The Wave 6 data were collected in 2021, during the COVID-19 pandemic. Public health restrictions on gatherings, movement, and face-to-face interaction could independently depress participation measures, particularly those involving physical contact such as community leader visits and demonstration attendance. This concern is legitimate but bounded. The key attitudinal shifts---the decline in democratic future expectations, the rise in authoritarian governance acceptance---are not plausibly explained by pandemic restrictions, which constrain behavior but have no direct mechanism for altering governance preferences. The decline in democratic expectations began in Wave 4 (2015), six years before the pandemic, establishing a trajectory that the 2021 data continue rather than initiate. Moreover, the participation items that most directly involve physical contact do not show the expected pandemic signature: demonstration attendance held essentially constant between Wave 4 and Wave 6 (`r round(stats$gate_demo_w4, 1)` to `r round(stats$gate_demo_w6, 1)` percent, a non-significant difference), while the sharpest participation decline occurred in contacting influential persons, an activity that need not involve face-to-face interaction. The pandemic may have amplified certain behavioral measures, but it cannot account for the breadth, direction, or temporal onset of the observed pattern.

One placebo item does register a notable decline: interpersonal trust fell from `r round(stats$placebo_trust_binary_w4, 0)` percent (Wave 4) to `r round(stats$placebo_trust_binary_w6, 0)` percent (Wave 6) on the binary "most people can be trusted" item. This drop is consistent with the social atomization documented in pandemic-era survey research across multiple countries and may also reflect broader erosion of social cohesion under intensified authoritarianism [@Aassve2021-dn]. However, the fact that this is the only non-political item showing a substantial decline---while national pride, family economic assessment, and perceived economic change all remain stable---suggests a localized effect rather than a general contamination of the survey environment.

*Temporal resolution and the dissolution-as-keystone interpretation.* A distinct methodological concern cuts across all of the preceding adjudication tests: the six-year gap between Wave 4 (2015) and Wave 6 (2021) means the analysis cannot determine the temporal profile of the observed shifts. Three models are consistent with the available data. The first---an acute shock model---holds that the sharpest attitudinal changes occurred in the immediate aftermath of the CNRP dissolution in November 2017 and the one-party election of July 2018, and that the 2021 data capture a population that has already partially recovered from the initial shock. Under this model, the 2021 values understate the peak magnitude of demobilization. The second---a gradual normalization model---holds that the shifts accumulated incrementally across the 2017–2021 period as the consequences of political closure compounded through informational feedback loops and the progressive atrophy of civic habits. Under this model, the 2021 data capture the endpoint of a process still in motion. The third---a repression-with-subtraction model---holds that the observed shifts are not primarily products of field restructuring but of the cumulative repressive apparatus that accompanied the dissolution: media closures, civil society restrictions, legal harassment, and the chilling effects of the criminalization of the opposition. Under this model, subtraction is a component of a standard authoritarian consolidation package rather than a distinctive mechanism.

The available data cannot definitively adjudicate among these three models, and the honest answer is that elements of all three likely operated simultaneously. Two considerations, however, favor the subtraction-centered interpretation over the repression-centered alternative. First, the attitudinal trajectory began before the most intensive period of repression: democratic future expectations fell by nearly two points and political interest declined measurably between Wave 3 (2012) and Wave 4 (2015), a period during which the CNRP was under selective pressure but still functionally intact and the broader repressive apparatus had not yet been deployed. If cumulative repression were the primary driver, the onset of attitudinal change should concentrate in the post-2017 period rather than emerging incrementally from the peak of mobilization. Second, the domain-specificity of the shifts---political attitudes and behaviors declining while non-political attitudes remain stable---is more consistent with a mechanism that operates through the structure of political alternatives than with generalized intimidation, which would be expected to depress expressed attitudes across domains. The interpretation advanced here treats the dissolution as the keystone of the closure process: the act that restructured the political field and set in motion the cascading adjustments documented in the data, even as the accompanying repressive measures amplified and accelerated those adjustments.

*Technological channel substitution.* The finding that participation through formal channels rose between Wave 4 and Wave 6---petition-signing, civil servant contact, and media contact all increased---while contact with influential persons outside the state apparatus declined, could reflect a shift in participation modalities driven by smartphone and internet penetration rather than political demobilization. If citizens simply migrated from informal, face-to-face civic engagement to digitally mediated formal channels, the resulting pattern would resemble demobilization even if total civic engagement were unchanged or increasing.

This interpretation is difficult to sustain for two reasons. First, it cannot explain the concurrent attitudinal shifts: channel substitution might redistribute participation across modes but provides no mechanism for the simultaneous decline in democratic expectations, rise in authoritarian acceptance, and collapse of political interest. If citizens were merely engaging through different channels, their attitudes toward democracy and authoritarianism should remain stable. They did not. Second, the formal channels that expanded---petitioning, contacting civil servants, contacting media---are precisely those most compatible with regime-managed participation under hegemonic authoritarianism. The rise of these channels alongside the collapse of independent engagement is more consistent with the substitution of participatory form for participatory substance than with genuine channel migration.

*Generational replacement.* Cambodia's young population and rapid demographic change raise the possibility that apparent attitudinal shifts across waves reflect cohort replacement rather than within-individual attitude change: the politically activated youth of 2012 aged into their thirties by 2021, while a new cohort with no memory of the CNRP mobilization entered the survey frame. If the post-2017 shifts are concentrated among younger respondents, compositional change rather than demobilization could be driving the aggregate pattern.

The subgroup decompositions do not support this interpretation. The key shifts between Wave 4 and Wave 6 are broad-based across age groups, education levels, and urban/rural residence. Contact with influential persons declined among under-30s (`r round(stats$sg_influential_under30_w4, 1)` to `r round(stats$sg_influential_under30_w6, 1)` percent), 30–49 year-olds (`r round(stats$sg_influential_3049_w4, 1)` to `r round(stats$sg_influential_3049_w6, 1)` percent), and over-50s (`r round(stats$sg_influential_50plus_w4, 1)` to `r round(stats$sg_influential_50plus_w6, 1)` percent) alike. Single-party rule acceptance rose across all three age groups. Political interest fell uniformly regardless of age, education, or urbanicity. If generational replacement were the primary mechanism, the shifts should concentrate in the youngest cohort or attenuate among older respondents who experienced the CNRP period directly; instead, the pattern is strikingly uniform. This breadth is consistent with a population-wide reorientation rather than a compositional artifact.

Taken together, the adjudication tests support the demobilization-by-subtraction interpretation while acknowledging that the 2021 data likely reflect some degree of pandemic amplification and elevated survey reticence on the most politically sensitive items. The core finding---synchronized, domain-specific shifts in political orientations that began before the pandemic, cut across all demographic subgroups, and leave non-political attitudes largely untouched---is more consistent with genuine attitudinal adjustment to political closure than with any single alternative mechanism.

# Conclusion

The story told in the preceding pages is, in one sense, particular to Cambodia---a product of specific historical circumstances, a specific regime, a specific opposition movement, and a specific act of political closure. In another sense, it is a story with much wider application, because the mechanism it illustrates---the restructuring of citizen orientations through the removal of political alternatives---operates wherever dominant-party regimes consolidate power. This concluding section draws out four theoretical implications of the Cambodian sequence and considers their relevance beyond the single case.

The first concerns the mechanism itself, and the expanded democratic orientation analysis permits a more precise formulation than earlier treatments. Demobilization by subtraction does not operate by eroding democratic values. Abstract democratic commitment---measured by the most widely used items in the cross-national survey toolkit---remained essentially stable across the Cambodian study period; a clear majority of respondents in 2021 continued to endorse democracy as the best form of government at rates statistically indistinguishable from the peak mobilization of 2012. What collapsed was not the ideal but the willingness to prioritize that ideal against competing goods---and, on the more tentatively measured expectation items, the perceived horizon within which the ideal could be realized. The pro-political-freedom share of the democracy-versus-equality trade-off contracted from 47 to 22 percent. Democracy satisfaction---paradoxically---rose to its highest recorded level, as citizens recalibrated their evaluative standards to accommodate a political reality in which the democratic alternative had been removed. Democratic expectations, measured on items carrying substantially higher nonresponse, fell by nearly three points on a ten-point scale---a decline that, while directionally robust even under extreme imputation assumptions, rests on a selected subsample and corroborates rather than independently establishes the normalization pattern. The divergence between resilient ideals and collapsing prioritization suggests that democratic attitudes in authoritarian contexts are not the unitary construct that much of the democratic values literature assumes. They have at least three partially independent dimensions---ideal endorsement, practical prioritization, and feasibility expectation---that respond to the structure of available political alternatives in characteristically different ways [@Jones2025-tf]. The mechanism, moreover, need not be attributed to the dissolution alone. The Cambodian closure involved the simultaneous elimination of the opposition party, independent media, and civil society organizations---a compressed sequence whose individual contributions cannot be isolated with the available data. The theoretical claim is that the *combined* effect of eliminating the political alternative and its supporting ecosystem produces the divergence pattern documented here, not that any single act within the sequence is independently sufficient. Cross-national surveys that aggregate these dimensions into a single "democratic support" index may be measuring a composite whose components move in opposite directions under precisely the conditions of greatest theoretical interest.

The second implication concerns the relationship between hope and democratic attitudes. The Wave 3 data, viewed alongside the expanded orientation battery, suggest that the extraordinary democratic optimism of 2012 was not accompanied by correspondingly extraordinary democratic commitment. The forced-choice democracy preference actually registered *lower* in 2012 than in 2015, when---counterintuitively---commitment surged to its highest level precisely as expectations were falling. This temporal decoupling suggests that democratic support in authoritarian contexts may be driven by two distinct psychological mechanisms that peak at different moments: feasibility perception, which responds to the structure of political opportunities and peaked when the CNRP made democracy appear imminent; and value crystallization, which responds to perceived threat and peaked when the post-2013 crackdown placed democratic values under visible siege. The Cambodian sequence---commitment lagging expectation on the upswing, then outlasting it on the downswing---challenges the foundational assumption of the political culture tradition that democratic values and democratic expectations move in tandem as expressions of a single underlying orientation.[^17]

[^17]: The canonical formulation of democratic values as a stable cultural attribute is found in the political culture tradition from @Inglehart1997-gw onward. The Cambodian data suggest a more situationally contingent model in which the components of democratic support are decoupled---closer to what Bratton and Mattes have termed "demand for democracy" as a function of perceived supply [@Bratton2004-ke; @Mattes2007-ej].

The third implication concerns corruption measurement. The finding that witnessed corruption declined precipitously alongside civic participation suggests a troubling methodological artifact: survey-based corruption indicators may systematically understate corruption in demobilized populations. When citizens withdraw from engagement with state institutions, they also withdraw from the sites where corruption is experienced and observed. A regime that successfully demobilizes its population will, by this logic, appear less corrupt on standard survey measures even if actual corruption is unchanged or increasing. This has practical consequences for the evaluation of anti-corruption campaigns in authoritarian contexts. The CPP’s selective anti-corruption efforts during this period may well have had some genuine effect, but the dramatic scale of the decline---from `r round(stats$corrupt_witnessed_w3, 0)` percent to `r round(stats$corrupt_witnessed_w6, 0)` percent (*p* \< 0.001)---is far more parsimoniously explained by the withdrawal of citizens from the institutional encounters where corruption becomes visible. Transparency International’s Corruption Perception Index and similar instruments may be particularly vulnerable to this dynamic in consolidating authoritarian regimes.

The fourth implication is diagnostic. The divergence between electoral participation and civic participation---turnout rising to `r round(stats$voted_w6, 0)` percent even as independent civic engagement collapsed---may serve as a generalizable indicator of authoritarian normalization. Where elections are maintained as regime rituals rather than competitive mechanisms, rising turnout signals not democratic health but its opposite: the transformation of voting from an act of political choice into an act of political compliance [@Schedler2002-eg; @Magaloni2010-wl]. The Cambodian case suggests that this turnout-participation divergence emerges specifically in the wake of opposition elimination, and that its appearance in other contexts---wherever rising turnout coincides with collapsing civic engagement---should be read as evidence that the electoral arena has been substantively emptied even if it remains formally intact.The

These findings carry inherent limitations. Four cross-sectional waves across thirteen years provide suggestive but not definitive evidence; the absence of Waves 1 and 5 from the Cambodian ABS series limits the precision with which structural breaks can be identified, and the six-year gap between Wave 4 (2015) and Wave 6 (2021) means the analysis cannot pinpoint whether the sharpest attitudinal shifts occurred immediately following the CNRP dissolution in 2017 or accumulated gradually across the broader closure process. Nor can the design isolate the specific contribution of the dissolution from the accompanying media shutdowns, civil society restrictions, and electoral manipulation that constituted the full authoritarian consolidation sequence. The elevated nonresponse on democratic expectation items in Wave 6---reaching `r round(stats$na_dem_future_w6, 0)` percent on the democratic future measure---means the attitudinal estimates from 2021 are based on the subset of respondents willing to venture an opinion, which may disproportionately exclude those with the strongest pro-democratic views or, alternatively, those least engaged with politics. A direct test of whether fear of speech drives this missingness is available from the ABS freedom-of-speech and freedom-of-assembly items: a logistic regression predicting nonresponse on the Wave 6 democratic future item, controlling for education and urbanicity, finds that neither perceived freedom of speech (OR = `r round(stats$nr_OR_speech, 2)`, *p* = `r round(stats$nr_p_speech, 2)`) nor perceived freedom to organize (OR = `r round(stats$nr_OR_org, 2)`, *p* = `r round(stats$nr_p_org, 2)`) predicts missingness. Respondents who feel less free to speak are no more likely to skip the democratic future question than those who feel more free, indicating that the elevated nonresponse rate is not the signature of a population censoring itself out of fear.

A methodological caution is warranted regarding the `r round(stats$na_dem_future_w6, 0)` percent non-response rate on the democratic future item in Wave 6. This level of missingness means that the reported 2021 mean is effectively a survivor statistic, reflecting only those respondents who retained enough political engagement---or perhaps political imagination---to formulate an expectation about Cambodia's democratic trajectory at all. Yet this limitation is also, in a sense, the study's most striking finding. If demobilization by subtraction operates by narrowing the horizon of political possibility, then the inability or unwillingness of nearly two-thirds of respondents to venture an opinion about Cambodia's democratic future is not merely a data problem to be corrected; it is the condition the framework predicts. The silence of the `r round(stats$na_dem_future_w6, 0)` percent and the diminished expectations of the `r round(100 - stats$na_dem_future_w6, 0)` percent who answered are not competing narratives but complementary expressions of the same process. The former have disengaged from the question entirely; the latter have revised their answer downward. Both constitute evidence of a citizenry educated out of democratic expectation, albeit through different registers of withdrawal.

The preceding section on alternative mechanisms has addressed the pandemic, generational, and preference-falsification concerns in detail; the conclusion here is that these factors may amplify specific estimates but cannot account for the overall pattern. Finally, any single-country study faces the inherent limitation of generalizability: the mechanisms identified here may be specific to Cambodia’s particular configuration of post-conflict politics, dominant-party rule, and opposition dynamics.

The most productive response to this limitation is comparative extension. The demobilization sequence observed in Cambodia---declining democratic expectations, followed by collapsing independent civic engagement, followed by rising authoritarian acceptance---generates a testable temporal ordering that can be examined in other cases of opposition elimination. Turkey after the suppression of the Peoples’ Democratic Party (HDP), Russia following the dismantling of Navalny’s organizational network, and Thailand in the years following the 2014 coup all represent contexts where credible political alternatives were systematically removed from the political field. The Asian Barometer dataset itself offers the most direct comparative leverage: Thailand’s post-coup trajectory and the Philippines under Duterte provide cases where similar dynamics may be observable using identical survey instruments. The turnout-participation divergence identified here---rising electoral compliance amid collapsing civic engagement---offers a particularly tractable diagnostic, as it requires only two commonly available survey measures and generates a clear empirical prediction. Future research should test whether the Cambodian sequence represents a general pattern of authoritarian normalization or a configuration specific to the conditions of this particular case.

This article began with a number: `r round(stats$dem_future_w3, 1)` out of 10. It is worth ending with the question of what that number meant. In 2012, Cambodian citizens looked at their country---a dominant-party authoritarian state with controlled media, a compromised judiciary, and a history of political violence---and saw a democratic future so bright they could barely imagine anything brighter. Nine years later, they had learned otherwise. The democratic future had not arrived. The party that promised it had been dissolved. The leaders who championed it were in prison or exile. And the population, having briefly imagined a different political life, had returned to the quiet arrangement their parents had known---participating in elections that offered no choice, accepting governance structures they had once rejected, and gradually, measurably, ceasing to pay attention.

Any reliance on longitudinal survey data carries inherent vulnerabilities, and this analysis of Cambodia’s trajectory is no exception. The dramatic peaks of 2012 and the stark valleys of 2021 are undeniably colored by the unique atmospheric conditions of those exact moments.[^18] Yet, the theoretical weight of this study does not rest on the absolute value of any single data point, but on the architecture of the sequence. By anchoring this narrative in within-case temporal variation---observing the same population across four distinct acts of a remarkably well-documented political closure---the noise of individual survey waves gives way to a clear, structural signal. It provides an unusually clean empirical window into how a population recalibrates its political reality when the horizon of possibility collapses.

[^18]: Two contextual anomalies warrant specific acknowledgment. First, the Wave 3 (2012) survey was fielded during an unprecedented window of opposition mobilization just prior to the 2013 elections, meaning the data likely capture a temporary zenith of democratic optimism rather than a stable baseline. Second, the Wave 6 (2021) data were collected during the COVID-19 pandemic, a period where public health restrictions inherently depressed physical civic participation (e.g., attending demonstrations) and may have temporarily inflated reliance on state authority. However, the fact that ideological demobilization (such as the drop in democratic future expectations) began in Wave 4, well before the pandemic, suggests the 2021 results reflect deeper authoritarian normalization rather than mere epidemiological artifacts. On the methodological challenges of surveying under authoritarian conditions more generally, see @Schedler2007-pv Furthermore, the single-country focus of this study naturally bounds its immediate generalizability, though the mechanisms observed here offer a framework for testing similar dynamics in other consolidating hegemonic regimes.

The fairy tale has no happy ending. But it has a lesson, and the lesson is not about Cambodia alone. It is about the fragility of democratic aspiration in the absence of democratic infrastructure, and about the remarkable efficiency with which authoritarian regimes can reshape citizen orientations simply by removing the object around which those orientations had formed. The education of Cambodia's citizenry between 2008 and 2021 was not accomplished through indoctrination or ideology. It was accomplished through subtraction---the removal of an alternative, and the long silence that followed.

\newpage

## AI assistance disclosure.

The author used Claude (Anthropic, version 4.5) to assist with R programming, including data cleaning scripts, model specification, and debugging of analysis code. All analytical decisions, theoretical framing, and interpretation of results are the author's own. AI-generated code was reviewed, tested, and modified by the author before inclusion in the analysis pipeline.

# References {.unnumbered}

::: {#refs}
:::
