% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper]{article}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother





\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\usepackage{setspace}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{float}
\usepackage{graphicx}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage{colortbl}
\usepackage{xcolor}
\captionsetup{width=\textwidth}
\setlength\LTcapwidth{\textwidth}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\raggedright
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{float}
\usepackage{graphicx}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage{colortbl}
\usepackage{xcolor}
\captionsetup{width=\textwidth}
\setlength\LTcapwidth{\textwidth}
\floatplacement{figure}{H}
\floatplacement{table}{H}
\raggedright
\AtBeginEnvironment{tablenotes}{\setstretch{1}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Online Appendix: The Education of a Citizenry},
  pdfauthor={Jeffrey Stark},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}


\title{Online Appendix: The Education of a Citizenry}
\author{Jeffrey Stark}
\date{March 4, 2026 at 5:38 PM}
\begin{document}
\maketitle


\setstretch{2}
\section*{Table A1: Cross-Wave Comparability of Survey
Items}\label{table-a1-cross-wave-comparability-of-survey-items}
\addcontentsline{toc}{section}{Table A1: Cross-Wave Comparability of
Survey Items}

\begingroup\fontsize{7}{9}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}[para]
\item \textit{Note:} 
\item Status: Full = identical or functionally equivalent wording across all waves used; Partial = minor wording shift or item not fielded in all waves (cross-wave comparison valid with caveats noted in text). No items are classified as Excluded from all comparisons; wave-pair restrictions are noted in the main text.
\end{TableNotes}

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.0cm}>{\raggedright\arraybackslash}p{2.3cm}>{\raggedright\arraybackslash}p{2.3cm}>{\raggedright\arraybackslash}p{2.3cm}>{\raggedright\arraybackslash}p{2.3cm}>{\raggedright\arraybackslash}p{1.5cm}}

\caption{\label{tbl-tableA1}}

\tabularnewline

\toprule
Variable & W2 (2008) & W3 (2012) & W4 (2015) & W6 (2021) & Status\\
\midrule
\endfirsthead
\multicolumn{6}{@{}l}{\textit{(continued)}}\\
\toprule
Variable & W2 (2008) & W3 (2012) & W4 (2015) & W6 (2021) & Status\\
\midrule
\endhead

\endfoot
\bottomrule
\insertTableNotes
\endlastfoot
Contacted elected official & Have you ever contacted an elected official to express your views? (broad) & Have you ever contacted an elected official to complain about or seek help with a problem? (narrow) & Have you ever contacted an elected official to express your views? (broad) & Have you ever contacted an elected official to express your views? (broad) & Partial\\
Contacted civil servant & Have you ever contacted a civil servant to express your views? (broad) & Have you ever contacted a government official to complain about or seek help with a problem? (narrow) & Have you ever contacted a civil servant to express your views? (broad) & Have you ever contacted a civil servant to express your views? (broad) & Partial\\
Contacted influential person & Have you ever contacted a person of influence? (broad) & Have you ever contacted a person of influence? (broad) & Have you ever contacted a person of influence? (broad) & Have you ever contacted a person of influence? (broad) & Full\\
Signed petition & Not collected in W2 & Have you ever signed a petition? (consistent) & Have you ever signed a petition? (consistent) & Have you ever signed a petition? (consistent) & Partial\\
Attended demonstration & Have you ever attended a demonstration or protest march? (consistent) & Have you ever attended a demonstration or protest march? (consistent) & Have you ever attended a demonstration or protest march? (consistent) & Have you ever attended a demonstration or protest march? (consistent) & Full\\
Contacted media & Have you ever contacted the media to express your views? (consistent) & Have you ever contacted the media to express your views? (consistent) & Have you ever contacted the media to express your views? (consistent) & Have you ever contacted the media to express your views? (consistent) & Full\\
Community leader contact & How often do you contact community leaders? (1–5 frequency) & How often do you contact community leaders? (1–5 frequency) & How often do you contact community leaders? (1–5 frequency) & How often do you contact community leaders? (1–5 frequency) & Full\\
Voted in last election & Did you vote in the last national election? (binary) & Did you vote in the last national election? (binary) & Did you vote in the last national election? (binary) & Did you vote in the last national election? (binary) & Full\\
Expert rule & Not collected in W2 & Should experts rather than government make decisions? (1–4) & Should experts rather than government make decisions? (1–4) & Should experts rather than government make decisions? (1–4) & Partial\\
Single-party rule & Should only one political party be allowed to contest elections? (1–4) & Should only one political party be allowed to contest elections? (1–4) & Should only one political party be allowed to contest elections? (1–4) & Should only one political party be allowed to contest elections? (1–4) & Full\\
Strongman rule & Should we get rid of parliament and elections for a strong leader? (1–4) & Should we get rid of parliament and elections for a strong leader? (1–4) & Should we get rid of parliament and elections for a strong leader? (1–4) & Should we get rid of parliament and elections for a strong leader? (1–4) & Full\\
Military rule & Should the military rule the country? (1–4) & Should the military rule the country? (1–4) & Should the military rule the country? (1–4) & Should the military rule the country? (1–4) & Full\\
Democratic future (0-10) & Not collected in W2 & How democratic will this country be in 10 years? (0–10) & How democratic will this country be in 10 years? (0–10) & How democratic will this country be in 10 years? (0–10) & Partial\\
Democratic past (0-10) & Not collected in W2 & How democratic was this country 10 years ago? (0–10) & How democratic was this country 10 years ago? (0–10) & How democratic was this country 10 years ago? (0–10) & Partial\\
Democratic present (0-10) & How democratic is this country today? (partial coverage, W2) & How democratic is this country today? (0–10 scale, government performance framing) & How democratic is this country today? (0–10 scale, government performance framing) & How democratic is this country today? (0–10 scale, government performance framing) & Partial\\
Witnessed corruption (binary) & Has any government official asked you for a bribe? (binary W2) & Have you or anyone in your family personally experienced or witnessed government corruption? (broader) & Have you or anyone in your family personally experienced or witnessed government corruption? & Have you or anyone in your family personally experienced or witnessed government corruption? & Partial\\
National govt corruption & How widespread is corruption in national government? (1–4) & How widespread is corruption in national government? (1–4) & How widespread is corruption in national government? (1–4) & How widespread is corruption in national government? (1–4) & Full\\
Local govt corruption & How widespread is corruption in local government? (1–4) & How widespread is corruption in local government? (1–4) & How widespread is corruption in local government? (1–4) & How widespread is corruption in local government? (1–4) & Full\\
Follows political news & How often do you follow political news? (1–5) & How often do you follow political news? (1–5) & How often do you follow political news? (1–5) & How often do you follow political news? (1–5) & Full\\
Internet news (harmonized) & How often do you get news from internet? (6-point scale) & How often do you get news from internet? (6-point scale) & How often do you get news from internet? (8-point scale — harmonized to 6pt) & How often do you get news from internet? (9-point scale — harmonized to 6pt) & Partial\\
Political interest & How interested are you in politics? (1–4) & How interested are you in politics? (1–4) & How interested are you in politics? (1–4) & How interested are you in politics? (1–4) & Full\\
Discusses politics & How often do you discuss politics with friends or family? (1–3) & How often do you discuss politics with friends or family? (1–3) & How often do you discuss politics with friends or family? (1–3) & How often do you discuss politics with friends or family? (1–3) & Full\\*

\end{longtable}

\end{ThreePartTable}
\endgroup{}

\newpage

\section*{Table A2: Subgroup Analysis of Key
Variables}\label{table-a2-subgroup-analysis-of-key-variables}
\addcontentsline{toc}{section}{Table A2: Subgroup Analysis of Key
Variables}

\begin{table}

\caption{\label{tbl-tableA2}Subgroup Analysis by Urban/Rural Residence}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{5.8cm}>{\raggedright\arraybackslash}p{5.8cm}}
\toprule
Wave & Urban & Rural\\
\midrule
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Gate: contacted influential person}}\\
\hspace{1em}Wave 2 (2008) & 38.6\% (32.2--45.3) & 44.3\% (40.8--47.9)\\
\hspace{1em}Wave 3 (2012) & 58.3\% (51.2--65.0) & 51.4\% (48.3--54.6)\\
\hspace{1em}Wave 4 (2015) & 65.5\% (58.4--72.0) & 54.0\% (50.9--57.1)\\
\hspace{1em}Wave 6 (2021) & 29.2\% (24.1--34.8) & 21.3\% (18.6--24.2)\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Democratic future (0-10 mean)}}\\
\hspace{1em}Wave 3 (2012) & 9.48 (9.22--9.74) & 9.60 (9.49--9.71)\\
\hspace{1em}Wave 4 (2015) & 7.32 (6.74--7.89) & 7.79 (7.53--8.05)\\
\hspace{1em}Wave 6 (2021) & 6.94 (6.53--7.36) & 6.57 (6.30--6.83)\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Witnessed corruption (\textbackslash{}\%)}}\\
\hspace{1em}Wave 2 (2008) & 34.7\% (28.5--41.5) & 25.7\% (22.6--28.9)\\
\hspace{1em}Wave 3 (2012) & 56.0\% (49.0--62.9) & 48.0\% (44.9--51.2)\\
\hspace{1em}Wave 4 (2015) & 65.6\% (58.3--72.3) & 62.3\% (59.0--65.4)\\
\hspace{1em}Wave 6 (2021) & 17.6\% (13.6--22.6) & 14.0\% (11.8--16.5)\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Single-party rule (1-4 mean)}}\\
\hspace{1em}Wave 2 (2008) & 1.53 (1.40--1.67) & 1.87 (1.78--1.95)\\
\hspace{1em}Wave 3 (2012) & 1.77 (1.64--1.91) & 2.01 (1.94--2.07)\\
\hspace{1em}Wave 4 (2015) & 1.64 (1.53--1.75) & 1.81 (1.75--1.86)\\
\hspace{1em}Wave 6 (2021) & 2.09 (1.98--2.21) & 2.24 (2.18--2.31)\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Voted in last election (\textbackslash{}\%)}}\\
\hspace{1em}Wave 3 (2012) & 70.9\% (64.1--76.9) & 80.3\% (77.7--82.7)\\
\hspace{1em}Wave 4 (2015) & 80.9\% (74.6--86.0) & 83.6\% (81.2--85.8)\\
\hspace{1em}Wave 6 (2021) & 92.1\% (88.4--94.8) & 86.8\% (84.3--88.9)\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Political interest (1-4 mean)}}\\
\hspace{1em}Wave 2 (2008) & 2.35 (2.21--2.48) & 2.62 (2.55--2.68)\\
\hspace{1em}Wave 3 (2012) & 2.44 (2.32--2.56) & 2.60 (2.54--2.65)\\
\hspace{1em}Wave 4 (2015) & 2.31 (2.17--2.45) & 2.29 (2.23--2.35)\\
\hspace{1em}Wave 6 (2021) & 1.94 (1.83--2.04) & 2.10 (2.04--2.16)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Estimates shown with 95\% CI in parentheses (Wilson interval for proportions; t-based for means). Gate proportions and voted/corruption displayed as percentages. '---' = item not fielded in wave. Wave 2 (2008), Wave 3 (2012), Wave 4 (2015), Wave 6 (2021).
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\begin{table}

\caption{\label{tbl-tableA2b}Subgroup Analysis by Age Group}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{3.8cm}>{\raggedright\arraybackslash}p{3.8cm}>{\raggedright\arraybackslash}p{3.8cm}}
\toprule
Wave & Under 30 & 30-49 & 50+\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Gate: contacted influential person}}\\
\hspace{1em}Wave 2 (2008) & 43.5\% (38.0--49.2) & 40.9\% (36.4--45.5) & 47.1\% (40.3--54.1)\\
\hspace{1em}Wave 3 (2012) & 50.6\% (45.6--55.7) & 53.6\% (49.2--57.9) & 53.5\% (47.5--59.4)\\
\hspace{1em}Wave 4 (2015) & 51.5\% (45.7--57.2) & 55.8\% (51.5--60.1) & 59.9\% (54.6--65.0)\\
\hspace{1em}Wave 6 (2021) & 22.5\% (17.7--28.2) & 24.8\% (21.3--28.7) & 21.5\% (17.4--26.4)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Democratic future (0-10 mean)}}\\
\hspace{1em}Wave 3 (2012) & 9.60 (9.43--9.77) & 9.53 (9.37--9.70) & 9.62 (9.43--9.82)\\
\hspace{1em}Wave 4 (2015) & 7.84 (7.38--8.30) & 7.42 (7.05--7.79) & 8.06 (7.63--8.49)\\
\hspace{1em}Wave 6 (2021) & 6.52 (6.07--6.97) & 6.93 (6.61--7.26) & 6.35 (5.94--6.77)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Witnessed corruption (\textbackslash{}\%)}}\\
\hspace{1em}Wave 2 (2008) & 29.7\% (24.7--35.2) & 28.2\% (24.3--32.6) & 23.3\% (17.8--29.8)\\
\hspace{1em}Wave 3 (2012) & 52.4\% (47.4--57.5) & 49.1\% (44.8--53.5) & 45.8\% (39.9--51.8)\\
\hspace{1em}Wave 4 (2015) & 59.9\% (53.9--65.6) & 67.1\% (62.7--71.1) & 58.9\% (53.3--64.3)\\
\hspace{1em}Wave 6 (2021) & 16.4\% (12.3--21.5) & 15.0\% (12.1--18.3) & 13.6\% (10.3--17.8)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Single-party rule (1-4 mean)}}\\
\hspace{1em}Wave 2 (2008) & 1.73 (1.61--1.86) & 1.85 (1.74--1.97) & 1.73 (1.56--1.90)\\
\hspace{1em}Wave 3 (2012) & 2.02 (1.92--2.13) & 1.95 (1.85--2.04) & 1.93 (1.80--2.06)\\
\hspace{1em}Wave 4 (2015) & 1.82 (1.72--1.92) & 1.69 (1.62--1.77) & 1.87 (1.77--1.96)\\
\hspace{1em}Wave 6 (2021) & 2.30 (2.18--2.42) & 2.19 (2.12--2.27) & 2.16 (2.06--2.26)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Voted in last election (\textbackslash{}\%)}}\\
\hspace{1em}Wave 3 (2012) & 52.4\% (47.4--57.5) & 88.9\% (85.8--91.4) & 95.8\% (92.6--97.7)\\
\hspace{1em}Wave 4 (2015) & 60.1\% (54.3--65.6) & 90.2\% (87.3--92.6) & 92.2\% (88.8--94.7)\\
\hspace{1em}Wave 6 (2021) & 80.3\% (74.9--84.8) & 90.2\% (87.4--92.5) & 90.7\% (87.1--93.4)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Political interest (1-4 mean)}}\\
\hspace{1em}Wave 2 (2008) & 2.54 (2.44--2.65) & 2.52 (2.44--2.61) & 2.65 (2.52--2.79)\\
\hspace{1em}Wave 3 (2012) & 2.55 (2.46--2.63) & 2.55 (2.47--2.62) & 2.65 (2.54--2.76)\\
\hspace{1em}Wave 4 (2015) & 2.14 (2.04--2.24) & 2.31 (2.23--2.39) & 2.40 (2.29--2.50)\\
\hspace{1em}Wave 6 (2021) & 2.08 (1.97--2.19) & 2.04 (1.96--2.11) & 2.08 (1.99--2.17)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Estimates shown with 95\% CI in parentheses (Wilson interval for proportions; t-based for means). Gate proportions and voted/corruption displayed as percentages. '---' = item not fielded in wave.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\begin{table}

\caption{\label{tbl-tableA2c}Subgroup Analysis by Education Level}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{3.8cm}>{\raggedright\arraybackslash}p{3.8cm}>{\raggedright\arraybackslash}p{3.8cm}}
\toprule
Wave & Primary or below & Secondary & Tertiary\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Gate: contacted influential person}}\\
\hspace{1em}Wave 2 (2008) & 43.1\% (39.3--46.9) & 44.2\% (38.5--50.1) & 46.2\% (27.1--66.3)\\
\hspace{1em}Wave 3 (2012) & 51.6\% (48.0--55.1) & 54.2\% (49.1--59.3) & 57.1\% (41.1--71.9)\\
\hspace{1em}Wave 4 (2015) & 55.0\% (51.3--58.5) & 57.9\% (52.9--62.8) & 54.5\% (39.0--69.3)\\
\hspace{1em}Wave 6 (2021) & 19.9\% (16.9--23.3) & 27.2\% (23.1--31.7) & 30.3\% (20.5--42.0)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Democratic future (0-10 mean)}}\\
\hspace{1em}Wave 3 (2012) & 9.51 (9.38--9.65) & 9.73 (9.59--9.87) & 9.38 (8.66--10.11)\\
\hspace{1em}Wave 4 (2015) & 7.93 (7.65--8.22) & 7.29 (6.83--7.74) & 7.47 (6.05--8.88)\\
\hspace{1em}Wave 6 (2021) & 6.77 (6.47--7.07) & 6.39 (6.04--6.74) & 7.34 (6.38--8.30)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Witnessed corruption (\textbackslash{}\%)}}\\
\hspace{1em}Wave 2 (2008) & 25.2\% (22.0--28.8) & 33.2\% (27.9--39.0) & 26.9\% (12.4--48.1)\\
\hspace{1em}Wave 3 (2012) & 46.5\% (42.9--50.1) & 51.8\% (46.7--57.0) & 81.0\% (65.4--90.9)\\
\hspace{1em}Wave 4 (2015) & 59.1\% (55.3--62.7) & 69.1\% (64.1--73.7) & 70.7\% (54.3--83.4)\\
\hspace{1em}Wave 6 (2021) & 13.9\% (11.5--16.9) & 16.3\% (13.0--20.2) & 15.6\% (8.7--26.0)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Single-party rule (1-4 mean)}}\\
\hspace{1em}Wave 2 (2008) & 1.94 (1.84--2.04) & 1.53 (1.42--1.65) & 1.19 (0.94--1.45)\\
\hspace{1em}Wave 3 (2012) & 2.12 (2.04--2.20) & 1.73 (1.62--1.83) & 1.36 (1.16--1.55)\\
\hspace{1em}Wave 4 (2015) & 1.91 (1.85--1.97) & 1.56 (1.48--1.64) & 1.53 (1.30--1.77)\\
\hspace{1em}Wave 6 (2021) & 2.22 (2.15--2.29) & 2.17 (2.08--2.25) & 2.34 (2.09--2.58)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Voted in last election (\textbackslash{}\%)}}\\
\hspace{1em}Wave 3 (2012) & 84.9\% (82.1--87.3) & 67.7\% (62.7--72.3) & 62.5\% (45.8--76.8)\\
\hspace{1em}Wave 4 (2015) & 84.8\% (82.0--87.3) & 81.0\% (76.7--84.7) & 74.4\% (58.5--86.0)\\
\hspace{1em}Wave 6 (2021) & 87.9\% (85.2--90.2) & 87.7\% (84.1--90.5) & 92.3\% (83.4--96.8)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Political interest (1-4 mean)}}\\
\hspace{1em}Wave 2 (2008) & 2.49 (2.42--2.57) & 2.72 (2.61--2.82) & 2.62 (2.22--3.01)\\
\hspace{1em}Wave 3 (2012) & 2.52 (2.46--2.59) & 2.65 (2.56--2.73) & 2.71 (2.44--2.99)\\
\hspace{1em}Wave 4 (2015) & 2.24 (2.18--2.31) & 2.37 (2.28--2.46) & 2.41 (2.07--2.75)\\
\hspace{1em}Wave 6 (2021) & 2.06 (1.99--2.13) & 2.07 (1.98--2.15) & 2.01 (1.79--2.23)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Estimates shown with 95\% CI in parentheses (Wilson interval for proportions; t-based for means). Gate proportions and voted/corruption displayed as percentages. '---' = item not fielded in wave.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A3: Placebo Battery---Non-Political
Variables}\label{table-a3-placebo-batterynon-political-variables}
\addcontentsline{toc}{section}{Table A3: Placebo Battery---Non-Political
Variables}

\begin{table}

\caption{\label{tbl-tableA3}Placebo Battery: Non-Political Variables by
Wave}

\centering{

\centering\begingroup\fontsize{10}{12}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4.0cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2.5cm}}
\toprule
label & W2 (2008) & W3 (2012) & W4 (2015) & W6 (2021)\\
\midrule
Interpersonal trust (binary) & 7.4\% (0.83) & 11.6\% (0.92) & 14.1\% (1.01) & 5.0\% (0.62)\\
Interpersonal trust (ordinal) & --- & 2.30 (0.026) & 2.44 (0.027) & 2.23 (0.022)\\
National pride & 3.60 (0.020) & 3.66 (0.017) & 3.63 (0.017) & 3.79 (0.015)\\
Pride in political system & --- & 3.03 (0.022) & 2.75 (0.024) & 2.95 (0.021)\\
HH income satisfaction & --- & 2.67 (0.029) & 1.91 (0.024) & 2.96 (0.028)\\
Family econ situation (now) & 2.92 (0.019) & 3.03 (0.020) & 3.05 (0.020) & 2.86 (0.020)\\
Family econ change (1yr) & 3.30 (0.026) & 3.10 (0.027) & 3.04 (0.024) & 3.27 (0.026)\\
National econ situation & 3.18 (0.026) & 3.33 (0.024) & 3.12 (0.023) & 3.05 (0.024)\\
Democracy satisfaction & 2.97 (0.024) & 2.98 (0.020) & 2.71 (0.023) & 3.10 (0.019)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Cell format: estimate (SE). Proportions (interpersonal trust) shown as percentage (SE in percentage points). Means shown as raw scale values (SE). '---' = item not fielded in wave. Items are non-political and should not be affected by opposition elimination; stable values support the domain-specificity of observed political attitude shifts.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A4: W3→W4 Partisan
Decomposition}\label{table-a4-w3w4-partisan-decomposition}
\addcontentsline{toc}{section}{Table A4: W3→W4 Partisan Decomposition}

\begin{table}

\caption{\label{tbl-tableA4}W3→W4 Partisan Decomposition of Key Outcomes
by Wave 3 Vote Choice}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{3.0cm}>{\centering\arraybackslash}p{3.0cm}>{\centering\arraybackslash}p{3.0cm}}
\toprule
Group & W3 (2012) & W4 (2015) & $\Delta$ W4$-$W3\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Contacted influential person (\textbackslash{}\%)}}\\
\hspace{1em}CPP voter & 53.7\% (50.0, 57.4) & 59.7\% (55.6, 63.6) & +5.9 pp\\
\hspace{1em}CNRP/SRP voter$^{\dagger}$ & 57.3\% (49.5, 64.8) & 54.5\% (48.7, 60.3) & -2.8 pp\\
\hspace{1em}Nonvoter & 45.1\% (38.9, 51.4) & 50.7\% (43.6, 57.8) & +5.7 pp\\
\hspace{1em}DK/Refuse & 58.0\% (43.3, 71.5) & 46.7\% (36.4, 57.4) & -11.3 pp\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Democratic future (0--10 mean)}}\\
\hspace{1em}CPP voter & 9.56 (9.43, 9.69) & 8.48 (8.19, 8.77) & -1.08\\
\hspace{1em}CNRP/SRP voter$^{\dagger}$ & 9.44 (9.09, 9.79) & 6.04 (5.52, 6.55) & -3.40***\\
\hspace{1em}Nonvoter & 9.69 (9.52, 9.86) & 7.80 (7.27, 8.33) & -1.89\\
\hspace{1em}DK/Refuse & 9.73 (9.34, 10.12) & 7.26 (6.15, 8.36) & -2.47\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Single-party rule acceptance (1--4)}}\\
\hspace{1em}CPP voter & 2.00 (1.92, 2.08) & 1.89 (1.82, 1.96) & -0.11\\
\hspace{1em}CNRP/SRP voter$^{\dagger}$ & 1.87 (1.71, 2.03) & 1.56 (1.47, 1.64) & -0.31\\
\hspace{1em}Nonvoter & 1.99 (1.86, 2.12) & 1.81 (1.70, 1.93) & -0.17\\
\hspace{1em}DK/Refuse & 1.78 (1.52, 2.04) & 1.70 (1.53, 1.87) & -0.09\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Political interest (1--4)}}\\
\hspace{1em}CPP voter & 2.57 (2.51, 2.63) & 2.25 (2.17, 2.32) & -0.32\\
\hspace{1em}CNRP/SRP voter$^{\dagger}$ & 2.74 (2.61, 2.88) & 2.60 (2.49, 2.71) & -0.14\\
\hspace{1em}Nonvoter & 2.53 (2.42, 2.64) & 2.04 (1.91, 2.17) & -0.49\\
\hspace{1em}DK/Refuse & 2.18 (1.91, 2.44) & 2.14 (1.96, 2.33) & -0.03\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Witnessed corruption (\textbackslash{}\%)}}\\
\hspace{1em}CPP voter & 44.8\% (41.2, 48.6) & 56.8\% (52.6, 60.9) & +11.9 pp\\
\hspace{1em}CNRP/SRP voter$^{\dagger}$ & 59.9\% (52.1, 67.2) & 76.3\% (70.7, 81.1) & +16.4 pp\\
\hspace{1em}Nonvoter & 55.1\% (48.8, 61.3) & 61.1\% (53.7, 68.0) & +5.9 pp\\
\hspace{1em}DK/Refuse & 50.0\% (36.6, 63.4) & 63.9\% (52.5, 73.9) & +13.9 pp\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Democracy satisfaction (1--4)}}\\
\hspace{1em}CPP voter & 3.12 (3.07, 3.16) & 3.00 (2.95, 3.05) & -0.12\\
\hspace{1em}CNRP/SRP voter$^{\dagger}$ & 2.53 (2.41, 2.66) & 2.21 (2.11, 2.30) & -0.33**\\
\hspace{1em}Nonvoter & 2.89 (2.82, 2.97) & 2.64 (2.53, 2.75) & -0.26\\
\hspace{1em}DK/Refuse & 2.94 (2.77, 3.11) & 2.55 (2.38, 2.72) & -0.39\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Estimates with 95\% CI in parentheses (Wilson interval for proportions; t-based for means). Proportions shown as percentages. $\Delta$ = W4 value minus W3 value. Groups defined by W3 reported vote choice: CPP = voted for winning party (Cambodia People's Party); CNRP/SRP = voted for losing party (Cambodia National Rescue Party / Sam Rainsy Party); Nonvoter = did not vote in the 2012 election; DK/Refuse = voted but declined to identify party or missing. $^{\dagger}$Stars in the $\Delta$ column for CNRP/SRP rows reflect significance of the Wave 4 $\times$ CNRP/SRP interaction term from pooled OLS on W3--W4 data (CPP as reference group): * $p < .05$, ** $p < .01$, *** $p < .001$. N per wave: W3 = 1,200; W4 = 1,200. Group Ns at W3: CPP = 722, CNRP/SRP = 172, Nonvoter = 255, DK/Refuse = 51.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A5: W6 DK/Refuse Group Attitudinal
Profile}\label{table-a5-w6-dkrefuse-group-attitudinal-profile}
\addcontentsline{toc}{section}{Table A5: W6 DK/Refuse Group Attitudinal
Profile}

\begin{table}

\caption{\label{tbl-tableA5}Wave 6 Attitudinal Profile: DK/Refuse
vs.~CPP and CNRP/SRP Voters}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5.0cm}>{\centering\arraybackslash}p{2.8cm}>{\centering\arraybackslash}p{2.8cm}>{\centering\arraybackslash}p{2.8cm}}
\toprule
Variable & CPP Voter & CNRP/SRP Voter & DK/Refuse\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Democratic Orientation}}\\
\hspace{1em}Democracy always preferable (\% = 1) & 68.2\% (64.6, 71.6) & 72.0\% (57.3, 83.3) & 59.3\% (52.1, 66.1)\\
\hspace{1em}Democracy is best form (1--4 mean) & 3.26 (3.22, 3.30) & 2.91 (2.73, 3.09) & 3.14 (3.05, 3.23)\\
\hspace{1em}Democracy vs. equality (1--5 mean) & 2.26 (2.15, 2.36) & 2.61 (2.23, 2.99) & 2.46 (2.28, 2.64)\\
\hspace{1em}Democracy satisfaction (1--4 mean) & 3.22 (3.17, 3.26) & 2.44 (2.23, 2.65) & 2.98 (2.91, 3.05)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Democratic Expectations}}\\
\hspace{1em}Democratic future (0--10 mean) & 7.06 (6.78, 7.34) & 4.96 (4.13, 5.79) & 5.86 (5.45, 6.27)\\
\hspace{1em}Democratic present (0--10 mean) & 6.15 (5.95, 6.35) & 4.04 (3.57, 4.52) & 5.26 (5.00, 5.51)\\
\hspace{1em}Democratic past (0--10 mean) & 4.92 (4.74, 5.10) & 3.96 (3.43, 4.48) & 4.50 (4.25, 4.75)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Authoritarian Preferences}}\\
\hspace{1em}Expert rule (1--4 mean) & 2.19 (2.12, 2.26) & 1.96 (1.72, 2.20) & 1.95 (1.83, 2.06)\\
\hspace{1em}Single-party rule (1--4 mean) & 2.27 (2.20, 2.34) & 1.89 (1.66, 2.13) & 2.05 (1.93, 2.17)\\
\hspace{1em}Strongman rule (1--4 mean) & 2.23 (2.16, 2.30) & 2.07 (1.81, 2.33) & 2.05 (1.93, 2.18)\\
\hspace{1em}Military rule (1--4 mean) & 2.24 (2.17, 2.31) & 2.28 (1.99, 2.57) & 2.18 (2.04, 2.33)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Civic Engagement}}\\
\hspace{1em}Contacted influential person (\%) & 24.9\% (21.9, 28.3) & 20.8\% (11.3, 34.5) & 20.4\% (15.3, 26.5)\\
\hspace{1em}Political interest (1--4 mean) & 2.09 (2.02, 2.15) & 2.37 (2.14, 2.60) & 1.93 (1.82, 2.04)\\
\hspace{1em}Follows political news (mean) & 2.05 (1.96, 2.14) & 2.48 (2.13, 2.84) & 1.96 (1.81, 2.11)\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Missingness}}\\
\hspace{1em}Nonresponse: democratic future (\%) & 62.6\% & 59.6\% & 65.1\%\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Wave 6 (2021) only. Groups defined by W6 reported vote choice. N: CPP voter = 794; CNRP/SRP voter = 57; DK/Refuse = 249. Estimates with 95\% CI in parentheses (Wilson interval for proportions; t-based for means). 'Democracy always preferable' = proportion selecting category 1 of 3-category forced-choice item (1 = always preferable, 2 = sometimes, 3 = sometimes non-dem better). 'Democracy vs. equality' (1--5): higher values = greater prioritization of economic equality over democracy. Nonresponse on democratic future item reported as raw percentage. CNRP/SRP N is small (n = 57); interpret CIs accordingly.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A6: OLS Attenuation Tests---Freedom
Perceptions}\label{table-a6-ols-attenuation-testsfreedom-perceptions}
\addcontentsline{toc}{section}{Table A6: OLS Attenuation Tests---Freedom
Perceptions}

\begin{table}

\caption{\label{tbl-tableA6}OLS Attenuation Tests: Wave Coefficients
With and Without Freedom Perceptions}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.0cm}>{\centering\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{3.5cm}>{\centering\arraybackslash}p{2.5cm}}
\toprule
Wave & Model A: Baseline & Model B: +Freedom & Attenuation (\%)\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Democratic future (0--10)}}\\
\hspace{1em}Wave 4 & -1.858 (0.120)*** & -1.796 (0.118)*** & 3.3\%\\
\hspace{1em}Wave 6 & -2.887 (0.128)*** & -2.876 (0.129)*** & 0.4\%\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Dem. satisfaction (1--4)}}\\
\hspace{1em}Wave 3 & 0.009 (0.031) & 0.023 (0.031) & -169.3\%\\
\hspace{1em}Wave 4 & -0.264 (0.031)*** & -0.232 (0.031)*** & 11.9\%\\
\hspace{1em}Wave 6 & 0.137 (0.031)*** & 0.159 (0.032)*** & -16.1\%\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Dem. preference (binary)}}\\
\hspace{1em}Wave 3 & -0.043 (0.021)* & -0.049 (0.021)* & -13.9\%\\
\hspace{1em}Wave 4 & 0.131 (0.021)*** & 0.122 (0.022)*** & 6.7\%\\
\hspace{1em}Wave 6 & 0.046 (0.022)* & 0.043 (0.022) & 8.1\%\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item OLS regression. DVs: Democratic future (0--10, W3/W4/W6, reference = W3); Democracy satisfaction (1--4, all waves, reference = W2); Democracy preference binary (1 = always preferable, reference = W2). Model A: wave dummies + age, gender, education, urbanicity. Model B adds perceived freedom of speech and freedom to organize. Cell format: coefficient (SE). Attenuation = \% reduction in $|$coefficient$|$ from Model A to Model B. * $p < .05$; ** $p < .01$; *** $p < .001$.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A7: Wave × Freedom Interaction
Models}\label{table-a7-wave-freedom-interaction-models}
\addcontentsline{toc}{section}{Table A7: Wave × Freedom Interaction
Models}

\begin{table}

\caption{\label{tbl-tableA7}Wave × Freedom of Speech Interaction
Coefficients}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{5.0cm}}
\toprule
Wave & Wave $\times$ Free Speech (high)\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Democratic future (0--10)}}\\
\hspace{1em}Wave 4 & -1.878 (0.249)***\\
\hspace{1em}Wave 6 & -1.181 (0.275)***\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Dem. satisfaction (1--4)}}\\
\hspace{1em}Wave 3 & 0.118 (0.065)\\
\hspace{1em}Wave 4 & -0.072 (0.064)\\
\hspace{1em}Wave 6 & -0.054 (0.067)\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Dem. preference (binary)}}\\
\hspace{1em}Wave 3 & -0.012 (0.045)\\
\hspace{1em}Wave 4 & 0.033 (0.045)\\
\hspace{1em}Wave 6 & -0.092 (0.047)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item OLS regression. DV regressed on wave dummies interacted with a high-freedom-of-speech indicator, plus perceived freedom to organize and controls (age, gender, education, urbanicity). High-freedom-of-speech indicator = 1 if respondent somewhat or strongly agrees that people are free to speak without fear (scale value 3 or 4 out of 4). Only the wave-by-freedom interaction terms are reported here. Reference wave: W3 for democratic future; W2 for satisfaction and preference. Cell format: coefficient (SE). * $p < .05$; ** $p < .01$; *** $p < .001$.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A8: Nonresponse Prediction---Logistic Regression (Wave
6)}\label{table-a8-nonresponse-predictionlogistic-regression-wave-6}
\addcontentsline{toc}{section}{Table A8: Nonresponse
Prediction---Logistic Regression (Wave 6)}

\begin{table}

\caption{\label{tbl-tableA8}Predictors of Nonresponse on Democratic
Future Item in Wave 6}

\centering{

\centering\begingroup\setstretch{1}\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.0cm}>{\centering\arraybackslash}p{3.0cm}>{\centering\arraybackslash}p{3.0cm}}
\toprule
Predictor & OR (95\% CI) & $p$\\
\midrule
Perceived freedom of speech (1--4) & 1.01 (0.82, 1.23) & 0.947NA\\
Perceived freedom to organize (1--4) & 0.90 (0.73, 1.12) & 0.362NA\\
Education (standardized) & 0.99 (0.88, 1.12) & 0.933NA\\
Urban residence (1 = urban) & 0.86 (0.65, 1.13) & 0.271NA\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Wave 6 (2021) only. N = 1,242. Dependent variable: 1 = missing on democratic future expectations item, 0 = responded. Missingness rate in W6: 63.1 percent. ORs above 1 indicate higher odds of nonresponse. * $p < .05$; ** $p < .01$; *** $p < .001$.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A9: Bounds Analysis---Democratic Future
Expectations}\label{table-a9-bounds-analysisdemocratic-future-expectations}
\addcontentsline{toc}{section}{Table A9: Bounds Analysis---Democratic
Future Expectations}

\begin{table}

\caption{\label{tbl-tableA9}Bounds Analysis: Democratic Future
Expectations Under Nonresponse Imputation Scenarios}

\centering{

\centering\begingroup\fontsize{9}{11}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6.5cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.4cm}>{\centering\arraybackslash}p{1.4cm}}
\toprule
Scenario & Wave & $N$ & Respondents & Mean & $\Delta$ from W3\\
\midrule
Observed (W3, 2012) & W3 & 1200 & 910 & 9.58 & —\\
\midrule
Observed (W6, respondents only) & W6 & 1242 & 458 & 6.67 & -2.91\\
Bound: nonrespondents imputed at max (10) & W6 & 1242 & 1242 & 8.77 & -0.80\\
Bound: nonrespondents imputed at min (0) & W6 & 1242 & 1242 & 2.46 & -7.12\\
Conservative: nonrespondents imputed at W3 mean & W6 & 1242 & 1242 & 8.51 & -1.07\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Dependent variable: democratic future expectations (0--10 scale). W3 (2012): N = 910 respondents of 1,200 (nonresponse 24.2 percent). W6 (2021): N = 458 respondents of 1,242 (nonresponse 63.1 percent). Imputation bounds assign W6 nonrespondents to the scale maximum (10), minimum (0), or the W3 population mean (9.58). Under the conservative imputation (nonrespondents at W3 mean), the observed decline of $-$2.91 points narrows to $-$1.07 points but remains in the same direction.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Table A10: Democracy vs.~Economic Development Trade-Off,
2008--2021}\label{table-a10-democracy-vs.-economic-development-trade-off-20082021}
\addcontentsline{toc}{section}{Table A10: Democracy vs.~Economic
Development Trade-Off, 2008--2021}

\begin{table}

\caption{\label{tbl-tableA10}Democracy vs.~Economic Development
Trade-Off Item, Wave Means (Cambodia)}

\centering{

\centering\begingroup\fontsize{10}{12}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{4.5cm}}
\toprule
Wave & Year & $N$ & Mean (95\% CI)\\
\midrule
W2 & 2008 & 941 & 2.43 (2.33, 2.53)\\
W3 & 2012 & 1194 & 2.11 (2.04, 2.18)\\
W4 & 2015 & 1191 & 2.18 (2.10, 2.26)\\
W6 & 2021 & 1174 & 2.25 (2.17, 2.34)\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Item: respondents choose between economic development and political democracy on a five-point scale (1 = economic development definitely more important, 5 = political democracy definitely more important), with a midpoint option. Wording is consistent across all four waves. Means reported with 95\% CI (t-based) in parentheses. The item is included here for completeness; it shows less dynamic variation than the democracy-vs-equality trade-off item reported in the main analysis.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%

\newpage

\section*{Appendix D: 2017 Cambodia Socio-Economic Survey --- Governance
Module}\label{appendix-d-2017-cambodia-socio-economic-survey-governance-module}
\addcontentsline{toc}{section}{Appendix D: 2017 Cambodia Socio-Economic
Survey --- Governance Module}

\begin{table}

\caption{\label{tbl-appendixD}Perceived Civic Freedoms and Corruption
Exposure, 2017 (SAIE)}

\centering{

\centering\begingroup\fontsize{10}{12}\selectfont

\begin{threeparttable}
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{9.0cm}>{\centering\arraybackslash}p{1.8cm}>{\centering\arraybackslash}p{2.5cm}}
\toprule
Item & \% Agree & 95\% CI\\
\midrule
Commune councillors generally helpful and responsive & 88.2 & {}[87.0, 89.2]\\
Village chief works for the benefit of citizens & 89.4 & {}[88.3, 90.4]\\
People free to speak what they think without fear & 86.1 & {}[84.9, 87.2]\\
People free to join any organisation without fear & 87.9 & {}[86.7, 88.9]\\
People free to vote for any political party & 94.0 & {}[93.1, 94.7]\\
\midrule
Paid additional money to officials to get things done faster (2016) & 34.9 & {}[33.3, 36.5]\\
\bottomrule
\end{tabular}
\begin{tablenotes}[para]
\item \textit{Note:} 
\item Source: Cambodia Socio-Economic Survey (KHM 2018 SAIE v01; N = 3,363). Fieldwork inferred early-to-mid 2017 based on internal temporal references (2016--17 school year as current; 2016 commune budget as most recently completed; tea money question references 2016 as preceding year). Binary response (Agree/Disagree); valid N ranges 3,285--3,339. 95\% CI: Wilson interval. The first five items measure perceived civic freedoms and local government responsiveness; the sixth (below the line) measures self-reported corruption exposure. These items are not comparable to ABS instruments and are used triangulatively to establish baseline civic conditions in the immediate pre-dissolution period.
\end{tablenotes}
\end{threeparttable}
\endgroup{}

}

\end{table}%




\end{document}
