---
title: "Online Appendix: [TITLE TBD]"
author:
  - name: Jeffrey Stark
date: now
date-format: "MMMM D, YYYY [at] h:mm A"
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{caption}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \usepackage{colortbl}
          \usepackage{xcolor}
          \captionsetup{width=\textwidth}
          \setlength\LTcapwidth{\textwidth}
          \floatplacement{figure}{H}
          \floatplacement{table}{H}
          \raggedright
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r setup}
#| include: false
library(tidyverse)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/04_cambodia_fairy_tale")
results_dir  <- file.path(paper_dir, "analysis/results")
fig_dir      <- file.path(paper_dir, "analysis/figures")
tbl_dir      <- file.path(paper_dir, "analysis/tables")
rr_dir       <- file.path(paper_dir, "analysis/reviewer_response")

source(file.path(paper_dir, "R", "helpers.R"))
```

## Table A1: Cross-Wave Comparability of Survey Items {.unnumbered}

```{r tbl-tableA1}
#| echo: false
comp <- read.csv(file.path(rr_dir, "comparability_matrix.csv"), stringsAsFactors = FALSE)

comp |>
  dplyr::select(label, w2_wording, w3_wording, w4_wording, w6_wording, status) |>
  kbl(
    col.names = c("Variable", "W2 (2008)", "W3 (2012)", "W4 (2015)", "W6 (2021)", "Status"),
    booktabs  = TRUE,
    linesep   = "",
    longtable = TRUE,
    escape    = TRUE
  ) |>
  kable_styling(
    latex_options = c("repeat_header"),
    font_size     = 7,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "2.0cm") |>
  column_spec(2, width = "2.3cm") |>
  column_spec(3, width = "2.3cm") |>
  column_spec(4, width = "2.3cm") |>
  column_spec(5, width = "2.3cm") |>
  column_spec(6, width = "1.5cm") |>
  footnote(
    general = "Status: Full = identical or functionally equivalent wording across all waves used; Partial = minor wording shift or item not fielded in all waves (cross-wave comparison valid with caveats noted in text). No items are classified as Excluded from all comparisons; wave-pair restrictions are noted in the main text.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    escape = FALSE,
    threeparttable = TRUE
  )
```

\newpage

## Table A2: Subgroup Analysis of Key Variables {.unnumbered}

```{r tbl-tableA2}
#| echo: false
#| tbl-cap: "Subgroup Analysis by Urban/Rural Residence"
A2 <- readRDS(file.path(tbl_dir, "tableA2_subgroups.rds"))

# Helper: print one panel as kbl
print_A2_panel <- function(df, title, footnote_txt) {
  data_cols <- setdiff(names(df), c("label", "wave_label"))
  tbl <- df |>
    dplyr::select(wave_label, all_of(data_cols)) |>
    kbl(
      col.names  = c("Wave", data_cols),
      booktabs   = TRUE,
      linesep    = "",
      escape     = FALSE
    ) |>
    kable_styling(
      latex_options = c("hold_position"),
      font_size     = 9,
      full_width    = FALSE
    ) |>
    column_spec(1, width = "2.5cm") |>
    column_spec(seq_along(data_cols) + 1L,
                width = paste0(round(11.5 / length(data_cols), 1), "cm")) |>
    pack_rows("Gate: contacted influential person", 1, 4) |>
    pack_rows("Democratic future (0-10 mean)", 5, 7) |>
    pack_rows("Witnessed corruption (\\%)", 8, 11) |>
    pack_rows("Single-party rule (1-4 mean)", 12, 15) |>
    pack_rows("Voted in last election (\\%)", 16, 18) |>
    pack_rows("Political interest (1-4 mean)", 19, 22) |>
    footnote(
      general           = footnote_txt,
      general_title     = "Note:",
      footnote_as_chunk = TRUE,
      escape            = FALSE,
      threeparttable    = TRUE
    )
  # Inject single-spacing so double-spaced document doesn't push rows off page
  tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl), fixed = TRUE)
  knitr::asis_output(tbl_str)
}

ci_note <- "Estimates shown with 95\\\\% CI in parentheses (Wilson interval for proportions; t-based for means). Gate proportions and voted/corruption displayed as percentages. '---' = item not fielded in wave. Wave 2 (2008), Wave 3 (2012), Wave 4 (2015), Wave 6 (2021)."

print_A2_panel(A2$urban_rural, "Subgroup Analysis by Urban/Rural Residence", ci_note)
```

```{r tbl-tableA2b}
#| echo: false
#| tbl-cap: "Subgroup Analysis by Age Group"
A2 <- readRDS(file.path(tbl_dir, "tableA2_subgroups.rds"))
ci_note <- "Estimates shown with 95\\\\% CI in parentheses (Wilson interval for proportions; t-based for means). Gate proportions and voted/corruption displayed as percentages. '---' = item not fielded in wave."

print_A2_panel <- function(df, title, footnote_txt) {
  data_cols <- setdiff(names(df), c("label", "wave_label"))
  tbl <- df |>
    dplyr::select(wave_label, all_of(data_cols)) |>
    kbl(
      col.names  = c("Wave", data_cols),
      booktabs   = TRUE,
      linesep    = "",
      escape     = FALSE
    ) |>
    kable_styling(
      latex_options = c("hold_position"),
      font_size     = 9,
      full_width    = FALSE
    ) |>
    column_spec(1, width = "2.5cm") |>
    column_spec(seq_along(data_cols) + 1L,
                width = paste0(round(11.5 / length(data_cols), 1), "cm")) |>
    pack_rows("Gate: contacted influential person", 1, 4) |>
    pack_rows("Democratic future (0-10 mean)", 5, 7) |>
    pack_rows("Witnessed corruption (\\%)", 8, 11) |>
    pack_rows("Single-party rule (1-4 mean)", 12, 15) |>
    pack_rows("Voted in last election (\\%)", 16, 18) |>
    pack_rows("Political interest (1-4 mean)", 19, 22) |>
    footnote(
      general           = footnote_txt,
      general_title     = "Note:",
      footnote_as_chunk = TRUE,
      escape            = FALSE,
      threeparttable    = TRUE
    )
  # Inject single-spacing so double-spaced document doesn't push rows off page
  tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl), fixed = TRUE)
  knitr::asis_output(tbl_str)
}

print_A2_panel(A2$age, "Subgroup Analysis by Age Group", ci_note)
```

```{r tbl-tableA2c}
#| echo: false
#| tbl-cap: "Subgroup Analysis by Education Level"
A2 <- readRDS(file.path(tbl_dir, "tableA2_subgroups.rds"))
ci_note <- "Estimates shown with 95\\\\% CI in parentheses (Wilson interval for proportions; t-based for means). Gate proportions and voted/corruption displayed as percentages. '---' = item not fielded in wave."

print_A2_panel <- function(df, title, footnote_txt) {
  data_cols <- setdiff(names(df), c("label", "wave_label"))
  tbl <- df |>
    dplyr::select(wave_label, all_of(data_cols)) |>
    kbl(
      col.names  = c("Wave", data_cols),
      booktabs   = TRUE,
      linesep    = "",
      escape     = FALSE
    ) |>
    kable_styling(
      latex_options = c("hold_position"),
      font_size     = 9,
      full_width    = FALSE
    ) |>
    column_spec(1, width = "2.5cm") |>
    column_spec(seq_along(data_cols) + 1L,
                width = paste0(round(11.5 / length(data_cols), 1), "cm")) |>
    pack_rows("Gate: contacted influential person", 1, 4) |>
    pack_rows("Democratic future (0-10 mean)", 5, 7) |>
    pack_rows("Witnessed corruption (\\%)", 8, 11) |>
    pack_rows("Single-party rule (1-4 mean)", 12, 15) |>
    pack_rows("Voted in last election (\\%)", 16, 18) |>
    pack_rows("Political interest (1-4 mean)", 19, 22) |>
    footnote(
      general           = footnote_txt,
      general_title     = "Note:",
      footnote_as_chunk = TRUE,
      escape            = FALSE,
      threeparttable    = TRUE
    )
  # Inject single-spacing so double-spaced document doesn't push rows off page
  tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl), fixed = TRUE)
  knitr::asis_output(tbl_str)
}

print_A2_panel(A2$education, "Subgroup Analysis by Education Level", ci_note)
```

\newpage

## Table A3: Placebo Battery---Non-Political Variables {.unnumbered}

```{r tbl-tableA3}
#| echo: false
#| tbl-cap: "Placebo Battery: Non-Political Variables by Wave"
A3 <- readRDS(file.path(tbl_dir, "tableA3_placebo.rds"))

A3 |>
  kbl(
    col.names = names(A3),
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 10,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "4.0cm") |>
  column_spec(2:5, width = "2.5cm") |>
  footnote(
    general = "Cell format: estimate (SE). Proportions (interpersonal trust) shown as percentage (SE in percentage points). Means shown as raw scale values (SE). '---' = item not fielded in wave. Items are non-political and should not be affected by opposition elimination; stable values support the domain-specificity of observed political attitude shifts.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    escape = FALSE,
    threeparttable = TRUE
  )
```

\newpage

## Table A4: W3→W4 Partisan Decomposition {.unnumbered}

```{r tbl-tableA4}
#| echo: false
#| tbl-cap: "W3→W4 Partisan Decomposition of Key Outcomes by Wave 3 Vote Choice"
A4 <- readRDS(file.path(tbl_dir, "tableA4_partisan_w3w4.rds"))

A4_display <- A4 |>
  dplyr::select(group, w3_fmt, w4_fmt, delta_fmt) |>
  dplyr::mutate(
    group = dplyr::recode(group,
      "CNRP/SRP voter" = "CNRP/SRP voter$^{\\dagger}$"
    )
  )

tbl_A4 <- A4_display |>
  kbl(
    col.names = c("Group", "W3 (2012)", "W4 (2015)",
                  "$\\Delta$ W4$-$W3"),
    align     = "lccc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 9,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "3.5cm") |>
  column_spec(2:4, width = "3.0cm") |>
  pack_rows("Contacted influential person (\\%)", 1, 4) |>
  pack_rows("Democratic future (0--10 mean)", 5, 8) |>
  pack_rows("Single-party rule acceptance (1--4)", 9, 12) |>
  pack_rows("Political interest (1--4)", 13, 16) |>
  pack_rows("Witnessed corruption (\\%)", 17, 20) |>
  pack_rows("Democracy satisfaction (1--4)", 21, 24) |>
  footnote(
    general = paste0(
      "Estimates with 95\\\\% CI in parentheses (Wilson interval for proportions; ",
      "t-based for means). Proportions shown as percentages. ",
      "$\\\\Delta$ = W4 value minus W3 value. Groups defined by W3 reported vote choice: ",
      "CPP = voted for winning party (Cambodia People's Party); ",
      "CNRP/SRP = voted for losing party (Cambodia National Rescue Party / Sam Rainsy Party); ",
      "Nonvoter = did not vote in the 2012 election; ",
      "DK/Refuse = voted but declined to identify party or missing. ",
      "$^{\\\\dagger}$Stars in the $\\\\Delta$ column for CNRP/SRP rows reflect significance of the ",
      "Wave 4 $\\\\times$ CNRP/SRP interaction term from pooled OLS on W3--W4 data ",
      "(CPP as reference group): * $p < .05$, ** $p < .01$, *** $p < .001$. ",
      "N per wave: W3 = 1,200; W4 = 1,200. Group Ns at W3: ",
      "CPP = 722, CNRP/SRP = 172, Nonvoter = 255, DK/Refuse = 51."
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    escape            = FALSE,
    threeparttable    = TRUE
  )

tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl_A4), fixed = TRUE)
knitr::asis_output(tbl_str)
```

\newpage

## Table A5: W6 DK/Refuse Group Attitudinal Profile {.unnumbered}

```{r tbl-tableA5}
#| echo: false
#| tbl-cap: "Wave 6 Attitudinal Profile: DK/Refuse vs. CPP and CNRP/SRP Voters"
A5_full <- readRDS(file.path(tbl_dir, "tableA5_dk_profile.rds"))
A5      <- A5_full$profile
A5_nr   <- A5_full$nonresponse

A5_display <- A5 |>
  dplyr::select(label, fmt_CPP_voter, fmt_CNRP_SRP_voter, fmt_DK_Refuse)

nr_cpp  <- A5_nr$na_dem_future[A5_nr$partisan == "CPP voter"]
nr_cnrp <- A5_nr$na_dem_future[A5_nr$partisan == "CNRP/SRP voter"]
nr_dk   <- A5_nr$na_dem_future[A5_nr$partisan == "DK/Refuse"]
nr_row  <- tibble::tibble(
  label              = "Nonresponse: democratic future (\\%)",
  fmt_CPP_voter      = sprintf("%.1f\\%%", nr_cpp),
  fmt_CNRP_SRP_voter = sprintf("%.1f\\%%", nr_cnrp),
  fmt_DK_Refuse      = sprintf("%.1f\\%%", nr_dk)
)
A5_display <- dplyr::bind_rows(A5_display, nr_row)

tbl_A5 <- A5_display |>
  kbl(
    col.names = c("Variable", "CPP Voter", "CNRP/SRP Voter", "DK/Refuse"),
    align     = "lccc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 9,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "5.0cm") |>
  column_spec(2:4, width = "2.8cm") |>
  pack_rows("Democratic Orientation", 1, 4) |>
  pack_rows("Democratic Expectations", 5, 7) |>
  pack_rows("Authoritarian Preferences", 8, 11) |>
  pack_rows("Civic Engagement", 12, 14) |>
  pack_rows("Missingness", 15, 15) |>
  footnote(
    general = paste0(
      "Wave 6 (2021) only. Groups defined by W6 reported vote choice. ",
      "N: CPP voter = 794; CNRP/SRP voter = 57; DK/Refuse = 249. ",
      "Estimates with 95\\\\% CI in parentheses (Wilson interval for proportions; ",
      "t-based for means). ",
      "'Democracy always preferable' = proportion selecting category 1 of 3-category ",
      "forced-choice item (1 = always preferable, 2 = sometimes, 3 = sometimes non-dem better). ",
      "'Democracy vs. equality' (1--5): higher values = greater prioritization of ",
      "economic equality over democracy. ",
      "Nonresponse on democratic future item reported as raw percentage. ",
      "CNRP/SRP N is small (n = 57); interpret CIs accordingly."
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    escape            = FALSE,
    threeparttable    = TRUE
  )

tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl_A5), fixed = TRUE)
knitr::asis_output(tbl_str)
```

\newpage

## Table A6: OLS Attenuation Tests---Freedom Perceptions {.unnumbered}

```{r tbl-tableA6}
#| echo: false
#| tbl-cap: "OLS Attenuation Tests: Wave Coefficients With and Without Freedom Perceptions"
mc <- read.csv(file.path(rr_dir, "freedom_models_coefs.csv"),
               stringsAsFactors = FALSE)

mc_sub <- mc |>
  dplyr::filter(grepl("^wave_f", term)) |>
  dplyr::mutate(
    dv_label = dplyr::recode(dv,
      "dem_country_future"    = "Democratic future (0--10)",
      "democracy_satisfaction" = "Dem. satisfaction (1--4)",
      "dem_pref_binary"       = "Dem. preference (binary)"
    ),
    wave_label  = paste0("Wave ", wave),
    model_label = dplyr::recode(model,
      "A_baseline"          = "A: No controls",
      "B_freedom_controls"  = "B: + Freedom"
    ),
    p_stars = dplyr::case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    cell = sprintf("%.3f (%.3f)%s", estimate, std.error, p_stars)
  )

mc_wide <- mc_sub |>
  dplyr::select(dv_label, wave_label, model_label, cell, attenuation_pct) |>
  tidyr::pivot_wider(names_from = model_label, values_from = cell) |>
  dplyr::mutate(
    attenuation_pct = dplyr::if_else(
      is.na(attenuation_pct), "\u2014",
      sprintf("%.1f\\%%", attenuation_pct)
    )
  ) |>
  dplyr::select(dv_label, wave_label, `A: No controls`, `B: + Freedom`, attenuation_pct)

dv_groups_mc <- rle(mc_wide$dv_label)

tbl_A6 <- mc_wide |>
  dplyr::select(-dv_label) |>
  kbl(
    col.names = c("Wave", "Model A: Baseline", "Model B: +Freedom", "Attenuation (\\%)"),
    align     = "lccc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 9,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "2.0cm") |>
  column_spec(2:3, width = "3.5cm") |>
  column_spec(4, width = "2.5cm")

row_idx <- 0
for (i in seq_along(dv_groups_mc$values)) {
  tbl_A6 <- pack_rows(tbl_A6, dv_groups_mc$values[i],
                       row_idx + 1, row_idx + dv_groups_mc$lengths[i])
  row_idx <- row_idx + dv_groups_mc$lengths[i]
}

tbl_A6 <- tbl_A6 |>
  footnote(
    general = paste0(
      "OLS regression. DVs: Democratic future (0--10, W3/W4/W6, reference = W3); ",
      "Democracy satisfaction (1--4, all waves, reference = W2); ",
      "Democracy preference binary (1 = always preferable, reference = W2). ",
      "Model A: wave dummies + age, gender, education, urbanicity. ",
      "Model B adds perceived freedom of speech and freedom to organize. ",
      "Cell format: coefficient (SE). ",
      "Attenuation = \\\\% reduction in $|$coefficient$|$ from Model A to Model B. ",
      "* $p < .05$; ** $p < .01$; *** $p < .001$."
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    escape            = FALSE,
    threeparttable    = TRUE
  )

tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl_A6), fixed = TRUE)
knitr::asis_output(tbl_str)
```

\newpage

## Table A7: Wave × Freedom Interaction Models {.unnumbered}

```{r tbl-tableA7}
#| echo: false
#| tbl-cap: "Wave × Freedom of Speech Interaction Coefficients"
mod <- read.csv(file.path(rr_dir, "freedom_moderator_coefs.csv"),
                stringsAsFactors = FALSE)

mod_int <- mod |>
  dplyr::filter(
    moderator == "free_speech_hi",
    grepl("wave_f.*free_speech_hi|free_speech_hi.*wave_f", term)
  ) |>
  dplyr::mutate(
    dv_label = dplyr::recode(dv,
      "dem_country_future"    = "Democratic future (0--10)",
      "democracy_satisfaction" = "Dem. satisfaction (1--4)",
      "dem_pref_binary"       = "Dem. preference (binary)"
    ),
    wave_label = dplyr::case_when(
      grepl("wave_f3", term) ~ "Wave 3",
      grepl("wave_f4", term) ~ "Wave 4",
      grepl("wave_f6", term) ~ "Wave 6",
      TRUE ~ term
    ),
    p_stars = dplyr::case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    cell = sprintf("%.3f (%.3f)%s", estimate, std.error, p_stars)
  ) |>
  dplyr::select(dv_label, wave_label, cell)

dv_groups_int <- rle(mod_int$dv_label)

tbl_A7 <- mod_int |>
  dplyr::select(wave_label, cell) |>
  kbl(
    col.names = c("Wave", "Wave $\\times$ Free Speech (high)"),
    align     = "lc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 9,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "2.5cm") |>
  column_spec(2, width = "5.0cm")

row_idx <- 0
for (i in seq_along(dv_groups_int$values)) {
  tbl_A7 <- pack_rows(tbl_A7, dv_groups_int$values[i],
                       row_idx + 1, row_idx + dv_groups_int$lengths[i])
  row_idx <- row_idx + dv_groups_int$lengths[i]
}

tbl_A7 <- tbl_A7 |>
  footnote(
    general = paste0(
      "OLS regression. DV regressed on wave dummies interacted with a high-freedom-of-speech ",
      "indicator, plus perceived freedom to organize and controls ",
      "(age, gender, education, urbanicity). ",
      "High-freedom-of-speech indicator = 1 if respondent somewhat or strongly agrees ",
      "that people are free to speak without fear (scale value 3 or 4 out of 4). ",
      "Only the wave-by-freedom interaction terms are reported here. ",
      "Reference wave: W3 for democratic future; W2 for satisfaction and preference. ",
      "Cell format: coefficient (SE). ",
      "* $p < .05$; ** $p < .01$; *** $p < .001$."
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    escape            = FALSE,
    threeparttable    = TRUE
  )

tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl_A7), fixed = TRUE)
knitr::asis_output(tbl_str)
```

\newpage

## Table A8: Nonresponse Prediction---Logistic Regression (Wave 6) {.unnumbered}

```{r tbl-tableA8}
#| echo: false
#| tbl-cap: "Predictors of Nonresponse on Democratic Future Item in Wave 6"
nr <- read.csv(file.path(rr_dir, "freedom_nonresponse_coefs.csv"),
               stringsAsFactors = FALSE)

nr_display <- nr |>
  dplyr::filter(term != "(Intercept)") |>
  dplyr::mutate(
    term_label = dplyr::recode(term,
      "dem_free_speech"      = "Perceived freedom of speech (1--4)",
      "gov_free_to_organize" = "Perceived freedom to organize (1--4)",
      "edu_z"                = "Education (standardized)",
      "urban"                = "Urban residence (1 = urban)"
    ),
    OR_fmt = sprintf("%.2f (%.2f, %.2f)", OR, OR_lo, OR_hi),
    p_fmt  = sprintf("%.3f%s", p.value, p_stars)
  ) |>
  dplyr::select(term_label, OR_fmt, p_fmt)

tbl_A8 <- nr_display |>
  kbl(
    col.names = c("Predictor", "OR (95\\% CI)", "$p$"),
    align     = "lcc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 9,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "6.0cm") |>
  column_spec(2:3, width = "3.0cm") |>
  footnote(
    general = paste0(
      "Wave 6 (2021) only. N = 1,242. Dependent variable: 1 = missing on ",
      "democratic future expectations item, 0 = responded. Missingness rate in W6: 63.1 percent. ",
      "ORs above 1 indicate higher odds of nonresponse. ",
      "* $p < .05$; ** $p < .01$; *** $p < .001$."
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    escape            = FALSE,
    threeparttable    = TRUE
  )

tbl_str <- sub("\\begingroup", "\\begingroup\\setstretch{1}", as.character(tbl_A8), fixed = TRUE)
knitr::asis_output(tbl_str)
```

\newpage

## Table A9: Bounds Analysis---Democratic Future Expectations {.unnumbered}

```{r tbl-tableA9}
#| echo: false
#| tbl-cap: "Bounds Analysis: Democratic Future Expectations Under Nonresponse Imputation Scenarios"
A9 <- readRDS(file.path(tbl_dir, "tableA9_bounds.rds"))

A9 |>
  kbl(
    col.names = c("Scenario", "Wave", "$N$", "Respondents",
                  "Mean", "$\\Delta$ from W3"),
    align     = "lccccc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 9,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "6.5cm") |>
  column_spec(2:6, width = "1.4cm") |>
  row_spec(1, hline_after = TRUE) |>
  footnote(
    general = paste0(
      "Dependent variable: democratic future expectations (0--10 scale). ",
      "W3 (2012): N = 910 respondents of 1,200 (nonresponse 24.2 percent). ",
      "W6 (2021): N = 458 respondents of 1,242 (nonresponse 63.1 percent). ",
      "Imputation bounds assign W6 nonrespondents to the scale maximum (10), ",
      "minimum (0), or the W3 population mean (9.58). ",
      "Under the conservative imputation (nonrespondents at W3 mean), ",
      "the observed decline of $-$2.91 points narrows to $-$1.07 points but remains ",
      "in the same direction."
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    escape            = FALSE,
    threeparttable    = TRUE
  )
```

\newpage

## Appendix D: 2017 Cambodia Socio-Economic Survey — Governance Module {.unnumbered}

```{r tbl-appendixD}
#| echo: false
#| tbl-cap: "Perceived Civic Freedoms and Corruption Exposure, 2017 (SAIE)"
# Data from survey-data-prep (tableD_saie_governance.rds); hardcoded pending delivery.
saie <- tibble::tribble(
  ~Item,                                                           ~Pct,   ~CI,
  "Commune councillors generally helpful and responsive",          "88.2", "[87.0, 89.2]",
  "Village chief works for the benefit of citizens",              "89.4", "[88.3, 90.4]",
  "People free to speak what they think without fear",            "86.1", "[84.9, 87.2]",
  "People free to join any organisation without fear",            "87.9", "[86.7, 88.9]",
  "People free to vote for any political party",                  "94.0", "[93.1, 94.7]",
  "Paid additional money to officials to get things done (2016)", "34.9", "[33.3, 36.5]"
)

saie |>
  kbl(
    col.names = c("Item", "\\% Agree", "95\\% CI"),
    align     = "lcc",
    booktabs  = TRUE,
    linesep   = "",
    escape    = FALSE
  ) |>
  kable_styling(
    latex_options = c("hold_position"),
    font_size     = 10,
    full_width    = FALSE
  ) |>
  column_spec(1, width = "9.0cm") |>
  column_spec(2, width = "1.8cm") |>
  column_spec(3, width = "2.5cm") |>
  row_spec(5, hline_after = TRUE) |>
  footnote(
    general = paste0(
      "Source: Cambodia Socio-Economic Survey (KHM 2018 SAIE v01; N = 3,363). ",
      "Fieldwork inferred early-to-mid 2017 based on internal temporal references ",
      "(2016--17 school year as current; 2016 commune budget as most recently completed; ",
      "tea money question references 2016 as preceding year). ",
      "Binary response (Agree/Disagree); valid N ranges 3,285--3,339. ",
      "95\\\\% CI: Wilson interval. ",
      "The first five items measure perceived civic freedoms and local government responsiveness; ",
      "the sixth (below the line) measures self-reported corruption exposure. ",
      "These items are not comparable to ABS instruments and are used triangulatively ",
      "to establish baseline civic conditions in the immediate pre-dissolution period."
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    escape            = FALSE,
    threeparttable    = TRUE
  )
```

\newpage

# References {.unnumbered}

::: {#refs}
:::
