---
title: "Online Appendix: The Limits of Borrowed Legitimacy"
subtitle: "Military Trust Collapse in Thailand, 2001–2022"
author:
  - name: Jeffrey Stark
    affiliation: Yonsei University, Seoul, South Korea
date: now
date-format: "MMMM D, YYYY [at] h:mm A"
format:
  pdf:
    keep-tex: true
  html:
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/econdev-authpref"
analysis_dir <- file.path(project_root, "papers/thailand-trust-collapse/analysis")
results_dir <- file.path(analysis_dir, "results")

d <- readRDS(file.path(analysis_dir, "thailand_panel.rds"))
res <- readRDS(file.path(results_dir, "all_results_summary.rds"))

h3_piecewise <- readRDS(file.path(results_dir, "h3_piecewise_slopes.rds"))
h3_quadratic <- readRDS(file.path(results_dir, "h3_quadratic.rds"))
h4_phil <- readRDS(file.path(results_dir, "h4_philippines_coercive_trust.rds"))
h5_taiwan <- readRDS(file.path(results_dir, "h5_taiwan_military_stability.rds"))
h6_ologit <- readRDS(file.path(results_dir, "h6_ordered_logit.rds"))
h6_subgroups <- readRDS(file.path(results_dir, "h6_subgroups.rds"))
h6_secondary <- readRDS(file.path(results_dir, "h6_secondary_institutions.rds"))
dem_commit <- readRDS(file.path(results_dir, "democratic_commitment.rds"))
```

This appendix presents supplementary analyses referenced in the main text. All models control for age (centered), gender, education (standardized), and urban residence unless otherwise noted.

# A1. Variable Descriptions

Table A1 describes the key variables used in the analysis, their original scales, and any transformations applied.

```{r tab-a1-variables}
var_table <- tribble(
  ~Variable, ~Source, ~`Original Scale`, ~Transformation, ~`Analysis Scale`,
  "trust_national_government", "ABS Q7", "1--4 (none -- a great deal)", "None", "1--4 (higher = more trust)",
  "trust_military", "ABS Q7", "1--4", "None", "1--4 (higher = more trust)",
  "trust_police", "ABS Q7", "1--4", "None", "1--4 (higher = more trust)",
  "trust_parliament", "ABS Q7", "1--4", "None", "1--4 (higher = more trust)",
  "trust_courts", "ABS Q7", "1--4", "None", "1--4 (higher = more trust)",
  "trust_political_parties", "ABS Q7", "1--4", "None", "1--4 (higher = more trust)",
  "trust_political (composite)", "Derived", "---", "Mean of govt, parliament, parties", "1--4",
  "trust_nonpolitical (composite)", "Derived", "---", "Mean of courts, military, police", "1--4",
  "military_rule", "ABS Q130", "1--4 (strongly disapprove -- strongly approve)", "None", "1--4",
  "reject_military", "Derived", "---", "5 -- military_rule", "1--4 (higher = more rejection)",
  "strongman_rule", "ABS Q131", "1--4", "None", "1--4",
  "reject_strongman", "Derived", "---", "5 -- strongman_rule", "1--4 (higher = more rejection)",
  "single_party_rule", "ABS Q132", "1--4", "None", "1--4",
  "reject_single_party", "Derived", "---", "5 -- single_party_rule", "1--4 (higher = more rejection)",
  "reject_authoritarian (composite)", "Derived", "---", "Mean of reject_military, reject_strongman, reject_single_party", "1--4 (higher = more rejection)",
  "dem_always_preferable", "ABS Q98", "1--3 (1 = democracy always preferable, 2 = authoritarian sometimes OK, 3 = doesn't matter)", "None", "1--3",
  "dem_commitment_01", "Derived", "---", "Rescaled from dem_always_preferable: 1 = always prefer democracy, 0.5 = sometimes OK with authoritarian, 0 = doesn't matter", "0--1 (higher = stronger democratic commitment)",
  "dem_vs_econ", "ABS Q119/Q126", "1--5 (1 = economy definitely more important, 2 = economy somewhat, 3 = democracy somewhat, 4 = democracy definitely, 5 = both equally)", "None", "1--5 (non-ordinal; 5 = midpoint)",
  "dem_priority_01", "Derived", "---", "Rescaled from dem_vs_econ: 0 = economy definitely, 0.25 = economy somewhat, 0.5 = both equally, 0.75 = democracy somewhat, 1 = democracy definitely", "0--1 (higher = stronger democracy priority)",
  "democracy_satisfaction", "ABS Q1", "1--4 (not at all -- very satisfied)", "None", "1--4 (higher = more satisfied)",
  "political_interest", "ABS Q22", "1--4 (not at all -- very interested)", "None; centered for interactions", "1--4 (higher = more interested)",
  "pol_discuss", "ABS Q24", "1--4 (never -- frequently)", "None; centered for interactions", "1--4 (higher = more frequent)",
  "econ_national_now", "ABS Q3", "1--5 (very bad -- very good)", "None", "1--5 (higher = better evaluation)",
  "age_centered", "ABS", "Years", "Centered at sample mean", "Continuous",
  "female", "ABS", "1 = male, 2 = female", "Recoded: 1 = female, 0 = male", "Binary",
  "education_z", "ABS", "Years of education", "Standardized (z-score)", "Continuous",
  "is_urban", "ABS", "1 = urban, 2 = rural", "Recoded: 1 = urban, 0 = rural", "Binary"
)

kbl(var_table,
    caption = "Variable descriptions and transformations",
    booktabs = TRUE, longtable = TRUE,
    label = "tab-a1-variables") %>%
  kable_styling(latex_options = c("repeat_header"),
                font_size = 8) %>%
  column_spec(1, width = "3.5cm") %>%
  column_spec(2, width = "1.2cm") %>%
  column_spec(3, width = "3cm") %>%
  column_spec(4, width = "4cm") %>%
  column_spec(5, width = "3cm")
```

# A2. Non-Linear Time Specifications

The main text models trust as a linear function of wave. Here we test whether non-linear specifications better capture Thailand's trajectory, using two approaches: piecewise period dummies aligned with Thailand's political timeline, and quadratic time trends.

## Piecewise Period Models

We define three political periods: pre-coup (Waves 1--2, 2001--2008), coup era (Waves 3--4, 2010--2016), and protest era (Waves 5--6, 2018--2022). These periods correspond to distinct phases of Thailand's political trajectory described in the main text.

```{r tab-a1-piecewise}
pw_key <- bind_rows(
  h3_piecewise$govt$tidy %>%
    filter(str_detect(term, "period|country")) %>%
    mutate(outcome = "Government Trust"),
  h3_piecewise$mil$tidy %>%
    filter(str_detect(term, "period|country")) %>%
    mutate(outcome = "Military Trust")
) %>%
  select(outcome, term, estimate, std.error, p.value) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

pw_key %>%
  kbl(caption = "Table A1: Piecewise period models. Reference period: pre-coup (Waves 1--2); reference country: Thailand. Weighted OLS with survey weights; SEs are heteroskedasticity-consistent. N = 21,036 (government), 21,049 (military).",
      col.names = c("Outcome", "Term", "Estimate", "SE", "p"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kable_styling(font_size = 9)
```

```{r pw-inline}
pw_govt_protest <- h3_piecewise$govt$tidy %>% filter(term == "periodprotest_era")
pw_mil_protest <- h3_piecewise$mil$tidy %>% filter(term == "periodprotest_era")
pw_govt_coup <- h3_piecewise$govt$tidy %>% filter(term == "periodcoup_era")
pw_mil_coup <- h3_piecewise$mil$tidy %>% filter(term == "periodcoup_era")
```

Table A1 reveals a clear pattern of acceleration. For government trust, the coup-era decline relative to the pre-coup baseline ($b$ = `r round(pw_govt_coup$estimate, 3)`) is substantially smaller than the protest-era decline ($b$ = `r round(pw_govt_protest$estimate, 3)`). For military trust, the asymmetry is even more striking: the coup era shows a modest decline ($b$ = `r round(pw_mil_coup$estimate, 3)`), while the protest era produces a much larger drop ($b$ = `r round(pw_mil_protest$estimate, 3)`). The country interactions confirm that this acceleration pattern is Thailand-specific.

## Quadratic Specifications

```{r tab-a1-quadratic}
q_key <- bind_rows(
  h3_quadratic$govt$tidy %>%
    filter(str_detect(term, "wave|country")) %>%
    mutate(outcome = "Government Trust"),
  h3_quadratic$mil$tidy %>%
    filter(str_detect(term, "wave|country")) %>%
    mutate(outcome = "Military Trust")
) %>%
  select(outcome, term, estimate, std.error, p.value) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

q_key %>%
  kbl(caption = "Table A1b: Quadratic time specification with country interactions. Weighted OLS with survey weights; SEs are heteroskedasticity-consistent. N = 21,036 (government), 21,049 (military).",
      col.names = c("Outcome", "Term", "Estimate", "SE", "p"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kable_styling(font_size = 9)
```

The quadratic specifications (Table A1b) confirm that Thailand's trust trajectory is not purely linear. For both government and military trust, the quadratic term and its country interactions suggest accelerating decline in later waves. However, the quadratic specification does not capture the period-specific dynamics as cleanly as the piecewise model, likely because the 2020--2021 disruption represents a discrete shock rather than a smooth acceleration.

# A3. Subgroup Analyses

To rule out the possibility that the aggregate patterns are driven by compositional shifts across waves—for example, if later waves disproportionately sampled younger, more urban, or more educated respondents who tend to report lower trust—we estimate the baseline model separately for demographic subgroups.

```{r tab-a3-subgroups}
sub_summary <- h6_subgroups$summary %>%
  mutate(
    institution = if_else(str_detect(subgroup, "_govt$"), "Government", "Military"),
    group = str_remove(subgroup, "_(govt|mil)$") %>%
      str_replace("age_", "Age: ") %>%
      str_replace("edu_", "Education: ") %>%
      str_to_title()
  ) %>%
  select(group, institution, n, thailand_wave_slope, se, p_value) %>%
  mutate(across(where(is.numeric) & !matches("^n$"), ~round(., 3))) %>%
  arrange(group, institution)

sub_summary %>%
  kbl(caption = "Table A2: Thailand's per-wave trust slope by demographic subgroup. All models include country × wave interactions; the reported slope is Thailand's (reference category). Weighted OLS with survey weights; SEs are heteroskedasticity-consistent. N ranges from 7,411 to 13,124 depending on subgroup.",
      col.names = c("Subgroup", "Institution", "N", "Slope", "SE", "p"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  kable_styling(font_size = 9)
```

```{r subgroup-inline}
edu_high_mil_slope <- sub_summary$thailand_wave_slope[
  sub_summary$group == "Education: High" & sub_summary$institution == "Military"
]
```

Table A2 confirms that the military-government differential holds across all demographic subgroups. In every subgroup, the military trust slope is more negative than the government trust slope. Several patterns merit note:

- *Age*: Younger respondents show the steepest declines for both institutions, but the military-government gap is present across all age terciles.
- *Urbanization*: Urban respondents decline more steeply than rural respondents, consistent with the urban character of the 2020--2021 protests. However, rural respondents also show significant declines.
- *Education*: The most educated respondents show the steepest military trust decline ($b$ = `r round(edu_high_mil_slope, 3)`), consistent with educated citizens being more attentive to democratic norms and more critical of military intervention.

The consistency of the pattern across subgroups rules out compositional explanations and strengthens the inference that the military-government differential reflects a genuine shift in public attitudes.

# A4. Philippines: Coercive Trust Under Duterte

The Philippines provides a useful contrast case. Under Rodrigo Duterte (2016--2022), the government pursued a highly visible drug war that relied heavily on police and military enforcement. If citizens who support a coercive policy agenda maintain trust in the institutions that implement it, we should observe stable or rising coercive trust during the Duterte period.

```{r tab-a4-philippines}
phil_key <- bind_rows(
  h4_phil$phil_military$tidy %>% mutate(outcome = "Military"),
  h4_phil$phil_police$tidy %>% mutate(outcome = "Police"),
  h4_phil$phil_govt$tidy %>% mutate(outcome = "Government")
) %>%
  filter(term == "wave_num") %>%
  select(outcome, estimate, std.error, p.value) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

phil_key %>%
  kbl(caption = "Table A3: Philippines trust trends during the Duterte period (Waves 4--6 only). Wave coefficient captures per-wave change. Weighted OLS with survey weights; SEs are heteroskedasticity-consistent. N = 3,517--3,571 depending on institution.",
      col.names = c("Institution", "Wave slope", "SE", "p"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position"))
```

```{r phil-inline}
phil_mil_b <- h4_phil$phil_military$tidy %>% filter(term == "wave_num")
phil_police_b <- h4_phil$phil_police$tidy %>% filter(term == "wave_num")
phil_govt_b <- h4_phil$phil_govt$tidy %>% filter(term == "wave_num")
```

Table A3 shows that during the Duterte period, Philippine military trust trends slightly positive ($b$ = `r round(phil_mil_b$estimate, 3)`, *p* = `r round(phil_mil_b$p.value, 3)`), and police trust shows a similar direction ($b$ = `r round(phil_police_b$estimate, 3)`, *p* = `r round(phil_police_b$p.value, 3)`). This pattern is consistent with a preference-alignment mechanism: when the political leadership endorses coercive institutions and a substantial portion of the public supports the associated policy agenda, trust in those institutions can remain stable even during periods of democratic backsliding.

The contrast with Thailand is instructive. In Thailand, the military's political role generated mass opposition and targeted trust erosion. In the Philippines, the coercive institutions operated with at least partial popular support for their core mission (anti-drug enforcement), insulating them from the kind of backlash observed in Thailand.

# A5. Taiwan: Military Depoliticization

Taiwan presents the opposite theoretical case: a successfully depoliticized military. Following democratization in the 1990s, the Taiwanese military transitioned from a party-state instrument to a professional defense force under firm civilian control. If depoliticization insulates military trust from political volatility, we should observe a flat trajectory.

```{r tab-a5-taiwan}
tw_key <- h5_taiwan$tidy %>%
  select(term, estimate, std.error, p.value) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

tw_key %>%
  kbl(caption = paste0("Table A4: Taiwan military trust model (all waves). Weighted OLS with survey weights; SEs are heteroskedasticity-consistent. N = 7,930. F-test for wave coefficient = 0: p = ", round(h5_taiwan$ftest[["Pr(>F)"]][2], 3), "."),
      col.names = c("Term", "Estimate", "SE", "p"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position"))
```

```{r tw-inline}
tw_wave <- h5_taiwan$tidy %>% filter(term == "wave_num")
tw_ftest_p <- h5_taiwan$ftest[["Pr(>F)"]][2]
```

Table A4 confirms the depoliticization prediction. The wave coefficient for Taiwan's military trust is near zero ($b$ = `r round(tw_wave$estimate, 3)`) and not statistically significant (*p* = `r round(tw_wave$p.value, 3)`). An F-test fails to reject the null hypothesis that the wave effect equals zero (*p* = `r round(tw_ftest_p, 3)`). Taiwan's military trust is essentially flat across two decades, despite substantial political turbulence (cross-strait tensions, party alternation, pandemic governance debates).

This stability contrasts sharply with Thailand and supports the theoretical claim that the military's *political role*, rather than general political instability, drives targeted trust erosion. Where the military remains above the political fray, its institutional trust is insulated from the partisan dynamics that corrode trust in elected institutions.

# A6. Full Secondary Institution Results

```{r tab-a6-secondary-full}
# Full interaction model results for all institutions
sec_full <- map_dfr(h6_secondary$models, function(mod) {
  mod$tidy %>%
    filter(str_detect(term, "wave_num|country")) %>%
    mutate(institution = str_remove(mod$dv, "trust_") %>%
             str_replace_all("_", " ") %>%
             str_to_title())
}) %>%
  select(institution, term, estimate, std.error, p.value) %>%
  mutate(across(where(is.numeric), ~round(., 3)))

sec_full %>%
  kbl(caption = "Table A5: Country × Wave interaction models for secondary trust measures (courts, police, parliament, political parties). Thailand is the reference category. Weighted OLS with survey weights; SEs are heteroskedasticity-consistent. N ranges from 20,630 (courts) to 21,465 (police).",
      col.names = c("Institution", "Term", "Estimate", "SE", "p"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kable_styling(font_size = 8)
```

Table A5 reports the full interaction models for the four secondary trust measures. Several patterns are worth noting. First, Thailand's decline is significant for all four institutions, but the magnitudes are substantially smaller than for military trust (main text, Table 7). Second, the Philippines and Taiwan interaction terms are generally positive, confirming Thailand's exceptionalism extends beyond the two primary institutions. Third, political parties and parliament show the smallest declines, possibly because these institutions already had lower baseline trust levels, leaving less room for erosion.

# A7. Institutional Breadth: NGOs, Local Government, National Government, and Military

To evaluate whether preference falsification inflates Philippine trust during the Duterte era, or whether Thailand's trust collapse is targeted at specific institutions rather than systemic, we compare trust in four institutional categories across all six waves.

```{r tab-a7-breadth}
ib <- readRDS(file.path(results_dir, "institutional_breadth.rds"))

ib_table <- ib %>%
  mutate(across(c(trust_ngo, trust_local_govt, trust_natl_govt, trust_military),
                ~round(., 2))) %>%
  select(country_name, wave, trust_ngo, trust_local_govt, trust_natl_govt, trust_military, n_total) %>%
  arrange(country_name, wave)

ib_table %>%
  kbl(caption = "Table A6: Mean trust in NGOs, local government, national government, and military by country and wave. Scale: 1 (none at all) to 4 (a great deal). Descriptive means (unweighted). N is the total number of respondents per country-wave cell; item-level missingness varies slightly across institutions.",
      col.names = c("Country", "Wave", "NGOs", "Local Govt", "Nat'l Govt", "Military", "N"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position")) %>%
  kable_styling(font_size = 9) %>%
  pack_rows("Philippines", 1, 6) %>%
  pack_rows("Thailand", 7, 12)
```

Table A6 shows that in the Philippines, the Wave 5 trust surge is uniform across all four institutional categories, including NGOs and local government, which are not politically sensitive in the same manner as the national government and military. This uniformity is inconsistent with a preference falsification account, which would predict selective inflation of sensitive measures only. In Thailand, Wave 6 trust erosion is similarly broad-based: NGOs and local government decline alongside national government and military trust, indicating systemic rather than institution-specific disillusionment.

# A8. Democratic Commitment and Military Trust Decline

Table A7 presents models testing whether democratic commitment moderates trust erosion. Models 1 and 1b examine the interaction between wave and democratic commitment (centered) for military and government trust in Thailand. Model 2 uses a stacked outcome (military and government trust) to test whether the commitment effect differs across institution type. Model 3 reports a Wave 6 cross-sectional model predicting military trust from democratic commitment and political interest. Model 4 tests the three-country interaction.

```{r tab-a7-dem-commitment}
# Build table from saved model tidy results
mod1 <- dem_commit$mod1_mil$tidy %>%
  filter(term %in% c("dem_commit_c", "wave_num:dem_commit_c")) %>%
  mutate(Model = "1", DV = "Military trust (Thailand)",
         Term = case_when(
           term == "dem_commit_c" ~ "Democratic commitment",
           term == "wave_num:dem_commit_c" ~ "Wave × Commitment"
         ))

mod1b <- dem_commit$mod1b_govt$tidy %>%
  filter(term %in% c("dem_commit_c", "wave_num:dem_commit_c")) %>%
  mutate(Model = "1b", DV = "Government trust (Thailand)",
         Term = case_when(
           term == "dem_commit_c" ~ "Democratic commitment",
           term == "wave_num:dem_commit_c" ~ "Wave × Commitment"
         ))

mod2 <- dem_commit$mod2_stacked$tidy %>%
  filter(term == "wave_num:dem_commit_c:is_military") %>%
  mutate(Model = "2", DV = "Stacked trust (Thailand)",
         Term = "Wave × Commitment × Military")

mod3 <- dem_commit$mod3_w6_xsec$tidy %>%
  filter(term %in% c("democratic_commitment", "political_interest")) %>%
  mutate(Model = "3", DV = "Military trust, W6 (Thailand)",
         Term = case_when(
           term == "democratic_commitment" ~ "Democratic commitment",
           term == "political_interest" ~ "Political interest"
         ))

mod4 <- dem_commit$mod4_threecountry$tidy %>%
  filter(term %in% c("wave_num:dem_commit_c:country_namePhilippines",
                      "wave_num:dem_commit_c:country_nameTaiwan")) %>%
  mutate(Model = "4", DV = "Military trust (three-country)",
         Term = case_when(
           str_detect(term, "Philippines") ~ "Wave × Commitment × Philippines",
           str_detect(term, "Taiwan") ~ "Wave × Commitment × Taiwan"
         ))

dc_table <- bind_rows(mod1, mod1b, mod2, mod3, mod4) %>%
  mutate(
    p_display = case_when(
      p.value < .001 ~ "< .001",
      TRUE ~ as.character(round(p.value, 3))
    )
  ) %>%
  select(Model, DV, Term, estimate, std.error, p_display) %>%
  mutate(across(c(estimate, std.error), ~round(., 3)))

dc_table %>%
  kbl(caption = "Table A7: Democratic commitment and institutional trust decline. Democratic commitment is a composite index (0--1) of democratic preference, democratic priority over economic development, and rejection of authoritarian alternatives, centered at the sample mean for interaction models (Models 1, 1b, 2, 4). All models are weighted OLS with survey weights; SEs are heteroskedasticity-consistent except Model 2 (respondent-clustered SEs). N: Model 1 = 5,947; Model 1b = 5,889; Model 2 = 11,836 (stacked); Model 3 = 1,082; Model 4 = 21,261.",
      col.names = c("Model", "Dependent Variable", "Key Coefficient", "b", "SE", "p"),
      booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position", "scale_down")) %>%
  kable_styling(font_size = 8)
```
