---
title: "Main Analysis"
subtitle: "Winner/Loser Effect on Procedural Democracy"
author: "Jeffrey Stark"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    df-print: paged
execute:
  warning: false
  message: false
---

# Overview

This script presents the main analysis of the loser effect on procedural democracy preferences:

1. Overall loser effect by wave
2. Country-level variation
3. Trajectory analysis for countries with 3+ waves
4. Case studies: Thailand, South Korea
5. Pooled regression models

# Setup

```{r setup}
library(tidyverse)
library(broom)
library(here)
library(knitr)
library(kableExtra)

here::i_am("papers/02_meaning_of_democracy_revision/analysis/02_main_analysis.qmd")

# Load analysis data (created by 01_data_preparation.qmd)
df <- readRDS(here("papers", "meaning-of-democracy", "analysis", "data", 
                   "winner_loser_4waves.rds"))

# Create binary loser variable for regression
df <- df %>%
  mutate(loser = if_else(winner_loser == "Loser", 1L, 0L))

cat("Analysis dataset:", nrow(df), "observations\n")
cat("Waves:", paste(unique(df$wave), collapse = ", "), "\n")
cat("Countries:", n_distinct(df$country_name), "\n")
```

# Part 1: Overall Loser Effect by Wave

## Descriptive Statistics

```{r wave-descriptives}
by_wave <- df %>%
  group_by(wave, wave_year, winner_loser) %>%
  summarise(
    n = n(),
    pct_procedural = mean(procedural_single) * 100,
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = winner_loser, 
    values_from = c(n, pct_procedural)
  ) %>%
  mutate(
    loser_effect = round(pct_procedural_Loser - pct_procedural_Winner, 1),
    total_n = n_Winner + n_Loser
  ) %>%
  arrange(wave_year)

by_wave %>%
  select(wave, wave_year, total_n, 
         `% Proc (Winner)` = pct_procedural_Winner,
         `% Proc (Loser)` = pct_procedural_Loser,
         `Loser Effect` = loser_effect) %>%
  kable(digits = 1, caption = "Procedural Orientation by Winner/Loser Status") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

## Statistical Tests by Wave

```{r wave-tests}
wave_tests <- df %>%
  group_by(wave) %>%
  summarise(
    n = n(),
    loser_pct = mean(procedural_single[winner_loser == "Loser"]) * 100,
    winner_pct = mean(procedural_single[winner_loser == "Winner"]) * 100,
    loser_effect = loser_pct - winner_pct,
    chi_sq = chisq.test(table(winner_loser, procedural_single))$statistic,
    p_value = chisq.test(table(winner_loser, procedural_single))$p.value,
    .groups = "drop"
  )

wave_tests %>%
  mutate(
    sig = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      TRUE ~ ""
    )
  ) %>%
  select(wave, n, loser_effect, chi_sq, p_value, sig) %>%
  kable(digits = c(0, 0, 1, 2, 4, 0),
        col.names = c("Wave", "N", "Loser Effect (pp)", "χ²", "p-value", ""),
        caption = "Chi-Square Tests of Loser Effect by Wave") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

**Key Finding**: The loser effect pattern shows losers are consistently more likely to prioritize procedural democracy than winners in most waves.

# Part 2: Country-Level Analysis

## Loser Effect by Country and Wave

```{r country-wave}
by_country_wave <- df %>%
  group_by(country_name, wave, wave_year) %>%
  filter(n() >= 50) %>%  # Minimum sample size for reliable estimates
  summarise(
    n = n(),
    pct_winner = round(sum(winner_loser == "Winner") / n() * 100, 1),
    pct_proc_winner = mean(procedural_single[winner_loser == "Winner"]) * 100,
    pct_proc_loser = mean(procedural_single[winner_loser == "Loser"]) * 100,
    loser_effect = round(pct_proc_loser - pct_proc_winner, 1),
    .groups = "drop"
  )

# Countries with 3+ waves for trajectory analysis
countries_3plus <- by_country_wave %>%
  count(country_name) %>%
  filter(n >= 3) %>%
  pull(country_name)

cat("Countries with 3+ waves:", paste(countries_3plus, collapse = ", "), "\n")
```

## Full Country × Wave Panel

```{r country-panel}
by_country_wave %>%
  select(country_name, wave, n, pct_winner, loser_effect) %>%
  arrange(country_name, wave) %>%
  kable(col.names = c("Country", "Wave", "N", "% Winners", "Loser Effect (pp)"),
        caption = "Loser Effect by Country and Wave") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  scroll_box(height = "400px")
```

## Trajectory Table (Countries with 3+ Waves)

```{r trajectory}
trajectory <- by_country_wave %>%
  filter(country_name %in% countries_3plus) %>%
  select(country_name, wave, loser_effect) %>%
  pivot_wider(names_from = wave, values_from = loser_effect)

# Calculate change from earliest to latest available wave
trajectory <- trajectory %>%
  rowwise() %>%
  mutate(
    earliest = coalesce(W2, W3),
    latest = coalesce(W6, W4),
    total_change = latest - earliest
  ) %>%
  ungroup() %>%
  arrange(desc(total_change))

trajectory %>%
  kable(digits = 1,
        caption = "Loser Effect Trajectory by Country (percentage points)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

# Part 3: Case Studies

## Thailand: The Democratic Erosion Arc

```{r thailand}
thailand <- by_country_wave %>%
  filter(country_name == "Thailand") %>%
  arrange(wave_year)

if (nrow(thailand) > 0) {
  thailand %>%
    mutate(
      context = case_when(
        wave == "W2" ~ "Post-2006 coup, interim govt",
        wave == "W3" ~ "Democrat government (Abhisit)",
        wave == "W4" ~ "May 2014 COUP, military govt",
        wave == "W6" ~ "Military-backed civilian govt",
        TRUE ~ ""
      )
    ) %>%
    select(wave, wave_year, context, n, pct_winner, loser_effect) %>%
    kable(col.names = c("Wave", "Year", "Political Context", "N", "% Winners", "Loser Effect"),
          caption = "Thailand: Loser Effect Over Time") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
  
  cat("\n**Thailand Interpretation**:\n")
  cat("As democracy eroded through successive coups, losers increasingly valued procedural democracy.\n")
  cat("The loser effect grew dramatically as citizens experienced democratic backsliding.\n")
} else {
  cat("Thailand data not available in the current dataset.\n")
}
```

## South Korea: Power Alternation

```{r korea}
korea <- by_country_wave %>%
  filter(country_name == "South Korea") %>%
  arrange(wave_year)

if (nrow(korea) > 0) {
  korea %>%
    mutate(
      context = case_when(
        wave == "W2" ~ "Roh Moo-hyun (progressive)",
        wave == "W3" ~ "Lee Myung-bak (conservative)",
        wave == "W4" ~ "Park Geun-hye (conservative)", 
        wave == "W6" ~ "Post-Moon Jae-in (progressive)",
        TRUE ~ ""
      )
    ) %>%
    select(wave, wave_year, context, n, pct_winner, loser_effect) %>%
    kable(col.names = c("Wave", "Year", "Context", "N", "% Winners", "Loser Effect"),
          caption = "South Korea: Loser Effect Over Time") %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
  
  cat("\n**South Korea Interpretation**:\n")
  cat("Shows the opposite trajectory from Thailand.\n")
  cat("Under conservative governments, progressive 'losers' showed large procedural orientation gaps.\n")
  cat("After Moon Jae-in's progressive government and power alternation, the loser effect collapsed.\n")
  cat("This demonstrates that power alternation can normalize democratic attitudes.\n")
} else {
  cat("South Korea data not available in the current dataset.\n")
}
```

## Other Trajectory Countries

```{r other-countries}
other_countries <- setdiff(countries_3plus, c("Thailand", "South Korea"))

for (ctry in other_countries) {
  ctry_data <- by_country_wave %>%
    filter(country_name == ctry) %>%
    arrange(wave_year)
  
  if (nrow(ctry_data) > 0) {
    cat("\n###", ctry, "\n")
    print(
      ctry_data %>%
        select(wave, wave_year, n, pct_winner, loser_effect) %>%
        kable(col.names = c("Wave", "Year", "N", "% Winners", "Loser Effect"),
              caption = paste(ctry, ": Loser Effect Over Time"))
    )
  }
}
```

# Part 4: Pooled Regression Analysis

## Main Model: Logistic Regression

```{r regression}
# Logistic regression: procedural ~ loser + country + wave FE
model <- glm(procedural_single ~ loser + country_name + wave,
             data = df, family = binomial)

# Extract loser coefficient
loser_results <- tidy(model) %>%
  filter(term == "loser") %>%
  mutate(
    odds_ratio = exp(estimate),
    ci_low = exp(estimate - 1.96 * std.error),
    ci_high = exp(estimate + 1.96 * std.error)
  )

cat("=== POOLED REGRESSION RESULTS ===\n\n")
cat("Model: procedural_single ~ loser + country_name + wave\n")
cat("Family: binomial (logistic)\n\n")
cat("Loser coefficient (log-odds):", round(loser_results$estimate, 3), "\n")
cat("Odds ratio:", round(loser_results$odds_ratio, 3), "\n")
cat("95% CI: [", round(loser_results$ci_low, 3), ",", round(loser_results$ci_high, 3), "]\n")
cat("p-value:", format(loser_results$p.value, digits = 3), "\n")
```

**Interpretation**: Controlling for country and wave fixed effects, electoral losers are **`r round((loser_results$odds_ratio - 1) * 100, 1)`% more likely** to prioritize procedural democracy than winners (OR = `r round(loser_results$odds_ratio, 2)`).

## Model Summary Table

```{r model-table}
tidy(model) %>%
  filter(!str_detect(term, "country_name")) %>%  # Hide country FE for readability
  mutate(
    odds_ratio = exp(estimate),
    term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "loser" ~ "Loser (vs. Winner)",
      str_detect(term, "wave") ~ str_replace(term, "wave", "Wave "),
      TRUE ~ term
    )
  ) %>%
  select(term, estimate, std.error, odds_ratio, p.value) %>%
  kable(digits = c(0, 3, 3, 3, 4),
        col.names = c("Term", "Log-Odds", "SE", "Odds Ratio", "p-value"),
        caption = "Logistic Regression: Procedural Democracy ~ Loser Status") %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  footnote(general = "Country fixed effects included but not shown. Reference: Winner, Wave W2.")
```

## Alternative Specification: Linear Probability Model

```{r lpm}
# LPM for coefficient interpretation
lpm <- lm(procedural_single ~ loser + country_name + wave, data = df)

lpm_results <- tidy(lpm) %>%
  filter(term == "loser")

cat("\n=== LINEAR PROBABILITY MODEL ===\n")
cat("Loser coefficient:", round(lpm_results$estimate * 100, 1), "percentage points\n")
cat("SE:", round(lpm_results$std.error * 100, 2), "\n")
cat("p-value:", format(lpm_results$p.value, digits = 3), "\n")
```

# Part 5: Summary of Key Findings

```{r summary}
cat("=== SUMMARY OF KEY FINDINGS ===\n\n")

cat("1. OVERALL LOSER EFFECT\n")
cat("   - Losers are more likely to prioritize procedural democracy\n")
cat("   - Odds ratio ~", round(loser_results$odds_ratio, 2), "(p < 0.001)\n")
cat("   - Effect size: ~", round(lpm_results$estimate * 100, 1), "percentage points\n\n")

cat("2. TEMPORAL PATTERN\n")
for (i in 1:nrow(wave_tests)) {
  w <- wave_tests[i, ]
  sig <- if (w$p_value < 0.001) "***" else if (w$p_value < 0.01) "**" else if (w$p_value < 0.05) "*" else ""
  cat("   - ", w$wave, ": ", sprintf("%+.1f", w$loser_effect), " pp ", sig, "\n", sep = "")
}

cat("\n3. COUNTRY TRAJECTORIES\n")
if (nrow(trajectory) > 0) {
  cat("   Largest increases:\n")
  top_increases <- trajectory %>% arrange(desc(total_change)) %>% head(3)
  for (i in 1:nrow(top_increases)) {
    cat("   - ", top_increases$country_name[i], ": ", 
        sprintf("%+.1f", top_increases$total_change[i]), " pp\n", sep = "")
  }
  
  cat("   Largest decreases:\n")
  top_decreases <- trajectory %>% arrange(total_change) %>% head(3)
  for (i in 1:nrow(top_decreases)) {
    cat("   - ", top_decreases$country_name[i], ": ", 
        sprintf("%+.1f", top_decreases$total_change[i]), " pp\n", sep = "")
  }
}

cat("\n4. KEY IMPLICATION\n")
cat("   Conceptions of democracy are not fixed cultural orientations—\n")
cat("   they respond dynamically to citizens' political position and\n")
cat("   the health of democratic institutions.\n")
```

# Save Results

```{r save}
output_dir <- here("papers", "meaning-of-democracy", "analysis", "data")
results_dir <- here("papers", "meaning-of-democracy", "analysis", "results")

# Save country-wave summary
write_csv(by_country_wave, file.path(output_dir, "loser_effect_by_country_wave.csv"))

# Save trajectory data
write_csv(trajectory, file.path(output_dir, "loser_effect_trajectory.csv"))

# Save wave summary
write_csv(wave_tests, file.path(output_dir, "loser_effect_by_wave.csv"))

# ============================================================
# COMPREHENSIVE RESULTS FOR MANUSCRIPT
# ============================================================

# Sample statistics by wave
sample_by_wave <- df %>%
  group_by(wave) %>%
  summarise(
    n = n(),
    n_countries = n_distinct(country_name),
    .groups = "drop"
  )

total_n <- nrow(df)
total_countries <- n_distinct(df$country_name)

# Wave tests with formatted values
wave_results <- wave_tests %>%
  mutate(
    loser_effect_fmt = sprintf("%+.1f", loser_effect),
    odds_ratio = exp(log(loser_pct / (100 - loser_pct)) - log(winner_pct / (100 - winner_pct))),
    odds_ratio_fmt = sprintf("%.2f", odds_ratio),
    p_fmt = case_when(
      p_value < 0.001 ~ "< 0.001",
      p_value < 0.01 ~ sprintf("%.3f", p_value),
      TRUE ~ sprintf("%.2f", p_value)
    ),
    sig = p_value < 0.05
  )

# Pooled regression statistics
pooled_stats <- list(
  odds_ratio = loser_results$odds_ratio,
  odds_ratio_fmt = sprintf("%.2f", loser_results$odds_ratio),
  pct_increase = round((loser_results$odds_ratio - 1) * 100, 0),
  ci_low = loser_results$ci_low,
  ci_high = loser_results$ci_high,
  ci_fmt = sprintf("[%.2f, %.2f]", loser_results$ci_low, loser_results$ci_high),
  lpm_effect = lpm_results$estimate * 100,
  lpm_effect_fmt = sprintf("%.1f", lpm_results$estimate * 100)
)

# Country-wave detailed results
country_wave_detail <- by_country_wave %>%
  mutate(
    pct_winner_fmt = sprintf("%.0f%%", pct_winner),
    loser_effect_fmt = sprintf("%+.1f", loser_effect)
  )

# Get specific country data for case studies
thailand_data <- country_wave_detail %>%
  filter(country_name == "Thailand") %>%
  arrange(wave_year)

korea_data <- country_wave_detail %>%
  filter(country_name == "South Korea") %>%
  arrange(wave_year)

taiwan_data <- country_wave_detail %>%
  filter(country_name == "Taiwan") %>%
  arrange(wave_year)

japan_data <- country_wave_detail %>%
  filter(country_name == "Japan") %>%
  arrange(wave_year)

philippines_data <- country_wave_detail %>%
  filter(country_name == "Philippines") %>%
  arrange(wave_year)

# Key statistics for abstract
# Thailand: change from W2 to W6
thai_w2 <- thailand_data %>% filter(wave == "W2") %>% pull(loser_effect)
thai_w6 <- thailand_data %>% filter(wave == "W6") %>% pull(loser_effect)
thai_change <- thai_w6 - thai_w2

# Korea: peak effect
korea_peak <- korea_data %>% summarise(max_effect = max(loser_effect)) %>% pull(max_effect)

# Effect range in pre-pandemic waves (W2-W4)
effect_range <- wave_results %>%
  filter(wave %in% c("W2", "W3", "W4")) %>%
  summarise(min_effect = min(loser_effect), max_effect = max(loser_effect))

# Compile all results into a list
manuscript_results <- list(
  # Sample
  total_n = total_n,
  total_n_fmt = format(total_n, big.mark = ","),
  total_countries = total_countries,
  sample_by_wave = sample_by_wave,

  # Wave-level results
  wave_results = wave_results,

  # Pooled regression
  pooled = pooled_stats,

  # Country-wave detail
  country_wave = country_wave_detail,

  # Case study data
  thailand = thailand_data,
  korea = korea_data,
  taiwan = taiwan_data,
  japan = japan_data,
  philippines = philippines_data,

  # Trajectory data
  trajectory = trajectory,

  # Key stats for abstract
  abstract = list(
    thailand_w6_effect = thai_w6,
    thailand_change = thai_change,
    korea_peak = korea_peak,
    effect_range_min = effect_range$min_effect,
    effect_range_max = effect_range$max_effect
  ),

  # Metadata
  generated = Sys.time(),
  waves_included = unique(df$wave)
)

# Save comprehensive results
saveRDS(manuscript_results, file.path(results_dir, "manuscript_results.rds"))

cat("\n=== RESULTS SAVED ===\n")
cat("- loser_effect_by_country_wave.csv\n")
cat("- loser_effect_trajectory.csv\n")
cat("- loser_effect_by_wave.csv\n")
cat("- results/manuscript_results.rds (comprehensive results for manuscript)\n")
```
