---
title: "The Accountability Gap: Mass Perceptions of Democratic Backsliding in South Korea"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: now
date-format: "MMMM D, YYYY [at] h:mm A"
abstract: |
  Democratic backsliding produces institutional erosion that elite-level indicators capture well but mass public opinion rarely illuminates. What such erosion looks like from below—how ordinary citizens perceive, process, and respond to accountability crises—remains poorly understood. Drawing on six waves of the Asian Barometer Survey (2001–2022) and two waves of the Korean Academic Multimode Open Survey (2016–2019), this study documents mass perceptual change in South Korea across a period of acute accountability crisis. The analysis identifies an accountability gap syndrome—a coherent configuration in which perceptions of governmental information withholding rise secularly, experiential corruption perceptions shift discontinuously following the 2016–2018 accountability shock, and trust in intermediary accountability institutions collapses while executive trust remains stable. This pattern is cross-partisan, replicates across independent survey instruments, and is inconsistent with generalized political cynicism. A perceptual reclassification argument accounts for the divergence between rising experiential and stable institutional corruption assessments, with implications for how anti-corruption reform is measured and interpreted comparatively.
keywords:
  - South Korea
  - accountability
  - democratic backsliding
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{caption}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \usepackage{colortbl}
          \usepackage{xcolor}
          \captionsetup{width=\textwidth}
          \setlength\LTcapwidth{\textwidth}
          \floatplacement{figure}{H}
          \floatplacement{table}{H}
          \raggedright
          \newfontfamily\koreanfont{Noto Sans KR}[Script=Hangul]
          \DeclareTextFontCommand{\textkorean}{\koreanfont}
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/econdev-authpref"
paper_dir    <- file.path(project_root, "papers/south-korea-accountability-gap")
analysis_dir <- file.path(paper_dir, "analysis")
fig_dir      <- file.path(analysis_dir, "figures")
table_dir    <- file.path(analysis_dir, "tables")
results_dir  <- file.path(analysis_dir, "results")

# ABS (Asian Barometer, Korea subset) and KAMOS (Korean domestic panel)
# kor_abs   <- readRDS(file.path(analysis_dir, "kor_abs.rds"))
# kor_kamos <- readRDS(file.path(analysis_dir, "kor_kamos.rds"))
```

```{r load-results}
load(file.path(results_dir, "table1_pub.RData"))             # wave_stats, diff_stats
load(file.path(results_dir, "table2_pub.RData"))             # kamos_trust_stats, kamos_ttests, tab2_wide
load(file.path(results_dir, "kamos_trust_means.RData"))      # kamos_change, pol_sat
load(file.path(results_dir, "table3_pub.RData"))             # ideo_cors, tab3
load(file.path(results_dir, "data_description_stats.RData")) # n_abs_total, abs_wave_n_min/max, n_kamos_total, n_kamos_vars
slopes_inline <- readRDS(file.path(results_dir, "slopes_inline.rds"))
cross_slopes  <- readRDS(file.path(results_dir, "cross_national_slopes.rds"))

wave_years <- c("1"=2003,"2"=2006,"3"=2011,"4"=2015,"5"=2019,"6"=2022)
n_2016 <- kamos_trust_stats |> dplyr::filter(variable=="trust_central_govt", wave_year==2016) |> dplyr::pull(n)
n_2019 <- kamos_trust_stats |> dplyr::filter(variable=="trust_central_govt", wave_year==2019) |> dplyr::pull(n)

abs_mn <- function(var, wv) round(wave_stats$mean[wave_stats$variable == var & wave_stats$wave == wv], 3)
transp_w2      <- abs_mn("govt_withholds_info", 2)
transp_w6      <- abs_mn("govt_withholds_info", 6)
cw_w1          <- abs_mn("corrupt_witnessed", 1)
cw_w4          <- abs_mn("corrupt_witnessed", 4)
cw_w5          <- abs_mn("corrupt_witnessed", 5)
cw_w6          <- abs_mn("corrupt_witnessed", 6)
cw_w4_to_w5    <- round(cw_w5 - cw_w4, 3)
cng_range_diff <- round(diff(range(wave_stats$mean[wave_stats$variable == "corrupt_national_govt"])), 2)

kc <- function(var) kamos_change[kamos_change$variable == var, ]
cgov_2016    <- round(kc("trust_central_govt")$y2016, 3)
cgov_2019    <- round(kc("trust_central_govt")$y2019, 3)
cgov_chg     <- abs(round(kc("trust_central_govt")$change, 3))
lgov_chg     <- abs(round(kc("trust_local_govt")$change, 3))
nassem_chg   <- abs(round(kc("trust_national_assembly")$change, 3))
leg_chg      <- abs(round(kc("trust_legislature")$change, 3))
media_chg    <- abs(round(kc("trust_media")$change, 3))
ngo_chg      <- abs(round(kc("trust_ngo")$change, 3))
intermed_min <- min(nassem_chg, leg_chg, media_chg, ngo_chg)
intermed_max <- max(nassem_chg, leg_chg, media_chg, ngo_chg)
polsat_chg   <- abs(round(pol_sat$mean_sat[pol_sat$wave == 4] - pol_sat$mean_sat[pol_sat$wave == 1], 3))

ideo_r_vals  <- unlist(ideo_cors |> dplyr::select(dplyr::starts_with("r_")))
ideo_cor_min <- round(min(ideo_r_vals), 3)
ideo_cor_max <- round(max(ideo_r_vals), 3)
```

# Introduction {#sec-introduction}

In December 2016, millions of South Korean citizens took to the streets in a series of candlelight vigils that ultimately precipitated the impeachment and removal of President Park Geun-hye. The constitutional mechanisms of democratic accountability functioned as designed: the National Assembly voted to impeach, the Constitutional Court upheld the decision, and Park was subsequently prosecuted and convicted. By most institutional measures, South Korean democracy had passed a severe test. Yet the survey evidence tells a more complicated story. In the years surrounding and following this episode, South Korean citizens became progressively more likely to report that their government withholds information from the public, more likely to report personal encounters with corruption in their everyday social environment, and substantially less trusting of the media, civil society organizations, and legislative institutions whose function is to hold power accountable. The institutional success of the impeachment did not translate into perceptual restoration at the mass public level. The accountability mechanisms worked; the accountability gap widened.

This paper investigates that paradox. Drawing on six waves of the Asian Barometer Survey spanning 2001 to 2022 and two waves of the Korean Academic Multimode Open Survey covering 2016 and 2019, it documents what democratic backsliding looks like from the perspective of mass publics—not the institutional erosion that elite-level indicators capture, but the perceptual reorganization that occurs among ordinary citizens as accountability crises unfold. The central argument is that backsliding episodes produce a distinctive mass public signature, here termed the accountability gap syndrome, characterized by the simultaneous accumulation of transparency concerns, the reclassification of everyday social practices as corrupt, and the selective collapse of trust in intermediary accountability institutions. This configuration is not generalized political disillusionment. It is a specific reconfiguration of the citizen-institution relationship that may sustain the conditions it reflects, by eroding the mass-level trust on which accountability mechanisms depend.

Two contributions follow from this analysis. The first is empirical: systematic survey evidence on the mass public correlates of South Korea's backsliding episode, a dimension of the crisis that has received considerably less attention than its institutional and electoral dimensions. The second is theoretical: the paper develops the reclassification mechanism as a distinct pathway through which accountability shocks shape mass corruption perception, distinguishing it from the secular legitimacy erosion that drives transparency concerns and demonstrating that the two mechanisms generate empirically separable observable implications. The divergence between rising experiential corruption perceptions and stable institutional corruption assessments—replicated across two independent survey instruments—constitutes the core empirical test of this distinction.

The remainder of the paper proceeds as follows. The next section develops the theoretical framework, distinguishing the legitimacy erosion and perceptual reclassification mechanisms and deriving the hypotheses that structure the empirical analysis. The following section reviews the relevant literatures on democratic backsliding, political trust, and corruption perception, situating the paper's contribution within each. The data and methods section describes the survey instruments, variable operationalization, and analytic strategy. The results section presents the ABS longitudinal trends and the KAMOS cross-wave comparisons in turn, followed by the partisan sorting analysis. The discussion section interprets the findings in light of the theoretical framework and addresses the study's limitations. The conclusion considers the broader implications of the accountability gap syndrome for the comparative study of democratic erosion and mass political belief.

# Theoretical Framework {#sec-theory}

Democratic backsliding poses a distinctive challenge for the study of mass political behavior. While the institutional dimensions of backsliding—the erosion of electoral integrity, the weakening of horizontal accountability, the subordination of courts and legislatures to executive prerogative—have attracted substantial scholarly attention [@Levitsky2018-dd; @Bermeo2016-zl; @Luhrmann2019-ow], comparatively less is known about how these processes register in mass public opinion. The question is not simply whether citizens notice democratic erosion, but how backsliding episodes restructure the terms through which citizens interpret their political environment. This paper addresses that gap through an analysis of public opinion in South Korea across a period of acute accountability crisis.

Two mechanisms through which backsliding episodes shape mass political perception structure the analysis. The first is legitimacy erosion: the gradual accumulation of evidence that governing institutions withhold information, manipulate accountability processes, and operate beyond public scrutiny. The second is perceptual reclassification: the process by which high-profile accountability shocks reorganize the categories through which citizens interpret everyday social and political life. These mechanisms operate on different timescales and produce distinct observable implications, yet both contribute to what this paper terms an accountability gap syndrome—a coherent pattern in which citizens perceive simultaneously that governing institutions are less transparent and that corruption is more pervasive in their immediate social environment.

Legitimacy erosion is a secular process. It does not require a single triggering event but accumulates across repeated encounters between citizens and institutions that fail to meet normative expectations of openness and responsiveness. Where governing institutions are perceived to systematically withhold information, trust in the accountability infrastructure erodes over time. This process is unlikely to reverse rapidly, since the perceptual updating it produces is driven by diffuse cumulative experience rather than discrete observable events. The Korean case offers a particularly clear window onto this dynamic: survey evidence from the Asian Barometer Survey indicates that perceptions of governmental information withholding rose monotonically across six waves spanning two decades, accelerating modestly under each successive crisis without reversing between them. This secular trajectory is the empirical signature of legitimacy erosion rather than event-driven cynicism.

Perceptual reclassification, by contrast, is discontinuous. It occurs when a high-salience public event—an impeachment, a high-profile prosecution, the enactment of landmark anti-corruption legislation—supplies citizens with new cognitive frameworks for categorizing practices they previously understood differently. The Park Geun-hye impeachment (2016–2017) and the near-simultaneous enactment of the Improper Solicitation and Graft Act, law 14813 [@Unknown2016-uy], hereafter referred to as the Kim Young-ran Act (September 2016) together constituted precisely such a moment in South Korea. The impeachment made elite corruption spectacularly visible, while the Kim Young-ran Act explicitly redrew the legal and normative boundary between acceptable relational practice and corruption, criminalizing gift-giving, hospitality, and monetary transfers that had previously occupied an ambiguous social status. Critically, the Act did not merely prohibit specific acts; it reorganized the conceptual landscape within which citizens evaluate everyday transactions. Practices previously understood as expressions of `\textkorean{온정주의}`{=latex}—warm relational obligation—were reframed as potential corruption, and an enforcement apparatus was created to institutionalize that reframing.

This reclassification mechanism generates a prediction that distinguishes it from simpler accounts of growing cynicism. If citizens were simply becoming more distrustful of institutions generally, one would expect perceptions of corruption to rise uniformly across domains: in national government, local government, and everyday social experience alike. Perceptual reclassification, by contrast, predicts an asymmetric pattern: experiential corruption perceptions, measuring what citizens encounter in their immediate social environment, should rise sharply following the accountability shock, while institutional corruption assessments, which tap more abstract evaluations of formal political actors, should remain comparatively stable. The divergence between these two perception types is the empirical signature of reclassification rather than generic cynicism, and it is precisely this divergence that the Korean data reveal.

Taken together, these two mechanisms yield the accountability gap syndrome: a configuration in which citizens perceive governing institutions as increasingly opaque while perceiving corruption as increasingly present in their immediate social world, yet without a corresponding rise in abstract institutional corruption assessments. The syndrome is not merely additive; rather, the combination of rising transparency concerns and rising experiential corruption perceptions constitutes a theoretically coherent mass-level response to a backsliding episode that both revealed elite misconduct and reorganized popular categories of civic evaluation.

From this framework, two hypotheses follow. First, perceptions that government withholds information should increase monotonically across the period of study, consistent with the secular legitimacy erosion mechanism (H1). Second, perceptions of corruption in everyday social experience should increase discontinuously, with the sharpest rise concentrated in the interval spanning the 2016–2018 accountability shock, while institutional corruption assessments remain comparatively stable (H2). The divergence predicted by H2—rising experiential perceptions alongside stable institutional assessments—constitutes the central empirical test of the reclassification argument.

# Literature Review

The study of democratic backsliding has undergone significant theoretical refinement over the past decade. Early scholarship tended to conceptualize democratic breakdown in terms of abrupt, visible ruptures—coups, constitutional suspensions, mass repression—that clearly demarcated the transition from democratic to authoritarian rule [@Linz1978-cr; @ODonnell1986-bf]. More recent work has drawn attention to a qualitatively different form of democratic erosion, one that proceeds incrementally through formally legal means and is therefore considerably harder to diagnose in real time [@Levitsky2018-dd; @Bermeo2016-zl]. Backsliding of this variety typically involves the systematic weakening of horizontal accountability institutions—courts, legislatures, audit bodies, independent media—rather than their outright abolition, and it frequently occurs under elected governments that retain genuine popular support even as they erode the structural conditions for democratic competition [@Luhrmann2019-ow; @Mechkova2017-na].

This institutional focus has generated important advances in the measurement and comparative analysis of backsliding, but it has also produced a relative neglect of the mass public dimension. The dominant empirical approaches—expert-coded indices such as V-Dem, Freedom House, and the Economist Intelligence Unit's Democracy Index—capture elite-level institutional change with considerable precision but reveal relatively little about how democratic erosion registers among ordinary citizens. This is a significant gap, for at least two reasons. First, mass public opinion shapes the political feasibility of backsliding: governments that erode accountability institutions face less resistance when citizens are disengaged, distrustful, or ideologically polarized along lines that cut across democratic commitments [@Foa2016-lq; @Norris2019-cr]. Second, and more fundamentally, the legitimacy of democratic institutions ultimately rests on mass belief—on the widespread perception that those institutions are capable of providing accountability, representation, and responsive governance. Understanding what backsliding looks like from the perspective of mass publics is therefore not merely an empirical complement to institutional analysis but a theoretically necessary component of any complete account of democratic erosion.

The literature on political trust and institutional legitimacy provides the most direct theoretical resources for this inquiry. A substantial body of scholarship has documented declining trust in political institutions across advanced democracies over recent decades, a trend variously attributed to rising educational attainment and critical citizenship [@Norris1999-sn], postmaterialist value change [@Inglehart1997-gw], growing economic inequality [@Uslaner2018-en], and the perceived failure of governing institutions to deliver on democratic promises [@Dalton2004-bq]. Within this literature, a useful distinction has emerged between specific and diffuse support for political systems [@Easton1965-ap; @Norris1999-sn]: specific support refers to evaluations of current incumbents and their performance, while diffuse support reflects deeper orientations toward regime institutions and democratic norms. Backsliding is theoretically most consequential when it erodes diffuse support—when citizens come to doubt not merely the competence of current officeholders but the capacity of institutions themselves to provide democratic accountability.

The Korean case has figured in this literature primarily through debates about the consolidation and quality of Korea's democratic transition rather than backsliding as such. Scholars have noted persistent gaps between formal democratic institutions and their effective functioning, particularly with respect to elite corruption, prosecutorial independence, and chaebol-state relations [@Mo1998-ly; @Croissant2002-korea; @Im2011-pj]. The Park Geun-hye scandal and subsequent impeachment have been analyzed as a test of institutional resilience [@Doucette2017-ta; @Han2021-nb; @Hahm2018-er], with most accounts emphasizing the ultimately successful operation of constitutional accountability mechanisms. Less attention has been paid to the mass public consequences of the crisis—to what the impeachment episode did to popular perceptions of transparency, corruption, and institutional integrity beyond the immediate electoral cycle.

The corruption perception literature offers a second set of theoretical resources, though one that requires careful handling. Survey-based measures of corruption perception are widely used as proxies for actual corruption levels, but the relationship between perception and reality is complex and context-dependent [@Treisman2007-vk; @Razafindrakoto2010-ur]. Perceptions of corruption are shaped not only by direct personal experience but by media coverage, elite cues, social norms, and the legal frameworks that define what counts as corrupt in the first place [@Manzetti2007-yz; @Tverdova2011-sz]. The definitional dimension is particularly consequential in contexts where anti-corruption legislation reorganizes the boundary between acceptable and unacceptable practice, as the Kim Young-ran Act did in Korea in 2016. When law explicitly recategorizes previously tolerated behavior as corruption, survey items asking respondents about their experience of corruption may capture genuine perceptual change—a reclassification of familiar practices—rather than changes in the underlying frequency of corrupt acts. Failure to account for this possibility risks misinterpreting attitudinal stability as behavioral stability, and attitudinal change as behavioral change.

The present study contributes to both literatures. Against the backsliding scholarship, it argues that mass public opinion exhibits a coherent syndrome of accountability perception—rising transparency concerns combined with rising experiential corruption perceptions and collapsing intermediary institution trust—that constitutes the citizen-level correlate of elite-level democratic erosion. Against the corruption perception literature, it argues that high-salience accountability shocks, particularly when accompanied by legal redefinition of corruption itself, can restructure mass perceptual categories in ways that standard survey instruments are liable to misread. The Korean case illustrates both dynamics with unusual clarity, offering a natural quasi-experiment in which the timing of institutional intervention—the Kim Young-ran Act, the impeachment, the subsequent prosecution—is known with precision relative to the survey waves used in the analysis.

# Data and Research Design {#sec-design}

This study draws on two complementary survey datasets to examine mass public perceptions of corruption and institutional trust in South Korea across a period of acute accountability crisis. The primary cross-national data come from the Asian Barometer Survey (ABS), a multi-wave comparative public opinion project covering democracies and semi-democracies across East and Southeast Asia. The Korean ABS sample spans six waves collected between 2001 and 2022, yielding `r formatC(n_abs_total, format="d", big.mark=",")` respondents drawn through consistent probability-based sampling, with wave sizes ranging from approximately `r formatC(abs_wave_n_min, format="d", big.mark=",")` to `r formatC(abs_wave_n_max, format="d", big.mark=",")`. The ABS data provide the longitudinal backbone of the analysis, enabling trend estimation across two decades and precise identification of the intervals in which perceptual change was most concentrated.

The secondary data come from the Korean Academic Multimode Open Survey (KAMOS), a domestic probability-based panel administered by Chungnam National University. Two waves of the KAMOS are used here: Wave 1 [@Center-for-Asian-Public-Opinion-Research-Collaboration-Initiative-CAPORCI-Chungnam-National-University2016-uk], fielded in February through May 2016, and Wave 4, fielded in May through June 2019 [@Center-for-Asian-Public-Opinion-Research-Collaboration-Initiative-CAPORCI-Chungnam-National-University2019-it] . The harmonized dataset combines `r formatC(n_kamos_total, format="d", big.mark=",")` South Korean respondents across these two waves and includes `r n_kamos_vars` variables covering institutional trust, political attitudes, economic evaluations, and demographic characteristics. The KAMOS data serve a triangulation function: they provide an independent domestic instrument measured at timepoints closely overlapping with ABS Waves 4 and 5, enabling assessment of whether the perceptual patterns identified in the ABS replicate on a separate survey with different question wording, sampling frame, and administration mode.

The theoretical argument developed above centers on three observable implications, each operationalized using variables available in one or both datasets. The first is the secular rise in perceptions of governmental information withholding, measured in the ABS by an item asking respondents whether they agree that the government withholds information from the public (govt_withholds_info, four-point scale, higher values indicating greater agreement). This variable is available across ABS Waves 2 through 6, providing five measurement points spanning 2003 to 2022. The second is the discontinuous rise in experiential corruption perceptions, measured by an ABS item asking whether the respondent has personally witnessed or experienced corrupt acts (corrupt_witnessed, four-point scale, higher values indicating greater frequency). This item is available across all six ABS waves. The third is the stability of institutional corruption assessments, measured by parallel ABS items asking respondents to evaluate corruption levels in the national and local governments (corrupt_national_govt, corrupt_local_govt). The divergence between the experiential and institutional series—rising perceptions of personal encounter with corruption alongside stable assessments of governmental corruption—constitutes the primary empirical test of the reclassification hypothesis.

Institutional trust is measured in the KAMOS using a ten-point battery (0 = completely untrustworthy, 10 = completely trustworthy) covering trust in the central government, local government, National Assembly, legislature, private enterprise, media and press, civil society organizations, and religious organizations. For the purposes of this analysis, five items are of primary theoretical interest: trust in the central government (cgotrust), the National Assembly (natrust), the legislature (legtrust), and the media and press (medtrust), as well as trust in civil society organizations (orgtrust). The first two capture trust in the core executive and legislative accountability institutions, while the latter three capture trust in the intermediary infrastructure through which citizens access and evaluate political information. The theoretical expectation is that the accountability shock of 2016–2018 should produce larger declines in intermediary institution trust than in central government trust, consistent with the accountability gap argument that citizens perceive a specific collapse of the mechanisms that should hold power accountable rather than a generalized withdrawal of political support.

The analytic strategy proceeds in three stages. The first stage examines the ABS longitudinal series for the three key perception variables, plotting wave means and assessing the timing and magnitude of change relative to the theoretical periodization established by the Kim Young-ran Act (September 2016) and the Park impeachment and conviction (March 2017 and April 2018 respectively). ABS Wave 4 (November 2015) provides the pre-treatment baseline; Wave 5 (June 2019) provides the post-treatment observation, with the intervening interval capturing the full accountability shock. The second stage examines the KAMOS institutional trust battery, comparing wave means across the 2016 and 2019 administrations and assessing the differential magnitude of trust decline across institutional categories. The third stage examines whether the trust collapse in the KAMOS is ideologically sorted—that is, whether it reflects partisan realignment following the change of government rather than a structural shift in accountability perceptions—by comparing trust trajectories across partisan and ideological subgroups.

Several methodological limitations warrant acknowledgment. The ABS data do not support individual-level panel estimation, since each wave draws an independent cross-sectional sample; trend claims therefore rest on repeated cross-sectional inference rather than within-person change. The KAMOS waves used here are similarly independent cross-sections rather than a true panel, which precludes direct estimation of individual-level attitude change. The analysis is consequently descriptive and inferential rather than causal in a strict experimental sense, and causal claims about the effects of the accountability shock on perceptual change are grounded in theoretical reasoning about timing and mechanism rather than identification from exogenous variation. Additionally, the question wording of the corrupt_witnessed item does not permit direct adjudication between behavioral and perceptual interpretations of rising scores—a limitation that the theoretical framework addresses by treating the ambiguity as substantively informative rather than merely as measurement error.

# Results {#sec-results}

## Longitudinal Trends in the Asian Barometer Survey

The ABS data reveal two distinct temporal patterns that together constitute the accountability gap syndrome described in the theoretical framework. The first is a monotonic secular rise in perceptions of governmental information withholding. Across five waves spanning 2003 to 2022, mean agreement that the government withholds information from the public increased without interruption, from `r transp_w2` in Wave 2 to `r transp_w6` in Wave 6 (Table 1). The rise is gradual rather than punctuated: no single inter-wave interval accounts for a disproportionate share of the total change, and the series does not flatten or reverse following either the Park impeachment or the subsequent transition to the Moon administration. This monotonic trajectory is consistent with the legitimacy erosion mechanism proposed in H1, and inconsistent with an event-driven account in which transparency perceptions respond primarily to the identity of the incumbent government. Had the series been driven by co-partisan cuing, one would expect a reversal following Moon's election in May 2017; no such reversal is evident.

The second pattern is more theoretically consequential. Perceptions of personally witnessed or experienced corruption follow a different trajectory from institutional corruption assessments, and it is the divergence between these two series that constitutes the primary empirical test of H2. Mean scores on the corrupt_witnessed item rose from `r cw_w1` in Wave 1 to `r cw_w6` in Wave 6, with the sharpest single-wave increase occurring between Wave 4 (November 2015, mean = `r cw_w4`) and Wave 5 (June 2019, mean = `r cw_w5`)—an increase of `r cw_w4_to_w5` points spanning the full accountability shock period and encompassing the Kim Young-ran Act, the Park impeachment, and the subsequent prosecution and conviction. Notably, the series plateaus between Waves 5 and 6, suggesting that the perceptual reorganization induced by the accountability shock had largely consolidated by the time of the 2019 fieldwork and did not continue to accumulate into the early Yoon period.

Against this backdrop, the institutional corruption assessment series is strikingly stable. Mean scores on both the corrupt_national_govt and corrupt_local_govt items fluctuate within a narrow range of approximately `r cng_range_diff` points across all six waves, with no directional trend and no pronounced response to the 2016–2018 accountability shock. This stability is not consistent with a generic cynicism account, under which rising perceptions of corruption in everyday life would be accompanied by rising assessments of governmental corruption. The divergence between the experiential and institutional series—a rise of `r cw_w4_to_w5` points in corrupt_witnessed across the critical interval against effectively flat institutional assessments—provides direct empirical support for the reclassification hypothesis. Citizens became more likely to report personal encounters with corruption precisely in the interval when the Kim Young-ran Act redrew the legal boundary of corrupt practice, while their abstract assessments of governmental corruption remained anchored. The pattern is consistent with perceptual reclassification of ambient social practices rather than a behavioral increase in corrupt acts or a generalized withdrawal of political trust.

```{r table1-prep, include=FALSE}
var_labels <- c(
  govt_withholds_info   = "Government withholds information",
  corrupt_witnessed     = "Personally witnessed/experienced corruption",
  corrupt_national_govt = "Corruption in national government",
  corrupt_local_govt    = "Corruption in local government"
)
scale_labels <- c(
  govt_withholds_info   = "1\u20134",
  corrupt_witnessed     = "1\u20132",
  corrupt_national_govt = "1\u20134",
  corrupt_local_govt    = "1\u20134"
)
abs_vars        <- names(var_labels)
all_waves       <- 1:6
all_wave_labels <- paste0("W", all_waves, "\n(~", wave_years[as.character(all_waves)], ")")
all_intervals   <- c("W1\u2192W2","W2\u2192W3","W3\u2192W4","W4\u2192W5","W5\u2192W6")

build_upper_row <- function(var) {
  row_cells <- purrr::map_chr(all_waves, function(wv) {
    d <- wave_stats |> dplyr::filter(variable == var, wave == wv)
    if (nrow(d) == 0) return("---")
    sprintf("%.3f\n(%.3f)\n[%s]", d$mean, d$se, formatC(d$n, format="d", big.mark=","))
  })
  c(var_labels[var], scale_labels[var], row_cells)
}
build_lower_row <- function(var) {
  row_cells <- purrr::map_chr(all_intervals, function(intv) {
    d <- diff_stats |> dplyr::filter(variable == var, interval == intv)
    if (nrow(d) == 0) return("---")
    sprintf("%+.3f\n(%.3f)", d$diff, d$se_diff)
  })
  c(var_labels[var], "", row_cells)
}

upper_df <- as.data.frame(do.call(rbind, purrr::map(abs_vars, build_upper_row)), stringsAsFactors=FALSE)
colnames(upper_df) <- c("Variable", "Scale", all_wave_labels)
lower_df <- as.data.frame(do.call(rbind, purrr::map(abs_vars, build_lower_row)), stringsAsFactors=FALSE)
colnames(lower_df) <- c("Variable", "Scale", all_intervals)
```

```{r table1-upper}
kableExtra::kbl(
  upper_df,
  caption  = "ABS Korea Wave Means (Upper Panel). Cell entries: mean (SE) [n]. Waves 1--2 use unweighted means (survey weights unavailable).",
  booktabs = TRUE, escape = FALSE,
  align    = c("l","c", rep("c",6))
) |>
  kableExtra::kable_styling(latex_options=c("hold_position","scale_down"), font_size=9, full_width=FALSE) |>
  kableExtra::column_spec(1, width="8em") |>
  kableExtra::column_spec(3:8, width="5em") |>
  kableExtra::footnote(
    general="W4 (2015) and W5 (2019) columns span the accountability shock interval. govt\\_withholds\\_info not administered in Wave 1.",
    general_title="Note.", threeparttable=TRUE
  )
```

```{r table1-lower}
kableExtra::kbl(
  lower_df,
  caption  = "ABS Korea Inter-Wave Differences (Lower Panel). Cell entries: difference (SE of difference). W4--W5 spans the accountability shock interval.",
  booktabs = TRUE, escape = FALSE,
  align    = c("l","c", rep("c",5))
) |>
  kableExtra::kable_styling(latex_options=c("hold_position","scale_down"), font_size=9, full_width=FALSE) |>
  kableExtra::column_spec(1, width="8em") |>
  kableExtra::column_spec(3:7, width="5em") |>
  kableExtra::footnote(
    general="--- indicates variable not available in that wave. W4--W5 column spans the accountability shock interval.",
    general_title="Note.", threeparttable=TRUE
  )
```

```{r fig1, fig.cap="ABS Longitudinal Trends: Transparency Erosion and Corruption Perceptions, South Korea 2003\\textendash 2022. Shaded corridor = accountability shock interval (Kim Young-ran Act, September 2016 to Park conviction, April 2018). Weighted means shown; Waves 1\\textendash 2 unweighted.", out.width="\\textwidth"}
knitr::include_graphics(file.path(fig_dir, "fig1_pub_abs_trends.pdf"))
```

## Institutional Trust in the KAMOS Data

The KAMOS data provide independent corroboration of these patterns from a domestic survey instrument administered at closely overlapping timepoints. Comparing Wave 1 (2016) and Wave 4 (2019) means across the institutional trust battery reveals a broad decline in trust, but one that is strikingly differentiated by institutional category in ways that align closely with the theoretical expectations derived from the accountability gap framework (Table 2).

Trust in the central government declined only marginally between 2016 and 2019, falling from `r cgov_2016` to `r cgov_2019`, a change of `r cgov_chg` points on the ten-point scale that is substantively negligible. 
Trust in local government registered a similarly modest decline of `r lgov_chg` points. By contrast, trust in the intermediary accountability infrastructure collapsed across the same interval. Trust in the National Assembly fell by `r nassem_chg` points, trust in the legislature by `r leg_chg` points, trust in media and press by `r media_chg` points, and trust in civil society organizations by `r ngo_chg` points, the largest single decline in the battery. Declines of this magnitude—concentrated in the institutions specifically responsible for holding executive power accountable—represent a substantial reorganization of the citizen-institution relationship.
This pattern of selective trust decline is inconsistent with a generalized withdrawal of political support. If citizens were simply becoming more cynical about politics broadly, one would expect roughly uniform trust declines across institutional categories. The observed pattern—near-stability in executive branch trust alongside sharp declines in legislative, media, and civil society trust—suggests instead that citizens distinguished between the institutions of government and the institutions of accountability, losing confidence specifically in the latter. This is precisely the configuration that the accountability gap framework predicts: a backsliding episode that weakens the mechanisms of horizontal accountability should register in mass opinion as a collapse of trust in accountability institutions rather than in government per se.

```{r table2}
kableExtra::kbl(
  tab2_wide |> dplyr::select(Institution = label, `2016 Mean (SD)`, `2019 Mean (SD)`, Change, t),
  caption  = sprintf(
    "KAMOS Institutional Trust, Wave 1 (2016, n = %s) and Wave 4 (2019, n = %s). Trust on 0\\textendash 10 scale. Change = 2019 minus 2016 mean. t = Welch two-sample t-statistic.",
    formatC(n_2016, format="d", big.mark=","),
    formatC(n_2019, format="d", big.mark=",")),
  booktabs = TRUE, escape = FALSE,
  align    = c("l","c","c","c","r")
) |>
  kableExtra::kable_styling(latex_options=c("hold_position"), font_size=11, full_width=FALSE) |>
  kableExtra::column_spec(1, width="12em") |>
  kableExtra::pack_rows("Block A: Executive Institutions",       1, 2, bold=TRUE, latex_align="l") |>
  kableExtra::pack_rows("Block B: Accountability Intermediaries", 3, 6, bold=TRUE, latex_align="l") |>
  kableExtra::pack_rows("Secondary (for completeness)",           7, 8, bold=FALSE, latex_align="l") |>
  kableExtra::footnote(
    general="\u2020 p < .10. * p < .05. ** p < .01. *** p < .001. Stars in Change column. Independent cross-sections; Welch t-test.",
    general_title="Note.", threeparttable=TRUE, escape=FALSE
  )
```

```{r fig2, fig.cap="KAMOS Institutional Trust Decline by Category, 2016\\textendash 2019. Blue = executive institutions; red = accountability intermediaries.", out.width="80%", fig.align="center"}
knitr::include_graphics(file.path(fig_dir, "fig2_pub_kamos_trust_decline.pdf"))
```

## Partisan Sorting and the Trust Collapse

A potential alternative explanation for the observed trust decline is partisan realignment: if the 2019 KAMOS wave was administered under a left-liberal Moon government whose supporters were more trusting of institutions than the conservative Park government's supporters had been in 2016, the aggregate trust decline might reflect compositional change in the electorate rather than a structural shift in accountability perceptions. The data do not support this interpretation. Correlations between ideological self-placement and trust across all five key institutional variables are near-zero in both waves, ranging from `r ideo_cor_min` to `r ideo_cor_max`, indicating that ideology explains a negligible share of trust variance at either timepoint. Furthermore, while co-partisan trust effects are evident in central government trust—Minjoo supporters report higher central government trust in 2019, just as Saenuri supporters did in 2016—no such co-partisan premium is evident for media trust, civil society trust, or legislative trust. The collapse of intermediary institution trust is cross-partisan, affecting supporters of both major parties and unaffiliated respondents in roughly equal measure. This cross-partisan character is the strongest evidence that the trust decline reflects a structural feature of the accountability environment rather than an artifact of electoral turnover.

A further observation reinforces this interpretation. Political satisfaction, measured on a four-point scale in both KAMOS waves, changed by only `r polsat_chg` points between 2016 and 2019, effectively flat despite the most dramatic political upheaval in Korea's recent democratic history. Citizens did not become meaningfully more or less satisfied with Korean politics in aggregate following the impeachment and change of government. The combination of flat political satisfaction with sharply declining intermediary institution trust suggests that the accountability shock reorganized citizens' perceptions of the institutional landscape without producing a corresponding shift in their overall political affect. This dissociation is itself theoretically informative: the accountability gap syndrome does not manifest as generalized political disillusionment but as a specific reconfiguration of the trust architecture, concentrated in the institutions whose function is to render power accountable.

```{r table3}
kableExtra::kbl(
  tab3,
  caption  = "Pearson Correlations between Ideological Self-Placement and Institutional Trust, KAMOS Korea 2016 and 2019. Ideology: 1 = extremely liberal to 5 = extremely conservative.",
  booktabs = TRUE, escape = FALSE,
  align    = c("l","c","c","l")
) |>
  kableExtra::kable_styling(latex_options=c("hold_position"), font_size=11, full_width=FALSE) |>
  kableExtra::column_spec(1, width="12em") |>
  kableExtra::footnote(
    general="Near-zero correlations (|r| < .10) indicate the trust decline is not an artifact of ideological composition shifts. No significance stars reported; the finding is the absence of meaningful association.",
    general_title="Note.", threeparttable=TRUE
  )
```

## Korea in Comparative Context

Figure 3 situates Korea's longitudinal trends within the broader ABS sample by plotting the Korean series against the unweighted regional average across all ABS countries with available data. Korea tracks above the regional mean on both variables across the full period, with the gap on the experiential corruption measure widening sharply following the accountability shock. Table 4 formalizes this comparative position, reporting weighted OLS slope estimates for all ABS countries with at least three survey waves. Korea ranks first among `r slopes_inline$gwi_n_cty` countries on the rate of increase in government information withholding (slope = `r slopes_inline$gwi_slope`, z = `r slopes_inline$gwi_z`) and first among `r slopes_inline$cw_n_cty` countries on the rate of increase in experiential corruption perceptions (slope = `r slopes_inline$cw_slope`, z = `r slopes_inline$cw_z`). This cross-national positioning suggests that the dynamics documented here are not merely a generic pattern of declining institutional confidence common across Asia but reflect a specifically Korean trajectory of perceptual change.

```{r fig3, fig.cap="Korea vs. ABS Regional Average: Government Information Withholding and Experiential Corruption Perceptions, 2003\\textendash 2022. Solid line = Korea; dashed = unweighted regional average across all available ABS countries.", out.width="\\textwidth"}
knitr::include_graphics(file.path(fig_dir, "fig3_pub_korea_vs_regional.pdf"))
```

```{r table4-prep, include=FALSE}
format_slope_panel <- function(var_name) {
  slopes_sub <- cross_slopes |> dplyr::filter(variable == var_name)
  summary_row <- slopes_sub |>
    dplyr::summarise(
      Country       = "Cross-national mean (SD)",
      `Slope (sig)` = sprintf("%.4f (%.4f)", dplyr::first(mean_slope), dplyr::first(sd_slope)),
      SE            = "---",
      `z-score`     = "---",
      Waves         = NA_integer_
    )
  slopes_sub |>
    dplyr::arrange(dplyr::desc(slope)) |>
    dplyr::mutate(
      sig = dplyr::case_when(
        p_value < 0.001 ~ "***", p_value < 0.01 ~ "**",
        p_value < 0.05  ~ "*",  p_value < 0.10  ~ "\u2020", TRUE ~ ""
      ),
      Country       = country_name,
      `Slope (sig)` = paste0(sprintf("%.4f", slope), sig),
      SE            = sprintf("%.4f", se),
      `z-score`     = sprintf("%.2f",  z_score),
      Waves         = as.integer(n_waves)
    ) |>
    dplyr::select(Country, `Slope (sig)`, SE, `z-score`, Waves) |>
    dplyr::bind_rows(summary_row)
}

tab4_gwi      <- format_slope_panel("govt_withholds_info")
tab4_cw       <- format_slope_panel("corrupt_witnessed")
korea_row_gwi <- which(tab4_gwi$Country == "Korea")
korea_row_cw  <- which(tab4_cw$Country  == "Korea")
summary_row_gwi <- nrow(tab4_gwi)
summary_row_cw  <- nrow(tab4_cw)
```

```{r table4a}
kableExtra::kbl(
  tab4_gwi,
  caption  = sprintf(
    "Cross-National ABS Slopes: Government Withholds Information. Korea slope = %.4f (SE = %.4f), z = %.2f, rank %d of %d countries. Countries ranked by slope, descending.",
    slopes_inline$gwi_slope, slopes_inline$gwi_se,
    slopes_inline$gwi_z, slopes_inline$gwi_rank, slopes_inline$gwi_n_cty),
  booktabs = TRUE, escape = FALSE,
  align    = c("l","r","r","r","r")
) |>
  kableExtra::kable_styling(latex_options=c("hold_position"), font_size=11, full_width=FALSE) |>
  kableExtra::column_spec(1, width="10em") |>
  kableExtra::row_spec(korea_row_gwi,   bold=TRUE, background="#FFF9C4") |>
  kableExtra::row_spec(summary_row_gwi, italic=TRUE) |>
  kableExtra::footnote(
    general=paste0("\u2020 p < .10. * p < .05. ** p < .01. *** p < .001 (slope estimate). ",
      "Slope = coefficient on wave number, weighted OLS. z-score standardised within variable across ABS sample. ",
      "Countries with <3 waves excluded. Yellow shading = Korea."),
    general_title="Note.", threeparttable=TRUE, escape=FALSE
  )
```

```{r table4b}
kableExtra::kbl(
  tab4_cw,
  caption  = sprintf(
    "Cross-National ABS Slopes: Personally Witnessed/Experienced Corruption. Korea slope = %.4f (SE = %.4f), z = %.2f, rank %d of %d countries.",
    slopes_inline$cw_slope, slopes_inline$cw_se,
    slopes_inline$cw_z, slopes_inline$cw_rank, slopes_inline$cw_n_cty),
  booktabs = TRUE, escape = FALSE,
  align    = c("l","r","r","r","r")
) |>
  kableExtra::kable_styling(latex_options=c("hold_position"), font_size=11, full_width=FALSE) |>
  kableExtra::column_spec(1, width="10em") |>
  kableExtra::row_spec(korea_row_cw,   bold=TRUE, background="#FFF9C4") |>
  kableExtra::row_spec(summary_row_cw, italic=TRUE) |>
  kableExtra::footnote(
    general="Same methodological caveats as Panel A. Yellow shading = Korea.",
    general_title="Note.", threeparttable=TRUE, escape=FALSE
  )
```

# Discussion {#sec-discussion}

The findings reported above cohere into a pattern that is theoretically legible as a mass public correlate of democratic backsliding. Taken together, the ABS longitudinal series and the KAMOS cross-wave comparisons reveal not a generalized withdrawal of political trust but a specific reconfiguration of the citizen-institution relationship—one in which perceptions of governmental opacity accumulate secularly, experiential corruption perceptions shift discontinuously following a high-salience accountability shock, and trust in the intermediary accountability infrastructure collapses while trust in the executive itself remains comparatively stable. This configuration—the accountability gap syndrome—is precisely what the theoretical framework predicts, and its empirical replication across two independent survey instruments strengthens confidence that it reflects a genuine feature of South Korean public opinion rather than a measurement artifact.

The secular trajectory of the govt_withholds_info series deserves particular theoretical attention. The monotonic rise across five waves, uninterrupted by the Park impeachment or the subsequent Moon administration, suggests that transparency perceptions in South Korea are not primarily driven by co-partisan cuing or incumbent-specific evaluation. Instead, they appear to reflect a cumulative process of legitimacy erosion in which each successive crisis adds to a stock of perceptual distrust that subsequent governments do not succeed in reversing. The implication for the study of backsliding extends well beyond the Korean case. If transparency perceptions are insensitive to changes in government—if they rise under conservative and progressive administrations alike—then the accountability gap they reflect may be structural rather than contingent on the identity or behavior of specific incumbents. Democratic backsliding, on this reading, leaves a perceptual residue that outlasts the episodes that produced it and constrains the legitimacy available to successor governments.

The reclassification argument, supported by the divergence between the corrupt_witnessed and institutional corruption assessment series, raises broader questions about how anti-corruption reform interacts with public perception. The Kim Young-ran Act was designed to change behavior by criminalizing practices previously understood as socially obligatory. The survey evidence suggests it also changed perception—that the legal redefinition of corruption reorganized the categories through which citizens evaluated their everyday social environment. This finding has implications beyond the Korean case. In comparative perspective, it suggests that landmark anti-corruption legislation may produce apparent increases in corruption perception even as it succeeds in reducing corrupt behavior, creating a paradox in which reform looks like deterioration in the very survey indicators most commonly used to evaluate it. Standard corruption perception indices, which treat rising scores as evidence of worsening corruption, may systematically misread the perceptual consequences of successful legal intervention. Future research would benefit from closer attention to the legislative and institutional context in which corruption perception surveys are administered, particularly in cases where the legal definition of corruption has recently changed.

The KAMOS findings on intermediary institution trust deepen this picture in theoretically important ways. To this author's knowledge, the differential trust collapse documented here—declines of `r intermed_min` to `r intermed_max` points on the ten-point scale in the National Assembly, the legislature, the media, and civil society organizations against near-stability in central government trust—has not previously been identified in the Korean literature. It suggests that the accountability shock of 2016–2018 damaged not the standing of government per se but the standing of the institutions whose function is to hold government accountable. This is a theoretically important distinction. A citizenry that distrusts the executive is one that may channel its dissatisfaction through accountability mechanisms: voting, protest, media engagement, civil society mobilization. A citizenry that distrusts the accountability mechanisms themselves faces a more fundamental problem: the very instruments through which democratic dissatisfaction is supposed to be expressed and resolved have lost legitimacy. The accountability gap syndrome, in this reading, is not merely a symptom of backsliding but a condition that may sustain it, by eroding the institutional trust necessary for accountability mechanisms to function effectively.

The cross-partisan character of the intermediary institution trust collapse reinforces this interpretation. That the decline in media trust, civil society trust, and legislative trust cuts across partisan lines—affecting Minjoo supporters, Liberty Korea supporters, and unaffiliated respondents in roughly equal measure—suggests that the accountability shock did not simply redistribute trust along partisan lines but eroded it structurally. This cross-partisan erosion is particularly consequential for democratic resilience. Partisan distrust of accountability institutions is a well-documented phenomenon that democratic systems have developed mechanisms to manage; structural distrust that transcends partisan identity is considerably harder to reverse, since it cannot be addressed simply by changing the government.

The difficulty of tracking partisan identity across Korean survey waves itself reflects a structural feature of the accountability environment under analysis. Since democratization in 1987, the major conservative party has undergone seven name changes, with rebranding episodes consistently following episodes of elite scandal or electoral defeat—most recently the transition from Saenuri to Liberty Korea following Park Geun-hye's impeachment, and from Liberty Korea to People Power Party after successive electoral losses. This pattern constitutes what might be termed reputational reconstitution: the deliberate dissolution and organizational rebirth of a party as a collective strategy for retiring reputational liability while preserving elite continuity. Individual-level incentives for this strategy are well-documented in weakly institutionalized party systems, where politicians facing a party brand crisis have incentives to switch parties or support rebranding to escape electoral consequences [@Kim2018-MISSING]. The organizational-level dynamic—in which the party itself is the unit of reputational management—has received less theoretical attention but is arguably the more consequential mechanism for understanding why rising perceptions of governmental opacity and corruption in Korea do not straightforwardly translate into durable partisan realignment. When the accountability carrier dissolves and reconstitutes, the perceptual residue of scandal remains without a stable organizational target.

Several limitations of this analysis merit acknowledgment beyond those noted in the methods section. The empirical window closes in mid-2022, just as the early Yoon administration was taking shape, and therefore does not capture the most acute subsequent episode of Korean democratic stress—the declaration and rapid reversal of martial law in December 2024. Whether the accountability gap syndrome documented here deepened, stabilized, or partially reversed in the period between Wave 6 and the martial law crisis cannot be assessed with the available data, and this represents a significant empirical gap that future research should address as new survey waves become available. Additionally, while the paper argues that the reclassification of corrupt practices under the Kim Young-ran Act contributed to rising corrupt_witnessed scores, this mechanism is inferred from the timing and theoretical logic of the intervention rather than directly demonstrated through evidence about what individual respondents understand the term "corruption" to mean. Qualitative research on the folk semantics of corruption in post-impeachment Korea would substantially strengthen the interpretive claim made here.

# Conclusion {#sec-conclusion}

The evidence assembled here traces the mass public imprint of South Korea's accountability crisis across two decades, combining longitudinal data from the Asian Barometer Survey with cross-wave comparisons from the Korean Academic Multimode Open Survey. The analysis reveals a coherent pattern—the accountability gap syndrome—in which perceptions of governmental information withholding rise secularly and without reversal, experiential corruption perceptions shift discontinuously following a high-salience accountability shock, and trust in intermediary accountability institutions collapses while trust in the executive itself remains comparatively stable. This pattern is replicated across two independent survey instruments, is not ideologically sorted, and is not explained by incumbent-specific evaluation or partisan realignment following the change of government.

The theoretical contribution is twofold. First, it argues that democratic backsliding produces a distinctive mass public signature—not generalized political disillusionment but a specific reconfiguration of the trust architecture in which accountability institutions lose legitimacy faster than the executive institutions they are meant to constrain. This configuration is theoretically consequential because it may sustain the conditions that produced it: a citizenry that distrusts the mechanisms of accountability is less able to deploy those mechanisms effectively, creating a feedback dynamic in which elite-level erosion and mass-level perceptual change reinforce one another. Second, the paper advances a reclassification argument that has implications beyond the Korean case: high-salience accountability shocks, particularly when accompanied by legal redefinition of corruption itself, may reorganize mass perceptual categories in ways that standard survey instruments are liable to misread as behavioral deterioration. The apparent paradox of rising corruption perceptions under a reform administration is, on this reading, evidence of the reform's cultural reach rather than its failure.

South Korea is a theoretically important case for the comparative study of democratic backsliding precisely because its institutional resilience has been held up as a model of successful democratic self-correction. The constitutional accountability mechanisms functioned: the impeachment proceeded, the prosecution succeeded, and electoral democracy was restored. Yet the survey evidence examined here suggests that this institutional success did not translate into perceptual restoration at the mass public level. Transparency concerns continued to accumulate, intermediary institution trust did not recover, and the accountability gap that opened during the crisis period shows no sign of closing in the available data. Democratic resilience, this case suggests, may be easier to achieve at the institutional level than at the level of mass belief—and the divergence between the two may itself represent an underappreciated dimension of backsliding's long-term democratic cost.

Future research should extend this analysis in at least three directions. The first is temporal: the empirical window examined here closes in mid-2022, before the most acute subsequent episode of Korean democratic stress, and new survey data covering the Yoon period and its aftermath will be essential for assessing whether the accountability gap syndrome deepened, stabilized, or partially reversed. The second is comparative: the accountability gap syndrome identified here may not be unique to South Korea, and systematic cross-national analysis of the mass public correlates of backsliding episodes across Asia and beyond would substantially advance understanding of the generalizability of the mechanisms proposed. The third is interpretive: the reclassification argument developed here rests on inferential reasoning about the effects of the Kim Young-ran Act on popular corruption categories, and direct qualitative evidence on how Korean citizens understand and apply the concept of corruption in the post-impeachment period would considerably strengthen the theoretical claim. These are limitations of the present analysis, but they are also a research agenda—one that the accountability gap framework is well positioned to organize.

# References {.unnumbered}

::: {#refs}
:::
