---
title: "Trust in Information and Pandemic Approval in Southeast Asia: Evidence from the Vietnam Paradox"
author:
  - name: Jeffrey Stark
    affiliation: Yonsei University
    email: jeffrey.stark@yonsei.ac.kr
date: today
abstract: |
  Vietnam's COVID-19 experience defies performance legitimacy theory. In Asian Barometer Wave 6 data (2021--2022, N = 3,685), 65.9% of Vietnamese respondents reported personal or family COVID-19 infection---more than seven times Cambodia's rate and the highest in Southeast Asia. Yet 97.5% rated the government's pandemic response favourably, also the region's highest. The highest infection burden coincided with the most favourable evaluations. Trust in government COVID-19 information, not personal experience, is the dominant predictor of these assessments. Moving from lowest to highest information trust corresponds to roughly 1.5 points on a 4-point approval scale; contracting the virus corresponds to less than 0.08 points. The pattern holds across Cambodia, Thailand, and Vietnam, survives controls for institutional trust, democratic attitudes, and demographics, and remains stable across alternative specifications; sensitivity analysis indicates that unmeasured confounders would need to be implausibly strong to account for the findings (see Online Appendix). Thailand provides a critical mechanism test: among citizens who distrust official information, infection is negatively associated with approval as performance theories predict. Among those who trust it, infection is irrelevant. Information credibility appears to function as a cognitive filter---when official narratives are deemed trustworthy, citizens may attribute suffering to forces beyond government control rather than to governmental shortcomings.
keywords:
  - COVID-19
  - government approval
  - information trust
  - Southeast Asia
  - pandemic politics
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    embed-resources: true
  pdf:
    toc: false
    pdf-engine: lualatex
    documentclass: article
    geometry: margin=1in
    fontsize: 12pt
    linestretch: 2
    number-sections: true
    fig-pos: 'H'
    tbl-pos: 'H'
    header-includes: |
      \usepackage{float}
      \usepackage{placeins}
      \usepackage{setspace}
      \raggedright
      \hyphenpenalty=10000
      \exhyphenpenalty=10000
      \AtBeginEnvironment{CSLReferences}{\singlespacing\raggedright}
  docx:
    toc: false
    number-sections: true
    reference-doc: reference.docx
citeproc: true
bibliography: references.bib
csl: sage-harvard.csl
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_knit:
    root.dir: "."
repo-actions: false
---

```{r setup, include=FALSE}
# ============================================================================
# MANUSCRIPT SETUP: Load Pre-Computed Analysis Results
# ============================================================================

# Load packages
library(tidyverse)
library(here)
library(gt)
library(gtExtras)
library(modelsummary)
library(knitr)
library(flextable)

# Set global options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6,
  dpi = 300
)

# Set theme for plots
theme_set(theme_minimal(base_size = 12))

# Detect output format for conditional table rendering
is_docx <- knitr::is_latex_output() == FALSE && knitr::is_html_output() == FALSE

# Helper: render table as flextable for docx, gt for others
render_table <- function(df, caption = NULL, col_labels = NULL) {
  if (is_docx) {
    # Use flextable for DOCX with auto-fit widths
    if (!is.null(col_labels)) {
      names(df) <- col_labels[names(df)]
    }
    df %>%
      flextable::flextable() %>%
      flextable::autofit() %>%
      flextable::set_caption(caption)
  } else {
    # Use gt for HTML/PDF
    tbl <- df %>% gt()
    if (!is.null(col_labels)) {
      tbl <- tbl %>% cols_label(!!!col_labels)
    }
    tbl
  }
}

# Define results directory path
results_dir <- file.path("..", "analysis", "results")

# ============================================================================
# FUNCTION: Safely load RDS files
# ============================================================================
safe_load <- function(filename) {
  filepath <- file.path(results_dir, filename)
  if (file.exists(filepath)) {
    readRDS(filepath)
  } else {
    NULL
  }
}

# ============================================================================
# FUNCTION: Format wide tables for both HTML and PDF output
# ============================================================================
format_wide_gt <- function(gt_table) {
  gt_table %>%
    tab_options(
      table.font.size = px(9),
      data_row.padding = px(2),
      table.width = pct(100),
      column_labels.font.size = px(10),
      heading.title.font.size = px(12),
      heading.subtitle.font.size = px(10),
      source_notes.font.size = px(8),
      footnotes.font.size = px(8)
    )
}

# ============================================================================
# LOAD ALL ANALYSIS RESULTS
# ============================================================================

# Descriptive results
paradox_summary <- safe_load("descriptive_paradox_summary.rds")
descriptive_stats <- safe_load("descriptive_sample_characteristics.rds")
scale_reliability <- safe_load("descriptive_scale_reliability.rds")
cor_matrix <- safe_load("descriptive_correlation_matrix.rds")
bootstrap_cors <- safe_load("descriptive_bootstrap_correlations.rds")
effect_sizes <- safe_load("descriptive_effect_sizes.rds")
economic_impacts <- safe_load("descriptive_economic_impacts.rds")

# Hypothesis test results
h1_results <- safe_load("h1_infection_effects.rds")
h1c_model <- safe_load("h1c_interaction_model.rds")
h2a_results <- safe_load("h2a_trust_direct_effects.rds")
h2b_results <- safe_load("h2b_mediation_paths.rds")
h3a_results <- safe_load("h3a_democracy_effects.rds")
h3b_results <- safe_load("h3b_authoritarianism_effects.rds")
h4a_results <- safe_load("h4a_economic_bivariate.rds")
h4b_results <- safe_load("h4b_trust_vs_economic.rds")
models_core <- safe_load("m1_core_models.rds")
models_full <- safe_load("m2_full_models.rds")
hypothesis_summary <- safe_load("hypothesis_summary.rds")

# Coefficient dataframes for consistent table generation (same as appendix)
m1_coefficients <- safe_load("m1_coefficients.rds")
m2_coefficients <- safe_load("m2_coefficients.rds")
m1_fit <- safe_load("m1_fit_statistics.rds")
m2_fit <- safe_load("m2_fit_statistics.rds")

# Mediation results
mediation_paths <- safe_load("med_path_coefficients.rds")
mediation_bootstrap <- safe_load("med_bootstrap_results.rds")
mediation_summary <- safe_load("med_summary.rds")
mediation_sensitivity <- safe_load("med_sensitivity.rds")
mediation_low_auth <- safe_load("med_low_auth.rds")
mediation_high_auth <- safe_load("med_high_auth.rds")
mediation_countries <- safe_load("med_cross_country.rds")

# Country comparison results
vietnam_correlations <- safe_load("country_vietnam_correlations.rds")
cambodia_correlations <- safe_load("country_cambodia_correlations.rds")
thailand_correlations <- safe_load("country_thailand_correlations.rds")
fisher_infection_approval <- safe_load("country_fisher_infection_approval.rds")
fisher_trust_approval <- safe_load("country_fisher_trust_approval.rds")
prop_mediated <- safe_load("country_proportion_mediated.rds")
pub_summary_table <- safe_load("country_summary_table.rds")

# Robustness check results
missing_data_summary <- safe_load("robust_missing_summary.rds")
ordinal_vs_ols <- safe_load("robust_ordinal_vs_ols.rds")
robust_comparison <- safe_load("robust_regression_comparison.rds")
quantile_comparison <- safe_load("robust_quantile_regression.rds")
outlier_diagnostics <- safe_load("robust_outlier_diagnostics.rds")
outlier_sensitivity <- safe_load("robust_excluding_outliers.rds")
complete_case_test <- safe_load("robust_complete_case_test.rds")
robustness_summary <- safe_load("robust_all_checks_summary.rds")

# Endogeneity diagnostics - using relative paths like other results
inline_stats_dir <- file.path("..", "analysis", "output", "inline_stats")
cor_trust_approval <- readRDS(file.path(inline_stats_dir, "cor_trust_approval.rds"))
vif_max <- readRDS(file.path(inline_stats_dir, "vif_max.rds"))
vif_min <- readRDS(file.path(inline_stats_dir, "vif_min.rds"))
pct_trust_disapprove <- readRDS(file.path(inline_stats_dir, "pct_trust_disapprove.rds"))

# ============================================================================
# NEW: EXTRACT KEY VALUES FOR INLINE USE THROUGHOUT MANUSCRIPT
# ============================================================================

# Helper function to safely extract values with fallback
get_value <- function(data, country, column, fallback) {
  if (is.null(data)) return(fallback)
  value <- data %>% dplyr::filter(country_name == country) %>% dplyr::pull({{ column }})
  if (length(value) == 0 || is.na(value[1])) return(fallback)
  value[1]
}

# VIETNAM VALUES (with fallbacks)
vietnam <- list(
  n = get_value(paradox_summary, "Vietnam", n, 1237),
  infection_pct = get_value(paradox_summary, "Vietnam", covid_infection_rate, 65.9),
  approval_pct = get_value(paradox_summary, "Vietnam", govt_approval_rate, 97.5),
  trust_pct = get_value(paradox_summary, "Vietnam", trust_covid_rate, 91.9)
)

# CAMBODIA VALUES (with fallbacks)
cambodia <- list(
  n = get_value(paradox_summary, "Cambodia", n, 1242),
  infection_pct = get_value(paradox_summary, "Cambodia", covid_infection_rate, 8.7),
  approval_pct = get_value(paradox_summary, "Cambodia", govt_approval_rate, 93.6),
  trust_pct = get_value(paradox_summary, "Cambodia", trust_covid_rate, 89.2)
)

# THAILAND VALUES (with fallbacks)
thailand <- list(
  n = get_value(paradox_summary, "Thailand", n, 1206),
  infection_pct = get_value(paradox_summary, "Thailand", covid_infection_rate, 40.1),
  approval_pct = get_value(paradox_summary, "Thailand", govt_approval_rate, 37.7),
  trust_pct = get_value(paradox_summary, "Thailand", trust_covid_rate, 34.0)
)

# TOTAL SAMPLE SIZE
total_n <- if (!is.null(paradox_summary)) sum(paradox_summary$n, na.rm = TRUE) else 3685

# EXTERNAL METADATA (Freedom House, Press Freedom, Response Rates)
metadata <- list(
  # Freedom House democracy scores (/100)
  cambodia_fh = 24,
  vietnam_fh = 19,
  thailand_fh = 29,
  # Press Freedom scores (/100)
  cambodia_pf = 23,
  vietnam_pf = 20,
  thailand_pf = 29,
  # Survey response rates (%)
  cambodia_response = 82,
  thailand_response = 68,
  vietnam_response = NA
)

# CALCULATED RATIOS (for narrative comparisons)
vietnam_vs_cambodia_infection_ratio <- vietnam$infection_pct / cambodia$infection_pct

# ============================================================================
# HELPER FUNCTIONS FOR CLEANER MANUSCRIPT CODE
# ============================================================================

# Format correlation with significance stars
format_cor <- function(cor_value, p_value) {
  stars <- if (p_value < 0.001) "***" else if (p_value < 0.01) "**" else if (p_value < 0.05) "*" else ""
  paste0("r = ", round(cor_value, 3), stars)
}

# Create correlation table from results list
create_cor_table <- function(results_list) {
  if (is.null(results_list)) {
    return(tibble(Note = "Data not available. Render analysis scripts first."))
  }
  
  tibble(
    Country = names(results_list),
    Correlation = sapply(results_list, function(x) x$coef),
    `p-value` = sapply(results_list, function(x) x$p),
    `95% CI Lower` = sapply(results_list, function(x) x$ci_lower),
    `95% CI Upper` = sapply(results_list, function(x) x$ci_upper),
    N = sapply(results_list, function(x) x$n)
  )
}

# Create placeholder table when data not available
create_placeholder <- function(table_name, script_to_run) {
  tbl_data <- tibble(
    Status = paste("Missing:", table_name),
    Action = paste("Render:", script_to_run)
  )
  if (is_docx) {
    tbl_data %>%
      flextable::flextable() %>%
      flextable::set_caption("Data Not Yet Available") %>%
      flextable::autofit()
  } else {
    tbl_data %>%
      gt() %>%
      tab_header(title = "Data Not Yet Available") %>%
      tab_style(
        style = cell_fill(color = "#fff3cd"),
        locations = cells_body()
      )
  }
}

# CALCULATE DESCRIPTIVE STATISTICS FOR VARIABLES
if (file.exists(here("papers", "01_vietnam_covid_paradox", "analysis", "data", "analysis_data.rds"))) {
  ab_anal_data <- readRDS(here("papers", "01_vietnam_covid_paradox", "analysis", "data", "analysis_data.rds"))
  
  var_stats <- list(
    # DV: Government approval
    approval_mean = mean(ab_anal_data$covid_govt_handling, na.rm = TRUE),
    approval_sd = sd(ab_anal_data$covid_govt_handling, na.rm = TRUE),
    
    # IV: Infection
    infection_rate = mean(ab_anal_data$covid_contracted, na.rm = TRUE) * 100,
    
    # IV: Economic severity
    econ_mean = mean(ab_anal_data$covid_impact_severity, na.rm = TRUE),
    econ_sd = sd(ab_anal_data$covid_impact_severity, na.rm = TRUE),
    
    # Individual economic impacts
    job_loss_pct = mean(ab_anal_data$covid_job_loss, na.rm = TRUE) * 100,
    income_loss_pct = mean(ab_anal_data$covid_income_loss, na.rm = TRUE) * 100,
    edu_disruption_pct = mean(ab_anal_data$covid_edu_disruption, na.rm = TRUE) * 100,
    
    # IV: Trust
    trust_mean = mean(ab_anal_data$covid_trust_info, na.rm = TRUE),
    trust_sd = sd(ab_anal_data$covid_trust_info, na.rm = TRUE),
    
    # Demographics
    age_mean = mean(ab_anal_data$age, na.rm = TRUE),
    age_sd = sd(ab_anal_data$age, na.rm = TRUE),
    female_pct = mean(ab_anal_data$gender == "Female", na.rm = TRUE) * 100,
    urban_pct = mean(ab_anal_data$urban == 1, na.rm = TRUE) * 100,

    # Scale statistics (for Methods section)
    institutional_trust_mean = mean(ab_anal_data$institutional_trust_index, na.rm = TRUE),
    institutional_trust_sd = sd(ab_anal_data$institutional_trust_index, na.rm = TRUE),
    auth_acceptance_mean = mean(ab_anal_data$auth_acceptance, na.rm = TRUE),
    auth_acceptance_sd = sd(ab_anal_data$auth_acceptance, na.rm = TRUE),
    emergency_powers_mean = mean(ab_anal_data$emergency_powers_index, na.rm = TRUE),
    emergency_powers_sd = sd(ab_anal_data$emergency_powers_index, na.rm = TRUE),
    regime_pref_mean = mean(ab_anal_data$regime_preference, na.rm = TRUE),
    regime_pref_sd = sd(ab_anal_data$regime_preference, na.rm = TRUE)
  )
} else {
  # Fallback values
  var_stats <- list(
    approval_mean = 3.29, approval_sd = 0.87,
    infection_rate = 39.8,
    econ_mean = 2.31, econ_sd = 0.94,
    job_loss_pct = 9.8, income_loss_pct = 55.2, edu_disruption_pct = 27.4,
    trust_mean = 3.21, trust_sd = 0.79,
    age_mean = 43.6, age_sd = 15.8,
    female_pct = 51.2, urban_pct = 38.4,
    # Scale statistics fallbacks
    institutional_trust_mean = 28.2, institutional_trust_sd = 5.8,
    auth_acceptance_mean = 2.34, auth_acceptance_sd = 0.68,
    emergency_powers_mean = 2.96, emergency_powers_sd = 0.61,
    regime_pref_mean = 2.52, regime_pref_sd = 0.59
  )
}

# ============================================================================
# SPEECH FREEDOM STATISTICS (for SDB defense in Discussion)
# ============================================================================

# Calculate speech freedom stats for Vietnam (used in SDB defense)
if (exists("ab_anal_data") && "freedom_of_speech" %in% names(ab_anal_data)) {
  # Filter to Vietnam with valid responses on both variables
  vietnam_sdb <- ab_anal_data %>%
    filter(country_name == "Vietnam",
           freedom_of_speech %in% c(1, 2, 3, 4),
           covid_govt_handling %in% c(1, 2, 3, 4))
  
  # Calculate key statistics
  speech_freedom_stats <- list(
    pct_constrained = mean(vietnam_sdb$freedom_of_speech <= 2, na.rm = TRUE) * 100,
    approval_among_constrained = vietnam_sdb %>%
      filter(freedom_of_speech <= 2) %>%
      summarise(pct = mean(covid_govt_handling >= 3, na.rm = TRUE) * 100) %>%
      pull(pct),
    approval_among_free = vietnam_sdb %>%
      filter(freedom_of_speech >= 3) %>%
      summarise(pct = mean(covid_govt_handling >= 3, na.rm = TRUE) * 100) %>%
      pull(pct)
  )
} else {
  # Fallback values based on analysis
  speech_freedom_stats <- list(
    pct_constrained = 40.3,
    approval_among_constrained = 97.5,
    approval_among_free = 97.3
  )
}

# Extract all key correlation values
cors <- list(
  # Infection-Approval
  vietnam_inf = if (!is.null(h1_results)) h1_results$Vietnam$coef else 0.013,
  cambodia_inf = if (!is.null(h1_results)) h1_results$Cambodia$coef else -0.014,
  thailand_inf = if (!is.null(h1_results)) h1_results$Thailand$coef else -0.079,
  
  # Trust-Approval
  vietnam_trust = if (!is.null(h2a_results)) h2a_results$Vietnam$coef else 0.507,
  cambodia_trust = if (!is.null(h2a_results)) h2a_results$Cambodia$coef else 0.576,
  thailand_trust = if (!is.null(h2a_results)) h2a_results$Thailand$coef else 0.672,
  
  # Economic-Approval
  vietnam_econ = if (!is.null(h4b_results)) h4b_results$Vietnam$econ_coef else -0.016,
  cambodia_econ = if (!is.null(h4b_results)) h4b_results$Cambodia$econ_coef else 0.156,
  thailand_econ = if (!is.null(h4b_results)) h4b_results$Thailand$econ_coef else -0.142
)

# Calculate ranges
econ_range_min <- min(cors$vietnam_econ, cors$cambodia_econ, cors$thailand_econ)
econ_range_max <- max(cors$vietnam_econ, cors$cambodia_econ, cors$thailand_econ)

# Calculate ratios (trust vs infection)
cors$vietnam_ratio <- abs(cors$vietnam_trust / cors$vietnam_inf)
cors$cambodia_ratio <- abs(cors$cambodia_trust / cors$cambodia_inf)
cors$thailand_ratio <- abs(cors$thailand_trust / cors$thailand_inf)

# ============================================================================
# CRONBACH'S ALPHA VALUES FROM SCALE RELIABILITY RESULTS
# ============================================================================

# Helper function to extract alpha values
get_alpha <- function(scale_name) {
  if (is.null(scale_reliability)) return(NA)
  alpha <- scale_reliability %>%
    filter(grepl(scale_name, scale, ignore.case = TRUE)) %>%
    pull(alpha)
  if (length(alpha) == 0) return(NA)
  return(alpha[1])
}

alphas <- list(
  institutional_trust = get_alpha("Institutional Trust"),
  auth_acceptance = get_alpha("Auth Acceptance"),
  emergency_powers = get_alpha("Emergency"),
  regime_pref = get_alpha("Regime")
)

# ============================================================================
# COMPUTED RATIOS, RANGES, AND HELPER FUNCTIONS
# ============================================================================

# Magnitude ratios for narrative (trust effects vs infection effects)
magnitude_ratios <- list(
  cambodia = abs(cors$cambodia_trust / cors$cambodia_inf),
  vietnam = abs(cors$vietnam_trust / cors$vietnam_inf),
  thailand = abs(cors$thailand_trust / cors$thailand_inf)
)

# Correlation ranges for "r = X to Y" statements in text
cor_ranges <- list(
  infection_max = max(abs(cors$vietnam_inf), abs(cors$cambodia_inf), abs(cors$thailand_inf)),
  economic_min = min(cors$vietnam_econ, cors$cambodia_econ, cors$thailand_econ),
  economic_max = max(cors$vietnam_econ, cors$cambodia_econ, cors$thailand_econ),
  trust_min = min(cors$vietnam_trust, cors$cambodia_trust, cors$thailand_trust),
  trust_max = max(cors$vietnam_trust, cors$cambodia_trust, cors$thailand_trust)
)

# Helper function for formatting regression coefficients in narrative
format_coef <- function(coef, p, digits = 3) {
  if (is.null(coef) || is.null(p) || length(coef) == 0 || length(p) == 0) {
    return("[Results not yet available]")
  }
  p_str <- if (p < 0.001) "p < 0.001"
           else if (p < 0.01) "p < 0.01"
           else if (p < 0.05) "p < 0.05"
           else paste0("p = ", round(p, 2))
  paste0("β = ", round(coef, digits), ", ", p_str)
}

# Bootstrap parameters
bootstrap_n <- 2000  # Number of bootstrap resamples used

# Model coefficient ranges from models_full (trust coefficients)
model_coef_ranges <- if (!is.null(models_full)) {
  list(
    trust_coef_min = min(
      coef(models_full$Cambodia)["covid_trust_info"],
      coef(models_full$Vietnam)["covid_trust_info"],
      coef(models_full$Thailand)["covid_trust_info"]
    ),
    trust_coef_max = max(
      coef(models_full$Cambodia)["covid_trust_info"],
      coef(models_full$Vietnam)["covid_trust_info"],
      coef(models_full$Thailand)["covid_trust_info"]
    )
  )
} else {
  list(trust_coef_min = 0.31, trust_coef_max = 0.61)
}

# ============================================================================
# PLACEBO TEST: Does COVID info trust predict non-pandemic outcomes equally?
# ============================================================================

if (exists("ab_anal_data")) {
  placebo_results <- list()
  
  for (cntry in c("Cambodia", "Vietnam", "Thailand")) {
    cdata <- ab_anal_data %>% filter(country_name == cntry)
    
    # Target model: trust -> pandemic approval (the real relationship)
    m_target <- lm(covid_govt_handling ~ covid_trust_info, data = cdata)
    
    # Placebo 1: trust -> democracy satisfaction
    m_placebo1 <- lm(dem_satisfaction ~ covid_trust_info, data = cdata)
    
    # Placebo 2: trust -> institutional trust index
    m_placebo2 <- lm(institutional_trust_index ~ covid_trust_info, data = cdata)
    
    placebo_results[[cntry]] <- list(
      target_r2 = summary(m_target)$r.squared,
      target_beta = coef(m_target)["covid_trust_info"],
      placebo_demsat_r2 = summary(m_placebo1)$r.squared,
      placebo_demsat_beta = coef(m_placebo1)["covid_trust_info"],
      placebo_insttrust_r2 = summary(m_placebo2)$r.squared,
      placebo_insttrust_beta = coef(m_placebo2)["covid_trust_info"],
      ratio_demsat = summary(m_target)$r.squared / summary(m_placebo1)$r.squared,
      ratio_insttrust = summary(m_target)$r.squared / summary(m_placebo2)$r.squared
    )
  }
  
  # Compute range of ratios for inline reporting
  all_ratios <- c(
    sapply(placebo_results, function(x) x$ratio_demsat),
    sapply(placebo_results, function(x) x$ratio_insttrust)
  )
  placebo_ratio_min <- round(min(all_ratios), 1)
  placebo_ratio_max <- round(max(all_ratios), 1)
  
} else {
  placebo_results <- NULL
  placebo_ratio_min <- 2
  placebo_ratio_max <- 4
}

cat("✓ Setup complete\n")
```

# Introduction

In August 2022, as the Omicron wave swept through Vietnam's cities, survey enumerators from the Asian Barometer project fanned out across the country carrying a standard battery of questions about governance and public life. One question was straightforward: how well or badly had the government handled the pandemic? The responses ought to trouble anyone who believes citizens punish governments for poor outcomes.

Almost two-thirds (`r round(vietnam$infection_pct, 1)`%) of Vietnamese respondents reported that they or a family member had contracted COVID-19---more than seven times Cambodia's `r round(cambodia$infection_pct, 1)`%, and substantially above Thailand's `r round(thailand$infection_pct, 1)`%. By any epidemiological standard, Vietnam bore the heaviest pandemic burden in the region. Yet `r round(vietnam$approval_pct, 1)`% of those respondents rated the government's response as "well" or "very well." Highest infection rate, highest approval. The correlation runs backwards.

This is not merely a Vietnamese peculiarity. It speaks to a broader puzzle that has occupied scholars of institutional trust across Asia, where lower-performing countries routinely report higher government trust than objectively better-governed democracies [@Baniamin2025-rx; @Wong2011-zv; @Zhao2017-yw]. Baniamin [-@Baniamin2025-rx], working with Asian Barometer data across fourteen countries, attributes this pattern primarily to authoritarian cultural orientation---citizens socialized into hierarchical deference evaluate authorities less critically. This article proposes a complementary but analytically distinct mechanism. Trust in *crisis-specific* government information may function as a cognitive filter that shapes how citizens interpret their own experiences, independently of whatever diffuse cultural values they hold.

The core claim is this: citizens who trust official COVID-19 information appear to attribute negative pandemic experiences---personal infection, economic hardship---to forces beyond government control. The virus, global circumstances, misfortune. Citizens who distrust official information read those same experiences as evidence of governmental shortcomings. What matters for political evaluation is not whether you got sick but how you make sense of getting sick. And that sense-making is powerfully shaped by whether you find the government's account of events credible.

I test this argument with Asian Barometer Wave 6 data from Cambodia, Thailand, and Vietnam (N = `r format(total_n, big.mark=",")`). The three countries share a regional context but diverge sharply on the dimensions that matter: pandemic outcomes, regime type, and information environment. Vietnam and Cambodia are authoritarian states with restricted media; Thailand is a hybrid regime with greater information pluralism. Vietnam experienced mass infection; Cambodia largely avoided it. If performance drives legitimacy, Cambodia should outperform Vietnam on approval. It does not. If information environment matters, Vietnam and Cambodia---both high-trust settings---should converge despite divergent outcomes, while Thailand should diverge. They do.

The empirical results are stark. Trust in government COVID-19 information predicts approval at β = `r round(model_coef_ranges$trust_coef_min, 2)` to `r round(model_coef_ranges$trust_coef_max, 2)` across countries, explaining 26--45% of variance. Personal infection? Essentially zero. A respondent moving from lowest to highest information trust gains approximately 1.5 points on the 4-point scale---nearly half the entire range. Whether that respondent contracted COVID-19 shifts their evaluation by less than 0.08 points.

The study makes three contributions. First, it documents an empirical pattern in which information trust dominates experiential predictors of pandemic approval, extending recent scholarship on trust paradoxes in authoritarian settings [@Baniamin2025-rx; @Pernia2022-gh; @Genova2025-gr] to crisis-specific evaluations where the stakes of performance assessment are unusually clear. Second, it proposes an information-credibility heuristic---grounded in theories of informational autocracy [@Guriev2019-xj; @Morgenbesser2020-ya] and the broader political economics of non-democratic survival [@Egorov2024-gq]---that specifies *how* citizens process negative experiences when official narratives are trusted. Third, the Thailand comparison demonstrates that the performance-legitimacy link is not absent in Southeast Asia but *conditional*: it operates normally among citizens who distrust official information, and breaks down where trust is high. The mechanism is informational, not cultural.

# Theoretical Framework

## Performance Legitimacy Under Stress

The foundational claim of performance legitimacy theory is disarmingly simple: citizens reward governments that deliver and punish those that fail [@Lipset1959-ur; @Rothstein2009-lr; @Easton1979-rc]. Institutional trust reflects institutional performance. When governments deliver public goods effectively, citizens reciprocate with support [@Bouckaert2003-ky; @Hetherington2003-oq]. During pandemics, the expectation follows directly---governments that protect citizens should enjoy higher approval than those that do not.

Evidence from Asia complicates this picture. Vietnam routinely reports the highest trust in government despite middling performance on most governance indices, while Japan and South Korea---by objective measures far better governed---report substantially lower trust [@Baniamin2025-rx; @Wong2011-zv]. Baniamin [-@Baniamin2025-rx] attributes the pattern to authoritarian cultural orientation: citizens socialized into hierarchical deference approach authority less critically, inflating trust measures independently of performance. A latent profile analysis across six Southeast Asian countries corroborates this, identifying a dominant high-trust profile in Vietnam that encompasses government, science, and vaccine confidence and cuts across demographic subgroups [@Genova2025-gr].

These findings sit uneasily with performance theory. But Vietnam's COVID-19 experience presents a starker challenge still, because the performance shortfall is not a matter of abstract governance indicators---it is concrete, personal, epidemiological. The country recorded the region's highest self-reported infection rate yet maintained its highest approval for pandemic handling. If performance legitimacy operates as theorized, Vietnam should show *lower* approval than Cambodia, which successfully controlled the virus. Cambodia's `r round(cambodia$infection_pct, 1)`% infection rate and `r round(cambodia$approval_pct, 1)`% approval are entirely consistent with performance theory. Vietnam's `r round(vietnam$infection_pct, 1)`% and `r round(vietnam$approval_pct, 1)`% are not.

## An Information-Credibility Heuristic

The mechanism proposed here is analytically distinct from diffuse cultural orientations, though it may operate alongside them. Citizens rarely have direct knowledge of government competence across policy domains. They rely on informational signals and cognitive shortcuts [@Lupia1994-jp; @Sniderman1991-uj; @Popkin1994-yv], and during crises---when uncertainty peaks and personal stakes are acute---these shortcuts matter disproportionately.

The specific shortcut I identify is an *information-credibility heuristic*: citizens use their trust in official crisis information as a proxy for evaluating government performance. This is not merely a restatement of the propaganda-works thesis. It is a claim about attribution. Negative pandemic experiences are inherently ambiguous---if I contract COVID-19, that event is open to multiple interpretations. Personal misfortune. The inevitable cost of a global catastrophe. Evidence that my government did not adequately protect me. Which frame dominates depends substantially on whether I find the government's account of the crisis convincing.

The theoretical resonance is with Guriev and Treisman's [-@Guriev2019-xj; -@Guriev2020-cu] work on informational autocracy. Modern authoritarian regimes, they argue, sustain power less through overt coercion than through information management and narrative dominance. Egorov and Sonin [-@Egorov2024-gq] locate information flows at the centre of authoritarian survival in their comprehensive review of non-democratic political economics---dictators restrict media, run propaganda, and organize censorship not as incidental features of repression but as the primary mechanisms through which regimes endure. Morgenbesser [-@Morgenbesser2020-ya] situates such informational strategies within a broader "menu of autocratic innovation," a repertoire of tools through which regimes cultivate the pretence of accountability without its practice.

The informational autocracy literature operates at the macro level: regimes deploy propaganda, citizens remain uninformed, regimes survive. The contribution here is to specify the micro-level cognitive pathway. How does trusted official information translate into sustained approval even when citizens have direct experiential evidence that things went badly? The answer lies in attributional ambiguity. Iyengar [-@Iyengar1994-fk] demonstrated that how political events are framed---as isolated incidents or systemic patterns---determines whether citizens assign responsibility to individuals or institutions. Healy and Malhotra [-@Healy2013-sy] showed that voters routinely punish incumbents for outcomes beyond governmental control: natural disasters, global economic shocks, even football results. Trusted official information may work precisely by helping citizens make the distinction that Healy and Malhotra's voters fail to make---attributing suffering to the virus rather than to governmental shortcomings.

This mechanism should operate asymmetrically across information environments. As used throughout this article, *information environment* refers to the constellation of structural conditions shaping citizens' access to and interpretation of political information: media pluralism (the number and independence of news sources), state capacity for narrative control (censorship infrastructure, press freedom restrictions, and official messaging reach), and the availability of oppositional or alternative frames. These structural features are not measured directly in the survey data---they serve as country-level contextual factors that I expect to shape the distribution of individual-level trust, which I do measure. In pluralistic settings with competing narratives, citizens must adjudicate between frames, and those who distrust official information will weight critical accounts more heavily. In restricted environments, the official frame faces less competition, consistent with research on how state-controlled media shape citizens' interpretive frameworks even where alternative sources technically exist [@Stockmann2012-xq; @Rozenas2019-jk].

Two features distinguish this mechanism from Baniamin's [-@Baniamin2025-rx] authoritarian cultural orientation. The first is domain specificity: the focus here is trust in crisis-specific government information, not generalized deference to authority. The second is mediation: information trust actively shapes how citizens interpret negative experiences, rather than simply inflating all government assessments uniformly. If the argument is correct, the coefficient on COVID information trust should remain substantial even after controlling for institutional trust and regime attitudes---proxies for the generalized deference Baniamin identifies. I test this directly.

These considerations yield two hypotheses:

> **H1 (Information Dominance):** Trust in official COVID-19 information will be positively associated with government pandemic approval, net of institutional trust, democratic attitudes, and demographic controls.

> **H2 (Conditional Accountability):** The relationship between personal COVID-19 infection and approval will be conditional on information trust---negative among citizens who distrust official information but attenuated or null among citizens who trust it.

H1 tests whether crisis-specific trust predicts approval above and beyond diffuse orientations toward the regime. H2 tests the core claim: that trusted narratives insulate governments from experiential accountability.

## Alternative Explanations

Several plausible alternatives cannot be cleanly eliminated with cross-sectional data.

*Authoritarian cultural orientation* may explain high approval in Vietnam and Cambodia independently of anything specific to information trust [@Dalton2005-su; @Shi2001-ga; @Ma2014-jj]. If so, controlling for institutional trust and authoritarian acceptance should substantially attenuate the information trust coefficient. This is a testable implication.

*Rally-around-the-flag effects* predict surges in government support during external threats [@Mueller1970-kf; @Baker2001-kz; @Baum2002-tc], though rally theory in its canonical form expects support to strengthen when governments *succeed* at managing threats---not when citizens bear the heaviest burden of the threat itself. Longitudinal evidence from Germany complicates the rally account further: Zoch and Wamsler [-@Zoch2024-kf] show that initial rally effects during COVID-19 eroded significantly over the pandemic's duration, implying that sustained high approval of the kind observed in Vietnam requires an active mechanism beyond a one-time boost. Vietnam's high infection rates combined with persistently high approval inverts the expected rally pattern.

*Prior regime legitimacy* from Vietnam's post-*Đổi Mới* economic transformation may account for baseline support levels [@Malesky2014-wz]. This is plausible as a contributor to generalized trust, but it cannot explain why COVID-*specific* infection and economic hardship show such feeble associations with COVID-*specific* approval within Vietnam. If legacy effects operate, they should manifest primarily through the institutional trust variable, which the models control for.

*Social desirability bias* is the most serious concern. Respondents in authoritarian settings may report approval they do not feel. Section 5 addresses this at length, but two observations bear noting here: the Asian Barometer employs standardised anonymity protocols across all three countries, and Vietnamese respondents openly reported the region's highest infection rates alongside substantial disagreement that people are free to speak without fear---responses difficult to reconcile with uniform social desirability driving all pro-government answers.

Thailand provides the crucial comparative leverage. If something common to Southeast Asian culture explains the pattern---hierarchical values, Confucian deference, Buddhist equanimity---then Thailand, sharing the regional context, should show similar approval dynamics. It does not. Thailand's more pluralistic information environment and lower baseline trust create the conditions under which H2 can be tested most cleanly: does infection reduce approval among citizens who distrust official information? The answer is yes.

# Data and Methods

## Data Source and Sampling

The analysis draws on Asian Barometer Survey Wave 6, collected between late 2021 and fall 2022 across Cambodia, Thailand, and Vietnam. The ABS employs multi-stage stratified random sampling designed to produce nationally representative samples of adult citizens [@Chu2008-st], with all interviews conducted face-to-face using standardised questionnaires.

The analytic sample comprises `r format(total_n, big.mark=",")` respondents: Cambodia (n = `r format(cambodia$n, big.mark=",")`), Vietnam (n = `r format(vietnam$n, big.mark=",")`), and Thailand (n = `r format(thailand$n, big.mark=",")`). Response rates ranged from `r metadata$thailand_response`% in Thailand to `r metadata$cambodia_response`% in Cambodia. Listwise deletion on the core variables---infection, approval, information trust, democracy satisfaction, institutional trust---retains 3,241 respondents (88% of the full sample). Adding demographic and economic controls reduces N modestly due to item-level missingness, with variable completeness ranging from 93% to 99%. Model-specific sample sizes are reported throughout; results are substantively unchanged under multiple imputation (Online Appendix L.3).

Case selection follows the logic of structured comparison. All three countries are Southeast Asian, share broad cultural affinities, and faced the same pandemic---but they vary on the dimensions the theory requires: regime type (Cambodia and Vietnam are authoritarian; Thailand is a hybrid regime with greater information pluralism), pandemic outcomes (infection rates diverged dramatically), and information environments [@Freedom-House2025-gx; @Reporters-Without-Borders2023-ks]. This combination of shared background characteristics and theoretically relevant variation is what makes the comparison productive.

One feature of the data that might appear a liability actually strengthens the design. Data collection occurred at different times: Cambodia primarily in December 2021, Thailand in April--May 2022, Vietnam in August--September 2022. Vietnam was surveyed during the Omicron wave, when infections peaked. Respondents were evaluating a government *while actively experiencing* the pandemic's most acute phase---not recalling distant events from the safety of hindsight. That they evaluated the government so favourably under those conditions deepens the paradox.

## Pandemic Trajectories

The three countries' pandemic experiences diverged so dramatically that cross-national comparison becomes theoretically informative in ways that similar-trajectory cases could not be.

Vietnam's trajectory unfolded in two acts. Through early 2021, the country was a global exemplar---fewer than 3,000 cumulative cases despite sharing a 1,400-kilometre border with China. Aggressive contact tracing, centralised quarantine, and strict mobility controls earned international praise [@Pollack2020-xq; @La2020-jv]. Then the Delta variant arrived in Ho Chi Minh City in April 2021, and the containment regime collapsed. By the time Omicron swept through in early 2022, Vietnam had recorded over 11.5 million confirmed cases [@WHO2023-js]. The transformation was total: from regional success story to the country with Southeast Asia's highest per-capita infection rate, in roughly eighteen months. The survey captures citizens evaluating a government that had first protected them brilliantly and then presided over mass infection.

Cambodia presents the opposite trajectory. Fewer than 140,000 total cases by late 2022, achieved through the authoritarian state's capacity to enforce strict lockdowns---particularly the severe restrictions imposed on Phnom Penh after the February 2021 outbreak cluster---and centralised vaccine distribution that achieved high coverage. Cambodian respondents, surveyed in December 2021, were evaluating a government that had by the numbers protected them.

Thailand fell between the two epidemiologically but diverged politically. Approximately 4.4 million cumulative cases by survey time---moderate by regional standards---but the government's pandemic response became deeply entangled with sustained anti-government protests of 2020--2021. Demonstrators explicitly linked pandemic mismanagement to broader critiques of military-backed governance [@Kongkirati2018-go]. Unlike in Vietnam and Cambodia, Thai citizens had sustained access to critical media coverage and opposition voices that challenged official accounts of the crisis.

These trajectories create the comparative leverage the analysis requires. Vietnam and Cambodia share authoritarian governance and restricted information environments but experienced opposite pandemic outcomes. Thailand shares Southeast Asian cultural context but differs in information pluralism. If performance drives approval, Cambodia should outperform Vietnam. If the information environment matters, Vietnam and Cambodia should converge despite divergent outcomes, while Thailand should diverge despite moderate performance.

## Measurement

**Dependent variable.** Government pandemic handling approval, measured on a four-point scale: "How well or badly do you think the government handled the pandemic?" (1 = Very badly to 4 = Very well; M = `r round(var_stats$approval_mean, 2)`, SD = `r round(var_stats$approval_sd, 2)`).

**Primary independent variables.** Three dimensions of pandemic experience are measured. *Personal COVID-19 infection* is a binary indicator from the question "Have you or your family members previously contracted the Covid-19 virus?" The overall rate is `r round(var_stats$infection_rate, 1)`%, but this average obscures enormous cross-national variation: Cambodia `r round(cambodia$infection_pct, 1)`%, Thailand `r round(thailand$infection_pct, 1)`%, Vietnam `r round(vietnam$infection_pct, 1)`%. This measure warrants a candid note on attenuation. By combining personal and family infection into a single binary indicator, it compresses substantial variation in exposure intensity: a respondent whose cousin tested positive likely received a weaker political signal than one who was hospitalised personally. The binary coding further groups mild and severe cases together. If anything, this measurement imprecision biases estimates of infection's effect *toward* zero, making null findings less surprising but also making any detected effects conservative. The overall pattern of information trust dominance is unlikely to reverse under finer-grained measurement, but the null infection coefficients should be interpreted with this attenuation in mind. *Economic impact severity* captures material hardship on a four-point scale (M = `r round(var_stats$econ_mean, 2)`, SD = `r round(var_stats$econ_sd, 2)`). *Trust in government COVID-19 information* is the hypothesised mediator: "How much do you trust the Covid-19 related information provided by the government?" (M = `r round(var_stats$trust_mean, 2)`, SD = `r round(var_stats$trust_sd, 2)`).

**Controls.** The models include institutional trust (a nine-item index, α = `r round(alphas$institutional_trust, 3)`), democracy satisfaction, authoritarian acceptance (α = `r round(alphas$auth_acceptance, 3)`), and standard demographics---age, gender, education, urban residence. These controls are not incidental: institutional trust and authoritarian acceptance capture precisely the diffuse regime orientations that Baniamin [-@Baniamin2025-rx] identifies as driving trust paradoxes. If information trust is merely a proxy for generalized deference, these controls should absorb most of its predictive power.

The trust measure also warrants conceptual elaboration. The survey item does not specify *which dimension* of trust respondents are evaluating. Political trust is a multidimensional construct [@Hardin2002-pb; @Levi2000-am]: citizens may trust information because they perceive the source as *credible* (the information is accurate), because they attribute *competence* to the source (the government possesses relevant expertise), or because they feel *loyalty* toward the source (identification with or deference to the regime). These dimensions carry distinct theoretical implications---credibility-based trust implies a cognitive assessment of information quality; competence-based trust implies an inference about institutional capacity; loyalty-based trust implies an affective attachment that may be orthogonal to information quality altogether. The single survey item cannot distinguish among these pathways, and the relative contribution of each likely differs across cases. In Vietnam, where state media dominates the information landscape, high trust may reflect some mixture of perceived accuracy, attributed competence from the government's early pandemic success, and socialised deference to party authority. In Thailand, where citizens access competing narratives, trust may depend more heavily on credibility assessments, since loyalty alone is insufficient to sustain confidence in official information when alternative frames are readily available. The distinction matters for interpretation: if Vietnam's high trust is primarily loyalty-driven, these findings speak to authoritarian resilience; if primarily credibility-driven, they speak to effective crisis communication. I return to this ambiguity in the Discussion.

A potential concern warrants attention upfront. Trust in government COVID-19 information and approval of government pandemic handling might appear tautological---variants of the same underlying attitude. The correlation between the two in Vietnam is r = `r cor_trust_approval`, meaning they share `r round(cor_trust_approval^2 * 100, 1)`% of variance. Substantial, but `r 100 - round(cor_trust_approval^2 * 100, 1)`% remains unexplained. VIF scores range from `r vif_min` to `r vif_max`, well below conventional thresholds. The constructs are related---it would be surprising if they were not---but empirically distinguishable.

## Analytic Strategy

The analysis proceeds in four stages. Descriptive statistics document the paradox and establish the patterns requiring explanation. Bivariate OLS regressions test the raw associations between each COVID-19 impact---infection, economic hardship, information trust---and approval within each country. Multivariate OLS models then add controls, testing whether information trust survives the inclusion of institutional trust, democratic attitudes, and demographics. Finally, extensive robustness testing (reported in the Online Appendix) subjects the findings to alternative specifications. All models are estimated separately by country, using heteroskedasticity-consistent (HC3) standard errors throughout.

A note on inferential scope is warranted. All regression models are estimated within individual countries; the causal inferences offered here---such as they are with cross-sectional data---rest on within-country variation in trust, infection, and approval. I do not pool countries or estimate cross-national causal effects, because survey timing, question administration, and unmeasured country-level confounders make strict cross-national causal comparison inappropriate. The cross-national dimension of the analysis is descriptive and theoretical: the same within-country pattern (information trust dominates infection as a predictor of approval) recurs across three countries with very different pandemic trajectories and information environments, and I interpret this patterning as evidence that the mechanism generalises beyond any single case. The divergence between Vietnam/Cambodia (high trust, high approval despite different outcomes) and Thailand (low trust, low approval) is treated as a theoretically informative contrast, not as a controlled comparison.

# Results

## The Aggregate Paradox

Table 1 lays out the basic facts. Vietnam's infection rate (`r round(vietnam$infection_pct, 1)`%) exceeds Cambodia's (`r round(cambodia$infection_pct, 1)`%) by a factor of more than seven and substantially exceeds Thailand's (`r round(thailand$infection_pct, 1)`%). Vietnam's approval (`r round(vietnam$approval_pct, 1)`%) exceeds Cambodia's (`r round(cambodia$approval_pct, 1)`%) and dwarfs Thailand's (`r round(thailand$approval_pct, 1)`%). The correlation between infection and approval runs in the wrong direction.

<!-- Table 1 rendered below -->

## Infection and Economic Hardship: Null Effects

The natural expectation---that Vietnamese citizens who contracted COVID-19 should evaluate their government more harshly---finds no support.

Cambodia shows effectively zero association between infection and approval (`r format_coef(cors$cambodia_inf, h1_results$Cambodia$p)`). Vietnam, remarkably, shows a *positive* point estimate (`r format_coef(cors$vietnam_inf, h1_results$Vietnam$p)`)---infected respondents were, if anything, marginally *more* approving, though the coefficient does not differ significantly from zero. Only Thailand displays a statistically significant negative relationship (`r format_coef(cors$thailand_inf, h1_results$Thailand$p)`), and even there the substantive magnitude is trivial. Whether a citizen contracted COVID-19 was essentially irrelevant to how they evaluated pandemic governance.

Economic hardship tells a somewhat more differentiated story, though it too is dominated by information trust. Thailand shows the expected negative pattern (`r format_coef(cors$thailand_econ, h4b_results$Thailand$econ_p)`). Vietnam and Cambodia show no significant economic effects.

Information trust, by contrast, is enormous. Cambodia: `r format_coef(cors$cambodia_trust, h2a_results$Cambodia$p)`. Vietnam: `r format_coef(cors$vietnam_trust, h2a_results$Vietnam$p)`. Thailand: `r format_coef(cors$thailand_trust, h2a_results$Thailand$p)`. Trust alone explains 26--45% of the variance in approval across the three countries (Figure 1). These are bivariate results that could in principle reflect confounding with generalized regime support. The multivariate models address that possibility.

<!-- Figure 1 rendered below -->

## The Trust Coefficient Survives Controls

H1 predicts that information trust will predict pandemic approval even after accounting for institutional trust, democratic attitudes, and demographics. Table 2 reports the results.

The trust coefficient does not merely survive---it dominates every model. Across all countries and specifications, information trust is the strongest predictor (β = `r round(model_coef_ranges$trust_coef_min, 2)` to `r round(model_coef_ranges$trust_coef_max, 2)`, all p < 0.001). Infection status remains non-significant everywhere. The critical test for Baniamin's [-@Baniamin2025-rx] alternative: adding institutional trust---the most direct proxy for diffuse regime support and authoritarian cultural orientation---reduces the COVID information trust coefficient by 16--26%. Approximately three-quarters of the bivariate relationship is not attributable to generalized deference. The information-credibility mechanism operates through a pathway largely independent of the diffuse orientations Baniamin identifies.

In concrete terms: a respondent moving from the lowest to the highest level of information trust gains approximately 1.5 points on the four-point approval scale---nearly half the entire range. Infection status corresponds to less than 0.1 points. H1 is strongly supported.

<!-- Table 2 rendered below -->

## Conditional Accountability: The Thailand Test

H2 is the mechanism test. Thailand provides the clearest evidence.

In Thailand, where baseline approval (`r round(thailand$approval_pct, 1)`%) and baseline trust (`r round(thailand$trust_pct, 1)`%) are both low, there is a significant negative interaction between infection and information trust (see Online Appendix Table D). Among Thai citizens who distrust official COVID-19 information, personal infection is associated with substantially lower approval---precisely the accountability response that performance legitimacy theories predict. Among those who trust official information, infection shows no relationship with approval. The accountability mechanism works. It just requires that citizens distrust the government's narrative.

Thailand's deeply polarized political climate during this period---the 2020--2021 anti-government protests explicitly linked pandemic governance to broader critiques of military-backed rule---raises the question of whether this interaction merely captures partisanship by another name. The ABS does not include a direct measure of party identification or protest participation, but the models control for democracy satisfaction and authoritarian acceptance, which serve as proxies for the political orientations that structured Thailand's protest cleavage. The infection--trust interaction survives these controls. Whatever partisan dynamics are at work in Thai politics, the information-credibility mechanism is not reducible to them.

Vietnam and Cambodia show null interactions---which, far from undermining H2, reflects ceiling effects. With `r round(vietnam$trust_pct, 1)`% of Vietnamese respondents and `r round(cambodia$trust_pct, 1)`% of Cambodian respondents trusting official COVID-19 information, there is simply insufficient variation in trust to observe the low-trust accountability mechanism that Thailand reveals. The information-credibility heuristic has already done its work---not as a conditional moderator but as the population-wide baseline. Even within Vietnam's compressed trust distribution, the standardised trust coefficient (β = 0.31--0.41 across specifications) remains among the model's largest effects (Online Appendix F).

Taken together, the three-country pattern strongly supports H2. Where trust varies, infection affects approval in the direction performance theory predicts, but only among low-trust citizens. Where trust is near-universal, infection and approval are decoupled entirely.

## Robustness

The most consequential test asks whether the trust-approval relationship reflects a domain-specific cognitive mechanism or merely generalized regime loyalty. This amounts to a placebo test: if COVID information trust is functioning as a crisis-specific cognitive filter, it should predict pandemic approval far more strongly than it predicts satisfaction with governance domains unrelated to the pandemic. It does. COVID information trust predicts pandemic approval `r placebo_ratio_min`--`r placebo_ratio_max` times more strongly than it predicts democracy satisfaction or institutional trust---outcomes that a generalized halo effect or reverse-causation account (in which approval simply inflates all pro-government responses) would predict equally well. The domain specificity of the trust coefficient is difficult to reconcile with the most obvious endogeneity concern.

Ordered logistic models respecting the ordinal structure of the approval scale yield identical conclusions, as do robust and quantile regression specifications. Survey-weighted estimation produces substantively indistinguishable results. E-value sensitivity analysis indicates that an unmeasured confounder would need to be implausibly strong---nearly tripling the risk of both high trust and high approval, simultaneously and independently of all observed covariates---to explain away the findings. Full E-value tables and interpretation appear in the Online Appendix.

The social desirability evidence warrants separate attention because it bears on the most likely objection. `r round(speech_freedom_stats$pct_constrained, 1)`% of Vietnamese respondents openly disagreed that "people are free to speak without fear"---a response that is itself critical of the regime---and among those who perceive speech constraints, `r round(speech_freedom_stats$approval_among_constrained, 1)`% still approved of pandemic handling. Among those who feel free to speak: `r round(speech_freedom_stats$approval_among_free, 1)`%. Virtually identical. Vietnamese respondents also reported the region's highest infection rates without apparent reluctance. These off-diagonal responses---critical of speech freedoms yet approving of pandemic performance---sit uneasily with a uniform fear-based account.

A fair objection is that not all sensitive topics carry equal risk: criticising abstract speech freedoms may feel safer than challenging the government's signature pandemic achievement. List experiment research on political trust in authoritarian settings provides useful context. In China---arguably a harder case for social desirability than Vietnam---list experiments have documented overreporting of trust in the central government on the order of 10--25 percentage points, though sincere trust levels remain at or above majority thresholds even after correction [@Nicholson2023-kt; @Robinson2019-qm]. No comparable list experiment exists for pandemic-specific attitudes in Vietnam, but even assuming a desirability gap at the upper end of the range documented in these studies, the trust-approval association observed here (β = `r round(model_coef_ranges$trust_coef_min, 2)` to `r round(model_coef_ranges$trust_coef_max, 2)`) substantially exceeds what desirability bias alone could plausibly generate---since bias would need to differentially inflate the trust-approval relationship relative to the infection-approval relationship to account for the pattern.

# Discussion

## How Information Trust Works

The evidence establishes that information trust dominates personal experience as a predictor of pandemic approval across three countries, all specifications, and extensive robustness testing. H1 is strongly supported. In Thailand---the one country with sufficient variation in trust---infection is negatively associated with approval only among citizens who distrust official information (H2). And controlling for institutional trust barely attenuates the information trust coefficient, indicating the mechanism is not generalized deference wearing a domain-specific mask.

Cross-sectional data cannot prove the causal pathway, but the pattern is consistent with a specific account: trusted official information shapes how citizens *interpret* negative experiences, directing attributions away from government responsibility and toward external forces. In Vietnam and Cambodia, where official narratives face little competition and command near-universal trust, this interpretive filter appears to have decoupled personal experience from political evaluation entirely. Not because Vietnamese citizens are irrational or uninformed---many freely reported high infection rates and restricted speech freedoms---but because they appear to have accepted a framing in which mass infection was something that happened *despite* government effort rather than *because of* inadequate government response.

Pandian and McGonigle [-@Pandian2025-ww], studying Singapore's COVID-19 response, describe an analogous dynamic they term "political technocracy"---a legitimation strategy in which the ruling party positions itself as a competent long-term trustee and deploys systematic rhetorical techniques to shape perceptions of national problems. Their taxonomy of state rhetorical strategies---ontological, moral, and statistical---resonates with the Southeast Asian cases, though the specific mechanisms differ. Vietnam's Communist Party is not Singapore's PAP, and the Vietnamese information environment operates through different institutional channels. But the underlying logic converges: states can actively engineer the cognitive conditions under which citizens process crisis outcomes.

An important alternative reading of the Vietnamese and Cambodian cases deserves direct engagement: that citizens are not "filtering" personal experience through trusted information so much as echoing the only available narrative frame. In a mono-narrative information environment, the distinction between a cognitive heuristic and simple frame adoption may collapse. During Vietnam's Omicron wave, state media consistently framed mass infection as the inevitable cost of a globally circulating variant---emphasising vaccination progress, international comparisons, and economic reopening rather than domestic policy shortcomings [@Le2024-sv; @Truong2022-sp]. Citizens may have adopted this attribution not because they weighed competing interpretations and found the official account credible, but because no competing interpretation was readily available. The Thailand comparison provides partial traction on this concern: there, where alternative frames circulated freely, the information trust variable still predicted approval, but the performance-accountability mechanism operated among low-trust citizens. The cognitive filter, in other words, appears to function even in a pluralistic information environment---it is not exclusively a product of narrative monopoly, though monopoly conditions clearly amplify it.

This poses a fundamental challenge to performance legitimacy theory. If the state can maintain a trusted narrative, adverse outcomes need not translate into political consequences. The argument requires that trust in official information causally shapes evaluation---a direction the present design cannot establish. But it does not require that the trust itself is unearned. Vietnam's early pandemic response was genuinely impressive, and citizens may have accurately perceived competent management even as the situation deteriorated---a kind of reputational reservoir that the information environment helped sustain.

## Limitations

Three limitations deserve attention.

The most consequential is causal direction. Trust may shape approval, approval may shape trust, or both may reflect an underlying orientation toward the regime. The 16--26% attenuation when controlling for institutional trust suggests the relationship is not *entirely* driven by common cause, but some portion plausibly is. The placebo test reported in Section 4.5---showing that COVID information trust predicts pandemic approval `r placebo_ratio_min`--`r placebo_ratio_max` times more strongly than non-pandemic outcomes---provides the strongest available evidence against a pure reverse-causation account, since a generalized halo effect should predict all government evaluations with similar intensity. But the test cannot rule out a pandemic-specific form of reverse causation in which citizens who approve of pandemic handling retrospectively report higher information trust to maintain attitudinal consistency. Isolating the causal pathway requires experimental or panel designs that this study cannot provide.

The analysis measures *trust* in government information, not information control, media censorship, or propaganda effectiveness. High trust in Vietnam could reflect successful narrative management by an authoritarian state, genuinely effective crisis communication, cultural deference to authority [@Thayer2011-te; @Kerkvliet2015-lp], or some combination. The paper maintains a distinction throughout between the measured construct---individual-level trust in official COVID-19 information---and unmeasured contextual factors (information environments, media systems, censorship regimes) that may explain why trust levels differ so dramatically across countries. The elite signaling alternative discussed in Section 5.1---that citizens in mono-narrative environments may simply echo the only available frame rather than actively filtering experience---remains a plausible partial account that cross-sectional data cannot fully adjudicate.

Social desirability bias has been partially addressed but cannot be fully resolved. The off-diagonal patterns described in Section 4.5---Vietnamese respondents critical of speech freedoms yet approving of pandemic handling---constitute the strongest evidence against a uniform desirability account, and the magnitude of the trust-approval association substantially exceeds the social desirability gaps documented in list experiment research on authoritarian political trust [@Nicholson2023-kt; @Robinson2019-qm]. But subtler forms of topic-specific bias remain possible: respondents may feel safer criticising abstract speech freedoms than challenging the government's signature pandemic achievement. Even granting this, the *relationships* between information trust, infection, and approval---the core empirical contribution---should be less affected than the levels, since desirability bias would need to differentially inflate the trust-approval association relative to the infection-approval association to account for the pattern.

How do these findings relate to Baniamin's [-@Baniamin2025-rx] authoritarian cultural orientation framework? The two accounts are not in tension. Authoritarian cultural orientation likely helps explain *why* citizens in Vietnam and Cambodia trust official COVID-19 information so readily---societies with stronger hierarchical orientations may be predisposed to accept official narratives. But once trust is established, the analysis here suggests it actively shapes how citizens interpret their experiences, functioning as a mediating mechanism rather than simply inflating all assessments uniformly. Authoritarian cultural orientation identifies a cultural precondition; the information-credibility heuristic identifies a cognitive pathway through which that precondition shapes political evaluations during crises.

A related alternative warrants consideration: that Vietnam's high approval reflects citizens rewarding the government's successful 2020 containment rather than evaluating present conditions. If such legacy effects dominated, institutional trust---which captures accumulated regime legitimacy---should absorb most of the explanatory power once both variables enter the model. Table 2 shows the opposite: COVID-specific information trust remains the stronger predictor even controlling for institutional trust. Citizens appear to be filtering *current* experiences through *current* assessments of official information credibility, not coasting on stored goodwill.

## Implications

The most significant implication concerns the informational preconditions for democratic accountability. The canonical retrospective voting model assumes that citizens can observe outcomes and assign responsibility [@Fiorina1978-wf; @Healy2013-sy]. These results suggest the assumption fails when trusted official narratives successfully redirect blame. Thailand is instructive: the accountability mechanism functioned among citizens who distrusted official information, suggesting that the performance-legitimacy link is not culturally absent in Southeast Asia but *informationally contingent*. Democratic accountability may depend not only on electoral institutions but on information environments that permit citizens to access---and have confidence in---alternative attributional frames.

For theories of authoritarian resilience, these findings provide individual-level survey evidence complementing macro-level arguments about informational autocracy [@Guriev2019-xj; @Egorov2024-gq]. Guriev and Treisman theorise that modern autocrats survive by convincing the public of their competence. The present analysis shows what this looks like at the level of individual attitudes: when citizens trust official crisis information, objective performance indicators become politically irrelevant. The survival mechanism is not repression or ideology but narrative credibility.

For rally-around-the-flag theory, health crises do not automatically generate support [@Hegewald2024-ii; @Kritzinger2021-im; @Zoch2024-kf]. The rally mechanism appears conditional on information environment---where citizens trust official narratives, support emerges regardless of outcomes; where they distrust them, the expected accountability pattern operates. Rally effects, if they contribute at all in these cases, are sustained by information trust rather than by the existence of an external threat alone.

The patterns documented here are troubling for theories of democratic accountability, but they do not demonstrate an "authoritarian advantage" in crisis management. Thailand's low approval could indicate democratic accountability functioning exactly as intended---citizens holding a government accountable for perceived shortcomings. The troubling finding is not that authoritarianism works better but that information trust can substitute for actual performance as a basis for political support, and that the conditions enabling this substitution are unevenly distributed across regime types.

# Conclusion

The Vietnam Paradox is real, robust, and resistant to the obvious alternative explanations. Despite recording Southeast Asia's highest COVID-19 infection rate, Vietnam maintained the region's highest approval for government pandemic handling. Personal infection predicted nothing. Information trust predicted nearly everything.

Trust in government COVID-19 information dominates personal experience by an order of magnitude as a predictor of pandemic approval---β = `r round(cor_ranges$trust_min, 2)` to `r round(cor_ranges$trust_max, 2)`, explaining 26--45% of variance across three countries. The Thailand interaction results transform these findings from a regional anomaly into evidence for a general theory: the performance-legitimacy link functions normally when citizens distrust official narratives, and breaks down when they trust them. Conditional, not absent.

The information-credibility heuristic proposed here---in which citizens who trust official narratives attribute negative experiences to forces beyond government control rather than to governmental shortcomings---extends scholarship on authoritarian trust paradoxes [@Baniamin2025-rx] by identifying a domain-specific cognitive pathway through which trust operates during crises, and complements macro-level theories of informational autocracy [@Guriev2019-xj; @Egorov2024-gq] with micro-level evidence from individual survey responses.

Whether the trust that Vietnamese citizens place in their government's COVID-19 information was earned or manufactured---whether it reflects genuinely competent communication or successful narrative control despite poor outcomes---remains an open question that awaits longitudinal or experimental designs.

But the evidence does establish, robustly, that if governments can maintain trusted narratives during crises, objective outcomes may become politically irrelevant. For scholars who ground democratic accountability in citizens' capacity to evaluate governmental performance, that is a deeply unsettling finding.

# References {.unnumbered}

::: {#refs}
:::

{{< pagebreak >}}

# Tables and Figures {.unnumbered}

```{r}
#| label: tbl-paradox
#| tbl-cap: "The Vietnam Paradox---COVID-19 Infection Rates and Government Approval by Country"
#| echo: false

if (!is.null(paradox_summary)) {
  tbl_data <- paradox_summary %>%
    select(country_name, n, covid_infection_rate, govt_approval_rate, trust_covid_rate)

  if (is_docx) {
    # Flextable for DOCX with explicit column widths
    tbl_data %>%
      rename(
        Country = country_name,
        N = n,
        `Infection Rate (%)` = covid_infection_rate,
        `Approval Rate (%)` = govt_approval_rate,
        `Trust COVID Info (%)` = trust_covid_rate
      ) %>%
      flextable::flextable() %>%
      flextable::width(j = 1, width = 1.2) %>%
      flextable::width(j = 2, width = 0.6) %>%
      flextable::width(j = 3:5, width = 1.3) %>%
      flextable::colformat_double(digits = 1) %>%
      flextable::autofit()
  } else {
    # gt for HTML/PDF
    tbl_data %>%
      gt() %>%
      cols_label(
        country_name = "Country",
        n = "N",
        covid_infection_rate = "Infection Rate (%)",
        govt_approval_rate = "Approval Rate (%)",
        trust_covid_rate = "Trust COVID Info (%)"
      ) %>%
      cols_width(
        country_name ~ px(100),
        n ~ px(60),
        covid_infection_rate ~ px(120),
        govt_approval_rate ~ px(120),
        trust_covid_rate ~ px(130)
      ) %>%
      fmt_number(columns = where(is.numeric), decimals = 1) %>%
      tab_header(
        title = "The Vietnam Paradox"
      ) %>%
      tab_source_note("Source: Asian Barometer Wave 6 (2021-2022). Infection = % reporting personal/family COVID infection. Approval = % rating handling as 'well'/'very well'. Trust = % reporting 'quite a lot'/'great deal' of trust.") %>%
      tab_options(
        table.font.size = px(10),
        heading.title.font.size = px(12),
        source_notes.font.size = px(9),
        data_row.padding = px(3),
        table.width = pct(100)
      )
  }
} else {
  cat("Data not available.\n")
}
```

{{< pagebreak >}}

```{r}
#| label: fig-coefplot
#| fig-cap: "Coefficient Plot with 95% Confidence Intervals---Bivariate Associations Between COVID-19 Impacts and Government Approval"
#| fig-width: 9
#| fig-height: 7
#| echo: false

# Build coefficient plot data from stored results (with CIs)
# All three use BIVARIATE models for apples-to-apples comparison
if (!is.null(h1_results) && !is.null(h2a_results) && !is.null(h4a_results)) {

  coef_data <- bind_rows(
    # Infection coefficients (bivariate)
    tibble(
      Country = names(h1_results),
      Variable = "Personal Infection",
      estimate = sapply(h1_results, function(x) x$coef),
      ci_lower = sapply(h1_results, function(x) x$ci_lower),
      ci_upper = sapply(h1_results, function(x) x$ci_upper),
      significant = sapply(h1_results, function(x) x$p < 0.05)
    ),
    # Trust coefficients (bivariate)
    tibble(
      Country = names(h2a_results),
      Variable = "Trust in COVID Info",
      estimate = sapply(h2a_results, function(x) x$coef),
      ci_lower = sapply(h2a_results, function(x) x$ci_lower),
      ci_upper = sapply(h2a_results, function(x) x$ci_upper),
      significant = sapply(h2a_results, function(x) x$p < 0.05)
    ),
    # Economic coefficients (bivariate - from h4a_results)
    tibble(
      Country = names(h4a_results),
      Variable = "Economic Hardship",
      estimate = sapply(h4a_results, function(x) x$coef),
      ci_lower = sapply(h4a_results, function(x) x$ci_lower),
      ci_upper = sapply(h4a_results, function(x) x$ci_upper),
      significant = sapply(h4a_results, function(x) x$p < 0.05)
    )
  ) %>%
    mutate(
      Country = factor(Country, levels = c("Cambodia", "Vietnam", "Thailand")),
      Variable = factor(Variable, levels = c("Personal Infection", "Economic Hardship", "Trust in COVID Info"))
    )

  # Cleveland dot plot with CIs
  ggplot(coef_data, aes(x = estimate, y = Variable, color = Country, shape = significant)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", linewidth = 0.5) +
    geom_errorbarh(
      aes(xmin = ci_lower, xmax = ci_upper),
      height = 0.15,
      linewidth = 0.7,
      position = position_dodge(width = 0.6)
    ) +
    geom_point(
      size = 3.5,
      position = position_dodge(width = 0.6)
    ) +
    scale_color_manual(
      values = c("Cambodia" = "#66C2A5", "Vietnam" = "#8DA0CB", "Thailand" = "#FC8D62"),
      name = "Country"
    ) +
    scale_shape_manual(
      values = c("TRUE" = 16, "FALSE" = 1),  # Filled vs hollow
      name = "p < 0.05",
      labels = c("TRUE" = "Yes", "FALSE" = "No")
    ) +
    labs(
      title = "Trust Dominates: Effect Sizes with Uncertainty",
      subtitle = "Standardized regression coefficients with 95% confidence intervals",
      x = expression(paste("Standardized Coefficient (", beta, ")")),
      y = NULL,
      caption = "Note: Filled points = p < 0.05. All coefficients from bivariate regressions."
    ) +
    scale_x_continuous(limits = c(-0.3, 0.85), breaks = seq(-0.2, 0.8, 0.2)) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "bottom",
      legend.box = "horizontal",
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 11, color = "gray30"),
      plot.caption = element_text(size = 9, hjust = 0),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      axis.text.y = element_text(size = 11, face = "bold")
    ) +
    guides(
      color = guide_legend(order = 1),
      shape = guide_legend(order = 2)
    )
} else {
  cat("Data not available. Run analysis scripts first.\n")
}
```

{{< pagebreak >}}

```{r}
#| label: tbl-multivariate
#| tbl-cap: "Multivariate OLS Regression---Determinants of Government Pandemic Approval"
#| echo: false

if (!is.null(m1_coefficients) && !is.null(m2_coefficients)) {

  term_labels <- c(
    "covid_contracted" = "COVID Infected",
    "covid_trust_info" = "Trust COVID Info",
    "institutional_trust_index" = "Institutional Trust",
    "dem_satisfaction" = "Democracy Satisfaction",
    "auth_acceptance" = "Authoritarian Acceptance",
    "covid_impact_severity" = "COVID Economic Impact",
    "income_quintile" = "Income Quintile",
    "econ_anxiety" = "Economic Anxiety",
    "age" = "Age",
    "genderMale" = "Male",
    "educ_levelSecondary" = "Education (Secondary)",
    "educ_levelUniversity/Tertiary" = "Education (University)",
    "urban" = "Urban"
  )

  format_coef <- function(estimate, std.error, p.value) {
    stars <- case_when(p.value < 0.001 ~ "***", p.value < 0.01 ~ "**", p.value < 0.05 ~ "*", TRUE ~ "")
    sprintf("%.3f%s (%.3f)", estimate, stars, std.error)
  }

  m1_processed <- m1_coefficients %>%
    filter(term != "(Intercept)") %>%
    mutate(Variable = term_labels[term], coef_display = format_coef(estimate, std.error, p.value)) %>%
    select(Country, Variable, coef_display) %>%
    pivot_wider(names_from = Country, values_from = coef_display)

  m2_processed <- m2_coefficients %>%
    filter(term != "(Intercept)") %>%
    mutate(Variable = ifelse(term %in% names(term_labels), term_labels[term], term),
           coef_display = format_coef(estimate, std.error, p.value)) %>%
    select(Country, Variable, coef_display) %>%
    pivot_wider(names_from = Country, values_from = coef_display)

  all_vars <- m2_processed$Variable

  m1_expanded <- tibble(Variable = all_vars) %>%
    left_join(m1_processed, by = "Variable") %>%
    mutate(across(where(is.character) & !matches("Variable"), ~replace_na(., "---")))

  summary_df <- tibble(
    Variable = all_vars,
    `Cambodia (1)` = m1_expanded$Cambodia, `Cambodia (2)` = m2_processed$Cambodia,
    `Vietnam (1)` = m1_expanded$Vietnam, `Vietnam (2)` = m2_processed$Vietnam,
    `Thailand (1)` = m1_expanded$Thailand, `Thailand (2)` = m2_processed$Thailand
  )

  fit_rows <- tibble(
    Variable = c("Observations", "R²"),
    `Cambodia (1)` = c(as.character(m1_fit$N[m1_fit$Country == "Cambodia"]), sprintf("%.3f", m1_fit$R_squared[m1_fit$Country == "Cambodia"])),
    `Cambodia (2)` = c(as.character(m2_fit$N[m2_fit$Country == "Cambodia"]), sprintf("%.3f", m2_fit$R_squared[m2_fit$Country == "Cambodia"])),
    `Vietnam (1)` = c(as.character(m1_fit$N[m1_fit$Country == "Vietnam"]), sprintf("%.3f", m1_fit$R_squared[m1_fit$Country == "Vietnam"])),
    `Vietnam (2)` = c(as.character(m2_fit$N[m2_fit$Country == "Vietnam"]), sprintf("%.3f", m2_fit$R_squared[m2_fit$Country == "Vietnam"])),
    `Thailand (1)` = c(as.character(m1_fit$N[m1_fit$Country == "Thailand"]), sprintf("%.3f", m1_fit$R_squared[m1_fit$Country == "Thailand"])),
    `Thailand (2)` = c(as.character(m2_fit$N[m2_fit$Country == "Thailand"]), sprintf("%.3f", m2_fit$R_squared[m2_fit$Country == "Thailand"]))
  )

  summary_df <- bind_rows(summary_df, fit_rows)

  if (is_docx) {
    # Flextable for DOCX with compact formatting
    # Total width: 1.3 + (6 × 0.85) = 6.4 inches (fits in 6.5" usable width)
    summary_df %>%
      flextable::flextable() %>%
      flextable::fontsize(size = 8, part = "all") %>%
      flextable::fontsize(size = 7, part = "header") %>%
      flextable::padding(padding = 1, part = "all") %>%
      flextable::width(j = 1, width = 1.3) %>%
      flextable::width(j = 2:7, width = 0.85) %>%
      flextable::align(align = "center", j = 2:7, part = "all") %>%
      flextable::align(align = "left", j = 1, part = "all")
  } else {
    # gt for HTML/PDF with narrower columns
    summary_df %>%
      gt() %>%
      tab_header(title = "Multivariate OLS Regression Results", subtitle = "DV: Government Pandemic Handling (1--4)") %>%
      tab_spanner(label = "Cambodia", columns = c(`Cambodia (1)`, `Cambodia (2)`)) %>%
      tab_spanner(label = "Vietnam", columns = c(`Vietnam (1)`, `Vietnam (2)`)) %>%
      tab_spanner(label = "Thailand", columns = c(`Thailand (1)`, `Thailand (2)`)) %>%
      cols_width(
        Variable ~ px(140),
        everything() ~ px(85)
      ) %>%
      tab_source_note("*** p < 0.001, ** p < 0.01, * p < 0.05. SE in parentheses. (1) Core; (2) Full with controls.") %>%
      tab_options(
        table.font.size = px(8),
        column_labels.font.size = px(8),
        heading.title.font.size = px(10),
        heading.subtitle.font.size = px(9),
        source_notes.font.size = px(7),
        data_row.padding = px(1),
        table.width = pct(100)
      )
  }
} else {
  cat("Data not available.\n")
}
```
