---
title: "EconDev-AuthPref: Paper Title"
subtitle: "Subtitle (Optional)"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
    email: jeffrey.stark@yonsei.ac.kr
    orcid: 0000-0000-0000-0000
date: today
date-format: long
abstract: |
  This is the abstract. It should concisely summarize the main question, 
  methodology, findings, and contribution of the paper. Typically 150-250 words.
keywords: [social science, survey, methodology]
thanks: |
  Acknowledgments and funding information go here.
bibliography: refs.bib
csl: ../../apa.csl
number-sections: true
---

# Introduction {#sec-introduction}

## Research Question

State your research question clearly and concisely.

## Motivation

Why is this question important? What gap in the literature does it address?

## Contribution

What does this paper contribute to the field?

# Literature Review {#sec-literature}

## Theoretical Framework

Discuss the theoretical foundations of your research.

## Previous Research

Review relevant empirical studies and their findings.

## Hypotheses

State your hypotheses explicitly:

**H1:** First hypothesis statement.

**H2:** Second hypothesis statement.

# Data and Methods {#sec-methods}

## Data Source

```{r}
#| label: load-data
#| include: false

# Load necessary R packages
library(tidyverse)
library(haven)
library(modelsummary)

# Source utility functions
source("../../src/r/utils/load_data.R")
source("../../src/r/utils/descriptive_stats.R")

# Load the processed data
data <- read_rds("../../data/processed/survey_data.rds")
```

Describe your data source, sampling strategy, and sample characteristics.

## Variables

### Dependent Variable

Describe your outcome variable(s).

### Independent Variables

Describe your key explanatory variables.

### Control Variables

List and describe control variables.

## Analytical Strategy

Describe your statistical approach and model specification.

# Results {#sec-results}

## Descriptive Statistics

@tbl-descriptives presents descriptive statistics for all variables in the analysis.

```{r}
#| label: tbl-descriptives
#| tbl-cap: "Descriptive Statistics"
#| echo: false

# Generate descriptive statistics table
datasummary_skim(
  data,
  output = "tables/descriptive_stats.html"
)
```

## Main Results

@tbl-main-results presents the main regression results.

```{r}
#| label: tbl-main-results
#| tbl-cap: "Main Regression Results"
#| echo: false

# Run main analysis
model1 <- lm(outcome ~ predictor1 + predictor2, data = data)
model2 <- lm(outcome ~ predictor1 + predictor2 + control1 + control2, data = data)

# Create regression table
modelsummary(
  list("Model 1" = model1, "Model 2" = model2),
  stars = TRUE,
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  output = "tables/main_results.html"
)
```

@fig-main-effect visualizes the main relationship of interest.

```{python}
#| label: fig-main-effect
#| fig-cap: "Main Effect Visualization"
#| echo: false

import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns

# Load Python data if needed
# data = pd.read_csv("../../data/processed/survey_data.csv")

# Create visualization
# plt.figure(figsize=(10, 6))
# sns.scatterplot(data=data, x='predictor', y='outcome')
# plt.savefig('figures/main_effect.png', dpi=300, bbox_inches='tight')
# plt.show()
```

## Robustness Checks

Describe robustness checks and sensitivity analyses.

# Discussion {#sec-discussion}

## Interpretation of Findings

Interpret your results in the context of your hypotheses and theoretical framework.

## Implications

Discuss theoretical and practical implications.

## Limitations

Acknowledge limitations of your study.

# Conclusion {#sec-conclusion}

Summarize main findings and contributions. Suggest directions for future research.

# References {.unnumbered}

::: {#refs}
:::

# Appendix {.appendix}

See [appendix](appendix.qmd) for additional materials.

