---
title: "Online Appendix: What Does Democracy Mean to Losers?"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
date: today
format:
  pdf:
    documentclass: article
    classoption: [letterpaper, 12pt]
    geometry: margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    linkcolor: blue
    urlcolor: blue
    citecolor: blue
    keep-tex: true
    include-in-header:
      text: |
        \renewcommand{\thesection}{\Alph{section}}
        \renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
        \renewcommand{\thetable}{\Alph{section}\arabic{table}}
        \renewcommand{\thefigure}{\Alph{section}\arabic{figure}}
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
bibliography: Winners-losers.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

here::i_am("papers/meaning-of-democracy/manuscript/online_appendix.qmd")

# Results directory (multinomial logit analysis outputs)
results_dir <- here("papers", "meaning-of-democracy", "analysis", "revised", "results")
data_dir <- here("papers", "meaning-of-democracy", "analysis", "revised", "data")

# Helper function to safely load files
safe_read <- function(path, type = "rds") {
  if (!file.exists(path)) {
    warning(paste("File not found:", path))
    return(NULL)
  }
  if (type == "rds") readRDS(path)
  else if (type == "csv") read_csv(path, show_col_types = FALSE)
  else stop("Unknown type")
}

# Load data files
item_meta <- safe_read(file.path(data_dir, "item_metadata.rds"))
w2_data <- safe_read(file.path(data_dir, "w2_baseline.rds"))
w346_data <- safe_read(file.path(data_dir, "w346_main.rds"))

# Load results files
cw_ames <- safe_read(file.path(results_dir, "country_wave_item_ames.csv"), "csv")
cw_gap <- safe_read(file.path(results_dir, "country_wave_proc_sub_gap.csv"), "csv")
pooled_results <- safe_read(file.path(results_dir, "loser_effects_all_items.csv"), "csv")
mlogit_results <- safe_read(file.path(results_dir, "mlogit_results.rds"))
```

# Survey Items and Coding {#sec-items}

This section documents the survey items used to measure conceptions of democracy across the Asian Barometer Survey waves.

## Item Reference Table

```{r tbl-items}
#| tbl-cap: "Survey Items by Set and Type"

if (!is.null(item_meta)) {
  item_meta %>%
    select(set, value, item_label, item_type, item_subtype) %>%
    rename(
      Set = set,
      Value = value,
      `Item Label` = item_label,
      `Item Type` = item_type,
      Subtype = item_subtype
    ) %>%
    kable(booktabs = TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped"),
                  font_size = 10) %>%
    pack_rows("Wave 2", 1, 4) %>%
    pack_rows("Set 1 (W3/W4/W6)", 5, 8) %>%
    pack_rows("Set 2 (W3/W4/W6)", 9, 12) %>%
    pack_rows("Set 3 (W3/W4/W6)", 13, 16) %>%
    pack_rows("Set 4 (W3/W4/W6)", 17, 20)
} else {
  cat("**Warning:** Item metadata file not found.")
}
```

**Note:** Wave 2 employed a single forced-choice item with four response options, while Waves 3, 4, and 6 used four separate item sets (Sets 1--4), each presenting four response options. The W2 instrument differs from the later waves and is analyzed separately.

# Sample Composition {#sec-sample}

## Sample Size by Country and Wave

```{r tbl-sample}
#| tbl-cap: "Sample Composition by Country and Wave"

if (!is.null(w2_data) && !is.null(w346_data)) {
  # Combine data
  w2_summary <- w2_data %>%
    mutate(wave_label = "W2") %>%
    group_by(country_name, wave_label) %>%
    summarise(
      N = n(),
      N_winner = sum(loser == 0),
      N_loser = sum(loser == 1),
      pct_loser = mean(loser) * 100,
      .groups = "drop"
    )

  w346_summary <- w346_data %>%
    group_by(country_name, wave_label) %>%
    summarise(
      N = n(),
      N_winner = sum(loser == 0),
      N_loser = sum(loser == 1),
      pct_loser = mean(loser) * 100,
      .groups = "drop"
    )

  sample_summary <- bind_rows(w2_summary, w346_summary) %>%
    arrange(country_name, wave_label)

  # Wide format for display
  sample_wide <- sample_summary %>%
    select(country_name, wave_label, N) %>%
    pivot_wider(names_from = wave_label, values_from = N, values_fill = 0) %>%
    mutate(Total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
    arrange(country_name)

  # Add column totals
  col_totals <- sample_wide %>%
    summarise(across(where(is.numeric), sum)) %>%
    mutate(country_name = "**Total**")

  sample_wide <- bind_rows(sample_wide, col_totals)

  sample_wide %>%
    rename(Country = country_name) %>%
    kable(booktabs = TRUE, format.args = list(big.mark = ",")) %>%
    kable_styling(latex_options = c("hold_position", "striped"))
} else {
  cat("**Warning:** Data files not found. Cannot generate sample table.")
}
```

## Response Distributions by Winner/Loser Status

```{r tbl-response-dist}
#| tbl-cap: "Response Distributions by Electoral Status (Pooled)"

if (!is.null(w346_data)) {
  # Function to compute cross-tabs for a set
  compute_crosstab <- function(data, set_num) {
    choice_var <- paste0("set", set_num, "_choice")
    valid_var <- paste0("set", set_num, "_valid")

    data %>%
      filter(.data[[valid_var]]) %>%
      mutate(choice = .data[[choice_var]]) %>%
      group_by(loser, choice) %>%
      summarise(n = n(), .groups = "drop") %>%
      group_by(loser) %>%
      mutate(pct = n / sum(n) * 100) %>%
      ungroup() %>%
      select(loser, choice, pct) %>%
      pivot_wider(names_from = loser, values_from = pct, names_prefix = "loser_") %>%
      mutate(
        Winner = loser_0,
        Loser = loser_1,
        Diff = loser_1 - loser_0,
        Set = paste0("Set ", set_num)
      ) %>%
      select(Set, Item = choice, Winner, Loser, Diff)
  }

  # Compute for all sets
  all_crosstabs <- map_dfr(1:4, ~compute_crosstab(w346_data, .x))

  all_crosstabs %>%
    mutate(
      Winner = sprintf("%.1f%%", Winner),
      Loser = sprintf("%.1f%%", Loser),
      Diff = sprintf("%+.1f", Diff)
    ) %>%
    kable(booktabs = TRUE, align = c("l", "l", "r", "r", "r")) %>%
    kable_styling(latex_options = c("hold_position", "striped"),
                  font_size = 9) %>%
    pack_rows("Set 1", 1, 4) %>%
    pack_rows("Set 2", 5, 8) %>%
    pack_rows("Set 3", 9, 12) %>%
    pack_rows("Set 4", 13, 16)
} else {
  cat("**Warning:** W346 data file not found.")
}
```

**Note:** Diff = Loser - Winner. Positive values indicate losers are more likely to choose that item.

## Non-Voter Rates by Country and Wave

The main analysis restricts attention to respondents who reported voting in the most recent national election. Table \@ref(tab:tbl-nonvoter-rates) reports the proportion of non-voters among respondents with valid democracy conception responses.

```{r tbl-nonvoter-rates}
#| tbl-cap: "Non-Voter Rates by Country and Wave"

# Load master data
abs <- readRDS(here("data", "processed", "abs_econdev_authpref.rds"))

# Country name mapping (11 analysis countries)
country_names <- c(
  `1` = "Japan", `3` = "South Korea", `5` = "Mongolia",
  `6` = "Philippines", `7` = "Taiwan", `8` = "Thailand",
  `9` = "Indonesia", `12` = "Cambodia", `13` = "Malaysia",
  `14` = "Myanmar", `15` = "Australia"
)

nonvoter_tab <- abs %>%
  filter(
    !is.na(voted_last_election),
    !country %in% c(2, 4, 10, 11)   # Exclude HK, China, Singapore, Vietnam
  ) %>%
  mutate(country_name = country_names[as.character(country)]) %>%
  filter(!is.na(country_name)) %>%
  group_by(country_name, wave) %>%
  summarise(
    N_total   = n(),
    N_voter   = sum(voted_last_election == 1, na.rm = TRUE),
    N_nonvote = sum(voted_last_election == 0, na.rm = TRUE),
    Pct_nonvote = round(100 * N_nonvote / N_total, 1),
    .groups = "drop"
  )

# Pivot to wide format: one row per country, columns per wave
tab_wide <- nonvoter_tab %>%
  select(country_name, wave, N_total, Pct_nonvote) %>%
  pivot_wider(
    names_from = wave,
    values_from = c(N_total, Pct_nonvote),
    names_sep = "_W"
  ) %>%
  arrange(country_name)

# Determine which waves are present
waves_present <- sort(unique(nonvoter_tab$wave))
n_waves <- length(waves_present)

# Build column names dynamically
col_labels <- c("Country")
for (w in waves_present) {
  col_labels <- c(col_labels, "N", "% NV")
}

# Build header grouping
header_groups <- c(" " = 1)
for (w in waves_present) {
  header_groups <- c(header_groups, setNames(2, paste0("Wave ", w)))
}

# Reorder columns to interleave N and % for each wave
col_order <- c("country_name")
for (w in waves_present) {
  col_order <- c(col_order, paste0("N_total_W", w), paste0("Pct_nonvote_W", w))
}
tab_wide <- tab_wide %>% select(any_of(col_order))

kbl(tab_wide,
    booktabs = TRUE,
    col.names = col_labels,
    align = c("l", rep("r", 2 * n_waves))) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"),
                font_size = 9) %>%
  add_header_above(header_groups)
```

**Note:** N = total respondents with valid vote status; % NV = percentage who reported not voting. Australia's low non-voter rate reflects compulsory voting.

### Sample Selection Considerations

Because the analysis conditions on reported vote choice to assign electoral status, non-voters are excluded by design. In most country-waves, non-voter rates fall between 8 and 25 percent, consistent with the moderately high turnout typical of Asian democracies. Two patterns merit attention.

First, Myanmar exhibits the highest non-voter rate among the analysis countries, reaching 26 percent in Wave 4. This likely reflects the challenging political environment during the transitional period following decades of military rule, where many citizens may have been uncertain about electoral participation or faced practical barriers to voting.

Second, Australia and Thailand represent near-compulsory or high-turnout contexts where non-voter rates remain below 5 percent across waves. In these cases, selection into voting is effectively eliminated as a confound, and the persistence of the loser effect in Thailand---where the gap grew dramatically despite minimal sample selection---provides strong evidence that the findings are not driven by differential turnout between winners and losers.

The exclusion of non-voters means the analysis characterizes democratic conceptions among the politically engaged electorate rather than the population at large. However, the consistency of the loser effect across countries with vastly different non-voter rates (from 3.7 percent in Australia to 26 percent in Myanmar) suggests the pattern is robust to variation in the scope of sample selection.

# Full Country-Wave Results {#sec-country-wave}

## Item-Level Average Marginal Effects

```{r tbl-cw-ames-sets12}
#| tbl-cap: "Country-Wave AMEs: Sets 1-2"

if (!is.null(cw_ames)) {
  # Create short item labels for column headers
  item_labels <- c(
    "Free elections" = "Elections",
    "Party competition" = "Parties",
    "Basic necessities" = "Necessities",
    "Reduce inequality" = "Inequality",
    "Gov explain" = "Explain",
    "Free speech" = "Speech",
    "Jobs for all" = "Jobs",
    "Narrow income gap" = "Income gap",
    "Free media" = "Media",
    "Fair elections" = "Fair elect",
    "Low unemployment" = "Unemp",
    "Welfare for poor" = "Welfare",
    "Check government" = "Check gov",
    "Minority rights" = "Minority",
    "Equal income" = "Equal inc",
    "No corruption" = "No corrupt"
  )

  # Format for display with short labels
  cw_display <- cw_ames %>%
    filter(set %in% c("Set1", "Set2")) %>%
    mutate(
      sig = replace_na(sig, ""),
      ame_fmt = sprintf("%+.1f%s", ame * 100, sig),
      item_short = if_else(item %in% names(item_labels),
                           item_labels[item],
                           substr(item, 1, 10))
    ) %>%
    select(country_name, wave_label, set, item_short, ame_fmt) %>%
    pivot_wider(
      names_from = c(set, item_short),
      values_from = ame_fmt,
      names_sep = "_"
    ) %>%
    arrange(country_name, wave_label)

  # Get column names for header grouping
  col_names <- names(cw_display)
  set1_cols <- sum(grepl("^Set1_", col_names))
  set2_cols <- sum(grepl("^Set2_", col_names))

  # Clean column names (remove Set prefix)
  names(cw_display) <- gsub("^Set[12]_", "", names(cw_display))

  cw_display %>%
    rename(Country = country_name, Wave = wave_label) %>%
    kable(booktabs = TRUE, longtable = TRUE, align = c("l", "c", rep("r", set1_cols + set2_cols))) %>%
    kable_styling(latex_options = c("hold_position", "repeat_header", "scale_down"),
                  font_size = 8) %>%
    add_header_above(c(" " = 2, "Set 1" = set1_cols, "Set 2" = set2_cols)) %>%
    column_spec(1, width = "1.8cm") %>%
    column_spec(2, width = "0.6cm") %>%
    row_spec(seq(3, nrow(cw_display), by = 3), extra_css = "border-bottom: 1px solid #ccc;",
             extra_latex_after = "\\midrule")
} else {
  cat("**Warning:** Country-wave AME file not found.")
}
```

```{r tbl-cw-ames-sets34}
#| tbl-cap: "Country-Wave AMEs: Sets 3-4"

if (!is.null(cw_ames)) {
  # Create short item labels for column headers
  item_labels <- c(
    "Free elections" = "Elections",
    "Party competition" = "Parties",
    "Basic necessities" = "Necessities",
    "Reduce inequality" = "Inequality",
    "Gov explain" = "Explain",
    "Free speech" = "Speech",
    "Jobs for all" = "Jobs",
    "Narrow income gap" = "Income gap",
    "Free media" = "Media",
    "Fair elections" = "Fair elect",
    "Low unemployment" = "Unemp",
    "Welfare for poor" = "Welfare",
    "Check government" = "Check gov",
    "Minority rights" = "Minority",
    "Equal income" = "Equal inc",
    "No corruption" = "No corrupt"
  )

  # Format for display with short labels
  cw_display <- cw_ames %>%
    filter(set %in% c("Set3", "Set4")) %>%
    mutate(
      sig = replace_na(sig, ""),
      ame_fmt = sprintf("%+.1f%s", ame * 100, sig),
      item_short = if_else(item %in% names(item_labels),
                           item_labels[item],
                           substr(item, 1, 10))
    ) %>%
    select(country_name, wave_label, set, item_short, ame_fmt) %>%
    pivot_wider(
      names_from = c(set, item_short),
      values_from = ame_fmt,
      names_sep = "_"
    ) %>%
    arrange(country_name, wave_label)

  # Get column names for header grouping
  col_names <- names(cw_display)
  set3_cols <- sum(grepl("^Set3_", col_names))
  set4_cols <- sum(grepl("^Set4_", col_names))

  # Clean column names (remove Set prefix)
  names(cw_display) <- gsub("^Set[34]_", "", names(cw_display))

  cw_display %>%
    rename(Country = country_name, Wave = wave_label) %>%
    kable(booktabs = TRUE, longtable = TRUE, align = c("l", "c", rep("r", set3_cols + set4_cols))) %>%
    kable_styling(latex_options = c("hold_position", "repeat_header", "scale_down"),
                  font_size = 8) %>%
    add_header_above(c(" " = 2, "Set 3" = set3_cols, "Set 4" = set4_cols)) %>%
    column_spec(1, width = "1.8cm") %>%
    column_spec(2, width = "0.6cm") %>%
    row_spec(seq(3, nrow(cw_display), by = 3), extra_css = "border-bottom: 1px solid #ccc;",
             extra_latex_after = "\\midrule")
} else {
  cat("**Warning:** Country-wave AME file not found.")
}
```

**Note:** Values are average marginal effects in percentage points. Significance: \*\*\* p<0.001, \*\* p<0.01, \* p<0.05, † p<0.10.

## Procedural-Substantive Gap by Country-Wave

```{r tbl-cw-gap}
#| tbl-cap: "Procedural-Substantive Gap by Country and Wave"

if (!is.null(cw_gap)) {
  cw_gap %>%
    select(country_name, wave_label, wave,
           starts_with("proc_sub_gap"), starts_with("mean_n")) %>%
    mutate(
      gap_pp = sprintf("%+.1f", proc_sub_gap * 100),
      ci = sprintf("[%+.1f, %+.1f]",
                   proc_sub_gap_ci_low * 100,
                   proc_sub_gap_ci_high * 100),
      N = if("mean_n_procedural" %in% names(.)) {
        round(mean_n_procedural)
      } else NA_integer_
    ) %>%
    select(Country = country_name, Wave = wave_label,
           `Gap (pp)` = gap_pp, `95% CI` = ci) %>%
    arrange(Country, Wave) %>%
    kable(booktabs = TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped"))
} else {
  cat("**Warning:** Country-wave gap file not found.")
}
```

**Note:** Gap = (Mean Procedural AME) - (Mean Substantive AME). Positive values indicate losers favor procedural items more than substantive items relative to winners.

# Country Trajectory Plots {#sec-trajectories}

```{r fig-trajectories}
#| fig-cap: "Procedural-Substantive Gap Trajectories by Country"
#| fig-width: 6.5
#| fig-height: 8
#| out-width: "100%"
#| fig-pos: "H"

if (!is.null(cw_gap)) {
  # Filter to countries with 2+ waves
  multi_wave <- cw_gap %>%
    group_by(country_name) %>%
    filter(n() >= 2) %>%
    ungroup()

  # Order by magnitude of change
  country_order <- multi_wave %>%
    group_by(country_name) %>%
    summarise(
      max_gap = max(proc_sub_gap, na.rm = TRUE),
      min_gap = min(proc_sub_gap, na.rm = TRUE),
      change = max_gap - min_gap
    ) %>%
    arrange(desc(change)) %>%
    pull(country_name)

  multi_wave <- multi_wave %>%
    mutate(country_name = factor(country_name, levels = country_order))

  ggplot(multi_wave, aes(x = wave, y = proc_sub_gap * 100)) +
    geom_ribbon(aes(ymin = proc_sub_gap_ci_low * 100,
                    ymax = proc_sub_gap_ci_high * 100),
                fill = "steelblue", alpha = 0.3) +
    geom_line(color = "steelblue", linewidth = 1) +
    geom_point(color = "steelblue", size = 2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    facet_wrap(~country_name, ncol = 3) +
    scale_x_continuous(breaks = c(3, 4, 6),
                       labels = c("W3", "W4", "W6")) +
    labs(
      x = "Wave",
      y = "Procedural - Substantive Gap (pp)"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      strip.text = element_text(face = "bold", size = 8),
      panel.grid.minor = element_blank(),
      axis.title = element_text(size = 9),
      axis.text = element_text(size = 7)
    )
} else {
  cat("**Warning:** Country-wave gap file not found. Cannot generate trajectory plot.")
}
```

**Note:** Countries are ordered by the magnitude of change in the procedural-substantive gap across waves, with Thailand (largest change) first. Positive gap values indicate losers prioritize procedural items more than substantive items, relative to winners.

# Demographic Controls Comparison {#sec-controls}

This section compares pooled average marginal effects estimated with and without demographic controls to assess the stability of the loser effect.

```{r tbl-controls-comparison}
#| tbl-cap: "Comparison of AMEs With and Without Demographic Controls"

# Load both results files
controlled <- safe_read(file.path(results_dir, "loser_effects_all_items.csv"), "csv")
uncontrolled <- safe_read(file.path(results_dir, "loser_effects_all_items_no_controls.csv"), "csv")

if (!is.null(controlled) && !is.null(uncontrolled)) {
  # Join and compare
  comparison <- controlled %>%
    select(set, item_label, item_type, ame_controlled = ame, sig_controlled = sig) %>%
    inner_join(
      uncontrolled %>% select(set, item_label, ame_uncontrolled = ame, sig_uncontrolled = sig),
      by = c("set", "item_label")
    ) %>%
    mutate(
      sig_controlled = replace_na(sig_controlled, ""),
      sig_uncontrolled = replace_na(sig_uncontrolled, ""),
      diff = ame_controlled - ame_uncontrolled
    )

  # Format for display
  comparison %>%
    mutate(
      `AME with controls (pp)` = sprintf("%+.1f%s", ame_controlled * 100, sig_controlled),
      `AME without controls (pp)` = sprintf("%+.1f%s", ame_uncontrolled * 100, sig_uncontrolled),
      `Difference (pp)` = sprintf("%+.2f", diff * 100)
    ) %>%
    select(
      Item = item_label,
      Type = item_type,
      `AME with controls (pp)`,
      `AME without controls (pp)`,
      `Difference (pp)`
    ) %>%
    kable(booktabs = TRUE, align = c("l", "l", "r", "r", "r")) %>%
    kable_styling(latex_options = c("hold_position", "striped"),
                  font_size = 9) %>%
    pack_rows("Set 1", 1, 4) %>%
    pack_rows("Set 2", 5, 8) %>%
    pack_rows("Set 3", 9, 12) %>%
    pack_rows("Set 4", 13, 16) %>%
    pack_rows("Wave 2", 17, 20)
} else {
  cat("**Warning:** Results files not found. Cannot generate comparison table.")
}
```

**Note:** Demographic controls include age, gender, education level, and urban/rural residence. The stability of estimates across specifications indicates that the loser effect is not driven by compositional differences between winner and loser populations. Significance: \*\*\* p<0.001, \*\* p<0.01, \* p<0.05, † p<0.10.

# Wave 2 Detailed Results {#sec-w2}

```{r tbl-w2-results}
#| tbl-cap: "Wave 2 Multinomial Logit Results"

if (!is.null(pooled_results)) {
  w2_results <- pooled_results %>%
    filter(set == "W2")

  if (nrow(w2_results) > 0) {
    w2_results %>%
      mutate(
        sig = replace_na(sig, ""),
        ame_pp = sprintf("%+.1f", ame * 100),
        se_pp = sprintf("(%.1f)", se * 100),
        ci = sprintf("[%+.1f, %+.1f]", ci_low * 100, ci_high * 100)
      ) %>%
      select(
        Item = item_label,
        Type = item_type,
        `AME (pp)` = ame_pp,
        SE = se_pp,
        `95% CI` = ci,
        Sig = sig
      ) %>%
      kable(booktabs = TRUE, align = c("l", "l", "r", "r", "c", "c")) %>%
      kable_styling(latex_options = c("hold_position"))
  } else {
    cat("No W2 results found in pooled results file.")
  }
} else {
  cat("**Warning:** Pooled results file not found.")
}
```

**Note:** Wave 2 employed a single forced-choice item asking respondents to identify the "most essential characteristic of democracy" from four options. Unlike Sets 1--4 in later waves, these options directly compared procedural (elections, criticize power) and substantive (income equality, basic necessities) conceptions within a single item. The AMEs represent the percentage-point difference in the probability of selecting each option between electoral losers and winners, controlling for country fixed effects and demographic characteristics.

# Non-Voter Rates and Sample Selection {#sec-appendix-nonvoters}

Because the analysis conditions on reported vote choice to assign electoral status, non-voters are excluded by design. This section assesses the scope and consequences of this exclusion.

Table F1 reports non-voter rates by country and wave for the eleven countries in the analysis sample. In most country-waves, non-voter rates fall between 8 and 25 percent, consistent with the moderately high turnout typical of Asian democracies. Rates are lowest in Thailand, where they never exceeded 10 percent, and highest in select waves of Malaysia, Myanmar, and Mongolia.

```{r tbl-F1-nonvoter-rates}
#| tbl-cap: "Non-Voter Rates by Country and Wave"

rates <- read_csv(here("papers", "meaning-of-democracy", "analysis",
                       "revised", "results", "nonvoter_preanalysis",
                       "nonvoter_rates_by_country_wave.csv"),
                  show_col_types = FALSE)

kbl(rates, booktabs = TRUE, align = c("l", rep("r", ncol(rates) - 1))) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

A more direct test examines whether non-voters' democratic conceptions systematically resemble those of winners or losers. Table F2 reports the procedural-substantive gap across three groups: winners, losers, and non-voters. Non-voters' gap (14.3 percentage points, 95% CI: 12.7--15.8) is virtually identical to that of winners (13.9 pp, 95% CI: 13.2--14.7) and clearly below that of losers (17.3 pp, 95% CI: 16.3--18.3).

```{r tbl-F2-procsub-gap}
#| tbl-cap: "Procedural-Substantive Gap by Electoral Status Group"

procsub <- read_csv(here("papers", "meaning-of-democracy", "analysis",
                         "revised", "results", "nonvoter_preanalysis",
                         "procsub_gap_by_group.csv"),
                    show_col_types = FALSE)

kbl(procsub, booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options = c("hold_position"))
```

Table F3 disaggregates these patterns by item set.

```{r tbl-F3-item-proportions}
#| tbl-cap: "Item Choice Proportions by Electoral Status Group"

three_group <- read_csv(here("papers", "meaning-of-democracy", "analysis",
                             "revised", "results", "nonvoter_preanalysis",
                             "three_group_item_proportions.csv"),
                        show_col_types = FALSE)

kbl(three_group, booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

**Note:** L - W = Loser minus Winner; NV - W = Non-voter minus Winner. Positive values on procedural items indicate greater procedural orientation relative to winners.

Table F4 reports pairwise chi-square tests confirming that non-voters differ more from losers than from winners across all item sets (Cramér's V for the loser--non-voter comparison ranges from 0.020 to 0.052, consistently exceeding the winner--non-voter comparison of 0.021 to 0.038).

```{r tbl-F4-pairwise-tests}
#| tbl-cap: "Pairwise Chi-Square Tests and Effect Sizes"

chisq <- read_csv(here("papers", "meaning-of-democracy", "analysis",
                       "revised", "results", "nonvoter_preanalysis",
                       "pairwise_chisq_tests.csv"),
                  show_col_types = FALSE)

kbl(chisq, booktabs = TRUE, digits = 3) %>%
  kable_styling(latex_options = c("hold_position"))
```

These patterns indicate that non-voters resemble winners who chose not to participate rather than disaffected citizens with distinctive democratic conceptions. Their exclusion does not inflate the estimated loser effect.

# References {-}

::: {#refs}
:::
