---
title: "Online Appendix: What Does Democracy Mean to Losers?"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
date: today
format:
  pdf:
    documentclass: article
    classoption: [letterpaper, 12pt]
    geometry: margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    linkcolor: blue
    urlcolor: blue
    citecolor: blue
    keep-tex: true
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
bibliography: Winners-losers.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

here::i_am("papers/meaning-of-democracy/manuscript/online_appendix.qmd")

# Results directory (multinomial logit analysis outputs)
results_dir <- here("papers", "meaning-of-democracy", "analysis", "revised", "results")
data_dir <- here("papers", "meaning-of-democracy", "analysis", "revised", "data")

# Helper function to safely load files
safe_read <- function(path, type = "rds") {
  if (!file.exists(path)) {
    warning(paste("File not found:", path))
    return(NULL)
  }
  if (type == "rds") readRDS(path)
  else if (type == "csv") read_csv(path, show_col_types = FALSE)
  else stop("Unknown type")
}

# Load data files
item_meta <- safe_read(file.path(data_dir, "item_metadata.rds"))
w2_data <- safe_read(file.path(data_dir, "w2_baseline.rds"))
w346_data <- safe_read(file.path(data_dir, "w346_main.rds"))

# Load results files
cw_ames <- safe_read(file.path(results_dir, "country_wave_item_ames.csv"), "csv")
cw_gap <- safe_read(file.path(results_dir, "country_wave_proc_sub_gap.csv"), "csv")
pooled_results <- safe_read(file.path(results_dir, "loser_effects_all_items.csv"), "csv")
mlogit_results <- safe_read(file.path(results_dir, "mlogit_results.rds"))
```

# Survey Items and Coding {#sec-items}

This section documents the survey items used to measure conceptions of democracy across the Asian Barometer Survey waves.

## Item Reference Table

```{r tbl-items}
#| tbl-cap: "Survey Items by Set and Type"

if (!is.null(item_meta)) {
  item_meta %>%
    select(set, value, item_label, item_type, item_subtype) %>%
    rename(
      Set = set,
      Value = value,
      `Item Label` = item_label,
      `Item Type` = item_type,
      Subtype = item_subtype
    ) %>%
    kable(booktabs = TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped"),
                  font_size = 10) %>%
    pack_rows("Wave 2", 1, 4) %>%
    pack_rows("Set 1 (W3/W4/W6)", 5, 8) %>%
    pack_rows("Set 2 (W3/W4/W6)", 9, 12) %>%
    pack_rows("Set 3 (W3/W4/W6)", 13, 16) %>%
    pack_rows("Set 4 (W3/W4/W6)", 17, 20)
} else {
  cat("**Warning:** Item metadata file not found.")
}
```

**Note:** Wave 2 employed a single forced-choice item with four response options, while Waves 3, 4, and 6 used four separate item sets (Sets 1--4), each presenting four response options. The W2 instrument differs from the later waves and is analyzed separately.

# Sample Composition {#sec-sample}

## Sample Size by Country and Wave

```{r tbl-sample}
#| tbl-cap: "Sample Composition by Country and Wave"

if (!is.null(w2_data) && !is.null(w346_data)) {
  # Combine data
  w2_summary <- w2_data %>%
    mutate(wave_label = "W2") %>%
    group_by(country_name, wave_label) %>%
    summarise(
      N = n(),
      N_winner = sum(loser == 0),
      N_loser = sum(loser == 1),
      pct_loser = mean(loser) * 100,
      .groups = "drop"
    )

  w346_summary <- w346_data %>%
    group_by(country_name, wave_label) %>%
    summarise(
      N = n(),
      N_winner = sum(loser == 0),
      N_loser = sum(loser == 1),
      pct_loser = mean(loser) * 100,
      .groups = "drop"
    )

  sample_summary <- bind_rows(w2_summary, w346_summary) %>%
    arrange(country_name, wave_label)

  # Wide format for display
  sample_wide <- sample_summary %>%
    select(country_name, wave_label, N) %>%
    pivot_wider(names_from = wave_label, values_from = N, values_fill = 0) %>%
    mutate(Total = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>%
    arrange(country_name)

  # Add column totals
  col_totals <- sample_wide %>%
    summarise(across(where(is.numeric), sum)) %>%
    mutate(country_name = "**Total**")

  sample_wide <- bind_rows(sample_wide, col_totals)

  sample_wide %>%
    rename(Country = country_name) %>%
    kable(booktabs = TRUE, format.args = list(big.mark = ",")) %>%
    kable_styling(latex_options = c("hold_position", "striped"))
} else {
  cat("**Warning:** Data files not found. Cannot generate sample table.")
}
```

## Response Distributions by Winner/Loser Status

```{r tbl-response-dist}
#| tbl-cap: "Response Distributions by Electoral Status (Pooled)"

if (!is.null(w346_data)) {
  # Function to compute cross-tabs for a set
  compute_crosstab <- function(data, set_num) {
    choice_var <- paste0("set", set_num, "_choice")
    valid_var <- paste0("set", set_num, "_valid")

    data %>%
      filter(.data[[valid_var]]) %>%
      mutate(choice = .data[[choice_var]]) %>%
      group_by(loser, choice) %>%
      summarise(n = n(), .groups = "drop") %>%
      group_by(loser) %>%
      mutate(pct = n / sum(n) * 100) %>%
      ungroup() %>%
      select(loser, choice, pct) %>%
      pivot_wider(names_from = loser, values_from = pct, names_prefix = "loser_") %>%
      mutate(
        Winner = loser_0,
        Loser = loser_1,
        Diff = loser_1 - loser_0,
        Set = paste0("Set ", set_num)
      ) %>%
      select(Set, Item = choice, Winner, Loser, Diff)
  }

  # Compute for all sets
  all_crosstabs <- map_dfr(1:4, ~compute_crosstab(w346_data, .x))

  all_crosstabs %>%
    mutate(
      Winner = sprintf("%.1f%%", Winner),
      Loser = sprintf("%.1f%%", Loser),
      Diff = sprintf("%+.1f", Diff)
    ) %>%
    kable(booktabs = TRUE, align = c("l", "l", "r", "r", "r")) %>%
    kable_styling(latex_options = c("hold_position", "striped"),
                  font_size = 9) %>%
    pack_rows("Set 1", 1, 4) %>%
    pack_rows("Set 2", 5, 8) %>%
    pack_rows("Set 3", 9, 12) %>%
    pack_rows("Set 4", 13, 16)
} else {
  cat("**Warning:** W346 data file not found.")
}
```

**Note:** Diff = Loser - Winner. Positive values indicate losers are more likely to choose that item.

# Full Country-Wave Results {#sec-country-wave}

## Item-Level Average Marginal Effects

```{r tbl-cw-ames-sets12}
#| tbl-cap: "Country-Wave AMEs: Sets 1-2"

if (!is.null(cw_ames)) {
  # Create short item labels for column headers
  item_labels <- c(
    "Free elections" = "Elections",
    "Party competition" = "Parties",
    "Basic necessities" = "Necessities",
    "Reduce inequality" = "Inequality",
    "Gov explain" = "Explain",
    "Free speech" = "Speech",
    "Jobs for all" = "Jobs",
    "Narrow income gap" = "Income gap",
    "Free media" = "Media",
    "Fair elections" = "Fair elect",
    "Low unemployment" = "Unemp",
    "Welfare for poor" = "Welfare",
    "Check government" = "Check gov",
    "Minority rights" = "Minority",
    "Equal income" = "Equal inc",
    "No corruption" = "No corrupt"
  )

  # Format for display with short labels
  cw_display <- cw_ames %>%
    filter(set %in% c("Set1", "Set2")) %>%
    mutate(
      sig = replace_na(sig, ""),
      ame_fmt = sprintf("%+.1f%s", ame * 100, sig),
      item_short = if_else(item %in% names(item_labels),
                           item_labels[item],
                           substr(item, 1, 10))
    ) %>%
    select(country_name, wave_label, set, item_short, ame_fmt) %>%
    pivot_wider(
      names_from = c(set, item_short),
      values_from = ame_fmt,
      names_sep = "_"
    ) %>%
    arrange(country_name, wave_label)

  # Get column names for header grouping
  col_names <- names(cw_display)
  set1_cols <- sum(grepl("^Set1_", col_names))
  set2_cols <- sum(grepl("^Set2_", col_names))

  # Clean column names (remove Set prefix)
  names(cw_display) <- gsub("^Set[12]_", "", names(cw_display))

  cw_display %>%
    rename(Country = country_name, Wave = wave_label) %>%
    kable(booktabs = TRUE, longtable = TRUE, align = c("l", "c", rep("r", set1_cols + set2_cols))) %>%
    kable_styling(latex_options = c("hold_position", "repeat_header", "scale_down"),
                  font_size = 8) %>%
    add_header_above(c(" " = 2, "Set 1" = set1_cols, "Set 2" = set2_cols)) %>%
    column_spec(1, width = "1.8cm") %>%
    column_spec(2, width = "0.6cm") %>%
    row_spec(seq(3, nrow(cw_display), by = 3), extra_css = "border-bottom: 1px solid #ccc;",
             extra_latex_after = "\\midrule")
} else {
  cat("**Warning:** Country-wave AME file not found.")
}
```

```{r tbl-cw-ames-sets34}
#| tbl-cap: "Country-Wave AMEs: Sets 3-4"

if (!is.null(cw_ames)) {
  # Create short item labels for column headers
  item_labels <- c(
    "Free elections" = "Elections",
    "Party competition" = "Parties",
    "Basic necessities" = "Necessities",
    "Reduce inequality" = "Inequality",
    "Gov explain" = "Explain",
    "Free speech" = "Speech",
    "Jobs for all" = "Jobs",
    "Narrow income gap" = "Income gap",
    "Free media" = "Media",
    "Fair elections" = "Fair elect",
    "Low unemployment" = "Unemp",
    "Welfare for poor" = "Welfare",
    "Check government" = "Check gov",
    "Minority rights" = "Minority",
    "Equal income" = "Equal inc",
    "No corruption" = "No corrupt"
  )

  # Format for display with short labels
  cw_display <- cw_ames %>%
    filter(set %in% c("Set3", "Set4")) %>%
    mutate(
      sig = replace_na(sig, ""),
      ame_fmt = sprintf("%+.1f%s", ame * 100, sig),
      item_short = if_else(item %in% names(item_labels),
                           item_labels[item],
                           substr(item, 1, 10))
    ) %>%
    select(country_name, wave_label, set, item_short, ame_fmt) %>%
    pivot_wider(
      names_from = c(set, item_short),
      values_from = ame_fmt,
      names_sep = "_"
    ) %>%
    arrange(country_name, wave_label)

  # Get column names for header grouping
  col_names <- names(cw_display)
  set3_cols <- sum(grepl("^Set3_", col_names))
  set4_cols <- sum(grepl("^Set4_", col_names))

  # Clean column names (remove Set prefix)
  names(cw_display) <- gsub("^Set[34]_", "", names(cw_display))

  cw_display %>%
    rename(Country = country_name, Wave = wave_label) %>%
    kable(booktabs = TRUE, longtable = TRUE, align = c("l", "c", rep("r", set3_cols + set4_cols))) %>%
    kable_styling(latex_options = c("hold_position", "repeat_header", "scale_down"),
                  font_size = 8) %>%
    add_header_above(c(" " = 2, "Set 3" = set3_cols, "Set 4" = set4_cols)) %>%
    column_spec(1, width = "1.8cm") %>%
    column_spec(2, width = "0.6cm") %>%
    row_spec(seq(3, nrow(cw_display), by = 3), extra_css = "border-bottom: 1px solid #ccc;",
             extra_latex_after = "\\midrule")
} else {
  cat("**Warning:** Country-wave AME file not found.")
}
```

**Note:** Values are average marginal effects in percentage points. Significance: \*\*\* p<0.001, \*\* p<0.01, \* p<0.05, † p<0.10.

## Procedural-Substantive Gap by Country-Wave

```{r tbl-cw-gap}
#| tbl-cap: "Procedural-Substantive Gap by Country and Wave"

if (!is.null(cw_gap)) {
  cw_gap %>%
    select(country_name, wave_label, wave,
           starts_with("proc_sub_gap"), starts_with("mean_n")) %>%
    mutate(
      gap_pp = sprintf("%+.1f", proc_sub_gap * 100),
      ci = sprintf("[%+.1f, %+.1f]",
                   proc_sub_gap_ci_low * 100,
                   proc_sub_gap_ci_high * 100),
      N = if("mean_n_procedural" %in% names(.)) {
        round(mean_n_procedural)
      } else NA_integer_
    ) %>%
    select(Country = country_name, Wave = wave_label,
           `Gap (pp)` = gap_pp, `95% CI` = ci) %>%
    arrange(Country, Wave) %>%
    kable(booktabs = TRUE) %>%
    kable_styling(latex_options = c("hold_position", "striped"))
} else {
  cat("**Warning:** Country-wave gap file not found.")
}
```

**Note:** Gap = (Mean Procedural AME) - (Mean Substantive AME). Positive values indicate losers favor procedural items more than substantive items relative to winners.

# Country Trajectory Plots {#sec-trajectories}

```{r fig-trajectories}
#| fig-cap: "Procedural-Substantive Gap Trajectories by Country"
#| fig-width: 6.5
#| fig-height: 8
#| out-width: "100%"
#| fig-pos: "H"

if (!is.null(cw_gap)) {
  # Filter to countries with 2+ waves
  multi_wave <- cw_gap %>%
    group_by(country_name) %>%
    filter(n() >= 2) %>%
    ungroup()

  # Order by magnitude of change
  country_order <- multi_wave %>%
    group_by(country_name) %>%
    summarise(
      max_gap = max(proc_sub_gap, na.rm = TRUE),
      min_gap = min(proc_sub_gap, na.rm = TRUE),
      change = max_gap - min_gap
    ) %>%
    arrange(desc(change)) %>%
    pull(country_name)

  multi_wave <- multi_wave %>%
    mutate(country_name = factor(country_name, levels = country_order))

  ggplot(multi_wave, aes(x = wave, y = proc_sub_gap * 100)) +
    geom_ribbon(aes(ymin = proc_sub_gap_ci_low * 100,
                    ymax = proc_sub_gap_ci_high * 100),
                fill = "steelblue", alpha = 0.3) +
    geom_line(color = "steelblue", linewidth = 1) +
    geom_point(color = "steelblue", size = 2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    facet_wrap(~country_name, ncol = 3) +
    scale_x_continuous(breaks = c(3, 4, 6),
                       labels = c("W3", "W4", "W6")) +
    labs(
      x = "Wave",
      y = "Procedural - Substantive Gap (pp)"
    ) +
    theme_minimal(base_size = 9) +
    theme(
      strip.text = element_text(face = "bold", size = 8),
      panel.grid.minor = element_blank(),
      axis.title = element_text(size = 9),
      axis.text = element_text(size = 7)
    )
} else {
  cat("**Warning:** Country-wave gap file not found. Cannot generate trajectory plot.")
}
```

**Note:** Countries are ordered by the magnitude of change in the procedural-substantive gap across waves, with Thailand (largest change) first. Positive gap values indicate losers prioritize procedural items more than substantive items, relative to winners.

# Demographic Controls Comparison {#sec-controls}

<!-- TODO: Add controlled vs uncontrolled comparison when both are available -->

This section would compare pooled average marginal effects estimated with and without demographic controls (age, gender, education, urban residence) to assess the stability of the loser effect. Currently, only the controlled estimates are available.

# Wave 2 Detailed Results {#sec-w2}

```{r tbl-w2-results}
#| tbl-cap: "Wave 2 Multinomial Logit Results"

if (!is.null(pooled_results)) {
  w2_results <- pooled_results %>%
    filter(set == "W2")

  if (nrow(w2_results) > 0) {
    w2_results %>%
      mutate(
        sig = replace_na(sig, ""),
        ame_pp = sprintf("%+.1f", ame * 100),
        se_pp = sprintf("(%.1f)", se * 100),
        ci = sprintf("[%+.1f, %+.1f]", ci_low * 100, ci_high * 100)
      ) %>%
      select(
        Item = item_label,
        Type = item_type,
        `AME (pp)` = ame_pp,
        SE = se_pp,
        `95% CI` = ci,
        Sig = sig
      ) %>%
      kable(booktabs = TRUE, align = c("l", "l", "r", "r", "c", "c")) %>%
      kable_styling(latex_options = c("hold_position"))
  } else {
    cat("No W2 results found in pooled results file.")
  }
} else {
  cat("**Warning:** Pooled results file not found.")
}
```

**Note:** Wave 2 employed a single forced-choice item asking respondents to identify the "most essential characteristic of democracy" from four options. Unlike Sets 1--4 in later waves, these options directly compared procedural (elections, criticize power) and substantive (income equality, basic necessities) conceptions within a single item. The AMEs represent the percentage-point difference in the probability of selecting each option between electoral losers and winners, controlling for country fixed effects and demographic characteristics.

# References {-}

::: {#refs}
:::
