---
title: "01 - Descriptive Analysis"
format: html
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

```{r setup}
library(tidyverse)
library(knitr)
```

# Load Prepared Data

```{r load-data}
load("tables/prepared_data.RData")
```

# 1. Cross-National Trajectories (Waves 1–5)

Mean values by country and wave for key variables. Focus: HK vs Taiwan vs Singapore vs regional mean.

```{r trajectory-table}
trajectory_vars <- c(
  "dem_always_preferable", "democracy_satisfaction", "democracy_suitability",
  "dem_extent_current", "dem_country_present_govt",
  "trust_national_government", "trust_president", "trust_parliament",
  "trust_police", "trust_courts",
  "efficacy_ability_participate", "political_interest", "pol_discuss",
  "gov_free_to_organize", "dem_free_speech", "govt_responds_people",
  "election_free_fair", "system_needs_change", "system_proud",
  "system_deserves_support", "nat_willing_emigrate"
)
trajectory_vars <- trajectory_vars[trajectory_vars %in% names(d)]

# Focus country means
focus_means <- d |>
  filter(country_name %in% focus_countries) |>
  group_by(country_name, wave) |>
  summarise(across(all_of(trajectory_vars), ~mean(.x, na.rm = TRUE)),
            .groups = "drop")

# Regional mean (excluding focus countries)
regional_means <- d |>
  filter(!country_name %in% focus_countries) |>
  group_by(wave) |>
  summarise(across(all_of(trajectory_vars), ~mean(.x, na.rm = TRUE)),
            .groups = "drop") |>
  mutate(country_name = "Regional mean")

trajectories <- bind_rows(focus_means, regional_means)

# Display for each variable
for (var in trajectory_vars) {
  cat("\n###", var, "\n\n")
  tbl <- trajectories |>
    select(country_name, wave, !!sym(var)) |>
    pivot_wider(names_from = wave, values_from = !!sym(var), names_prefix = "W") |>
    mutate(across(where(is.numeric), ~round(.x, 3)))
  print(kable(tbl))
  cat("\n")
}
```

# 2. W4→W5 Change Comparison

```{r w4w5-change}
compare_countries <- c("Hong Kong", "Taiwan", "Singapore", "South Korea", "Japan")

w4w5 <- d |>
  filter(country_name %in% compare_countries, wave %in% c(4, 5)) |>
  group_by(country_name, wave) |>
  summarise(across(all_of(trajectory_vars), ~mean(.x, na.rm = TRUE)),
            .groups = "drop") |>
  pivot_wider(names_from = wave, values_from = all_of(trajectory_vars),
              names_sep = "_W")

# Calculate deltas
change_vars <- c("democracy_suitability", "trust_police", "trust_president",
                 "gov_free_to_organize", "dem_free_speech", "system_needs_change",
                 "dem_always_preferable", "democracy_satisfaction")
change_vars <- change_vars[change_vars %in% names(d)]

w4w5_deltas <- d |>
  filter(country_name %in% compare_countries, wave %in% c(4, 5)) |>
  group_by(country_name, wave) |>
  summarise(across(all_of(change_vars), ~mean(.x, na.rm = TRUE)),
            .groups = "drop") |>
  pivot_longer(cols = all_of(change_vars), names_to = "variable", values_to = "value") |>
  pivot_wider(names_from = wave, values_from = value, names_prefix = "W") |>
  mutate(delta = W5 - W4) |>
  select(country_name, variable, delta) |>
  pivot_wider(names_from = country_name, values_from = delta)

kable(w4w5_deltas, digits = 3,
      caption = "W4→W5 change by country (positive = increase)")
```

# 3. Cross-National Z-Scores (Wave 5)

```{r zscore-matrix}
w5 <- d |> filter(wave == 5)

zscore_vars <- c(
  "democracy_suitability", "dem_extent_current", "dem_always_preferable",
  "trust_president", "trust_police", "trust_courts", "trust_national_government",
  "gov_free_to_organize", "dem_free_speech", "election_free_fair",
  "system_needs_change", "system_proud", "democracy_satisfaction"
)
zscore_vars <- zscore_vars[zscore_vars %in% names(d)]

z_matrix <- w5 |>
  group_by(country_name) |>
  summarise(across(all_of(zscore_vars), ~mean(.x, na.rm = TRUE)),
            .groups = "drop")

# Compute z-scores
z_scored <- z_matrix |>
  mutate(across(all_of(zscore_vars),
                ~(. - mean(., na.rm = TRUE)) / sd(., na.rm = TRUE)))

# Display with HK highlighted
z_display <- z_scored |>
  arrange(country_name == "Hong Kong") |>  # HK first
  mutate(across(where(is.numeric), ~round(.x, 2)))

kable(z_display, caption = "Wave 5 z-scores (std. devs from cross-national mean)")
```

# 4. Pre/Post NSL Comparison (HK Wave 5)

```{r pre-post-nsl}
nsl_comparison <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL")) |>
  pivot_longer(cols = all_of(key_vars), names_to = "variable", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(variable, period) |>
  summarise(
    mean = mean(value),
    sd = sd(value),
    n = n(),
    .groups = "drop"
  )

# T-tests and effect sizes
nsl_tests <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL")) |>
  pivot_longer(cols = all_of(key_vars), names_to = "variable", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(variable) |>
  summarise(
    protest_mean = mean(value[period == "Protest"]),
    postnsl_mean = mean(value[period == "Post-NSL"]),
    delta = postnsl_mean - protest_mean,
    protest_sd = sd(value[period == "Protest"]),
    postnsl_sd = sd(value[period == "Post-NSL"]),
    protest_n = sum(period == "Protest"),
    postnsl_n = sum(period == "Post-NSL"),
    t_stat = t.test(value ~ period)$statistic,
    p_value = t.test(value ~ period)$p.value,
    .groups = "drop"
  ) |>
  mutate(
    # Cohen's d
    pooled_sd = sqrt(((protest_n - 1) * protest_sd^2 + (postnsl_n - 1) * postnsl_sd^2) /
                       (protest_n + postnsl_n - 2)),
    cohens_d = delta / pooled_sd,
    sig = case_when(
      p_value < 0.001 ~ "***",
      p_value < 0.01 ~ "**",
      p_value < 0.05 ~ "*",
      TRUE ~ ""
    )
  ) |>
  filter(protest_n >= 30, postnsl_n >= 30) |>
  arrange(p_value)

kable(
  nsl_tests |>
    select(variable, protest_mean, postnsl_mean, delta, cohens_d, p_value, sig) |>
    mutate(across(c(protest_mean, postnsl_mean, delta, cohens_d), ~round(.x, 3)),
           p_value = format.pval(p_value, digits = 3)),
  caption = "Pre/Post NSL comparison (HK Wave 5)"
)
```

# 5. Age Cohort Analysis

```{r age-analysis}
age_vars <- c(
  "democracy_suitability", "trust_police", "trust_national_government",
  "gov_free_to_organize", "dem_free_speech", "dem_always_preferable",
  "system_needs_change", "dem_extent_current"
)
age_vars <- age_vars[age_vars %in% names(d)]

age_means <- hk5 |>
  filter(!is.na(age_group)) |>
  pivot_longer(cols = all_of(age_vars), names_to = "variable", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(variable, age_group) |>
  summarise(mean = mean(value), n = n(), .groups = "drop")

# ANOVA for each variable
age_anova <- map_dfr(age_vars, function(var) {
  dat <- hk5 |> filter(!is.na(age_group), !is.na(!!sym(var)))
  if (nrow(dat) < 10) return(NULL)
  fit <- aov(as.formula(paste(var, "~ age_group")), data = dat)
  s <- summary(fit)[[1]]
  tibble(
    variable = var,
    F_stat = s$`F value`[1],
    p_value = s$`Pr(>F)`[1]
  )
})

# Display
age_display <- age_means |>
  pivot_wider(names_from = age_group, values_from = mean) |>
  left_join(age_anova, by = "variable") |>
  mutate(across(where(is.numeric), ~round(.x, 3)),
         sig = case_when(
           p_value < 0.001 ~ "***",
           p_value < 0.01 ~ "**",
           p_value < 0.05 ~ "*",
           TRUE ~ ""
         ))

kable(age_display, caption = "Age cohort means (HK W5) with ANOVA F-tests")
```

# 6. Pre/Post NSL × Age Interaction

```{r age-nsl-interaction}
interaction_vars <- c("democracy_suitability", "trust_police", "gov_free_to_organize")
interaction_vars <- interaction_vars[interaction_vars %in% names(d)]

age_nsl <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL"), !is.na(age_group)) |>
  pivot_longer(cols = all_of(interaction_vars), names_to = "variable", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(variable, age_group, period) |>
  summarise(mean = mean(value), n = n(), .groups = "drop") |>
  pivot_wider(names_from = period, values_from = c(mean, n)) |>
  mutate(delta = `mean_Post-NSL` - mean_Protest)

kable(age_nsl |>
        select(variable, age_group, mean_Protest, `mean_Post-NSL`, delta) |>
        mutate(across(where(is.numeric), ~round(.x, 3))),
      caption = "Pre/Post NSL shift by age group")
```

# 7. Flipped Correlation: dem_preferable → dem_satisfaction

```{r flipped-correlation}
corr_countries <- c("Hong Kong", "Taiwan", "Singapore",
                    "South Korea", "Japan", "Mongolia", "Philippines")

flipped <- d |>
  filter(wave %in% c(3, 4, 5)) |>
  filter(!is.na(dem_always_preferable), !is.na(democracy_satisfaction)) |>
  group_by(country_name) |>
  summarise(
    r = cor(dem_always_preferable, democracy_satisfaction, use = "complete.obs"),
    n = n(),
    .groups = "drop"
  ) |>
  arrange(desc(r))

kable(flipped |> mutate(r = round(r, 3)),
      caption = "Correlation: dem_always_preferable → democracy_satisfaction (W3-5)")
```

# 8. Save Tables for Manuscript

```{r save-tables}
save(nsl_tests, age_display, z_display, w4w5_deltas, flipped,
     age_nsl, nsl_comparison, trajectories,
     file = "tables/descriptive_results.RData")
```
