---
title: "00 — Data Preparation: The Democratic Aspiration Gap"
subtitle: "ABS · LBS · Afrobarometer — variable construction, indices, gap scores"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

**Purpose**: Load harmonized ABS, LBS, and Afrobarometer data; construct the
Democratic Aspiration Gap score and its components; build trust divergence
indices (executive vs. intermediary); construct authoritarian-alternatives
variables; add demographic controls; and save three analysis-ready datasets
plus a cross-survey summary frame.

- **Inputs**:
  - `abs_harmonized_path` → `abs_harmonized.rds`
  - `lbs_harmonized_path` → `lbs_harmonized.rds`
  - `afro_harmonized_path` → `afro_harmonized.rds`
  (all via `_data_config.R`)
- **Outputs**:
  - `analysis/results/dat_abs.rds`
  - `analysis/results/dat_lbs.rds`
  - `analysis/results/dat_afro.rds`
  - `analysis/results/dat_gap_summary.rds`  ← country-wave gap scores, all surveys

```{r setup}
library(tidyverse)

project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
source(file.path(project_root, "_data_config.R"))

paper_dir    <- file.path(project_root, "papers/10_democratic_aspiration_gap")
analysis_dir <- file.path(paper_dir, "analysis")
results_dir  <- file.path(analysis_dir, "results")
fig_dir      <- file.path(analysis_dir, "figures")
tables_dir   <- file.path(analysis_dir, "tables")

dir.create(results_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(fig_dir,     showWarnings = FALSE, recursive = TRUE)
dir.create(tables_dir,  showWarnings = FALSE, recursive = TRUE)

source(file.path(paper_dir, "R", "helpers.R"))
```

---

## Load raw harmonized data

```{r load-data}
abs_raw  <- readRDS(abs_harmonized_path)
lbs_raw  <- readRDS(lbs_harmonized_path)
afro_raw <- readRDS(afro_harmonized_path)

cat("ABS:  ", nrow(abs_raw),  "obs,", n_distinct(abs_raw$country),  "countries\n")
cat("LBS:  ", nrow(lbs_raw),  "obs,", n_distinct(lbs_raw$country),  "countries\n")
cat("Afro: ", nrow(afro_raw), "obs,", n_distinct(afro_raw$country), "countries\n")
```

---

## ABS: variable construction

### Trust indices

Executive trust = president + military (+ police as sensitivity check).
Intermediary trust = parliament + political parties + courts + election commission.
All items 1–4, higher = more trust (already harmonized).

Note: `trust_president` is missing from ABS Wave 1. The executive trust index
starts at Wave 2 for all ABS countries.

```{r abs-trust}
dat_abs <- abs_raw |>
  rowwise() |>
  mutate(
    # Executive trust (load-bearing for mechanism claim)
    trust_executive = mean(c_across(c(trust_president, trust_military)),
                           na.rm = TRUE),

    # Intermediary trust
    trust_intermediary = mean(c_across(c(trust_parliament,
                                         trust_political_parties,
                                         trust_courts,
                                         trust_election_commission)),
                              na.rm = TRUE),

    # Trust divergence: positive = executive > intermediary (hollowing pattern)
    trust_divergence = trust_executive - trust_intermediary,

    # Civil-society cluster (for robustness)
    trust_civil = mean(c_across(c(trust_ngos, trust_newspapers,
                                  trust_television)),
                       na.rm = TRUE)
  ) |>
  ungroup()
```

### Democratic Aspiration Gap score

Gap = democratic preference (demand) minus democratic supply assessment.
Both normalized 0–1 before differencing so the gap is on a common scale.

- Preference (supply-independent): `dem_best_form` (1–4, higher = pro-democracy)
- Supply assessment: `dem_sat_national` (1–4, higher = more satisfied)

```{r abs-gap}
dat_abs <- dat_abs |>
  mutate(
    dem_pref_n   = normalize_01(dem_best_form),
    dem_supply_n = normalize_01(democracy_satisfaction),
    dem_gap      = dem_pref_n - dem_supply_n
    # Positive gap: wants more democracy than they assess they have
    # Negative gap: satisfied or over-satisfied (rare)
  )
```

### Authoritarian alternatives

ABS items: `strongman_rule`, `expert_rule`, `military_rule`, `single_party_rule`
Scale: 1–4, **higher = more authoritarian support** (already harmonized).

For cross-survey comparability these will be rescaled to 0–1 within the paper
itself (see `02_models.qmd`). Raw harmonized values preserved here.

```{r abs-authalt}
dat_abs <- dat_abs |>
  rowwise() |>
  mutate(
    # Composite authoritarian alternatives index (0-1)
    authalt_index = mean(c_across(c(strongman_rule, expert_rule,
                                    military_rule, single_party_rule)),
                         na.rm = TRUE) |> (\(x) (x - 1) / 3)(),

    # Technocratic channel (theoretically distinct for Singapore/Cambodia/Malaysia)
    expert_rule_n = (expert_rule - 1) / 3,

    # Personalist channel
    strongman_rule_n = (strongman_rule - 1) / 3
  ) |>
  ungroup()
```

### Economic controls and demographics

```{r abs-controls}
dat_abs <- dat_abs |>
  mutate(
    econ_national_n = normalize_01(econ_national_now),
    econ_family_n   = normalize_01(econ_family_now),
    age_n           = normalize_01(age),
    edu_n           = normalize_01(education_level),
    urban           = if_else(urban_rural == 1, 1L, 0L)
  )
```

### Diagnostics

```{r abs-diag}
cat("=== ABS: Gap score coverage by country ===\n")
dat_abs |>
  group_by(country) |>
  summarise(
    n           = n(),
    waves       = paste(sort(unique(wave)), collapse = ", "),
    gap_nonmiss = sum(!is.na(dem_gap)),
    gap_mean    = round(mean(dem_gap, na.rm = TRUE), 3),
    trust_div_mean = round(mean(trust_divergence, na.rm = TRUE), 3),
    .groups = "drop"
  ) |>
  arrange(desc(gap_mean)) |>
  print(n = Inf)
```

---

## LBS: variable construction

### Trust indices

LBS trust battery: all items 1–4, higher = more trust (reversed from raw).

Executive cluster: `trust_president`, `trust_armed_forces`
Intermediary cluster: `trust_congress`, `trust_judiciary`, `trust_political_parties`,
`trust_electoral_authority`

```{r lbs-trust}
dat_lbs <- lbs_raw |>
  rowwise() |>
  mutate(
    trust_executive = mean(c_across(c(trust_president, trust_armed_forces)),
                           na.rm = TRUE),

    trust_intermediary = mean(c_across(c(trust_parliament,
                                          trust_courts,
                                          trust_political_parties,
                                          trust_elections)),
                              na.rm = TRUE),

    trust_divergence = trust_executive - trust_intermediary,

    # Media trust cluster (LBS-only; information environment moderator)
    trust_media = mean(c_across(c(trust_newspapers, trust_television,
                                   trust_radio)),
                       na.rm = TRUE)
  ) |>
  ungroup()
```

### Democratic Aspiration Gap score

- Preference: `dem_always_preferable` (higher = democracy preferable)
- Supply: `dem_satisfaction` (higher = more satisfied)

```{r lbs-gap}
dat_lbs <- dat_lbs |>
  mutate(
    dem_pref_n   = normalize_01(dem_always_preferable),
    dem_supply_n = normalize_01(dem_satisfaction),
    dem_gap      = dem_pref_n - dem_supply_n
  )
```

### Authoritarian alternatives

LBS does NOT have a strong-leader / expert-rule item in ordinal form.
Available items and their coverage:

- `dem_nondem_ok`: tolerance for non-democracy if it solves problems
  (1–4)
- `dem_military_support`: support military govt if things get difficult
  (binary 0/1)

NOTE on cross-survey harmonization: `dem_nondem_ok` is rescaled to 0–1
and compared to ABS `authalt_index` and Afro `authalt_index_rev`
in `02_models.qmd`. The Afro reject items run opposite in polarity — see
the Afro section below for the reversal formula.

```{r lbs-authalt}
dat_lbs <- dat_lbs |>
  mutate(
    # nondem_ok: normalize to 0-1
    nondem_ok_n = normalize_01(dem_nondem_ok)
  )
```

### Economic controls and corruption

Note: LBS harmonized data does not include demographic variables (age,
education, urban_rural). Models using LBS will omit demographic controls.

```{r lbs-controls}
dat_lbs <- dat_lbs |>
  mutate(
    econ_national_n    = normalize_01(econ_national_current),
    econ_personal_n    = normalize_01(econ_personal_current),
    # corruption_experience is binary (1=Yes, 2=No) — recode to 0/1
    corrupt_exp        = if_else(corruption_experience == 1, 1L, 0L, NA_integer_),
    corrupt_progress_n = normalize_01(corruption_progress)
  )
```

### Diagnostics

```{r lbs-diag}
cat("=== LBS: Gap score coverage by country ===\n")
dat_lbs |>
  group_by(country) |>
  summarise(
    n              = n(),
    waves          = paste(sort(unique(wave)), collapse = ", "),
    gap_nonmiss    = sum(!is.na(dem_gap)),
    gap_mean       = round(mean(dem_gap, na.rm = TRUE), 3),
    trust_div_mean = round(mean(trust_divergence, na.rm = TRUE), 3),
    .groups = "drop"
  ) |>
  arrange(desc(gap_mean)) |>
  print(n = Inf)
```

---

## Afrobarometer: variable construction

### Trust indices

Afro trust battery: all items 1–4, higher = more trust (shifted from 0–3 raw).

Executive cluster: `trust_president`, `trust_armed_forces`
Intermediary cluster: `trust_parliament`, `trust_political_parties`,
`trust_courts`, `trust_elections`
Note: `trust_traditional_leaders` is available R2/R4/R6–R9 — kept separate
as it represents an alternative legitimacy source.

```{r afro-trust}
dat_afro <- afro_raw |>
  rowwise() |>
  mutate(
    trust_executive = mean(c_across(c(trust_president, trust_armed_forces)),
                           na.rm = TRUE),

    trust_intermediary = mean(c_across(c(trust_parliament,
                                          trust_political_parties,
                                          trust_courts,
                                          trust_elections)),
                              na.rm = TRUE),

    trust_divergence = trust_executive - trust_intermediary
  ) |>
  ungroup()
```

### Democratic Aspiration Gap score

- Preference: `dem_support_preferable` (higher = democracy preferable)
- Supply: `dem_satisfaction` (higher = more satisfied)

Note: `dem_support_preferable` produced z=3.70 in Gambia — the strongest signal
in the entire dataset. Verify this item's wave coverage for Gambia before
running the main models.

```{r afro-gap}
dat_afro <- dat_afro |>
  mutate(
    dem_pref_n   = normalize_01(dem_support_preferable),
    dem_supply_n = normalize_01(dem_satisfaction),
    dem_gap      = dem_pref_n - dem_supply_n
  )
```

### Authoritarian alternatives

Afro items: `reject_one_party`, `reject_military_rule`, `reject_one_man_rule`
Scale: 1–5, **higher = MORE rejection (pro-democracy)** — OPPOSITE polarity
to ABS authoritarian items. Reversal formula for cross-survey comparison:
`authoritarian_support = (6 - reject_item)`, then rescale `(x - 1) / 4` → 0–1.

Raw harmonized direction preserved here. Reversal applied in `02_models.qmd`.

```{r afro-authalt}
dat_afro <- dat_afro |>
  rowwise() |>
  mutate(
    # Pro-democracy composite (as harmonized — higher = more democratic)
    prodem_index = mean(c_across(c(reject_one_party,
                                    reject_military_rule,
                                    reject_one_man_rule)),
                        na.rm = TRUE) |> (\(x) (x - 1) / 4)(),

    # Pre-reversed for cross-survey models (higher = more authoritarian,
    # matching ABS direction). Apply only where needed.
    authalt_index_rev = mean(c_across(c(reject_one_party,
                                         reject_military_rule,
                                         reject_one_man_rule)),
                             na.rm = TRUE) |> (\(x) 1 - ((x - 1) / 4))()
  ) |>
  ungroup()
```

### Corruption: behavioral bribery battery

Four items available across most rounds (R2–R9): `bribe_document`,
`bribe_police`, `bribe_school`, `bribe_medical`. These measure *experienced*
frequency (0–3), not perception — stronger for causal argument than perception-only.

```{r afro-corruption}
dat_afro <- dat_afro |>
  rowwise() |>
  mutate(
    bribery_index = mean(c_across(c(bribe_document, bribe_police,
                                     bribe_school, bribe_medical)),
                         na.rm = TRUE) |> (\(x) x / 3)()  # rescale to 0-1
  ) |>
  ungroup()
```

### Economic controls and demographics

```{r afro-controls}
dat_afro <- dat_afro |>
  mutate(
    econ_national_n = normalize_01(econ_national_current),
    econ_living_n   = normalize_01(econ_living_conditions),
    age_n           = normalize_01(age),
    edu_n           = normalize_01(education_level),
    urban           = if_else(urban_rural == 1, 1L, 0L)
  )
```

### Diagnostics

```{r afro-diag}
cat("=== Afro: Gap score coverage by country ===\n")
dat_afro |>
  group_by(country) |>
  summarise(
    n              = n(),
    rounds         = paste(sort(unique(wave)), collapse = ", "),
    gap_nonmiss    = sum(!is.na(dem_gap)),
    gap_mean       = round(mean(dem_gap, na.rm = TRUE), 3),
    trust_div_mean = round(mean(trust_divergence, na.rm = TRUE), 3),
    .groups = "drop"
  ) |>
  arrange(desc(gap_mean)) |>
  print(n = Inf)
```

---

## Cross-survey gap summary frame

Country-wave mean gap scores and trust divergence across all three surveys,
stacked into a single long frame. Used for Figure 1 (gap trajectories by region)
and the country typology.

```{r gap-summary}
# ABS
sum_abs <- dat_abs |>
  group_by(country, wave) |>
  summarise(
    survey         = "ABS",
    n              = n(),
    dem_gap        = mean(dem_gap,         na.rm = TRUE),
    dem_pref       = mean(dem_pref_n,      na.rm = TRUE),
    dem_supply     = mean(dem_supply_n,    na.rm = TRUE),
    trust_exec     = mean(trust_executive,    na.rm = TRUE),
    trust_int      = mean(trust_intermediary, na.rm = TRUE),
    trust_div      = mean(trust_divergence,   na.rm = TRUE),
    .groups = "drop"
  )

# LBS
sum_lbs <- dat_lbs |>
  group_by(country, wave) |>
  summarise(
    survey         = "LBS",
    n              = n(),
    dem_gap        = mean(dem_gap,         na.rm = TRUE),
    dem_pref       = mean(dem_pref_n,      na.rm = TRUE),
    dem_supply     = mean(dem_supply_n,    na.rm = TRUE),
    trust_exec     = mean(trust_executive,    na.rm = TRUE),
    trust_int      = mean(trust_intermediary, na.rm = TRUE),
    trust_div      = mean(trust_divergence,   na.rm = TRUE),
    .groups = "drop"
  )

# Afro
sum_afro <- dat_afro |>
  group_by(country, wave) |>
  summarise(
    survey         = "Afro",
    n              = n(),
    dem_gap        = mean(dem_gap,         na.rm = TRUE),
    dem_pref       = mean(dem_pref_n,      na.rm = TRUE),
    dem_supply     = mean(dem_supply_n,    na.rm = TRUE),
    trust_exec     = mean(trust_executive,    na.rm = TRUE),
    trust_int      = mean(trust_intermediary, na.rm = TRUE),
    trust_div      = mean(trust_divergence,   na.rm = TRUE),
    .groups = "drop"
  )

dat_gap_summary <- bind_rows(
  sum_abs  |> mutate(country = as.character(country)),
  sum_lbs  |> mutate(country = as.character(country)),
  sum_afro |> mutate(country = as.character(country))
) |>
  mutate(survey = factor(survey, levels = c("ABS", "LBS", "Afro")))

cat("Gap summary frame:", nrow(dat_gap_summary), "country-wave observations\n")
cat("By survey:\n")
count(dat_gap_summary, survey) |> print()
```

---

## Save

```{r save}
saveRDS(dat_abs,         file.path(results_dir, "dat_abs.rds"))
saveRDS(dat_lbs,         file.path(results_dir, "dat_lbs.rds"))
saveRDS(dat_afro,        file.path(results_dir, "dat_afro.rds"))
saveRDS(dat_gap_summary, file.path(results_dir, "dat_gap_summary.rds"))

cat("\n✓ Saved:\n")
cat("  dat_abs.rds        —", nrow(dat_abs), "obs,",
    n_distinct(dat_abs$country), "countries\n")
cat("  dat_lbs.rds        —", nrow(dat_lbs), "obs,",
    n_distinct(dat_lbs$country), "countries\n")
cat("  dat_afro.rds       —", nrow(dat_afro), "obs,",
    n_distinct(dat_afro$country), "countries\n")
cat("  dat_gap_summary.rds —", nrow(dat_gap_summary),
    "country-wave rows\n")
```
