---
title: "Data Exploration: Cambodia Elite Capture"
subtitle: "Initial inspection of variables for trust and elite capture analysis"
author: "Jeffrey Stark"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    df-print: paged
    theme: cosmo
    embed-resources: true
execute:
  warning: false
  message: false
---

# Overview

This document explores the Asian Barometer Wave 6 dataset to identify and assess variables relevant for analyzing elite capture through trust patterns in Cambodia (compared to Vietnam and Thailand).

**Key Research Questions:**

1. Do Cambodians show higher local trust but lower national trust (the "trust ladder")?
2. Does this pattern differ from Vietnam (bureaucratic) and Thailand (polarized)?
3. What variables can serve as proxies for elite capture and patronage?

# Setup
```{r 02-setup, message=FALSE, warning=FALSE}
# ============================================
# PACKAGE LOADING (Order matters!)
# ============================================

# Clear environment to prevent conflicts
rm(list = ls())

# Set CRAN mirror
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Load packages in CAREFUL ORDER to avoid masking
# Rule: Load general packages first, specific packages last

# 1. Core tidyverse (loads dplyr, ggplot2, tidyr, etc.)
library(tidyverse)

# 2. Data import/export
library(haven)        # For reading SPSS/Stata files
library(here)         # For project-relative paths

# 3. Data manipulation (reload to ensure dplyr is active)
library(dplyr)        # Explicit reload to prevent masking

# 4. Labelled data
library(labelled)     # For handling labeled data

# 5. Visualization and reporting
library(ggplot2)      # Already in tidyverse, but explicit
library(gt)           # For tables

# 6. Statistical analysis
library(psych)        # For factor analysis

# 7. Data quality and cleaning
library(naniar)       # For missing data visualization
library(skimr)        # For data summaries
library(janitor)      # For data cleaning

# ============================================
# PREVENT COMMON CONFLICTS
# ============================================

# Ensure we're using dplyr's key functions
select <- dplyr::select
filter <- dplyr::filter
mutate <- dplyr::mutate
summarise <- dplyr::summarise

# Set global options
options(scipen = 999)  # Disable scientific notation
set.seed(2025)         # Reproducibility

cat("✓ Packages loaded successfully\n")
cat("  dplyr::select is active:", exists("select", mode = "function"), "\n")

# ============================================
# SOURCE CUSTOM FUNCTIONS
# ============================================

t0 <- Sys.time()
cat("[setup] Sourcing custom functions at", format(t0), "\n")

source_success <- tryCatch({
  source(here("R", "utils", "_load_functions.R"))
  TRUE
}, error = function(e) {
  cat("❌ Error sourcing R/utils/_load_functions.R:", conditionMessage(e), "\n")
  FALSE
})

t1 <- Sys.time()
cat(sprintf("[setup] Sourcing completed in %.2f sec\n", 
            as.numeric(difftime(t1, t0, units = "secs"))))

if (!source_success) {
  stop("Failed to source R/utils/_load_functions.R")
}

# ============================================
# VERIFY CUSTOM FUNCTIONS
# ============================================

# Note: Reversal functions are sourced from R/utils/_load_functions.R
# - safe_reverse_3pt(): Reverses 1-3 scales
# - safe_reverse_4pt(): Reverses 1-4 scales  
# - safe_reverse_5pt(): Reverses 1-5 scales
# - verify_reversal(): Cross-tabulates original vs recoded
# - verify_no_invalid_codes(): Checks for invalid codes in variables
# - create_composite(): Creates mean composites with minimum item requirement

required_functions <- c(
  "safe_reverse_3pt", 
  "safe_reverse_4pt", 
  "safe_reverse_5pt", 
  "verify_reversal", 
  "verify_no_invalid_codes", 
  "create_composite"
)

cat("[setup] Verifying required functions exist...\n")

exists_fun <- vapply(
  required_functions,
  function(f) exists(f, mode = "function", inherits = TRUE),
  logical(1)
)

missing_functions <- required_functions[!exists_fun]

if (length(missing_functions) > 0) {
  stop("❌ Missing required functions: ", 
       paste(missing_functions, collapse = ", "))
}

cat("✓ All required functions available\n")

# ============================================
# FINAL VERIFICATION
# ============================================

# Test that select() is from dplyr
test_select <- tryCatch({
  # This should work if dplyr::select is active
  mtcars %>% select(mpg, cyl)
  TRUE
}, error = function(e) {
  cat("⚠ Warning: select() may not be from dplyr\n")
  FALSE
})

if (test_select) {
  cat("✓ dplyr::select() working correctly\n")
}

# Display package versions for reproducibility
cat("\n=== Key Package Versions ===\n")
cat("dplyr:", as.character(packageVersion("dplyr")), "\n")
cat("tidyverse:", as.character(packageVersion("tidyverse")), "\n")
cat("here:", as.character(packageVersion("here")), "\n")

cat("\n✓ Setup complete!\n")
```


```{r load_data}
# Load data
data <- readRDS(here("papers/02_cambodia_elite_capture/data/processed/ab_analysis.rds"))
```

# Basic Dataset Structure
```{r overview}
# Dimensions
cat("Dataset dimensions:\n")
cat("Rows:", nrow(data), "\n")
cat("Columns:", ncol(data), "\n\n")

# Countries included
cat("Countries in dataset:\n")
data %>% 
  count(country_name) %>% 
  arrange(desc(n))

```

# Variable Inventory

## All Available Variables
```{r all_variables}
# Get variable names and labels
var_labels <- data %>% 
  map_chr(~{
    label <- attr(.x, "label")
    if (is.null(label)) return("")
    return(label)
  })

# Create inventory
var_inventory <- tibble(
  variable = names(data),
  label = var_labels,
  type = map_chr(data, ~paste(class(.x), collapse = ", ")),
  n_unique = map_int(data, ~length(unique(.x))),
  n_missing = map_int(data, ~sum(is.na(.x)))
) %>%
  mutate(
    pct_missing = round(n_missing / nrow(data) * 100, 1)
  )

# Display full inventory
var_inventory %>%
  arrange(variable) %>%
  gt() %>%
  tab_header(
    title = "Complete Variable Inventory",
    subtitle = paste0("n = ", nrow(data), " observations")
  ) %>%
  cols_label(
    variable = "Variable Name",
    label = "Label",
    type = "Type",
    n_unique = "Unique Values",
    n_missing = "Missing (n)",
    pct_missing = "Missing (%)"
  )
```

# Key Variable Categories

## Trust Variables

Looking for variables measuring trust in different institutions/levels:
```{r trust_variables}
# Search for trust-related variables
trust_vars <- var_inventory %>%
  filter(str_detect(tolower(label), "trust|confidence") | 
         str_detect(tolower(variable), "trust|conf"))

trust_vars %>%
  gt() %>%
  tab_header(title = "Trust & Confidence Variables")
```

### Trust Variable Distributions (Cambodia)
```{r trust_distributions_cambodia}
if (nrow(trust_vars) > 0) {
  # Filter Cambodia only
  cambodia_data <- data %>% filter(country_name == "Cambodia")
  
  # Get trust variable names
  trust_var_names <- trust_vars$variable
  
  # Plot distributions
  cambodia_data %>%
    select(any_of(trust_var_names)) %>%
    pivot_longer(everything(), names_to = "variable", values_to = "value") %>%
    filter(!is.na(value)) %>%
    ggplot(aes(x = value)) +
    geom_bar() +
    facet_wrap(~variable, scales = "free") +
    labs(title = "Distribution of Trust Variables (Cambodia)",
         x = "Response", y = "Count") +
    theme_minimal()
} else {
  cat("No trust variables found yet.\n")
}
```

## Government Performance & Service Delivery
```{r government_performance}
# Search for government performance, service delivery
govt_vars <- var_inventory %>%
  filter(str_detect(tolower(label), "government|performance|service|deliver|satis") |
         str_detect(tolower(variable), "govt|perf|serv|satis"))

govt_vars %>%
  gt() %>%
  tab_header(title = "Government Performance & Service Variables")
```

## Corruption & Elite Capture Proxies
```{r corruption_vars}
# Search for corruption, elite, patron, local leader variables
capture_vars <- var_inventory %>%
  filter(str_detect(tolower(label), "corrupt|elite|patron|leader|village|local|chief") |
         str_detect(tolower(variable), "corr|elite|patron|leader|village|local"))

capture_vars %>%
  gt() %>%
  tab_header(title = "Potential Elite Capture Proxy Variables")
```

## Contact with Officials/Leaders
```{r contact_vars}
# Search for contact, approach, help-seeking
contact_vars <- var_inventory %>%
  filter(str_detect(tolower(label), "contact|approach|help|assist|request") |
         str_detect(tolower(variable), "contact|approach"))

contact_vars %>%
  gt() %>%
  tab_header(title = "Contact with Officials/Leaders Variables")
```

## Socioeconomic & Demographics
```{r demographics}
# Key demographics
demo_vars <- var_inventory %>%
  filter(str_detect(tolower(label), "age|gender|sex|education|income|urban|rural|employ") |
         str_detect(tolower(variable), "age|gender|sex|educ|income|urban|rural|employ"))

demo_vars %>%
  gt() %>%
  tab_header(title = "Demographic & Socioeconomic Variables")
```

# country_name-Specific Sample Sizes
```{r country_name_samples}
# Focus on Cambodia, Vietnam, Thailand
data %>%
  filter(country_name %in% c("Cambodia", "Vietnam", "Thailand")) %>%
  count(country_name) %>%
  gt() %>%
  tab_header(title = "Sample Sizes: Three-country_name Comparison")
```

# Missing Data Assessment
```{r missing_data}
# For Cambodia specifically, what's the missing data situation?
cambodia_data <- data %>% filter(country_name == "Cambodia")

missing_summary <- tibble(
  variable = names(cambodia_data),
  n_missing = map_int(cambodia_data, ~sum(is.na(.x))),
  pct_missing = round(map_int(cambodia_data, ~sum(is.na(.x))) / nrow(cambodia_data) * 100, 1)
) %>%
  filter(pct_missing > 0) %>%
  arrange(desc(pct_missing))

missing_summary %>%
  gt() %>%
  tab_header(
    title = "Missing Data (Cambodia Only)",
    subtitle = paste0("n = ", nrow(cambodia_data))
  ) %>%
  data_color(
    columns = pct_missing,
    colors = scales::col_numeric(
      palette = c("green", "yellow", "red"),
      domain = c(0, 100)
    )
  )
```

# Next Steps

Based on this exploration, we need to:

1. **Identify core trust variables** for the "trust ladder" hypothesis
   - Local level trust (village, neighbors, community)
   - National level trust (government, parliament, courts)
   
2. **Find or construct elite capture proxies**
   - Service delivery mechanisms (formal vs. informal)
   - Contact patterns with local leaders
   - Perceptions of local vs. national corruption
   
3. **Check variable consistency** across Cambodia, Vietnam, Thailand

4. **Create analysis-ready variables** with proper coding and labels

5. **Develop initial descriptive statistics** comparing the three countries

---

**Session Info**
```{r session_info}
sessionInfo()
```
---
