---
title: "Data Preparation and Cleaning"
subtitle: "Asian Barometer Wave 6 - Elite Capture Project"
author: "Jeffrey Stark"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: show
    df-print: paged
---

# Setup

```{r 00-header-info}
# ==============================================================================
# ASIAN BAROMETER WAVE 6 - DATA PREPARATION
# ==============================================================================
# Purpose: Clean and recode AB Wave 6 data for Elite Capture project
# Countries: Cambodia, Thailand, Vietnam, Phillipines, Indonesia
# N = ~6500 respondents
# 
# Key outputs:
#   - ab_analysis.rds: Main analysis dataset
#   - ab_analysis.csv: For sharing/Stata
#   - Figures: Missing data heatmap
#   - Tables: Reliability summary
#
# Dependencies:
#   - R/utils/_load_functions.R (reversal and composite functions)
#   - data/raw/w6_all_countries_merged.rds
#
# Last updated: [DATE]
# ==============================================================================
```

```{r 01-devtools, eval=interactive(), message=FALSE, warning=FALSE}
# Development-only tools; skipped during rendering
suppressPackageStartupMessages({
  library(lintr)
  library(goodpractice)
})
```

```{r 02-setup, message=FALSE, warning=FALSE}
# ============================================
# PACKAGE LOADING (Order matters!)
# ============================================

# Clear environment to prevent conflicts
rm(list = ls())

# Set CRAN mirror
options(repos = c(CRAN = "https://cloud.r-project.org"))

# Load packages in CAREFUL ORDER to avoid masking
# Rule: Load general packages first, specific packages last

# 1. Core tidyverse (loads dplyr, ggplot2, tidyr, etc.)
library(tidyverse)

# 2. Data import/export
library(haven)        # For reading SPSS/Stata files
library(here)         # For project-relative paths

# 3. Data manipulation (reload to ensure dplyr is active)
library(dplyr)        # Explicit reload to prevent masking

# 4. Labelled data
library(labelled)     # For handling labeled data

# 5. Visualization and reporting
library(ggplot2)      # Already in tidyverse, but explicit
library(gt)           # For tables

# 6. Statistical analysis
library(psych)        # For factor analysis

# 7. Data quality and cleaning
library(naniar)       # For missing data visualization
library(skimr)        # For data summaries
library(janitor)      # For data cleaning

# ============================================
# PREVENT COMMON CONFLICTS
# ============================================

# Ensure we're using dplyr's key functions
select <- dplyr::select
filter <- dplyr::filter
mutate <- dplyr::mutate
summarise <- dplyr::summarise

# Set global options
options(scipen = 999)  # Disable scientific notation
set.seed(2025)         # Reproducibility

cat("✓ Packages loaded successfully\n")
cat("  dplyr::select is active:", exists("select", mode = "function"), "\n")

# ============================================
# SOURCE CUSTOM FUNCTIONS
# ============================================

t0 <- Sys.time()
cat("[setup] Sourcing custom functions at", format(t0), "\n")

source_success <- tryCatch({
  source(here("R", "utils", "_load_functions.R"))
  TRUE
}, error = function(e) {
  cat("❌ Error sourcing R/utils/_load_functions.R:", conditionMessage(e), "\n")
  FALSE
})

t1 <- Sys.time()
cat(sprintf("[setup] Sourcing completed in %.2f sec\n", 
            as.numeric(difftime(t1, t0, units = "secs"))))

if (!source_success) {
  stop("Failed to source R/utils/_load_functions.R")
}

# ============================================
# VERIFY CUSTOM FUNCTIONS
# ============================================

# Note: Reversal functions are sourced from R/utils/_load_functions.R
# - safe_reverse_3pt(): Reverses 1-3 scales
# - safe_reverse_4pt(): Reverses 1-4 scales  
# - safe_reverse_5pt(): Reverses 1-5 scales
# - verify_reversal(): Cross-tabulates original vs recoded
# - verify_no_invalid_codes(): Checks for invalid codes in variables
# - create_composite(): Creates mean composites with minimum item requirement

required_functions <- c(
  "safe_reverse_3pt", 
  "safe_reverse_4pt", 
  "safe_reverse_5pt", 
  "verify_reversal", 
  "verify_no_invalid_codes", 
  "create_composite"
)

cat("[setup] Verifying required functions exist...\n")

exists_fun <- vapply(
  required_functions,
  function(f) exists(f, mode = "function", inherits = TRUE),
  logical(1)
)

missing_functions <- required_functions[!exists_fun]

if (length(missing_functions) > 0) {
  stop("❌ Missing required functions: ", 
       paste(missing_functions, collapse = ", "))
}

cat("✓ All required functions available\n")

# ============================================
# FINAL VERIFICATION
# ============================================

# Test that select() is from dplyr
test_select <- tryCatch({
  # This should work if dplyr::select is active
  mtcars %>% select(mpg, cyl)
  TRUE
}, error = function(e) {
  cat("⚠ Warning: select() may not be from dplyr\n")
  FALSE
})

if (test_select) {
  cat("✓ dplyr::select() working correctly\n")
}

# Display package versions for reproducibility
cat("\n=== Key Package Versions ===\n")
cat("dplyr:", as.character(packageVersion("dplyr")), "\n")
cat("tidyverse:", as.character(packageVersion("tidyverse")), "\n")
cat("here:", as.character(packageVersion("here")), "\n")

cat("\n✓ Setup complete!\n")
```

# Load Data

```{r 03-load-data}
# Load Asian Barometer Wave 6 data
# Adjust the path and file format based on your actual data file
# Common formats: .sav (SPSS), .dta (Stata), .csv

# Example for SPSS file:
ab_whole <- read_rds(here("data", "processed", "w6_all_countries_merged.rds"))

# Display basic information
cat("Raw data dimensions:", nrow(ab_whole), "rows x", ncol(ab_whole), "columns\n")
cat("Countries in dataset:", unique(ab_whole$country), "\n")
```

# Filter to Countries of Interest

```{r 04-filter-countries}
# Filter to Cambodia, Thailand, and Vietnam

# Identify country variable and values
cat("Country variable values:\n")
table(ab_whole$country)

# Filter to three countries
# Adjust country codes based on your data
ab_data <- ab_whole %>%
  filter(country %in% c(6, 9, 12, 11, 8))

cat("\nFiltered data dimensions:", nrow(ab_data), "rows x", ncol(ab_data), "columns\n")
cat("\nSample sizes by country:\n")
print(table(ab_data$country))
```

## Country-Specific Notes

```{r 05-country-notes}
# ============================================================================
# COUNTRY-SPECIFIC DATA NOTES
# ============================================================================

cat("\n=== COUNTRY-SPECIFIC NOTES ===\n\n")

cat("VIETNAM (country code = 11):\n")
cat("  - Missing q143a-e (COVID-specific restrictions) - not asked in survey\n")
cat("  - All other measures present\n")
cat("  - Sample size:", sum(ab_data$country == 11), "\n\n")

cat("CAMBODIA (country code = 12):\n")
cat("  - Complete data on all measures\n")
cat("  - Sample size:", sum(ab_data$country == 12), "\n\n")

cat("THAILAND (country code = 8):\n")
cat("  - Complete data on all measures\n")
cat("  - Sample size:", sum(ab_data$country == 8), "\n\n")

cat("PHILLIPINES (country code = 6):\n")
cat("  - Complete data on all measures\n")
cat("  - Sample size:", sum(ab_data$country == 6), "\n\n")

cat("INDONESIA (country code = 9):\n")
cat("  - Complete data on all measures\n")
cat("  - Sample size:", sum(ab_data$country == 9), "\n\n")

cat("✓ Country filtering complete\n")
```

# Variable Selection and Recoding

## Core Variables

```{r 06-select-variables}
# Trust in Institutions (q7-q15)
# Democracy Variables (q90-q92, q95, q128)
# Political Interest (q47-q49)
# Democracy/Freedom Trade-off (q126-q127)
# Authoritarianism Support (q129-q132)
# Acceptance of Authoritarian Actions (q168-q171)
# Emergency Powers General (q172a-e)
# COVID (q138, q140-q142)
# Family Impact from COVID (q139a-d)
# Pandemic-Specific Government Powers (q143a-e) - Note: Missing in Vietnam
# Corruption (q115-q119c)
# Clientelism (q79)
# Rich-poor treated equally (q104)
# Quality of governance
# Government responds to what people want (Responsiveness (or lack thereof) (q112)
# Leaders break law/abuse powerImpunity perceptionq (q109) 
# Holding government accountable between elections (q101)
# Accountability gaps
# Legislature keeps leaders in checkInstitutional constraints (q111)
# Ethnic equality in treatmentGroup-based capture (q103)

trust_vars <- paste0("q", 7:15)
democracy_vars <- c("q90", "q91", "q128", "q92", "q95")
political_interest_vars <- c("q47", "q48", "q49")
tradeoff_vars <- c("q126", "q127")
authoritarianism_vars <- paste0("q", 129:132)
acceptance_of_auth_vars <- paste0("q", 168:171)
emergency_powers_vars <- paste0("q172", letters[1:5])
covid_vars <- c("q138", "q140", "q141", "q142")
family_impact_vars <- paste0("q139", letters[1:4])
pandemic_gov_powers_vars <- paste0("q143", letters[1:5])
demographics <- c("level", "se2", "se3_1", "se5", "se5a")
corruption <- c("q115", "q116", "q117", "q118", "q119a", "q119b", "q119c")
vote_buying <- c("q79")
rich_poor_treated_equally <- c("q104")
govt_responsiveness <- c("q112")
govt_accountability <- c("q109", "q101", "q111")
ethnic_equality <- c("q103")
public_goods_ease_of_access <- paste0("q", 39:45)
govt_openness <- c("q108")
suffixes <- setdiff(letters[1:11], "j")
# There is no "j"
neighborhood_service_avail <- paste0("ir12", suffixes)
subjective_ses <- c("se13a")
political_efficacy <- c("q135")
elite_vs_mass <- c("se9c_3","se10c_3")

# Combine all variable names for selection
all_vars <- c(
  "country", "idnumber",  # Identifiers
  trust_vars,
  democracy_vars,
  political_interest_vars,
  tradeoff_vars,
  authoritarianism_vars,
  acceptance_of_auth_vars,
  emergency_powers_vars,
  covid_vars,
  family_impact_vars,
  pandemic_gov_powers_vars,
  demographics,
  corruption,
  vote_buying,
  rich_poor_treated_equally,
  govt_accountability,
  govt_responsiveness,
  ethnic_equality,
  public_goods_ease_of_access,
  neighborhood_service_avail,
  subjective_ses,
  political_efficacy,
  elite_vs_mass,
  govt_openness
)

ab_selected <- ab_data %>%
  select(all_of(all_vars))

cat("Selected", ncol(ab_selected), "variables for analysis\n")

# Extract question labels for validation
question_text <- sapply(ab_selected, function(x) {
  label <- attr(x, "label")
  if (is.null(label)) "" else label
})

cat("Extracted labels for", length(question_text), "variables\n")
```

## Recode Country Variable

```{r 07-recode-country}
# Create standardized country variable
ab_selected <- ab_selected %>%
  mutate(
    country_name = case_when(
      country == 12 ~ "Cambodia",
      country == 11 ~ "Vietnam",
      country == 8 ~ "Thailand",
      country == 6 ~ "Phillipines",
      country == 9 ~ "Indonesia",
      TRUE ~ as.character(country) # This correctly handles any unexpected codes
    ),
    # Convert to factor to control order in plots/tables
    country_name = factor(country_name, levels = c("Cambodia", "Vietnam", "Thailand", "Phillipines", "Indonesia"))
  )

# Create dummy variables for regression models
ab_selected <- ab_selected %>%
  mutate(
    cambodia = if_else(country_name == "Cambodia", 1, 0),
    vietnam = if_else(country_name == "Vietnam", 1, 0),
    thailand = if_else(country_name == "Thailand", 1, 0),
    phillipines = if_else(country_name == "Phillipines", 1, 0),
    indonesia = if_else(country_name == "Indonesia", 1, 0)
  )

# Verify the chunk ran
cat("=== Country Name Recoding Complete ===\n")
table(ab_selected$country_name)
```

## Missing Data Recode

```{r 08-missing-data-recode}
# ============================================================================
# RECODE SPECIAL MISSING CODES AS NA (SCALE-SPECIFIC)
# ============================================================================

cat("=== RECODING SPECIAL MISSING VALUES AS NA ===\n")

# Step 1: Universal missing codes (apply to ALL variables)
universal_missing <- c(-1, 0, 97, 98, 99)

ab_selected <- ab_selected %>%
  mutate(
    across(where(is.numeric), # Applies to ALL numeric columns
    ~if_else(. %in% universal_missing, NA_real_, .)
  )
)

cat("✓ Step 1 complete: Universal missing codes recoded\n")

# Step 2: Additional missing codes (7, 8, 9) for SHORT SCALE variables only
# EXCLUDE 10-point scales like q92 and q95 where 7,8,9 are valid!
short_scale_vars <- c(
  # Trust (4-point scales)
  paste0("q", 7:15),
  
  # Democracy (4-point only - EXCLUDE q92 and q95 which are 10-point!)
  "q90", "q91", "q128",
  
  # Trade-offs (5-point scales)
  "q126", "q127",
  
  # Authoritarianism Support (4-point)
  paste0("q", 129:132),
  
  # Acceptance of Authoritarian Actions (4-point)
  paste0("q", 168:171),
  
  # Emergency Powers (4-point)
  paste0("q172", letters[1:5]),
  
  # COVID Variables (most are 4-point or 3-point)
  "q138", "q140", "q141", "q142",
  paste0("q139", letters[1:4]),
  paste0("q143", letters[1:5]),
  
  # Political engagement (various scales)
  "q47", "q48", "q49",

  # Corruption
  "q115", "q116", "q117", "q118",
  paste0("119", letters[1:3]), 

  "q108"
)

# Apply 7,8,9 as missing ONLY to short scale variables
ab_selected <- ab_selected %>%
  mutate(
    across(
      all_of(short_scale_vars[short_scale_vars %in% names(.)]),
      ~if_else(. %in% c(7, 8, 9), NA_real_, .)
    )
  )

cat("✓ Step 2 complete: Short-scale missing codes (7,8,9) recoded\n")

cat("=== MISSING DATA RECODING COMPLETE ===\n")
```

## Demographics Recoding

```{r 09-demographics-recode}

# ============================================================================
# DEMOGRAPHICS RECODING
# ============================================================================

cat("\n=== DEMOGRAPHICS ===\n")

ab_selected <- ab_selected %>% 
  mutate(
    # Gender
    gender = case_when(
      se2 == 1 ~ "Male",
      se2 == 2 ~ "Female",
      TRUE ~ NA_character_
    ),
    
    # Urban/Rural
    urban = case_when(
      level == 1 ~ 0,  # Rural
      level == 2 ~ 1,  # Urban
      TRUE ~ NA_real_
    ),
    
    # Age (continuous)
    age = if_else(
      se3_1 > 0 & se3_1 < 120,  # Added upper bound check
      se3_1,
      NA_real_
    ),
    
    # Age groups for analysis
    age_group = case_when(
      age < 30 ~ "18-29",
      age < 45 ~ "30-44",
      age < 60 ~ "45-59",
      age >= 60 ~ "60+",
      TRUE ~ NA_character_
    ),
    age_group = factor(age_group, levels = c("18-29", "30-44", "45-59", "60+")),
    
    # Education level (categorical)
    educ_level = case_when(
      se5 %in% c(1, 2, 3) ~ "Primary or less",
      se5 %in% c(4, 5, 6, 7) ~ "Secondary",
      se5 %in% c(8, 9, 10) ~ "University/Tertiary",
      TRUE ~ NA_character_
    ),
    educ_level = factor(educ_level, levels = c(
      "Primary or less",
      "Secondary",
      "University/Tertiary"
    )),
    
    # Years of education (continuous)
    educ_years = case_when(
      se5a > 0 & se5a < 99 ~ se5a,
      TRUE ~ NA_real_
    )
  )

# Verify demographics
cat("\n=== DEMOGRAPHICS VERIFICATION ===\n")
table(ab_selected$gender, ab_selected$country_name, useNA="ifany")
summary(ab_selected$age)
table(ab_selected$educ_level, ab_selected$country_name, useNA="ifany")

cat("✓ Demographics recoded\n")
```

## Verification of Missing Data Recoding

```{r 10-verify-missing-data-recoding}
# ============================================================================
# VERIFICATION
# ============================================================================

cat("\n=== VERIFYING MISSING DATA RECODING ===\n")

# Check for any remaining invalid codes in numeric variables
cat("\n1. Checking for negative values (should be 0):\n")
negative_check <- ab_selected %>%
  summarise(across(where(is.numeric), 
                   ~sum(. < 0, na.rm=TRUE),
                   .names = "{.col}_negative")) %>%
  pivot_longer(everything()) %>%
  filter(value > 0)

if(nrow(negative_check) > 0) {
  cat("⚠ WARNING: Negative values found in:\n")
  print(negative_check)
} else {
  cat("✓ No negative values found\n")
}

# Verify 10-point scales retained 7,8,9
cat("\n2. Verifying 10-point scales (q92, q95) have full range:\n")
cat("\nq92 distribution:\n")
table(ab_selected$q92, useNA = "ifany") %>% print()

cat("\nq95 distribution:\n")
table(ab_selected$q95, useNA = "ifany") %>% print()

# Check if 7,8,9 are present
q92_has_789 <- any(ab_selected$q92 %in% 7:9, na.rm=TRUE)
q95_has_789 <- any(ab_selected$q95 %in% 7:9, na.rm=TRUE)

if(q92_has_789 & q95_has_789) {
  cat("\n✓ 10-point scales correctly retain values 7,8,9\n")
} else {
  cat("\n⚠ WARNING: 10-point scales missing values 7,8,9!\n")
}

# Verify Vietnam q143 variables are 100% missing
cat("\n3. Verifying Vietnam q143 variables (should be 100% missing):\n")
vietnam_q143_check <- ab_selected %>%
  filter(country_name == "Vietnam") %>%
  summarise(across(starts_with("q143"), 
                   ~round(100 * sum(is.na(.)) / n(), 1),
                   .names = "{.col}_pct_missing"))

print(vietnam_q143_check)

if(all(vietnam_q143_check == 100, na.rm=TRUE)) {
  cat("✓ Vietnam q143 variables correctly 100% missing\n")
} else {
  cat("⚠ WARNING: Vietnam q143 variables should be 100% missing\n")
}

# Verify demographics
cat("\n4. Demographics summary:\n")
cat("\nGender distribution:\n")
table(ab_selected$gender, ab_selected$country_name, useNA = "ifany") %>% 
  addmargins() %>% print()

cat("\nUrban/Rural distribution:\n")
table(ab_selected$urban, ab_selected$country_name, useNA = "ifany") %>%
  addmargins() %>% print()

cat("\nAge summary:\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    age_min = min(age, na.rm=TRUE),
    age_mean = round(mean(age, na.rm=TRUE), 1),
    age_max = max(age, na.rm=TRUE),
    age_missing = sum(is.na(age)),
    n = n()
  ) %>%
  print()

cat("\nEducation level distribution:\n")
table(ab_selected$educ_level, ab_selected$country_name, useNA = "ifany") %>%
  addmargins() %>% print()

cat("\nEducation years summary:\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    educ_years_mean = round(mean(educ_years, na.rm=TRUE), 1),
    educ_years_sd = round(sd(educ_years, na.rm=TRUE), 1),
    educ_years_missing = sum(is.na(educ_years)),
    n = n()
  ) %>%
  print()

# Check for impossible values in demographics
cat("\n5. Checking for impossible demographic values:\n")
impossible_age <- sum(ab_selected$age > 120, na.rm=TRUE)
impossible_educ_years <- sum(ab_selected$educ_years > 30, na.rm=TRUE)

if(impossible_age > 0) {
  cat("⚠ WARNING:", impossible_age, "cases with age > 120\n")
} else {
  cat("✓ No impossible age values\n")
}

if(impossible_educ_years > 0) {
  cat("⚠ WARNING:", impossible_educ_years, "cases with education years > 30\n")
} else {
  cat("✓ No impossible education values\n")
}

# Summary by country
cat("\n6. Overall data quality by country:\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    gender_complete = sum(!is.na(gender)),
    age_complete = sum(!is.na(age)),
    educ_complete = sum(!is.na(educ_level)),
    urban_complete = sum(!is.na(urban)),
    all_demo_complete = sum(!is.na(gender) & !is.na(age) & 
                            !is.na(educ_level) & !is.na(urban)),
    pct_all_demo_complete = round(100 * all_demo_complete / n, 1)
  ) %>%
  print()

cat("\n=== 08 MISSING DATA RECODING COMPLETE ===\n")
```

## Recode Trust Variables

```{r 11-recode-trust-variables}
# Document the original scales and recode to consistent direction
# Trust variables (q7-q15): Typically 1=Great deal, 4=None at all
# We'll recode so higher = more trust

# === INSTITUTIONAL TRUST ===

# Recode institutional trust variables (q7-q15) so higher = more trust
# Assuming original: 1=Great deal, 2=Quite a lot, 3=Not very much, 4=None at all
# Recode to: 4=Great deal, 3=Quite a lot, 2=Not very much, 1=None at all

# Recode trust variables
ab_selected <- ab_selected %>%
  mutate(
    across(q7:q15,
           safe_reverse_4pt,
           .names = "trust_{.col}")
  )

# Verify recoding worked
cat("\n--- Verifying Reversals ---\n")
verify_reversal(ab_data$q7, ab_selected$trust_q7)
verify_reversal(ab_data$q8, ab_selected$trust_q8)
verify_reversal(ab_data$q9, ab_selected$trust_q9)
verify_reversal(ab_data$q10, ab_selected$trust_q10)
verify_reversal(ab_data$q11, ab_selected$trust_q11)
verify_reversal(ab_data$q12, ab_selected$trust_q12)
verify_reversal(ab_data$q13, ab_selected$trust_q13)
verify_reversal(ab_data$q14, ab_selected$trust_q14)
verify_reversal(ab_data$q15, ab_selected$trust_q15)

# Verify no invalid codes remain (should all be 1-4 or NA)
cat("\n--- Checking for Invalid Codes ---\n")
invalid_check <- verify_no_invalid_codes(ab_selected, 
                                         paste0("trust_q", 7:15), 
                                         c(1,4))

if(nrow(invalid_check) > 0) {
  cat("⚠ WARNING: Invalid codes found!\n")
  print(invalid_check)
} else {
  cat("✓ All trust variables have valid codes (1-4 or NA)\n")
}

# --- Reliability analysis ---
cat("\n=== TRUST RELIABILITY ANALYSIS ===\n")
trust_items <- ab_selected %>%
  select(paste0("trust_q", 7:15)) %>%
  na.omit()
alpha_trust <- alpha(trust_items, check.keys = TRUE)
print(alpha_trust)

# Create composite
ab_selected <- ab_selected %>%
  mutate(
    institutional_trust_index = create_composite(
      ., 
      vars = paste0("trust_q", 7:15),
      min_items = 7
    )
  )
```

## Recode Democracy Variables

```{r 12-recode-4point-democracy-variables}
cat("\n========================================\n")
cat("SECTION 3: DEMOCRACY VARIABLES")
cat("\n========================================\n\n")
# ============================================================================
# DEMOCRACY VARIABLES - RECODING AND COMPOSITE CREATION
# ============================================================================

# --- Step 1: Recode 4-point scales ---
# q90: Satisfaction with democracy (1=very dissatisfied -> 4=very satisfied)
# q91: How much of a democracy (1=not a democracy -> 4=complete democracy)
# q128: Democracy may have problems but still best (1=strongly disagree -> 4=strongly agree)

ab_selected <- ab_selected %>%
  mutate(
    across(c(q90, q91, q128),
           safe_reverse_4pt,
           .names = "dem_{.col}")
  )

# Verify reversals
cat("=== VERIFYING 4-POINT DEMOCRACY SCALES ===\n")
verify_reversal(ab_data$q90, ab_selected$dem_q90)
verify_reversal(ab_data$q91, ab_selected$dem_q91)
verify_reversal(ab_data$q128, ab_selected$dem_q128)

# Verify no invalid codes remain (should all be 1-4 or NA)
cat("\n--- Checking for Invalid Codes ---\n")
invalid_check <- verify_no_invalid_codes(ab_selected, 
                                         c("q90", "q91", "q128"),
                                         c(1,4))

if(nrow(invalid_check) > 0) {
  cat("⚠ WARNING: Invalid codes found!\n")
  print(invalid_check)
} else {
  cat("✓ All trust variables have valid codes (1-4 or NA)\n")
}

# Check distributions
cat("\n=== DISTRIBUTION OF RECODED VARIABLES ===\n")
ab_selected %>%
  select(dem_q90, dem_q91, dem_q128) %>%
  summary() %>%
  print()
```

## Clean 10-Point Democracy Scales

```{r 13-clean-10point-democracy-scales}
# ============================================================================
# CLEAN 10-POINT DEMOCRACY SCALES
# ============================================================================
# q92: How democratic is current government [1..10]
# q95: How suitable is democracy for our country [1..10]
# 
# CRITICAL: These scales use 1-10, so 7,8,9 are VALID responses
#           Only recode universal missing codes (-1, 0, 97, 98, 99)

cat("\n=== CLEANING 10-POINT DEMOCRACY SCALES ===\n")

# --- Step 1: Verify universal missing codes were already applied ---
cat("\n1. Checking if universal missing codes were already removed:\n")
ab_selected %>%
  summarise(
    q92_has_negative = sum(q92 < 0, na.rm=TRUE),
    q92_has_97_99 = sum(q92 >= 97, na.rm=TRUE),
    q95_has_negative = sum(q95 < 0, na.rm=TRUE),
    q95_has_97_99 = sum(q95 >= 97, na.rm=TRUE)
  ) %>%
  print()

# --- Step 2: Check BEFORE cleaning ---
cat("\n2. BEFORE cleaning:\n")
cat("q92 range:", range(ab_selected$q92, na.rm=TRUE), "\n")
cat("q95 range:", range(ab_selected$q95, na.rm=TRUE), "\n")

cat("\nq92 unique values:\n")
sort(unique(ab_selected$q92[!is.na(ab_selected$q92)])) %>% print()

cat("\nq95 unique values:\n")
sort(unique(ab_selected$q95[!is.na(ab_selected$q95)])) %>% print()

# --- Step 3: Clean (only keep 1-10) ---
ab_selected <- ab_selected %>%
  mutate(
    # Clean q92: Only keep valid 1-10 responses
    q92_clean = case_when(
      q92 >= 1 & q92 <= 10 ~ q92,
      TRUE ~ NA_real_
    ),
    # Clean q95: Only keep valid 1-10 responses
    q95_clean = case_when(
      q95 >= 1 & q95 <= 10 ~ q95,
      TRUE ~ NA_real_
    )
  )

# --- Step 4: Verify AFTER cleaning ---
cat("\n3. AFTER cleaning:\n")
cat("q92_clean range:", range(ab_selected$q92_clean, na.rm=TRUE), "\n")
cat("q95_clean range:", range(ab_selected$q95_clean, na.rm=TRUE), "\n")

cat("\nq92_clean unique values (should be 1-10 only):\n")
sort(unique(ab_selected$q92_clean[!is.na(ab_selected$q92_clean)])) %>% print()

cat("\nq95_clean unique values (should be 1-10 only):\n")
sort(unique(ab_selected$q95_clean[!is.na(ab_selected$q95_clean)])) %>% print()

# --- Step 5: Missing data check ---
cat("\n4. Missing data:\n")
ab_selected %>%
  summarise(
    q92_original_na = sum(is.na(q92)),
    q92_clean_na = sum(is.na(q92_clean)),
    q92_newly_na = sum(is.na(q92_clean)) - sum(is.na(q92)),
    q95_original_na = sum(is.na(q95)),
    q95_clean_na = sum(is.na(q95_clean)),
    q95_newly_na = sum(is.na(q95_clean)) - sum(is.na(q95))
  ) %>%
  print()

cat("\nCases recoded to missing (these had invalid codes):\n")
cat("q92:", sum(is.na(ab_selected$q92_clean)) - sum(is.na(ab_selected$q92)), "\n")
cat("q95:", sum(is.na(ab_selected$q95_clean)) - sum(is.na(ab_selected$q95)), "\n")

# --- Step 6: Full value distributions ---
cat("\n5. Full value distributions:\n")
cat("\nq92_clean frequency table:\n")
table(ab_selected$q92_clean, useNA = "ifany") %>% 
  addmargins() %>% 
  print()

cat("\nq95_clean frequency table:\n")
table(ab_selected$q95_clean, useNA = "ifany") %>% 
  addmargins() %>% 
  print()

# --- Step 7: Distribution by country ---
cat("\n6. Distribution by country:\n")
cat("\nq92_clean by country:\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    mean = round(mean(q92_clean, na.rm=TRUE), 2),
    sd = round(sd(q92_clean, na.rm=TRUE), 2),
    min = min(q92_clean, na.rm=TRUE),
    max = max(q92_clean, na.rm=TRUE),
    n_missing = sum(is.na(q92_clean)),
    pct_missing = round(100 * sum(is.na(q92_clean)) / n(), 1)
  ) %>%
  print()

cat("\nq95_clean by country:\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    mean = round(mean(q95_clean, na.rm=TRUE), 2),
    sd = round(sd(q95_clean, na.rm=TRUE), 2),
    min = min(q95_clean, na.rm=TRUE),
    max = max(q95_clean, na.rm=TRUE),
    n_missing = sum(is.na(q95_clean)),
    pct_missing = round(100 * sum(is.na(q95_clean)) / n(), 1)
  ) %>%
  print()

# --- Step 8: Critical verification checks ---
cat("\n7. CRITICAL VERIFICATION:\n")

# Check 1: Verify 7, 8, 9 are present (not accidentally removed)
q92_has_789 <- all(c(7, 8, 9) %in% ab_selected$q92_clean)
q95_has_789 <- all(c(7, 8, 9) %in% ab_selected$q95_clean)

cat("\nq92_clean contains values 7, 8, 9:", 
    ifelse(q92_has_789, "✓ YES", "⚠ NO - PROBLEM!"), "\n")
cat("q95_clean contains values 7, 8, 9:", 
    ifelse(q95_has_789, "✓ YES", "⚠ NO - PROBLEM!"), "\n")

if(!q92_has_789 | !q95_has_789) {
  cat("\n⚠⚠⚠ WARNING: 10-point scales missing mid-range values! ⚠⚠⚠\n")
  cat("This suggests 7,8,9 were incorrectly coded as missing.\n")
  cat("Check your missing data recoding in the previous chunk.\n")
}

# Check 2: Verify no values outside 1-10
invalid_q92 <- verify_no_invalid_codes(ab_selected, "q92_clean", c(1, 10))
invalid_q95 <- verify_no_invalid_codes(ab_selected, "q95_clean", c(1, 10))

if(nrow(invalid_q92) > 0 | nrow(invalid_q95) > 0) {
  cat("\n⚠ WARNING: Invalid codes found!\n")
  if(nrow(invalid_q92) > 0) print(invalid_q92)
  if(nrow(invalid_q95) > 0) print(invalid_q95)
  stop("Fix invalid codes before proceeding")
} else {
  cat("\n✓ All values are within valid range (1-10 or NA)\n")
}

# Check 3: Verify full range is used
q92_range <- range(ab_selected$q92_clean, na.rm=TRUE)
q95_range <- range(ab_selected$q95_clean, na.rm=TRUE)

cat("\nFinal ranges:\n")
cat("q92_clean: ", q92_range[1], "to", q92_range[2], 
    ifelse(q92_range[1]==1 & q92_range[2]==10, "✓", "⚠"), "\n")
cat("q95_clean: ", q95_range[1], "to", q95_range[2], 
    ifelse(q95_range[1]==1 & q95_range[2]==10, "✓", "⚠"), "\n")

# Check 4: Compare with original to see what was removed
cat("\n8. What was removed during cleaning?\n")

# Show examples of cases that were recoded to NA
removed_q92 <- ab_selected %>%
  filter(!is.na(q92) & is.na(q92_clean)) %>%
  select(country_name, q92, q92_clean) %>%
  head(10)

removed_q95 <- ab_selected %>%
  filter(!is.na(q95) & is.na(q95_clean)) %>%
  select(country_name, q95, q95_clean) %>%
  head(10)

if(nrow(removed_q92) > 0) {
  cat("\nExamples of q92 values recoded to NA:\n")
  print(removed_q92)
} else {
  cat("\nNo q92 values were recoded to NA (good if universal missing already handled)\n")
}

if(nrow(removed_q95) > 0) {
  cat("\nExamples of q95 values recoded to NA:\n")
  print(removed_q95)
} else {
  cat("\nNo q95 values were recoded to NA (good if universal missing already handled)\n")
}

cat("\n=== 10-POINT SCALE CLEANING COMPLETE ===\n")
cat("Proceed only if all checks passed ✓\n")
```

## Create Democracy Composites

```{r 14-create-democracy-composites}
# --- Step 3: Standardize all variables (z-scores) ---
ab_selected <- ab_selected %>%
  mutate(
    z_dem_q90 = scale(dem_q90)[,1],
    z_dem_q91 = scale(dem_q91)[,1],
    z_dem_q128 = scale(dem_q128)[,1],
    z_dem_q92 = scale(q92_clean)[,1],      # Use cleaned version
    z_dem_q95 = scale(q95_clean)[,1]       # Use cleaned version
  )

cat("\n=== STANDARDIZED VARIABLES ===\n")
ab_selected %>%
  select(starts_with("z_dem")) %>%
  summary() %>%
  print()

# Verify z-scores have reasonable ranges
cat("\nChecking z-score ranges (should be roughly -3 to +3):\n")
cat("z_dem_q92 range:", range(ab_selected$z_dem_q92, na.rm=TRUE), "\n")
cat("z_dem_q95 range:", range(ab_selected$z_dem_q95, na.rm=TRUE), "\n")

# --- Step 4: Create democracy composites ---
ab_selected <- ab_selected %>%
  mutate(
    # Dimension 1: Democratic satisfaction (performance legitimacy)
    # Based on q90 (satisfaction) + q91 (level of democracy)
    dem_satisfaction = rowMeans(
      cbind(z_dem_q90, z_dem_q91),
      na.rm = TRUE
    ),
    
    # Dimension 2: Democratic legitimacy (normative support)
    # Based on q92 (current govt democratic) + q95 (democracy suitable)
    # Note: q128 excluded due to poor reliability
    dem_legitimacy = rowMeans(
      cbind(z_dem_q92, z_dem_q95),
      na.rm = TRUE
    )
  )

# --- Step 5: Reliability checks ---
cat("\n=== RELIABILITY ANALYSIS ===\n")

# Satisfaction scale (2 items)
cat("\n--- Democratic Satisfaction Scale ---\n")
r_satisfaction <- cor(ab_selected$z_dem_q90, ab_selected$z_dem_q91, 
                      use = "complete.obs")
sb_satisfaction <- (2 * r_satisfaction) / (1 + r_satisfaction)
cat("Inter-item correlation:", round(r_satisfaction, 3), "\n")
cat("Spearman-Brown reliability:", round(sb_satisfaction, 3), "\n")
cat("N with complete data:", sum(!is.na(ab_selected$z_dem_q90) & 
                                   !is.na(ab_selected$z_dem_q91)), "\n")

# Legitimacy scale (2 items)
cat("\n--- Democratic Legitimacy Scale ---\n")
r_legitimacy <- cor(ab_selected$z_dem_q92, ab_selected$z_dem_q95, 
                    use = "complete.obs")
sb_legitimacy <- (2 * r_legitimacy) / (1 + r_legitimacy)
cat("Inter-item correlation:", round(r_legitimacy, 3), "\n")
cat("Spearman-Brown reliability:", round(sb_legitimacy, 3), "\n")
cat("N with complete data:", sum(!is.na(ab_selected$z_dem_q92) & 
                                   !is.na(ab_selected$z_dem_q95)), "\n")

# --- Step 6: Check construct validity ---
cat("\n=== CONSTRUCT VALIDITY ===\n")
cat("Correlation between satisfaction and legitimacy:\n")
cor_test <- cor.test(ab_selected$dem_satisfaction, 
                     ab_selected$dem_legitimacy, 
                     use = "complete.obs")
print(cor_test)

if (abs(cor_test$estimate) < 0.3) {
  cat("\n✓ Low correlation (r =", round(cor_test$estimate, 3), 
      ") confirms these are distinct constructs.\n")
} else if (abs(cor_test$estimate) > 0.7) {
  cat("\n⚠ High correlation (r =", round(cor_test$estimate, 3), 
      ") suggests constructs may overlap.\n")
} else {
  cat("\n→ Moderate correlation (r =", round(cor_test$estimate, 3), 
      ") suggests related but distinct constructs.\n")
}

# --- Step 7: Descriptive statistics ---
cat("\n=== DESCRIPTIVE STATISTICS FOR COMPOSITES ===\n")
ab_selected %>%
  select(dem_satisfaction, dem_legitimacy, dem_q128) %>%
  summary() %>%
  print()

cat("\nMissing data:\n")
ab_selected %>%
  summarise(
    satisfaction_na = sum(is.na(dem_satisfaction)),
    legitimacy_na = sum(is.na(dem_legitimacy)),
    q128_na = sum(is.na(dem_q128))
  ) %>%
  print()

# --- Step 8: Distribution checks ---
cat("\n=== CHECKING FOR OUTLIERS ===\n")
n_sat_outliers <- sum(abs(ab_selected$dem_satisfaction) > 3, na.rm=TRUE)
n_leg_outliers <- sum(abs(ab_selected$dem_legitimacy) > 3, na.rm=TRUE)

cat("Satisfaction outliers (|z| > 3):", n_sat_outliers, "\n")
cat("Legitimacy outliers (|z| > 3):", n_leg_outliers, "\n")

if (n_sat_outliers > 0 | n_leg_outliers > 0) {
  cat("\n⚠ Warning: Outliers detected. Review data cleaning.\n")
} else {
  cat("\n✓ No extreme outliers detected.\n")
}

# --- Final summary ---
cat("\n=== 10 DEMOCRACY VARIABLES READY ===\n")
cat("Variables created:\n")
cat("  - dem_satisfaction: Performance legitimacy (α ≈", 
    round(sb_satisfaction, 2), ")\n")
cat("  - dem_legitimacy: Normative support (α ≈", 
    round(sb_legitimacy, 2), ")\n")
cat("  - dem_q128: Democratic preference (separate item)\n")
cat("\nSample sizes:\n")
cat("  - Satisfaction:", sum(!is.na(ab_selected$dem_satisfaction)), "\n")
cat("  - Legitimacy:", sum(!is.na(ab_selected$dem_legitimacy)), "\n")
cat("  - DemStillBest:", sum(!is.na(ab_selected$dem_q128)), "\n")
```

## Recode Political Engagement Variables

```{r 15-recode-political-variables}
cat("\n========================================\n")
cat("SECTION 3: POLITICAL ENGAGEMENT VARIABLES")
cat("\n========================================\n\n")
## POLITICAL ENGAGEMENT

ab_selected <- ab_selected %>%
  mutate(
    # q47: Interest in politics (1=Very interested, 4=Not at all)
    # Reverse so: 4=Very interested, 1=Not at all
    interest_politics = safe_reverse_4pt(q47),

    # q48: Follow news (1=Everyday, 5=Practically never)
    # Reverse so: 5=Everyday, 1=Practically never
    follow_news = safe_reverse_5pt(q48),

    # q49: Discuss politics
    # Reverse so: higher = more discussion
    discuss_politics = safe_reverse_3pt(q49)
  )

cat("=== VERIFYING VARIOUS POLITICAL INTEREST SCALES ===\n")
verify_reversal(ab_data$q47, ab_selected$interest_politics)
verify_reversal(ab_data$q48, ab_selected$follow_news)
verify_reversal(ab_data$q49, ab_selected$discuss_politics)
```

Corruption Variables

```{r 15b-recode-corruption-variables}
cat("\n========================================\n")
cat("SECTION 3: CORRUPTION VARIABLES")
cat("\n========================================\n\n")
## CORRUPTION

ab_selected <- ab_selected %>%
  mutate(
    corruption_local = case_when(
      q115 == 5 ~ 0,  # No one involved = least corrupt
      q115 == 1 ~ 1,  # Hardly anyone
      q115 == 2 ~ 2,  # Not a lot
      q115 == 3 ~ 3,  # Most
      q115 == 4 ~ 4,  # Almost everyone = most corrupt
      TRUE ~ NA_real_
    ),
    corruption_national = case_when(
      q116 == 5 ~ 0,  # No one involved = least corrupt
      q116 == 1 ~ 1,  # Hardly anyone
      q116 == 2 ~ 2,  # Not a lot
      q116 == 3 ~ 3,  # Most
      q116 == 4 ~ 4,  # Almost everyone = most corrupt
      TRUE ~ NA_real_
    )
  )

# Check distribution
table(ab_selected$q115, ab_selected$corruption_local, useNA = "ifany")
table(ab_selected$q116, ab_selected$corruption_national, useNA = "ifany")

ab_selected <- ab_selected %>%
  mutate(

    vote_buying = safe_reverse_3pt(q79,
                                     validate_keyword = "food, cash",
                                     question_text = question_text,
                                     var_name = "q79"),

    # q117: How effective is the government in cracking down on corruption?
    # Keyword validation: 'cracking'
    corruption_crackdown = safe_reverse_4pt(q117,
                                     validate_keyword = "cracking",
                                     question_text = question_text,
                                     var_name = "q117"),


    # q118: Know first hand of gov't corruption (1=Yes, 0=No)
    corruption_intimate_knowledge = case_when(
      q118 == 1 ~ 1,
      q118 == 2 ~ 0,
      TRUE ~ NA_real_
    ),

    # q119a: Personally witnessed corruption (1=Yes, 0=No)
    corruption_personally_witnessed = case_when(
      q119a == 1 ~ 1,
      q119a == 2 ~ 0,
      TRUE ~ NA_real_
    ),

    # q119b: Family witnessed corruption (1=Yes, 0=No)
    corruption_family_witnessed = case_when(
      q119b == 1 ~ 1,
      q119b == 2 ~ 0,
      TRUE ~ NA_real_
    ),

    # q119a: Friend witnessed corruption (1=Yes, 0=No)
    corruption_friend_witnessed = case_when(
      q119c == 1 ~ 1,
      q119c == 2 ~ 0,
      TRUE ~ NA_real_
    ),
  )

cat("=== VERIFYING VARIOUS CORRUPTION SCALES ===\n")
verify_reversal(ab_data$q79, ab_selected$vote_buying)
verify_reversal(ab_data$q117, ab_selected$corruption_crackdown)
verify_reversal(ab_data$q118, ab_selected$corruption_intimate_knowledge)
verify_reversal(ab_data$q119a, ab_selected$corruption_personally_witnessed)
verify_reversal(ab_data$q119b, ab_selected$corruption_family_witnessed)
verify_reversal(ab_data$q119c, ab_selected$corruption_friend_witnessed)
```

```{r 15c-elite-capture-addl-variables}
cat("\n========================================\n")
cat("SECTION 3: ELITE CAPTURE ADDITIONAL VARIABLES")
cat("\n========================================\n\n")
## ELITE CAPTURE ADD'L

ab_selected <- ab_selected %>%
  mutate(

    # q101  Between elections, the people have no way of holding the government responsible for its actions. 
    # Keyword validation: 'holding'
    govt_accountable_people = safe_reverse_4pt(q101,
      validate_keyword = "elections", question_text = question_text,
                                     var_name = "q101"),

    # q103: All citizens from different ethnic communities in the [Country X] are treated equally by the government. 
    # Keyword validation: 'ethnic communities'
    ethnic_groups_equal = safe_reverse_4pt(q103,
      validate_keyword = "ethnic communities", question_text = question_text,
                                     var_name = "q103"),

    # q104  Rich and poor people are treated equally by the government
    # Keyword validation: 'rich'
    rich_and_poor_equal = safe_reverse_4pt(q104,
                                     validate_keyword = "rich",
                                     question_text = question_text,
                                     var_name = "q104"),

    # q108 How often do government officials withhold important information from the public? 
    # Keyword validation: 'withhold'
    govt_withholds_info = safe_reverse_4pt(q108,
      validate_keyword = "withhold", question_text = question_text,
                                     var_name = "q108"),

    # q109 How often do you think government leaders break the law or abuse their power
    # Keyword validation: 'abuse'
    leaders_abuse_power = safe_reverse_4pt(q109,
      validate_keyword = "abuse", question_text = question_text,
                                     var_name = "q109"),

    # q111: Legislature capable of holding government accountable
    # Keyword validation: 'in check'
    govt_accountable_leg = safe_reverse_4pt(q111,
                                     validate_keyword = "in check",
                                     question_text = question_text,
                                     var_name = "q111"),

    # q112: How well do you think the government responds to what people want? 
    # Keyword validation: 'responds'
    govt_responsive = safe_reverse_4pt(q112,
      validate_keyword = "responds", question_text = question_text,
                                     var_name = "q112"),

    # q135 People like me don't have any influence over what the government does. 
    # Keyword validation: 'influence over'
    political_efficacy = safe_reverse_4pt(q135,
    validate_keyword = "influence over", question_text = question_text,
                                     var_name = "q135")             
)

cat("=== VERIFYING VARIOUS CORRUPTION SCALES ===\n")
verify_reversal(ab_data$q101, ab_selected$govt_accountable_people)
verify_reversal(ab_data$q103, ab_selected$ethnic_groups_equal)
verify_reversal(ab_data$q104, ab_selected$rich_and_poor_equal)
verify_reversal(ab_data$q108, ab_selected$govt_withholds_info)
verify_reversal(ab_data$q109, ab_selected$leaders_abuse_power)
verify_reversal(ab_data$q111, ab_selected$govt_accountable_leg)
verify_reversal(ab_data$q112, ab_selected$govt_responsive)
verify_reversal(ab_data$q135, ab_selected$political_efficacy)
```

## Recode Regime Preference Variables

```{r 16-recode-regime-preference}
cat("\n========================================\n")
cat("SECTION 3: REGIME PREFERENCE VARIABLES")
cat("\n========================================\n\n")

# Recode authoritarianism support (q129-q132)
# Q129 We should get rid of parliament and elections and have a strong leader to decide things.
# Q130 Only one political party is allowed to stand for election and hold office.
# Q131 The army (military) should come in to govern the country. 
# Q132 We should get rid of elections and parliaments and have experts make decisions on behalf of the people. 
# Ensure higher = more support for authoritarianism
# Original: 1=Strongly agree, 4=Strongly disagree
# Recode so: 4=Strongly agree, 1=Strongly disagree

ab_selected <- ab_selected %>%
  mutate(
    across(q129:q132,
      safe_reverse_4pt,
      .names = "regimepref_{.col}")
  )

cat("=== VERIFYING 4-POINT AUTHORITARIAN SUPPORT SCALES ===\n")
verify_reversal(ab_data$q129, ab_selected$regimepref_q129)
verify_reversal(ab_data$q130, ab_selected$regimepref_q130)
verify_reversal(ab_data$q131, ab_selected$regimepref_q131)
verify_reversal(ab_data$q132, ab_selected$regimepref_q132)

# ============================================================================
# REGIME PREFERENCE VARIABLES - ANALYSIS AND COMPOSITE CREATION
# ============================================================================

# Check distributions
cat("\n=== DISTRIBUTION OF RECODED VARIABLES ===\n")
ab_selected %>%
  select(regimepref_q129, regimepref_q130, regimepref_q131, regimepref_q132) %>%
  summary() %>%
  print()

# Check by country
cat("\n=== DISTRIBUTION BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    regimepref_q129_mean = mean(regimepref_q129, na.rm=TRUE),
    regimepref_q130_mean = mean(regimepref_q130, na.rm=TRUE),
    regimepref_q131_mean = mean(regimepref_q131, na.rm=TRUE),
    regimepref_q132_mean = mean(regimepref_q132, na.rm=TRUE),
    n = n()
  ) %>%
  print()

# --- Reliability analysis ---
cat("\n=== RELIABILITY ANALYSIS ===\n")

library(psych)

# Create item matrix for alpha calculation
regime_items <- ab_selected %>%
  select(regimepref_q129, regimepref_q130, regimepref_q131, regimepref_q132) %>%
  na.omit()

cat("N with complete data:", nrow(regime_items), "\n\n")

# Calculate Cronbach's alpha
alpha_regime <- alpha(regime_items, check.keys = TRUE)
print(alpha_regime)

# --- Create regime preference composite ---
ab_selected <- ab_selected %>%
  mutate(
    # Composite: Mean of all four items
    regime_preference = rowMeans(
      select(., regimepref_q129, regimepref_q130, regimepref_q131, regimepref_q132),
      na.rm = TRUE
    )
  )

# --- Descriptive statistics ---
cat("\n=== DESCRIPTIVE STATISTICS FOR COMPOSITE ===\n")
ab_selected %>%
  select(regime_preference) %>%
  summary() %>%
  print()

cat("\nMissing data:", sum(is.na(ab_selected$regime_preference)), "\n")

# By country
cat("\n=== REGIME PREFERENCE BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>% # <-- This is the fix
  summarise(
    mean = mean(regime_preference, na.rm = TRUE),
    sd = sd(regime_preference, na.rm = TRUE),
    median = median(regime_preference, na.rm = TRUE),
    n = sum(!is.na(regime_preference))
  ) %>%
  print()

# --- Inter-item correlations ---
cat("\n=== INTER-ITEM CORRELATIONS ===\n")
cor(regime_items, use="complete.obs") %>%
  round(3) %>%
  print()

# --- Final summary ---
cat("\n=== REGIME PREFERENCE SCALE READY ===\n")
cat("Variable created: regime_preference\n")
cat("Components:\n")
cat("  - regimepref_q129: Strong leader, no parliament/elections\n")
cat("  - regimepref_q130: One-party system\n")
cat("  - regimepref_q131: Military rule\n")
cat("  - regimepref_q132: Expert rule, no elections\n")
cat("\nInterpretation: Higher scores = Greater preference for authoritarian regimes\n")
```

## Recode Authoritarian Acceptance Variables

```{r 17-recode-authoritarian-acceptance-variables}
cat("\n========================================\n")
cat("SECTION 3: AUTHORITARIAN ACCEPTANCE VARIABLES")
cat("\n========================================\n\n")
# AUTHORITARIAN ACCEPTANCE VARIABLES - RECODING AND COMPOSITE CREATION

# --- Step 1: Recode 4-point authoritarian acceptance scales ---
# q168: Leader can govern by decree and disregard parliament
# q169: Economic performance > regular elections
# q170: Country needs rule-breaking leader
# q171: Order/stability > democracy
# Original: 1=Strongly agree, 4=Strongly disagree
# Recode so: 4=Strongly agree (more authoritarian acceptance), 1=Strongly disagree

ab_selected <- ab_selected %>%
  mutate(
    across(c(q168, q169, q170, q171),
           safe_reverse_4pt,
           .names = "auth_{.col}")
  )

# Verify reversals
cat("=== VERIFYING 4-POINT AUTHORITARIAN ACCEPTANCE SCALES ===\n")
verify_reversal(ab_data$q168, ab_selected$auth_q168)
verify_reversal(ab_data$q169, ab_selected$auth_q169)
verify_reversal(ab_data$q170, ab_selected$auth_q170)
verify_reversal(ab_data$q171, ab_selected$auth_q171)

# Check distributions
cat("\n=== DISTRIBUTION OF RECODED VARIABLES ===\n")
ab_selected %>%
  select(auth_q168, auth_q169, auth_q170, auth_q171) %>%
  summary() %>%
  print()

invalid_check <- verify_no_invalid_codes(ab_selected, 
                        paste0("auth_q", 168:171),
                        c(1,4))

if(nrow(invalid_check) > 0) {
  cat("⚠ WARNING: Invalid codes found!\n")
  print(invalid_check)
} else {
  cat("✓ All authoritarian acceptance variables valid (1-4)\n")
}

# Check by country_name
cat("\n=== DISTRIBUTION BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    auth_q168_mean = mean(auth_q168, na.rm=TRUE),
    auth_q169_mean = mean(auth_q169, na.rm=TRUE),
    auth_q170_mean = mean(auth_q170, na.rm=TRUE),
    auth_q171_mean = mean(auth_q171, na.rm=TRUE),
    n = n()
  ) %>%
  print()

# --- Step 2: Reliability analysis ---
cat("\n=== RELIABILITY ANALYSIS ===\n")

library(psych)

# Create item matrix for alpha calculation
auth_items <- ab_selected %>%
  select(auth_q168, auth_q169, auth_q170, auth_q171) %>%
  na.omit()

cat("N with complete data:", nrow(auth_items), "\n\n")

# Calculate Cronbach's alpha
alpha_auth <- alpha(auth_items, check.keys = TRUE)
print(alpha_auth)

# --- Step 3: Create authoritarian acceptance composite ---
ab_selected <- ab_selected %>%
  mutate(
    # Composite: Mean of all four items
    auth_acceptance = rowMeans(
      select(., auth_q168, auth_q169, auth_q170, auth_q171),
      na.rm = TRUE
    )
  )

# --- Step 4: Descriptive statistics ---
cat("\n=== DESCRIPTIVE STATISTICS FOR COMPOSITE ===\n")
ab_selected %>%
  select(auth_acceptance) %>%
  summary() %>%
  print()

cat("\nMissing data:", sum(is.na(ab_selected$auth_acceptance)), "\n")

# By country_name
cat("\n=== AUTHORITARIAN ACCEPTANCE BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    mean = mean(auth_acceptance, na.rm=TRUE),
    sd = sd(auth_acceptance, na.rm=TRUE),
    median = median(auth_acceptance, na.rm=TRUE),
    n = sum(!is.na(auth_acceptance))
  ) %>%
  print()

# --- Step 5: Inter-item correlations ---
cat("\n=== INTER-ITEM CORRELATIONS ===\n")
cor(auth_items, use="complete.obs") %>%
  round(3) %>%
  print()

# --- Final summary ---
cat("\n=== AUTHORITARIAN ACCEPTANCE SCALE READY ===\n")
cat("Variable created: auth_acceptance\n")
cat("Components:\n")
cat("  - auth_q168: Decree power vs parliament\n")
cat("  - auth_q169: Performance vs elections\n")
cat("  - auth_q170: Rule-breaking leader\n")
cat("  - auth_q171: Order/stability vs democracy\n")
cat("\nInterpretation: Higher scores = Greater acceptance of authoritarian governance\n")
```

## Compare the Two Authoritarian Constructs

```{r 18-comparison-authoritarian-constructs}
cat("\n========================================\n")
cat("SECTION 3: AUTHORITARIAN CONSTRUCTS COMPARISON")
cat("\n========================================\n\n")

cat("\n\n=== CORRELATION BETWEEN AUTH_ACCEPTANCE AND REGIME_PREFERENCE ===\n")

# Correlation test
cor_test_auth <- cor.test(ab_selected$auth_acceptance, 
                          ab_selected$regime_preference, 
                          use = "complete.obs")
print(cor_test_auth)

# Interpretation
if (abs(cor_test_auth$estimate) < 0.4) {
  cat("\n✓ Low-moderate correlation (r =", round(cor_test_auth$estimate, 3), 
      ") - These are DISTINCT constructs.\n")
  cat("Recommendation: Use both as separate DVs or compare them.\n")
} else if (abs(cor_test_auth$estimate) < 0.7) {
  cat("\n→ Moderate-strong correlation (r =", round(cor_test_auth$estimate, 3), 
      ") - Related but distinguishable constructs.\n")
  cat("Recommendation: These tap related dimensions - consider both.\n")
} else {
  cat("\n⚠ Very high correlation (r =", round(cor_test_auth$estimate, 3), 
      ") - May be measuring same underlying construct.\n")
  cat("Recommendation: Could combine into single scale or use one as main DV.\n")
}

# Descriptive comparison
cat("\n=== DESCRIPTIVE COMPARISON ===\n")
ab_selected %>%
  select(auth_acceptance, regime_preference) %>%
  summary() %>%
  print()

# Cross-tabulation: Who scores high on one but not the other?
cat("\n=== CROSS-CLASSIFICATION ===\n")
ab_selected <- ab_selected %>%
  mutate(
    auth_accept_high = ifelse(auth_acceptance >= median(auth_acceptance, na.rm=TRUE), 
                              "High", "Low"),
    regime_pref_high = ifelse(regime_preference >= median(regime_preference, na.rm=TRUE), 
                              "High", "Low")
  )

table(ab_selected$auth_accept_high, ab_selected$regime_pref_high) %>%
  addmargins() %>%
  print()

cat("\nInterpretation:\n")
cat("- High Auth Accept + Low Regime Pref = 'Soft authoritarians' (accept practices, not regimes)\n")
cat("- Low Auth Accept + High Regime Pref = 'Ideological authoritarians' (prefer regimes conceptually)\n")
cat("- High on both = 'Strong authoritarians'\n")
cat("- Low on both = 'Committed democrats'\n")

```

# Scatterplot of the Two Constructs

```{r 19-scatterplot-authoritarian-constructs}
cat("\n=== SCATTERPLOT OF AUTHORITARIAN CONSTRUCTS ===\n")

library(ggplot2)

ggplot(ab_selected, aes(x = auth_acceptance, y = regime_preference)) + 
  geom_point(aes(color = haven::as_factor(country_name)), alpha = 0.3) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Relationship Between Two Forms of Authoritarianism",
    x = "Authoritarian Acceptance (Practices)",
    y = "Regime Preference (Anti-democracy)",
    color = "Country"
  ) +
  theme_minimal()
```

## Recode Emegency Powers Acceptance Variables

```{r 20-recode-emergency-powers}
# ============================================================================
# EMERGENCY POWERS - COMPLETE ANALYSIS
# ============================================================================

# Recode emergency powers (q172a-e)
# Q172 Which circumstance would justify the governments's use of emergency powers to constrain rights and freedom: 
# Q172a public health crisis like the Covid-19 pandemic
# Q172b An economic crisis that has caused the loss of many jobs
# Q172c Widespread corruption that the president [PM] claims can only be reduced by increasing executive power
# Q172d A security crisis due to social unrest or terrorism
# Q172e The country is at war

# Ensure higher = more support for emergency powers
# Original: 1=Very justified, 4=Not at all justified
# Recode so: 4=Not at all justified, 4=Very justified

ab_selected <- ab_selected %>%
  mutate(
    across(q172a:q172e,
      safe_reverse_4pt,
      .names = "emergency_{.col}")
  )

cat("=== VERIFYING 4-POINT EMERGENCY POWERS SCALES ===\n")
verify_reversal(ab_data$q172a, ab_selected$emergency_q172a)
verify_reversal(ab_data$q172b, ab_selected$emergency_q172b)
verify_reversal(ab_data$q172c, ab_selected$emergency_q172c)
verify_reversal(ab_data$q172d, ab_selected$emergency_q172d)
verify_reversal(ab_data$q172e, ab_selected$emergency_q172e)

# Create composite
ab_selected <- ab_selected %>%
  mutate(
    emergency_powers_support = rowMeans(
      select(., emergency_q172a:emergency_q172e),
      na.rm = TRUE
    )
  )

cat("\n=== EMERGENCY POWERS RELIABILITY ===\n")

library(psych)

emergency_items <- ab_selected %>%
  select(emergency_q172a, emergency_q172b, emergency_q172c, 
         emergency_q172d, emergency_q172e) %>%
  na.omit()

cat("N with complete data:", nrow(emergency_items), "\n\n")

alpha_emergency <- alpha(emergency_items, check.keys = TRUE)
print(alpha_emergency)

# Descriptive stats
cat("\n=== EMERGENCY POWERS BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    covid_health_mean = mean(emergency_q172a, na.rm=TRUE),
    economic_mean = mean(emergency_q172b, na.rm=TRUE),
    corruption_mean = mean(emergency_q172c, na.rm=TRUE),
    security_mean = mean(emergency_q172d, na.rm=TRUE),
    war_mean = mean(emergency_q172e, na.rm=TRUE),
    composite_mean = mean(emergency_powers_support, na.rm=TRUE),
    n = sum(!is.na(emergency_powers_support))
  ) %>%
  print()

cat("\n=== EMERGENCY POWERS INTER-ITEM CORRELATIONS ===\n")
cor(emergency_items, use="complete.obs") %>%
  round(3) %>%
  print()
```

## Recode COVID Variables and Create Composites

```{r 21-covid-complete-version}
cat("\n========================================\n")
cat("SECTION 3: COVID VARIABLES")
cat("\n========================================\n\n")
# COVID VARIABLES - RECODING AND COMPOSITE CREATION

# --- Step 1: Check response rates by country ---
cat("=== COVID VARIABLES RESPONSE RATES BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    q138_valid = sum(q138 %in% 1:2),
    q140_valid = sum(q140 %in% 1:4),
    q141_valid = sum(q141 %in% 1:4),
    q142_valid = sum(q142 %in% 1:4),
    q143a_valid = sum(q143a %in% 1:3),
    n_total = n()
  ) %>%
  print()

# --- Step 2: Personal COVID Impact (Binary Variables) ---
ab_selected <- ab_selected %>%
  mutate(
    # Contracted COVID (1=Yes, 0=No)
    covid_contracted = case_when(
      q138 == 1 ~ 1,
      q138 == 2 ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Serious illness/death (1=Yes, 0=No)
    covid_illness_death = case_when(
      q139a == 1 ~ 1,
      q139a == 2 ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Job loss (1=Yes, 0=No)
    covid_job_loss = case_when(
      q139b == 1 ~ 1,
      q139b == 2 ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Income loss (1=Yes, 0=No)
    covid_income_loss = case_when(
      q139c == 1 ~ 1,
      q139c == 2 ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Education disruption (1=Yes, 0=No)
    covid_edu_disruption = case_when(
      q139d == 1 ~ 1,
      q139d == 2 ~ 0,
      TRUE ~ NA_real_
    ),
    
    # Cumulative impact index (0-5 impacts)
    covid_impact_count = rowSums(
      cbind(covid_contracted, covid_illness_death, covid_job_loss, 
            covid_income_loss, covid_edu_disruption),
      na.rm = FALSE
    )
  )

cat("\n=== PERSONAL COVID IMPACT FREQUENCIES ===\n")
ab_selected %>%
  summarise(
    contracted = sum(covid_contracted == 1, na.rm=TRUE),
    illness_death = sum(covid_illness_death == 1, na.rm=TRUE),
    job_loss = sum(covid_job_loss == 1, na.rm=TRUE),
    income_loss = sum(covid_income_loss == 1, na.rm=TRUE),
    edu_disruption = sum(covid_edu_disruption == 1, na.rm=TRUE)
  ) %>%
  print()

# --- Step 3: COVID Impact Severity ---
# Use safe_reverse_4pt with custom missing codes
ab_selected <- ab_selected %>%
  mutate(
    covid_impact_severity = safe_reverse_4pt(q140, missing_codes = c(-1, 0, 7, 8, 9, 10))
  )

cat("\n=== COVID IMPACT SEVERITY ===\n")
table(ab_selected$covid_impact_severity, useNA = "ifany") %>% print()

# --- Step 4: Government COVID Performance ---
ab_selected <- ab_selected %>%
  mutate(
    # Trust in govt COVID info - use safe_reverse_4pt
    covid_trust_info = safe_reverse_4pt(q141, missing_codes = c(-1, 0, 5, 7, 8, 9)),
    
    # Government pandemic handling - use safe_reverse_4pt
    covid_govt_handling = safe_reverse_4pt(q142, missing_codes = c(-1, 0, 97, 7, 8, 9)),
    
    # Composite: COVID government performance
    covid_govt_performance = rowMeans(
      cbind(covid_trust_info, covid_govt_handling),
      na.rm = TRUE
    )
  )

cat("\n=== COVID GOVERNMENT PERFORMANCE ===\n")
cat("Trust in info:\n")
table(ab_selected$covid_trust_info, useNA = "ifany") %>% print()
cat("\nGovt handling:\n")
table(ab_selected$covid_govt_handling, useNA = "ifany") %>% print()

# Check reliability
covid_perf_items <- ab_selected %>%
  select(covid_trust_info, covid_govt_handling) %>%
  na.omit()

cat("\nN with complete COVID performance data:", nrow(covid_perf_items), "\n")
r_covid_perf <- cor(ab_selected$covid_trust_info, ab_selected$covid_govt_handling, 
                    use = "complete.obs")
sb_covid_perf <- (2 * r_covid_perf) / (1 + r_covid_perf)
cat("Inter-item correlation:", round(r_covid_perf, 3), "\n")
cat("Spearman-Brown reliability:", round(sb_covid_perf, 3), "\n")

# --- Step 5: COVID-Justified Restrictions ---
# Use safe_reverse_3pt
ab_selected <- ab_selected %>%
  mutate(
    across(q143a:q143e,
           ~safe_reverse_3pt(.x),
           .names = "covid_restrict_{.col}")
  )

# Rename for clarity
ab_selected <- ab_selected %>%
  rename(
    covid_restrict_elections = covid_restrict_q143a,
    covid_restrict_speech = covid_restrict_q143b,
    covid_restrict_media = covid_restrict_q143c,
    covid_restrict_tracking = covid_restrict_q143d,
    covid_restrict_lockdown = covid_restrict_q143e
  )

cat("\n=== COVID RESTRICTION ACCEPTANCE ===\n")
ab_selected %>%
  select(starts_with("covid_restrict_")) %>%
  summary() %>%
  print()

# Check availability by country
cat("\n=== COVID RESTRICTIONS AVAILABILITY BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    n_valid = sum(!is.na(covid_restrict_elections)),
    n_total = n(),
    pct_valid = round(100 * n_valid / n_total, 1)
  ) %>%
  print()

# --- Reliability check for COVID restrictions ---
covid_restrict_items <- ab_selected %>%
  select(covid_restrict_elections, covid_restrict_speech, covid_restrict_media,
         covid_restrict_tracking, covid_restrict_lockdown) %>%
  na.omit()

cat("\n=== COVID RESTRICTIONS RELIABILITY ===\n")
cat("N with complete data:", nrow(covid_restrict_items), "\n\n")

library(psych)
alpha_covid_restrict <- alpha(covid_restrict_items, check.keys = TRUE)
print(alpha_covid_restrict)

# Create composite
ab_selected <- ab_selected %>%
  mutate(
    covid_restrict_composite = rowMeans(
      select(., covid_restrict_elections, covid_restrict_speech, 
             covid_restrict_media, covid_restrict_tracking, covid_restrict_lockdown),
      na.rm = TRUE
    )
  )

# --- Step 6: Descriptive statistics by country ---
cat("\n=== COVID VARIABLES BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    impact_count_mean = mean(covid_impact_count, na.rm=TRUE),
    impact_severity_mean = mean(covid_impact_severity, na.rm=TRUE),
    govt_performance_mean = mean(covid_govt_performance, na.rm=TRUE),
    restrict_composite_mean = mean(covid_restrict_composite, na.rm=TRUE),
    n = n()
  ) %>%
  print()

# --- Final Summary ---
cat("\n=== COVID CODING COMPLETE ===\n")
cat("Personal Impact:\n")
cat("  - covid_contracted, covid_illness_death, covid_job_loss,\n")
cat("    covid_income_loss, covid_edu_disruption (all binary)\n")
cat("  - covid_impact_count (0-5 cumulative)\n")
cat("  - covid_impact_severity (1-4, higher = more severe)\n")
cat("\nGovernment Performance:\n")
cat("  - covid_trust_info (1-4, higher = more trust)\n")
cat("  - covid_govt_handling (1-4, higher = better)\n")
cat("  - covid_govt_performance (composite, 1-4 scale)\n")
cat("\nCOVID Restrictions:\n")
cat("  - covid_restrict_composite (1-3, higher = more acceptance)\n")
cat("  - Note: Vietnam did not ask q143 questions\n")
```

```{r 22-democratic-tradeoffs}
# ============================================================================
# DEMOCRATIC TRADE-OFFS - q126 & q127
# ============================================================================

# --- Step 1: Check response rates ---
cat("=== TRADE-OFF QUESTIONS RESPONSE RATES BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    q126_valid = sum(q126 %in% 1:5),
    q126_missing = sum(!q126 %in% 1:5),
    q127_valid = sum(q127 %in% 1:5),
    q127_missing = sum(!q127 %in% 1:5),
    n_total = n()
  ) %>%
  print()

# --- Step 2: Recode for different analytical uses ---
ab_selected <- ab_selected %>%
  mutate(
    # === Q126: Democracy vs Economic Development ===
    
    # Version 1: Linear scale (1=econ dev, 4=democracy), excluding "both equal"
    tradeoff_dem_econ_linear = case_when(
      q126 %in% 1:4 ~ as.numeric(q126),
      q126 == 5 ~ NA_real_,  # Exclude "both equal"
      TRUE ~ NA_real_
    ),
    
    # Version 2: Include "both equal" as midpoint (1-5 scale)
    tradeoff_dem_econ_full = case_when(
      q126 %in% 1:5 ~ as.numeric(q126),
      TRUE ~ NA_real_
    ),
    
    # Version 3: Binary - Economic development priority (1=yes, 0=no)
    prioritize_econ_dev = case_when(
      q126 %in% 1:2 ~ 1,  # Econ dev more important
      q126 %in% 3:4 ~ 0,  # Democracy more important
      q126 == 5 ~ NA_real_,  # Exclude "both equal"
      TRUE ~ NA_real_
    ),
    
    # Version 4: Three-way categorical
    tradeoff_dem_econ_cat = case_when(
      q126 %in% 1:2 ~ "Economic Development",
      q126 %in% 3:4 ~ "Democracy",
      q126 == 5 ~ "Both Equal",
      TRUE ~ NA_character_
    ),
    
    # === Q127: Equality vs Freedom ===
    
    # Version 1: Linear scale (1=equality, 4=freedom), excluding "both equal"
    tradeoff_equal_freedom_linear = case_when(
      q127 %in% 1:4 ~ as.numeric(q127),
      q127 == 5 ~ NA_real_,
      TRUE ~ NA_real_
    ),
    
    # Version 2: Include "both equal" as midpoint (1-5 scale)
    tradeoff_equal_freedom_full = case_when(
      q127 %in% 1:5 ~ as.numeric(q127),
      TRUE ~ NA_real_
    ),
    
    # Version 3: Binary - Equality priority (1=yes, 0=no)
    prioritize_equality = case_when(
      q127 %in% 1:2 ~ 1,  # Equality more important
      q127 %in% 3:4 ~ 0,  # Freedom more important
      q127 == 5 ~ NA_real_,
      TRUE ~ NA_real_
    ),
    
    # Version 4: Three-way categorical
    tradeoff_equal_freedom_cat = case_when(
      q127 %in% 1:2 ~ "Economic Equality",
      q127 %in% 3:4 ~ "Political Freedom",
      q127 == 5 ~ "Both Equal",
      TRUE ~ NA_character_
    )
  )

# --- Step 3: Descriptive statistics by country ---
cat("\n=== Q126: DEMOCRACY VS ECONOMIC DEVELOPMENT BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    econ_dev_definitely = sum(q126 == 1, na.rm=TRUE),
    econ_dev_somewhat = sum(q126 == 2, na.rm=TRUE),
    democracy_somewhat = sum(q126 == 3, na.rm=TRUE),
    democracy_definitely = sum(q126 == 4, na.rm=TRUE),
    both_equal = sum(q126 == 5, na.rm=TRUE),
    pct_prioritize_econ = mean(prioritize_econ_dev == 1, na.rm=TRUE) * 100,
    n = n()
  ) %>%
  print()

cat("\n=== Q127: EQUALITY VS FREEDOM BY COUNTRY ===\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    equality_definitely = sum(q127 == 1, na.rm=TRUE),
    equality_somewhat = sum(q127 == 2, na.rm=TRUE),
    freedom_somewhat = sum(q127 == 3, na.rm=TRUE),
    freedom_definitely = sum(q127 == 4, na.rm=TRUE),
    both_equal = sum(q127 == 5, na.rm=TRUE),
    pct_prioritize_equality = mean(prioritize_equality == 1, na.rm=TRUE) * 100,
    n = n()
  ) %>%
  print()

# --- Step 4: Cross-tabs with categorical versions ---
cat("\n=== Q126 DISTRIBUTION (ALL COUNTRIES) ===\n")
table(ab_selected$tradeoff_dem_econ_cat, useNA = "ifany") %>%
  addmargins() %>%
  print()

cat("\n=== Q127 DISTRIBUTION (ALL COUNTRIES) ===\n")
table(ab_selected$tradeoff_equal_freedom_cat, useNA = "ifany") %>%
  addmargins() %>%
  print()

# --- Step 5: Validation - Correlation with auth_acceptance ---
cat("\n=== VALIDATION: CORRELATION WITH AUTHORITARIAN ACCEPTANCE ===\n")

# Q126: Those who prioritize econ dev should have higher auth_acceptance
cor_q126 <- cor.test(ab_selected$tradeoff_dem_econ_linear, 
                     ab_selected$auth_acceptance,
                     use = "complete.obs")
cat("\nQ126 (Econ Dev > Democracy) vs Auth Acceptance:\n")
print(cor_q126)

# Q127: Those who prioritize equality should have higher auth_acceptance?
cor_q127 <- cor.test(ab_selected$tradeoff_equal_freedom_linear,
                     ab_selected$auth_acceptance,
                     use = "complete.obs")
cat("\nQ127 (Equality > Freedom) vs Auth Acceptance:\n")
print(cor_q127)

# Cross-tab: Do those who prioritize econ dev also accept auth practices?
cat("\n=== CROSS-TAB: ECONOMIC DEV PRIORITY × AUTH ACCEPTANCE ===\n")
ab_selected <- ab_selected %>%
  mutate(
    auth_accept_high = ifelse(auth_acceptance >= median(auth_acceptance, na.rm=TRUE),
                              "High Auth Accept", "Low Auth Accept")
  )

table(ab_selected$prioritize_econ_dev, ab_selected$auth_accept_high) %>%
  addmargins() %>%
  print()

# --- Step 6: Correlation between the two trade-offs ---
cat("\n=== CORRELATION BETWEEN TWO TRADE-OFFS ===\n")
cor_tradeoffs <- cor.test(ab_selected$tradeoff_dem_econ_linear,
                          ab_selected$tradeoff_equal_freedom_linear,
                          use = "complete.obs")
print(cor_tradeoffs)

# --- Final summary ---
cat("\n=== TRADE-OFF VARIABLES READY ===\n")
cat("Q126 Variables Created:\n")
cat("  - tradeoff_dem_econ_linear (1-4, excluding 'both equal')\n")
cat("  - tradeoff_dem_econ_full (1-5, including 'both equal')\n")
cat("  - prioritize_econ_dev (binary: 1=econ, 0=democracy)\n")
cat("  - tradeoff_dem_econ_cat (categorical)\n")
cat("\nQ127 Variables Created:\n")
cat("  - tradeoff_equal_freedom_linear (1-4, excluding 'both equal')\n")
cat("  - tradeoff_equal_freedom_full (1-5, including 'both equal')\n")
cat("  - prioritize_equality (binary: 1=equality, 0=freedom)\n")
cat("  - tradeoff_equal_freedom_cat (categorical)\n")
cat("\nInterpretation:\n")
cat("  - Higher values = prioritize democracy/freedom\n")
cat("  - Lower values = prioritize economic development/equality\n")
cat("\nRecommended use:\n")
cat("  1. Descriptive context (% who prioritize econ dev)\n")
cat("  2. Construct validation (correlation with auth_acceptance)\n")
cat("  3. Robustness check (alternative operationalization)\n")
```

```{r 23-convert-nan-to-na}
# ============================================================================
# CONVERT NaN TO NA (CLEANUP) - MASS CONVERSION
# ============================================================================

cat("\n=== CONVERTING NaN TO NA ===\n")

# Check for NaN values before conversion
nan_check_before <- ab_selected %>%
  summarise(across(where(is.numeric), ~sum(is.nan(.)), .names = "{col}")) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_nan") %>%
  filter(n_nan > 0) %>%
  arrange(desc(n_nan))

if(nrow(nan_check_before) > 0) {
  cat("Found NaN values in", nrow(nan_check_before), "variables:\n")
  print(nan_check_before, n = 20)
  cat("\nTotal NaN cases:", sum(nan_check_before$n_nan), "\n")
} else {
  cat("No NaN values found\n")
}

# MASS CONVERT: All NaN to NA across entire dataset
ab_selected <- ab_selected %>%
  mutate(across(where(is.numeric), ~ifelse(is.nan(.), NA_real_, .)))

# Verify conversion
nan_check_after <- ab_selected %>%
  summarise(across(where(is.numeric), ~sum(is.nan(.)), .names = "{col}")) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_nan") %>%
  filter(n_nan > 0)

if(nrow(nan_check_after) > 0) {
  cat("\n⚠ WARNING: Still have NaN values after conversion!\n")
  print(nan_check_after)
} else {
  cat("\n✓ All NaN values successfully converted to NA\n")
}
```

## Summarize Missing Data

```{r 24-missing-data-summary}
# ============================================================================
# MISSING DATA SUMMARY
# ============================================================================

cat("\n\n=== Missing Data Overview ===\n")
cat("Total observations:", nrow(ab_selected), "\n\n")

# 1. DEFINE YOUR ID VARS ONCE
id_vars <- c("country", "country_name", "cambodia", "vietnam", "thailand", "idnumber")

# 2. Define variables to EXCLUDE from missing data reports
derived_vars <- c(
  # ID variables
  id_vars,
  
  # Recoded Demographics
  "gender", "age", "age_group", "urban", "educ_level", "educ_years",
  
  # Recoded Political Engagement
  "interest_politics", "follow_news", "discuss_politics",

  # Recoded Trust
  paste0("trust_q", 7:15), "institutional_trust_index",

  # Recoded Democracy
  "dem_q90", "dem_q91", "dem_q128", "q92_clean", "q95_clean",
  "z_dem_q90", "z_dem_q91", "z_dem_q128", "z_dem_q92", "z_dem_q95",
  "dem_satisfaction", "dem_legitimacy",
  
  # Recoded Authoritarianism
  paste0("regimepref_q", 129:132), "regime_preference",
  paste0("auth_q", 168:171), "auth_acceptance",
  "auth_accept_high", "regime_pref_high",
  
  # Recoded Emergency Powers
  paste0("emergency_q", letters[1:5]), "emergency_powers_support",

  # Recoded Trade-offs
  "tradeoff_dem_econ_linear", "tradeoff_dem_econ_full", "prioritize_econ_dev",   "tradeoff_dem_econ_cat",
  "tradeoff_equal_freedom_linear", "tradeoff_equal_freedom_full",
  "prioritize_equality", "tradeoff_equal_freedom_cat",
  
  # Recoded COVID
  "covid_contracted", "covid_illness_death", "covid_job_loss",
  "covid_income_loss", "covid_edu_disruption", "covid_impact_count",
  "covid_impact_severity", "covid_trust_info", "covid_govt_handling",
  "covid_govt_performance",
  "covid_restrict_elections", "covid_restrict_speech", "covid_restrict_media",
  "covid_restrict_tracking", "covid_restrict_lockdown",
  "covid_restrict_composite"

  # Recoded Corruption
   "vote_buying", "corruption_crackdown",
   "corruption_intimate_knowledge", "corruption_personally_witnessed",
   "corruption_family_witnessed", "corruption_friend_witnessed"
)

# 3. Calculate missing data excluding derived variables
missing_summary <- ab_selected %>%
  summarise(across(-any_of(derived_vars),
                   ~sum(is.na(.)) / n() * 100,
                   .names = "{col}_pct_missing")) %>%
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "pct_missing") %>%
  mutate(variable = str_remove(variable, "_pct_missing")) %>%
  arrange(desc(pct_missing))

print(missing_summary, n = 50)

# 4. Calculate missing data by country
missing_by_country <- ab_selected %>%
  group_by(country_name) %>%
  summarise(
    across(-any_of(derived_vars),
           ~sum(is.na(.)) / n() * 100,
           .names = "{col}_pct_missing")
  ) %>%
  pivot_longer(-country_name,
               names_to = "variable",
               values_to = "pct_missing") %>%
  mutate(variable = str_remove(variable, "_pct_missing"))

# Focus on key missing patterns
cat("\n\nMissing Data for Vietnam q143 series (pandemic-specific powers):\n")
missing_by_country %>%
  filter(variable %in% paste0("q143", letters[1:5])) %>%
  pivot_wider(names_from = country_name, values_from = pct_missing) %>%
  print()
```

## Visualize Missing Data

```{r 25-visualize-missing}
# Create heatmap of missing data by country (BASE R VERSION - Variables double-checked)
missing_plot_data <- missing_by_country %>%
  filter(variable %in% c(
    # Trust in Institutions
    paste0("q", 7:15),
    
    # Democracy Variables
    "q90", "q91", "q92", "q95",      # Core democracy items
    "q126", "q127", "q128",          # Trade-offs and best system
    
    # Authoritarianism Support (regime types)
    paste0("q", 129:132),
    
    # Acceptance of Authoritarian Actions
    paste0("q", 168:171),
    
    # Emergency Powers
    paste0("q172", letters[1:5]),
    
    # COVID Variables
    "q138",                          # COVID infection
    paste0("q139", letters[1:4]),    # Family impacts (a-d)
    "q140",                          # Livelihood impact
    "q141", "q142",                  # Trust and handling
    paste0("q143", letters[1:5])     # Pandemic powers
  ))

# Reshape to wide format (matrix needed for image())
library(tidyr)
missing_matrix <- missing_plot_data %>%
  select(country_name, variable, pct_missing) %>%
  pivot_wider(names_from = country_name, values_from = pct_missing) %>%
  column_to_rownames("variable") %>%
  as.matrix()

# Create color palette (lightblue to darkred)
color_palette <- colorRampPalette(c("lightblue", "darkred"))(100)

# Create the heatmap
png(here("figures", "missing_data_heatmap.png"), width = 10, height = 12, units = "in", res = 300)

par(mar = c(6, 8, 4, 6))  # Margins: bottom, left, top, right

# Plot the heatmap
image(1:ncol(missing_matrix), 
      1:nrow(missing_matrix), 
      t(missing_matrix),  # Transpose for correct orientation
      col = color_palette,
      xlab = "", ylab = "",
      main = "Missing Data Patterns by Country",
      xaxt = "n", yaxt = "n",
      zlim = c(0, 100))

# Add country names (x-axis)
axis(1, at = 1:ncol(missing_matrix), labels = colnames(missing_matrix), 
     las = 2, cex.axis = 0.8)

# Add variable names (y-axis)
axis(2, at = 1:nrow(missing_matrix), labels = rownames(missing_matrix), 
     las = 2, cex.axis = 0.7)

# Add color legend
legend_values <- seq(0, 100, length.out = 11)
legend("right", 
       legend = paste0(legend_values, "%"),
       fill = color_palette[seq(1, 100, length.out = 11)],
       title = "% Missing",
       xpd = TRUE, inset = c(-0.15, 0),
       cex = 0.8)

dev.off()

# Also display on screen
par(mar = c(6, 8, 4, 6))
image(1:ncol(missing_matrix), 
      1:nrow(missing_matrix), 
      t(missing_matrix),
      col = color_palette,
      xlab = "", ylab = "",
      main = "Missing Data Patterns by Country",
      xaxt = "n", yaxt = "n",
      zlim = c(0, 100))
axis(1, at = 1:ncol(missing_matrix), labels = colnames(missing_matrix), 
     las = 2, cex.axis = 0.8)
axis(2, at = 1:nrow(missing_matrix), labels = rownames(missing_matrix), 
     las = 2, cex.axis = 0.7)

# Save a summary table too
missing_summary_table <- missing_by_country %>%
  filter(pct_missing > 10) %>%  # Only show substantial missing
  pivot_wider(names_from = country_name, values_from = pct_missing) %>%
  arrange(desc(Cambodia))

write_csv(missing_summary_table, 
          here("tables", "missing_data_summary.csv"))
```

# Construct Validation

```{r 26-construct-validation}
# ============================================================================
# CONSTRUCT VALIDATION
# ============================================================================

cat("\n=== VALIDATING MEASUREMENT CONSTRUCTS ===\n")

# --- A. Inter-Construct Correlations ---
cat("\n1. INTER-CONSTRUCT CORRELATIONS\n")
cat("   (Checking for multicollinearity and construct validity)\n\n")

key_constructs <- ab_selected %>%
  select(
    institutional_trust_index,
    dem_satisfaction,
    dem_legitimacy,
    regime_preference,
    auth_acceptance,
    emergency_powers_support,
    covid_govt_performance,
    covid_restrict_composite
  )

# Correlation matrix
cor_matrix <- cor(key_constructs, use = "pairwise.complete.obs")

cat("Correlation Matrix:\n")
print(round(cor_matrix, 3))

# Export for paper
write.csv(round(cor_matrix, 3), 
          here("tables", "construct_correlations.csv"))

# Flag problematic correlations
cat("\n--- Correlation Diagnostics ---\n")

# High correlations (potential multicollinearity)
high_cors <- which(abs(cor_matrix) > 0.7 & cor_matrix != 1, arr.ind = TRUE)
if(nrow(high_cors) > 0) {
  cat("\n⚠ High correlations (r > 0.7) - May indicate multicollinearity:\n")
  for(i in 1:nrow(high_cors)) {
    cat(sprintf("  %s <-> %s: r = %.3f\n",
                rownames(cor_matrix)[high_cors[i,1]],
                colnames(cor_matrix)[high_cors[i,2]],
                cor_matrix[high_cors[i,1], high_cors[i,2]]))
  }
  cat("\nRecommendation: Consider combining these or using only one in models.\n")
} else {
  cat("✓ No problematic high correlations (all r < 0.7)\n")
}

# Very low correlations (potential validity issues)
low_cors <- which(abs(cor_matrix) < 0.1 & cor_matrix != 1, arr.ind = TRUE)
if(nrow(low_cors) > 0) {
  cat("\n⚠ Very low correlations (r < 0.1) - May indicate construct issues:\n")
  # Remove duplicate pairs (only show each pair once)
  low_cors_unique <- low_cors[low_cors[,1] < low_cors[,2], , drop = FALSE]
  if(nrow(low_cors_unique) > 0) {
    for(i in 1:nrow(low_cors_unique)) {
      cat(sprintf("  %s <-> %s: r = %.3f\n",
                  rownames(cor_matrix)[low_cors_unique[i,1]],
                  colnames(cor_matrix)[low_cors_unique[i,2]],
                  cor_matrix[low_cors_unique[i,1], low_cors_unique[i,2]]))
    }
  }
} else {
  cat("✓ All constructs show some relationship (r > 0.1)\n")
}

# --- B. Theoretical Expectations ---
cat("\n2. TESTING THEORETICAL EXPECTATIONS\n")

# Should be negatively correlated
cat("\nExpected NEGATIVE correlations:\n")
cat(sprintf("  Democracy Satisfaction <-> Auth Acceptance: r = %.3f %s\n",
            cor_matrix["dem_satisfaction", "auth_acceptance"],
            ifelse(cor_matrix["dem_satisfaction", "auth_acceptance"] < 0, "✓", "⚠")))
cat(sprintf("  Democracy Legitimacy <-> Regime Preference: r = %.3f %s\n",
            cor_matrix["dem_legitimacy", "regime_preference"],
            ifelse(cor_matrix["dem_legitimacy", "regime_preference"] < 0, "✓", "⚠")))

# Should be positively correlated
cat("\nExpected POSITIVE correlations:\n")
cat(sprintf("  Auth Acceptance <-> Regime Preference: r = %.3f %s\n",
            cor_matrix["auth_acceptance", "regime_preference"],
            ifelse(cor_matrix["auth_acceptance", "regime_preference"] > 0, "✓", "⚠")))
cat(sprintf("  Trust <-> COVID Govt Performance: r = %.3f %s\n",
            cor_matrix["institutional_trust_index", "covid_govt_performance"],
            ifelse(cor_matrix["institutional_trust_index", "covid_govt_performance"] > 0, "✓", "⚠")))

# --- C. Distribution Checks ---
cat("\n3. DISTRIBUTION CHECKS\n")
cat("   (Checking for severe skewness that may need transformation)\n\n")

for(var in names(key_constructs)) {
  x <- key_constructs[[var]]
  skewness <- psych::skew(x, na.rm=TRUE)
  kurtosis <- psych::kurtosi(x, na.rm=TRUE)
  
  flag <- ""
  if(abs(skewness) > 2) flag <- "⚠ SEVERE SKEW"
  if(abs(kurtosis) > 7) flag <- paste(flag, "⚠ SEVERE KURTOSIS")
  
  cat(sprintf("  %-30s: Skew = %6.2f, Kurtosis = %6.2f %s\n",
              var, skewness, kurtosis, flag))
}

cat("\nNote: |Skewness| > 2 or |Kurtosis| > 7 may indicate need for transformation\n")

# --- D. Sample Sizes for Key Constructs ---
cat("\n4. SAMPLE SIZES WITH COMPLETE DATA\n")
ab_selected %>%
  summarise(
    across(all_of(names(key_constructs)),
           ~sum(!is.na(.)),
           .names = "{.col}_n")
  ) %>%
  pivot_longer(everything(),
               names_to = "construct",
               values_to = "n_complete") %>%
  mutate(
    construct = str_remove(construct, "_n"),
    pct_complete = round(100 * n_complete / nrow(ab_selected), 1)
  ) %>%
  arrange(desc(n_complete)) %>%
  print()

# --- E. Reliability Summary Table ---
cat("\n5. RELIABILITY SUMMARY\n")

# Create comprehensive reliability table
reliability_summary <- tibble(
  Construct = c(
    "Institutional Trust",
    "Democratic Satisfaction", 
    "Democratic Legitimacy",
    "Regime Preference",
    "Authoritarian Acceptance",
    "Emergency Powers Support",
    "COVID Govt Performance",
    "COVID Restrictions Accept"
  ),
  N_Items = c(9, 2, 2, 4, 4, 5, 2, 5),
  
  # This pulls the 'std.alpha' value from each saved object
  Alpha = c(
    alpha_trust$total$std.alpha,
    sb_satisfaction,  # Spearman-Brown is correct for 2 items
    sb_legitimacy,    # Spearman-Brown is correct for 2 items
    alpha_regime$total$std.alpha,
    alpha_auth$total$std.alpha,
    alpha_emergency$total$std.alpha,
    sb_covid_perf,    # Spearman-Brown is correct for 2 items
    alpha_covid_restrict$total$std.alpha
  ),
  # ===============================================

  Sample_N = c(
    sum(!is.na(ab_selected$institutional_trust_index)),
    sum(!is.na(ab_selected$dem_satisfaction)),
    sum(!is.na(ab_selected$dem_legitimacy)),
    sum(!is.na(ab_selected$regime_preference)),
    sum(!is.na(ab_selected$auth_acceptance)),
    sum(!is.na(ab_selected$emergency_powers_support)),
    sum(!is.na(ab_selected$covid_govt_performance)),
    sum(!is.na(ab_selected$covid_restrict_composite))
  ),
  Mean = round(sapply(key_constructs, mean, na.rm=TRUE), 2),
  SD = round(sapply(key_constructs, sd, na.rm=TRUE), 2)
)

# Round the Alpha column
reliability_summary <- reliability_summary %>%
  mutate(Alpha = round(Alpha, 3))

print(reliability_summary)

# Save for paper
write_csv(reliability_summary, 
          here("tables", "construct_summary.csv"))

cat("\n=== 21 CONSTRUCT VALIDATION COMPLETE ===\n")
cat("All constructs validated. Proceeding to standardization.\n")
```

# Sample Size Flow Chart

```{r 27-sample-size-flow}
# ============================================================================
# SAMPLE SIZE FLOW CHART
# ============================================================================

cat("\n=== SAMPLE SIZE FLOW ===\n")
cat("1. Starting N (all countries):", nrow(ab_whole), "\n")
cat("2. After filtering to Cambodia/Thailand/Vietnam:", nrow(ab_data), "\n")
cat("3. After variable selection:", nrow(ab_selected), "\n\n")

cat("4. Complete data for key constructs:\n")
ab_selected %>%
  summarise(
    complete_trust = sum(!is.na(institutional_trust_index)),
    complete_dem_sat = sum(!is.na(dem_satisfaction)),
    complete_dem_leg = sum(!is.na(dem_legitimacy)),
    complete_auth_accept = sum(!is.na(auth_acceptance)),
    complete_regime_pref = sum(!is.na(regime_preference)),
    complete_emergency = sum(!is.na(emergency_powers_support)),
    complete_covid_perf = sum(!is.na(covid_govt_performance)),
    
    # Most restrictive: all key variables present
    complete_all_key = sum(
      !is.na(institutional_trust_index) & 
      !is.na(dem_satisfaction) &
      !is.na(dem_legitimacy) &
      !is.na(auth_acceptance) &
      !is.na(regime_preference) &
      !is.na(covid_govt_performance)
    )
  ) %>%
  pivot_longer(everything(), names_to = "measure", values_to = "n") %>%
  mutate(
    pct_of_total = round(n / nrow(ab_selected) * 100, 1)
  ) %>%
  print()

cat("\n5. Complete cases by country:\n")
ab_selected %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    complete_all_key = sum(
      !is.na(institutional_trust_index) & 
      !is.na(dem_satisfaction) &
      !is.na(auth_acceptance) &
      !is.na(regime_preference) &
      !is.na(covid_govt_performance)
    ),
    pct_complete = round(complete_all_key / n * 100, 1)
  ) %>%
  print()

cat("\n✓ Sample size tracking complete\n")
```

# Create Composite Indices

# Standardize Variables for Regression

```{r 28-standardize-variables}
# Create standardized (z-score) versions of continuous variables

ab_selected <- ab_selected %>%
  mutate(
    # Standardize indices
    institutional_trust_std = scale(institutional_trust_index)[,1],
    dem_satisfaction_std = scale(dem_satisfaction)[,1],
    dem_legitimacy_std = scale(dem_legitimacy)[,1],
    interest_politics_std = scale(interest_politics)[,1],
    
    regime_preference_std = scale(regime_preference)[,1],
    auth_acceptance_std = scale(auth_acceptance)[,1],
    emergency_powers_support_std = scale(emergency_powers_support)[,1],

    # Standardize COVID variables
    covid_impact_severity_std = scale(covid_impact_severity)[,1],
    covid_trust_info_std = scale(covid_trust_info)[,1],
    covid_govt_handling_std = scale(covid_govt_handling)[,1]
  )

cat("✓ 22: Standardized variables created\n")
```

# Create Analysis Dataset

```{r 29-create-analysis-dataset}
# Select final variables for analysis
ab_analysis <- ab_selected %>%
  select(
    # --- Identifiers ---
    country_name, cambodia, thailand, vietnam,
    matches("^id|^ID|^respondent"), # Your respondent ID

    # --- Demographics ---
    gender, age, age_group, urban, educ_level, educ_years,

    # --- Trust ---
    starts_with("trust_q"),
    institutional_trust_index, institutional_trust_std,

    # --- Democracy ---
    # Recoded items
    dem_q90, dem_q91, dem_q128,
    q92_clean, q95_clean,
    # Composites
    dem_satisfaction, dem_satisfaction_std,
    dem_legitimacy, dem_legitimacy_std,

    # --- Political Engagement ---
    interest_politics, follow_news, discuss_politics,
    
    # --- Authoritarianism (Two Scales) ---
    # Regime Preference
    starts_with("regimepref_q"),
    regime_preference, regime_preference_std,
    # Authoritarian Acceptance
    starts_with("auth_q"),
    auth_acceptance, auth_acceptance_std,

    # --- Emergency Powers ---
    starts_with("emergency_q"),
    emergency_powers_support, emergency_powers_support_std,

    # --- Trade-offs ---
    tradeoff_dem_econ_linear, tradeoff_dem_econ_full, prioritize_econ_dev, tradeoff_dem_econ_cat,
    tradeoff_equal_freedom_linear, tradeoff_equal_freedom_full, prioritize_equality, tradeoff_equal_freedom_cat,

    # --- COVID Variables ---
    # Recoded items
    covid_contracted, covid_illness_death, covid_job_loss,
    covid_income_loss, covid_edu_disruption, covid_impact_count,
    covid_impact_severity, covid_trust_info, covid_govt_handling,
    starts_with("covid_restrict_"),
    # Composites
    covid_govt_performance, covid_restrict_composite,
    
    # --- Standardized COVID items ---
    covid_impact_severity_std, covid_trust_info_std, covid_govt_handling_std
  )

# Final dataset summary
cat("\n=== FINAL ANALYSIS DATASET ===\n")
cat("Dimensions:", nrow(ab_analysis), "rows x", ncol(ab_analysis), "columns\n\n")

cat("Sample sizes by country:\n")
print(table(ab_analysis$country_name))

cat("\nDemographics completeness:\n")
ab_analysis %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    gender_complete = sum(!is.na(gender)),
    age_complete = sum(!is.na(age)),
    educ_complete = sum(!is.na(educ_level)),
    urban_complete = sum(!is.na(urban))
  ) %>%
  print()

cat("\nComplete cases (no missing on key variables):\n")
key_vars <- c("institutional_trust_index", 
              "regime_preference",
              "auth_acceptance",
              "covid_contracted",
              "covid_govt_handling")

ab_analysis %>%
  mutate(complete = complete.cases(select(., all_of(key_vars)))) %>%
  group_by(country_name) %>%
  summarise(
    n_complete = sum(complete),
    pct_complete = round(mean(complete) * 100, 1)
  ) %>%
  print()

cat("\n✓ Analysis dataset created with", ncol(ab_analysis), "variables\n")
```

# Final Data Quality Checks and Save

```{r 30-final-quality-checks}
# ============================================
# FINAL DATA QUALITY CHECKS
# ============================================

# Ensure country_name is character (not factor)
if(is.factor(ab_analysis$country_name)) {
  cat("✓ Converting country_name to character\n")
  ab_analysis$country_name <- as.character(ab_analysis$country_name)
}

# Save Processed Data

# Save as RDS (preserves R data types)
saveRDS(ab_analysis, here("data", "processed", "ab_analysis.rds"))

# Save as CSV (for sharing/other software)
write_csv(ab_analysis, here("data", "processed", "ab_analysis.csv"))

# For Stata users
library(haven)
write_dta(ab_analysis, here("data", "processed", "ab_analysis.dta"))

# ============================================================================
# CREATE SIMPLE CODEBOOK
# ============================================================================

cat("\nCreating codebook...\n")

# Create codebook dataframe
codebook <- tibble(
  variable = names(ab_analysis),
  class = sapply(ab_analysis, function(x) class(x)[1]),
  n_total = nrow(ab_analysis),
  n_missing = sapply(ab_analysis, function(x) sum(is.na(x))),
  pct_missing = round(sapply(ab_analysis, function(x) mean(is.na(x)) * 100), 1),
  n_unique = sapply(ab_analysis, function(x) length(unique(x[!is.na(x)]))),
  min = sapply(ab_analysis, function(x) {
    if(is.numeric(x)) round(min(x, na.rm=TRUE), 2) else NA_real_
  }),
  max = sapply(ab_analysis, function(x) {
    if(is.numeric(x)) round(max(x, na.rm=TRUE), 2) else NA_real_
  }),
  mean = sapply(ab_analysis, function(x) {
    if(is.numeric(x)) round(mean(x, na.rm=TRUE), 2) else NA_real_
  }),
  sd = sapply(ab_analysis, function(x) {
    if(is.numeric(x)) round(sd(x, na.rm=TRUE), 2) else NA_real_
  })
)

# Add variable descriptions
codebook <- codebook %>%
  mutate(
    description = case_when(
      # Identifiers
      variable == "country_name" ~ "Country name (Cambodia, Vietnam, Thailand)",
      variable == "cambodia" ~ "Cambodia dummy (1=Cambodia, 0=other)",
      variable == "vietnam" ~ "Vietnam dummy (1=Vietnam, 0=other)",
      variable == "thailand" ~ "Thailand dummy (1=Thailand, 0=other)",
      
      # Demographics
      variable == "gender" ~ "Gender (Male/Female)",
      variable == "age" ~ "Age in years",
      variable == "age_group" ~ "Age group (18-29, 30-44, 45-59, 60+)",
      variable == "urban" ~ "Urban residence (1=urban, 0=rural)",
      variable == "educ_level" ~ "Education level (Primary/Secondary/Tertiary)",
      variable == "educ_years" ~ "Years of formal education",
      
      # Trust
      str_starts(variable, "trust_q") ~ "Trust in institution (recoded, 1=none to 4=great deal)",
      variable == "institutional_trust_index" ~ "Composite institutional trust (mean of q7-q15)",
      variable == "institutional_trust_std" ~ "Institutional trust (standardized)",
      
      # Democracy
      variable == "dem_q90" ~ "Satisfaction with democracy (1=low to 4=high)",
      variable == "dem_q91" ~ "Level of democracy (1=low to 4=high)",
      variable == "dem_q128" ~ "Democracy best system (1=disagree to 4=agree)",
      variable == "q92_clean" ~ "How democratic is government (1-10 scale)",
      variable == "q95_clean" ~ "Suitability of democracy (1-10 scale)",
      variable == "dem_satisfaction" ~ "Democratic satisfaction composite (z-scores)",
      variable == "dem_satisfaction_std" ~ "Democratic satisfaction (standardized)",
      variable == "dem_legitimacy" ~ "Democratic legitimacy composite (z-scores)",
      variable == "dem_legitimacy_std" ~ "Democratic legitimacy (standardized)",
      
      # Political engagement
      variable == "interest_politics" ~ "Interest in politics (1=low to 4=high)",
      variable == "follow_news" ~ "Follow news (1=never to 5=daily)",
      variable == "discuss_politics" ~ "Discuss politics (1=never to 3=often)",
      
      # Authoritarianism - Regime preference
      str_starts(variable, "regimepref_q") ~ "Regime preference item (1=disagree to 4=agree)",
      variable == "regime_preference" ~ "Authoritarian regime preference (1=low to 4=high)",
      variable == "regime_preference_std" ~ "Regime preference (standardized)",
      
      # Authoritarianism - Acceptance
      str_starts(variable, "auth_q") ~ "Authoritarian acceptance item (1=disagree to 4=agree)",
      variable == "auth_acceptance" ~ "Authoritarian acceptance (1=low to 4=high)",
      variable == "auth_acceptance_std" ~ "Authoritarian acceptance (standardized)",
      
      # Emergency powers
      str_starts(variable, "emergency_q") ~ "Emergency powers item (1=not justified to 4=very justified)",
      variable == "emergency_powers_support" ~ "Emergency powers support (1=low to 4=high)",
      variable == "emergency_powers_support_std" ~ "Emergency powers support (standardized)",
      
      # Trade-offs
      str_starts(variable, "tradeoff_dem_econ") ~ "Democracy vs economic development trade-off",
      str_starts(variable, "prioritize_econ") ~ "Prioritize economic development (1=yes, 0=no)",
      str_starts(variable, "tradeoff_equal_freedom") ~ "Equality vs freedom trade-off",
      str_starts(variable, "prioritize_equality") ~ "Prioritize equality (1=yes, 0=no)",
      
      # COVID
      variable == "covid_contracted" ~ "Contracted COVID-19 (1=yes, 0=no)",
      variable == "covid_illness_death" ~ "Family illness/death from COVID (1=yes, 0=no)",
      variable == "covid_job_loss" ~ "Job loss due to COVID (1=yes, 0=no)",
      variable == "covid_income_loss" ~ "Income loss due to COVID (1=yes, 0=no)",
      variable == "covid_edu_disruption" ~ "Education disruption due to COVID (1=yes, 0=no)",
      variable == "covid_impact_count" ~ "Cumulative COVID impacts (0-5)",
      variable == "covid_impact_severity" ~ "COVID impact severity (1=low to 4=high)",
      variable == "covid_trust_info" ~ "Trust government COVID info (1=low to 4=high)",
      variable == "covid_govt_handling" ~ "Government pandemic handling (1=poor to 4=excellent)",
      variable == "covid_govt_performance" ~ "COVID government performance composite (1=low to 4=high)",
      str_starts(variable, "covid_restrict_") ~ "COVID restriction acceptance (1=low to 3=high)",
      str_ends(variable, "_std") ~ paste0("Standardized version of ", str_remove(variable, "_std")),
      
      TRUE ~ ""
    )
  )

# Reorder columns for readability
codebook <- codebook %>%
  select(variable, description, class, n_total, n_missing, pct_missing, 
         n_unique, min, max, mean, sd)

# Save codebook as CSV
write_csv(codebook, here("docs", "codebook.csv"))
cat("✓ Saved: docs/codebook.csv\n")

# Create HTML version using gt
library(gt)

codebook_gt <- codebook %>%
  gt() %>%
  tab_header(
    title = md("**Asian Barometer Wave 6 Analysis Dataset**"),
    subtitle = "Codebook: Cambodia, Thailand, Vietnam | COVID-Democracy Project"
  ) %>%
  fmt_number(
    columns = c(min, max, mean, sd),
    decimals = 2
  ) %>%
  fmt_number(
    columns = pct_missing,
    decimals = 1
  ) %>%
  cols_label(
    variable = "Variable",
    description = "Description",
    class = "Type",
    n_total = "N Total",
    n_missing = "N Missing",
    pct_missing = "% Missing",
    n_unique = "Unique Values",
    min = "Min",
    max = "Max",
    mean = "Mean",
    sd = "SD"
  ) %>%
  tab_style(
    style = cell_fill(color = "#f0f0f0"),
    locations = cells_body(rows = class == "factor")
  ) %>%
  tab_style(
    style = cell_fill(color = "#e6f2ff"),
    locations = cells_body(rows = class == "character")
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  tab_options(
    table.font.size = px(12),
    data_row.padding = px(3)
  ) %>%
  cols_width(
    variable ~ px(200),
    description ~ px(350),
    everything() ~ px(80)
  )

# Save HTML codebook
gtsave(codebook_gt, here("docs", "codebook.html"))
cat("✓ Saved: docs/codebook.html\n")

# Print summary
cat("\n=== CODEBOOK SUMMARY ===\n")
cat("Total variables:", nrow(codebook), "\n")
cat("Variable types:\n")
print(table(codebook$class))

cat("\nVariables with >20% missing:\n")
high_missing <- codebook %>%
  filter(pct_missing > 20) %>%
  select(variable, pct_missing, description)

if(nrow(high_missing) > 0) {
  print(high_missing)
} else {
  cat("  None\n")
}

cat("\nProcessed data saved to:\n")
cat("  ✓ data/processed/ab_analysis.rds (R format)\n")
cat("  ✓ data/processed/ab_analysis.csv (CSV format)\n")
cat("  ✓ data/processed/ab_analysis.dta (Stata format)\n")
cat("  ✓ docs/codebook.csv (Codebook CSV)\n")
cat("  ✓ docs/codebook.html (Codebook HTML)\n")
```

# Data Quality Report

```{r 31-data-quality-report}
# ============================================================================
# COMPREHENSIVE DATA QUALITY REPORT
# ============================================================================

cat("\n=== DATA QUALITY REPORT ===\n\n")

# 1. Sample characteristics
cat("1. SAMPLE CHARACTERISTICS\n")
cat("   Total N:", nrow(ab_analysis), "\n\n")
ab_analysis %>%
  count(country_name) %>%
  mutate(pct = round(n / sum(n) * 100, 1)) %>%
  print()

# 2. Demographics by country
cat("\n2. DEMOGRAPHICS BY COUNTRY\n")
# ... (This section is good as-is) ...

# 3. Variable completeness (using correct names)
cat("\n3. VARIABLE COMPLETENESS (Key Composites)\n")
ab_analysis %>%
  summarise(
    across(c(institutional_trust_index, 
             dem_satisfaction, dem_legitimacy,
             regime_preference, auth_acceptance,
             emergency_powers_support,
             covid_govt_performance, covid_restrict_composite),
           ~sum(!is.na(.)) / n() * 100,
           .names = "{col}_pct_complete")
  ) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "pct_complete") %>%
  mutate(variable = str_remove(variable, "_pct_complete"),
         pct_complete = round(pct_complete, 1)) %>%
  arrange(pct_complete) %>%
  print(n = 20)

# 4. Data quality by country (using correct names)
cat("\n4. DATA QUALITY BY COUNTRY (% Complete)\n")
quality_by_country <- ab_analysis %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    trust_complete = round(sum(!is.na(institutional_trust_index)) / n() * 100, 1),
    dem_sat_complete = round(sum(!is.na(dem_satisfaction)) / n() * 100, 1),
    dem_leg_complete = round(sum(!is.na(dem_legitimacy)) / n() * 100, 1),
    auth_accept_complete = round(sum(!is.na(auth_acceptance)) / n() * 100, 1),
    regime_pref_complete = round(sum(!is.na(regime_preference)) / n() * 100, 1),
    emergency_complete = round(sum(!is.na(emergency_powers_support)) / n() * 100, 1),
    covid_perf_complete = round(sum(!is.na(covid_govt_performance)) / n() * 100, 1),
    covid_restrict_complete = round(sum(!is.na(covid_restrict_composite)) / n() * 100, 1)
  )
print(quality_by_country)

# 5. Key descriptive statistics (using correct names)
cat("\n5. KEY DESCRIPTIVE STATISTICS BY COUNTRY (Means)\n")
descriptives <- ab_analysis %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    mean_trust = round(mean(institutional_trust_index, na.rm = TRUE), 2),
    mean_dem_sat = round(mean(dem_satisfaction, na.rm = TRUE), 2),
    mean_dem_leg = round(mean(dem_legitimacy, na.rm = TRUE), 2),
    mean_auth_accept = round(mean(auth_acceptance, na.rm = TRUE), 2),
    mean_regime_pref = round(mean(regime_preference, na.rm = TRUE), 2),
    mean_emergency = round(mean(emergency_powers_support, na.rm = TRUE), 2),
    mean_covid_perf = round(mean(covid_govt_performance, na.rm = TRUE), 2),
    covid_rate = round(mean(covid_contracted, na.rm = TRUE) * 100, 1)
  )
print(descriptives)

# 6. Cross-country comparisons (statistical tests)
cat("\n6. CROSS-COUNTRY DIFFERENCES (ANOVA)\n")

# Test if countries differ significantly on key measures
key_measures <- c("institutional_trust_index", "auth_acceptance", 
                  "regime_preference", "covid_govt_performance")

for(measure in key_measures) {
  cat(sprintf("\n%s:\n", measure))
  formula_str <- paste(measure, "~ country_name")
  anova_result <- aov(as.formula(formula_str), data = ab_analysis)
  anova_summary <- summary(anova_result)
  
  f_val <- anova_summary[[1]]$`F value`[1]
  p_val <- anova_summary[[1]]$`Pr(>F)`[1]
  
  cat(sprintf("  F = %.2f, p = %.4f", f_val, p_val))
  if(p_val < 0.001) {
    cat(" ***\n")
  } else if(p_val < 0.01) {
    cat(" **\n")
  } else if(p_val < 0.05) {
    cat(" *\n")
  } else {
    cat(" (ns)\n")
  }
}

# 7. Save summary tables
cat("\n7. SAVING SUMMARY TABLES\n")

# Demographics table
demo_table <- ab_analysis %>%
  group_by(country_name) %>%
  summarise(
    n = n(),
    pct_female = round(mean(gender == "Female", na.rm = TRUE) * 100, 1),
    mean_age = round(mean(age, na.rm = TRUE), 1),
    pct_urban = round(mean(urban, na.rm = TRUE) * 100, 1),
    pct_tertiary = round(mean(educ_level == "University/Tertiary", na.rm = TRUE) * 100, 1)
  )

write_csv(demo_table, here("tables", "demographics_summary.csv"))
cat("✓ Saved: tables/demographics_summary.csv\n")

# Descriptives table
write_csv(descriptives, here("tables", "descriptives_by_country.csv"))
cat("✓ Saved: tables/descriptives_by_country.csv\n")

# Quality table
write_csv(quality_by_country, here("tables", "data_quality_by_country.csv"))
cat("✓ Saved: tables/data_quality_by_country.csv\n")

cat("\n=== DATA PREPARATION COMPLETE ===\n")
cat("\n✓ Dataset ready for analysis\n")
cat("✓ All quality checks passed\n")
cat("✓ Summary tables exported\n")
cat("\nNext step: Proceed to Script 03 - Descriptive Analysis\n")
```

# Session Information

```{r 32-session-info}
sessionInfo()
```
