---
title: "Data Preparation"
subtitle: "Meaning of Democracy Paper"
author: "Jeffrey Stark"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    df-print: paged
execute:
  warning: false
  message: false
---

# Overview

This script prepares the analysis dataset from the harmonized Asian Barometer master file (`abs_econdev_authpref.rds`, 336 variables). The master file now contains all harmonized variables including:

- `electoral_status`: Winner/loser status (1=Winner, 2=Loser)
- `procedural_preference_index`: 0-4 scale (count of procedural choices)
- `dem_procedural_vs_substantive`: Binary procedural/substantive measure

# Setup

```{r setup}
library(tidyverse)
library(here)

here::i_am("papers/meaning-of-democracy/analysis/01_data_preparation.qmd")

# Output directory for paper-specific analysis data
output_dir <- here("papers", "meaning-of-democracy", "analysis", "data")
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# Master dataset location
master_file <- here("data", "processed", "abs_econdev_authpref.rds")
```

# Load Master Dataset

```{r load}
abs <- readRDS(master_file)

cat("Master dataset:", nrow(abs), "rows,", ncol(abs), "columns\n")
cat("Waves:", paste(sort(unique(abs$wave)), collapse = ", "), "\n")
```

# Helper Functions

```{r helpers}
# Country name lookup (harmonized country codes)
get_country_name <- function(code) {
  code <- as.integer(code)
  case_when(
    code == 1 ~ "Japan",
    code == 2 ~ "Hong Kong",
    code == 3 ~ "South Korea",
    code == 4 ~ "China",
    code == 5 ~ "Mongolia",
    code == 6 ~ "Philippines",
    code == 7 ~ "Taiwan",
    code == 8 ~ "Thailand",
    code == 9 ~ "Indonesia",
    code == 10 ~ "Singapore",
    code == 11 ~ "Vietnam",
    code == 12 ~ "Cambodia",
    code == 13 ~ "Malaysia",
    code == 14 ~ "Myanmar",
    code == 15 ~ "Australia",
    code == 18 ~ "India",
    TRUE ~ NA_character_
  )
}

# Wave year lookup (approximate midpoint of fieldwork)
get_wave_year <- function(wave) {
  # Approximate midpoint of fieldwork for each wave
  # Based on actual interview dates from 05_interview_dates.qmd
  case_when(
    wave == 1 ~ 2001L,
    wave == 2 ~ 2006L,  # Jul 2004 - Dec 2008, bulk 2006-07
    wave == 3 ~ 2011L,  # Jan 2010 - Nov 2012
    wave == 4 ~ 2015L,  # Jun 2014 - Apr 2016
    wave == 5 ~ 2017L,
    wave == 6 ~ 2022L,  # Jul 2021 - Feb 2023
    TRUE ~ NA_integer_
  )
}
```

# Check Key Variables

```{r check-vars}
# Verify electoral_status and procedural_preference_index are present
cat("=== KEY VARIABLES CHECK ===\n\n")

# electoral_status
cat("electoral_status:\n")
if ("electoral_status" %in% names(abs)) {
  cat("  Present: YES\n")
  cat("  Values:", paste(sort(unique(abs$electoral_status[!is.na(abs$electoral_status)])), collapse = ", "), "\n")
  cat("  Non-missing by wave:\n")
  abs %>% 
    filter(!is.na(electoral_status)) %>% 
    count(wave) %>% 
    mutate(label = paste0("    W", wave, ": ", n)) %>% 
    pull(label) %>% 
    cat(sep = "\n")
} else {
  cat("  Present: NO - check harmonization\n")
}

cat("\n")

# procedural_preference_index
cat("procedural_preference_index:\n")
if ("procedural_preference_index" %in% names(abs)) {
  cat("  Present: YES\n")
  cat("  Range:", min(abs$procedural_preference_index, na.rm = TRUE), "-", 
      max(abs$procedural_preference_index, na.rm = TRUE), "\n")
  cat("  Non-missing by wave:\n")
  abs %>% 
    filter(!is.na(procedural_preference_index)) %>% 
    count(wave) %>% 
    mutate(label = paste0("    W", wave, ": ", n)) %>% 
    pull(label) %>% 
    cat(sep = "\n")
} else {
  cat("  Present: NO - check harmonization\n")
}
```

# Process Dataset

```{r process}
# Filter to waves with both electoral_status and procedural_preference_index
# W2, W3, W4, W6 have these variables (W1 and W5 do not)
combined <- abs %>%
  filter(wave %in% c(2, 3, 4, 6)) %>%
  filter(!is.na(electoral_status), !is.na(procedural_preference_index)) %>%
  mutate(
    # Format wave label
    wave_label = paste0("W", wave),
    wave_year = get_wave_year(wave),
    country_name = get_country_name(country),
    
    # Winner/loser as character (1=Winner, 2=Loser in harmonized data)
    winner_loser = case_when(
      electoral_status == 1 ~ "Winner",
      electoral_status == 2 ~ "Loser",
      TRUE ~ NA_character_
    ),
    
    # Procedural index (already 0-4 scale from harmonization)
    procedural_index = procedural_preference_index,
    
    # Binary procedural measure (>=2 procedural choices = procedural)
    # For W2: 0 or 4 only, so threshold works correctly
    procedural_single = if_else(procedural_index >= 2, 1L, 0L),
    
    # Measure type (single item for W2, four-set battery for W3/W4/W6)
    measure_type = if_else(wave == 2, "single_item", "four_set")
  ) %>%
  filter(!is.na(country_name))

cat("\n=== COMBINED DATASET ===\n")
cat("Total observations:", nrow(combined), "\n")
cat("Waves:", paste(unique(combined$wave_label), collapse = ", "), "\n")
cat("Countries:", n_distinct(combined$country_name), "\n")
```

# Summary Statistics

```{r summary}
# By wave
cat("=== SAMPLE BY WAVE ===\n")
combined %>%
  group_by(wave_label, wave_year) %>%
  summarise(
    n = n(),
    n_winner = sum(winner_loser == "Winner"),
    n_loser = sum(winner_loser == "Loser"),
    pct_procedural = round(mean(procedural_single) * 100, 1),
    .groups = "drop"
  ) %>%
  knitr::kable(caption = "Sample by Wave")

# By country
cat("\n=== SAMPLE BY COUNTRY AND WAVE ===\n")
combined %>%
  count(country_name, wave_label) %>%
  pivot_wider(names_from = wave_label, values_from = n, values_fill = 0) %>%
  mutate(total = rowSums(across(starts_with("W")))) %>%
  arrange(desc(total)) %>%
  knitr::kable(caption = "Sample by Country and Wave")
```

# Country Coverage Analysis

```{r coverage}
# Which countries have data in which waves?
coverage <- combined %>%
  distinct(country_name, wave_label) %>%
  mutate(present = 1) %>%
  pivot_wider(names_from = wave_label, values_from = present, values_fill = 0) %>%
  mutate(
    n_waves = rowSums(across(starts_with("W"))),
    trajectory_possible = n_waves >= 3
  ) %>%
  arrange(desc(n_waves))

cat("=== COUNTRY COVERAGE ===\n")
print(coverage, n = 20)

cat("\n=== COUNTRIES WITH 3+ WAVES (trajectory analysis possible) ===\n")
trajectory_countries <- coverage %>% filter(trajectory_possible) %>% pull(country_name)
cat(paste(trajectory_countries, collapse = ", "), "\n")
```

# Save Analysis Dataset

```{r save}
# Select and reorder columns for analysis dataset
analysis_data <- combined %>%
  select(
    wave = wave_label,
    wave_year,
    country_name,
    winner_loser,
    procedural_single,
    procedural_index,
    measure_type
  )

# Save
saveRDS(analysis_data, file.path(output_dir, "winner_loser_4waves.rds"))
write_csv(analysis_data, file.path(output_dir, "winner_loser_4waves.csv"))

cat("\n=== SAVED ===\n")
cat("RDS:", file.path(output_dir, "winner_loser_4waves.rds"), "\n")
cat("CSV:", file.path(output_dir, "winner_loser_4waves.csv"), "\n")
cat("Rows:", nrow(analysis_data), "\n")
cat("Cols:", ncol(analysis_data), "\n")
```

# Variable Codebook

| Variable | Description | Source |
|----------|-------------|--------|
| `wave` | Survey wave (W2, W3, W4, W6) | `wave` column, reformatted |
| `wave_year` | Approximate midpoint year of fieldwork | Derived from wave |
| `country_name` | Country name | `country` column, recoded |
| `winner_loser` | Electoral status: "Winner" or "Loser" | `electoral_status` |
| `procedural_single` | Binary: 1 = procedural, 0 = substantive | `procedural_preference_index >= 2` |
| `procedural_index` | 0-4 scale (W3/W4/W6) or 0/4 (W2) | `procedural_preference_index` |
| `measure_type` | "single_item" (W2) or "four_set" (W3/W4/W6) | Derived from wave |

# Data Notes

## Source Variables

- **Data source**: Uses harmonized variables from the main ABS data pipeline (`abs_econdev_authpref.rds`)
- **electoral_status**: Harmonized winner/loser variable
  - W2: q39a
  - W3: q33a
  - W4: q34a
  - W6: q34a
- **procedural_preference_index**: Harmonized 0-4 index
  - W2: Single item (q92) scaled to 0/4
  - W3: 4-set battery (q85-q88)
  - W4: 4-set battery (q88-q91)
  - W6: 4-set battery (q85-q88)

## Measurement Notes

- **W2 caveat**: Uses single forced-choice item (q92), not the 4-set battery. Values are only 0 or 4.
- **Missing data**: Observations with missing winner/loser status or procedural measure are excluded.
- **W1 and W5**: Do not have the electoral_status (winner/loser) question

## Country Coverage

```{r coverage-table}
#| echo: false
coverage %>%
  select(country_name, W2, W3, W4, W6, n_waves) %>%
  knitr::kable(caption = "Country × Wave Coverage Matrix")
```
