---
title: "Supplementary Materials"
subtitle: "I Don't Care How the Sausage Gets Made"
author: "Jeffrey Stark"
date: today
format:
  pdf:
    documentclass: article
    geometry: margin=1in
    fontsize: 10pt
    number-sections: true
  html:
    toc: true
    code-fold: true
execute:
  warning: false
  message: false
  echo: false
---

```{r setup}
#| include: false
library(tidyverse)
library(here)
library(gt)
library(knitr)

here::i_am("papers/sausage-paper/supplementary/appendix.qmd")
```

# Appendix A: Variable Definitions

## Dependent Variables

| Variable | Question Text | Scale | Waves |
|----------|---------------|-------|-------|
| `econ_over_democracy` | [Question text] | 1-4 | W1-W6 |
| `democracy_solves_problems` | [Question text] | 1-4 | W2-W6 |

## Independent Variables

| Variable | Definition | Coding |
|----------|------------|--------|
| Regime Type | Based on V-Dem classification | Democracy, Hybrid, Autocracy |
| Education | Years of formal schooling | 0-25 years |
| Survey Wave | ABS wave number | W1-W6 (2001-2020) |

## Control Variables

[Table of controls]

# Appendix B: Country Classifications

## V-Dem Regime Classifications

```{r vdem-table}
# TODO: Insert V-Dem classification table by country-year
```

## Alternative Classification Schemes

[Discussion of alternative classification approaches]

# Appendix C: Full Regression Tables

## H1: Main Effects Models

```{r h1-full}
# TODO: Full model output
```

## H2: Interaction Models

```{r h2-full}
# TODO: Full model output
```

## H3: Time Trend Models

```{r h3-full}
# TODO: Full model output
```

# Appendix D: Robustness Checks

## Alternative Specifications

[Summary of robustness results]

## Sensitivity to Sample Restrictions

[Results with different sample restrictions]

# Appendix E: Survey Question Wording

## Wave 1 (2001-2003)

[Question wording for W1]

## Wave 2 (2005-2008)

[Question wording for W2]

## Waves 3-6

[Changes in question wording across waves]

# Appendix F: Missing Data Analysis

## Missing Data Patterns

```{r missing}
# TODO: Missing data visualization
```

## Comparison of Complete vs. Full Sample

[Analysis comparing results with different missing data treatments]
