---
title: "Online Appendix: The Satisfaction Paradox"
subtitle: "Economic Performance and the Structure of Democratic Support in South Korea"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: today
date-format: "MMMM D, YYYY"
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{caption}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \captionsetup{width=\textwidth}
          \setlength\LTcapwidth{\textwidth}
          \floatplacement{figure}{H}
          \floatplacement{table}{H}
          \raggedright
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
    embed-resources: true
  docx:
    toc: true
    number-sections: true
bibliography: references.bib
csl: apa.csl
execute:
  warning: false
  message: false
  echo: false
  cache: false
---

```{r setup}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/03b_sk_satisfaction_paradox")
results_dir  <- file.path(paper_dir, "analysis", "results")

source(file.path(paper_dir, "R", "helpers.R"))

dat  <- readRDS(file.path(results_dir, "analysis_data.rds"))
mr   <- readRDS(file.path(results_dir, "model_results.rds"))
ar   <- readRDS(file.path(results_dir, "auth_rejection_results.rds"))
ir   <- readRDS(file.path(results_dir, "item_specificity_results.rds"))
desc <- readRDS(file.path(results_dir, "descriptive_stats.rds"))

yr_kr <- c(`1`=2003L, `2`=2006L, `3`=2011L, `4`=2015L, `5`=2019L, `6`=2022L)
yr_tw <- c(`1`=2001L, `2`=2006L, `3`=2010L, `4`=2014L, `5`=2019L, `6`=2022L)

sfn <- function(p) case_when(p<.001~"***", p<.01~"**", p<.05~"*", TRUE~"")
```

# Appendix A: Data and Measurement

## A.1 Survey Coverage

The analysis draws on six waves of the Asian Barometer Survey (ABS), a comparative
survey program administering nationally representative face-to-face interviews across
East and Southeast Asia [@Asian-Barometer-Survey2023-uh]. @tbl-coverage reports
wave-level sample sizes for Korea and Taiwan.

```{r tbl-coverage}
#| label: tbl-coverage
#| tbl-cap: "Sample coverage: Korea and Taiwan, ABS waves 1–6"

dat |>
  group_by(country_label, wave) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(year = case_when(
    country_label == "Korea"  ~ yr_kr[as.character(wave)],
    country_label == "Taiwan" ~ yr_tw[as.character(wave)]
  )) |>
  pivot_wider(names_from = country_label, values_from = c(year, n), names_sep = "_") |>
  select(wave, year_Korea, n_Korea, year_Taiwan, n_Taiwan) |>
  kbl(
    col.names = c("Wave", "Year", "N", "Year", "N"),
    booktabs = TRUE, align = "lrrrr", format.args = list(big.mark = ",")
  ) |>
  add_header_above(c(" " = 1, "Korea" = 2, "Taiwan" = 2)) |>
  kable_styling(latex_options = "hold_position")
```

## A.2 Variable Construction and Question Wording

All survey items were normalized to a 0–1 scale within country using min–max
normalization. Indices are row-wise means of their normalized components.

**Satisfaction index** (2 items, all waves):

- *Democracy satisfaction* (`democracy_satisfaction`): "On the whole, how satisfied
  or dissatisfied are you with the way democracy works in our country?" (1 = very
  dissatisfied, 4 = very satisfied)
- *Government performance* (`gov_sat_national`): "How satisfied are you with the
  performance of the current national government?" (1 = very dissatisfied, 4 = very
  satisfied)

**Democratic quality index** (2 items, all waves):

- *Always preferable* (`dem_always_preferable`): "Which of the following statements
  comes closest to your own opinion? (a) Democracy is always preferable to any other
  kind of government; (b) Under some circumstances, an authoritarian government can
  be preferable; (c) For people like me, it doesn't matter whether we have a
  democratic or a non-democratic government." (Coded 1 = always preferable,
  0 otherwise, then normalized.)
- *Extent democratic* (`dem_extent_current`): "How much of a democracy is [country]
  today?" (1–10 scale)

**Authoritarian rejection index** (4 items; higher = more rejection of authoritarianism):

"I'm going to describe some political systems. For each, tell me whether you think
it would be very good, fairly good, fairly bad, or very bad for governing [country]."

- "Having a strong leader who does not have to bother with parliament and elections" (strongman rule)
- "Having the army rule the country" (military rule)
- "Having experts, not government, make decisions according to what they think is
  best for the country" (expert rule)
- "Having only one political party allowed to stand for election and rule the country"
  (single-party rule)

Each item scored 1 = very good … 4 = very bad. Items are reverse-coded before
normalization so that higher values indicate stronger rejection of authoritarianism.

**Economic evaluation index** (6 items, all waves):

Six items spanning sociotropic and pocketbook assessments across current,
retrospective, and prospective time horizons: national economy now; household economy
now; national economic outlook; household economic outlook; national economic change;
household economic change. All scored 1–5 (worse to better) and normalized 0–1.

**Controls** [@Dalton2004-bq]: age (normalized 0–1), gender (1 = male), education
level (normalized 0–1), urban/rural residence (dummy), political interest
(normalized 0–1).

## A.3 Wave-Level Descriptive Statistics

```{r tbl-desc-kr}
#| label: tbl-desc-kr
#| tbl-cap: "Korea: wave-level index means (normalized 0–1, SE in parentheses)"

desc$wave_means_kr |>
  mutate(
    sat  = sprintf("%.3f (%.3f)", sat_mean,  sat_se),
    qual = sprintf("%.3f (%.3f)", qual_mean, qual_se),
    econ = sprintf("%.3f (%.3f)", econ_mean, econ_se)
  ) |>
  select(wave, year, n, sat, qual, econ) |>
  kbl(
    col.names = c("Wave", "Year", "N", "Satisfaction", "Dem. quality", "Econ. eval."),
    booktabs = TRUE, align = "lrrlll", format.args = list(big.mark = ",")
  ) |>
  add_footnote("SE = standard error of the mean.",
               notation = "none") |>
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

```{r tbl-desc-tw}
#| label: tbl-desc-tw
#| tbl-cap: "Taiwan: wave-level index means (normalized 0–1, SE in parentheses)"

desc$wave_means_tw |>
  mutate(
    sat  = sprintf("%.3f (%.3f)", sat_mean,  sat_se),
    qual = sprintf("%.3f (%.3f)", qual_mean, qual_se)
  ) |>
  select(wave, year, n, sat, qual) |>
  kbl(
    col.names = c("Wave", "Year", "N", "Satisfaction", "Dem. quality"),
    booktabs = TRUE, align = "lrrrr", format.args = list(big.mark = ",")
  ) |>
  add_footnote("Econ. eval. index not shown for Taiwan as it is not the focus of the paper.",
               notation = "none") |>
  kable_styling(latex_options = "hold_position")
```

# Appendix B: Full Wave-by-Wave OLS Results

@tbl-wave-all presents OLS coefficients on `econ_index` for all five dependent
variables across all six waves, for Korea and Taiwan separately. All models include
controls for age, gender, education, urban/rural residence, and political interest.
The table underpins Figures 3 and 4 in the main text.

```{r tbl-wave-all}
#| label: tbl-wave-all
#| tbl-cap: "Wave-by-wave OLS: econ to DV for all outcomes, Korea and Taiwan"

wave_wide <- mr$wave_by_wave |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s", estimate, sig),
    year = case_when(
      country == "Korea"  ~ yr_kr[as.character(wave)],
      country == "Taiwan" ~ yr_tw[as.character(wave)]
    )
  ) |>
  select(country, year, dv_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell) |>
  arrange(country, year)

wave_wide |>
  kbl(
    col.names = c("Country", "Year", names(wave_wide)[-(1:2)]),
    booktabs = TRUE, align = paste0("ll", strrep("r", ncol(wave_wide) - 2))
  ) |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  add_footnote(
    "OLS coefficient on econ_index. Controls: age, gender, education, urban/rural, political interest. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

# Appendix C: Pooled Individual-DV Models

@tbl-pooled reports pooled OLS results (with wave fixed effects) for each individual
dependent variable in Korea and Taiwan. This complements the wave-by-wave table by
presenting a single summary estimate for each outcome across the full survey period.

```{r tbl-pooled}
#| label: tbl-pooled
#| tbl-cap: "Pooled OLS: econ$\\to$individual DVs, Korea and Taiwan"

mr$pooled_indiv |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(country, dv_label, cell, n) |>
  pivot_wider(names_from = country, values_from = c(cell, n), names_sep = "_") |>
  select(dv_label, cell_Korea, n_Korea, cell_Taiwan, n_Taiwan) |>
  kbl(
    col.names = c("Dependent variable", "$\\beta$ (SE)", "N", "$\\beta$ (SE)", "N"),
    booktabs = TRUE, align = "lrrrr", escape = FALSE,
    format.args = list(big.mark = ",")
  ) |>
  add_header_above(c(" " = 1, "Korea" = 2, "Taiwan" = 2)) |>
  column_spec(1, width = "4.5cm") |>
  column_spec(c(2, 4), width = "2.8cm") |>
  footnote(
    general = "OLS b (econ_index) with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE
  ) |>
  kable_styling(latex_options = "hold_position")
```

# Appendix D: Cross-Country Interaction Test

@tbl-cross-country tests the H4 prediction that the Korea–Taiwan divergence in the
economic–quality relationship is statistically significant. Taiwan is the reference
category; a positive `econ_index:is_korea` interaction indicates that the positive
economic effect on satisfaction is stronger in Korea than in Taiwan; a negative
coefficient indicates the reverse (i.e., Taiwan shows a stronger negative effect on
quality, as expected under the critical-citizens pattern).

```{r tbl-cross-country}
#| label: tbl-cross-country
#| tbl-cap: "Cross-country interaction: econ x Korea, all five outcomes"

term_labels <- c(
  "econ_index"          = "Econ. eval.",
  "is_korea"            = "Korea (vs. Taiwan)",
  "econ_index:is_korea" = "Econ. x Korea"
)

xc_wide <- mr$cross_country |>
  filter(term %in% names(term_labels)) |>
  mutate(
    term_label = recode(term, !!!term_labels),
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(dv_label, term_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell)

xc_wide |>
  kbl(
    col.names = c("Term", names(xc_wide)[-1]),
    booktabs = TRUE, align = paste0("l", strrep("r", ncol(xc_wide)-1)),
    escape = FALSE
  ) |>
  add_footnote(
    "OLS. Taiwan = reference category. Pooled model with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

# Appendix E: Subgroup Heterogeneity

## E.1 Political Interest Subgroups

H5 predicts the satisfaction–quality decoupling is not concentrated in a single
demographic group. @tbl-polint presents results by political interest tertile
(low, medium, high) for Korea. All subgroups should show positive econ → satisfaction
and near-zero or negative econ → quality.

```{r tbl-polint}
#| label: tbl-polint
#| tbl-cap: "Korea: econ to DV by political interest subgroup"

# 3 DVs x 2 groups: pivot to wide, just show cell (estimate + SE + stars)
polint_wide <- mr$polint_subgroup |>
  filter(country == "Korea") |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(group, dv_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell)

polint_wide |>
  kbl(
    col.names = c("Group", names(polint_wide)[-1]),
    booktabs = TRUE, align = paste0("l", strrep("r", ncol(polint_wide) - 1))
  ) |>
  add_footnote(
    "OLS coefficient on econ_index with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = "hold_position")
```

## E.2 Winner–Loser Subgroups

Electoral winners and losers often display different patterns of democratic support.
@tbl-winner tests whether the satisfaction–quality decoupling is symmetric across
winners (supporters of the governing party) and losers (supporters of the opposition).

```{r tbl-winner}
#| label: tbl-winner
#| tbl-cap: "Korea and Taiwan: econ to DV by electoral winner/loser status"

# 2 DVs x 2 groups x 2 countries: country + group + 2 cell cols = 4 cols
winner_wide <- mr$winner_loser_subgroup |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error),
    dv_label = recode(dv_label,
      "Dem always preferable"      = "Dem. always pref.",
      "Satisfaction with democracy" = "Sat. w/ democracy"
    )
  ) |>
  select(country, group, dv_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell) |>
  arrange(country, group)

winner_wide |>
  kbl(
    col.names = c("Country", "Group", names(winner_wide)[-(1:2)]),
    booktabs = TRUE, align = paste0("ll", strrep("r", ncol(winner_wide) - 2))
  ) |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  column_spec(3:ncol(winner_wide), width = "3.5cm") |>
  footnote(
    general = "OLS coefficient on econ_index with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE
  ) |>
  kable_styling(latex_options = "hold_position")
```

# Appendix F: Authoritarian Rejection — Full Results

## F.1 Pooled Models

@tbl-auth-pooled reports pooled OLS estimates of econ → authoritarian rejection
(index and four component items) for Korea and Taiwan. The sign test is: if
the Korean decoupling reflects democratic erosion, better economic evaluations should
weaken rejection of authoritarian alternatives. If it reflects only a
performance-satisfaction response, they should strengthen rejection.

```{r tbl-auth-pooled}
#| label: tbl-auth-pooled
#| tbl-cap: "Pooled econ$\\to$authoritarian rejection: index and component items"

ar$pooled |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(country, dv_label, cell, n) |>
  pivot_wider(names_from = country, values_from = c(cell, n), names_sep = "_") |>
  select(dv_label, cell_Korea, n_Korea, cell_Taiwan, n_Taiwan) |>
  kbl(
    col.names = c("Dependent variable", "$\\beta$ (SE)", "N", "$\\beta$ (SE)", "N"),
    booktabs = TRUE, align = "lllll", escape = FALSE,
    format.args = list(big.mark = ",")
  ) |>
  add_header_above(c(" " = 1, "Korea" = 2, "Taiwan" = 2)) |>
  add_footnote(
    "OLS b (econ_index) with wave FE and standard controls. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = "hold_position")
```

## F.2 Wave-by-Wave Authoritarian Rejection

@tbl-auth-wave presents the wave-by-wave results for the authoritarian rejection
index and its four component items in both countries.

```{r tbl-auth-wave}
#| label: tbl-auth-wave
#| tbl-cap: "Wave-by-wave econ to authoritarian rejection"

auth_wave_wide <- ar$wave_by_wave |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s", estimate, sig),
    year = case_when(
      country == "Korea"  ~ yr_kr[as.character(wave)],
      country == "Taiwan" ~ yr_tw[as.character(wave)]
    )
  ) |>
  select(country, year, dv_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell) |>
  arrange(country, year)

auth_wave_wide |>
  kbl(
    col.names = c("Country", "Year", names(auth_wave_wide)[-(1:2)]),
    booktabs = TRUE, align = paste0("ll", strrep("r", ncol(auth_wave_wide) - 2))
  ) |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  add_footnote(
    "OLS coefficient on econ_index. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

## F.3 Cross-Country Interaction

@tbl-auth-xc tests whether Korea's positive econ → auth rejection relationship is
significantly different from Taiwan's, using the same cross-country interaction
framework as Appendix D.

```{r tbl-auth-xc}
#| label: tbl-auth-xc
#| tbl-cap: "Cross-country interaction: econ x Korea on authoritarian rejection outcomes"

auth_term_labels <- c(
  "econ_index"          = "Econ. eval.",
  "is_korea"            = "Korea",
  "econ_index:is_korea" = "Econ. x Korea"
)

auth_xc_wide <- ar$cross_country |>
  filter(term %in% names(auth_term_labels)) |>
  mutate(
    term_label = recode(term, !!!auth_term_labels),
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(dv_label, term_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell)

auth_xc_wide |>
  kbl(
    col.names = c("Term", names(auth_xc_wide)[-1]),
    booktabs = TRUE, align = paste0("l", strrep("r", ncol(auth_xc_wide) - 1))
  ) |>
  add_footnote(
    "OLS. Taiwan = reference. Pooled with wave FE and controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

# Appendix G: Item Specificity

## G.1 All Outcomes Side-by-Side

@tbl-item-spec presents OLS coefficients for `econ_index` on all nine individual
dependent variables for Korea and Taiwan, plus the Korea–Taiwan gap. The key
result is that the gap between Korea's positive economic effect on the abstract
normative item ("dem always preferable") and its effect on the authoritarian
rejection items is much larger than the corresponding gap in Taiwan—confirming
that the abstract item is capturing something distinct from substantive democratic
commitment.

```{r tbl-item-spec}
#| label: tbl-item-spec
#| tbl-cap: "Item specificity: econ$\\to$all DVs, Korea and Taiwan"

ir$pooled |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(country, dv_label, cell, n) |>
  pivot_wider(names_from = country, values_from = c(cell, n), names_sep = "_") |>
  select(dv_label, cell_Korea, n_Korea, cell_Taiwan, n_Taiwan) |>
  kbl(
    col.names = c("Dependent variable", "$\\beta$ (SE)", "N", "$\\beta$ (SE)", "N"),
    booktabs = TRUE, align = "lrrrr", escape = FALSE,
    format.args = list(big.mark = ",")
  ) |>
  add_header_above(c(" " = 1, "Korea" = 2, "Taiwan" = 2)) |>
  column_spec(1, width = "4.5cm") |>
  column_spec(c(2, 4), width = "2.8cm") |>
  footnote(
    general = "Pooled OLS b (econ_index) with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE
  ) |>
  kable_styling(latex_options = "hold_position")
```

# References {.unnumbered}

::: {#refs}
:::
