---
title: "Online Appendix: The Satisfaction Paradox"
subtitle: "Economic Performance and the Structure of Democratic Support in South Korea"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: today
date-format: "MMMM D, YYYY"
format:
  pdf:
    documentclass: article
    classoption: [12pt, letterpaper]
    geometry:
      - margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{setspace}
          \usepackage{booktabs}
          \usepackage{caption}
          \usepackage{longtable}
          \usepackage{array}
          \usepackage{multirow}
          \usepackage{float}
          \usepackage{graphicx}
          \usepackage{threeparttable}
          \usepackage{threeparttablex}
          \captionsetup{width=\textwidth}
          \setlength\LTcapwidth{\textwidth}
          \floatplacement{figure}{H}
          \floatplacement{table}{H}
          \raggedright
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
    code-fold: true
    embed-resources: true
prefer-html: true
bibliography: references.bib
csl: apa.csl
execute:
  warning: false
  message: false
  echo: false
  cache: false
---

```{r setup}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)

project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/03b_sk_satisfaction_paradox")
results_dir  <- file.path(paper_dir, "analysis", "results")

source(file.path(paper_dir, "R", "helpers.R"))

dat  <- readRDS(file.path(results_dir, "analysis_data.rds"))
mr   <- readRDS(file.path(results_dir, "model_results.rds"))
ar   <- readRDS(file.path(results_dir, "auth_rejection_results.rds"))
ir   <- readRDS(file.path(results_dir, "item_specificity_results.rds"))
desc <- readRDS(file.path(results_dir, "descriptive_stats.rds"))

yr_kr <- c(`1`=2003L, `2`=2006L, `3`=2011L, `4`=2015L, `5`=2019L, `6`=2022L)
yr_tw <- c(`1`=2001L, `2`=2006L, `3`=2010L, `4`=2014L, `5`=2019L, `6`=2022L)

sfn <- function(p) case_when(p<.001~"***", p<.01~"**", p<.05~"*", TRUE~"")
```

# Appendix A: Data and Measurement

## A.1 Survey Coverage

The analysis draws on six waves of the Asian Barometer Survey (ABS), a comparative
survey program administering nationally representative face-to-face interviews across
East and Southeast Asia [@Asian-Barometer-Survey2023-uh]. @tbl-coverage reports
wave-level sample sizes for Korea and Taiwan.

```{r tbl-coverage}
#| label: tbl-coverage
#| tbl-cap: "Sample coverage: Korea and Taiwan, ABS waves 1–6"

dat |>
  group_by(country_label, wave) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(year = case_when(
    country_label == "Korea"  ~ yr_kr[as.character(wave)],
    country_label == "Taiwan" ~ yr_tw[as.character(wave)]
  )) |>
  pivot_wider(names_from = country_label, values_from = c(year, n), names_sep = "_") |>
  select(wave, year_Korea, n_Korea, year_Taiwan, n_Taiwan) |>
  kbl(
    col.names = c("Wave", "Year", "N", "Year", "N"),
    booktabs = TRUE, align = "lrrrr", format.args = list(big.mark = ",")
  ) |>
  add_header_above(c(" " = 1, "Korea" = 2, "Taiwan" = 2)) |>
  kable_styling(latex_options = "hold_position")
```

## A.2 Variable Construction and Question Wording

All survey items were normalized to a 0–1 scale within country using min–max
normalization. Indices are row-wise means of their normalized components.

**Satisfaction index** (2 items, all waves):

- *Democracy satisfaction* (`democracy_satisfaction`): "On the whole, how satisfied
  or dissatisfied are you with the way democracy works in our country?" (1 = very
  dissatisfied, 4 = very satisfied)
- *Government performance* (`gov_sat_national`): "How satisfied are you with the
  performance of the current national government?" (1 = very dissatisfied, 4 = very
  satisfied)

**Democratic quality index** (2 items, all waves):

- *Always preferable* (`dem_always_preferable`): "Which of the following statements
  comes closest to your own opinion? (a) Democracy is always preferable to any other
  kind of government; (b) Under some circumstances, an authoritarian government can
  be preferable; (c) For people like me, it doesn't matter whether we have a
  democratic or a non-democratic government." (Coded 1 = always preferable,
  0 otherwise, then normalized.)
- *Extent democratic* (`dem_extent_current`): "How much of a democracy is [country]
  today?" (1–10 scale)

**Authoritarian rejection index** (4 items; higher = more rejection of authoritarianism):

"I'm going to describe some political systems. For each, tell me whether you think
it would be very good, fairly good, fairly bad, or very bad for governing [country]."

- "Having a strong leader who does not have to bother with parliament and elections" (strongman rule)
- "Having the army rule the country" (military rule)
- "Having experts, not government, make decisions according to what they think is
  best for the country" (expert rule)
- "Having only one political party allowed to stand for election and rule the country"
  (single-party rule)

Each item scored 1 = very good … 4 = very bad. Items are reverse-coded before
normalization so that higher values indicate stronger rejection of authoritarianism.

**Economic evaluation index** (6 items, all waves):

Six items spanning sociotropic and pocketbook assessments across current,
retrospective, and prospective time horizons: national economy now; household economy
now; national economic outlook; household economic outlook; national economic change;
household economic change. All scored 1–5 (worse to better) and normalized 0–1.

**Controls** [@Dalton2004-bq]: age (normalized 0–1), gender (1 = male), education
level (normalized 0–1), urban/rural residence (dummy), political interest
(normalized 0–1).

## A.3 Wave-Level Descriptive Statistics

```{r tbl-desc-kr}
#| label: tbl-desc-kr
#| tbl-cap: "Korea: wave-level index means (normalized 0–1, SE in parentheses)"

desc$wave_means_kr |>
  mutate(
    sat  = sprintf("%.3f (%.3f)", sat_mean,  sat_se),
    qual = sprintf("%.3f (%.3f)", qual_mean, qual_se),
    econ = sprintf("%.3f (%.3f)", econ_mean, econ_se)
  ) |>
  select(wave, year, n, sat, qual, econ) |>
  kbl(
    col.names = c("Wave", "Year", "N", "Satisfaction", "Dem. quality", "Econ. eval."),
    booktabs = TRUE, align = "lrrlll", format.args = list(big.mark = ",")
  ) |>
  add_footnote("SE = standard error of the mean.",
               notation = "none") |>
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

```{r tbl-desc-tw}
#| label: tbl-desc-tw
#| tbl-cap: "Taiwan: wave-level index means (normalized 0–1, SE in parentheses)"

yr_tw_vec <- c(`1`=2001L, `2`=2006L, `3`=2010L, `4`=2014L, `5`=2019L, `6`=2022L)

dat |>
  filter(country_label == "Taiwan") |>
  group_by(wave) |>
  summarise(
    n        = n(),
    sat_mean = mean(sat_index,  na.rm = TRUE),
    sat_se   = sd(sat_index,    na.rm = TRUE) / sqrt(sum(!is.na(sat_index))),
    qual_mean= mean(qual_index, na.rm = TRUE),
    qual_se  = sd(qual_index,   na.rm = TRUE) / sqrt(sum(!is.na(qual_index))),
    econ_mean= mean(econ_index, na.rm = TRUE),
    econ_se  = sd(econ_index,   na.rm = TRUE) / sqrt(sum(!is.na(econ_index))),
    .groups  = "drop"
  ) |>
  mutate(
    year = yr_tw_vec[as.character(wave)],
    sat  = sprintf("%.3f (%.3f)", sat_mean,  sat_se),
    qual = sprintf("%.3f (%.3f)", qual_mean, qual_se),
    econ = sprintf("%.3f (%.3f)", econ_mean, econ_se)
  ) |>
  select(wave, year, n, sat, qual, econ) |>
  kbl(
    col.names = c("Wave", "Year", "N", "Satisfaction", "Dem. quality", "Econ. eval."),
    booktabs = TRUE, align = "lrrlll", format.args = list(big.mark = ",")
  ) |>
  add_footnote("SE = standard error of the mean.", notation = "none") |>
  kable_styling(latex_options = "hold_position")
```

# Appendix B: Full Wave-by-Wave OLS Results

@tbl-wave-all presents OLS coefficients on `econ_index` for all five dependent
variables across all six waves, for Korea and Taiwan separately. All models include
controls for age, gender, education, urban/rural residence, and political interest.
The table underpins Figures 3 and 4 in the main text.

```{r tbl-wave-all}
#| label: tbl-wave-all
#| tbl-cap: "Wave-by-wave OLS: econ to DV for all outcomes, Korea and Taiwan"

wave_wide <- mr$wave_by_wave |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s", estimate, sig),
    year = case_when(
      country == "Korea"  ~ yr_kr[as.character(wave)],
      country == "Taiwan" ~ yr_tw[as.character(wave)]
    )
  ) |>
  select(country, year, dv_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell) |>
  arrange(country, year)

wave_wide |>
  kbl(
    col.names = c("Country", "Year", names(wave_wide)[-(1:2)]),
    booktabs = TRUE, align = paste0("ll", strrep("r", ncol(wave_wide) - 2))
  ) |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  add_footnote(
    "OLS coefficient on econ_index. Controls: age, gender, education, urban/rural, political interest. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

# Appendix C: Pooled Individual-DV Models

@tbl-pooled reports pooled OLS results (with wave fixed effects) for each individual
dependent variable in Korea and Taiwan. This complements the wave-by-wave table by
presenting a single summary estimate for each outcome across the full survey period.

```{r tbl-pooled}
#| label: tbl-pooled
#| tbl-cap: "Pooled OLS: econ$\\to$individual DVs, Korea and Taiwan"

mr$pooled_indiv |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(country, dv_label, cell, n) |>
  pivot_wider(names_from = country, values_from = c(cell, n), names_sep = "_") |>
  select(dv_label, cell_Korea, n_Korea, cell_Taiwan, n_Taiwan) |>
  kbl(
    col.names = c("Dependent variable", "$\\beta$ (SE)", "N", "$\\beta$ (SE)", "N"),
    booktabs = TRUE, align = "lrrrr", escape = FALSE,
    format.args = list(big.mark = ",")
  ) |>
  add_header_above(c(" " = 1, "Korea" = 2, "Taiwan" = 2)) |>
  column_spec(1, width = "4.5cm") |>
  column_spec(c(2, 4), width = "2.8cm") |>
  footnote(
    general = "OLS b (econ_index) with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE
  ) |>
  kable_styling(latex_options = "hold_position")
```

# Appendix D: Cross-Country Interaction Test

@tbl-cross-country tests the H4 prediction that the Korea–Taiwan divergence in the
economic–quality relationship is statistically significant. Taiwan is the reference
category; a positive `econ_index:is_korea` interaction indicates that the positive
economic effect on satisfaction is stronger in Korea than in Taiwan; a negative
coefficient indicates the reverse (i.e., Taiwan shows a stronger negative effect on
quality, as expected under the critical-citizens pattern).

```{r tbl-cross-country}
#| label: tbl-cross-country
#| tbl-cap: "Cross-country interaction: econ x Korea, all five outcomes"

term_labels <- c(
  "econ_index"          = "Econ. eval.",
  "is_korea"            = "Korea (vs. Taiwan)",
  "econ_index:is_korea" = "Econ. x Korea"
)

xc_wide <- mr$cross_country |>
  filter(term %in% names(term_labels)) |>
  mutate(
    term_label = recode(term, !!!term_labels),
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(dv_label, term_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell)

xc_wide |>
  kbl(
    col.names = c("Term", names(xc_wide)[-1]),
    booktabs = TRUE, align = paste0("l", strrep("r", ncol(xc_wide)-1)),
    escape = FALSE
  ) |>
  add_footnote(
    "OLS. Taiwan = reference category. Pooled model with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

# Appendix E: Subgroup Heterogeneity

## E.1 Political Interest Subgroups

H5 predicts the satisfaction–quality decoupling is not concentrated in a single
demographic group. @tbl-polint presents results by political interest tertile
(low, medium, high) for Korea. All subgroups should show positive econ → satisfaction
and near-zero or negative econ → quality.

```{r tbl-polint}
#| label: tbl-polint
#| tbl-cap: "Korea: econ to DV by political interest subgroup"

# 3 DVs x 2 groups: pivot to wide, just show cell (estimate + SE + stars)
polint_wide <- mr$polint_subgroup |>
  filter(country == "Korea") |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(group, dv_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell)

polint_wide |>
  kbl(
    col.names = c("Group", names(polint_wide)[-1]),
    booktabs = TRUE, align = paste0("l", strrep("r", ncol(polint_wide) - 1))
  ) |>
  add_footnote(
    "OLS coefficient on econ_index with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = "hold_position")
```

## E.2 Winner–Loser Subgroups

Electoral winners and losers often display different patterns of democratic support.
@tbl-winner tests whether the satisfaction–quality decoupling is symmetric across
winners (supporters of the governing party) and losers (supporters of the opposition).

```{r tbl-winner}
#| label: tbl-winner
#| tbl-cap: "Korea and Taiwan: econ to DV by electoral winner/loser status"

# 2 DVs x 2 groups x 2 countries: country + group + 2 cell cols = 4 cols
winner_wide <- mr$winner_loser_subgroup |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error),
    dv_label = recode(dv_label,
      "Dem always preferable"      = "Dem. always pref.",
      "Satisfaction with democracy" = "Sat. w/ democracy"
    )
  ) |>
  select(country, group, dv_label, cell) |>
  pivot_wider(names_from = dv_label, values_from = cell) |>
  arrange(country, group)

winner_tbl <- winner_wide |>
  kbl(
    col.names = c("Country", "Group", names(winner_wide)[-(1:2)]),
    booktabs = TRUE, align = paste0("ll", strrep("r", ncol(winner_wide) - 2))
  )

# column_spec before collapse_rows; guard against HTML column-merge mismatch
if (knitr::is_latex_output()) {
  winner_tbl <- winner_tbl |>
    column_spec(seq(3, ncol(winner_wide)), width = "3.5cm")
}

winner_tbl |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  footnote(
    general = "OLS coefficient on econ_index with wave FE and standard controls. SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

# Appendix F: Authoritarian Rejection — Full Results

## F.1 Pooled Models

@tbl-auth-pooled reports pooled OLS estimates of econ → authoritarian rejection
(index and four component items) for Korea and Taiwan. The sign test is: if
the Korean decoupling reflects democratic erosion, better economic evaluations should
weaken rejection of authoritarian alternatives. If it reflects only a
performance-satisfaction response, they should strengthen rejection.

```{r tbl-auth-pooled}
#| label: tbl-auth-pooled
#| tbl-cap: "Pooled econ$\\to$authoritarian rejection: index and component items"

auth_pooled_all <- bind_rows(
  ar$pooled_index |> mutate(dv_label = "Auth. rejection index"),
  ar$pooled_items
) |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  ) |>
  select(country, dv_label, cell, n) |>
  pivot_wider(names_from = country, values_from = c(cell, n), names_sep = "_")

auth_pooled_all |>
  select(dv_label, cell_Korea, n_Korea, cell_Taiwan, n_Taiwan) |>
  kbl(
    col.names = c("Dependent variable", "$\\beta$ (SE)", "N", "$\\beta$ (SE)", "N"),
    booktabs = TRUE, align = "lllll", escape = FALSE,
    format.args = list(big.mark = ",")
  ) |>
  add_header_above(c(" " = 1, "Korea" = 2, "Taiwan" = 2)) |>
  add_footnote(
    "Pooled OLS b (econ_index) with wave FE and standard controls. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = "hold_position")
```

## F.2 Wave-by-Wave Authoritarian Rejection

@tbl-auth-wave presents the wave-by-wave results for the authoritarian rejection
index and its four component items in both countries.

```{r tbl-auth-wave}
#| label: tbl-auth-wave
#| tbl-cap: "Wave-by-wave econ to authoritarian rejection index"

ar$wave_by_wave |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error),
    year = case_when(
      country == "Korea"  ~ yr_kr[as.character(wave)],
      country == "Taiwan" ~ yr_tw[as.character(wave)]
    ),
    n_fmt = formatC(n, format = "d", big.mark = ",")
  ) |>
  select(country, year, cell, n_fmt) |>
  kbl(
    col.names = c("Country", "Year", "$\\beta$ (SE)", "N"),
    booktabs = TRUE, align = "llrr", escape = FALSE,
    format.args = list(big.mark = ",")
  ) |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  add_footnote(
    "OLS coefficient on econ_index, auth. rejection index as DV. Controls: age, gender, education, urban/rural, political interest. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = "hold_position", font_size = 9)
```

## F.3 Cross-Country Interaction

@tbl-auth-xc tests whether Korea's positive econ → auth rejection relationship is
significantly different from Taiwan's, using the same cross-country interaction
framework as Appendix D.

```{r tbl-auth-xc}
#| label: tbl-auth-xc
#| tbl-cap: "Cross-country interaction: econ $\\times$ Korea on authoritarian rejection index"

auth_term_labels <- c(
  "econ_index"          = "Econ. eval.",
  "is_korea"            = "Korea (1 = Korea)",
  "econ_index:is_korea" = "Econ. $\\times$ Korea"
)

ar$cross_country |>
  filter(term %in% names(auth_term_labels)) |>
  mutate(
    term_label = recode(term, !!!auth_term_labels),
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s", estimate, sig),
    se   = sprintf("(%.3f)", std.error)
  ) |>
  select(term_label, cell, se) |>
  kbl(
    col.names = c("Term", "$\\beta$", "SE"),
    booktabs = TRUE, align = "lrr", escape = FALSE
  ) |>
  add_footnote(
    "OLS pooled with wave FE and standard controls. Auth. rejection index as DV. Taiwan = reference category. *p < .05, **p < .01, ***p < .001.",
    notation = "none"
  ) |>
  kable_styling(latex_options = "hold_position", font_size = 9)
```

# Appendix G: Item Specificity

## G.1 All Outcomes Side-by-Side

@tbl-item-spec-kr and @tbl-item-spec-tw present the complete coefficient evidence
for Korea and Taiwan respectively. For each outcome: $\beta$ = unstandardized OLS
coefficient on `econ_index`; SE = HC2 robust standard error; $p$ = two-sided
*p*-value; $\beta^\star$ = standardized coefficient (units: outcome SD per SD of
predictor); $R^2$ = model $R^2$ (including wave FE and all controls). The key
diagnostic is that the abstract normative item ("Dem always preferable") is the
only row where Korea's coefficient is near zero while all satisfaction and
authoritarian rejection rows show consistent, significant effects.

```{r tbl-item-spec-kr}
#| label: tbl-item-spec-kr
#| tbl-cap: "Evidence locker: econ$\\to$all DVs, Korea. All statistics from pooled OLS with wave FE and standard controls."

ir$all_dvs |>
  filter(country == "Korea") |>
  arrange(category, dv) |>
  mutate(
    sig      = sfn(p.value),
    beta_fmt = sprintf("%.3f%s", estimate, sig),
    se_fmt   = sprintf("%.3f", std.error),
    p_fmt    = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value)),
    bstd_fmt = sprintf("%.3f", beta_std),
    r2_fmt   = sprintf("%.3f", r_squared),
    n_fmt    = formatC(n, format = "d", big.mark = ",")
  ) |>
  select(category, dv_label, beta_fmt, se_fmt, p_fmt, bstd_fmt, r2_fmt, n_fmt) |>
  kbl(
    col.names = c("Category", "Outcome", "$\\beta$", "SE", "$p$",
                  "$\\beta^\\star$", "$R^2$", "$N$"),
    booktabs = TRUE, align = "llrrrrrl", escape = FALSE
  ) |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  column_spec(1, width = "3cm") |>
  column_spec(2, width = "4.5cm") |>
  footnote(
    general = "$\\\\beta$ = unstandardized; $\\\\beta^\\\\star$ = standardized (SD units); $R^2$ includes wave FE and controls. HC2 robust SE. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE,
    escape = FALSE
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

```{r tbl-item-spec-tw}
#| label: tbl-item-spec-tw
#| tbl-cap: "Evidence locker: econ$\\to$all DVs, Taiwan. All statistics from pooled OLS with wave FE and standard controls."

ir$all_dvs |>
  filter(country == "Taiwan") |>
  arrange(category, dv) |>
  mutate(
    sig      = sfn(p.value),
    beta_fmt = sprintf("%.3f%s", estimate, sig),
    se_fmt   = sprintf("%.3f", std.error),
    p_fmt    = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value)),
    bstd_fmt = sprintf("%.3f", beta_std),
    r2_fmt   = sprintf("%.3f", r_squared),
    n_fmt    = formatC(n, format = "d", big.mark = ",")
  ) |>
  select(category, dv_label, beta_fmt, se_fmt, p_fmt, bstd_fmt, r2_fmt, n_fmt) |>
  kbl(
    col.names = c("Category", "Outcome", "$\\beta$", "SE", "$p$",
                  "$\\beta^\\star$", "$R^2$", "$N$"),
    booktabs = TRUE, align = "llrrrrrl", escape = FALSE
  ) |>
  collapse_rows(columns = 1, latex_hline = "major") |>
  column_spec(1, width = "3cm") |>
  column_spec(2, width = "4.5cm") |>
  footnote(
    general = "$\\\\beta$ = unstandardized; $\\\\beta^\\\\star$ = standardized (SD units); $R^2$ includes wave FE and controls. HC2 robust SE. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE,
    escape = FALSE
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

# Appendix H: Identity-Signaling Moderation Tests

The identity-signaling interpretation predicts that the economic → abstract preference
relationship should be stronger among Taiwanese respondents who perceive greater
threat from China, and among those with stronger national pride. Two moderation
tests address this. Both use pooled OLS with wave fixed effects and standard controls,
restricted to the Taiwan sample.

## H.1 China threat moderation

`china_harm_bin` is a binary indicator (1 = respondent views China's influence
as harmful; 0 = beneficial or neutral), derived from the ABS item asking whether
China's growing influence is beneficial or harmful for the country. The interaction
term `econ_index × china_harm_bin` tests whether the critical citizens pattern is
stronger among respondents who perceive China as a threat (Taiwan only, pooled
across waves).

```{r tbl-china-mod}
#| label: tbl-china-mod
#| tbl-cap: "China threat moderation: econ$\\times$China-harm on democratic support (Taiwan)"

cm <- mr$china_moderation |>
  filter(is.na(group), !is.na(term)) |>   # interaction model rows
  mutate(
    term_label = recode(term,
      "econ_index"              = "Econ. eval.",
      "china_harm_bin"          = "China harmful (0/1)",
      "econ_index:china_harm_bin" = "Econ. \\times China harmful"
    ),
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error),
    n_fmt = formatC(n, format = "d", big.mark = ",")
  )

dv_order <- c("Quality", "Democracy always preferable", "Extent democratic")

n_per_dv <- cm |>
  group_by(dv) |>
  slice(1) |>
  select(dv, n)

cm |>
  select(dv, term_label, cell) |>
  pivot_wider(names_from = dv, values_from = cell) |>
  select(term_label, all_of(dv_order)) |>
  kbl(
    col.names = c("Term", "Quality", "Dem. always pref.", "Extent dem."),
    booktabs = TRUE, align = "lrrr", escape = FALSE
  ) |>
  add_footnote(
    paste0(
      "OLS pooled with wave FE and standard controls. Taiwan only. ",
      "N = ", paste(formatC(n_per_dv$n, format="d", big.mark=","), collapse=" / "), " by DV. ",
      "SE in parentheses. *p < .05, **p < .01, ***p < .001."
    ),
    notation = "none"
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

## H.2 National pride moderation

`nat_proud_n` is the normalized national pride item (0–1; higher = more proud).
The moderation is estimated separately for Korea and Taiwan.

```{r tbl-pride-mod}
#| label: tbl-pride-mod
#| tbl-cap: "National pride moderation: econ$\\times$national pride on democratic support"

pi <- mr$pride_interaction |>
  filter(term == "econ_index:nat_proud_n") |>
  mutate(
    sig  = sfn(p.value),
    cell = sprintf("%.3f%s (%.3f)", estimate, sig, std.error)
  )

dv_order2 <- c("Quality", "Democracy always preferable", "Extent democratic")

pi |>
  select(country, dv, cell, n) |>
  pivot_wider(names_from = dv, values_from = cell) |>
  select(country, all_of(dv_order2), n) |>
  mutate(n = formatC(n, format = "d", big.mark = ",")) |>
  kbl(
    col.names = c("Country", "Quality", "Dem. always pref.", "Extent dem.", "N"),
    booktabs = TRUE, align = "lrrrr", escape = FALSE
  ) |>
  footnote(
    general = "OLS pooled with wave FE and standard controls. Interaction term only (econ_index × nat_proud_n). SE in parentheses. *p < .05, **p < .01, ***p < .001.",
    general_title = "",
    threeparttable = TRUE
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```


# Appendix I: Index Reliability and Construct Diagnostics

```{r setup-reliability}
#| include: false
rel <- readRDS(file.path(results_dir, "reliability_results.rds"))
```

## I.1 Cronbach's Alpha and McDonald's Omega

@tbl-reliability reports internal consistency for the economic evaluation index
(6 items) and the authoritarian rejection index (4 items) by country and wave.
Wave 2 authoritarian rejection results are missing because those items were not
administered in the 2006 wave of the ABS.

```{r tbl-reliability}
#| label: tbl-reliability
#| tbl-cap: "Index reliability: Cronbach's $\\alpha$ and McDonald's $\\omega$ by country and wave"

rel_econ <- rel$alpha_econ |>
  left_join(rel$omega_econ |> select(country, wave, omega = omega_t),
            by = c("country", "wave")) |>
  mutate(index = "Econ. eval. (6 items)")

rel_auth <- rel$alpha_auth |>
  left_join(rel$omega_auth |> select(country, wave, omega = omega_t),
            by = c("country", "wave")) |>
  mutate(index = "Auth. rejection (4 items)")

bind_rows(rel_econ, rel_auth) |>
  mutate(
    alpha_fmt = ifelse(is.na(alpha) | n == 0, "—", sprintf("%.3f", alpha)),
    omega_fmt = ifelse(is.na(omega) | n == 0, "—", sprintf("%.3f", omega)),
    n_fmt     = ifelse(n == 0, "—", formatC(n, format = "d", big.mark = ","))
  ) |>
  select(index, country, wave, year, n_fmt, alpha_fmt, omega_fmt) |>
  kbl(
    col.names = c("Index", "Country", "Wave", "Year", "N", "$\\alpha$", "$\\omega$"),
    booktabs = TRUE, align = "llrrrrr", escape = FALSE
  ) |>
  collapse_rows(columns = c(1, 2), latex_hline = "major") |>
  footnote(
    general = "$\\\\alpha$ = Cronbach's alpha; $\\\\omega$ = McDonald's omega (total). Wave 2 auth. rejection is NA: items not administered in ABS 2006.",
    general_title = "",
    threeparttable = TRUE,
    escape = FALSE
  ) |>
  kable_styling(latex_options = c("hold_position", "scale_down"), font_size = 9)
```

## I.2 Abstract vs. Substantive Item Correlations

@tbl-polychoric reports the Pearson correlation between the abstract preference
item ("democracy is always preferable") and each of the four authoritarian rejection
items, pooled across waves by country. Negative correlations are expected: higher
abstract endorsement correlates with stronger rejection of authoritarian alternatives,
but the modest magnitudes (|r| ≈ 0.17–0.20) confirm the two constructs are
related but empirically distinct.

```{r tbl-polychoric}
#| label: tbl-polychoric
#| tbl-cap: "Pearson r: abstract preference vs. authoritarian rejection items, by country"

item_labels <- c(
  strongman_reject_n   = "Reject: strongman rule",
  military_reject_n    = "Reject: military rule",
  expert_reject_n      = "Reject: expert/technocratic rule",
  singleparty_reject_n = "Reject: single-party rule"
)
auth_items <- names(item_labels)

# Extract first row (dem_always_preferable) vs auth columns from each country's cor_mat
poly_tbl <- purrr::map_dfr(rel$polychoric_by_country, function(x) {
  row_vals <- x$cor_mat["dem_always_preferable", auth_items]
  tibble::tibble(
    country = x$country,
    auth_item = auth_items,
    r = as.numeric(row_vals)
  )
})

poly_tbl |>
  mutate(
    item_label = item_labels[auth_item],
    r_fmt = sprintf("%.3f", r)
  ) |>
  select(item_label, country, r_fmt) |>
  pivot_wider(names_from = country, values_from = r_fmt) |>
  kbl(
    col.names = c("Authoritarian rejection item", "Korea", "Taiwan"),
    booktabs = TRUE, align = "lrr"
  ) |>
  footnote(
    general = "Pearson r between dem_always_preferable and each auth rejection item (both normalized 0–1). Pooled across all waves within country, complete cases only.",
    general_title = "",
    threeparttable = TRUE
  ) |>
  kable_styling(latex_options = "hold_position")
```


# References {.unnumbered}

::: {#refs}
:::
