---
title: "The Satisfaction Paradox: Economic Performance and the Structure of Democratic Support in South Korea"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
    corresponding: true
date: now
date-format: "MMMM D, YYYY [at] h:mm A"
format:
  pdf:
    documentclass: article
    geometry: margin=1in
    fontsize: 12pt
    linestretch: 2
    number-sections: true
    toc: false
    cite-method: citeproc
    include-in-header:
      text: |
        \AtBeginDocument{\author{Jeffrey Stark \\[0.3em] \normalsize Department of Sociology, Yonsei University}}
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
project_root <- "/Users/jeffreystark/Development/Research/paper-bank"
paper_dir    <- file.path(project_root, "papers/03b_sk_satisfaction_paradox")
fig_dir      <- file.path(paper_dir, "analysis", "figures")
```

# Abstract {.unnumbered}

```{=latex}
\setstretch{1}
```

Democratic consolidation theory predicts that satisfaction with democratic performance and normative commitment to democracy reinforce each other, with economic performance as the engine. South Korean aggregate trends appear to confirm this expectation: over two decades, both democratic satisfaction and abstract democratic preference have risen. Six waves of Asian Barometer Survey data (2003--2022, *n* = 7,603) reveal that the co-movement is coincidental. Wave-by-wave OLS regressions show that economic evaluations powerfully predict satisfaction with democracy ($\beta$ = 0.26--0.46, *p* < 0.001 in every wave) but show a shifting and ultimately reversed relationship with normative democratic preference: positive in 2003, null in the middle waves, and significantly negative by 2015--2019. The pooled relationship is zero---not because the connection is absent, but because a positive early relationship cancels a negative later one. The 30-percentage-point rise in the belief that democracy is "always preferable"---from 49 to 78 percent---is a uniform period effect visible in every age group and birth cohort, driven by forces entirely unconnected to economic performance. The decoupling extends further. On substantive measures---rejection of strongman rule, military governance, single-party systems---economically comfortable Koreans are marginally *more tolerant* of authoritarian alternatives, not less. The performance-legitimation mode runs through the entire structure of Korean democratic support: economic comfort increases satisfaction with democracy, shows a relationship with principled democratic commitment that reversed from positive to negative over the observation window---suggesting an emerging critical citizens dynamic invisible in the pooled estimates---and weakly increases openness to the authoritarian governance modes that the consolidation literature treats as the clearest markers of normative depth. A parallel analysis of Taiwan (*n* = 9,042) sharpens the point: Taiwan exhibits no period-level rise in abstract preference but shows strong individual-level economic sensitivity---the structural inverse of the Korean pattern, consistent with a context where democracy carries identity weight that makes economic conditions relevant to normative evaluation. South Korea's December 2024 martial law crisis---resolved within six hours by institutional mechanisms and organized political opposition---is consistent with a democracy whose demonstrated resilience is institutional, though the attitudinal data predate the crisis and cannot establish the counterfactual.

**Keywords:** democratic consolidation, democratic satisfaction, performance legitimacy, authoritarian rejection, South Korea, Taiwan, Asian Barometer Survey

```{=latex}
\setstretch{2}
```

# Introduction

On the evening of December 3, 2024, President Yoon Suk Yeol appeared on national television to declare emergency martial law in South Korea, citing the need to protect the country from "anti-state forces" and "North Korean communist sympathizers" allegedly operating through the opposition-controlled National Assembly. Troops were dispatched to the parliament building. Political activity was banned. The press was placed under the authority of the Martial Law Command. It was the first declaration of martial law since Chun Doo-hwan's 1980 coup---and it lasted barely six hours. Lawmakers pushed past soldiers, convened an emergency session, and voted unanimously to overturn the decree. Yoon was impeached within days. In February 2026, a Seoul court sentenced him to life in prison for insurrection [@Kim2024-ud; @Hankyoreh2026-aa].

South Korean democracy, by any institutional measure, held. The question is what held it.

The standard answer points to institutional design: constitutional safeguards, an independent judiciary, a legislature willing to act. These matter. But institutional mechanisms are activated by people who choose to activate them, and the nature of that choice---whether it reflects principled democratic conviction or instrumental calculation about which system works---has consequences for how resilient the democracy proves against subtler threats than a clumsy late-night putsch. The mass public opinion data spanning the two decades before the crisis suggest a more complicated picture of Korean democratic commitment than the institutional response alone conveys.

The empirical basis for this claim is a pattern documented across six waves of the Asian Barometer Survey. Between 2003 and 2022, democratic satisfaction in South Korea rose steadily---roughly five percentage points per decade. Over the same period, the belief that democracy is always preferable rose even more sharply, from 49 to 78 percent of respondents. System support declined modestly. A researcher examining these aggregate trends might conclude that South Korea is consolidating on schedule---rising satisfaction accompanied by rising democratic preference, just as the literature predicts. The individual-level evidence tells a different story.

The individual-level evidence contradicts a foundational expectation of the consolidation literature. From Easton's [-@Easton1965-tz] distinction between diffuse and specific support through Linz and Stepan's [-@Linz1996-qs] insistence that democracy must become "the only game in town," the dominant theoretical framework predicts that economic performance builds both satisfaction and normative commitment---that the co-movement observable at the aggregate level reflects a causal mechanism operating on individuals. The Korean data show that it does not. Economic evaluations powerfully and consistently predict democratic satisfaction across all six waves. Their relationship with normative democratic preference is more complex: positive in 2003, null in the middle waves, and significantly negative by 2015---a sign reversal that the pooled null ($\beta$ = --0.036) conceals. The 30-percentage-point aggregate rise in "always preferable" is a period effect---visible in every age group, every birth cohort, and every education level simultaneously---driven by forces entirely unconnected to economic performance. The aggregate co-movement masks an individual-level structure far more complex than either simple decoupling or the co-movement the consolidation literature predicts.

But the story extends further than the headline finding suggests, and in a more troubling direction. The decoupling between economic performance and democratic commitment is not uniform across measures of commitment. On substantive indicators---rejection of strongman rule, military governance, expert rule, single-party systems---economically comfortable Koreans are marginally *more tolerant* of authoritarian alternatives, not less. The performance-legitimation mode does not stop at satisfaction. It penetrates the authoritarian rejection battery, weakly but consistently, in the direction of reduced vigilance.

The distinction carries consequences for how we assess democratic resilience---and the December 2024 crisis illustrates the point. Martial law was reversed by institutional mechanisms: legislators who refused military authority, judges who asserted constitutional supremacy, a military that chose institutional loyalty over presidential command. Not by mass normative mobilization organized around abstract democratic principle. The system held because constitutional safeguards activated as designed. The attitudinal data suggest the mass public was not the primary agent of that activation.

A parallel analysis of Taiwan (*n* = 9,042 across the corresponding six ABS waves) provides comparative leverage---not as a co-equal case study but as a diagnostic tool. Taiwan shares South Korea's developmental state heritage, Confucian cultural context, and third-wave democratization experience. The Taiwanese data reveal the structural inverse of the Korean pattern. Taiwan exhibits no period-level rise in abstract preference---it has been flat for two decades---but shows strong individual-level economic sensitivity: economically comfortable Taiwanese are substantially more likely to endorse democracy as always preferable ($\beta$ = 0.286, *p* < 0.001, pooled), a critical citizens pattern that grows stronger over time. On substantive authoritarian rejection, however, Taiwan's economic sensitivity is actually weaker than Korea's. The abstract item diverges sharply across the two countries (cross-country interaction $\beta$ = --0.278, *p* < 0.001); the substantive items do not. That pattern suggests the abstract measure is capturing something specific to Taiwan---where democratization became constitutive of national identity in ways it did not in Korea---rather than a deeper normative commitment generalizable across contexts.

The paper makes three contributions. First, it documents the satisfaction paradox with systematic longitudinal evidence spanning two decades, showing precisely where economic evaluations reach into the structure of Korean democratic support and where they do not. Second---and this is the contribution that emerged from the diagnostic analysis rather than the initial theoretical framework---it demonstrates that the decoupling is item-specific, concentrated on the abstract normative measure rather than on substantive authoritarian rejection. Korean democratic support has more substance than the standard consolidation indicators suggest. Third, the Taiwan comparison raises a measurement question about the field's most widely used indicator of normative democratic commitment, suggesting it captures democratic identity signaling in contexts where democracy and national identity have fused, rather than the deep principled attachment Easton [-@Easton1965-tz] theorized as diffuse support.

Section 2 develops the theoretical framework, distinguishing abstract from substantive democratic commitment within the consolidation literature's co-movement assumption. Section 3 describes the data and analytic strategy. Section 4 presents the core results, the diagnostic findings, the cross-country comparison, and robustness checks. Section 5 interprets the findings in light of the Yoon crisis and the measurement implications.


# Literature Review and Theoretical Framework

## Democratic consolidation and the co-movement assumption

The analytical scaffolding for democratic consolidation rests on a distinction Easton [-@Easton1965-tz] drew between two modes of citizen support. Diffuse support is a generalized attachment to the political system that persists regardless of specific policy outcomes---the reservoir that keeps democracies afloat when governments stumble. Specific support is the day-to-day evaluation: the economy, government competence, service delivery. The two are conceptually independent but, in Easton's account, not causally so. Sustained specific support was expected to deposit gradually into the reservoir of diffuse support, building legitimacy over time [@Easton1975-bg].

This expectation hardened into a prediction. Linz and Stepan [-@Linz1996-qs] gave it a memorable formulation---democracy must become "the only game in town"---but the underlying logic was more consequential than the phrase. Democratic norms internalize through democratic practice. Citizens who experience accountable governance develop principled attachment to the system that produces it. Diamond [-@Diamond1999-wd] theorized a parallel deepening process. The prediction common to these accounts, sometimes stated explicitly but more often operating as a tacit assumption, is co-movement. Satisfaction and commitment should trend together. The term encompasses two related predictions that the literature does not always distinguish. At the aggregate level, satisfaction and normative commitment should trend in the same direction over time---both rising as democracies mature. At the individual level, the same predictors---economic evaluations chief among them---should be significantly associated with both outcomes. The empirical strategy tests both: §4.1 documents aggregate trends; §§4.2--4.3 test individual-level associations. Korean aggregate trends appear to satisfy the first prediction---satisfaction and abstract preference both rose---but the individual-level analysis reveals the co-movement to be coincidental: the causal mechanism the literature posits (economic performance building both satisfaction and commitment) operates on satisfaction alone. The individual-level violation is the more analytically consequential finding because it shows that apparent aggregate co-movement can mask complete causal decoupling.

Norris [-@Norris1999-sn; -@Norris2011-oc] complicated the picture by documenting citizens who were simultaneously more committed to democratic ideals and more dissatisfied with democratic practice---but note what remained intact. The normative dimension was secure and strengthening. Only performance evaluation grew more demanding. Critical citizenship, as Norris theorized it, is a sign of democratic health. The more radical challenge---that satisfaction and normative commitment might rise together yet be driven by entirely unconnected mechanisms, with the economic engine of legitimation reaching satisfaction but not commitment---has received surprisingly little systematic attention. Whether a mature democracy can exhibit apparent co-movement at the aggregate level while its individual-level causal structure reveals complete decoupling is a question the literature has largely left unasked.


## Performance legitimacy and the East Asian context

The co-movement assumption grew out of the Western experience, where democratic legitimacy was established before or alongside industrialization. Civic education, constitutional veneration, foundational political narratives---these gave normative commitment a basis independent of economic performance. East Asia's democratization experience upends this sequence. In South Korea, Taiwan, and the region's other third-wave democracies, rapid economic development under authoritarian auspices preceded democratic transition. Citizens had already experienced effective governance without democracy. The developmental state worked. Democracy's legitimacy was therefore inevitably comparative---measured, at least in part, against the system it replaced.

Park and Shin [-@Park2006-bv] identified the resulting configuration: widespread democratic preference in Korea coexisting with authoritarian nostalgia, limited procedural understanding, and what they termed "incomplete democratization." Chu and colleagues pursued the point across the region [@Chu2008-fk; @Chu2020-ir], documenting citizens who define democracy in substantive rather than procedural terms. Economic outcomes and good governance take priority over elections and civil liberties. When citizens have experienced competent authoritarianism, democracy must justify itself on output. Mauk [-@Mauk2022-dm] confirmed that even trust in democratic institutions across East and Southeast Asian democracies is driven primarily by short-term economic performance evaluations and incumbent support rather than by the long-term liberal democratic value orientations that consolidation theory posits as the deeper foundation. Bomhoff and Gu [-@Bomhoff2012-oq] pressed the broader point that the Inglehart-Welzel self-expression values framework---which grounds democratic commitment in post-materialist cultural change---simply does not capture how democratic attitudes operate in East Asia.

The implication for the co-movement assumption is significant but has not been fully drawn out. If citizens evaluate their political system primarily on what it delivers rather than how it delivers, satisfaction and normative commitment need not track one another. A citizen can be highly satisfied with democratic performance while developing no deeper attachment to democratic principles, because the attachment was never principled in the first place. It was conditional on performance and always was.


## Abstract endorsement versus substantive anti-authoritarianism

The consolidation literature operationalizes normative commitment primarily through abstract endorsement items. The "democracy is always preferable" question---asking respondents to choose among democracy as always preferable, authoritarian government as sometimes preferable, and regime type as irrelevant---has become the field's standard indicator of Easton's diffuse support. Linz and Stepan's "only game in town" is typically measured through this item or close variants. Foa and Mounk [-@Foa2017-oq] built their deconsolidation argument on generational trends in responses to it. The item does important work.

But it also attracts criticism. Inglehart [-@Inglehart2003-do] warned that respondents may pay mere "lip service" to abstract pro-democracy questions, producing "questionnaire democrats" whose endorsement carries little substantive weight [see also @Schedler2007-pv]. More recently, the measurement debate has sharpened. Tai, Hu, and Solt [-@Tai2022-yr] demonstrated that incorporating measurement uncertainty into Claassen's [-@Claassen2020-hk] influential latent variable estimates of democratic support eliminates the evidence that support helps democracy survive---a finding that underscores just how much turns on measurement choices. Hu [-@Hu2025-yj] extended this critique, arguing that one-dimensional latent variables "do not capture the full flavor of democratic support"---that support is inherently multidimensional and cannot be collapsed into a single index without losing analytically consequential variation. Most comprehensively, Claassen and colleagues [-@Claassen2025-xl] developed a new multidimensional measurement framework for democratic support using 17 items across multiple survey programs, finding that items cohere readily into unidimensional measures in higher-income countries but that "in less-developed settings, citizens are more likely to oppose certain of these rights and institutions even while they support liberal democracy as a whole."

The present analysis contributes to this measurement debate at the item level rather than the latent-variable level. Abstract endorsement and substantive anti-authoritarianism are conceptually distinct, and conflating them risks misdiagnosing the structure of democratic support. The ABS authoritarian rejection battery---which asks respondents to evaluate specific governance alternatives: strongman rule, military government, expert rule, single-party systems---taps a different dimension. It requires cognitive engagement with concrete institutional alternatives rather than an abstract statement of principle. A citizen who rejects strongman rule is evaluating a specific governance mode on its merits. A citizen who affirms that democracy is always preferable may be doing the same---or may be expressing an identity, invoking a socialized norm, signaling membership in a political community where democratic self-description carries weight.

A clarification about the scope of this distinction is warranted. Rejecting authoritarian alternatives is not equivalent to affirming democratic processes. A citizen who opposes military rule need not value judicial independence, minority rights protection, or legislative deliberation---the affirmative content of democratic governance. The authoritarian rejection battery captures what citizens oppose, not what they endorse. The construct is better understood as defensive anti-authoritarianism than as positive democratic commitment. This matters for interpretation. A citizenry high on authoritarian rejection but low on abstract endorsement may be one that knows what it does not want without having articulated what it does---a configuration with real protective value against overt authoritarian seizure but uncertain resilience against democratic erosion that never presents itself as authoritarianism.

The distinction generates different predictions about democratic resilience [@Mauk2020-cn]. High abstract endorsement with low substantive rejection suggests fragile commitment---cultural performance without evaluative content. Low abstract endorsement with high substantive rejection suggests defensive anti-authoritarianism more robust than standard indicators imply. Chu, Huang, Lagos, and Mattes [@Chu2020-ir] recognized the multidimensional character of democratic support in their ABS decomposition but did not isolate the abstract-substantive divergence as analytically consequential. The analysis presented here does.

The framing generates testable predictions. If the Korean decoupling reflects genuinely shallow foundations, it should appear across both abstract and substantive measures. If it is instead specific to abstract commitment, economic evaluations should be null on the abstract item but significant on the substantive items.


## Hypotheses

H1 (Performance--satisfaction linkage): Economic evaluations are positively and significantly associated with democratic satisfaction in both Korea and Taiwan.

H2 (Abstract decoupling): In South Korea, economic evaluations have no significant relationship with the abstract normative preference that democracy is always preferable.

H3a (Substantive null): If the performance-legitimation mode is confined to satisfaction, economic evaluations should bear no significant relationship to rejection of specific authoritarian alternatives---strongman rule, military governance, and single-party systems.

H3b (Substantive penetration): If, alternatively, the instrumental orientation penetrates the full structure of democratic support, economic comfort should be associated with *reduced* rejection of authoritarian governance modes.

H4 (Item specificity): The Korea--Taiwan divergence on the abstract normative item is substantially larger than the divergence on substantive authoritarian rejection items.

H5 (Cross-country divergence): The relationship between economic evaluations and abstract democratic preference is significantly different between Korea and Taiwan.

H6 (Universality): The abstract decoupling holds across age groups, education levels, and model specifications in Korea, indicating a society-wide pattern.


# Data and Methods

## Data

The analysis draws on the Asian Barometer Survey, a comparative program administering nationally representative face-to-face interviews across East and Southeast Asia since 2001 [@Asian-Barometer-Survey2023-uh]. Six waves are currently available. South Korea and Taiwan are among the few countries surveyed in all six, providing an unusually long panel of repeated cross-sections. The Korean sample comprises 7,603 respondents; the Taiwanese 9,042. Both use multi-stage stratified probability sampling. Survey weights are not applied in the main analysis, following convention in the ABS literature; the Korean and Taiwanese samples match census margins on age, gender, education, and region within the tolerances reported in the ABS technical documentation. Weighted estimates produce substantively identical results.

## Dependent variables

The analysis tracks four categories of dependent variable, each tapping a conceptually distinct dimension of democratic support. Primary specifications treat items separately rather than combining them into indices---a decision that proves analytically consequential, since the items respond to economic evaluations in different and theoretically informative ways.

The *satisfaction* dimension includes two items corresponding to Easton's [-@Easton1965-tz] specific support: satisfaction with the way democracy works and satisfaction with the national government. These are reported separately because government satisfaction is so tightly coupled to incumbent evaluation that combining it with democratic satisfaction obscures the distinction between system-level and government-level assessment.

The *abstract normative* dimension is captured by agreement that democracy is always preferable to any other form of government---the cleanest unconditional normative indicator available across all six waves. A second item asks respondents to assess the extent to which their country is currently democratic; this is evaluative rather than normative and is treated as a distinct perceived-quality outcome.

The *substantive commitment* dimension draws on the ABS authoritarian rejection battery: rejection of strongman rule, military government, expert rule, and single-party systems (four items, all waves).[^rej] An authoritarian rejection index is computed as the mean of these items after normalization. These items require respondents to evaluate specific governance modes rather than endorse an abstract principle, and their inclusion constitutes the primary analytic addition relative to the initial analysis. Chu, Huang, Lagos, and Mattes [@Chu2020-ir] introduced the "authoritarian detachment" construct as a conceptually distinct dimension of democratic support, separate from both democratic preference and democratic satisfaction; the present analysis exploits this decomposition by testing whether the dimensions respond differently to economic evaluations.

[^rej]: I use "authoritarian rejection" to describe what Chu et al. [-@Chu2008-fk] termed "authoritarian detachment." Higher values on all items indicate stronger rejection of authoritarian alternatives.

Two supplementary regime legitimacy items---agreement that the political system deserves support and disagreement that the system needs major change---are available from Wave 3 onward. All items are normalized to 0--1 scales.

## Independent variable

The key independent variable is an economic evaluation index constructed from six ABS items spanning sociotropic and pocketbook assessments across current, retrospective, and prospective time horizons. Each is normalized to 0--1; the index is their mean. Including both levels is deliberate---the theoretical question is not which level of economic evaluation drives democratic attitudes but whether economic performance as a class of explanation links to normative commitment at all.

## Controls

All models include age (normalized), gender, education level (normalized), urban-rural residence, and political interest (normalized)---the major demographic and attitudinal correlates identified in the comparative literature [@Dalton2004-bq]. Institutional trust is excluded from the primary specification because it is plausibly endogenous to both economic evaluations and democratic satisfaction; it enters as a robustness check.

## Analytic strategy

The analysis proceeds in four stages. The starting point is visual: wave-level means for satisfaction, abstract commitment, and substantive commitment plotted over time to document the divergence.

The core engine is wave-by-wave OLS regression. For each wave in each country, parallel models predict satisfaction, abstract preference, and authoritarian rejection from the economic evaluation index with standard controls. The key comparison is the economic evaluation coefficient across these dependent variables within the same wave.

Running models within each wave rather than pooling with wave fixed effects allows the economic-attitude relationship to vary freely over time, making visible any temporal instability that pooled specifications would average away. Pooled models with centered wave interaction terms are estimated separately.

The cross-country comparison pools both countries and estimates:

$Y_i = \alpha + \beta_1 \text{EconIndex}_i + \beta_2 \text{Korea}_i + \beta_3 (\text{EconIndex}_i \times \text{Korea}_i) + \delta \text{Wave}_i + \gamma \mathbf{X}_i + \varepsilon_i$

where $\beta_3$ captures the differential effect of economic evaluations in Korea relative to Taiwan. That is, $\beta_3$ captures the difference in how strongly economic comfort predicts democratic attitudes, not the total country difference at any given level of economic evaluation. H5 predicts a significant $\beta_3$ on the abstract item. H4 predicts that $\beta_3$ is substantially larger for the abstract item than for the substantive authoritarian rejection items.

Standard errors are HC2 heteroscedasticity-robust throughout [@MacKinnon1985-rb]. The Asian Barometer Survey does not release primary sampling unit or stratum identifiers in its harmonized public release, precluding cluster-robust inference; HC2 provides a conservative alternative that does not require distributional assumptions about within-cluster correlation. Effective sample sizes vary across outcomes due to item nonresponse; exact *N* is reported in each table.

The repeated cross-sectional design cannot establish within-person causality. The analytic strategy exploits an asymmetric prediction to narrow the range of plausible confounding accounts, though it cannot eliminate them. If a single confound---such as dispositional optimism coloring both economic perceptions and democratic evaluations---drove the results, the bias should inflate $\beta_1$ similarly across outcome types. A pattern in which $\beta_1$ is large for satisfaction, null for abstract preference, and significant (in the opposite direction) for authoritarian rejection is inconsistent with any symmetric confound. The asymmetric-prediction test cannot rule out differential confounding across outcome types---social desirability bias, for instance, likely operates more strongly on the abstract 'always preferable' item, where a correct answer exists in a democratic society, than on the satisfaction or authoritarian rejection items. The cross-sectional associations documented here are therefore best read as structural regularities in the relationship between economic evaluations and democratic attitudes, not as causal estimates.

## AI assistance disclosure. 

The author used Claude (Anthropic, version 4.5) to assist with R programming, including data cleaning scripts, model specification, and debugging of analysis code. All analytical decisions, theoretical framing, and interpretation of results are the author's own. AI-generated code was reviewed, tested, and modified by the author before inclusion in the analysis pipeline.

# Results

## Descriptive trends: co-movement at the surface, decoupling beneath

Aggregate trends in South Korea present a picture that, at first glance, vindicates the consolidation literature. Democratic satisfaction rose over the two decades spanned by the Asian Barometer Survey: the satisfaction cluster increased from 0.456 to 0.525 between 2011 and 2022, roughly seven percentage points on the normalized scale. Government satisfaction exhibited the steepest trajectory, climbing from approximately 0.33 in 2011 to 0.53 by 2022.

Abstract democratic preference rose even more dramatically. The proportion of Korean respondents affirming that democracy is "always preferable" increased from 49.4 percent in 2003 to 77.5 percent in 2022---a 28-percentage-point gain that, taken at face value, suggests deepening normative commitment alongside rising satisfaction. Other indicators sent mixed signals: the democratic quality cluster drifted slightly downward (0.366 to 0.351 between 2011 and 2022), and system support declined modestly. But the headline numbers---satisfaction up, abstract preference up---look like co-movement.

The regression evidence will puncture this reading. But the aggregate rise itself has two features worth documenting first.

The rise is a period effect, not cohort replacement. Every age group tracks nearly identically: respondents aged 18--29 rose from 49.5 to 78.2 percent; respondents aged 60 and older rose from 46.5 to 76.2 percent. Birth cohort tracking confirms the pattern---every cohort, from those born before 1955 to those born after 1985, shows a rise of roughly 25--30 percentage points between their first and last observed waves. The aggregate trend is not an artifact of younger, more democratic cohorts entering the sample. Something changed for everyone. The Korean sample actually aged substantially over the observation window---respondents 60 and older grew from 15 to 30 percent of the sample---which would suppress the aggregate rise if it were driven by younger cohorts. It rose anyway, in every stratum simultaneously.

The period effect is unexplained by individual-level predictors. Wave fixed effects in the regression models are large: the Wave 6 coefficient corresponds to a 26-percentage-point increase in the probability of endorsing "always preferable" relative to Wave 1, and even the most saturated individual-level model---including education, political interest, news consumption, democratic values, and institutional attitudes---reduces this coefficient by only two-thirds. The aggregate trend is not reducible to compositional shifts in the variables the survey measures.

Taiwan presents the structural inverse. Abstract preference there has been flat for two decades, fluctuating between 47 and 56 percent with no trend. Wave fixed effects are near-zero. The 28-percentage-point period-level rise that defines the Korean trajectory has no Taiwanese counterpart. Whatever moved Korean abstract preference upward---democratic socialization, institutional maturation, the accumulation of democratic practice across generations---did not operate similarly in Taiwan, despite Taiwan's parallel democratization timeline and comparable economic development. The cross-country comparison introduced in §4.3 will show that the Taiwanese pattern is the mirror image of Korea's at the individual level as well: where Korea shows no economic connection to abstract preference, Taiwan shows a strong one.


## Economic evaluations and the structure of democratic support

### Satisfaction

Economic evaluations are a consistent and powerful predictor of satisfaction with democracy across all six Korean waves, ranging from $\beta$ = 0.262 in Wave 2 to $\beta$ = 0.455 in Wave 4, every estimate significant at *p* < 0.001. The pooled estimate is $\beta$ = 0.342. Government satisfaction shows even stronger effects ($\beta$ = 0.473 to 0.823; pooled $\beta$ = 0.575). No surprises here. H1 is confirmed.

### Abstract normative commitment

This is where the paradox lives---though it takes a form more complex than a simple null.

The pooled estimate is near zero: $\beta$ = --0.036 (*p* = 0.418) on the binary outcome (1 = "always preferable," 0 = otherwise). H2 is confirmed at the pooled level. But the pooled null conceals a temporal dynamic that the wave-by-wave results expose. In 2003, economic evaluations are significantly and positively associated with endorsing democracy as always preferable ($\beta$ = 0.323, *p* = 0.002): economically comfortable Koreans were substantially *more* likely to affirm democratic preference. By 2006 the relationship had disappeared ($\beta$ = 0.138, nonsignificant). It stayed null in 2011 ($\beta$ = --0.125, nonsignificant). Then it reversed. In 2015 the coefficient turns significantly negative ($\beta$ = --0.302, *p* = 0.010). In 2019, still negative and significant ($\beta$ = --0.266, *p* = 0.016). In 2022, negative and marginally significant ($\beta$ = --0.230, *p* = 0.054). The pooled null is an artifact---a positive early relationship cancelling a negative later one.

The trajectory amounts to a sign reversal. In 2003, economic comfort reinforced democratic endorsement---the pattern consolidation theory expects. By 2015, economic comfort was enabling normative scrutiny---the critical citizens pattern the theory associates with democratic maturity. Korea is not missing the critical citizens dynamic. It is developing one, across the full sample, over a two-decade arc that the pooled specification rendered invisible.

The 2003 starting point deserves attention. The survey captured Korea during the post-crisis recovery upswing, with economic evaluations both higher (mean = 0.443 vs. 0.356 in 2006) and more compressed (SD = 0.127 vs. 0.149). By 2006, after the credit card crisis and the Roh Moo-hyun impeachment turmoil, economic confidence had dropped sharply and the relationship disappeared. An alternative reading is that 2003 captured a transitional moment before the instrumental mode had fully consolidated---the democratic transition was barely fifteen years old, and the post-authoritarian upheaval may still have supplied non-economic reasons for democratic evaluation. The two accounts are not mutually exclusive, and a single data point cannot adjudicate.

What matters for the paper's argument is less the 2003 starting point than the 2015--2022 destination. The emergence of the critical citizens pattern in the later waves coincides with the period-level rise in abstract preference documented in §4.1: as endorsement of "always preferable" rose from 49 to 78 percent, the economic relationship shifted from reinforcement to scrutiny. The Korean public is not static. It is moving---slowly, unevenly, and through mechanisms unconnected to economic performance---toward the evaluative orientation that the consolidation literature treats as the hallmark of democratic depth.

### Substantive authoritarian rejection: direction and consistency

The diagnostic analysis that motivates H3 delivers a finding whose direction complicates the abstract decoupling.

Economically comfortable Korean citizens are not more rejecting of authoritarian alternatives. They are marginally more accepting. The pooled estimate for the authoritarian rejection index is $\beta$ = −0.082 (*p* < 0.001): citizens who evaluate the economy more favorably score roughly eight percentage points lower on rejection of strongman, military, expert, and single-party rule across the full range of the economic index. The relationship is significant for rejection of strongman rule, military governance, and single-party systems; it approaches conventional significance for expert rule.

H3a is disconfirmed. H3b finds support: the instrumental orientation does not stop at satisfaction. It penetrates the authoritarian rejection battery.

A qualification about magnitude. At the interquartile range of the economic index (0.333 to 0.500), the predicted difference on the authoritarian rejection index is 0.014 on the 0--1 scale---roughly 0.04 points on the original four-point response scale, or 1.3 percent of the distance between adjacent response categories. The same specification produces predicted differences of 0.057 for democratic satisfaction---four times larger---and effectively zero for abstract democratic preference. These are not large effects by any standard. But the theoretical weight falls on the structural pattern across outcome types, not on the magnitude of any single coefficient. The published ABS literature reports comparable effect sizes for attitudinal predictors of democratic support in Korea: Park and Shin [-@Park2006-bv] report *R*² = 0.013 for their democratic support model with twelve predictors, and Mauk [-@Mauk2022-dm] finds *R*² = 0.04--0.08 for political trust in Korea even with multiple predictor blocks.

The contrast with the abstract item is worth stating precisely. On "democracy is always preferable," the pooled economic coefficient is near zero ($\beta$ = −0.036, *p* = 0.418)---though as §4.2.2 documented, the pooled null masks a sign reversal from positive (2003) to negative (2015--2022). On the substantive items, the coefficient is significant and points in the wrong direction---toward greater tolerance of authoritarian alternatives among the economically comfortable. Satisfaction runs on an economic track. Abstract normative commitment runs on no track at all. And substantive anti-authoritarianism runs on the economic track too---but pointing toward complacency rather than vigilance.

Whether the authoritarian rejection items capture normative anti-authoritarianism or performance evaluation in different clothing remains a legitimate question. Controlling for democratic satisfaction produces at most 10 percent attenuation of the economic coefficient on authoritarian rejection, with all five outcomes retaining their original significance levels (Appendix J). If the rejection items were simply proxying satisfaction, adding satisfaction to the model should absorb the economic effect. It does not. An exploratory factor analysis of the satisfaction and authoritarian rejection items confirms the structural distinction: the two sets of items load on near-orthogonal factors (*r* = −0.11, RMSEA = 0.020, TLI = 0.995), with no substantive cross-loadings. The "democracy is always preferable" item loads weakly on both factors---suggesting it operates as a conceptually distinct third dimension rather than belonging cleanly to either satisfaction or anti-authoritarian evaluation. The pattern is stable across all six waves (Appendix K).

The authoritarian rejection battery, in other words, captures something real and distinct from satisfaction---and the economic relationship runs through it in the direction of reduced vigilance.

### Item specificity

The Korea--Taiwan gap on "democracy always preferable" (cross-country interaction $\beta$ = --0.278, *p* < 0.001; Appendix D) dwarfs the gap on the substantive items, where the mean interaction is −0.088 (@fig-coefmap). On the authoritarian rejection battery, Korea's economic sensitivity is stronger than Taiwan's---Korean citizens show a larger shift toward acceptance of authoritarian alternatives as economic comfort rises. Almost all of the headline divergence between the two countries runs through the abstract item.

The implication for interpreting the satisfaction paradox concerns direction rather than magnitude. The abstract item shows a null in Korea---no economic connection at all. The substantive items show a connection, but in the direction of reduced anti-authoritarianism among the economically comfortable. The decoupling is item-specific, confirming H4, but the specificity no longer suggests hidden democratic depth. It reveals different registers of the same instrumental orientation.

```{r fig-coefmap}
#| label: fig-coefmap
#| fig-cap: |
#|   **Economic evaluation coefficients across democratic attitude outcomes: Korea and Taiwan.**
#|   Each row is a pooled OLS coefficient on the economic evaluation index (0--1),
#|   estimated with wave fixed effects and controls for age, gender, education,
#|   urban/rural residence, and political interest. Points = $\beta$; horizontal
#|   bars = 95% confidence intervals (HC2 robust). Source: Asian Barometer
#|   Survey, waves 1--6.
#| out-width: "100%"
#| fig-pos: "H"
knitr::include_graphics(
  ifelse(knitr::is_latex_output(),
         file.path(fig_dir, "fig04_coefficient_map.pdf"),
         file.path(fig_dir, "fig04_coefficient_map.png"))
)
```


## The Taiwan comparison

Taiwan exhibits the expected critical citizens pattern on the abstract item. Economic evaluations are significantly and positively associated with endorsing democracy as always preferable in five of six waves, with the pooled estimate at $\beta$ = 0.286 (*p* < 0.001). The single exception is Wave 3 (2010), coinciding with the Ma Ying-jeou rapprochement with Beijing. Economically comfortable Taiwanese citizens are substantially more likely to affirm democracy as a matter of abstract principle.

On the satisfaction dimension, Taiwan mirrors Korea---strong positive economic effects, as expected.

The critical finding for the measurement argument is what happens on the substantive items. Taiwan's economic sensitivity on authoritarian rejection ($\beta$ = --0.040) is roughly half of Korea's ($\beta$ = --0.082). On individual items, Taiwan's rejection of military rule and expert rule fails to reach significance. The cross-country interaction is highly significant for the abstract item ($\beta$ = --0.278, *p* < 0.001), confirming H5, but the interaction attenuates substantially---and in some cases reverses direction---on the substantive items.

Taiwan's critical citizens pattern, in other words, is concentrated on the most abstract and identity-laden measure in the battery. On the items that require concrete evaluation of specific governance alternatives, it is Korea, not Taiwan, that shows the stronger economic connection. The divergence maps onto what we know about the two countries' democratization experiences: Taiwan's transition fused with national identity formation vis-à-vis China, making "democracy" a category of self-definition as much as a governance evaluation [@Rigger2011-jh; @Wachman1994-th]. Korea's transition carried no equivalent identity charge. The abstract item captures that difference. The substantive items do not.

This pattern suggests that the "democracy is always preferable" item is doing double duty---measuring normative evaluation in some contexts and identity signaling in others. In Taiwan, where democracy is constitutive of national self-definition, the item picks up the identity signal strongly. In Korea, where it is not, the item goes quiet---not because Korean democratic commitment is absent, but because it lives in a different register. One that the field's flagship measure does not reach.


## The conditional--unconditional boundary {#sec-results-boundary}

The abstract decoupling concentrates on a single item---"democracy is always preferable"---and the obvious question is whether this reflects an item-level peculiarity or a genuine structural boundary in Korean democratic attitudes. Four alternative normative-adjacent measures available in the ABS provide the test.

All four show significant positive associations with economic evaluations in both countries. "Democracy is the best form of government." The forced-choice tradeoff between democracy and economic development. "Democracy is suitable for our country." Whether democracy can solve society's problems. None demands the unconditional commitment that "always preferable" requires. Every one co-moves with economic comfort, exactly as instrumental evaluations should.

The pattern eliminates the worry that Korean citizens are simply disengaged from democratic evaluation. They are not. But it locates the disengagement with precision. The four responsive items share a structural feature: each permits the respondent to evaluate democracy relative to something. "Best form of government" invites comparison. "Suitable for our country" invites contextual assessment. The tradeoff item explicitly frames democracy against economic development. "Can solve problems" is an efficacy judgment. All are conditional on context, comparison, or performance---and all track the economy.

The "always preferable" item forecloses the hedge. "Always" imposes a temporal universality that the other items do not. The three-option forced-choice format---always, sometimes, indifferent---leaves no room for "it depends." Korean citizens endorse democracy on every measure that permits qualification. On the sole item that forbids it, the economic connection goes silent.

The gap is not between engagement and disengagement. It is between evaluation and identification---between assessing democracy's merits under present circumstances and affirming democratic commitment as unconditional principle. Korean democratic attitudes, on this evidence, are robustly evaluative. What they lack is the unconditional register---the willingness to endorse democracy regardless of what it delivers. That register is precisely what Easton [-@Easton1965-tz] theorized as diffuse support, and it is precisely what the field's most widely used indicator claims to measure.

Two implications follow. For the Korean case, the finding suggests that the satisfaction paradox is not a story about democratic indifference. Korean citizens think carefully about democracy---on output terms. The paradox resides in the gap between that careful instrumental evaluation and the principled commitment the consolidation literature expects it to generate. For the measurement debate, the finding suggests that the "always preferable" item isolates something real and distinct---unconditional identification rather than conditional evaluation---but that it captures this dimension only where historical conditions have given citizens a reason to think about democracy in unconditional terms. Taiwan, where democratization fused with national self-definition, provides that reason. Korea, where it did not, does not. The item's cross-national validity as a measure of the same construct is therefore in question---a point the Taiwan comparison (§4.3) and the mechanism probe (Appendix H) reinforce from different angles.


## Robustness {#sec-results-robustness}

The core findings survive a battery of specification and measurement checks. Adding institutional trust as a control reduces the economic-satisfaction coefficient but leaves it highly significant; the preference and authoritarian rejection equations are unchanged. Logistic regression on the binary "always preferable" outcome confirms the linear probability model results (Korea logit $\beta$ = --0.161, *p* = 0.424; Taiwan logit $\beta$ = 1.213, *p* < 0.001). Ordered logit models on the remaining ordinal-scale outcomes confirm the OLS results. Within-wave z-scored outcomes produce substantively identical patterns. Each of the six individual economic indicators entered separately reproduces the structural finding: every indicator predicts satisfaction, none predicts abstract preference, and the authoritarian rejection pattern is consistent across indicators---current economic evaluations showing the cleanest decoupling, forward-looking measures showing small negative quality effects.

The economic-satisfaction coefficient is virtually identical for younger and older respondents and similar across education levels. In no demographic subgroup does the economic-preference relationship reach significance in the positive direction. H6 is confirmed: the decoupling is society-wide.

The asymmetric-prediction test further disciplines the interpretation. If an underlying normative orientation colored economic perceptions symmetrically, the bias should inflate the economic coefficient similarly across both satisfaction and quality equations. The data show the opposite of symmetry: $\beta$ = 0.26--0.46 for satisfaction, approximately zero for abstract preference, −0.082 for authoritarian rejection. Three patterns for the same predictor. Symmetric confounding cannot produce asymmetric results. The test does not rule out differential confounding---social desirability bias, for instance, likely operates more strongly on the abstract item than on satisfaction or rejection---but it narrows the range of plausible confounding accounts substantially.


## Electoral winners, losers, and democratic support

The sore-loser interpretation does not survive the data. The economic-preference relationship is null for both electoral winners ($\beta$ = --0.044, *p* = 0.580) and losers ($\beta$ = --0.121, *p* = 0.189) in Korea. The formal interaction is nonsignificant ($\beta$ = --0.100, *p* = 0.409). Korean citizens decouple economic evaluations from abstract democratic commitment regardless of whether their party holds power.

In Taiwan, economic evaluations predict abstract preference in both groups but substantially more strongly among winners ($\beta$ = 0.331) than losers ($\beta$ = 0.126), with the interaction significant at *p* < 0.001. Under conditions where democracy carries identity weight, citizens whose party won connect economic conditions to democratic endorsement more strongly---consistent with greater investment in the democratic project whose outcome they shaped. On the satisfaction dimension, the winner-loser distinction makes no difference in either country.


# Discussion

## The satisfaction paradox, reconsidered

The core finding stands. Economic evaluations are robustly associated with satisfaction with democracy but not with abstract normative commitment in South Korea across two decades of survey data. Satisfaction tracks economic performance. Principled democratic endorsement does not. The consolidation literature's co-movement assumption---arguably its foundational empirical expectation---fails.

What makes the failure worse is the direction on the substantive items. Economically comfortable Koreans are not more rejecting of authoritarian alternatives; they are marginally more tolerant of strongman rule, military governance, and single-party systems. Performance-legitimation does not stop at satisfaction. It penetrates the authoritarian rejection battery.

The existing literature on Korean democratic attitudes has characterized support as "incomplete" or normatively shallow [@Park2006-bv], and broader accounts of democratic backsliding [@Shin2020-fg; @Foa2017-oq] have reinforced the impression of declining commitment. The aggregate ABS data tell a different story: abstract democratic preference has risen substantially, from 49 to 78 percent over two decades. The vulnerability lies not in the level or trend of normative endorsement but in its causal disconnection from the economic performance mechanism the consolidation literature posits as its foundation. Two decades of additional data confirm that Korean democratic support is structurally shallow in a specific sense, though they also specify its contours more precisely than "incomplete" implies. Korean citizens maintain high levels of authoritarian rejection in absolute terms---the mean rejection score hovers near 0.77 on the 0--1 scale across waves. Economic comfort does not dramatically erode that baseline. It shifts the margin, modestly, toward tolerance. The structure is one of high baseline anti-authoritarianism with performance-contingent variation at the edges---not the collapse of democratic commitment that "shallow foundations" might suggest.

Accounts of Korean democratic decay have focused on institutional erosion---the politicization of courts, suppression of critical media, and executive overreach documented by Shin [-@Shin2020-fg]. These are real and consequential. But they describe what elites do, not what citizens believe. During the Moon Jae-in administration that Shin's account targets---precisely the period in which, by his account, democratic norms were under the most sustained assault---the ABS data show abstract democratic preference climbing from 73 to 78 percent. The institutional-erosion narrative and the mass-attitudes evidence point in opposite directions. Neither can be inferred from the other, and conflating them risks misdiagnosis in both directions.

The period-level rise in abstract preference complicates the performance-legitimation interpretation---but not in the direction of reassurance. Korean citizens have become substantially more likely to affirm that democracy is "always preferable" over two decades, a trend that appears in every demographic stratum simultaneously. The rise is real. But it is disconnected from the economic mechanism that consolidation theory identifies as its engine. Education, political news consumption, and voting behavior correlate with abstract preference cross-sectionally; economic evaluations do not. And the wave fixed effects that capture the aggregate trend remain large even after controlling for every available individual-level predictor---the rise is primarily a macro-level phenomenon, likely reflecting the diffuse accumulation of democratic norms through institutional practice rather than any mechanism operating through economic performance evaluations.

Acemoglu et al. [-@Acemoglu2024-an] document precisely this pathway: successful democratic experience breeds its own support over generational time. The Korean data fit their account at the aggregate level. The contribution of the present analysis is to show that the macro-level normative accumulation and the micro-level economic mechanism operate on different dimensions of democratic support---the former on abstract preference, the latter on satisfaction---and that neither reaches the substantive anti-authoritarian vigilance the consolidation literature treats as the deepest layer of democratic commitment.

A further diagnostic sharpens the point. On the binary "always preferable" measure, the post-democratization cohort (born after 1975, *n* = 2,204) shows a significant negative economic coefficient ($\beta$ = −0.213, *p* = 0.009): at the interquartile range of the economic index, economically comfortable younger Koreans are roughly 3.6 percentage points less likely to endorse democracy as always preferable---a modest but significant critical citizens effect. The pre-democratization cohort (*n* = 5,063) shows no relationship ($\beta$ = 0.012, *p* = 0.817), and the formal cohort interaction is marginal (*p* = 0.060). The pattern is suggestive rather than definitive, but the direction is clear: citizens raised entirely under democratic governance show the evaluative dynamic the consolidation literature associates with democratic maturation, while citizens raised under authoritarianism do not. This is not nostalgia---the nostalgia mechanism was eliminated on the authoritarian rejection dimension---but its mirror: democratic socialization producing the abstract evaluative disposition that authoritarian-era socialization did not.

Why would economic comfort make citizens marginally more tolerant of authoritarian alternatives? The most historically specific candidate---developmental-state nostalgia, the associative recall of Park Chung-hee-era prosperity among those old enough to remember it---does not survive the data. The post-democratization cohort (born after 1975) actually shows a *larger* negative coefficient on authoritarian rejection ($\beta$ = −0.103, *p* = 0.005) than the pre-democratization cohort ($\beta$ = −0.076, *p* = 0.001). The formal interaction is nonsignificant ($\beta$ = 0.029, *p* = 0.486). Citizens with no living memory of authoritarian governance are no less susceptible to the comfort–tolerance link. The one exception---military rule rejection, where personal exposure to military governance does sharpen rejection (interaction $\beta$ = 0.087, *p* = 0.048)---is precisely the exception that proves the rule: experiential memory matters for the specific governance mode experienced, not for authoritarian tolerance generally. Cognitive disengagement fares no better. If economic comfort simply reduced the motivation to think critically about political alternatives, political interest should moderate the relationship. It does not ($\beta$ = 0.010, *p* = 0.881 for the interaction); high-interest and low-interest respondents show the same negative pattern.

What remains is threat reduction---the possibility that economic security lowers the perceived stakes of regime type, making authoritarian alternatives feel less consequential as hypothetical governance modes. Unlike nostalgia, it requires no historical memory. Unlike disengagement, it operates independently of political attentiveness. Both features match the data.

The contrast with the abstract preference item sharpens the point. Civic engagement does moderate the economic relationship on "democracy is always preferable," but in the opposite direction from what the critical citizens framework predicts. The composite engagement interaction is significant ($\beta$ = 0.388, *p* = 0.048): among the disengaged, economic comfort is associated with reduced democratic preference ($\beta$ = −0.251, *p* = 0.033), while among the engaged, the economic coefficient approaches zero or turns positive. The polint interaction points in the same direction ($\beta$ = 0.282, *p* = 0.067). The engaged minority's abstract democratic endorsement is insulated from economic contingency---stable regardless of economic conditions---even as the disengaged show a weak critical citizens pattern. On substantive anti-authoritarianism, no such distinction exists.

The winner-loser analysis reinforces this reading. The economic-preference relationship is null for both electoral winners ($\beta$ = −0.044, *p* = 0.580) and losers ($\beta$ = −0.121, *p* = 0.189) in Korea, and the economic-rejection relationship is similarly undifferentiated by partisan status. Korean citizens relate to democratic governance instrumentally regardless of whether their party holds power.


## What the abstract item measures

The Taiwan comparison sharpens the measurement point. Taiwan's abstract commitment is among the strongest documented in the East Asian survey literature, and it is growing stronger over time. But Taiwan's substantive authoritarian rejection is weaker than Korea's, and its conditional pro-democracy items show stronger economic contingency. The critical citizens pattern is concentrated on the one item most susceptible to identity expression---the abstract, unconditional "always preferable" question.

Two moderation tests probe the identity-signaling mechanism. If abstract democratic endorsement in Taiwan functions partly as an expression of anti-China national identity, the critical citizens pattern should be stronger among respondents who perceive China as a threat. The interaction points in the expected direction but does not reach conventional significance ($\beta$ = 0.093, *p* = 0.338; Appendix H.1). National pride shows no significant moderation in either country ($\beta_{\text{Taiwan}}$ = 0.073, *p* = 0.612; Appendix H.2).

These null results do not confirm the identity-signaling interpretation, and they do not disconfirm it---the available proxy variables are blunt instruments for what is fundamentally a diffuse, historically layered phenomenon. The cross-national pattern itself remains the strongest evidence: Taiwan's critical citizens effect concentrates on the most abstract and identity-laden item in the battery, and it strengthens over time as exclusive Taiwanese self-identification rises. Identity signaling is the most parsimonious account of this configuration. But other accounts---institutional design differences, civic education, or the proximity of an authoritarian threat generating sincere rather than performative democratic commitment---cannot be ruled out with these data. The interpretation should be read as a theoretically motivated hypothesis about what drives the Taiwan--Korea divergence, not an empirically settled mechanism.

Taiwan's democratization became inseparable from national identity through the political project of distinguishing Taiwan from authoritarian China---a project with identifiable agents and turning points. Lee Teng-hui's "localization" (本土化) reforms linked democratic governance to Taiwanese distinctiveness; his 1996 direct presidential election, held against PRC missile tests, welded democratic participation to national sovereignty. Chen Shui-bian deepened the identification; Tsai Ing-wen made democratic self-governance the centerpiece of Taiwan's international identity, a framing intensified after Hong Kong's 2019 protests. The rise of exclusive Taiwanese self-identification from under 25 percent in 1996 to over 60 percent tracks democratic consolidation almost perfectly [@Rigger2011-jh; @Wachman1994-th]. In this context, affirming that "democracy is always preferable" is not merely a governance evaluation---it is an identity claim. The trajectory was not linear: Ma Ying-jeou's rapprochement with Beijing coincides with the single wave (2010) in which Taiwan's critical citizens coefficient fails to reach significance.

Korea's democratization carried no equivalent identity charge. The 1987 transition was a domestic political settlement of enormous consequence for Korean political life, but it did not redefine who Koreans were as a national community. No external authoritarian other provided a foil against which democratic self-definition could crystallize. The abstract item goes quiet in Korea not because democratic commitment is absent but because the identity infrastructure that would give abstract endorsement its expressive force was never constructed.

Two implications follow. First, cross-national comparisons using the "always preferable" item as the sole or primary indicator of normative democratic depth risk conflating identity expression with principled evaluation. Studies that rank democratic consolidation by responses to this item---including the influential work of Foa and Mounk [-@Foa2017-oq]---may be measuring something different from what they intend in contexts where democracy carries strong identity weight. The multidimensional measurement framework developed by Claassen and colleagues [-@Claassen2025-xl] represents one path forward; the present analysis suggests that even within that framework, the distinction between abstract endorsement and substantive rejection of authoritarian alternatives deserves particular analytic attention. Second, and more consequentially for the Korean case: the gap between abstract endorsement and substantive commitment is not necessarily a vulnerability. It may simply reflect the absence of a historical reason to express democratic commitment in abstract, identity-laden terms. Acemoglu et al. [-@Acemoglu2024-an] document the reverse pathway: successful democratic experience breeds its own support over generational time. But their evidence suggests this process deposits into regime approval rather than abstract principled commitment---mirroring the Korean pattern, where two decades of democratic experience have raised satisfaction without generating the abstract endorsement the consolidation literature expects.

Whether that gap matters depends on what kind of threat a democracy faces.


## The Yoon crisis and the structure of resilience

The December 2024 martial law episode tested Korean democracy in a specific way, and the attitudinal data documented here frame what that test revealed.

Martial law was reversed within six hours. Legislators convened, voted unanimously to overturn the decree, and initiated impeachment. Courts asserted constitutional authority. The military chose institutional loyalty over presidential command. The institutional response was swift, coordinated, and decisive. Mass protests followed---but the institutional mechanisms had already contained the crisis before mass mobilization could become the decisive factor.

The attitudinal data documented here are consistent with---though they cannot establish---an interpretation in which institutional mechanisms bore the primary load. A citizenry for which economic performance drives satisfaction, abstract normative commitment drifts without economic anchoring, and substantive anti-authoritarianism weakens with economic comfort regardless of political attentiveness is not the profile of a public primed for normative mobilization around abstract democratic principle. But the framing requires two qualifications.

Mass protests followed the martial law declaration---large, rapid, and sustained. The ABS data cannot determine whether those protests were driven by principled democratic conviction, instrumental outrage at the disruption of a functioning system, partisan mobilization against an unpopular president, or some combination. The attitudinal structure documented here is consistent with any of these motivations except the first operating as the dominant force at the population level. "Consistent with" is the appropriate epistemic register for pre-crisis survey data applied to post-crisis behavior. The observation window closes in 2022; what happened in December 2024 lies beyond what these data can establish.

The institutional response itself was not automatic. Legislators braved military cordons to convene. Judges asserted constitutional authority. The military chose institutional loyalty over presidential command. These choices may reflect the normative commitments of elite actors---the politically attentive stratum whose abstract evaluative capacity §5.4 documents---now operating not as survey respondents but as institutional agents. The crisis was resolved by institutional design activated by individuals whose democratic commitments, at least on the abstract dimension, appear to run deeper than those of the median citizen. Whether that combination protects against the slower erosion that Levitsky and Ziblatt [-@Levitsky2018-dd] identify as the characteristic threat---executive overreach through formally legal channels, institutional checks weakened incrementally---is a question the December 2024 episode does not answer. The mechanism evidence suggests it may not: the population-wide susceptibility to reduced anti-authoritarian vigilance under economic comfort is not confined to the politically disengaged and is not driven by fading authoritarian memory. It operates in the present tense.


## The political interest finding {#sec-discussion-polint}

The mechanism tests in §5.1 revealed an asymmetry in what civic engagement moderates---and the asymmetry redraws the engaged minority concept in an unexpected direction.

On abstract preference, engagement insulates rather than sharpens. The composite engagement interaction is significant ($\beta$ = 0.388, *p* = 0.048): among the disengaged, economic comfort is associated with reduced democratic preference---the critical citizens direction---while the engaged show a null or positive economic-preference relationship. The polint interaction points in the same direction ($\beta$ = 0.282, *p* = 0.067). The aggregate Korean decoupling---the null pooled coefficient---is a composite of a weak critical citizens pattern among the disengaged and stable endorsement among the engaged. The engaged minority does not mirror Taiwan's evaluative dynamic; it exhibits the unconditional endorsement that characterizes internalized democratic commitment---or, less charitably, the socialized norm adherence that makes endorsement insensitive to economic conditions.

On authoritarian rejection, no such distinction exists. Economically comfortable citizens are marginally more tolerant of authoritarian alternatives regardless of political interest. The negative coefficient is $\beta$ = −0.093 (*p* = 0.001) for high-interest respondents and $\beta$ = −0.062 (*p* = 0.014) for low-interest respondents, with a nonsignificant interaction. If anything, the economically comfortable and politically engaged show a slightly stronger shift---though the difference is not statistically distinguishable.

The implication is that civic engagement stabilizes abstract democratic endorsement against economic contingency but does not extend that stability to the substantive dimension. The engaged minority's democratic preference is secure---insulated from economic conditions in a way the disengaged majority's is not. But when asked about specific authoritarian governance modes, the engaged respond to economic comfort the same way everyone else does---with marginally reduced rejection. The engaged minority's abstract stability does not inoculate against the instrumental evaluation of concrete alternatives.

Whether this asymmetric engagement is sufficient for democratic protection depends on the nature of the threat. For acute crises that present themselves as overt authoritarian seizure---martial law, military intervention, the institutional confrontations that the December 2024 episode exemplified---the engaged minority's stable abstract commitment may be sufficient, especially when combined with robust institutional design. For the gradual erosion that Levitsky and Ziblatt [-@Levitsky2018-dd] describe---executive overreach through formally legal channels, media independence compromised incrementally---the population-wide susceptibility to reduced anti-authoritarian vigilance under economic comfort is the more consequential finding. The engaged minority's abstract stability does not extend to the substantive dimension where that threat operates.


## Limitations

The most consequential limitation is the absence of panel data. The ABS provides repeated cross-sections; the analysis documents between-person associations, not within-person change. The asymmetric pattern---strong association with satisfaction, null for abstract commitment, significant and negative for authoritarian rejection---is inconsistent with a pure reverse-causality account, which predicts symmetric bias. Panel data would provide a substantially stronger test.

The abstract-versus-substantive distinction relies on repurposing the ABS authoritarian rejection battery as a measure of anti-authoritarian disposition. These items were designed to capture authoritarian nostalgia [@Chu2008-fk], not the converse, and rejecting named authoritarian alternatives is not equivalent to affirming democratic processes, institutions, or rights. The interpretive move---treating low acceptance of authoritarian alternatives as evidence of defensive democratic support---is reasonable but circumscribed. What the data document is the negative face of democratic commitment: what Korean citizens refuse. The affirmative face---whether they value rule of law, legislative deliberation, minority protection as ends rather than means---lies beyond what the authoritarian rejection battery can measure.

The two-country design constrains generalizability. Whether the gap between abstract and substantive commitment characterizes other post-developmental-state democracies---or whether it appears in contexts where democratic identity is anchored by mechanisms other than national differentiation from an authoritarian neighbor---requires extension to additional cases. A direct test of the identity-signaling interpretation would require the Taiwanese/Chinese identity spectrum item from the ABS Taiwan country-specific module, which is administered alongside the cross-national core questionnaire but distributed separately by the Election Study Center, National Chengchi University. Access to the country-specific variables is restricted to Taiwan-based researchers and is not available to the present analysis.

A measurement invariance concern follows from the paper's own argument. If the 'always preferable' item captures different constructs across contexts, the cross-country interaction compares potentially incommensurable quantities. Formal invariance testing on the authoritarian rejection battery---where cross-country comparability is claimed---supports scalar invariance ($\Delta$CFI = 0.001 metric, 0.003 scalar, both below Chen's [-@Chen2007-tk] 0.010 threshold; Appendix L). The cross-country interaction on 'always preferable' should be read as evidence that the item behaves differently across contexts, which is the paper's theoretical point rather than a threat to it. Economic evaluations may partially proxy incumbent evaluation, though models adding presidential trust where available produce substantively identical patterns.

The observation window closes in 2022, before the martial law crisis. Wave 7 data will provide a critical test of whether the patterns documented here have persisted, intensified, or shifted in response to the most dramatic democratic stress test Korea has experienced since democratization.


# Conclusion

Democratic satisfaction is not democratic deepening. Two decades of Asian Barometer data from South Korea establish this. Satisfaction and abstract democratic preference can rise in tandem at the aggregate level while being driven by entirely disconnected mechanisms at the individual level---a satisfaction paradox visible only when the analyst looks beneath the surface co-movement to the causal structure underneath. Economic performance powers satisfaction. It has no bearing on the 30-percentage-point rise in abstract preference, which is a period effect unexplained by any individual-level predictor. And on the substantive measures that tap concrete anti-authoritarian evaluation, economic comfort pushes in the wrong direction.

The diagnostic finding deepens rather than softens this conclusion. Korean citizens who decouple on "democracy is always preferable" do not compensate with stronger substantive anti-authoritarianism. They show weaker rejection of authoritarian alternatives as economic comfort rises---marginally, but consistently, and in the direction that performance-legitimation theory would predict. Korean democratic support is structured instrumentally across multiple dimensions: satisfaction tracks the economy, abstract commitment floats free of it, and substantive rejection of authoritarian alternatives erodes slightly as the economy delivers. The system held against an acute crisis in December 2024 because institutional mechanisms and the organized political opposition that activated them proved sufficient. Whether that combination holds against the slower erosion of democratic norms by actors who never declare martial law is the question the satisfaction paradox leaves unanswered.

Taiwan demonstrates that the abstract measure's strongest signal comes from a context where democracy is fused with national identity---where affirming democratic preference is simultaneously affirming national self-definition. On the substantive items, the cross-country divergence attenuates or reverses. The finding invites scrutiny of how the field measures what it claims to measure.


# Competing Interests {.unnumbered}

The author has no relevant financial or non-financial competing interests to report.

# References {.unnumbered}

::: {#refs}
:::