---
title: "Online Appendix: When Do Surveys Produce False Positives for Regime Support?"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
date: "`r format(Sys.time(), '%B %d, %Y, %H:%M')`"
format:
  pdf:
    documentclass: article
    classoption: [letterpaper, 12pt]
    geometry: margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    linkcolor: blue
    urlcolor: blue
    citecolor: blue
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{threeparttable}
        \usepackage{threeparttablex}
        \renewcommand{\thesection}{\Alph{section}}
        \renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
        \renewcommand{\thetable}{\Alph{section}\arabic{table}}
        \renewcommand{\thefigure}{\Alph{section}\arabic{figure}}
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
  docx:
    toc: true
    number-sections: true
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r load-data}
#| include: false
load("../analysis/results/prepared_data.RData")
load("../analysis/results/descriptive_results.RData")
load("../analysis/results/robustness_results.RData")

# Load main analysis results (falsification, FDR, sensitivity)
if (file.exists("../analysis/results/main_analysis_results.RData")) {
  load("../analysis/results/main_analysis_results.RData")
}
```

# Variable Definitions and Coding {#sec-variables}

@tbl-var-definitions presents the full set of variables used in the analysis, with metadata extracted directly from the YAML harmonization specifications used in the data pipeline.

```{r var-metadata-helpers}
#| include: false

library(yaml)

# --- Helper: extract variable metadata from YAML spec files ---
extract_var_metadata <- function(yaml_dir, var_ids) {
  yaml_files <- list.files(yaml_dir, pattern = "\\.yml$", full.names = TRUE)

  all_vars <- list()
  for (f in yaml_files) {
    spec <- yaml::read_yaml(f)
    if (!is.null(spec$variables)) {
      for (v in spec$variables) {
        all_vars[[v$id]] <- v
      }
    }
  }

  map_dfr(var_ids, function(vid) {
    v <- all_vars[[vid]]
    if (is.null(v)) {
      return(tibble(var_id = vid, w5_item = NA_character_,
                    scale_text = NA_character_, recode_note = NA_character_))
    }

    # W5 item code
    w5_item <- if (!is.null(v$source$w5)) as.character(v$source$w5) else NA_character_

    # Scale text
    smin <- v$scale$min
    smax <- v$scale$max
    scale_text <- if (!is.null(smin) && !is.null(smax)) {
      lo_lab <- v$scale$labels[[as.character(smin)]]
      hi_lab <- v$scale$labels[[as.character(smax)]]
      if (!is.null(lo_lab) && !is.null(hi_lab)) {
        paste0(smin, "--", smax, " (", lo_lab, " to ", hi_lab, ")")
      } else {
        paste0(smin, "--", smax)
      }
    } else { NA_character_ }

    # Recoding note for W5
    recode_note <- NA_character_
    harm <- v$harmonize
    if (!is.null(harm$exceptions$w5)) {
      exc <- harm$exceptions$w5
      fn_name <- exc$fn %||% exc$method %||% "custom"
      recode_note <- switch(fn_name,
        "identity" = "None",
        "safe_reverse_4pt" = "Reversed (4pt)",
        "safe_reverse_5pt" = "Reversed (5pt)",
        "safe_reverse_3pt" = "Reversed (3pt)",
        "safe_6pt_to_4pt" = "6pt rescaled to 4pt",
        "safe_4pt_none" = "None",
        "safe_reverse_2pt" = "Reversed (2pt)",
        fn_name
      )
    } else if (!is.null(harm$w5)) {
      w5h <- harm$w5
      fn_name <- w5h$fn %||% w5h$method %||% "custom"
      recode_note <- switch(fn_name,
        "identity" = "None",
        "recode" = "Recoded (see spec)",
        fn_name
      )
    } else if (!is.null(harm$default)) {
      def <- harm$default
      fn_name <- def$fn %||% def$method %||% "custom"
      recode_note <- switch(fn_name,
        "identity" = "None",
        "safe_reverse_4pt" = "Reversed (4pt)",
        "safe_reverse_5pt" = "Reversed (5pt)",
        "safe_reverse_3pt" = "Reversed (3pt)",
        "safe_6pt_to_4pt" = "6pt rescaled to 4pt",
        "safe_4pt_none" = "None",
        "safe_reverse_2pt" = "Reversed (2pt)",
        "middle_identity_5pt" = "Midpoint recoded",
        fn_name
      )
    }

    tibble(var_id = vid, w5_item = w5_item,
           scale_text = scale_text, recode_note = recode_note)
  })
}
```

```{r}
#| label: tbl-var-definitions
#| tbl-cap: "Variable definitions, scales, and coding for all variables used in the analysis. Metadata extracted from YAML harmonization specifications; N statistics computed from the full Hong Kong Wave 5 sample."

source("../../../_data_config.R")
yaml_dir <- file.path(survey_data_prep, "src/config/abs/harmonize_validated/")

# --- Display name and category mapping (controls sort order) ---
var_meta <- tribble(
  ~var_id,                         ~display_name,                          ~category,
  # Institutional Trust (6)
  "trust_police",                  "Trust in police",                      "Institutional Trust",
  "trust_national_government",     "Trust in national government",         "Institutional Trust",
  "trust_president",               "Trust in president/CE",                "Institutional Trust",
  "trust_parliament",              "Trust in parliament",                  "Institutional Trust",
  "trust_courts",                  "Trust in courts",                      "Institutional Trust",
  "trust_civil_service",           "Trust in civil service",               "Institutional Trust",
  # Democratic Attitudes (5)
  "dem_always_preferable",         "Democracy always preferable",          "Democratic Attitudes",
  "democracy_satisfaction",        "Democratic satisfaction",              "Democratic Attitudes",
  "democracy_suitability",         "Democracy suitability",               "Democratic Attitudes",
  "dem_extent_current",            "Extent of current democracy",          "Democratic Attitudes",
  "dem_country_present_govt",      "Current government dem.\\ rating",     "Democratic Attitudes",
  # Governance Perceptions (5)
  "gov_free_to_organize",          "Freedom to organize",                  "Governance Perceptions",
  "dem_free_speech",               "Freedom of speech",                    "Governance Perceptions",
  "election_free_fair",            "Elections free and fair",              "Governance Perceptions",
  "rich_poor_treated_equally",     "Rich and poor treated equally",        "Governance Perceptions",
  "gov_opposition_opportunities",  "Opposition opportunities",             "Governance Perceptions",
  # System Support (3)
  "system_deserves_support",       "System deserves support",              "System Support",
  "system_proud",                  "System pride",                         "System Support",
  "system_needs_change",           "System needs change",                  "System Support",
  # Authoritarian Alternatives (3)
  "strongman_rule",                "Strongman rule",                       "Authoritarian Alternatives",
  "military_rule",                 "Military rule",                        "Authoritarian Alternatives",
  "single_party_rule",             "Single-party rule",                    "Authoritarian Alternatives",
  # Political Efficacy & Participation (4)
  "political_interest",            "Political interest",                   "Political Efficacy \\& Participation",
  "efficacy_ability_participate",  "Efficacy: ability to participate",     "Political Efficacy \\& Participation",
  "efficacy_no_influence",         "Efficacy: no influence",               "Political Efficacy \\& Participation",
  "action_demonstration",          "Demonstration attendance",             "Political Efficacy \\& Participation",
  # Alternative Explanations & Controls (9)
  "dem_rating_china",              "China democratic rating",              "Alternative Explanations \\& Controls",
  "sat_president_govt",            "Satisfaction w/ president/govt",       "Alternative Explanations \\& Controls",
  "dem_country_future",            "Democratic future (10 yrs)",           "Alternative Explanations \\& Controls",
  "dem_essential_harmonized",      "Essential element of democracy",       "Alternative Explanations \\& Controls",
  "dem_procedural_vs_substantive", "Procedural vs.\\ substantive",        "Alternative Explanations \\& Controls",
  "nat_willing_emigrate",          "Willingness to emigrate",             "Alternative Explanations \\& Controls",
  "dem_best_form",                 "Democracy is best form",              "Alternative Explanations \\& Controls",
  "gov_courts_powerless",          "Courts powerless vs.\\ leaders",      "Alternative Explanations \\& Controls",
  "govt_responds_people",          "Government responds to people",        "Alternative Explanations \\& Controls"
)

# --- Extract YAML metadata ---
yaml_info <- extract_var_metadata(yaml_dir, var_meta$var_id)

# --- Compute N valid / N missing from hk5 ---
n_stats <- map_dfr(var_meta$var_id, function(vid) {
  if (vid %in% names(hk5)) {
    n_valid <- sum(!is.na(hk5[[vid]]))
    n_miss  <- sum(is.na(hk5[[vid]]))
  } else {
    n_valid <- NA_integer_
    n_miss  <- NA_integer_
  }
  tibble(var_id = vid, n_valid = n_valid, n_missing = n_miss)
})

# --- Merge everything ---
tbl_data <- var_meta |>
  left_join(yaml_info, by = "var_id") |>
  left_join(n_stats, by = "var_id") |>
  mutate(
    w5_item = ifelse(is.na(w5_item) | w5_item == "NULL", "---", w5_item),
    recode_note = replace_na(recode_note, "---"),
    question_wording = ""
  )

# --- Build table ---
display_tbl <- tbl_data |>
  select(display_name, w5_item, scale_text, recode_note, n_valid, n_missing, question_wording)

# Compute pack_rows groups
cat_counts <- tbl_data |>
  group_by(category) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(end = cumsum(n), start = end - n + 1)

# Preserve category order from var_meta
cat_order <- unique(tbl_data$category)
cat_counts <- cat_counts |> arrange(match(category, cat_order))

kbl <- display_tbl |>
  kable(
    col.names = c("Variable", "ABS W5 Item", "Response Scale",
                   "Recoding", "N Valid", "N Missing", "Question Wording"),
    booktabs = TRUE,
    longtable = TRUE
  ) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position", "repeat_header"),
    font_size = 9
  ) |>
  column_spec(7, width = "3cm")

# Apply pack_rows dynamically
for (i in seq_len(nrow(cat_counts))) {
  kbl <- kbl |>
    pack_rows(cat_counts$category[i], cat_counts$start[i], cat_counts$end[i])
}

kbl |>
  footnote(
    general = "N statistics refer to the full Hong Kong Wave 5 sample (all fieldwork periods). Scale descriptions reflect the harmonized coding direction after any recoding. Question wording column reserved for final manuscript preparation.",
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```


# Mann-Whitney U Tests {#sec-mannwhitney}

All two-sample $t$-tests reported in the main text were cross-validated with non-parametric Mann-Whitney $U$ (Wilcoxon rank-sum) tests. This guards against violations of normality assumptions, which are plausible given that many indicators are measured on short ordinal scales (1--4). @tbl-mw-comparison reports both test results side by side.

```{r}
#| label: tbl-mw-comparison
#| tbl-cap: "Parametric (t-test) vs. non-parametric (Mann-Whitney U) comparison for all pre/post NSL variables."

if (exists("mw_results")) {
  mw_results |>
    select(variable, cohens_d, t_p, t_sig, mw_p, mw_sig, agree) |>
    mutate(across(c(cohens_d), ~round(.x, 3)),
           t_p = format.pval(t_p, digits = 3),
           mw_p = format.pval(mw_p, digits = 3)) |>
    kable(
      col.names = c("Variable", "Cohen's d", "t p-value", "", "MW p-value", "", "Agree?"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("scale_down", "hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("Mann-Whitney results not yet computed. Run 04_robustness_checks.qmd first.")
}
```


# Demographic Balance Across Periods {#sec-balance}

A key concern with the within-wave quasi-experiment is that the Protest and Post-NSL sub-samples may differ systematically on demographic covariates, which could account for some of the observed attitudinal shifts. @tbl-balance reports means or proportions for age, gender, and education across the two periods, along with $t$-tests (continuous) or $\chi^2$ tests (categorical) for between-period differences.

```{r}
#| label: tbl-balance
#| tbl-cap: "Demographic balance across fieldwork periods (Protest vs. Post-NSL). SMD = standardized mean difference."

hk5_balance <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL"))

balance_rows <- list()

# Age (continuous)
age_protest <- hk5_balance |> filter(period == "Protest") |> pull(age)
age_postnsl <- hk5_balance |> filter(period == "Post-NSL") |> pull(age)
age_t <- t.test(age_protest, age_postnsl)
pooled_sd_age <- sqrt(((length(age_protest) - 1) * sd(age_protest, na.rm = TRUE)^2 +
                        (length(age_postnsl) - 1) * sd(age_postnsl, na.rm = TRUE)^2) /
                       (length(age_protest) + length(age_postnsl) - 2))
smd_age <- (mean(age_protest, na.rm = TRUE) - mean(age_postnsl, na.rm = TRUE)) / pooled_sd_age

balance_rows[[1]] <- tibble(
  Covariate = "Age (mean)",
  Protest = as.character(round(mean(age_protest, na.rm = TRUE), 1)),
  `Post-NSL` = as.character(round(mean(age_postnsl, na.rm = TRUE), 1)),
  SMD = round(smd_age, 3),
  Test = "t-test",
  `p-value` = round(age_t$p.value, 3)
)

# Gender
gender_var <- if ("female" %in% names(hk5_balance)) "female" else
              if ("gender" %in% names(hk5_balance)) "gender" else NULL

if (!is.null(gender_var)) {
  gender_tab <- hk5_balance |>
    filter(!is.na(!!sym(gender_var))) |>
    count(period, !!sym(gender_var)) |>
    group_by(period) |>
    mutate(prop = n / sum(n)) |>
    ungroup()

  gender_chi <- chisq.test(table(hk5_balance[[gender_var]], hk5_balance$period))

  if (gender_var == "female") {
    p1 <- mean(hk5_balance$female[hk5_balance$period == "Protest"], na.rm = TRUE)
    p2 <- mean(hk5_balance$female[hk5_balance$period == "Post-NSL"], na.rm = TRUE)
  } else {
    protest_tab <- gender_tab |> filter(period == "Protest")
    postnsl_tab <- gender_tab |> filter(period == "Post-NSL")
    p1 <- if (nrow(protest_tab) >= 2) protest_tab$prop[2] else NA
    p2 <- if (nrow(postnsl_tab) >= 2) postnsl_tab$prop[2] else NA
  }
  pooled_p <- (p1 + p2) / 2
  smd_gender <- if (!is.na(p1) && !is.na(p2)) (p1 - p2) / sqrt(pooled_p * (1 - pooled_p)) else NA

  balance_rows[[2]] <- tibble(
    Covariate = "Gender (% female)",
    Protest = paste0(round(p1 * 100, 1), "%"),
    `Post-NSL` = paste0(round(p2 * 100, 1), "%"),
    SMD = round(smd_gender, 3),
    Test = "Chi-squared",
    `p-value` = round(gender_chi$p.value, 3)
  )
}

# Education
edu_var <- if ("education_level_clean" %in% names(hk5_balance)) "education_level_clean" else
           if ("education_level" %in% names(hk5_balance)) "education_level" else NULL

if (!is.null(edu_var)) {
  edu_data <- hk5_balance |> mutate(
    edu_clean = case_when(
      !!sym(edu_var) %in% c(-1, 11, 97, 98, 99) ~ NA_real_,
      TRUE ~ as.numeric(!!sym(edu_var))
    )
  )
  edu_protest <- edu_data |> filter(period == "Protest") |> pull(edu_clean)
  edu_postnsl <- edu_data |> filter(period == "Post-NSL") |> pull(edu_clean)
  edu_t <- t.test(edu_protest, edu_postnsl)
  pooled_sd_edu <- sqrt(((sum(!is.na(edu_protest)) - 1) * sd(edu_protest, na.rm = TRUE)^2 +
                          (sum(!is.na(edu_postnsl)) - 1) * sd(edu_postnsl, na.rm = TRUE)^2) /
                         (sum(!is.na(edu_protest)) + sum(!is.na(edu_postnsl)) - 2))
  smd_edu <- (mean(edu_protest, na.rm = TRUE) - mean(edu_postnsl, na.rm = TRUE)) / pooled_sd_edu

  balance_rows[[length(balance_rows) + 1]] <- tibble(
    Covariate = "Education level (SE5, ordinal 1-10)",
    Protest = as.character(round(mean(edu_protest, na.rm = TRUE), 2)),
    `Post-NSL` = as.character(round(mean(edu_postnsl, na.rm = TRUE), 2)),
    SMD = round(smd_edu, 3),
    Test = "t-test",
    `p-value` = round(edu_t$p.value, 3)
  )
}

# Age groups (categorical)
if ("age_group" %in% names(hk5_balance)) {
  age_grp_tab <- hk5_balance |>
    filter(!is.na(age_group)) |>
    count(period, age_group) |>
    group_by(period) |>
    mutate(prop = round(n / sum(n) * 100, 1)) |>
    ungroup()

  age_chi <- chisq.test(table(hk5_balance$age_group, hk5_balance$period))

  for (ag in sort(unique(age_grp_tab$age_group))) {
    p_val <- age_grp_tab |> filter(period == "Protest", age_group == ag) |> pull(prop)
    pn_val <- age_grp_tab |> filter(period == "Post-NSL", age_group == ag) |> pull(prop)
    p_val <- if (length(p_val) == 0) 0 else p_val
    pn_val <- if (length(pn_val) == 0) 0 else pn_val

    balance_rows[[length(balance_rows) + 1]] <- tibble(
      Covariate = paste0("  Age ", ag),
      Protest = paste0(p_val, "%"),
      `Post-NSL` = paste0(pn_val, "%"),
      SMD = NA_real_,
      Test = if (ag == sort(unique(age_grp_tab$age_group))[1]) "Chi-squared" else "",
      `p-value` = if (ag == sort(unique(age_grp_tab$age_group))[1]) round(age_chi$p.value, 3) else NA_real_
    )
  }
}

bind_rows(balance_rows) |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("hold_position")) |>
  footnote(
    general = "SMD computed as (Protest \u2212 Post-NSL) / pooled SD for continuous variables and as the difference in proportions / pooled SD for binary variables. Values |SMD| < 0.1 indicate good balance.",
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```


# Stratified Bootstrap and Manski Bounds {#sec-robustness}

## Stratified Percentile Bootstrap {#sec-bootstrap}

To assess the robustness of the pre/post NSL shifts to sampling variability, I employed a stratified percentile bootstrap with 5,000 iterations, resampling independently within the Protest and Post-NSL periods to preserve the temporal structure of the quasi-experiment. @tbl-bootstrap reports the observed Cohen's $d$ alongside 95% bias-corrected and accelerated (BCa) confidence intervals.

```{r}
#| label: tbl-bootstrap
#| tbl-cap: "Stratified bootstrap results (5,000 iterations): BCa 95% confidence intervals for Cohen's d."

if (exists("boot_results")) {
  boot_results |>
    select(variable, observed_d, boot_d_lo, boot_d_hi, ci_type, sig_boot) |>
    mutate(across(where(is.numeric), ~round(.x, 3))) |>
    kable(
      col.names = c("Variable", "Observed d", "d CI Lo", "d CI Hi",
                     "CI Type", "Sig?"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("scale_down", "hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("Bootstrap results not yet computed. Run 04_robustness_checks.qmd first.")
}
```

```{r}
#| label: fig-bootstrap-forest
#| fig-cap: "Stratified bootstrap 95% BCa confidence intervals for Cohen's d (Post-NSL − Protest). Positive values indicate higher post-NSL responses. Variables where the CI excludes zero are shown in colour."
#| fig-width: 10
#| fig-height: 8

if (exists("boot_results")) {
  boot_results |>
    mutate(
      variable = str_replace_all(variable, "_", " ") |> str_to_title(),
      sig_label = ifelse(sig_boot, "Significant", "Not significant")
    ) |>
    ggplot(aes(x = observed_d, y = reorder(variable, observed_d), color = sig_label)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3) +
    geom_errorbarh(aes(xmin = boot_d_lo, xmax = boot_d_hi), height = 0.25) +
    scale_color_manual(values = c("Significant" = "#E53935", "Not significant" = "#9E9E9E")) +
    labs(
      x = "Cohen's d (Post-NSL - Protest)",
      y = NULL,
      color = NULL
    ) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")
}
```


## Informal Bounding Analysis {#sec-manski}

A key concern with the trust increase is that compositional selection --- whereby regime critics disproportionately declined to participate in the post-NSL survey --- could entirely account for the observed shift. To assess the plausibility of this account, I conduct an informal bounding exercise in the spirit of @Manski2003-wo.

The logic is as follows: if we assume that all non-respondents in the post-NSL period would have reported the minimum possible trust score (1 on the 1--4 scale), what proportion of the target population would need to have been "missing critics" for the trust increase to disappear? I present two scenarios: a worst-case bound (imputing the scale minimum) and a more empirically grounded scenario (imputing the 25th percentile of the protest-period distribution, representing respondents drawn from the bottom quartile of critics). @tbl-manski-bounds-police and @tbl-manski-bounds-govt present the results for trust in police and trust in the national government, respectively.

```{r}
#| label: tbl-manski-bounds-police
#| tbl-cap: "Bounding analysis for trust in police: worst-case (impute minimum) and bottom-quartile (impute 25th percentile of protest distribution) scenarios."

hk5_manski <- hk5 |> filter(period %in% c("Protest", "Post-NSL"))

protest_police <- hk5_manski |> filter(period == "Protest", !is.na(trust_police)) |> pull(trust_police)
postnsl_police <- hk5_manski |> filter(period == "Post-NSL", !is.na(trust_police)) |> pull(trust_police)
protest_mean_police <- mean(protest_police, na.rm = TRUE)
q25_police <- quantile(protest_police, 0.25, na.rm = TRUE)
scale_min_police <- min(c(protest_police, postnsl_police), na.rm = TRUE)

manski_police_ext <- map_dfr(seq(0, 40, by = 5), function(pct) {
  n_obs <- length(postnsl_police)
  n_miss <- round(n_obs * pct / (100 - pct))
  total_n <- n_obs + n_miss
  adj_mean_min <- (sum(postnsl_police) + n_miss * scale_min_police) / total_n
  adj_mean_q25 <- (sum(postnsl_police) + n_miss * q25_police) / total_n
  tibble(
    `% Missing` = pct,
    `N Obs` = n_obs,
    `N Missing` = n_miss,
    `Adj Mean (min)` = round(adj_mean_min, 3),
    `Still > Protest? (min)` = adj_mean_min > protest_mean_police,
    `Adj Mean (Q25)` = round(adj_mean_q25, 3),
    `Still > Protest? (Q25)` = adj_mean_q25 > protest_mean_police
  )
})

manski_police_ext |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("scale_down", "hold_position")) |>
  footnote(
    general = paste0(
      "Protest-period mean: ", round(protest_mean_police, 3),
      ". Worst-case imputes scale minimum (", scale_min_police,
      "); Q25 imputes the 25th percentile of the protest-period distribution (",
      q25_police, "), a more empirically grounded assumption."
    ),
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```

```{r}
#| label: tbl-manski-bounds-govt
#| tbl-cap: "Bounding analysis for trust in national government: worst-case (impute minimum) and bottom-quartile (impute 25th percentile of protest distribution) scenarios."

protest_govt <- hk5_manski |> filter(period == "Protest", !is.na(trust_national_government)) |> pull(trust_national_government)
postnsl_govt <- hk5_manski |> filter(period == "Post-NSL", !is.na(trust_national_government)) |> pull(trust_national_government)
protest_mean_govt <- mean(protest_govt, na.rm = TRUE)
q25_govt <- quantile(protest_govt, 0.25, na.rm = TRUE)
scale_min_govt <- min(c(protest_govt, postnsl_govt), na.rm = TRUE)

manski_govt_ext <- map_dfr(seq(0, 40, by = 5), function(pct) {
  n_obs <- length(postnsl_govt)
  n_miss <- round(n_obs * pct / (100 - pct))
  total_n <- n_obs + n_miss
  adj_mean_min <- (sum(postnsl_govt) + n_miss * scale_min_govt) / total_n
  adj_mean_q25 <- (sum(postnsl_govt) + n_miss * q25_govt) / total_n
  tibble(
    `% Missing` = pct,
    `N Obs` = n_obs,
    `N Missing` = n_miss,
    `Adj Mean (min)` = round(adj_mean_min, 3),
    `Still > Protest? (min)` = adj_mean_min > protest_mean_govt,
    `Adj Mean (Q25)` = round(adj_mean_q25, 3),
    `Still > Protest? (Q25)` = adj_mean_q25 > protest_mean_govt
  )
})

manski_govt_ext |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("scale_down", "hold_position")) |>
  footnote(
    general = paste0(
      "Protest-period mean: ", round(protest_mean_govt, 3),
      ". Worst-case imputes scale minimum (", scale_min_govt,
      "); Q25 imputes the 25th percentile of the protest-period distribution (",
      q25_govt, "), a more empirically grounded assumption."
    ),
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```

The results suggest that the observed trust increase is robust to all but implausibly extreme selection scenarios. Under worst-case imputation (all non-respondents scored at the scale minimum), the post-NSL trust mean for police remains above its protest-period baseline until approximately 35--40% of the target population is assumed to consist of missing extreme critics. Under the more empirically grounded bottom-quartile scenario, the trust increase persists even at higher assumed non-response rates. Selection alone would therefore need to reach implausibly extreme levels---removing over one-third of the survey-eligible population from the respondent pool, all of whom would have to hold the most negative possible view---to fully account for the observed shift.

## Ruling Out Differential Baseline Distrust {#sec-baseline-distrust}

A compositional selection account offers a plausible alternative to the preference falsification interpretation: if protest-aligned respondents disproportionately exited the survey frame following NSL implementation, and if those respondents happened to distrust coercive institutions more than they distrusted political or democratic institutions at baseline, observed gradient-consistent shifts could reflect compositional change rather than strategic misreporting. We test this directly.

The selection account generates a specific prediction: items with greater baseline variance in trust---reflecting deeper pre-existing disagreement---should shift more sharply after NSL implementation, as the critics most likely to exit were clustered at the low end of these items. We operationalize baseline variance as the pre-period standard deviation for each survey item and test whether it independently predicts the magnitude of post-NSL shifts (Cohen's *d*) after accounting for sensitivity rank.

We conduct the analysis at the item level ($n = 6$: five individual trust items plus a democratic attitudes index formed from three z-scored items re-expressed in trust-scale units).^[The democratic attitudes index averages z-scored versions of democracy suitability, democracy always preferable, and democratic satisfaction. Collapsing these three rank-5 items to a single index is theoretically motivated---all three are "low-sensitivity abstract evaluations" in the framework's typology---and avoids pseudo-replication from treating operationalizations of the same construct as independent observations.] Cross-item regressions regress Cohen's *d* on sensitivity rank (M1), baseline SD (M2), and both jointly (M3). Results are summarized in @tbl-baseline-distrust and @tbl-gradient-regression.

```{r}
#| include: false
if (file.exists("../analysis/results/baseline_distrust_interaction.RData")) {
  load("../analysis/results/baseline_distrust_interaction.RData")
}
```

```{r}
#| label: tbl-baseline-distrust
#| tbl-cap: "Baseline disagreement vs. post-NSL shift: item-level statistics. Protest-period SD measures pre-existing disagreement (the key lever for the selection account). Sensitivity rank is the pre-specified theoretical ordering. Cohen's $d$ is the observed post-NSL effect size."
#| echo: false

if (exists("primary_stats")) {
  primary_stats |>
    select(label, sensitivity_rank, baseline_sd, cohens_d) |>
    mutate(across(c(baseline_sd, cohens_d), ~round(.x, 3))) |>
    kable(
      col.names = c("Item", "Sensitivity rank", "Protest-period SD", "Cohen's $d$"),
      booktabs = TRUE, escape = FALSE
    ) |>
    kable_styling(latex_options = c("hold_position")) |>
    footnote(
      general = "Sensitivity rank: 1 = most coercive/sensitive (police, national government), 5 = least sensitive (democratic attitudes index). Cohen's $d$ = (Post-NSL $-$ Protest) / pooled SD; positive values indicate post-NSL increase. Democratic attitudes index = mean of z-scored democracy suitability, democracy always preferable, and democratic satisfaction items.",
      general_title = "Note: ", footnote_as_chunk = TRUE, escape = FALSE
    )
} else {
  cat("Run analysis/12_baseline_distrust_interaction.R to generate these results.")
}
```

The results are unambiguous. Sensitivity rank alone explains 79% of cross-item variance in post-NSL shifts (M1: $\beta = -0.094$, $p = .007$). Adding baseline SD to the model leaves sensitivity rank essentially unchanged ($\beta = -0.102$, $p = .053$) while contributing no independent explanatory power (M2 baseline SD coefficient: $\beta = -0.088$, $p = .80$). The partial correlation for sensitivity rank, controlling for baseline SD, is $r = -0.805$ ($p = .029$); the partial correlation for baseline SD, controlling for rank, is $r = -0.062$ ($p = .14$). The Spearman rank correlation between sensitivity and shift is $\rho = -0.946$ ($p = .001$).

The selection account's key prediction---that baseline variance should positively predict post-NSL shifts---is directly contradicted. The clearest discriminating observation involves police and courts. These two items have nearly identical pre-period standard deviations (police SD = 1.16; courts SD = 1.02), indicating similar degrees of baseline disagreement. Yet their post-NSL shifts diverge dramatically: police confidence increased by $d = 0.46$, while court confidence changed by only $d = 0.05$. If baseline heterogeneity were driving the gradient, items with similar variance should shift similarly. Instead, the difference precisely tracks coercive sensitivity: police represent the front line of NSL enforcement; courts retained meaningful judicial independence. Baseline distrust cannot explain this divergence; institutional sensitivity can.

```{r}
#| label: tbl-gradient-regression
#| tbl-cap: "Cross-item regression: sensitivity rank and baseline disagreement as predictors of Cohen's $d$. M1: sensitivity rank alone. M2: baseline SD alone. M3: both jointly (key test). M4: M3 with standardized predictors."
#| echo: false

if (exists("regression_table") && exists("model_fit")) {
  regression_table |>
    filter(model %in% c("M1: Sensitivity rank only",
                        "M2: Baseline SD only",
                        "M3: Both jointly (key test)",
                        "M4: Standardized")) |>
    mutate(
      term = case_when(
        term == "(Intercept)"          ~ "Intercept",
        term == "sensitivity_rank"     ~ "Sensitivity rank",
        term == "sensitivity_rank_z"   ~ "Sensitivity rank (std.)",
        term == "baseline_sd"          ~ "Baseline SD",
        term == "baseline_sd_z"        ~ "Baseline SD (std.)",
        TRUE ~ term
      )
    ) |>
    select(Model = model, Term = term,
           `$b` = estimate, `$SE` = std.error,
           `$t` = statistic, `$p` = p.value) |>
    kable(booktabs = TRUE, escape = FALSE) |>
    kable_styling(latex_options = c("hold_position", "scale_down")) |>
    footnote(
      general = paste0(
        "$n = 6$ observations (5 trust items + democratic attitudes index). ",
        "Model fit: M1 adj. $R^2$ = ", model_fit$adj_r_squared[1],
        "; M2 adj. $R^2$ = ", model_fit$adj_r_squared[2],
        "; M3 adj. $R^2$ = ", model_fit$adj_r_squared[3], "."
      ),
      general_title = "Note: ", footnote_as_chunk = TRUE, escape = FALSE
    )
} else {
  cat("Run analysis/12_baseline_distrust_interaction.R to generate these results.")
}
```

A supplementary analysis disaggregating the democratic attitudes index into its three component items ($n = 9$) confirms the main finding ($\rho = -0.846$, $p = .004$) and yields an additional piece of evidence. *Democracy Suitability* has the highest baseline SD in the battery (2.98)---indicating deep pre-existing disagreement---yet posts the sharpest *negative* shift ($d = -0.522$). Under the selection account, high baseline variance should predict upward shift (as critics exit); instead, this item moves in the wrong direction entirely. The selection account cannot accommodate this pattern. The divergence of *Democracy Suitability* from other democratic items is consistent with item asymmetry: evaluating whether democracy "suits" Hong Kong implicitly references regime performance, making falsification psychologically costly in a way that endorsing democracy as an abstract ideal ("democracy is always preferable") is not (see @sec-item-asymmetry for the theoretical argument).

Taken together, these tests support the conclusion that the sensitivity gradient reflects strategic response processes rather than differential attrition of baseline critics.


# Supplementary Figures {#sec-figures}

## Age-Stratified Pre/Post NSL Shifts {#sec-age-stratified}

@fig-age-strat-appendix presents the pre/post NSL mean comparison within each age cohort for trust in police, trust in government, democracy suitability, and freedom to organize. If the trust increase were driven entirely by compositional selection --- specifically, by younger, more critical respondents declining to participate in the post-NSL period --- we would expect the shift to appear primarily or exclusively in younger age groups. The figure shows that while the magnitude of change varies across cohorts, the direction of the trust increase is broadly consistent across age groups, which is more consistent with preference falsification operating across the population than with a purely compositional account.

```{r}
#| label: fig-age-strat-appendix
#| fig-cap: "Pre/Post NSL mean shifts within age cohorts for key indicators. Error bars show 95% confidence intervals."
#| fig-width: 12
#| fig-height: 7

hk5_analysis <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL")) |>
  mutate(period = droplevels(period))

age_strat_vars <- c("trust_police", "democracy_suitability",
                     "trust_national_government", "gov_free_to_organize")

age_strat_labels <- c(
  trust_police = "Trust in Police",
  democracy_suitability = "Democracy Suitability",
  trust_national_government = "Trust in Government",
  gov_free_to_organize = "Freedom to Organize"
)

age_strat_data <- hk5_analysis |>
  filter(!is.na(age_group)) |>
  pivot_longer(cols = all_of(age_strat_vars), names_to = "variable", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(variable, period, age_group) |>
  summarise(
    mean = mean(value),
    se = sd(value) / sqrt(n()),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(var_label = age_strat_labels[variable])

ggplot(age_strat_data, aes(x = age_group, y = mean, fill = period)) +
  geom_col(position = position_dodge(0.8), width = 0.7, alpha = 0.85) +
  geom_errorbar(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se),
                position = position_dodge(0.8), width = 0.25) +
  facet_wrap(~var_label, scales = "free_y", ncol = 2) +
  scale_fill_manual(values = c("Protest" = "#FF7043", "Post-NSL" = "#5C6BC0")) +
  labs(x = "Age Cohort", y = "Mean", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
```


# Falsification Tests and Multiple Comparisons {#sec-falsification}

## Placebo-Adjacent Items

To assess whether the observed attitudinal shifts are specifically attributable to politically sensitive mechanisms rather than generalized confounding (e.g., pandemic effects, time trends), I pre-specified a set of items with lower expected political sensitivity. @tbl-falsification reports the pre/post NSL comparison for these placebo-adjacent items.

```{r}
#| label: tbl-falsification
#| tbl-cap: "Placebo-adjacent items: pre-specified items with lower expected political sensitivity. Small or non-significant effects are consistent with NSL-specific mechanisms rather than generalized confounding."

if (exists("falsification_results") && !is.null(falsification_results)) {
  falsification_results |>
    mutate(across(c(protest_mean, postnsl_mean, delta, cohens_d), ~round(.x, 3)),
           p_value = format.pval(p_value, digits = 3)) |>
    kable(
      col.names = c("Variable", "Protest Mean", "Post-NSL Mean",
                     "Delta", "Cohen's d", "p-value", "Sig"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("Falsification results not yet computed. Run 02_main_analysis.qmd first.")
}
```

## Benjamini-Hochberg FDR Correction

The main analysis tests multiple variables simultaneously. To guard against inflated Type I error rates, @tbl-fdr reports raw and Benjamini-Hochberg adjusted $p$-values for all pre/post NSL comparisons. The four pre-specified primary outcomes are flagged.

```{r}
#| label: tbl-fdr
#| tbl-cap: "Raw and FDR-corrected (Benjamini-Hochberg) p-values for all pre/post NSL comparisons."

if (exists("fdr_results")) {
  fdr_results |>
    select(variable, primary, cohens_d, p_value, sig, p_bh, sig_bh) |>
    mutate(across(c(cohens_d), ~round(.x, 3)),
           p_value = format.pval(p_value, digits = 3),
           p_bh = format.pval(p_bh, digits = 3)) |>
    kable(
      col.names = c("Variable", "Primary?", "Cohen's d", "p (raw)", "",
                     "p (BH-adj)", ""),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("scale_down", "hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("FDR results not yet computed. Run 02_main_analysis.qmd first.")
}
```


# Critical Citizens Correlation: CI and Period Decomposition {#sec-critical-citizens}

The main text reports that Hong Kong is the only country among the sixteen ABS Wave 5 polities where the correlation between democratic preference ("democracy is always preferable") and democratic satisfaction is positive. This section reports the precise estimate with 95% confidence intervals and decomposes the correlation by fieldwork period.

```{r}
#| label: tbl-critical-citizens-ci
#| tbl-cap: "Critical citizens correlation (democratic preference × democratic satisfaction): overall and by fieldwork period, with 95% confidence intervals."

if (exists("hk5")) {
  hk5_cc <- hk5 |>
    filter(!is.na(dem_always_preferable), !is.na(democracy_satisfaction))

  # Overall correlation with CI
  overall_test <- cor.test(hk5_cc$dem_always_preferable, hk5_cc$democracy_satisfaction,
                           method = "pearson")

  # By period
  protest_cc <- hk5_cc |> filter(period == "Protest")
  postnsl_cc <- hk5_cc |> filter(period == "Post-NSL")

  protest_test <- if (nrow(protest_cc) > 10) {
    cor.test(protest_cc$dem_always_preferable, protest_cc$democracy_satisfaction)
  } else { NULL }

  postnsl_test <- if (nrow(postnsl_cc) > 10) {
    cor.test(postnsl_cc$dem_always_preferable, postnsl_cc$democracy_satisfaction)
  } else { NULL }

  cc_results <- tibble(
    Period = c("Overall (Wave 5)", "Protest period", "Post-NSL period"),
    N = c(nrow(hk5_cc), nrow(protest_cc), nrow(postnsl_cc)),
    r = c(overall_test$estimate,
          if (!is.null(protest_test)) protest_test$estimate else NA,
          if (!is.null(postnsl_test)) postnsl_test$estimate else NA),
    CI_lo = c(overall_test$conf.int[1],
              if (!is.null(protest_test)) protest_test$conf.int[1] else NA,
              if (!is.null(postnsl_test)) postnsl_test$conf.int[1] else NA),
    CI_hi = c(overall_test$conf.int[2],
              if (!is.null(protest_test)) protest_test$conf.int[2] else NA,
              if (!is.null(postnsl_test)) postnsl_test$conf.int[2] else NA),
    p = c(overall_test$p.value,
          if (!is.null(protest_test)) protest_test$p.value else NA,
          if (!is.null(postnsl_test)) postnsl_test$p.value else NA)
  )

  cc_results |>
    mutate(across(c(r, CI_lo, CI_hi), ~round(.x, 3)),
           p = format.pval(p, digits = 3)) |>
    kable(
      col.names = c("Period", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position"))
} else {
  cat("Data not loaded. Ensure prepared_data.RData is available.")
}
```

The positive correlation is concentrated in the post-NSL period, consistent with a mechanism that operates more strongly under authoritarian constraint.

## Cross-National Comparison {#sec-cc-crossnational}

@tbl-cc-crossnational reports the Pearson correlation between democratic preference ("democracy is always preferable") and democratic satisfaction for all fifteen ABS Wave 5 polities with sufficient data ($N \geq 50$).

```{r}
#| label: tbl-cc-crossnational
#| tbl-cap: "Critical citizens correlation (democratic preference × democratic satisfaction) across ABS Wave 5 polities, sorted by correlation magnitude."

if (file.exists("../analysis/results/cc_country.RData")) {
  load("../analysis/results/cc_country.RData")
  cc_country |>
    mutate(across(c(r, ci_lo, ci_hi), ~round(.x, 3)),
           p = format.pval(p, digits = 3)) |>
    kable(
      col.names = c("Polity", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position")) |>
    row_spec(1, bold = TRUE)
} else {
  cat("cc_country.RData not found.")
}
```

Hong Kong is the only polity with a positive correlation ($r = 0.279$). Singapore shows a near-zero positive value ($r = 0.013$, $p = .70$); all other polities are negative, with established democracies (Australia, Taiwan, Japan, South Korea) showing the strongest negative correlations as expected by the critical citizens framework.

## Decomposition by Democratic Conception Type {#sec-cc-conception}

@tbl-cc-conception decomposes the critical citizens correlation by respondents' conception of democracy, using the ABS forced-choice item asking which element is most essential to democracy. Respondents are classified as holding procedural conceptions (free expression or free elections) or substantive conceptions (basic necessities or clean governance).

```{r}
#| label: tbl-cc-conception
#| tbl-cap: "Critical citizens correlation decomposed by democratic conception type and fieldwork period."

if (exists("hk5")) {
  conception_labels <- c("0" = "Substantive", "1" = "Procedural")
  
  hk5_cc2 <- hk5 |> 
    filter(period %in% c("Protest", "Post-NSL"),
           !is.na(dem_procedural_vs_substantive))
  
  cc_conception <- expand_grid(
    period = c("Protest", "Post-NSL"),
    conception = c(0, 1)
  ) |> pmap_dfr(function(period, conception) {
    sub <- hk5_cc2 |> 
      filter(period == !!period, dem_procedural_vs_substantive == conception,
             !is.na(dem_always_preferable), !is.na(democracy_satisfaction))
    if (nrow(sub) < 20) return(NULL)
    ct <- cor.test(sub$dem_always_preferable, sub$democracy_satisfaction)
    tibble(
      Period = period,
      Conception = conception_labels[as.character(conception)],
      N = nrow(sub),
      r = ct$estimate,
      CI_lo = ct$conf.int[1],
      CI_hi = ct$conf.int[2],
      p = ct$p.value
    )
  })
  
  cc_conception |>
    mutate(across(c(r, CI_lo, CI_hi), ~round(.x, 3)),
           p = format.pval(p, digits = 3)) |>
    kable(
      col.names = c("Period", "Conception", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position"))
} else {
  cat("Data not loaded.")
}
```

The positive correlation is concentrated entirely among respondents with procedural conceptions of democracy ($r = 0.355$, $p < .001$). Respondents with substantive conceptions show near-zero correlations in both periods. A Fisher $z$-test confirms the difference between conception groups in the post-NSL period is statistically significant ($z = 3.55$, $p < .001$). See Sections 4.6 and 5.3 of the main text for discussion.

## Fine-Grained Decomposition by Essential Element {#sec-cc-fine}

@tbl-cc-fine provides a finer decomposition using the four specific response options.

```{r}
#| label: tbl-cc-fine
#| tbl-cap: "Critical citizens correlation by specific essential element of democracy and fieldwork period."

if (exists("hk5")) {
  essential_labels <- c(
    "1" = "Free expression", "2" = "Free elections",
    "4" = "Basic necessities", "5" = "Clean governance"
  )
  
  cc_fine <- expand_grid(
    period = c("Protest", "Post-NSL"),
    essential = c(1, 2, 4, 5)
  ) |> pmap_dfr(function(period, essential) {
    sub <- hk5 |> 
      filter(period == !!period, dem_essential_harmonized == essential,
             !is.na(dem_always_preferable), !is.na(democracy_satisfaction))
    if (nrow(sub) < 20) return(NULL)
    ct <- cor.test(sub$dem_always_preferable, sub$democracy_satisfaction)
    tibble(
      Period = period,
      Element = essential_labels[as.character(essential)],
      N = nrow(sub),
      r = round(ct$estimate, 3),
      CI_lo = round(ct$conf.int[1], 3),
      CI_hi = round(ct$conf.int[2], 3),
      p = format.pval(ct$p.value, digits = 3)
    )
  })
  
  cc_fine |>
    kable(
      col.names = c("Period", "Essential Element", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position"))
} else {
  cat("Data not loaded.")
}
```

# Differential Item Non-Response {#sec-nonresponse}

@tbl-nonresponse reports item-level response rates for key survey variables, comparing the protest period and post-NSL period sub-samples.

```{r}
#| label: tbl-nonresponse
#| tbl-cap: "Item-level response rates by fieldwork period."

if (exists("hk5")) {
  hk5_nr <- hk5 |> filter(period %in% c("Protest", "Post-NSL"))
  
  key_vars <- c(
    "trust_police", "trust_national_government", "trust_president",
    "trust_parliament", "trust_courts", "trust_civil_service",
    "dem_always_preferable", "democracy_satisfaction",
    "dem_extent_current", "democracy_suitability",
    "system_deserves_support",
    "gov_free_to_organize", "dem_free_speech", "election_free_fair",
    "political_interest", "rich_poor_treated_equally",
    "nat_willing_emigrate"
  )
  
  item_categories <- c(
    trust_police = "Trust", trust_national_government = "Trust",
    trust_president = "Trust", trust_parliament = "Trust",
    trust_courts = "Trust", trust_civil_service = "Trust",
    dem_always_preferable = "Dem. (normative)",
    democracy_satisfaction = "Dem. (normative)",
    dem_extent_current = "Dem. (evaluative)",
    democracy_suitability = "Dem. (evaluative)",
    system_deserves_support = "System support",
    gov_free_to_organize = "Governance", dem_free_speech = "Governance",
    election_free_fair = "Governance",
    political_interest = "Control", rich_poor_treated_equally = "Control",
    nat_willing_emigrate = "Control"
  )
  
  item_labels <- c(
    trust_police = "Trust: police",
    trust_national_government = "Trust: national govt",
    trust_president = "Trust: president/CE",
    trust_parliament = "Trust: parliament",
    trust_courts = "Trust: courts",
    trust_civil_service = "Trust: civil service",
    dem_always_preferable = "Democracy always preferable",
    democracy_satisfaction = "Democratic satisfaction",
    dem_extent_current = "Extent of current democracy",
    democracy_suitability = "Democracy suitability",
    system_deserves_support = "System deserves support",
    gov_free_to_organize = "Freedom to organize",
    dem_free_speech = "Free to speak without fear",
    election_free_fair = "Elections free and fair",
    political_interest = "Political interest",
    rich_poor_treated_equally = "Rich/poor treated equally",
    nat_willing_emigrate = "Willing to emigrate"
  )
  
  nr_table <- map_dfr(key_vars, function(v) {
    if (!v %in% names(hk5_nr)) return(NULL)
    protest <- hk5_nr |> filter(period == "Protest")
    postnsl <- hk5_nr |> filter(period == "Post-NSL")
    tibble(
      Category = item_categories[v],
      Item = item_labels[v],
      `% Protest` = round(100 * sum(!is.na(protest[[v]])) / nrow(protest), 1),
      `% Post-NSL` = round(100 * sum(!is.na(postnsl[[v]])) / nrow(postnsl), 1),
      Delta = round(`% Post-NSL` - `% Protest`, 1)
    )
  })
  
  nr_table |>
    arrange(Category, Delta) |>
    select(-Category) |>
    kable(booktabs = TRUE,
          col.names = c("Item", "Response % (Protest)", "Response % (Post-NSL)", 
                        "\u0394 (pp)")) |>
    kable_styling(latex_options = c("hold_position", "scale_down"))
} else {
  cat("Data not loaded.")
}
```

Trust items maintain near-identical response rates across periods (mean change: $+0.3$ pp). Normative democracy items decline by a mean of $4.6$ pp, governance items by $2.7$ pp, while control items remain stable. This differential pattern is consistent with politically motivated non-response on items requiring explicit democratic commitments.


# Robust Standard Errors {#sec-robust-se}

To address potential heteroskedasticity and assess sensitivity to standard error computation, @tbl-robust-se reports covariate-adjusted estimates with both conventional OLS standard errors and heteroskedasticity-consistent (HC2) robust standard errors for the four primary outcomes.

```{r}
#| label: tbl-robust-se
#| tbl-cap: "Covariate-adjusted Post-NSL effect estimates with conventional and HC2 robust standard errors."

library(sandwich)
library(lmtest)

hk5_robust <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL")) |>
  mutate(post_nsl = as.integer(period == "Post-NSL"))

outcomes <- c("trust_police", "trust_national_government",
              "democracy_suitability", "gov_free_to_organize")
labels_robust <- c("Trust in Police", "Trust in National Government",
            "Democracy Suitability", "Freedom to Organize")

# Build covariate formula dynamically based on available variables
covs <- c()
if ("age" %in% names(hk5_robust) && sum(!is.na(hk5_robust$age)) > 100) covs <- c(covs, "age")
if ("female" %in% names(hk5_robust) && sum(!is.na(hk5_robust$female)) > 100) {
  covs <- c(covs, "female")
} else if ("gender" %in% names(hk5_robust) && sum(!is.na(hk5_robust$gender)) > 100) {
  covs <- c(covs, "gender")
}
if ("education_level_clean" %in% names(hk5_robust) && sum(!is.na(hk5_robust$education_level_clean)) > 100) {
  covs <- c(covs, "education_level_clean")
} else if ("education_level" %in% names(hk5_robust)) {
  hk5_robust <- hk5_robust |> mutate(
    edu_clean = case_when(
      education_level %in% c(-1, 11, 97, 98, 99) ~ NA_real_,
      TRUE ~ as.numeric(education_level)
    )
  )
  if (sum(!is.na(hk5_robust$edu_clean)) > 100) covs <- c(covs, "edu_clean")
}

cov_string <- if (length(covs) > 0) paste("+", paste(covs, collapse = " + ")) else ""

robust_results <- map2_dfr(outcomes, labels_robust, function(outcome, label) {
  f <- as.formula(paste(outcome, "~ post_nsl", cov_string))
  m <- lm(f, data = hk5_robust)
  robust_test <- coeftest(m, vcov = vcovHC(m, type = "HC2"))

  tibble(
    Outcome = label,
    b = round(coef(m)["post_nsl"], 3),
    `SE (conv)` = round(summary(m)$coefficients["post_nsl", "Std. Error"], 3),
    `SE (HC2)` = round(robust_test["post_nsl", "Std. Error"], 3),
    `p (conv)` = format.pval(summary(m)$coefficients["post_nsl", "Pr(>|t|)"], digits = 3),
    `p (HC2)` = format.pval(robust_test["post_nsl", "Pr(>|t|)"], digits = 3)
  )
})

robust_results |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("hold_position")) |>
  footnote(
    general = paste0("All models control for ", paste(covs, collapse = ", "),
                     ". HC2 standard errors are heteroskedasticity-consistent. ",
                     "Substantive conclusions are unchanged across specifications."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```

The HC2 robust standard errors are substantively indistinguishable from conventional OLS standard errors across all four primary outcomes, confirming that heteroskedasticity does not meaningfully affect inference.

# Cross-National Sensitivity Gradient: Turkey and Russia {#sec-wvs-gradient}

The main analysis documents a sensitivity gradient in Hong Kong's ABS Wave 5 data: trust in coercive institutions surged post-NSL ($d \approx +0.49$ for police) while democratic evaluations collapsed ($d = -0.54$), with the magnitude of inflation tracking each institution's coercive role ($r = -0.85$ between pre-specified sensitivity rankings and observed effect sizes). Differential item non-response reinforced this pattern, with normative democracy items showing a mean response rate decline of 4.6 percentage points while trust items remained stable across fieldwork periods. A natural question is whether these patterns are specific to Hong Kong's particular political context and the ABS instrument, or whether the sensitivity gradient captures a more general phenomenon of survey measurement distortion under autocratization.

To assess this, we apply the same logic to two independent cases using an entirely different survey program: the World Values Survey (WVS). Turkey (Wave 6: 2011, Wave 7: 2018) experienced dramatic autocratization following the 2016 failed coup attempt, with mass purges of civil servants, academics, and journalists, declaration of a state of emergency, and constitutional changes consolidating executive power. Russia (Wave 6: 2011, Wave 7: 2017) underwent intensified authoritarian consolidation following the 2014 Crimea annexation, including the suppression of independent media, criminal prosecution of opposition figures, and expansion of surveillance and censorship infrastructure. Both cases are well-documented in the comparative autocratization literature and offer variation in the *type* of repressive mechanism---Turkey's post-coup emergency powers versus Russia's incremental media and civil society squeeze---while sharing the core theoretical condition: a sharp increase in the cost of expressing regime-critical attitudes between survey waves.

We classify WVS items into three categories by expected political sensitivity, paralleling the ABS classification used in the main analysis. *Democracy and political system items* (e.g., importance of democracy, evaluations of democratic governance, attitudes toward strong-leader or military rule) correspond to the low-sensitivity evaluative items in the ABS framework---items where respondents assess the political system in abstract terms. *Institutional confidence items* (e.g., confidence in government, police, courts) correspond to the high-sensitivity trust items in the ABS---items requiring direct evaluation of specific state actors with coercive capacity. *Apolitical control items* (e.g., happiness, life satisfaction, self-reported health) serve as placebo benchmarks with no expected political sensitivity, analogous to the falsification tests reported in Online Appendix E.1.

Note that the WVS and ABS frameworks classify the *direction* of the gradient differently: in the ABS analysis, trust in coercive institutions is coded as high-sensitivity because respondents face direct pressure to express confidence in the police and government, while abstract democratic evaluations are lower-sensitivity. In the WVS analysis, democracy items show *higher non-response* than trust items, which may appear to reverse the ABS pattern. The reconciliation is straightforward: the WVS democracy items ask respondents to evaluate *regime type preferences* (e.g., whether democracy or strong-leader rule is desirable), which in an autocratizing context amounts to a direct assessment of the regime's legitimacy---a high-cost response. The ABS trust items and WVS confidence items both evaluate specific institutions, but the *response distortion* operates in opposite directions: trust items are inflated (respondents overstate confidence) while democracy items are suppressed (respondents either refuse to answer or understate democratic commitment). Both patterns are consistent with preference falsification; they simply manifest differently depending on whether the "safe" response is agreement (trust) or non-response (democratic values).

```{r wvs-gradient-load}
#| include: false

# Load pre-computed results from 08_wvs_sensitivity_gradient.R
if (file.exists("../analysis/results/wvs_sensitivity_gradient.RData")) {
  load("../analysis/results/wvs_sensitivity_gradient.RData")
} else {
  warning("WVS sensitivity gradient results not found. Run analysis/08_wvs_sensitivity_gradient.R first.")
}
```

```{r}
#| label: tbl-wvs-gradient
#| tbl-cap: "Sensitivity gradient in autocratizing contexts: item non-response rates (%) and mean shifts across WVS Waves 6–7 for Turkey and Russia. Items are grouped by expected political sensitivity. Non-response includes don't know, refusal, and item-level missing."

if (exists("wvs_gradient_results")) {
  tbl_data <- wvs_gradient_results |>
    filter(label != "Army rule") |>  # Turkey: not fielded in W7
    mutate(
      sensitivity = factor(sensitivity, levels = c("Democracy", "Trust", "Control")),
      w6_nr = round(w6_nr, 1),
      w7_nr = round(w7_nr, 1),
      nr_change = round(nr_change, 1),
      w6_mean = round(w6_mean, 2),
      w7_mean = round(w7_mean, 2),
      mean_change = round(mean_change, 2)
    ) |>
    select(country, sensitivity, label, w6_nr, w7_nr, nr_change, w6_mean, w7_mean, mean_change) |>
    arrange(country, sensitivity, desc(w7_nr))

  # Build table for each country
  for (cty in c("Turkey", "Russia")) {
    cty_data <- tbl_data |>
      filter(country == cty) |>
      select(-country) |>
      rename(
        Category = sensitivity,
        Item = label,
        `W6 NR%` = w6_nr,
        `W7 NR%` = w7_nr,
        `Δ NR` = nr_change,
        `W6 Mean` = w6_mean,
        `W7 Mean` = w7_mean,
        `Δ Mean` = mean_change
      )

    cat(sprintf("\n**%s (W6: %s, W7: %s)**\n\n",
                cty,
                ifelse(cty == "Turkey", "2011", "2011"),
                ifelse(cty == "Turkey", "2018", "2017")))

    pack_groups <- cty_data |>
      group_by(Category) |>
      summarise(n = n(), .groups = "drop")

    print(
      cty_data |>
        select(-Category) |>
        kable(booktabs = TRUE, align = c("l", rep("r", 6))) |>
        kable_styling(latex_options = c("hold_position", "scale_down")) |>
        pack_rows(index = setNames(pack_groups$n, pack_groups$Category))
    )
    cat("\n")
  }
}
```

@tbl-wvs-gradient presents the results. @tbl-wvs-gradient-tests confirms that the differences are statistically significant.

```{r}
#| label: tbl-wvs-gradient-tests
#| tbl-cap: "Statistical tests for the sensitivity gradient. Panel A reports omnibus chi-squared tests for the association between item category and non-response (pooled across items within each category). Panel B reports pairwise comparisons with odds ratios quantifying the relative odds of non-response for democracy items versus trust items."

if (exists("wvs_gradient_respondent_tests") && exists("wvs_gradient_odds_ratios")) {

  # Panel A: Omnibus tests
  omnibus <- bind_rows(
    wvs_gradient_respondent_tests$Turkey$omnibus,
    wvs_gradient_respondent_tests$Russia$omnibus
  ) |>
    mutate(
      chi_sq = round(chi_sq, 1),
      p_value = format.pval(p_value, digits = 3)
    ) |>
    rename(Country = country, `Chi-sq` = chi_sq, df = df, p = p_value)

  cat("**Panel A: Omnibus chi-squared tests**\n\n")
  print(
    omnibus |>
      kable(booktabs = TRUE) |>
      kable_styling(latex_options = c("hold_position"))
  )

  # Panel B: Odds ratios (Democracy vs Trust NR)
  cat("\n\n**Panel B: Democracy vs Trust non-response odds ratios**\n\n")
  or_tbl <- wvs_gradient_odds_ratios |>
    mutate(across(c(odds_ratio, or_ci_lo, or_ci_hi), ~round(.x, 2))) |>
    rename(Country = country, OR = odds_ratio, `95% CI lo` = or_ci_lo, `95% CI hi` = or_ci_hi)

  print(
    or_tbl |>
      kable(booktabs = TRUE) |>
      kable_styling(latex_options = c("hold_position"))
  )
}
```

In both Turkey and Russia, democracy-related items exhibit substantially higher non-response rates in Wave 7 than institutional confidence items, which in turn exceed apolitical control items. In Turkey, the average Wave 7 non-response rate for democracy items (excluding the army rule item, which was not fielded) is approximately 7% compared to 1.4% for trust items and 0.4% for control items. In Russia, the corresponding figures are approximately 12%, 4.5%, and 1.3%. This monotonic gradient---democracy items > trust items > control items---parallels the pattern documented in Hong Kong using the ABS.

Notably, Turkey's "army rule" item (Q237) was entirely absent from the Wave 7 questionnaire, constituting the most extreme form of item suppression: the survey organization itself removed the most politically sensitive question in the post-coup context. This absence underscores the practical relevance of the sensitivity gradient for survey practitioners working in autocratizing environments.

The trust paradox is also visible in both cases. Despite substantial democratic backsliding as measured by V-Dem indices, confidence in state institutions either increased or remained stable between waves. In Turkey, confidence in government, police, courts, and armed forces all shifted toward greater trust. In Russia, the same pattern holds, with confidence in parliament showing the largest increase. Meanwhile, Russia's democracy evaluation score *increased* by over one full point on the 1--10 scale---a pattern more consistent with rally-around-the-flag effects and preference falsification than with genuine improvements in democratic governance.

```{r}
#| label: fig-wvs-gradient
#| fig-cap: "Sensitivity gradient in autocratizing contexts: Wave 7 item non-response rates by category. Democracy-related items exhibit systematically higher non-response than institutional trust items, which exceed apolitical controls. Turkey: 2011→2018 (post-2016 coup attempt); Russia: 2011→2017 (post-Crimea annexation). 'Army rule' item excluded (not fielded in Turkey Wave 7)."
#| fig-width: 10
#| fig-height: 7

if (exists("wvs_gradient_results")) {
  fig_data <- wvs_gradient_results |>
    filter(label != "Army rule") |>
    mutate(
      sensitivity = factor(sensitivity, levels = c("Democracy", "Trust", "Control")),
      label = fct_reorder2(label, as.numeric(sensitivity), -w7_nr,
                           .fun = function(s, nr) s[1] * 1000 - mean(nr))
    )

  ggplot(fig_data, aes(x = w7_nr, y = label, fill = sensitivity)) +
    geom_col(width = 0.7, alpha = 0.85) +
    facet_wrap(~country, scales = "free_x") +
    scale_fill_manual(
      values = c(Democracy = "#CC3311", Trust = "#0077BB", Control = "#009988"),
      name = "Item category"
    ) +
    labs(
      x = "Item non-response rate (%, Wave 7)",
      y = NULL
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      strip.text = element_text(face = "bold", size = 13),
      panel.grid.major.y = element_line(color = "grey92"),
      panel.grid.minor = element_blank()
    )
}
```

These findings demonstrate that the sensitivity gradient is not idiosyncratic to the Hong Kong context or the Asian Barometer Survey instrument. The same pattern---systematic differential non-response scaling with item political sensitivity---emerges independently in WVS data from two additional autocratizing contexts with distinct political histories, repressive mechanisms, and survey implementations. This cross-national replication strengthens the case that the sensitivity gradient constitutes a generalizable diagnostic for survey validity under political pressure.

# Trust vs. Confidence Wording and Cross-Instrument Validation {#sec-trust-confidence}

## The Trust/Confidence Distinction {#sec-trust-confidence-distinction}

The primary analysis uses ABS items asking how much respondents "trust" each institution (1--4 scale), while the Turkey replication uses WVS items asking how much "confidence" respondents have in each institution (1--4 scale). A potential concern is that these constructs differ---"trust" may tap relational or interpersonal judgments, while "confidence" implies assessments of institutional competence [@Newton2000-ce; @Zmerli2011-ji].

Three considerations mitigate this concern. First, the sensitivity gradient is computed within each instrument: ABS trust items are compared to ABS democratic evaluation items in Hong Kong, and WVS confidence items are compared to WVS democratic evaluation items in Turkey. The trust/confidence distinction therefore does not confound the core within-instrument test---it matters only for the cross-case comparison of gradient magnitudes ($r = -0.85$ in Hong Kong vs. $r = -0.68$ in Turkey).

Second, the measurement invariance literature provides reassurance. Factor-analytic studies consistently find that trust and confidence items load on the same latent factor in cross-national comparisons [@Marien2011-ms; @Zmerli2011-ji]. Item response theory analyses similarly show that the trust/confidence distinction does not generate meaningful differential item functioning across survey programs [@Ariely2011-ht].

Third, the direction of any bias is conservative for the cross-case comparison. If "confidence" taps a more cognitive, competence-based dimension, it should be *less* susceptible to preference falsification than the more affectively loaded "trust" items---respondents might strategically report trust in a feared institution without genuinely believing it is competent. This would attenuate the WVS gradient relative to the ABS gradient, meaning the observed Turkey gradient ($r = -0.68$) represents a lower bound.

## Same-Country, Same-Period Cross-Instrument Comparison {#sec-cross-instrument}

To directly assess cross-instrument equivalence, we exploit the temporal overlap between ABS Wave 5 and WVS Wave 7 fieldwork. We focus on Thailand (ABS W5: 2019, WVS W7: 2018; ~1 year gap) and the Philippines (ABS W5: 2019, WVS W7: 2019; ~0 years gap), computing weighted means for the five institutions present in both instruments: police, military/armed forces, government, courts, and parliament.

```{r}
#| label: tbl-cross-instrument
#| tbl-cap: "Cross-instrument comparison of institutional trust (ABS) and confidence (WVS) means. All items scored 1--4, higher = more trust/confidence. Weighted estimates. Rank = descending order by mean within each instrument (1 = highest)."
#| echo: false
#| warning: false

load("../analysis/results/cross_instrument_validation.RData")

cross_instrument_results$means |>
  select(Country = country, Institution = institution,
         `ABS Mean` = abs_mean, `ABS Rank` = abs_rank,
         `WVS Mean` = wvs_mean, `WVS Rank` = wvs_rank,
         `Diff (WVS−ABS)` = mean_diff) |>
  mutate(across(where(is.numeric) & !contains("Rank"),
                ~ sprintf("%.2f", .x)),
         across(contains("Rank"), as.integer)) |>
  kableExtra::kbl(align = c("l", "l", "r", "r", "r", "r", "r")) |>
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9, full_width = FALSE
  ) |>
  kableExtra::pack_rows("Thailand", 1, 5) |>
  kableExtra::pack_rows("Philippines", 6, 10)
```

Three patterns emerge from @tbl-cross-instrument. First, *mean levels differ systematically* between instruments: ABS trust means in Thailand are 0.25--0.61 points higher than WVS confidence means for the same institutions, consistent with known scale calibration differences between survey programs. This mean-level gap is expected and does not threaten the gradient methodology, which relies on *within-instrument* rank orderings rather than cross-instrument level comparisons.

```{r}
#| echo: false

rc <- cross_instrument_results$rank_correlations
tha_r <- rc |> filter(country == "Thailand")
phl_r <- rc |> filter(country == "Philippines")
```

Second, *institutional rank orderings are broadly preserved across instruments*. In Thailand, the Spearman rank correlation between ABS and WVS institutional means is $\rho$ = `r sprintf("%.2f", tha_r$spearman_r)`. Both instruments identify military and courts as the most trusted institutions and parliament as the least trusted, with police and government occupying intermediate positions. The Philippine rank-order correlation is weaker ($\rho$ = `r sprintf("%.2f", phl_r$spearman_r)`), though both instruments agree that military ranks highest and parliament lowest---the key anchors for gradient inference.

```{r}
#| echo: false

ng <- cross_instrument_results$null_gradients
tha_ng <- ng |> filter(country == "Thailand")
phl_ng <- ng |> filter(country == "Philippines")
```

Third, we compute *null sensitivity gradients*---the correlation between institutional coercive salience (police = most coercive, parliament = least) and mean trust/confidence levels---for these non-autocratizing contexts.^[Thailand under Prayuth and the Philippines under Duterte experienced democratic erosion during this period, complicating the "null" interpretation. The gradients should therefore be interpreted cautiously as comparative baselines rather than clean nulls.] In Thailand, the ABS null gradient is $r$ = `r sprintf("%.2f", tha_ng$abs_gradient_r)` and the WVS null gradient is $r$ = `r sprintf("%.2f", tha_ng$wvs_gradient_r)` (neither statistically significant, $n = 5$ institutions). Critically, both instruments produce *similar gradient magnitudes within the same country*, suggesting that the trust/confidence wording distinction does not systematically bias the gradient test. The cross-case gradient magnitude comparison (Hong Kong $r = -0.85$ vs. Turkey $r = -0.68$) can therefore be interpreted with greater confidence, since the attenuation in Turkey is unlikely to be an artifact of item wording.

# Extending the Sensitivity Gradient to Africa: Burkina Faso {#sec-burkina-faso}

```{r}
#| echo: false
#| warning: false

load("../analysis/results/afro_sensitivity_gradient.RData")
res <- afro_gradient_results
items <- res$item_results
```

## The Burkina Faso Natural Experiment {#sec-bfa-experiment}

On September 30, 2022, Captain Ibrahim Traoré led Burkina Faso's second coup of the year, deposing the military government of Lieutenant Colonel Paul-Henri Sandaogo Damiba that had itself seized power in January 2022. The September coup dissolved the national government, suspended the Constitution, sealed state borders, and imposed a strict curfew [@Brailey2024-mi]. Critically, Afrobarometer Round 9 fieldwork in Burkina Faso was underway during the coup, enabling a within-wave pre/post comparison analogous to the Hong Kong NSL split.

Following @Brailey2024-mi, we drop the transition period (September 30--October 2) and compare respondents interviewed before the coup ($n$ = `r res$n_pre`, September 20--29) with those interviewed after ($n$ = `r res$n_post`, October 4--12). This design exploits the same identification strategy as the Hong Kong analysis---survey fieldwork that happened to straddle a discrete autocratization shock---but in a different continent, using a different survey program (Afrobarometer rather than ABS or WVS), and with a different type of shock (military coup rather than legislative crackdown or executive consolidation).

The available item battery includes six institutional trust items (army, president, police, courts, electoral commission, churches; all 1--4 scales) and three democratic evaluation items (satisfaction with democracy, perceived extent of democracy, and preference for democracy).^[Afrobarometer did not field the parliament, ruling party, or opposition party trust items in Burkina Faso Round 9, reducing the available battery relative to Hong Kong and Turkey. The core coercive/political institutions (army, president, police, courts) remain available.] We assign sensitivity rankings following the same logic: army and president (now the coup leader) are most coercive (ranks 1--2), police rank 3, courts and elections are medium-sensitivity political institutions (ranks 4--5), churches serve as a non-political benchmark (rank 6), and democratic evaluations are lowest sensitivity (ranks 7--9).

## Results {#sec-bfa-results}

```{r}
#| label: tbl-bfa-gradient
#| tbl-cap: "Burkina Faso pre/post coup comparison. Weighted means on original scales (trust items: 1--4, higher = more trust; dem. satisfaction and how democratic: 1--4; dem. preferable: 1--3). Cohen's d computed from pooled standard deviations. Positive d indicates post-coup increase."
#| echo: false
#| warning: false

items |>
  select(Item = label, Category = sensitivity,
         `Pre mean` = pre_mean, `Post mean` = post_mean,
         `d` = cohens_d, `p` = p_value) |>
  mutate(
    across(c(`Pre mean`, `Post mean`), ~ sprintf("%.2f", .x)),
    d = sprintf("%+.3f", d),
    p = ifelse(p < 0.001, "< .001",
        ifelse(p < 0.01, sprintf("%.3f", p),
        ifelse(p < 0.05, sprintf("%.3f", p),
               sprintf("%.3f", p))))
  ) |>
  kableExtra::kbl(align = c("l", "l", "r", "r", "r", "r")) |>
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9, full_width = FALSE
  ) |>
  kableExtra::pack_rows("High sensitivity (coercive)", 1, 3) |>
  kableExtra::pack_rows("Medium sensitivity (political)", 4, 6) |>
  kableExtra::pack_rows("Low sensitivity (democratic evaluations)", 7, 9)
```

@tbl-bfa-gradient presents the item-level results. The pattern is notably weaker and more mixed than in Hong Kong or Turkey, consistent with the theoretical complications outlined below.

**Trust--democracy divergence.** The core divergence signature is present but attenuated. Institutional trust items show a small average post-coup increase (mean $d$ = `r sprintf("%+.3f", res$trust_avg_d)`), while democratic evaluation items decline (mean $d$ = `r sprintf("%+.3f", res$dem_avg_d)`), producing a trust--democracy divergence of `r sprintf("%.3f", res$trust_avg_d - res$dem_avg_d)`. The largest single effect is the decline in satisfaction with democracy ($d$ = `r sprintf("%+.3f", items$cohens_d[items$label == "Dem. satisfaction"])`, $p$ = `r sprintf("%.3f", items$p_value[items$label == "Dem. satisfaction"])`). Among trust items, the president---now coup leader Traoré---shows the largest increase ($d$ = `r sprintf("%+.3f", items$cohens_d[items$label == "Trust president"])`, $p$ = `r sprintf("%.3f", items$p_value[items$label == "Trust president"])`).

**Sensitivity gradient.** The gradient correlation across all nine items is $r$ = `r sprintf("%.2f", res$gradient_pearson$estimate)` ($p$ = `r sprintf("%.2f", res$gradient_pearson$p.value)`), in the predicted negative direction but far from statistical significance with only nine items. Within trust items alone, the gradient reverses ($r$ = `r sprintf("%.2f", cor(items$sensitivity_rank[items$sensitivity == "Trust"], items$cohens_d[items$sensitivity == "Trust"]))`), driven by the large church trust increase and the near-zero army shift.

**Critical citizens correlation.** The correlation between democratic preference and democratic satisfaction is near zero in both periods (pre-coup: $r$ = `r sprintf("%.3f", res$cc_pre$estimate)`; post-coup: $r$ = `r sprintf("%.3f", res$cc_post$estimate)`), providing no evidence of the coherence disruption observed in Hong Kong.

**Non-response.** Unlike Hong Kong and Turkey, post-coup non-response rates do not differ between trust and democracy items (OR = 1.00, $\chi^2$ < 0.1, $p$ = 1.00). The absence of differential non-response suggests that respondents were not selectively avoiding politically sensitive questions.

## Interpretation: An Informative Intermediate Case {#sec-bfa-interpretation}

The Burkina Faso results occupy a theoretically informative position between the strong gradient cases (Hong Kong, Turkey) and the null cases (Venezuela, Nicaragua). The trust--democracy divergence is present---trust edges up while democratic satisfaction declines---but the full sensitivity gradient does not emerge, and the diagnostic signatures of preference falsification (differential non-response, critical citizens disruption) are absent.

Several features of the Burkina Faso context likely explain this attenuation. First, the September 2022 coup had substantial popular support, driven by frustration with the previous military government's failure to address jihadist violence. @Brailey2024-mi report that 66% of Burkinabè supported military intervention "when leaders abuse power." Genuine increases in trust in the coup leader are therefore expected alongside any falsification-driven increases, and these two mechanisms cannot be separated in the data. Second, trust in the army was already near ceiling before the coup (mean = `r sprintf("%.2f", items$pre_mean[items$label == "Trust army"])` on a 1--4 scale), leaving little room for a post-coup increase regardless of mechanism. Third, the repressive environment following the coup---border closures and curfews rather than targeted surveillance of political expression---may generate qualitatively different pressures on survey responses than the targeted repression in Hong Kong (arrests of activists and journalists) or Turkey (purges of specific sectors).

These complications are analytically productive. The sensitivity gradient framework identifies a *specific* distortion signature associated with targeted political repression of dissent---not a universal feature of all political shocks. The Burkina Faso case, alongside the Venezuela and Nicaragua null results, helps establish the boundary conditions: the gradient emerges most clearly when (a) repression specifically targets political expression, (b) institutions differ meaningfully in coercive salience, and (c) there is not a concurrent genuine rally effect that overwhelms the falsification signal. Military coups with broad popular support---where genuine attitude change and strategic misrepresentation point in the same direction---represent a harder test for the framework, and the attenuated results are consistent with this theoretical expectation.

```{r}
#| label: fig-bfa-gradient
#| fig-cap: "Burkina Faso sensitivity gradient. Each point represents one survey item, plotted by sensitivity rank (1 = most coercive) and Cohen's d (pre→post coup mean shift). Dashed line shows the OLS fit. The gradient is in the predicted negative direction but weak and non-significant, consistent with the theoretical complications discussed in the text."
#| echo: false
#| warning: false
#| fig-width: 7
#| fig-height: 5

ggplot(items, aes(x = sensitivity_rank, y = cohens_d,
                  color = sensitivity)) +
  geom_hline(yintercept = 0, color = "grey80") +
  geom_smooth(method = "lm", se = TRUE, color = "grey40",
              linetype = "dashed", linewidth = 0.5) +
  geom_point(size = 3) +
  ggrepel::geom_text_repel(aes(label = label), size = 3,
                           show.legend = FALSE, max.overlaps = 15) +
  scale_color_manual(
    values = c(Trust = "#0077BB", Democracy = "#CC3311"),
    name = "Category"
  ) +
  labs(
    x = "Sensitivity rank (1 = most coercive)",
    y = expression(paste("Cohen's ", italic(d), " (pre ", symbol('\256'), " post coup)"))
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top")
```

# Sensitivity Gradient in Latin America: Venezuela and Nicaragua {#sec-latin-america}

```{r}
#| echo: false
#| warning: false

load("../analysis/results/venezuela_gradient.RData")
load("../analysis/results/nicaragua_gradient.RData")
```

## Venezuela: Uniform Institutional Collapse {#sec-venezuela}

### Case Background and Data

Venezuela's authoritarian consolidation under Nicolás Maduro accelerated dramatically following the July 2017 National Constituent Assembly, which effectively superseded the opposition-controlled National Assembly and concentrated executive authority. The period from 2017 onward combined targeted political repression---arrests of opposition leaders, suppression of protests, criminalization of dissent---with catastrophic economic mismanagement producing hyperinflation exceeding one million percent by 2018. Approximately 7.7 million Venezuelans emigrated between 2015 and 2023, constituting the largest displacement crisis in the Western Hemisphere. This case provides the strongest available test of the compositional selection mechanism: if emigration of regime critics were the dominant driver of the sensitivity gradient, Venezuela's massive politically non-random exodus should produce the strongest gradient in the analysis.

Latinobarómetro data for Venezuela are available for 2015 (pre-shock baseline), 2017 (partially overlapping with the July constituent assembly), 2018, 2020, and 2023. Following the preregistration filed on OSF prior to data access (February 14, 2026), the analysis compares pooled pre-shock waves (2015; $N \approx$ `r round(mean(ven_gradient_results$w_before_n), 0)`) against pooled post-shock waves (2018, 2020, 2023; $N \approx$ `r round(mean(ven_gradient_results$w_after_n), 0)` per item). The 2017 wave is excluded from the primary comparison due to fieldwork timing uncertainty relative to the July constituent assembly. All estimates incorporate Latinobarómetro survey weights.

### Item Classification

Items are classified following the sensitivity gradient framework, adapted to the Latinobarómetro instrument. High-sensitivity items evaluate institutions with direct coercive capacity: confidence in police, national government, and armed forces (all coded 1--4, higher = more confidence). Medium-sensitivity items assess political institutions: confidence in parliament, courts, electoral authority, and political parties. Low-sensitivity items capture abstract democratic evaluations: satisfaction with democracy and support for democracy as the best system of government. A control item (confidence in the church) serves as an apolitical benchmark.

### Results

```{r}
#| label: tbl-ven-gradient
#| tbl-cap: "Venezuela sensitivity gradient results (Latinobarómetro, pre-2017 vs. post-2017). Items coded 1--4 (higher = more confidence), except Dem. Best System (1--4). Cohen's $d$ = (post $-$ pre) / pooled SD; positive values indicate higher post-shock means. Survey weights applied. Pre $N \\approx$ 1,195; Post $N \\approx$ 1,185 per item."
#| echo: false
#| warning: false

ven_gradient_results |>
  mutate(
    delta = w_after_mean - w_before_mean,
    pre   = sprintf("%.2f", w_before_mean),
    post  = sprintf("%.2f", w_after_mean),
    delta_fmt = sprintf("%+.3f", delta),
    d_fmt = sprintf("%+.3f", cohens_d),
    p_fmt = ifelse(p_value < 0.001, "< .001",
            ifelse(p_value < 0.01,  sprintf("%.3f", p_value),
                                    sprintf("%.3f", p_value)))
  ) |>
  select(Item = label, Category = category,
         `Pre mean` = pre, `Post mean` = post,
         `$\\Delta$` = delta_fmt, `$d$` = d_fmt, `$p$` = p_fmt) |>
  kableExtra::kbl(align = c("l", "l", "r", "r", "r", "r", "r"),
                  escape = FALSE) |>
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9, full_width = FALSE
  ) |>
  kableExtra::pack_rows("High sensitivity (coercive)",       1, 3) |>
  kableExtra::pack_rows("Medium sensitivity (political)",    4, 7) |>
  kableExtra::pack_rows("Low sensitivity (democratic)",      8, 9) |>
  kableExtra::pack_rows("Control",                          10, 10)
```

@tbl-ven-gradient presents the full item-level results. The pattern is strikingly inconsistent with the sensitivity gradient prediction. All institutional confidence items declined significantly, including high-sensitivity items (armed forces $d =$ `r sprintf("%.2f", ven_gradient_results$cohens_d[ven_gradient_results$label == "Conf. Armed Forces"])`, government $d =$ `r sprintf("%.2f", ven_gradient_results$cohens_d[ven_gradient_results$label == "Conf. Government"])`, police $d =$ `r sprintf("%.2f", ven_gradient_results$cohens_d[ven_gradient_results$label == "Conf. Police"])`). Democratic satisfaction fell nearly as steeply ($d =$ `r sprintf("%.2f", ven_gradient_results$cohens_d[ven_gradient_results$label == "Dem. Satisfaction"])`), and support for democracy as the best system collapsed substantially ($d =$ `r sprintf("%.2f", ven_gradient_results$cohens_d[ven_gradient_results$label == "Dem. Best System"])`). The one item that did not decline significantly is the apolitical control: confidence in churches changed marginally ($d =$ `r sprintf("%+.2f", ven_gradient_results$cohens_d[ven_gradient_results$label == "Conf. Churches"])`, $p =$ `r sprintf("%.2f", ven_gradient_results$p_value[ven_gradient_results$label == "Conf. Churches"])`). The gradient correlation across all items is $r =$ `r sprintf("%.2f", ven_gradient_r_all)`---statistically indistinguishable from zero.

The null gradient is theoretically informative. Venezuela's combination of targeted political repression and catastrophic economic failure produced disillusionment so severe and evenly distributed that genuine attitude change overwhelmed any strategic compliance signal. The massive emigration removed critics from the survey frame, but those who remained were not falsifying regime support: they too had lost faith in all institutions. This confirms a boundary condition of the framework: the sensitivity gradient requires that strategic compliance remain a viable respondent strategy, which presupposes that the regime retains sufficient coercive authority and apparent legitimacy to make strategic response rational. When regime failure is too total and visible, this condition fails.

The gradient figure for Venezuela is presented in the main text.

## Nicaragua: Inverted Gradient Under Overt Repression {#sec-nicaragua}

### Case Background and Data

Nicaragua's April 2018 crackdown on anti-government protests---which left over 300 civilians dead and triggered the arrest or exile of opposition leaders, independent journalists, and civil society figures under President Daniel Ortega---provides a second Latin American test. Unlike Venezuela's gradual economic-political collapse, Nicaragua's repression was rapid, targeted, and highly visible: live ammunition against protesters, paramilitary violence captured on video, and international condemnation were immediate features of the crisis. Nicaragua's V-Dem liberal democracy index fell from 0.38 in 2017 to 0.17 by 2020.

Latinobarómetro data are available for Nicaragua in 2017 (pre-shock baseline; $N \approx$ `r round(mean(nic_gradient_results$w_before_n), 0)`) and 2019--2021 (post-shock; $N \approx$ `r round(mean(nic_gradient_results$w_after_n), 0)` per wave). The analysis compares the 2017 wave against pooled 2019--2021 waves, following the same preregistration logic as the Venezuela analysis. All estimates incorporate survey weights.

### Results

```{r}
#| label: tbl-nic-gradient
#| tbl-cap: "Nicaragua sensitivity gradient results (Latinobarómetro, 2017 vs. 2019--2021). Items coded 1--4 (higher = more confidence), except Dem. Best System (1--4). Cohen's $d$ = (post $-$ pre) / pooled SD; all values negative (universal decline). Survey weights applied. Pre $N \\approx$ 1,200; Post $N \\approx$ 1,200 per wave."
#| echo: false
#| warning: false

nic_gradient_results |>
  mutate(
    delta = w_after_mean - w_before_mean,
    pre   = sprintf("%.2f", w_before_mean),
    post  = sprintf("%.2f", w_after_mean),
    delta_fmt = sprintf("%+.3f", delta),
    d_fmt = sprintf("%+.3f", cohens_d),
    p_fmt = ifelse(p_value < 0.001, "< .001",
            ifelse(p_value < 0.01,  sprintf("%.3f", p_value),
                                    sprintf("%.3f", p_value)))
  ) |>
  select(Item = label, Category = category,
         `Pre mean` = pre, `Post mean` = post,
         `$\\Delta$` = delta_fmt, `$d$` = d_fmt, `$p$` = p_fmt) |>
  kableExtra::kbl(align = c("l", "l", "r", "r", "r", "r", "r"),
                  escape = FALSE) |>
  kableExtra::kable_styling(
    latex_options = c("hold_position"),
    font_size = 9, full_width = FALSE
  ) |>
  kableExtra::pack_rows("High sensitivity (coercive)",       1, 3) |>
  kableExtra::pack_rows("Medium sensitivity (political)",    4, 7) |>
  kableExtra::pack_rows("Low sensitivity (democratic)",      8, 9) |>
  kableExtra::pack_rows("Control",                          10, 10)
```

@tbl-nic-gradient presents the full item-level results. Every item in the battery declined significantly, including the control item (confidence in churches, $d =$ `r sprintf("%.2f", nic_gradient_results$cohens_d[nic_gradient_results$label == "Conf. Churches"])`). Crucially, democratic satisfaction declined most steeply ($d =$ `r sprintf("%.2f", nic_gradient_results$cohens_d[nic_gradient_results$label == "Dem. Satisfaction"])`), while high-sensitivity coercive institution items declined substantially but less than democratic evaluations. The gradient correlation is $r =$ `r sprintf("%+.2f", nic_gradient_r_all)`---positive, meaning the most coercive institutions declined least while low-sensitivity democratic evaluation items declined most. This is the precise inverse of the falsification-consistent pattern.

The inverted gradient has a theoretically coherent interpretation. When repression involves live-fire violence against civilian protesters---events that are publicly documented, internationally condemned, and directly experienced by a large fraction of the population---the strategic calculus of survey response changes fundamentally. In the Hong Kong and Turkey cases, respondents faced diffuse surveillance risk and the possibility of legal consequences for expressed dissent, making strategic compliance rational. In Nicaragua's April 2018 context, the regime's violence was so overt that the "safety" of expressing loyalty through survey responses was psychologically unavailable to respondents who had witnessed or experienced direct state violence. Under these conditions, genuine disillusionment dominates: even institutions normally protected by fear-driven falsification collapsed in measured confidence.

The inverted gradient does not indicate that Nicaragua is "beyond" measurement distortion in a technical sense. It indicates that the conditions generating the falsification-consistent gradient were absent: targeted differential cost of honest response requires that respondents believe expressing dissatisfaction is more costly than expressing loyalty, a belief that becomes difficult to sustain when the regime is conducting mass killings in full public view.

The gradient figure for Nicaragua is presented alongside Venezuela in the main text.

## Comparison: Venezuela, Nicaragua, and the Positive Cases {#sec-latam-comparison}

The contrast between the Latin American cases and Hong Kong/Turkey can be summarized along three dimensions. First, the direction of institutional trust shifts: upward in Hong Kong and Turkey (falsification-consistent), uniformly downward in Venezuela and Nicaragua (genuine collapse). Second, the gradient correlation: strongly negative in Hong Kong ($r = -0.85$) and Turkey ($r = -0.68$), near-zero in Venezuela ($r =$ `r sprintf("%.2f", ven_gradient_r_all)`), and inverted in Nicaragua ($r =$ `r sprintf("%+.2f", nic_gradient_r_all)`). Third, the diagnostic signatures: differential non-response favoring trust items in Hong Kong and Turkey, identical non-response across categories in Burkina Faso, and no gradient-consistent pattern in Venezuela or Nicaragua.

These contrasts establish that the sensitivity gradient is not a universal feature of survey data from autocratizing regimes. It is a signature of a specific type of autocratization: targeted political repression that preserves procedural legitimacy claims, creates differential costs of honest response across item types, and does not destroy the regime's credibility so completely that strategic compliance becomes irrational. Venezuela and Nicaragua represent the failure mode of that condition---regimes whose visible incompetence or overt violence eliminated the informational and coercive preconditions for the gradient to emerge.
