---
title: "Online Appendix: When the Levee Breaks"
author:
  - name: Jeffrey Stark
    affiliations:
      - name: Yonsei University
        department: Department of Sociology
    email: jeffrey.stark@yonsei.ac.kr
date: today
format:
  pdf:
    documentclass: article
    classoption: [letterpaper, 12pt]
    geometry: margin=1in
    linestretch: 2
    number-sections: true
    colorlinks: true
    linkcolor: blue
    urlcolor: blue
    citecolor: blue
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{threeparttable}
        \usepackage{threeparttablex}
        \renewcommand{\thesection}{\Alph{section}}
        \renewcommand{\thesubsection}{\Alph{section}.\arabic{subsection}}
        \renewcommand{\thetable}{\Alph{section}\arabic{table}}
        \renewcommand{\thefigure}{\Alph{section}\arabic{figure}}
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    theme: cosmo
  docx:
    toc: true
    number-sections: true
bibliography: references.bib
csl: apa.csl
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r load-data}
#| include: false
load("../analysis/results/prepared_data.RData")
load("../analysis/results/descriptive_results.RData")
load("../analysis/results/robustness_results.RData")

# Load main analysis results (falsification, FDR, sensitivity)
if (file.exists("../analysis/results/main_analysis_results.RData")) {
  load("../analysis/results/main_analysis_results.RData")
}
```

# Variable Definitions and Coding {#sec-variables}

@tbl-var-definitions presents the full set of variables used in the analysis, with metadata extracted directly from the YAML harmonization specifications used in the data pipeline.

```{r var-metadata-helpers}
#| include: false

library(yaml)

# --- Helper: extract variable metadata from YAML spec files ---
extract_var_metadata <- function(yaml_dir, var_ids) {
  yaml_files <- list.files(yaml_dir, pattern = "\\.yml$", full.names = TRUE)

  all_vars <- list()
  for (f in yaml_files) {
    spec <- yaml::read_yaml(f)
    if (!is.null(spec$variables)) {
      for (v in spec$variables) {
        all_vars[[v$id]] <- v
      }
    }
  }

  map_dfr(var_ids, function(vid) {
    v <- all_vars[[vid]]
    if (is.null(v)) {
      return(tibble(var_id = vid, w5_item = NA_character_,
                    scale_text = NA_character_, recode_note = NA_character_))
    }

    # W5 item code
    w5_item <- if (!is.null(v$source$w5)) as.character(v$source$w5) else NA_character_

    # Scale text
    smin <- v$scale$min
    smax <- v$scale$max
    scale_text <- if (!is.null(smin) && !is.null(smax)) {
      lo_lab <- v$scale$labels[[as.character(smin)]]
      hi_lab <- v$scale$labels[[as.character(smax)]]
      if (!is.null(lo_lab) && !is.null(hi_lab)) {
        paste0(smin, "--", smax, " (", lo_lab, " to ", hi_lab, ")")
      } else {
        paste0(smin, "--", smax)
      }
    } else { NA_character_ }

    # Recoding note for W5
    recode_note <- NA_character_
    harm <- v$harmonize
    if (!is.null(harm$exceptions$w5)) {
      exc <- harm$exceptions$w5
      fn_name <- exc$fn %||% exc$method %||% "custom"
      recode_note <- switch(fn_name,
        "identity" = "None",
        "safe_reverse_4pt" = "Reversed (4pt)",
        "safe_reverse_5pt" = "Reversed (5pt)",
        "safe_reverse_3pt" = "Reversed (3pt)",
        "safe_6pt_to_4pt" = "6pt rescaled to 4pt",
        "safe_4pt_none" = "None",
        "safe_reverse_2pt" = "Reversed (2pt)",
        fn_name
      )
    } else if (!is.null(harm$w5)) {
      w5h <- harm$w5
      fn_name <- w5h$fn %||% w5h$method %||% "custom"
      recode_note <- switch(fn_name,
        "identity" = "None",
        "recode" = "Recoded (see spec)",
        fn_name
      )
    } else if (!is.null(harm$default)) {
      def <- harm$default
      fn_name <- def$fn %||% def$method %||% "custom"
      recode_note <- switch(fn_name,
        "identity" = "None",
        "safe_reverse_4pt" = "Reversed (4pt)",
        "safe_reverse_5pt" = "Reversed (5pt)",
        "safe_reverse_3pt" = "Reversed (3pt)",
        "safe_6pt_to_4pt" = "6pt rescaled to 4pt",
        "safe_4pt_none" = "None",
        "safe_reverse_2pt" = "Reversed (2pt)",
        "middle_identity_5pt" = "Midpoint recoded",
        fn_name
      )
    }

    tibble(var_id = vid, w5_item = w5_item,
           scale_text = scale_text, recode_note = recode_note)
  })
}
```

```{r}
#| label: tbl-var-definitions
#| tbl-cap: "Variable definitions, scales, and coding for all variables used in the analysis. Metadata extracted from YAML harmonization specifications; N statistics computed from the full Hong Kong Wave 5 sample."

source("../../../_data_config.R")
yaml_dir <- file.path(survey_data_prep, "src/config/abs/harmonize_validated/")

# --- Display name and category mapping (controls sort order) ---
var_meta <- tribble(
  ~var_id,                         ~display_name,                          ~category,
  # Institutional Trust (6)
  "trust_police",                  "Trust in police",                      "Institutional Trust",
  "trust_national_government",     "Trust in national government",         "Institutional Trust",
  "trust_president",               "Trust in president/CE",                "Institutional Trust",
  "trust_parliament",              "Trust in parliament",                  "Institutional Trust",
  "trust_courts",                  "Trust in courts",                      "Institutional Trust",
  "trust_civil_service",           "Trust in civil service",               "Institutional Trust",
  # Democratic Attitudes (5)
  "dem_always_preferable",         "Democracy always preferable",          "Democratic Attitudes",
  "democracy_satisfaction",        "Democratic satisfaction",              "Democratic Attitudes",
  "democracy_suitability",         "Democracy suitability",               "Democratic Attitudes",
  "dem_extent_current",            "Extent of current democracy",          "Democratic Attitudes",
  "dem_country_present_govt",      "Current government dem.\\ rating",     "Democratic Attitudes",
  # Governance Perceptions (5)
  "gov_free_to_organize",          "Freedom to organize",                  "Governance Perceptions",
  "dem_free_speech",               "Freedom of speech",                    "Governance Perceptions",
  "election_free_fair",            "Elections free and fair",              "Governance Perceptions",
  "rich_poor_treated_equally",     "Rich and poor treated equally",        "Governance Perceptions",
  "gov_opposition_opportunities",  "Opposition opportunities",             "Governance Perceptions",
  # System Support (3)
  "system_deserves_support",       "System deserves support",              "System Support",
  "system_proud",                  "System pride",                         "System Support",
  "system_needs_change",           "System needs change",                  "System Support",
  # Authoritarian Alternatives (3)
  "strongman_rule",                "Strongman rule",                       "Authoritarian Alternatives",
  "military_rule",                 "Military rule",                        "Authoritarian Alternatives",
  "single_party_rule",             "Single-party rule",                    "Authoritarian Alternatives",
  # Political Efficacy & Participation (4)
  "political_interest",            "Political interest",                   "Political Efficacy \\& Participation",
  "efficacy_ability_participate",  "Efficacy: ability to participate",     "Political Efficacy \\& Participation",
  "efficacy_no_influence",         "Efficacy: no influence",               "Political Efficacy \\& Participation",
  "action_demonstration",          "Demonstration attendance",             "Political Efficacy \\& Participation",
  # Alternative Explanations & Controls (9)
  "dem_rating_china",              "China democratic rating",              "Alternative Explanations \\& Controls",
  "sat_president_govt",            "Satisfaction w/ president/govt",       "Alternative Explanations \\& Controls",
  "dem_country_future",            "Democratic future (10 yrs)",           "Alternative Explanations \\& Controls",
  "dem_essential_harmonized",      "Essential element of democracy",       "Alternative Explanations \\& Controls",
  "dem_procedural_vs_substantive", "Procedural vs.\\ substantive",        "Alternative Explanations \\& Controls",
  "nat_willing_emigrate",          "Willingness to emigrate",             "Alternative Explanations \\& Controls",
  "dem_best_form",                 "Democracy is best form",              "Alternative Explanations \\& Controls",
  "gov_courts_powerless",          "Courts powerless vs.\\ leaders",      "Alternative Explanations \\& Controls",
  "govt_responds_people",          "Government responds to people",        "Alternative Explanations \\& Controls"
)

# --- Extract YAML metadata ---
yaml_info <- extract_var_metadata(yaml_dir, var_meta$var_id)

# --- Compute N valid / N missing from hk5 ---
n_stats <- map_dfr(var_meta$var_id, function(vid) {
  if (vid %in% names(hk5)) {
    n_valid <- sum(!is.na(hk5[[vid]]))
    n_miss  <- sum(is.na(hk5[[vid]]))
  } else {
    n_valid <- NA_integer_
    n_miss  <- NA_integer_
  }
  tibble(var_id = vid, n_valid = n_valid, n_missing = n_miss)
})

# --- Merge everything ---
tbl_data <- var_meta |>
  left_join(yaml_info, by = "var_id") |>
  left_join(n_stats, by = "var_id") |>
  mutate(
    w5_item = ifelse(is.na(w5_item) | w5_item == "NULL", "---", w5_item),
    recode_note = replace_na(recode_note, "---"),
    question_wording = ""
  )

# --- Build table ---
display_tbl <- tbl_data |>
  select(display_name, w5_item, scale_text, recode_note, n_valid, n_missing, question_wording)

# Compute pack_rows groups
cat_counts <- tbl_data |>
  group_by(category) |>
  summarise(n = n(), .groups = "drop") |>
  mutate(end = cumsum(n), start = end - n + 1)

# Preserve category order from var_meta
cat_order <- unique(tbl_data$category)
cat_counts <- cat_counts |> arrange(match(category, cat_order))

kbl <- display_tbl |>
  kable(
    col.names = c("Variable", "ABS W5 Item", "Response Scale",
                   "Recoding", "N Valid", "N Missing", "Question Wording"),
    booktabs = TRUE,
    longtable = TRUE
  ) |>
  kable_styling(
    latex_options = c("scale_down", "hold_position", "repeat_header"),
    font_size = 9
  ) |>
  column_spec(7, width = "3cm")

# Apply pack_rows dynamically
for (i in seq_len(nrow(cat_counts))) {
  kbl <- kbl |>
    pack_rows(cat_counts$category[i], cat_counts$start[i], cat_counts$end[i])
}

kbl |>
  footnote(
    general = "N statistics refer to the full Hong Kong Wave 5 sample (all fieldwork periods). Scale descriptions reflect the harmonized coding direction after any recoding. Question wording column reserved for final manuscript preparation.",
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```


# Mann-Whitney U Tests {#sec-mannwhitney}

All two-sample $t$-tests reported in the main text were cross-validated with non-parametric Mann-Whitney $U$ (Wilcoxon rank-sum) tests. This guards against violations of normality assumptions, which are plausible given that many indicators are measured on short ordinal scales (1--4). @tbl-mw-comparison reports both test results side by side.

```{r}
#| label: tbl-mw-comparison
#| tbl-cap: "Parametric (t-test) vs. non-parametric (Mann-Whitney U) comparison for all pre/post NSL variables."

if (exists("mw_results")) {
  mw_results |>
    select(variable, cohens_d, t_p, t_sig, mw_p, mw_sig, agree) |>
    mutate(across(c(cohens_d), ~round(.x, 3)),
           t_p = format.pval(t_p, digits = 3),
           mw_p = format.pval(mw_p, digits = 3)) |>
    kable(
      col.names = c("Variable", "Cohen's d", "t p-value", "", "MW p-value", "", "Agree?"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("scale_down", "hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("Mann-Whitney results not yet computed. Run 04_robustness_checks.qmd first.")
}
```


# Demographic Balance Across Periods {#sec-balance}

A key concern with the within-wave quasi-experiment is that the Protest and Post-NSL sub-samples may differ systematically on demographic covariates, which could account for some of the observed attitudinal shifts. @tbl-balance reports means or proportions for age, gender, and education across the two periods, along with $t$-tests (continuous) or $\chi^2$ tests (categorical) for between-period differences.

```{r}
#| label: tbl-balance
#| tbl-cap: "Demographic balance across fieldwork periods (Protest vs. Post-NSL). SMD = standardized mean difference."

hk5_balance <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL"))

balance_rows <- list()

# Age (continuous)
age_protest <- hk5_balance |> filter(period == "Protest") |> pull(age)
age_postnsl <- hk5_balance |> filter(period == "Post-NSL") |> pull(age)
age_t <- t.test(age_protest, age_postnsl)
pooled_sd_age <- sqrt(((length(age_protest) - 1) * sd(age_protest, na.rm = TRUE)^2 +
                        (length(age_postnsl) - 1) * sd(age_postnsl, na.rm = TRUE)^2) /
                       (length(age_protest) + length(age_postnsl) - 2))
smd_age <- (mean(age_protest, na.rm = TRUE) - mean(age_postnsl, na.rm = TRUE)) / pooled_sd_age

balance_rows[[1]] <- tibble(
  Covariate = "Age (mean)",
  Protest = as.character(round(mean(age_protest, na.rm = TRUE), 1)),
  `Post-NSL` = as.character(round(mean(age_postnsl, na.rm = TRUE), 1)),
  SMD = round(smd_age, 3),
  Test = "t-test",
  `p-value` = round(age_t$p.value, 3)
)

# Gender
gender_var <- if ("female" %in% names(hk5_balance)) "female" else
              if ("gender" %in% names(hk5_balance)) "gender" else NULL

if (!is.null(gender_var)) {
  gender_tab <- hk5_balance |>
    filter(!is.na(!!sym(gender_var))) |>
    count(period, !!sym(gender_var)) |>
    group_by(period) |>
    mutate(prop = n / sum(n)) |>
    ungroup()

  gender_chi <- chisq.test(table(hk5_balance[[gender_var]], hk5_balance$period))

  if (gender_var == "female") {
    p1 <- mean(hk5_balance$female[hk5_balance$period == "Protest"], na.rm = TRUE)
    p2 <- mean(hk5_balance$female[hk5_balance$period == "Post-NSL"], na.rm = TRUE)
  } else {
    protest_tab <- gender_tab |> filter(period == "Protest")
    postnsl_tab <- gender_tab |> filter(period == "Post-NSL")
    p1 <- if (nrow(protest_tab) >= 2) protest_tab$prop[2] else NA
    p2 <- if (nrow(postnsl_tab) >= 2) postnsl_tab$prop[2] else NA
  }
  pooled_p <- (p1 + p2) / 2
  smd_gender <- if (!is.na(p1) && !is.na(p2)) (p1 - p2) / sqrt(pooled_p * (1 - pooled_p)) else NA

  balance_rows[[2]] <- tibble(
    Covariate = "Gender (% female)",
    Protest = paste0(round(p1 * 100, 1), "%"),
    `Post-NSL` = paste0(round(p2 * 100, 1), "%"),
    SMD = round(smd_gender, 3),
    Test = "Chi-squared",
    `p-value` = round(gender_chi$p.value, 3)
  )
}

# Education
edu_var <- if ("education_level_clean" %in% names(hk5_balance)) "education_level_clean" else
           if ("education_level" %in% names(hk5_balance)) "education_level" else NULL

if (!is.null(edu_var)) {
  edu_data <- hk5_balance |> mutate(
    edu_clean = case_when(
      !!sym(edu_var) %in% c(-1, 11, 97, 98, 99) ~ NA_real_,
      TRUE ~ as.numeric(!!sym(edu_var))
    )
  )
  edu_protest <- edu_data |> filter(period == "Protest") |> pull(edu_clean)
  edu_postnsl <- edu_data |> filter(period == "Post-NSL") |> pull(edu_clean)
  edu_t <- t.test(edu_protest, edu_postnsl)
  pooled_sd_edu <- sqrt(((sum(!is.na(edu_protest)) - 1) * sd(edu_protest, na.rm = TRUE)^2 +
                          (sum(!is.na(edu_postnsl)) - 1) * sd(edu_postnsl, na.rm = TRUE)^2) /
                         (sum(!is.na(edu_protest)) + sum(!is.na(edu_postnsl)) - 2))
  smd_edu <- (mean(edu_protest, na.rm = TRUE) - mean(edu_postnsl, na.rm = TRUE)) / pooled_sd_edu

  balance_rows[[length(balance_rows) + 1]] <- tibble(
    Covariate = "Education level (SE5, ordinal 1-10)",
    Protest = as.character(round(mean(edu_protest, na.rm = TRUE), 2)),
    `Post-NSL` = as.character(round(mean(edu_postnsl, na.rm = TRUE), 2)),
    SMD = round(smd_edu, 3),
    Test = "t-test",
    `p-value` = round(edu_t$p.value, 3)
  )
}

# Age groups (categorical)
if ("age_group" %in% names(hk5_balance)) {
  age_grp_tab <- hk5_balance |>
    filter(!is.na(age_group)) |>
    count(period, age_group) |>
    group_by(period) |>
    mutate(prop = round(n / sum(n) * 100, 1)) |>
    ungroup()

  age_chi <- chisq.test(table(hk5_balance$age_group, hk5_balance$period))

  for (ag in sort(unique(age_grp_tab$age_group))) {
    p_val <- age_grp_tab |> filter(period == "Protest", age_group == ag) |> pull(prop)
    pn_val <- age_grp_tab |> filter(period == "Post-NSL", age_group == ag) |> pull(prop)
    p_val <- if (length(p_val) == 0) 0 else p_val
    pn_val <- if (length(pn_val) == 0) 0 else pn_val

    balance_rows[[length(balance_rows) + 1]] <- tibble(
      Covariate = paste0("  Age ", ag),
      Protest = paste0(p_val, "%"),
      `Post-NSL` = paste0(pn_val, "%"),
      SMD = NA_real_,
      Test = if (ag == sort(unique(age_grp_tab$age_group))[1]) "Chi-squared" else "",
      `p-value` = if (ag == sort(unique(age_grp_tab$age_group))[1]) round(age_chi$p.value, 3) else NA_real_
    )
  }
}

bind_rows(balance_rows) |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("hold_position")) |>
  footnote(
    general = "SMD computed as (Protest \u2212 Post-NSL) / pooled SD for continuous variables and as the difference in proportions / pooled SD for binary variables. Values |SMD| < 0.1 indicate good balance.",
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```


# Stratified Bootstrap and Manski Bounds {#sec-robustness}

## Stratified Percentile Bootstrap {#sec-bootstrap}

To assess the robustness of the pre/post NSL shifts to sampling variability, I employed a stratified percentile bootstrap with 5,000 iterations, resampling independently within the Protest and Post-NSL periods to preserve the temporal structure of the quasi-experiment. @tbl-bootstrap reports the observed Cohen's $d$ alongside 95% bias-corrected and accelerated (BCa) confidence intervals.

```{r}
#| label: tbl-bootstrap
#| tbl-cap: "Stratified bootstrap results (5,000 iterations): BCa 95% confidence intervals for Cohen's d."

if (exists("boot_results")) {
  boot_results |>
    select(variable, observed_d, boot_d_lo, boot_d_hi, ci_type, sig_boot) |>
    mutate(across(where(is.numeric), ~round(.x, 3))) |>
    kable(
      col.names = c("Variable", "Observed d", "d CI Lo", "d CI Hi",
                     "CI Type", "Sig?"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("scale_down", "hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("Bootstrap results not yet computed. Run 04_robustness_checks.qmd first.")
}
```

```{r}
#| label: fig-bootstrap-forest
#| fig-cap: "Stratified bootstrap 95% BCa confidence intervals for Cohen's d (Post-NSL − Protest). Positive values indicate higher post-NSL responses. Variables where the CI excludes zero are shown in colour."
#| fig-width: 10
#| fig-height: 8

if (exists("boot_results")) {
  boot_results |>
    mutate(
      variable = str_replace_all(variable, "_", " ") |> str_to_title(),
      sig_label = ifelse(sig_boot, "Significant", "Not significant")
    ) |>
    ggplot(aes(x = observed_d, y = reorder(variable, observed_d), color = sig_label)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3) +
    geom_errorbarh(aes(xmin = boot_d_lo, xmax = boot_d_hi), height = 0.25) +
    scale_color_manual(values = c("Significant" = "#E53935", "Not significant" = "#9E9E9E")) +
    labs(
      x = "Cohen's d (Post-NSL \u2212 Protest)",
      y = NULL,
      color = NULL
    ) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "bottom")
}
```


## Informal Bounding Analysis {#sec-manski}

A key concern with the trust increase is that compositional selection --- whereby regime critics disproportionately declined to participate in the post-NSL survey --- could entirely account for the observed shift. To assess the plausibility of this account, I conduct an informal bounding exercise in the spirit of @Manski2003-wo.

The logic is as follows: if we assume that all non-respondents in the post-NSL period would have reported the minimum possible trust score (1 on the 1--4 scale), what proportion of the target population would need to have been "missing critics" for the trust increase to disappear? I present two scenarios: a worst-case bound (imputing the scale minimum) and a more empirically grounded scenario (imputing the 25th percentile of the protest-period distribution, representing respondents drawn from the bottom quartile of critics). @tbl-manski-bounds-police and @tbl-manski-bounds-govt present the results for trust in police and trust in the national government, respectively.

```{r}
#| label: tbl-manski-bounds-police
#| tbl-cap: "Bounding analysis for trust in police: worst-case (impute minimum) and bottom-quartile (impute 25th percentile of protest distribution) scenarios."

hk5_manski <- hk5 |> filter(period %in% c("Protest", "Post-NSL"))

protest_police <- hk5_manski |> filter(period == "Protest", !is.na(trust_police)) |> pull(trust_police)
postnsl_police <- hk5_manski |> filter(period == "Post-NSL", !is.na(trust_police)) |> pull(trust_police)
protest_mean_police <- mean(protest_police, na.rm = TRUE)
q25_police <- quantile(protest_police, 0.25, na.rm = TRUE)
scale_min_police <- min(c(protest_police, postnsl_police), na.rm = TRUE)

manski_police_ext <- map_dfr(seq(0, 40, by = 5), function(pct) {
  n_obs <- length(postnsl_police)
  n_miss <- round(n_obs * pct / (100 - pct))
  total_n <- n_obs + n_miss
  adj_mean_min <- (sum(postnsl_police) + n_miss * scale_min_police) / total_n
  adj_mean_q25 <- (sum(postnsl_police) + n_miss * q25_police) / total_n
  tibble(
    `% Missing` = pct,
    `N Obs` = n_obs,
    `N Missing` = n_miss,
    `Adj Mean (min)` = round(adj_mean_min, 3),
    `Still > Protest? (min)` = adj_mean_min > protest_mean_police,
    `Adj Mean (Q25)` = round(adj_mean_q25, 3),
    `Still > Protest? (Q25)` = adj_mean_q25 > protest_mean_police
  )
})

manski_police_ext |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("scale_down", "hold_position")) |>
  footnote(
    general = paste0(
      "Protest-period mean: ", round(protest_mean_police, 3),
      ". Worst-case imputes scale minimum (", scale_min_police,
      "); Q25 imputes the 25th percentile of the protest-period distribution (",
      q25_police, "), a more empirically grounded assumption."
    ),
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```

```{r}
#| label: tbl-manski-bounds-govt
#| tbl-cap: "Bounding analysis for trust in national government: worst-case (impute minimum) and bottom-quartile (impute 25th percentile of protest distribution) scenarios."

protest_govt <- hk5_manski |> filter(period == "Protest", !is.na(trust_national_government)) |> pull(trust_national_government)
postnsl_govt <- hk5_manski |> filter(period == "Post-NSL", !is.na(trust_national_government)) |> pull(trust_national_government)
protest_mean_govt <- mean(protest_govt, na.rm = TRUE)
q25_govt <- quantile(protest_govt, 0.25, na.rm = TRUE)
scale_min_govt <- min(c(protest_govt, postnsl_govt), na.rm = TRUE)

manski_govt_ext <- map_dfr(seq(0, 40, by = 5), function(pct) {
  n_obs <- length(postnsl_govt)
  n_miss <- round(n_obs * pct / (100 - pct))
  total_n <- n_obs + n_miss
  adj_mean_min <- (sum(postnsl_govt) + n_miss * scale_min_govt) / total_n
  adj_mean_q25 <- (sum(postnsl_govt) + n_miss * q25_govt) / total_n
  tibble(
    `% Missing` = pct,
    `N Obs` = n_obs,
    `N Missing` = n_miss,
    `Adj Mean (min)` = round(adj_mean_min, 3),
    `Still > Protest? (min)` = adj_mean_min > protest_mean_govt,
    `Adj Mean (Q25)` = round(adj_mean_q25, 3),
    `Still > Protest? (Q25)` = adj_mean_q25 > protest_mean_govt
  )
})

manski_govt_ext |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("scale_down", "hold_position")) |>
  footnote(
    general = paste0(
      "Protest-period mean: ", round(protest_mean_govt, 3),
      ". Worst-case imputes scale minimum (", scale_min_govt,
      "); Q25 imputes the 25th percentile of the protest-period distribution (",
      q25_govt, "), a more empirically grounded assumption."
    ),
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```

The results suggest that the observed trust increase is moderately robust to worst-case selection assumptions. Under the more empirically plausible bottom-quartile scenario, the trust increase persists at even higher levels of missing critics. The trust increase disappears only under extreme non-response scenarios with worst-case imputation. This exercise does not resolve the ambiguity between preference falsification and compositional selection --- both remain plausible contributors --- but it provides a concrete benchmark for evaluating the magnitude of selection required to explain the full trust shift.


# Supplementary Figures {#sec-figures}

## Age-Stratified Pre/Post NSL Shifts {#sec-age-stratified}

@fig-age-strat-appendix presents the pre/post NSL mean comparison within each age cohort for trust in police, trust in government, democracy suitability, and freedom to organize. If the trust increase were driven entirely by compositional selection --- specifically, by younger, more critical respondents declining to participate in the post-NSL period --- we would expect the shift to appear primarily or exclusively in younger age groups. The figure shows that while the magnitude of change varies across cohorts, the direction of the trust increase is broadly consistent across age groups, which is more consistent with preference falsification operating across the population than with a purely compositional account.

```{r}
#| label: fig-age-strat-appendix
#| fig-cap: "Pre/Post NSL mean shifts within age cohorts for key indicators. Error bars show 95% confidence intervals."
#| fig-width: 12
#| fig-height: 7

hk5_analysis <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL")) |>
  mutate(period = droplevels(period))

age_strat_vars <- c("trust_police", "democracy_suitability",
                     "trust_national_government", "gov_free_to_organize")

age_strat_labels <- c(
  trust_police = "Trust in Police",
  democracy_suitability = "Democracy Suitability",
  trust_national_government = "Trust in Government",
  gov_free_to_organize = "Freedom to Organize"
)

age_strat_data <- hk5_analysis |>
  filter(!is.na(age_group)) |>
  pivot_longer(cols = all_of(age_strat_vars), names_to = "variable", values_to = "value") |>
  filter(!is.na(value)) |>
  group_by(variable, period, age_group) |>
  summarise(
    mean = mean(value),
    se = sd(value) / sqrt(n()),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(var_label = age_strat_labels[variable])

ggplot(age_strat_data, aes(x = age_group, y = mean, fill = period)) +
  geom_col(position = position_dodge(0.8), width = 0.7, alpha = 0.85) +
  geom_errorbar(aes(ymin = mean - 1.96 * se, ymax = mean + 1.96 * se),
                position = position_dodge(0.8), width = 0.25) +
  facet_wrap(~var_label, scales = "free_y", ncol = 2) +
  scale_fill_manual(values = c("Protest" = "#FF7043", "Post-NSL" = "#5C6BC0")) +
  labs(x = "Age Cohort", y = "Mean", fill = NULL) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
```


# Falsification Tests and Multiple Comparisons {#sec-falsification}

## Placebo-Adjacent Items

To assess whether the observed attitudinal shifts are specifically attributable to politically sensitive mechanisms rather than generalized confounding (e.g., pandemic effects, time trends), I pre-specified a set of items with lower expected political sensitivity. @tbl-falsification reports the pre/post NSL comparison for these placebo-adjacent items.

```{r}
#| label: tbl-falsification
#| tbl-cap: "Placebo-adjacent items: pre-specified items with lower expected political sensitivity. Small or non-significant effects are consistent with NSL-specific mechanisms rather than generalized confounding."

if (exists("falsification_results") && !is.null(falsification_results)) {
  falsification_results |>
    mutate(across(c(protest_mean, postnsl_mean, delta, cohens_d), ~round(.x, 3)),
           p_value = format.pval(p_value, digits = 3)) |>
    kable(
      col.names = c("Variable", "Protest Mean", "Post-NSL Mean",
                     "Delta", "Cohen's d", "p-value", "Sig"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("Falsification results not yet computed. Run 02_main_analysis.qmd first.")
}
```

## Benjamini-Hochberg FDR Correction

The main analysis tests multiple variables simultaneously. To guard against inflated Type I error rates, @tbl-fdr reports raw and Benjamini-Hochberg adjusted $p$-values for all pre/post NSL comparisons. The four pre-specified primary outcomes are flagged.

```{r}
#| label: tbl-fdr
#| tbl-cap: "Raw and FDR-corrected (Benjamini-Hochberg) p-values for all pre/post NSL comparisons."

if (exists("fdr_results")) {
  fdr_results |>
    select(variable, primary, cohens_d, p_value, sig, p_bh, sig_bh) |>
    mutate(across(c(cohens_d), ~round(.x, 3)),
           p_value = format.pval(p_value, digits = 3),
           p_bh = format.pval(p_bh, digits = 3)) |>
    kable(
      col.names = c("Variable", "Primary?", "Cohen's d", "p (raw)", "",
                     "p (BH-adj)", ""),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("scale_down", "hold_position")) |>
    footnote(
      general = "Cohen's d computed as (Post-NSL \u2212 Protest) / pooled SD; positive values indicate higher post-NSL responses.",
      general_title = "Note: ",
      footnote_as_chunk = TRUE
    )
} else {
  cat("FDR results not yet computed. Run 02_main_analysis.qmd first.")
}
```


# Critical Citizens Correlation: CI and Period Decomposition {#sec-critical-citizens}

The main text reports that Hong Kong is the only country among the sixteen ABS Wave 5 polities where the correlation between democratic preference ("democracy is always preferable") and democratic satisfaction is positive. This section reports the precise estimate with 95% confidence intervals and decomposes the correlation by fieldwork period.

```{r}
#| label: tbl-critical-citizens-ci
#| tbl-cap: "Critical citizens correlation (democratic preference × democratic satisfaction): overall and by fieldwork period, with 95% confidence intervals."

if (exists("hk5")) {
  hk5_cc <- hk5 |>
    filter(!is.na(dem_always_preferable), !is.na(democracy_satisfaction))

  # Overall correlation with CI
  overall_test <- cor.test(hk5_cc$dem_always_preferable, hk5_cc$democracy_satisfaction,
                           method = "pearson")

  # By period
  protest_cc <- hk5_cc |> filter(period == "Protest")
  postnsl_cc <- hk5_cc |> filter(period == "Post-NSL")

  protest_test <- if (nrow(protest_cc) > 10) {
    cor.test(protest_cc$dem_always_preferable, protest_cc$democracy_satisfaction)
  } else { NULL }

  postnsl_test <- if (nrow(postnsl_cc) > 10) {
    cor.test(postnsl_cc$dem_always_preferable, postnsl_cc$democracy_satisfaction)
  } else { NULL }

  cc_results <- tibble(
    Period = c("Overall (Wave 5)", "Protest period", "Post-NSL period"),
    N = c(nrow(hk5_cc), nrow(protest_cc), nrow(postnsl_cc)),
    r = c(overall_test$estimate,
          if (!is.null(protest_test)) protest_test$estimate else NA,
          if (!is.null(postnsl_test)) postnsl_test$estimate else NA),
    CI_lo = c(overall_test$conf.int[1],
              if (!is.null(protest_test)) protest_test$conf.int[1] else NA,
              if (!is.null(postnsl_test)) postnsl_test$conf.int[1] else NA),
    CI_hi = c(overall_test$conf.int[2],
              if (!is.null(protest_test)) protest_test$conf.int[2] else NA,
              if (!is.null(postnsl_test)) postnsl_test$conf.int[2] else NA),
    p = c(overall_test$p.value,
          if (!is.null(protest_test)) protest_test$p.value else NA,
          if (!is.null(postnsl_test)) postnsl_test$p.value else NA)
  )

  cc_results |>
    mutate(across(c(r, CI_lo, CI_hi), ~round(.x, 3)),
           p = format.pval(p, digits = 3)) |>
    kable(
      col.names = c("Period", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position"))
} else {
  cat("Data not loaded. Ensure prepared_data.RData is available.")
}
```

The positive correlation is concentrated in the post-NSL period, consistent with a mechanism that operates more strongly under authoritarian constraint.

## Cross-National Comparison {#sec-cc-crossnational}

@tbl-cc-crossnational reports the Pearson correlation between democratic preference ("democracy is always preferable") and democratic satisfaction for all fifteen ABS Wave 5 polities with sufficient data ($N \geq 50$).

```{r}
#| label: tbl-cc-crossnational
#| tbl-cap: "Critical citizens correlation (democratic preference × democratic satisfaction) across ABS Wave 5 polities, sorted by correlation magnitude."

if (file.exists("../analysis/results/cc_country.RData")) {
  load("../analysis/results/cc_country.RData")
  cc_country |>
    mutate(across(c(r, ci_lo, ci_hi), ~round(.x, 3)),
           p = format.pval(p, digits = 3)) |>
    kable(
      col.names = c("Polity", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position")) |>
    row_spec(1, bold = TRUE)
} else {
  cat("cc_country.RData not found.")
}
```

Hong Kong is the only polity with a positive correlation ($r = 0.279$). Singapore shows a near-zero positive value ($r = 0.013$, $p = .70$); all other polities are negative, with established democracies (Australia, Taiwan, Japan, South Korea) showing the strongest negative correlations as expected by the critical citizens framework.

## Decomposition by Democratic Conception Type {#sec-cc-conception}

@tbl-cc-conception decomposes the critical citizens correlation by respondents' conception of democracy, using the ABS forced-choice item asking which element is most essential to democracy. Respondents are classified as holding procedural conceptions (free expression or free elections) or substantive conceptions (basic necessities or clean governance).

```{r}
#| label: tbl-cc-conception
#| tbl-cap: "Critical citizens correlation decomposed by democratic conception type and fieldwork period."

if (exists("hk5")) {
  conception_labels <- c("0" = "Substantive", "1" = "Procedural")
  
  hk5_cc2 <- hk5 |> 
    filter(period %in% c("Protest", "Post-NSL"),
           !is.na(dem_procedural_vs_substantive))
  
  cc_conception <- expand_grid(
    period = c("Protest", "Post-NSL"),
    conception = c(0, 1)
  ) |> pmap_dfr(function(period, conception) {
    sub <- hk5_cc2 |> 
      filter(period == !!period, dem_procedural_vs_substantive == conception,
             !is.na(dem_always_preferable), !is.na(democracy_satisfaction))
    if (nrow(sub) < 20) return(NULL)
    ct <- cor.test(sub$dem_always_preferable, sub$democracy_satisfaction)
    tibble(
      Period = period,
      Conception = conception_labels[as.character(conception)],
      N = nrow(sub),
      r = ct$estimate,
      CI_lo = ct$conf.int[1],
      CI_hi = ct$conf.int[2],
      p = ct$p.value
    )
  })
  
  cc_conception |>
    mutate(across(c(r, CI_lo, CI_hi), ~round(.x, 3)),
           p = format.pval(p, digits = 3)) |>
    kable(
      col.names = c("Period", "Conception", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position"))
} else {
  cat("Data not loaded.")
}
```

The positive correlation is concentrated entirely among respondents with procedural conceptions of democracy ($r = 0.355$, $p < .001$). Respondents with substantive conceptions show near-zero correlations in both periods. A Fisher $z$-test confirms the difference between conception groups in the post-NSL period is statistically significant ($z = 3.55$, $p < .001$). See Sections 4.6 and 5.3 of the main text for discussion.

## Fine-Grained Decomposition by Essential Element {#sec-cc-fine}

@tbl-cc-fine provides a finer decomposition using the four specific response options.

```{r}
#| label: tbl-cc-fine
#| tbl-cap: "Critical citizens correlation by specific essential element of democracy and fieldwork period."

if (exists("hk5")) {
  essential_labels <- c(
    "1" = "Free expression", "2" = "Free elections",
    "4" = "Basic necessities", "5" = "Clean governance"
  )
  
  cc_fine <- expand_grid(
    period = c("Protest", "Post-NSL"),
    essential = c(1, 2, 4, 5)
  ) |> pmap_dfr(function(period, essential) {
    sub <- hk5 |> 
      filter(period == !!period, dem_essential_harmonized == essential,
             !is.na(dem_always_preferable), !is.na(democracy_satisfaction))
    if (nrow(sub) < 20) return(NULL)
    ct <- cor.test(sub$dem_always_preferable, sub$democracy_satisfaction)
    tibble(
      Period = period,
      Element = essential_labels[as.character(essential)],
      N = nrow(sub),
      r = round(ct$estimate, 3),
      CI_lo = round(ct$conf.int[1], 3),
      CI_hi = round(ct$conf.int[2], 3),
      p = format.pval(ct$p.value, digits = 3)
    )
  })
  
  cc_fine |>
    kable(
      col.names = c("Period", "Essential Element", "N", "r", "95% CI Lo", "95% CI Hi", "p-value"),
      booktabs = TRUE
    ) |>
    kable_styling(latex_options = c("hold_position"))
} else {
  cat("Data not loaded.")
}
```

# Differential Item Non-Response {#sec-nonresponse}

@tbl-nonresponse reports item-level response rates for key survey variables, comparing the protest period and post-NSL period sub-samples.

```{r}
#| label: tbl-nonresponse
#| tbl-cap: "Item-level response rates by fieldwork period."

if (exists("hk5")) {
  hk5_nr <- hk5 |> filter(period %in% c("Protest", "Post-NSL"))
  
  key_vars <- c(
    "trust_police", "trust_national_government", "trust_president",
    "trust_parliament", "trust_courts", "trust_civil_service",
    "dem_always_preferable", "democracy_satisfaction",
    "dem_extent_current", "democracy_suitability",
    "system_deserves_support",
    "gov_free_to_organize", "dem_free_speech", "election_free_fair",
    "political_interest", "rich_poor_treated_equally",
    "nat_willing_emigrate"
  )
  
  item_categories <- c(
    trust_police = "Trust", trust_national_government = "Trust",
    trust_president = "Trust", trust_parliament = "Trust",
    trust_courts = "Trust", trust_civil_service = "Trust",
    dem_always_preferable = "Dem. (normative)",
    democracy_satisfaction = "Dem. (normative)",
    dem_extent_current = "Dem. (evaluative)",
    democracy_suitability = "Dem. (evaluative)",
    system_deserves_support = "System support",
    gov_free_to_organize = "Governance", dem_free_speech = "Governance",
    election_free_fair = "Governance",
    political_interest = "Control", rich_poor_treated_equally = "Control",
    nat_willing_emigrate = "Control"
  )
  
  item_labels <- c(
    trust_police = "Trust: police",
    trust_national_government = "Trust: national govt",
    trust_president = "Trust: president/CE",
    trust_parliament = "Trust: parliament",
    trust_courts = "Trust: courts",
    trust_civil_service = "Trust: civil service",
    dem_always_preferable = "Democracy always preferable",
    democracy_satisfaction = "Democratic satisfaction",
    dem_extent_current = "Extent of current democracy",
    democracy_suitability = "Democracy suitability",
    system_deserves_support = "System deserves support",
    gov_free_to_organize = "Freedom to organize",
    dem_free_speech = "Free to speak without fear",
    election_free_fair = "Elections free and fair",
    political_interest = "Political interest",
    rich_poor_treated_equally = "Rich/poor treated equally",
    nat_willing_emigrate = "Willing to emigrate"
  )
  
  nr_table <- map_dfr(key_vars, function(v) {
    if (!v %in% names(hk5_nr)) return(NULL)
    protest <- hk5_nr |> filter(period == "Protest")
    postnsl <- hk5_nr |> filter(period == "Post-NSL")
    tibble(
      Category = item_categories[v],
      Item = item_labels[v],
      `% Protest` = round(100 * sum(!is.na(protest[[v]])) / nrow(protest), 1),
      `% Post-NSL` = round(100 * sum(!is.na(postnsl[[v]])) / nrow(postnsl), 1),
      Delta = round(`% Post-NSL` - `% Protest`, 1)
    )
  })
  
  nr_table |>
    arrange(Category, Delta) |>
    select(-Category) |>
    kable(booktabs = TRUE,
          col.names = c("Item", "Response % (Protest)", "Response % (Post-NSL)", 
                        "\u0394 (pp)")) |>
    kable_styling(latex_options = c("hold_position", "scale_down"))
} else {
  cat("Data not loaded.")
}
```

Trust items maintain near-identical response rates across periods (mean change: $+0.3$ pp). Normative democracy items decline by a mean of $4.6$ pp, governance items by $2.7$ pp, while control items remain stable. This differential pattern is consistent with politically motivated non-response on items requiring explicit democratic commitments.


# Robust Standard Errors {#sec-robust-se}

To address potential heteroskedasticity and assess sensitivity to standard error computation, @tbl-robust-se reports covariate-adjusted estimates with both conventional OLS standard errors and heteroskedasticity-consistent (HC2) robust standard errors for the four primary outcomes.

```{r}
#| label: tbl-robust-se
#| tbl-cap: "Covariate-adjusted Post-NSL effect estimates with conventional and HC2 robust standard errors."

library(sandwich)
library(lmtest)

hk5_robust <- hk5 |>
  filter(period %in% c("Protest", "Post-NSL")) |>
  mutate(post_nsl = as.integer(period == "Post-NSL"))

outcomes <- c("trust_police", "trust_national_government",
              "democracy_suitability", "gov_free_to_organize")
labels_robust <- c("Trust in Police", "Trust in National Government",
            "Democracy Suitability", "Freedom to Organize")

# Build covariate formula dynamically based on available variables
covs <- c()
if ("age" %in% names(hk5_robust) && sum(!is.na(hk5_robust$age)) > 100) covs <- c(covs, "age")
if ("female" %in% names(hk5_robust) && sum(!is.na(hk5_robust$female)) > 100) {
  covs <- c(covs, "female")
} else if ("gender" %in% names(hk5_robust) && sum(!is.na(hk5_robust$gender)) > 100) {
  covs <- c(covs, "gender")
}
if ("education_level_clean" %in% names(hk5_robust) && sum(!is.na(hk5_robust$education_level_clean)) > 100) {
  covs <- c(covs, "education_level_clean")
} else if ("education_level" %in% names(hk5_robust)) {
  hk5_robust <- hk5_robust |> mutate(
    edu_clean = case_when(
      education_level %in% c(-1, 11, 97, 98, 99) ~ NA_real_,
      TRUE ~ as.numeric(education_level)
    )
  )
  if (sum(!is.na(hk5_robust$edu_clean)) > 100) covs <- c(covs, "edu_clean")
}

cov_string <- if (length(covs) > 0) paste("+", paste(covs, collapse = " + ")) else ""

robust_results <- map2_dfr(outcomes, labels_robust, function(outcome, label) {
  f <- as.formula(paste(outcome, "~ post_nsl", cov_string))
  m <- lm(f, data = hk5_robust)
  robust_test <- coeftest(m, vcov = vcovHC(m, type = "HC2"))

  tibble(
    Outcome = label,
    b = round(coef(m)["post_nsl"], 3),
    `SE (conv)` = round(summary(m)$coefficients["post_nsl", "Std. Error"], 3),
    `SE (HC2)` = round(robust_test["post_nsl", "Std. Error"], 3),
    `p (conv)` = format.pval(summary(m)$coefficients["post_nsl", "Pr(>|t|)"], digits = 3),
    `p (HC2)` = format.pval(robust_test["post_nsl", "Pr(>|t|)"], digits = 3)
  )
})

robust_results |>
  kable(booktabs = TRUE) |>
  kable_styling(latex_options = c("hold_position")) |>
  footnote(
    general = paste0("All models control for ", paste(covs, collapse = ", "),
                     ". HC2 standard errors are heteroskedasticity-consistent. ",
                     "Substantive conclusions are unchanged across specifications."),
    general_title = "Note: ",
    footnote_as_chunk = TRUE
  )
```

The HC2 robust standard errors are substantively indistinguishable from conventional OLS standard errors across all four primary outcomes, confirming that heteroskedasticity does not meaningfully affect inference.

# Cross-National Sensitivity Gradient: Turkey and Russia {#sec-wvs-gradient}

The main analysis documents a sensitivity gradient in Hong Kong's ABS Wave 5 data: trust in coercive institutions surged post-NSL ($d \approx +0.49$ for police) while democratic evaluations collapsed ($d = -0.54$), with the magnitude of inflation tracking each institution's coercive role ($r = -0.85$ between pre-specified sensitivity rankings and observed effect sizes). Differential item non-response reinforced this pattern, with normative democracy items showing a mean response rate decline of 4.6 percentage points while trust items remained stable across fieldwork periods. A natural question is whether these patterns are specific to Hong Kong's particular political context and the ABS instrument, or whether the sensitivity gradient captures a more general phenomenon of survey measurement distortion under autocratization.

To assess this, we apply the same logic to two independent cases using an entirely different survey program: the World Values Survey (WVS). Turkey (Wave 6: 2011, Wave 7: 2018) experienced dramatic autocratization following the 2016 failed coup attempt, with mass purges of civil servants, academics, and journalists, declaration of a state of emergency, and constitutional changes consolidating executive power. Russia (Wave 6: 2011, Wave 7: 2017) underwent intensified authoritarian consolidation following the 2014 Crimea annexation, including the suppression of independent media, criminal prosecution of opposition figures, and expansion of surveillance and censorship infrastructure. Both cases are well-documented in the comparative autocratization literature and offer variation in the *type* of repressive mechanism---Turkey's post-coup emergency powers versus Russia's incremental media and civil society squeeze---while sharing the core theoretical condition: a sharp increase in the cost of expressing regime-critical attitudes between survey waves.

We classify WVS items into three categories by expected political sensitivity, paralleling the ABS classification used in the main analysis. *Democracy and political system items* (e.g., importance of democracy, evaluations of democratic governance, attitudes toward strong-leader or military rule) correspond to the low-sensitivity evaluative items in the ABS framework---items where respondents assess the political system in abstract terms. *Institutional confidence items* (e.g., confidence in government, police, courts) correspond to the high-sensitivity trust items in the ABS---items requiring direct evaluation of specific state actors with coercive capacity. *Apolitical control items* (e.g., happiness, life satisfaction, self-reported health) serve as placebo benchmarks with no expected political sensitivity, analogous to the falsification tests reported in Online Appendix E.1.

Note that the WVS and ABS frameworks classify the *direction* of the gradient differently: in the ABS analysis, trust in coercive institutions is coded as high-sensitivity because respondents face direct pressure to express confidence in the police and government, while abstract democratic evaluations are lower-sensitivity. In the WVS analysis, democracy items show *higher non-response* than trust items, which may appear to reverse the ABS pattern. The reconciliation is straightforward: the WVS democracy items ask respondents to evaluate *regime type preferences* (e.g., whether democracy or strong-leader rule is desirable), which in an autocratizing context amounts to a direct assessment of the regime's legitimacy---a high-cost response. The ABS trust items and WVS confidence items both evaluate specific institutions, but the *response distortion* operates in opposite directions: trust items are inflated (respondents overstate confidence) while democracy items are suppressed (respondents either refuse to answer or understate democratic commitment). Both patterns are consistent with preference falsification; they simply manifest differently depending on whether the "safe" response is agreement (trust) or non-response (democratic values).

```{r wvs-gradient-load}
#| include: false

# Load pre-computed results from 08_wvs_sensitivity_gradient.R
if (file.exists("../analysis/results/wvs_sensitivity_gradient.RData")) {
  load("../analysis/results/wvs_sensitivity_gradient.RData")
} else {
  warning("WVS sensitivity gradient results not found. Run analysis/08_wvs_sensitivity_gradient.R first.")
}
```

```{r}
#| label: tbl-wvs-gradient
#| tbl-cap: "Sensitivity gradient in autocratizing contexts: item non-response rates (%) and mean shifts across WVS Waves 6–7 for Turkey and Russia. Items are grouped by expected political sensitivity. Non-response includes don't know, refusal, and item-level missing."

if (exists("wvs_gradient_results")) {
  tbl_data <- wvs_gradient_results |>
    filter(label != "Army rule") |>  # Turkey: not fielded in W7
    mutate(
      sensitivity = factor(sensitivity, levels = c("Democracy", "Trust", "Control")),
      w6_nr = round(w6_nr, 1),
      w7_nr = round(w7_nr, 1),
      nr_change = round(nr_change, 1),
      w6_mean = round(w6_mean, 2),
      w7_mean = round(w7_mean, 2),
      mean_change = round(mean_change, 2)
    ) |>
    select(country, sensitivity, label, w6_nr, w7_nr, nr_change, w6_mean, w7_mean, mean_change) |>
    arrange(country, sensitivity, desc(w7_nr))

  # Build table for each country
  for (cty in c("Turkey", "Russia")) {
    cty_data <- tbl_data |>
      filter(country == cty) |>
      select(-country) |>
      rename(
        Category = sensitivity,
        Item = label,
        `W6 NR%` = w6_nr,
        `W7 NR%` = w7_nr,
        `Δ NR` = nr_change,
        `W6 Mean` = w6_mean,
        `W7 Mean` = w7_mean,
        `Δ Mean` = mean_change
      )

    cat(sprintf("\n**%s (W6: %s, W7: %s)**\n\n",
                cty,
                ifelse(cty == "Turkey", "2011", "2011"),
                ifelse(cty == "Turkey", "2018", "2017")))

    pack_groups <- cty_data |>
      group_by(Category) |>
      summarise(n = n(), .groups = "drop")

    print(
      cty_data |>
        select(-Category) |>
        kable(booktabs = TRUE, align = c("l", rep("r", 6))) |>
        kable_styling(latex_options = c("hold_position", "scale_down")) |>
        pack_rows(index = setNames(pack_groups$n, pack_groups$Category))
    )
    cat("\n")
  }
}
```

@tbl-wvs-gradient presents the results. @tbl-wvs-gradient-tests confirms that the differences are statistically significant.

```{r}
#| label: tbl-wvs-gradient-tests
#| tbl-cap: "Statistical tests for the sensitivity gradient. Panel A reports omnibus chi-squared tests for the association between item category and non-response (pooled across items within each category). Panel B reports pairwise comparisons with odds ratios quantifying the relative odds of non-response for democracy items versus trust items."

if (exists("wvs_gradient_respondent_tests") && exists("wvs_gradient_odds_ratios")) {

  # Panel A: Omnibus tests
  omnibus <- bind_rows(
    wvs_gradient_respondent_tests$Turkey$omnibus,
    wvs_gradient_respondent_tests$Russia$omnibus
  ) |>
    mutate(
      chi_sq = round(chi_sq, 1),
      p_value = format.pval(p_value, digits = 3)
    ) |>
    rename(Country = country, `Chi-sq` = chi_sq, df = df, p = p_value)

  cat("**Panel A: Omnibus chi-squared tests**\n\n")
  print(
    omnibus |>
      kable(booktabs = TRUE) |>
      kable_styling(latex_options = c("hold_position"))
  )

  # Panel B: Odds ratios (Democracy vs Trust NR)
  cat("\n\n**Panel B: Democracy vs Trust non-response odds ratios**\n\n")
  or_tbl <- wvs_gradient_odds_ratios |>
    mutate(across(c(odds_ratio, or_ci_lo, or_ci_hi), ~round(.x, 2))) |>
    rename(Country = country, OR = odds_ratio, `95% CI lo` = or_ci_lo, `95% CI hi` = or_ci_hi)

  print(
    or_tbl |>
      kable(booktabs = TRUE) |>
      kable_styling(latex_options = c("hold_position"))
  )
}
```

In both Turkey and Russia, democracy-related items exhibit substantially higher non-response rates in Wave 7 than institutional confidence items, which in turn exceed apolitical control items. In Turkey, the average Wave 7 non-response rate for democracy items (excluding the army rule item, which was not fielded) is approximately 7% compared to 1.4% for trust items and 0.4% for control items. In Russia, the corresponding figures are approximately 12%, 4.5%, and 1.3%. This monotonic gradient---democracy items > trust items > control items---parallels the pattern documented in Hong Kong using the ABS.

Notably, Turkey's "army rule" item (Q237) was entirely absent from the Wave 7 questionnaire, constituting the most extreme form of item suppression: the survey organization itself removed the most politically sensitive question in the post-coup context. This absence underscores the practical relevance of the sensitivity gradient for survey practitioners working in autocratizing environments.

The trust paradox is also visible in both cases. Despite substantial democratic backsliding as measured by V-Dem indices, confidence in state institutions either increased or remained stable between waves. In Turkey, confidence in government, police, courts, and armed forces all shifted toward greater trust. In Russia, the same pattern holds, with confidence in parliament showing the largest increase. Meanwhile, Russia's democracy evaluation score *increased* by over one full point on the 1--10 scale---a pattern more consistent with rally-around-the-flag effects and preference falsification than with genuine improvements in democratic governance.

```{r}
#| label: fig-wvs-gradient
#| fig-cap: "Sensitivity gradient in autocratizing contexts: Wave 7 item non-response rates by category. Democracy-related items exhibit systematically higher non-response than institutional trust items, which exceed apolitical controls. Turkey: 2011→2018 (post-2016 coup attempt); Russia: 2011→2017 (post-Crimea annexation). 'Army rule' item excluded (not fielded in Turkey Wave 7)."
#| fig-width: 10
#| fig-height: 7

if (exists("wvs_gradient_results")) {
  fig_data <- wvs_gradient_results |>
    filter(label != "Army rule") |>
    mutate(
      sensitivity = factor(sensitivity, levels = c("Democracy", "Trust", "Control")),
      label = fct_reorder2(label, as.numeric(sensitivity), -w7_nr,
                           .fun = function(s, nr) s[1] * 1000 - mean(nr))
    )

  ggplot(fig_data, aes(x = w7_nr, y = label, fill = sensitivity)) +
    geom_col(width = 0.7, alpha = 0.85) +
    facet_wrap(~country, scales = "free_x") +
    scale_fill_manual(
      values = c(Democracy = "#CC3311", Trust = "#0077BB", Control = "#009988"),
      name = "Item category"
    ) +
    labs(
      x = "Item non-response rate (%, Wave 7)",
      y = NULL
    ) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "top",
      strip.text = element_text(face = "bold", size = 13),
      panel.grid.major.y = element_line(color = "grey92"),
      panel.grid.minor = element_blank()
    )
}
```

These findings demonstrate that the sensitivity gradient is not idiosyncratic to the Hong Kong context or the Asian Barometer Survey instrument. The same pattern---systematic differential non-response scaling with item political sensitivity---emerges independently in WVS data from two additional autocratizing contexts with distinct political histories, repressive mechanisms, and survey implementations. This cross-national replication strengthens the case that the sensitivity gradient constitutes a generalizable diagnostic for survey validity under political pressure.
